### 2017-07-28 JStatSoft_forestinventory_master.tex
first commit for initializing

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Andreas Hill \\ETH Z\"urich \And 
        Alexander Massey}
\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable and evidence based information to assess the state and development of forests over time. The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods. This sample is then used to provide estimates of a target variable for a given spatial unit. Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs. With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient. The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs. Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata. This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}

\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Andreas Hill\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{andreas.hill@usys.ethz.ch}\\
  Telephone: +41/44/632 32 36\\
  URL: \url{http://www.lue.ethz.ch/people/hilla}

  Alexander Massey\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{afmass@gmail.com}\\}
}

%% for those who use Sweave please include the following line (with % symbols):

%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Tex Settings ------------------------------------------ %

% used packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathptmx} 
\usepackage{latexsym}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{geometry}
 \geometry{
 a4paper,
 left=28mm,
 bottom=28mm
 }
\usepackage[labelfont=bf]{caption}
\usepackage[font=footnotesize]{caption}
\usepackage[font=footnotesize]{subcaption}

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\var}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{RE}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}

%\setlength{\parindent}{1em}

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Main Document------------------------------------------ %

\begin{document}

%------------------------------------------------------------------------------------------------%
% -------------------------------------- R Settings -------------------------------------------- %

% set global options for R-output:

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Introduction ---------------------------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Introdution}
\label{sec:intro}

In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods. They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level. Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions. For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e. information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}. This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}. The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size. However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited. Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par

Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}. A method which has become particularly attractive is so called \textit{multiphase sampling}. The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size. This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered. These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area. Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}. Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par

% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software. One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach. However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing. Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total. The estimators can be used for stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e. rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).

% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
%              available to a large user community, to facilitate their application in science as well as 
%              operational forest management (why is this important? --> because the methods have demonstrate
%              the potential of considerably increasing the estimation precision)
% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
%                and their implementation in our package
%             b) illustrate the application of the various estimators by the respsective functions in our package 
%                to real-world inventory scenarios
%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
%                deal with such situations (error-checking functions and data-adjustments)

% NOTES:
% --> take parts of the old introduction version of the diploma-thesis!

% NOTES:
% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
%                model-dependent vs. design-based
% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
%                 and seperate them from our package (focus, applied areas, approach)
%              + important reasoning for our package: we are the first that implemented and provide 
%                a large range of design-based estimators
% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
%              available to a large user community, to facilitate their application in science as well as 
%              operational forest management (why is this important? --> because the methods have demonstrate
%              the potential of considerably increasing the estimation precision)
% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
%                and their implementation in our package
%             b) illustrate the application of the various estimators by the respsective functions in our package 
%                to real-world inventory scenarios
%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
%                deal with such situations (error-checking functions and data-adjustments)

% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
%
% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)

% Explain the problem that leads to sae estimations (in generell and in forestry)
%
%

% NOTES:
% --> take parts of the old introduction version of the diploma-thesis!

% NOTES:
% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
%                model-dependent vs. design-based
% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
%                 and seperate them from our package (focus, applied areas, approach)
%              + important reasoning for our package: we are the first that implemented and provide 
%                a large range of design-based estimators
% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
%              available to a large user community, to facilitate their application in science as well as 
%              operational forest management (why is this important? --> because the methods have demonstrate
%              the potential of considerably increasing the estimation precision)
% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
%                and their implementation in our package
%             b) illustrate the application of the various estimators by the respsective functions in our package 
%                to real-world inventory scenarios
%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
%                deal with such situations (error-checking functions and data-adjustments)

% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
%
% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)

%------------------------------------------------------------------------------------------------%
% --------------------------- Structure of the Package ----------------------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

%\section{Structure of the Package}
\section{Methods and Structure of the Package}
\label{sec:str_and_mod}

%Here comes the general structure of the package, or at least of the two and three-phase functions

\subsection{Two-Phase Sampling}

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ such as the timber volume density per hectare. The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model. In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function. Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.

\subsection{Three-Phase Sampling}

The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}). The basic assumption is that auxiliary information is available in two different sample sizes.

A first phase of auxiliary information (e.g. taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares. A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions. A further subsample produces a third final phase based on terrestrial observations (i.e. the local densities of the ground truth) and is used to correct for bias in the design-based sense. The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e. a census). Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.

Explain three phase sampling, giving refrences to existing literature and studies

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth} 
		\centering
    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth} 
		\centering
   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{\textbf{(a)} Concept of multiphase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate \textit{available} information. \textbf{(b)} Illustration of the Small Area Estimation problem}
\label{fig:concmphase_and_sae}
\end{figure}

% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case. The auxiliary information can be any information that exhibits a correlation to the local density, i.e. target variable, $Y(x)$.
% 
% e.g. past inventory data \citep{massey2014a}

\subsection{Small Area Estimation}

Explain:
\begin{itemize}
  \item What is the difference between \textit{global} and \textit{small area} estimation?
  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
\end{itemize}

\begin{figure}[htb]
\centering
\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
\label{fig:struct_package}
\end{figure}

% NOTES:
% - Pyramide / decision tree Graphik
% - Confidence Interval calculation
% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
%
% --> explain each point briefly without to much (mathematical) detail

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Estimators and their Application -------------------------- %

%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section[Global Estimators and their Application in R]{Global Estimators and their Application}
\label{sec:globest_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Double Sampling (Two-Phase) Estimators}

% Note: As the mathematical form of estimators is still quite simple and short, we can here
%       focus on presenting and explaining important 'components' of the estimators, such as Z, C, ...)
% 
% \subsection{Mathematical Background}
% 
% Here, we give the \textbf{mathematical background} of:
% \begin{itemize}
%   \item the classical two-phase estimator, including:
%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
%   \item ...
% \end{itemize}
% 
% 
% 
% \subsubsection{Application}
% 
% Here, we give the \textbf{application} examples:
% \begin{itemize}
%   \item example for non-exhaustive case: use grison dataset
%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
%   \item example for exhaustive case: use cluster set
% \end{itemize}

\subsubsection{Mathematical Background}

Mention:
\begin{itemize}
  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
\end{itemize}

The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as

\begin{equation}\label{normequ_simple}
  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
\end{equation}

The design-based variance-covariance matrix of the regression coefficients is then calculated as

\begin{equation}\label{eq:estvarmatrix}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
\end{equation}

with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
\begin{equation}\label{eq:resids}
  \hat{R}(x)=Y(x)-\hat{Y}(x)
\end{equation}

The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}. Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models. Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to

\begin{equation}\label{eq:pointest_simple}
\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
\end{equation}

The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}. Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}. The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

\begin{equation}\label{eq:gw_var_simple}
\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
\end{equation}

\begin{equation}\label{eq:varexternalsimple}
\hat{\var}(\hat{Y}_{reg})=
\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
\end{equation}

\subsubsection{Application}

\subsection{Triple Sampling (Three-Phase) Estimators}

\subsubsection{Mathematical Background}

\subsubsection{Application}

% The psynth estimator:
% 
% \begin{equation}\label{pseudosynth1}
% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
% \end{equation}
% 
% \begin{equation}\label{estvarpseudosynth1}
% \hat{\var}(\hat{Y}_{G,psynth}): =
% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
% \end{equation}

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
\label{sec:saeest_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Double Sampling (Two-Phase) Estimators}

% - Merge Mathematical Background and Application
% - Quite Likely, we cannot go through all estimators ...
%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
%
% - 
%
% - We could give the function call for all cases, but we should only show the output
%   for demonstrating particular things

% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
% 1) nex, simple sampling, pseudo extended sae estimator
% 2) nex, simple sampling, pseudo sae estimator
% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
% 
% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
%    (mention that in cluster case, the cluster means shall be used)
% always use bw-adjustment (?)
%

\subsection{Triple Sampling (Three-Phase) Estimators}

% The psynth estimator:
% 
% \begin{equation}\label{pseudosynth1}
% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
% \end{equation}
% 
% \begin{equation}\label{estvarpseudosynth1}
% \hat{\var}(\hat{Y}_{G,psynth}): =
% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
% \end{equation}

% NOTES:
% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
%

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Special Cases and Scenarios ------------------------------- %

% NOTES:
% - Error handling and adjustments: Sampling designs not nested
% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
%                   or s2 or/and s1 sample in 3-phase case
% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Calculation of Confidence Intervals ----------------------- %

% NOTES:
% - demonstrate calculation of confidence intervals (give the formulas)

%------------------------------------------------------------------------------------------------%
% ----------------------------- Visualization and Analysis of Estimation Results --------------- %

% NOTES:
% - demonstrate visualization of errors and point-ests + CIs
% - demonstrate a) releff-calculation and gain-analysis

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Literature ------------------------------------------ %

\bibliography{bib/literature}\newpage\cleardoublepage

% -------------------------------------------------------------------------- %
\end{document}

Hallo. Here comes a first try:

\begin{Schunk}
\begin{Sinput}
R> library(forestinventory)
R> op <- onephase(formula = tvol~1 ,data = grisons,
 +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
R> summary(op)
\end{Sinput}
\begin{Soutput}
One-phase estimation

Call: 
onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
    terrgrid.id = 2))

Method used:
One-phase estimator

Estimation results:
 estimate variance n2
 399.4321 567.2001 67
\end{Soutput}
\end{Schunk}

This was the result

### 2017-07-28 JStatSoft_forestinventory_master.rnw
upload all files

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
 \Abstract{
-Forest inventories provide reliable and evidence based information to assess the state and development of forests over time.
+Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.
 The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
@@
 % set global options for R-output:
+<<echo=FALSE>>=
+options(prompt = "R> ", continue = " + ")
+@
 
@@
 
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section{Introdution}
-\label{sec:intro}
-
-In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods.
-They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
-information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
-However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par
-
-Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multiphase sampling}.
-The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
-This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
-Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
-
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
-The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
-The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
-
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
-
-% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-% Explain the problem that leads to sae estimations (in generell and in forestry)
-%
-%
+\SweaveInput{rnw/Introduction.Rnw}
 
@@
 
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-%\section{Structure of the Package}
-\section{Methods and Structure of the Package}
-\label{sec:str_and_mod}
-
-%Here comes the general structure of the package, or at least of the two and three-phase functions
-
-\subsection{Two-Phase Sampling}
-
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
-These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ such as the timber volume density per hectare.
-The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function.
-Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
-
-\subsection{Three-Phase Sampling}
-
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
-The basic assumption is that auxiliary information is available in two different sample sizes.
-
-A first phase of auxiliary information (e.g.
-taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares.
-A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions.
-A further subsample produces a third final phase based on terrestrial observations (i.e.
-the local densities of the ground truth) and is used to correct for bias in the design-based sense.
-The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e.
-a census).
-Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.
-
-Explain three phase sampling, giving refrences to existing literature and studies
-
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_b}
-	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
-The square represents the forest area for which an inventory is being conducted.
-The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
-\label{fig:concmphase_and_sae}
-\end{figure}
-
-% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
-The auxiliary information can be any information that exhibits a correlation to the local density, i.e.
-target variable, $Y(x)$.
-% 
-% e.g. past inventory data \citep{massey2014a}
-
-\subsection{Small Area Estimation}
-
-Explain:
-\begin{itemize}
-  \item What is the difference between \textit{global} and \textit{small area} estimation?
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-\begin{figure}[htb]
-\centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
-\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
-\label{fig:struct_package}
-\end{figure}
+\SweaveInput{rnw/Main structure and modules of the Package.Rnw}
 
@@
 
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
+\SweaveInput{rnw/Global Estimators and their Application in R (2).Rnw}
 
-\section[Global Estimators and their Application in R]{Global Estimators and their Application}
-\label{sec:globest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% Note: As the mathematical form of estimators is still quite simple and short, we can here
-%       focus on presenting and explaining important 'components' of the estimators, such as Z, C, ...)
-% 
-% \subsection{Mathematical Background}
-% 
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-% 
-% 
-% \subsubsection{Application}
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
-
-\subsubsection{Mathematical Background}
-
-Mention:
-\begin{itemize}
-  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
-
-\begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
-\end{equation}
-
-The design-based variance-covariance matrix of the regression coefficients is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
-\end{equation}
-
-with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
-\begin{equation}\label{eq:resids}
-  \hat{R}(x)=Y(x)-\hat{Y}(x)
-\end{equation}
-
-The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
-Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to
-
-\begin{equation}\label{eq:pointest_simple}
-\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
-\end{equation}
-
-The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
-The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{equation}\label{eq:gw_var_simple}
-\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-\end{equation}
-
-\begin{equation}\label{eq:varexternalsimple}
-\hat{\var}(\hat{Y}_{reg})=
-\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
-\end{equation}
-
-\subsubsection{Application}
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-\subsubsection{Mathematical Background}
-
-\subsubsection{Application}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
-\label{sec:saeest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% - 
-%
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
-
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
-%
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
+\SweaveInput{rnw/SAE Estimators and their Application in R.Rnw}
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
+<<>>=
+library(forestinventory)
+op <- onephase(formula = tvol~1 ,data = grisons,
+              phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
+summary(op)
+@
 

### 2017-07-28 rnw/Global Estimators and their Application in R (2).Rnw
upload all files

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
-\documentclass[article]{jss}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% almost as usual
-\author{Andreas Hill \\ETH Z\"urich \And 
-        Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{
-Forest inventories provide reliable and evidence based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient.
-The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}
-
-\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-  Andreas Hill\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{andreas.hill@usys.ethz.ch}\\
-  Telephone: +41/44/632 32 36\\
-  URL: \url{http://www.lue.ethz.ch/people/hilla}
-
-  Alexander Massey\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{afmass@gmail.com}\\}
-}
-
-%% for those who use Sweave please include the following line (with % symbols):
-
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Tex Settings ------------------------------------------ %
-
-% used packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathptmx} 
-\usepackage{latexsym}
-\usepackage{a4}
-\usepackage{graphicx}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{flafter}
-\usepackage{bm}
-\usepackage{setspace}
-\usepackage{lineno}
-\usepackage{natbib}
-\usepackage{geometry}
- \geometry{
- a4paper,
- left=28mm,
- bottom=28mm
- }
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage[font=footnotesize]{subcaption}
-
-% new commands
-\newcommand{\bwi}{BWI3}
-\newcommand{\LF}{\ensuremath{\lambda(F)}}
-\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
-\newcommand{\EX}{\mathbb{E}}
-\newcommand{\PR}{\mathbb{P}}
-\newcommand{\var}{\mathbb{V}}
-\newcommand{\COV}{\mathbb{COV}}
-\newcommand{\MAV}{\mathbb{MAV}}
-\newcommand{\MRAV}{\mathbb{MRAV}}
-\newcommand{\POP}{\mathcal{P}}
-\newcommand{\SAMP}{\mathcal{S}}
-\newcommand{\RE}{\mathbb{RE}}
-\newcommand{\PLAN}{\Re^2}
-\newcommand{\SUR}{\mathbb{S}}
-\newcommand{\ING}{\mathbb{I}}
-\newcommand{\DEP}{\mathbb{D}}
-
-%\setlength{\parindent}{1em}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Main Document------------------------------------------ %
-
-\begin{document}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- R Settings -------------------------------------------- %
-
-% set global options for R-output:
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Introduction ---------------------------------------------- %
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section{Introdution}
-\label{sec:intro}
-
-In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods.
-They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
-information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
-However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par
-
-Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multiphase sampling}.
-The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
-This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
-Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
-
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
-The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
-The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
-
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
-
-% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-% Explain the problem that leads to sae estimations (in generell and in forestry)
-%
-%
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-%------------------------------------------------------------------------------------------------%
-% --------------------------- Structure of the Package ----------------------------------------- %
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-%\section{Structure of the Package}
-\section{Methods and Structure of the Package}
-\label{sec:str_and_mod}
-
-%Here comes the general structure of the package, or at least of the two and three-phase functions
-
-\subsection{Two-Phase Sampling}
-
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
-These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ such as the timber volume density per hectare.
-The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function.
-Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
-
-\subsection{Three-Phase Sampling}
-
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
-The basic assumption is that auxiliary information is available in two different sample sizes.
-
-A first phase of auxiliary information (e.g.
-taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares.
-A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions.
-A further subsample produces a third final phase based on terrestrial observations (i.e.
-the local densities of the ground truth) and is used to correct for bias in the design-based sense.
-The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e.
-a census).
-Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.
-
-Explain three phase sampling, giving refrences to existing literature and studies
-
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_b}
-	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
-The square represents the forest area for which an inventory is being conducted.
-The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
-\label{fig:concmphase_and_sae}
-\end{figure}
-
-% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
-The auxiliary information can be any information that exhibits a correlation to the local density, i.e.
-target variable, $Y(x)$.
-% 
-% e.g. past inventory data \citep{massey2014a}
-
-\subsection{Small Area Estimation}
-
-Explain:
-\begin{itemize}
-  \item What is the difference between \textit{global} and \textit{small area} estimation?
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-\begin{figure}[htb]
-\centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
-\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
-\label{fig:struct_package}
-\end{figure}
-
-% NOTES:
-% - Pyramide / decision tree Graphik
-% - Confidence Interval calculation
-% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
-%
-% --> explain each point briefly without to much (mathematical) detail
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Estimators and their Application -------------------------- %
-
-%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
-
 % !Rnw root = JStatSoft_forestinventory_master.Rnw
@@
 % \end{equation}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
-\label{sec:saeest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% - 
-%
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
-
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
-%
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% NOTES:
-% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
-% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
-%
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Special Cases and Scenarios ------------------------------- %
-
-% NOTES:
-% - Error handling and adjustments: Sampling designs not nested
-% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
-%                   or s2 or/and s1 sample in 3-phase case
-% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
-
-% NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
-
-%------------------------------------------------------------------------------------------------%
-% ----------------------------- Visualization and Analysis of Estimation Results --------------- %
-
-% NOTES:
-% - demonstrate visualization of errors and point-ests + CIs
-% - demonstrate a) releff-calculation and gain-analysis
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------------- Literature ------------------------------------------ %
-
-\bibliography{bib/literature}\newpage\cleardoublepage
-
-% -------------------------------------------------------------------------- %
-\end{document}
-
-Hallo. Here comes a first try:
-
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
-
-This was the result

### 2017-07-28 rnw/Global Estimators and their Application in R.Rnw
upload all files

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
 
-Mention:
-\begin{itemize}
-  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
+The double sampling estimators use inventory information from \textbf{two} nested samples (\textit{phases}).
+The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
+These explanatory variables have been derived from auxiliary information, often remote sensing data, that is available in high quantity within the forest area $F$.
+The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
+the local density $Y(x)$ such as the timber volume density per hectare.
+The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.The regression coefficients of the OLS regression model are found by solving the sample-based normal equation.
+In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
 
@@
 
-The design-based variance-covariance matrix of the regression coefficients is then calculated as
+In the special case of \textbf{cluster sampling}, we have
 
-\begin{equation}\label{eq:estvarmatrix}
+\begin{equation}\label{normequ_cluster}
+  \hat{\pmb{\beta}}_{c,s_2}&=&
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}_c(x)\pmb{Z}_c^t(x)\Big)^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
+\end{equation}
+
+with $M(x)$ being the number of samples per cluster that can vary due to forest/non-forest decision on the plot level, and $\pmb{Z}_c(x)$ containing the weighted mean of each explanatory variable on the cluster level.
+Note that in the case of cluster sampling, $n_2$ denotes the number of clusters.
+
+In order to calculate the point estimate and its estimated variance, we further have to calulate the design-based variance-covariance matrix of the regression coefficients, defined as
+
+\begin{eqnarray}\label{eq:estvarmatrix}
   \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
-\end{equation}
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}_c(x)\pmb{Z}_c^t(x) \Big)^{-1}
+  \Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}^2_c(x)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}_c(x)\pmb{Z}_c^t(x) \Big)^{-1}
+\end{eqnarray}
 
-with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
-\begin{equation}\label{eq:resids}
-  \hat{R}(x)=Y(x)-\hat{Y}(x)
-\end{equation}
+as well as the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ as
+\begin{eqnarray}\label{eq:resids}
+  \hat{R}(x)=Y(x)-\hat{Y}(x) \\
+  \hat{R}_c(x)=Y_c(x)-\hat{Y}_c(x)
+\end{eqnarray}
 
-The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
-Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to
+We can subsequently calculate the \textbf{point estimate} for simpe sampling according to equation \ref{eq:pointest_simple}.
+Note that this form result under the prerequisite that the regression coefficients are derived using the data from the current inventory, i.e.
+the regression model is \textbf{internal}.
 
@@
 
-The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
+We can specify the estimation precision of the point estimate by calcualting the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
 Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
-The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
+The package \pkg{forestinventory} additonally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
 
 \begin{equation}\label{eq:gw_var_simple}
-\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
+\hat{\var}(\hat{Y}_{G,synth})=\bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
 \end{equation}

### 2017-07-28 rnw/Introduction.Rnw
upload all files

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
-\documentclass[article]{jss}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% almost as usual
-\author{Andreas Hill \\ETH Z\"urich \And 
-        Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{
-Forest inventories provide reliable and evidence based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient.
-The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}
-
-\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-  Andreas Hill\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{andreas.hill@usys.ethz.ch}\\
-  Telephone: +41/44/632 32 36\\
-  URL: \url{http://www.lue.ethz.ch/people/hilla}
-
-  Alexander Massey\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{afmass@gmail.com}\\}
-}
-
-%% for those who use Sweave please include the following line (with % symbols):
-
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Tex Settings ------------------------------------------ %
-
-% used packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathptmx} 
-\usepackage{latexsym}
-\usepackage{a4}
-\usepackage{graphicx}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{flafter}
-\usepackage{bm}
-\usepackage{setspace}
-\usepackage{lineno}
-\usepackage{natbib}
-\usepackage{geometry}
- \geometry{
- a4paper,
- left=28mm,
- bottom=28mm
- }
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage[font=footnotesize]{subcaption}
-
-% new commands
-\newcommand{\bwi}{BWI3}
-\newcommand{\LF}{\ensuremath{\lambda(F)}}
-\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
-\newcommand{\EX}{\mathbb{E}}
-\newcommand{\PR}{\mathbb{P}}
-\newcommand{\var}{\mathbb{V}}
-\newcommand{\COV}{\mathbb{COV}}
-\newcommand{\MAV}{\mathbb{MAV}}
-\newcommand{\MRAV}{\mathbb{MRAV}}
-\newcommand{\POP}{\mathcal{P}}
-\newcommand{\SAMP}{\mathcal{S}}
-\newcommand{\RE}{\mathbb{RE}}
-\newcommand{\PLAN}{\Re^2}
-\newcommand{\SUR}{\mathbb{S}}
-\newcommand{\ING}{\mathbb{I}}
-\newcommand{\DEP}{\mathbb{D}}
-
-%\setlength{\parindent}{1em}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Main Document------------------------------------------ %
-
-\begin{document}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- R Settings -------------------------------------------- %
-
-% set global options for R-output:
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Introduction ---------------------------------------------- %
-
 % !Rnw root = JStatSoft_forestinventory_master.Rnw
@@
 %
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-%------------------------------------------------------------------------------------------------%
-% --------------------------- Structure of the Package ----------------------------------------- %
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-%\section{Structure of the Package}
-\section{Methods and Structure of the Package}
-\label{sec:str_and_mod}
-
-%Here comes the general structure of the package, or at least of the two and three-phase functions
-
-\subsection{Two-Phase Sampling}
-
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
-These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ such as the timber volume density per hectare.
-The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function.
-Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
-
-\subsection{Three-Phase Sampling}
-
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
-The basic assumption is that auxiliary information is available in two different sample sizes.
-
-A first phase of auxiliary information (e.g.
-taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares.
-A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions.
-A further subsample produces a third final phase based on terrestrial observations (i.e.
-the local densities of the ground truth) and is used to correct for bias in the design-based sense.
-The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e.
-a census).
-Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.
-
-Explain three phase sampling, giving refrences to existing literature and studies
-
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_b}
-	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
-The square represents the forest area for which an inventory is being conducted.
-The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
-\label{fig:concmphase_and_sae}
-\end{figure}
-
-% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
-The auxiliary information can be any information that exhibits a correlation to the local density, i.e.
-target variable, $Y(x)$.
-% 
-% e.g. past inventory data \citep{massey2014a}
-
-\subsection{Small Area Estimation}
-
-Explain:
-\begin{itemize}
-  \item What is the difference between \textit{global} and \textit{small area} estimation?
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-\begin{figure}[htb]
-\centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
-\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
-\label{fig:struct_package}
-\end{figure}
-
-% NOTES:
-% - Pyramide / decision tree Graphik
-% - Confidence Interval calculation
-% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
-%
-% --> explain each point briefly without to much (mathematical) detail
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Estimators and their Application -------------------------- %
-
-%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Global Estimators and their Application in R]{Global Estimators and their Application}
-\label{sec:globest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% Note: As the mathematical form of estimators is still quite simple and short, we can here
-%       focus on presenting and explaining important 'components' of the estimators, such as Z, C, ...)
-% 
-% \subsection{Mathematical Background}
-% 
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-% 
-% 
-% \subsubsection{Application}
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
-
-\subsubsection{Mathematical Background}
-
-Mention:
-\begin{itemize}
-  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
-
-\begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
-\end{equation}
-
-The design-based variance-covariance matrix of the regression coefficients is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
-\end{equation}
-
-with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
-\begin{equation}\label{eq:resids}
-  \hat{R}(x)=Y(x)-\hat{Y}(x)
-\end{equation}
-
-The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
-Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to
-
-\begin{equation}\label{eq:pointest_simple}
-\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
-\end{equation}
-
-The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
-The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{equation}\label{eq:gw_var_simple}
-\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-\end{equation}
-
-\begin{equation}\label{eq:varexternalsimple}
-\hat{\var}(\hat{Y}_{reg})=
-\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
-\end{equation}
-
-\subsubsection{Application}
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-\subsubsection{Mathematical Background}
-
-\subsubsection{Application}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
-\label{sec:saeest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% - 
-%
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
-
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
-%
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% NOTES:
-% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
-% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
-%
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Special Cases and Scenarios ------------------------------- %
-
-% NOTES:
-% - Error handling and adjustments: Sampling designs not nested
-% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
-%                   or s2 or/and s1 sample in 3-phase case
-% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
-
-% NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
-
-%------------------------------------------------------------------------------------------------%
-% ----------------------------- Visualization and Analysis of Estimation Results --------------- %
-
-% NOTES:
-% - demonstrate visualization of errors and point-ests + CIs
-% - demonstrate a) releff-calculation and gain-analysis
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------------- Literature ------------------------------------------ %
-
-\bibliography{bib/literature}\newpage\cleardoublepage
-
-% -------------------------------------------------------------------------- %
-\end{document}
-
-Hallo. Here comes a first try:
-
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
-
-This was the result

### 2017-07-28 rnw/Main structure and modules of the Package.Rnw
upload all files

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
-\documentclass[article]{jss}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% almost as usual
-\author{Andreas Hill \\ETH Z\"urich \And 
-        Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{
-Forest inventories provide reliable and evidence based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient.
-The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}
-
-\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-  Andreas Hill\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{andreas.hill@usys.ethz.ch}\\
-  Telephone: +41/44/632 32 36\\
-  URL: \url{http://www.lue.ethz.ch/people/hilla}
-
-  Alexander Massey\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{afmass@gmail.com}\\}
-}
-
-%% for those who use Sweave please include the following line (with % symbols):
-
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Tex Settings ------------------------------------------ %
-
-% used packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathptmx} 
-\usepackage{latexsym}
-\usepackage{a4}
-\usepackage{graphicx}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{flafter}
-\usepackage{bm}
-\usepackage{setspace}
-\usepackage{lineno}
-\usepackage{natbib}
-\usepackage{geometry}
- \geometry{
- a4paper,
- left=28mm,
- bottom=28mm
- }
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage[font=footnotesize]{subcaption}
-
-% new commands
-\newcommand{\bwi}{BWI3}
-\newcommand{\LF}{\ensuremath{\lambda(F)}}
-\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
-\newcommand{\EX}{\mathbb{E}}
-\newcommand{\PR}{\mathbb{P}}
-\newcommand{\var}{\mathbb{V}}
-\newcommand{\COV}{\mathbb{COV}}
-\newcommand{\MAV}{\mathbb{MAV}}
-\newcommand{\MRAV}{\mathbb{MRAV}}
-\newcommand{\POP}{\mathcal{P}}
-\newcommand{\SAMP}{\mathcal{S}}
-\newcommand{\RE}{\mathbb{RE}}
-\newcommand{\PLAN}{\Re^2}
-\newcommand{\SUR}{\mathbb{S}}
-\newcommand{\ING}{\mathbb{I}}
-\newcommand{\DEP}{\mathbb{D}}
-
-%\setlength{\parindent}{1em}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Main Document------------------------------------------ %
-
-\begin{document}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- R Settings -------------------------------------------- %
-
-% set global options for R-output:
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Introduction ---------------------------------------------- %
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section{Introdution}
-\label{sec:intro}
-
-In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods.
-They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
-information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
-However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par
-
-Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multiphase sampling}.
-The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
-This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
-Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
-
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
-The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
-The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
-
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
-
-% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-% Explain the problem that leads to sae estimations (in generell and in forestry)
-%
-%
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-%------------------------------------------------------------------------------------------------%
-% --------------------------- Structure of the Package ----------------------------------------- %
-
 % !Rnw root = JStatSoft_forestinventory_master.Rnw
@@
 The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
-The basic assumption is that auxiliary information is available in two different sample sizes.
+The basic assumption is that auxiliary information is available in two different frequencies, i.e.
+sample sizes.
+The phase $s_0$ provides
 
@@
 \end{figure}
-
-% NOTES:
-% - Pyramide / decision tree Graphik
-% - Confidence Interval calculation
-% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
-%
-% --> explain each point briefly without to much (mathematical) detail
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Estimators and their Application -------------------------- %
-
-%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Global Estimators and their Application in R]{Global Estimators and their Application}
-\label{sec:globest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% Note: As the mathematical form of estimators is still quite simple and short, we can here
-%       focus on presenting and explaining important 'components' of the estimators, such as Z, C, ...)
-% 
-% \subsection{Mathematical Background}
-% 
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-% 
-% 
-% \subsubsection{Application}
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
-
-\subsubsection{Mathematical Background}
-
-Mention:
-\begin{itemize}
-  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
-
-\begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
-\end{equation}
-
-The design-based variance-covariance matrix of the regression coefficients is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
-\end{equation}
-
-with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
-\begin{equation}\label{eq:resids}
-  \hat{R}(x)=Y(x)-\hat{Y}(x)
-\end{equation}
-
-The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
-Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to
-
-\begin{equation}\label{eq:pointest_simple}
-\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
-\end{equation}
-
-The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
-The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{equation}\label{eq:gw_var_simple}
-\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-\end{equation}
-
-\begin{equation}\label{eq:varexternalsimple}
-\hat{\var}(\hat{Y}_{reg})=
-\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
-\end{equation}
-
-\subsubsection{Application}
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-\subsubsection{Mathematical Background}
-
-\subsubsection{Application}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
-\label{sec:saeest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% - 
-%
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
-
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
-%
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% NOTES:
-% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
-% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
-%
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Special Cases and Scenarios ------------------------------- %
-
-% NOTES:
-% - Error handling and adjustments: Sampling designs not nested
-% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
-%                   or s2 or/and s1 sample in 3-phase case
-% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
-
-% NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
-
-%------------------------------------------------------------------------------------------------%
-% ----------------------------- Visualization and Analysis of Estimation Results --------------- %
-
-% NOTES:
-% - demonstrate visualization of errors and point-ests + CIs
-% - demonstrate a) releff-calculation and gain-analysis
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------------- Literature ------------------------------------------ %
-
-\bibliography{bib/literature}\newpage\cleardoublepage
-
-% -------------------------------------------------------------------------- %
-\end{document}
-
-Hallo. Here comes a first try:
-
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
-
-This was the result

### 2017-07-28 texfiles Daniel - orig/kernelestimator3.tex
upload all files

\documentclass[a4paper,12pt,leqno, titlepage]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage[margin=0.5in]{geometry}
\bibliographystyle{mystyle2}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\VAR}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\PLAN}{\R^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}
\setlength{\topmargin}{0cm}
\setlength{\headheight}{3mm}
\setlength{\headsep}{1cm}
\setlength{\topskip}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{22cm}
\setlength{\oddsidemargin}{0cm}
\makeatletter
\def\tagform@#1{\maketag@@@{[\ignorespaces#1\unskip\@@italiccorr]}}
\makeatother

\begin{document}
\pagestyle{empty}
\begin{center}
\large
$ $\\[5cm]
\textbf{Regression and non-parametric estimators for two-phase forest inventories in the design-based Monte-Carlo approach}\\[2cm]
\normalsize
Daniel Mandallaz and Alexander Massey \\[2cm]
Chair of Land Use Engineering \\[1cm]
Department of Environmental Systems Science \\[1cm]
ETH Zurich\\[1cm]
April 2015
\end{center}
\newpage

\nolinenumbers
\begin{abstract}
This technical report adapts rank based nearest neighbor methods to classical kernel estimators using neighborhoods defined by a metric in the feature space resting upon the predictions obtained from a linear model. The approach is purely design-based. The main concepts of kernel-based density and regression estimators are summarized and reformulated in the design-based Monte-Carlo approach to forest inventory. New estimators are proposed which are based on the conditional expectation of the response variable given its prediction and are compared to the classical two-phase regression estimator and to two-phase k-nearest neighbor (k-NN) estimators, using either a multidimensional metric or a one-dimensional metric based on the Euclidian distance between the predictions. The results suggest that the commonly used analytical external variance formula may systematically underestimate the true variance for a variety of nonparametric kernel based estimators including kNN, but is still adequate for the classical regression estimator. Although using a bootstrap variance can help to correct this underestimation it was also found that the bootstrap variance estimates can be unstable if the bandwidth or the number of neighbors are recalculated in each bootstrap sample. These findings suggest that if the model captures the main features of the underlying process, which is clearly the case for timber volume estimates based on canopy characteristics obtained from LiDAR measurements, then it is advisable to use the classical regression estimator because it performs at least as well as the other techniques and is by far the simplest to implement.
\end{abstract}
\clearpage
\pagestyle{plain}
\pagenumbering{arabic}
\section{Introduction}\label{intro}
\pagenumbering{arabic} \setcounter{page}{1}
In this note we use the Monte-Carlo approach (infinite population model), as described in \cite{mandallaz} (Chapter 5). We consider the forested area $F$ of surface area $\lambda(F)$ for which the sum $Y_F=\sum_{i=1}^N Y_i$ of the response variable $Y_i$ for a well defined population of $N$ trees, e.g. the total volume of the $N$ trees, and the spatial mean $\bar{Y}_F=\frac{1}{\lambda(F)}\sum_{i=1}^N Y_i$ must be estimated.

\section{Two-phase sampling}\label{twophase}
We have a sample $s_1$ of $n_1$ points independently and uniformly distributed in $F$ out of which $n_2$ are selected by simple random sampling without replacement. At all points $x\in{s_{1}}$ we have the vector of auxiliary information $\pmb{Z}(x)\in{\R^p}$ and at all terrestrial points $x\in{s_{2}}$  we have the local density $Y(x)$ which is essentially a Horwitz-Thompson estimator at point $x\in{\R^2}$. In the Monte-Carlo approach we have the fundamental relationship
   \begin{equation}\label{montecarlo}
   \frac{1}{\lambda(F)}\int_F Y(x)dx=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i
   \end{equation}
    We consider the linear model

   \begin{equation}\label{linearmodel1}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R(x)
 \end{equation}
 In the \textbf{model-dependent approach} the point $x$ is fixed and $R(x)$ is a random variable with zero mean and a given covariance structure (in most instances very simple and not really adapted to the spatial nature of the problem in contrast to geostatistical techniques). In the \textbf{design-based approach} $Y(x)$, $\pmb{Z}(x)$ and $R(x)$ are random variables  because $x$ is random. According to the terminology of \cite{sarndal} our approach is \textbf{model-assisted}, i.e. we use models to reduce the variance but we do not assume that they are correct. The inference is design-based and valid because it rests upon the randomization principle. Some authors call model-based the model-dependent approach, which is in our opinion a source of confusion because the inference is valid only if the model is true. Here, the true regression coefficient $\pmb{\beta}$ is by definition the theoretical least squares estimate minimizing
 $$\int_F R^2(x)dx=\int_F(Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2 dx$$
 It satisfies the normal equation
 \begin{equation}\label{normaleq1}
 \Big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\Big)\pmb{\beta}=\int_F Y(x)\pmb{Z}(x)dx
 \end{equation}
 and the orthogonality relationship
 \begin{equation}\label{linearmodel2}
 \int_F R(x)\pmb{Z}(x)dx=\pmb{0}
 \end{equation}
  We shall assume that $\pmb{Z}(x)$ contains the intercept term $1$, or, more generally, that the intercept can be expressed as a linear combination of the components of $\pmb{Z}(x)$, which then ensures that the mean residual is zero, i.e. $\int_F R(x)dx=0$. To simplify the notation we define the following
  \begin{eqnarray}
  \pmb{U}(x)&:=&Y(x)\pmb{Z}(x) \nonumber \\
  \pmb{A}_{s_2}&:=&\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x)\nonumber \\
  \pmb{U}_{s_2}&:=&\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{U}(x)
  \end{eqnarray}
 The theoretical and empirical regression vector parameters can then be written as
 \begin{eqnarray}\label{linearmodel3}
 \pmb{\beta}&:=&\pmb{A}^{-1}\pmb{U} \nonumber \\
 \hat{\pmb{\beta}}_{s_2}&:=&\pmb{A}_{s_2}^{-1}\pmb{U}_{s_2}
 \end{eqnarray}
 where we have set $\pmb{A}=\EX_{x\in{F}}\pmb{Z}(x)\pmb{Z}^t(x)$ and $\pmb{U}=\EX_{x\in{F}}\pmb{Z}(x)Y(x)$.
$\hat{\pmb{\beta}}_{s_2}$ is asymptotically design-unbiased for $\pmb{\beta}$. \\
 The empirical predictions and residuals are $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and $\hat{R}(x)=Y(x)-\hat{Y}(x)$. The asymptotic design-based variance-covariance matrix of $\hat{\pmb{\beta}}_{s_2}$ is
 \begin{equation}\label{estvarmatrix}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\pmb{A}_{s_2}^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}_{s_2}^{-1}
\end{equation}
and the two-phase classical regression estimator is
  \begin{eqnarray}\label{regressionestimators}
\hat{Y}_{reg}&=&\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\hat{Y}(x)+\frac{1}{n_{2}}
\sum_{x\in{s_{2}}}\hat{R}(x)\nonumber
\end{eqnarray}
\noindent
In the following we shall work under the external model assumption, i.e. we neglect the error in $\hat{\pmb{\beta}}_{s_2}$ and set formally $\hat{\pmb{\beta}}_{s_2}\equiv \pmb{\beta}$, which is asymptotically equivalent to the variance obtained via the design-based variance covariance matrix (\cite{mandallaz}, p. 125). One has the theoretical external variance
\begin{eqnarray}\label{varreg2}
\VAR(\hat{Y}_{reg})& =&
\frac{1}{n_{1}}\VAR_{x\in{F}}(Y(x))+(1-\frac{n_{2}}{n_{1}})\frac{1}{n_{2}}\VAR_{x\in{F}}\hat{R}(x)\nonumber \\
& =&
\frac{1}{n_{1}}\VAR_{x\in{F}}(\hat{Y}(x))+\frac{1}{n_{2}}\VAR_{x\in{F}}\hat{R}(x)
\end{eqnarray}
Sample copies of the second version yields the external variance estimate asymptotically equivalent to the g-weight variance based on $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ discussed in
(\cite{mandallaz} and \cite{mandallaz3},), which has slightly better statistical properties from a theoretical point of view. The working model $\pmb{Z}(x)$ yielding $\hat{Y}(x)$ does not have to be identical with the true model model generating $Y(x)$. On the other hand, the closer the working model comes to the true model the smaller the residual variance and therefore the error of the regression estimator.\\
 In this report the predictions are based on linear models, external or internal, the difference being asymptotically negligible. It is intuitively clear that the predictions $\hat{Y}(x)$ could also be based on non-linear models, which leads to so-called calibration model discussed by \cite{wusitter1}. The proof of this result is sketched in Appendix \ref{calibration}.

\section{Kernel-based estimators}\label{kernelbased}
In this section we consider non-parametric kernel based regression estimator. The idea is very simple: if the working model departs from the true unknown model, e.g. because relevant explanatory variables have not been incorporated into the working model due to either inadequate or unavailable hidden variables, then one can attempt to combine locally $\hat{Y}(x)$ and $Y(x)$ to reduce the residual size $\mid Y(x)-\hat{Y}(x)\mid $.\\
\subsection{Preliminaries}\label{preliminaries}
We first consider one dimensional kernels
\begin{equation}
K: u\in{\R}\rightarrow K(u)>0
\end{equation}
The kernel $K()$ is a probability density function with zero mean and finite variance, i.e $\int_\R K(u)du=1$, $\int_\R uk(u)du=0$ and $\int_\R u^2K(u)du <\infty$. Popular choices are
\begin{eqnarray}\label{threekernels}
K(u)&=& 0.5 I_{[-1,1]}(u)\quad \text{uniform kernel} \nonumber \\
K(u)&=& \frac{1}{\sqrt{2\pi}}\exp(-u^2/2)  \quad \text{normal kernel} \nonumber \\
K(u)&=& \frac{3}{4\sqrt{5}}(1-\frac{u^2}{5})I_{\mid u \mid \le \sqrt{5}}(u)\quad \text{Epanechnikov kernel}
\end{eqnarray}
The first and third kernels have a finite support whereas the normal kernel does not. The concept can be easily generalized to multidimensional kernel $K(\pmb{u})$ with $\pmb{u}=(u_1,u_2,\ldots u_p)\in{\R}^p$, by setting $k(\pmb{u})= \prod_{l=1}^pK_l(u_l)$ where the $K_l(\cdot)$ are 1-d kernels.\\
For a sample size $n_2$ (or respectively $n_1$ ) we can use a global bandwidth $\epsilon(n_2)\rightarrow 0$ or a local bandwidth sequence $\epsilon(n_2,x_0)$ ($x_0$ being a given point $x_0\in{F}$ for which one wants to get estimates) where the dependence on $x_0$ is usually via $Y(x_0)$, $\hat{Y}(x_0)$ or $R(x_0)$ depending on the context. Many results follow intuitively from the fact that $K(\frac{u}{\varepsilon(n_2,x_0)})$ tends to the Dirac $\delta$ function as $\epsilon(n_2,x_0)$ tends to $0$. \\
In the nearest neighbor literature a wide choice of metric in the feature space (i.e. distances between points $\pmb{Z}(u)$ and $\pmb{Z}(v)$ for any two $u,v\in{F}$) have been used. From a geometric point of view one can always orthogonalize the $\pmb{Z}(u)$ so that the Euclidian and Mahalanobis distances are essentially equivalent. Also, it is clear that
$$\mid\mid \pmb{Z}(u)-\pmb{Z}(v)\mid \mid=:\mid \pmb{\beta}^t(\pmb{Z}(u)-\pmb{Z}(v))\mid=\mid \hat{Y}(u)-\hat{Y}(v)\mid$$ also defines a metric in the feature space, which is for any given $\pmb{\beta}$ and in particular for the theoretical least squares estimate, a natural choice for the problem at hand. Furthermore, the multidimensional feature space is reduced to the one-dimensional space of the predictions, thus avoiding the use of multidimensional kernels, the curse of dimensionality, as well as the difficulties arising from the different scales of the explanatory variables. One should also recall that in finite dimensional spaces, all norms are equivalent from a topological point of view ($\mid \mid\pmb{Z}(u)\mid \mid:=\mid \pmb{\beta}^t\pmb{Z}(u)\mid$ is a norm). Of course, in practice one will usually replace $\pmb{\beta}$ by its estimate and do as if it were a fixed value, as under the external model assumption. \\
We shall consider the kernel-based prediction $\hat{Y}^{(1)}_{\epsilon}(x_0)$ defined point-wise for a point $x_0\in{s_1}$ as
\begin{equation}\label{Yregepsilon}
\hat{Y}^{(1)}_{\epsilon}(x_0):=\frac{ \frac{1}{n_2\epsilon(n_2,\hat{y}_0)}
\sum_{x\in{s_2}}Y(x)K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,\hat{y}_0)}\big)}
{\frac{1}{n_2\epsilon(n_2,\hat{y}_0)}\sum_{x\in{s_2}}K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,\hat{y}_0)}\big)}
\end{equation}
where $\hat{y}_0=\hat{Y}(x_0)$. If the numerator in [\ref{Yregepsilon}] is zero then one sets $\hat{Y}^{(1)}_{\epsilon}(x_0)=\hat{Y}(x_0)$.\\
$\hat{Y}^{(1)}_{\epsilon}(x_0)$ is the Nadaraya-Watson estimator based on the euclidian distance between linear predictions and can be viewed as an estimate of the conditional expectation of $Y(x_0)$ given $\hat{Y}(x_0)$ (see section \ref{dbestimators} and Appendix \ref{appendixnadaraya}).\\
It is interesting to consider the special case in which the $\hat{Y}(x)$ can only take a finite number of different values, such as in post-stratification, still an important and widely used technique in forest inventory. Recall that in post-stratification $\hat{Y}(x)$ is simply the observed stratum mean. For $\epsilon(n_2,\hat{y}_0)$ small enough (very moderate smoothing) it is easy to see that $\hat{Y}^{(1)}_{\epsilon}(\hat{y}_0)$ will tend to the mean of the $Y(x)$ for the $x$ in the same stratum $F_k\subset F$ as $x_0$. If the $\hat{Y}(x)$ (for $x\in{s_1}$ ) are all distinct then the limit for $\epsilon(n_2,\hat{y}_0)\rightarrow 0$ is $Y(x_0)$ for $x_0\in{s_2}$. when the bandwidth $\epsilon(n_2,\hat{y}_0)$ tends to infinity (strong smoothing) then $\hat{Y}^{(1)}_{\epsilon}(x_0)$ will tend to the overall sample mean $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)=\bar{Y}_{s_2}$, so that the model is completely ignored. \\
Note that if all explanatory variables are discrete then $\pmb{Z}(u)$  can be viewed as a set of $0-1$ indicator variables and the distance between two points will also take a finite number of values (e.g. say $0,1$ in the case of one-way analysis of variance and $0,1,\ldots q$ in general ANOVA models). Hence, kernel estimators are likely to be potentially useful only if some of the explanatory variables are continuous so that the set of local densities and predictions $(Y(x),\hat{Y}(x))\in{\R^2}$ can be viewed as a continuum from a mathematical point of view (in practice everything is discrete!).\\
For the uniform kernel $\hat{Y}^{(1)}_{\epsilon}(x_0)$ is simply the mean of all observations in the $\varepsilon(n_2,x_0)$ neighborhood, whereas the k-nearest neighbors methods takes the mean of the observations whose ranked distances are less than the integer $k$. With discrete metrics one has ties in the ranked distances and one could raise the question whether it is meaningful to take the mean of only $k$ observations if there are more with the same ranks. \\
For the above reasons and for mathematical convenience we shall assume in the following that $\hat{Y}(x)$ and $Y(x)$ have sufficiently regular probability density functions (e.g. twice differentiable), even if the proposed kernel estimators can be used in practice when this is not the case. In other words we have
\begin{equation}\label{regulardensities1}
\PR(\hat{Y}(x)\in{A})=\int_A f_{\hat{Y}}(y)dy, \;\;\forall A\subset \R
\end{equation}
\begin{equation}\label{regulardensities2}
\PR( Y(x)\in{A})=\int_A f_{Y}(y)dy, \;\;\forall A\subset \R
\end{equation}
Note that this implies that $\PR(Y(x_1)=Y(x_2))=0$ for all $x_1 \ne x_2$ and likewise for $\hat{Y}(\cdot)$.
Under some regularity assumptions one can show for $\hat{Y}(x)$ with a continuous distribution function on $\R$ that one has $\hat{Y}_{\epsilon}(x_0)\rightarrow Y(x_0)$ as $\epsilon(n_2,x_0)\rightarrow 0$, convergence being in the mean, probability, almost surely, point-wise or uniform, depending on the regularity assumptions (see \cite{Tsybakov1} for a modern introduction).\\
We define the two-phase kernel-based regression estimator as
\begin{equation}\label{kernelyreg}
\hat{Y}^{(1)}_{\epsilon,reg}:=\frac{1}{n_1}\sum_{x_0\in{s_1}}\hat{Y}^{(1)}_{\epsilon}(x_0)
+\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}^{(1)}_{\epsilon}(x)
\end{equation}
where $\hat{R}^{(1)}_{\epsilon}(x)=Y(x)-\hat{Y}^{(1)}_{\epsilon}(x)$. Note that the sum of the residuals is no longer zero in general.
 \\By analogy with the classical theory one can hope that $\hat{Y}^{(1)}_{\epsilon,reg}$ is asymptotically design unbiased with the following design-based variance estimate
\begin{equation}\label{estvarkernelyreg}
\hat{\VAR}(\hat{Y}^{1)}_{\epsilon,reg})=\frac{1}{n_1}\hat{\VAR}(Y(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}
\sum_{x\in{s_2}}(\hat{R}^{(1)}_{\epsilon}(x)-\bar{\hat{R}}^{(1)}_{\epsilon,2})^2
\end{equation}
where
$$\hat{\VAR}(Y(x))=\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2, \; \bar{Y}_2=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x),\;\bar{\hat{R}}^{(1)}_{\epsilon,2}=\frac{1}{n_2}
\sum_{x\in{s_2}}\hat{R}^{(1)}_{\epsilon}(x)$$.

\subsection{Design-based kernel regression estimators}\label{dbestimators}
There is an immense literature on kernel-based estimators but the overwhelming majority of the papers are in the model-dependent one-dimensional framework, i.e. $Y(x_i)=m(x_i)+\epsilon(x_i), \;x_i\in{Re^1}$, $i=1,2\ldots n$, with $m(x_i)=\EX(Y(x_i)\mid x_i)$ and $\VAR(Y(x_i)\mid x_i)=\sigma^2(x_i)$, the $x_i$ may be fixed or randomly independently distributed on $\R^1$. Informal proofs of the most important results in the standard set-up are given in Appendix \ref{appendixnadaraya} and Appendix \ref{appendixdesignbased}.\\
Recall that in general the best predictor is given by the conditional expectation. Ideally one would therefore consider the conditional expectation of $Y(x)$ given the auxiliary information, i.e. $\EX(Y(x)\mid\pmb{Z}(x)\in\R^p)$ which requires the use of multivariate kernels to estimate multivariate densities and, as afore mentioned, is subject to the curse of dimensionality problem when $p$ is large (in practice one rarely goes beyond $p=2$). In some sense one can view $\hat{Y}(x)\in\R^1$ as a concise summary of $\pmb{Z}(x)\in\R^p$ which leads us to consider instead
\begin{equation}\label{startconditionalexpectation}
 \EX(Y(x)\mid\hat{Y}(x))=\EX(\hat{Y}(x)+R(x)\mid\hat{Y}(x))= \hat{Y}(x)+\EX(R(x)\mid\hat{Y}(x))
 \end{equation}
 To have an approximation of this conditional expectation we need estimates of the joint bivariate density of $(R(x),\hat{Y}(x))$ and of the marginal density of $\hat{Y}(x)$. This is done with bivariate and univariate Nadaraya-Watson estimators to obtain first an estimate of the conditional density and then, by integration, the estimate of the conditional expectation (see Appendix \ref{appendixnadaraya} for details). Asymptotically we replace the true predictions $\hat{Y}(x)=\pmb{\beta}^t\pmb{Z}(x)$ and residuals $R(x)=Y(x)-\pmb{\beta}^t\pmb{Z}(x)$ by their empirical versions, e.g. after replacing $\pmb{\beta}$ by $\hat{\pmb{\beta}}_{s_2}$. Using the first term in [\ref{startconditionalexpectation}] this leads to the estimator $\hat{Y}^{(1)}_{\epsilon}(x_0)$  given in [\ref{Yregepsilon}] and using the third  we have the following estimate for the conditional expectation of the residual
 \begin{eqnarray}\label{conditionalmeanofresiduals2}
 \hat{\EX}(\hat{R}(x_0)\mid\hat{Y}(x_0))&=&\frac{\frac{1}{n_2\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\sum_{x\in{s_2}}\hat{R}(x)
 K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\big)}{\frac{1}{n_2\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\sum_{x\in{s_2}}
 K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\big)}\nonumber\\
 &=&\frac{\sum_{x\in{s_2}}\hat{R}(x)
 K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\big)}{\sum_{x\in{s_2}}
 K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon_{\hat{y}}(n_2,\hat{y}_0)}\big)} \nonumber\\
 &=&:\hat{R}^{(2)}_{\epsilon}(x_0) \;\; \forall x_0\in{s_1}
 \end{eqnarray}
 and consequently the following kernel-based estimators at $x_0$ at any point $x_0\in{s_1}$ in the large sample
 \begin{equation}\label{Yepsilonstar}
 \hat{Y}^{(2)}_{\epsilon}(x_0)= \hat{Y}(x_0)+ \hat{R}^{(2)}_{\epsilon}(x_0)
 \end{equation}
From a formal point of view $\hat{Y}_{\epsilon}^{(2)}(x_0)$ has a slight advantage over $\hat{Y}_{\epsilon}^{(1)}(x_0)$ when the bandwidth goes to $\infty$ as it tends to $\hat{Y}(x_0)$ whereas
$\hat{Y}_{kernel}^{(1)}(x_0)$ tends to the overall sample mean $\bar{Y}_{s_2}$. This also means that
$\hat{Y}_{\epsilon}^{(2)}(x_0)$ can escape the range of observations $Y(x)$ where $x\in{s_2}$ whereas $\hat{Y}_{\epsilon}^{(1)}(x_0)$  can not.\\
 Note that the probability density function of the predictions could also be estimated in the large sample $s_1$, a possibility we shall not pursue further here, primarily because it has not been investigated yet in the literature and because it is not available in standard software (such as \textbf{R}), at least not to our knowledge.\\

 \noindent \textbf{Mathematical difficulties}\\
The true predictions $\hat{Y}_0(x)=\pmb{\beta}^t\pmb{Z}(x)$ and observations $Y(x)$ are i.i.d.  but the points $(\hat{Y}_0(x),Y(x))$ are on a 1-d curve in $\R^2$ (possibly with discontinuities) so that one has a degenerate bivariate distribution. This is not the case with the empirical predictions $\hat{Y}(x)=\hat{\pmb{\beta}}_{s_2}^t\pmb{Z}(x)$, which are only asymptotically uncorrelated so that the formulae for the variances of the various kernel-based estimators (given in Appendix \ref{appendixnadaraya} and Appendix \ref{appendixdesignbased}) are probably at best only asymptotically valid, possibly  with a lower convergence rate. On the other hand, the kernel-estimators have a strong intuitive background, and they remain meaningful even if the set of predictions is discrete.\\

Neglecting the correlation between the predictions (asymptotically of order $O(\frac{\varepsilon(n_2,\hat{y}_0)}{n_2})$), the results given in Appendix \ref{appendixdesignbased} show that the convergence rate at $x_0$ is of order $O(\frac{\varepsilon(n_2,\hat{y}_0)}{n_2})$. It can be conjectured that the convergence rate will be slower if one takes the correlations of the predictions into account, but this seems to be beyond simple analytical treatment. In any case, one often relies in practice on the external model approach and use as usual [\ref{kernelyreg}] and [\ref{estvarkernelyreg}] with
$\hat{Y}^{(2)}_{\epsilon}(x_0)$ and $\hat{R}^{(2)}_{\epsilon}(x_0)$. It turns out that the external variance can severely underestimates the true variance, except for the classical regression estimator $\hat{Y}_{reg}$, and that one must use bootstrap to obtain correct variance estimates and confidence intervals. \\[0.5cm]
\noindent
 \textbf{Remarks:}\\
 \begin{enumerate}
 \item
 In the case of post-stratification $\hat{R}^{(2)}_{\epsilon}(x_0)$ will tend to the sum of the residuals in the stratum of $x_0$ as $\epsilon_{\hat{y}}(n_2,\hat{y}_0)\to 0$,and therefore to zero. Hence, in this case, $\hat{Y}^{(2)}_{\epsilon}(x_0)\to \hat{Y}(x_0)$ and nothing is gained.  If the predictions are all distinct it will tend to $\hat{R}(x_0)$ for $x_0\in{s_2}$ and
 $\hat{Y}^{(2)}_{\epsilon}(x_0)\to Y(x_0)$.
 \item
 If the predictions are all different and $\epsilon_{\hat{y}}(n_2,\hat{y}_0)\to 0$ then $\hat{R}^{(2)}_{\epsilon}(x_0)$ will tend to $\hat{R}(x_0)$ for $x_0\in{s_2}$ and to $0$ if $x_0\in{s_1\setminus s_2}$. Hence, $\hat{Y}^{(2)}_{\epsilon}(x_0)\rightarrow Y(x_0) \;\forall x_0\in{s_2}$ and $\hat{Y}^{(2)}_{\epsilon}(x_0)\rightarrow \hat{Y}(x_0) \;\forall x_0\in{s_1\setminus s_2}$. The resulting regression estimator is then $\frac{1}{n_1}\sum_{x_0\in{s_1}}\hat{Y}(x_0)$ with estimated external variance $\frac{1}{n_1}\hat{\VAR}(Y(x))$ because all the residuals in $s_2$ are zero! This is of course to good to be true! The reason is that to ensure consistency of the estimated densities we cannot simply set $\epsilon_{\hat{y}}(n_2,\hat{y}_0)=0$, one needs also
$\lim_{n_2\to\infty}n_2\varepsilon_{\hat{y}}(n_2,\hat{y}_0)=\infty$. Hence, we cannot determine the bandwidth by minimizing the resulting variance of the regression estimator (the minimum being achieved with zero residual variance) but we have to rely instead on cross-validation procedures. Note also that for $\epsilon_{\hat{y}}(n_2,\hat{y}_0)\to \infty$ one has $\hat{Y}^{(2)}_{\epsilon}(x_0)\to \hat{Y}(x_0)$ and $\hat{Y}^{(1)}_{\epsilon}(x_0)\to \bar{Y}_{s_2}$.
 \item
  If some important explanatory variables are not accounted for it may well be that locally we do not have zero mean residual (i.e. a pattern in the structure of the residuals), in which case the new kernel-based regression estimators might be better. Note, that correct exploratory data analysis should identify such problems. If the choice of the explanatory variables is adequate one can hope that the kernel-based point estimates will be close to the classical regression estimate. The simulations and the case study discussed in section \ref{simandcasestudy} and others in the literature, provide some evidence for that, as well as for the \textbf{k-NN} methods. For the variance estimate, the situation is more intricate and analytical derivations of the design-based variance are, so far, only available for $\hat{Y}_{reg}$.

 \end{enumerate}

 Note that the kernel-based regression estimator $\hat{Y}^{(1)}_{\epsilon}(x_0)$ is a convex linear combinations of the observations (i.e. a weighted average of the observations,  where the weights are positive and sum up to $1$). Hence, the range of the predictions is always contained in the range of the observations, which can be a disadvantage if  the prediction model can be trusted. The estimator
 $\hat{Y}^{(2)}_{\epsilon}(x_0)$ does not have this disadvantage. Other estimators using kernels but not in the same way as $\hat{Y}^{(1)}_{\epsilon}(x_0)$ lead to non-convex linear combinations, such as the widely used \textbf{R}-functions \textbf{glkern} (with global optimal bandwidth) and \textbf{lokern} (with local optimal bandwidth) where the option "derivative=0" is used. The resulting estimators are known to have good performances (see \cite{gasser1} and \cite{gasser2} for details). In this work we have used the \textbf{R}-function \textbf{npreg} to calculate $\hat{Y}^{(1)}_{\epsilon}(x_0)$ and the residual component of $\hat{Y}^{(2)}_{\epsilon}(x_0)$. The global optimal bandwidth is obtained by the standard leave one out cross-validation technique, which can be calculated very efficiently for $\hat{Y}^{(1)}_{\epsilon}(x_0)$ but not for $\hat{Y}^{(2)}_{\epsilon}(x_0)$, an unfortunate development for variance estimation via bootstrap.\\
 The overwhelming majority of kernel based non-parametric curve-fitting procedures are primarily concerned with interpolation, whereas in the forest inventory context with two-phase sampling the extrapolation aspects are also very important. In the case study below we will see that the mean-canopy height obtained by remote sensing (LiDAR) in the large sample $s_1$ is the most important explanatory variable, leading single-handedly to $R^2$ above $50\%$, and it clear that the highest values will most likely be in $s_1$ and not in $s_2$. Thus, forcing the prediction range to be contained in the observation range is a disadvantage. As afore mentioned, this is not the case for $\hat{Y}^{(2)}_{\epsilon}(x_0)$.

 \section{Design-based k-NN}\label{knn}
As for kernel-based estimators the literature on k-Nearest Neighbors (\textbf{k-NN}) is immense and largely model-dependent. In the forest inventory context \cite{bafetta1,bafetta2} are key references in the design-based context whereas \cite{McRoberts1}, \cite{McRoberts2}, \cite{magnussen1}, \cite{MacRoberts3}, \cite{breidenbach1} are more model-dependent. Also, most authors work in the finite population framework of pixels in which the first phase is exhaustive.\\
We shall use here the procedure described by \cite{hechenbichler1}, which is implemented in the \textbf{R} program \textbf{kknn}. For an arbitrary point $x_0\in{s_1}$ the nearest neighbor in $s_2$, with respect to a distance $d(\cdot,\cdot)$ in $\R^p$ is the point $x_{(1)}\in{s_2}$ such that $d(\pmb{Z}(x_0),\pmb{Z}(x_{(1)})=\min_{x\in{s_2}}d(\pmb{Z}(x_0),\pmb{Z}(x))$. The second nearest neighbor $x_{(2)}$ is defined by
$d(\pmb{Z}(x_0),\pmb{Z}(x_{(2)})=\min_{ x\in{s_2}\setminus{x_{(1)}}}d(\pmb{Z}(x_0),\pmb{Z}(x))$
and so on until we have obtained the $k+1$ nearest neighbors. We emphasize the fact that if $x_0$ is in $s_2$ then we set $x_{(1)}=x_0$. If the components of $\pmb{Z}(x)$ have very different scales it is usually recommended to standardize the component so that they all have standard deviation equal to $1$.\\
The simulations and case study presented in section \ref{simandcasestudy} contain only continuous variables so that ties do not occur. The k-nn methods with the uniform kernel simply take the mean $\frac{1}{k}\sum_{l=1}^k Y(x_{(l)})=\hat{Y}_{knn}(x_0)$. In this case it is possible, in principle, to calculate the theoretical variance of the corresponding regression estimator, albeit in the finite population case with exhaustive first phase (see \cite{bafetta1}). However, it is usually more efficient to give larger weights to the closest neighbors. To achieve that we have to standardize
the distances according to
\begin{equation}\label{knnstandarddist}
D(\pmb{Z}(x_0),\pmb{Z}(x_{(i)})=\frac{d(\pmb{Z}(x_0),\pmb{Z}(x_{(i)})}{d(\pmb{Z}(x_0),\pmb{Z}(x_{(k+1)})}
\end{equation}
and the weighted k-NN estimator is given by
\begin{equation}\label{kernelknn}
\hat{Y}_{knn}(x_0)=\frac{\sum_{l=1}^k K\big(D(\pmb{Z}(x_0),\pmb{Z}(x_{(i)})\big)Y(x_{(i)})}
{\sum_{l=1}^k K\big(D(\pmb{Z}(x_0),\pmb{Z}(x_{(i)})\big)}
\end{equation}
for an arbitrary one-dimensional kernel $K(\cdot)$ (Note that the kernel is one-dimensional whereas the distance is calculated in the multidimensional feature space). This leads to the \textbf{k-NN regression estimator}
\begin{equation}\label{knnreg}
\hat{Y}_{knn}=\frac{1}{n_1}\sum_{x_0\in{s_1}}\hat{Y}_{knn}(x_0)+
\frac{1}{n_2}\sum_{x_0\in{s_2}}(Y(x_0)-\hat{Y}_{knn}(x_0))
\end{equation}
Note that the predictions, $\hat{Y}_{knn}(x_0)$, are a weighted mean of the observations and therefore constrained to be in the observed range of the observations. Thus \textbf{k-nn} is a purely interpolation technique and never extrapolates, which, as already mentioned, can be a disadvantage.\\
As for kernel-based regression estimator one can use the one dimensional metric based on the predictions alone, i.e. $d(\pmb{Z}(x_0),\pmb{Z}(x_{(i)}))=\mid \hat{\pmb{\beta}}^t\big(\pmb{Z}(x_0)-\pmb{Z}(x_{(i)})\big)\mid$. The resulting estimator is denoted by
$\hat{Y}_{pred,knn}$.\\
The external variance estimates are obtained as usual via sample copies of [\ref{varreg2}].\\
In practice one has to be careful about the tuning options available in software packages, in particular
with respect to the definition of nearest neighbors when $x_0\in{s_2}$. If the point itself is viewed as its nearest neighbor the optimal choices $k=1$ can result from cross-validation and should of course not be retained as it its leads to constant $0$ residuals. We recommend at least $k \ge 3$ and to plot $\hat{Y}_{knn}$ and $\hat{\VAR}(\hat{Y}_{knn})$ as a function of $k$. In our experience, except for $\hat{Y}_{reg}$, the external variance estimate is to small and the bootstrap variance estimate should be preferred.
\clearpage\newpage

 \section{Examples}\label{simandcasestudy}
To illustrate the various techniques we consider the real case study discussed in \cite{mandallaz4} and the artificial simulation example used in \cite{mandallaz3}.\\
We have compared several kernel-based estimators calculated with the widely used \textbf{R}-functions \textbf{glkern}, \textbf{lokern} (both from the \textbf{R}-package \textbf{lokern}), \textbf{npreg} (from the \textbf{R}-package \textbf{np}) and the \textbf{k-NN} estimators calculated with the \textbf{R}-package \textbf{kknn}. In the main text we do not discuss the function \textbf{lokern} because it was found that the local optimal bandwidth led to very spurious results near the boundaries, particularly at the lower end (see e.g. Fig. \ref{predict1} of Appendix \ref{appendixfigures} obtained from the case study).\\
The two-phase bootstrap procedure is very simple: first a bootstrap sample $s_{1boot}$ is generated from $s_1$ and the points $x\in{s_{1boot}}$ that are in $s_2$ define the the second-phase bootstrap sample $s_{2boot}$.\\
The bootstrap variances with optimal bandwidth recalculated in each bootstrap sample were very unstable and frequently led to extreme outliers (above 1000 times the mean), particularly for \textbf{lokern} as can be seen from Fig. \ref{boot1} and \ref{boot2} in the Appendix \ref{appendixfigures} based on the simulation example. This was also the case, but not as extreme, for the case study as can be seen from Tables \ref{appcasestudy1} and \ref{appcasestudy2} given in the Appendix \ref{appendixtables}. \\
For $\hat{Y}^{(2)}_{glkern}$ the \textbf{R}-function \textbf{glkern} was applied to the residual part only and the optimal bandwidth was obtained via the imbedded cross-validation procedure. For
$\hat{Y}^{(2)}_{npreg}$ the optimal bandwidth obtained via the imbedded cross-validation procedure and applied to only the residual part led essentially to a flat horizontal line close to $0$ (i.e. making $\hat{Y}^{(2)}_{npreg}\approx \hat{Y}_{reg}$). For this reason we wrote our own very simple cross-validation procedure based on the mean squared sequentially deleting each $x_0\in{s_2}$ and refitting both $\hat{Y}(x_0)$ using the linear model as well as the smoothed residual part (using \textbf{npreg} on the second term) to get the optimal bandwidth out of a small set of feasible values. \\ The resulting $\hat{Y}^{(2)}_{npreg}$ then differs slightly from $\hat{Y}_{reg}$. The same procedure was used for the simulation presented in section \ref{simul}.

 \subsection{Case study}\label{casestudy}

  The auxiliary vector $\pmb{Z}(x)\in{\R^p}$ has $p=7$ components: $Z_0(x)\equiv 1$, mean canopy height $Z_1(x)$, maximal canopy height $Z_2(x)$, $75\%$ quantile of the canopy height, $Z_3(x)$, standard deviation of canopy height $Z_4(x)$, the LiDAR estimated volume density $Z_5(x)$ and the LiDAR estimated density of stems $Z_6(x)$. The sample sizes were $n_1=306,n_2=67$. The fitted model is
 \begin{equation*}
 \hat{Y}(x)=322.57 + 52.55Z_1(x) -19.24Z_2(x) -33.04Z_3(x)+71.06Z_4(x) +0.19Z_5(x)-0.09Z_6
 \end{equation*}
 The coefficient of determination for this model is $R^2=0.64$. \\
 For the estimator $\hat{Y}_{knn}$ we use the gaussian kernel with $k=7$, which is the optimal value obtained from the leave-one-out cross-validation procedure. Fig. \ref{choiceofkcasestudy} in the Appendix \ref{appendixfigures} shows the relation between the variance and the parameter $k$ and justifies the choice $k=7$ for $\hat{Y}_{knn}$. The corresponding value for the $\hat{Y}_{pred,knn}$, also based on the gaussian kernel, was found to be $k=9$.\\
 \textbf{To stabilize the bootstrap variances, the optimal global bandwidth obtained from the original full sample was used in all the bootstrap samples, and likewise for the optimal number $k$ of neighbors}. The regression coefficients, on the other hand, were re-calculated in each bootstrap sample.\\
 There are very slight differences in the point estimates $\hat{Y}_{reg}$  between the results presented here and those discussed in \cite{mandallaz4} because the boundary adjustments at the forest edge had to be simplified in order to reduce the computing time for the bootstrap calculations, based on 1000 replicates.\\
\noindent
 Table \ref{casestudy1} displays the results for the entire domain.
 \clearpage\newpage
 \begin{table}[ht]
\centering
\caption{ \label{casestudy1}\textbf{Global estimation for the case study in Canton Grisons}}\\[0.5cm]
\small
 \begin{tabular}{lccc}\hline
  Estimator & point estimate & external s.e.   & bootstrap s.e.\\ \hline\hline
 $\bar{Y}_F$               & 399.43  & 23.82    & 23.90    \\ \hline
 $\hat{Y}_{reg}$           & 384.95  & 16.52    & 18.34  \\ \hline\hline
 $\hat{Y}^{(1)}_{glkern}$  & 389.89  & 16.08    & 19.36      \\ \hline
 $\hat{Y}^{(1)}_{npreg}$   & 388.58  & 16.08    & 17.79    \\ \hline \hline
 $\hat{Y}^{(2)}_{glkern}$  & 389.79  & 16.08    & 19.41 \\ \hline
 $\hat{Y}^{(2)}_{npreg}$   & 387.51  & 16.28    & 18.55    \\ \hline \hline
 $\hat{Y}_{knn}$           & 390.53  & 16.35    & 17.83 \\ \hline
 $\hat{Y}_{pred,knn}$      & 389.22  & 15.69    & 17.42 \\ \hline\hline
 \end{tabular}
 \end{table}
 \normalsize
\noindent\textbf{Discussion:}\\[0.5cm]
The differences between the estimators are not statistically significant, which is reassuring. The bootstrap standard errors are roughly $15\%$ larger than the external standard errors. For $\hat{Y}_{reg}$ the g-weight variance is $16.96$ and closer to the bootstrap variance (see \cite{mandallaz4}). To calculate confidence intervals with the external variance one can use the normal distribution or the Student's distribution on $n_2-p$ degrees of freedom. \\[1cm]

We now consider the small-area estimation problem: we have a partition of the entire domain $F$, in 4 small areas $G_k$, i.e. $F=\cup_{k=1}^4 G_k$, with approximately equal surface areas $\lambda(G_k)$.\\
We use the simple external model approach after restricting the samples to the small areas. The regression coefficients, predictions and nearest neighbors are always based on the entire domain $F$. For the bootstrap calculations the re-sampling takes place in $F$ and the samples are then restricted to the small areas $G\in F$. Table \ref{casestudy2} displays the results.
\clearpage\newpage
\begin{table}[ht]
\begin{center}
\caption{ \label{casestudy2}\textbf{Small area estimation for the case study in Canton Grisons}}\\[0.5cm]
\small
\begin{tabular}{lrrrr}
  \hline
small area               & $G_1$ & $G_2$ & $G_3$ & $G_4$ \\
$n_1:n_2$                & 94:19   & 81:17   & 66:15   & 65:16 \\ \hline \hline
$\bar{Y}_G$              & 410.40  & 461.44  & 318.00  & 396.85  \\
                         & (44.58) & (56.35) & (34.36) & (47.86) \\
                         & [44.17] & [55.37] & [33.93] & [47.88] \\  \hline
$\hat{Y}_{reg}$          & 397.27  & 426.81  & 327.64  & 366.44 \\
                         & (28.80) & (35.01) & (31.64) & (36.01) \\
                         & [31.61] & [34.93] & [32.06] & [35.53] \\ \hline \hline
$\hat{Y}^{(1)}_{glkern}$ & 405.60  & 427.16  & 340.50  & 364.63 \\
                         & (28.82) & (34.63) & (29.65) & (34.65) \\
                         & [34.40] & [34.30] & [30.80] & [34.76] \\  \hline
$\hat{Y}^{(1)}_{npreg}$  & 397.36  & 428.37  & 341.08  & 369.36   \\
                         & (28.65) & (35.11) & (29.17) & (34.76) \\
                         & [30.72] & [33.27] & [29.21] & [33.67] \\  \hline \hline
$\hat{Y}^{(2)}_{glkern}$ & 405.46  & 426.94  & 340.58  & 364.59 \\
                         & (28.81) & (34.62) & (29.67) & (34.64) \\
                         & [34.42] & [34.25] & [30.86] & [34.70] \\ \hline
$\hat{Y}^{(2)}_{npreg}$  & 400.43  & 430.39  & 334.59  & 363.70   \\
                         & (29.61) & (34.73) & (30.09) & (35.01) \\
                         & [32.72] & [34.23] & [30.91] & [34.98] \\  \hline \hline
$\hat{Y}_{knn}$          & 382.46  & 464.14  & 319.42  & 376.96 \\
                         & (27.07) & (36.64) & (32.59) & (32.14) \\
                         & [28.28] & [34.77] & [28.04] & [30.02] \\ \hline
$\hat{Y}_{pred,knn}      & 398.60  & 424.02  & 353.21  & 364.48 \\
                         & (28.87) & (34.09) & (29.96) & (32.14) \\
                         & [29.36] & [32.42] & [28.41] & [30.62] \\ \hline \hline
\end{tabular}
\end{center}
\end{table}
\noindent\normalsize
\textbf{Legend}: \\
The standard error based on the external model assumption are given in $()$ and the standard errors based on the bootstrap variance estimates are given in $[]$.
\end{table}
\clearpage\newpage

\noindent\textbf{Discussion:}
\begin{itemize}
\item
All points estimators are close to each other and  not significantly different.
\item
The external variance estimates and bootstrap variances of all estimators are comparable.
\item
The small-area point estimates in the extended model and their g-weight variance estimates (as discussed in \cite{mandallaz4}) are also very close to to their counterparts in Table \ref{casestudy2}.
\item
One may eventually obtain better bootstrap results for the small areas by using a modified balanced replication as suggested in \cite{magnussen2}. However, this was not the case here.
\item
To calculate confidence intervals with the external variance we recommend to use the Student's distribution on $n_{2,G}-1$ degrees of freedom rather than the normal distribution.

\end{itemize}

 \subsection{Simulations}\label{simul}
 To illustrate the theory and check empirically the validity of the various mathematical approximation used to derive the variance estimates we present simulations performed on a purely artificial example, already used in \cite{mandallaz3}. The local density $Y(x)$ is defined according to the following procedure: at point $x=(x_1,x_2)^t\in\R^2$, the auxiliary vector is defined as $\pmb{Z}(x)=(1,x_1,x_2,x_1^2,x_1x_2,x_2^2)^t\in\R^p$ ($p=6$ in this example). The true parameter is
 $\pmb{\beta}_0=(30,13,-6,-4,3,2)^t\in\R^6$ and the local density over the domain $F=[0,2]\times [0,3]$ is given by the function
 \begin{equation}\label{simul1}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}_0+6\cos(\pi x_1)\sin(2\pi x_2):=\hat{Y}_0(x)+R(x)
 \end{equation}
 Figure \ref{plotdensity2} in Appendix \ref{appendixfigures} displays the local density.\\
 $\hat{Y}_{knn}$ is based on the Euclidian distance for $\pmb{Z}(x)$. We fixed the number of neighbors to $k=3$ because if the software is allowed to choose the optimal value of $k$ (on average $2.7$) we obtain $k_{opt}=1$ in about $30\%$ of the runs, which is totally absurd since this leads to constant zero residuals (essentially the best prediction at $x_0\in{s_2}$ is then the true observation $Y(x_0)$). This corresponds roughly to setting the local bandwidth $h(x_0)$ to zero for kernel-based estimators, which is not allowed because consistency requires $n_2h(x_0) \to \infty$ (see the remarks at the end of Appendix \ref{appendixdesignbased}). Fig. \ref{choiceofksim} in Appendix \ref{appendixfigures} illustrates the fact that the external variance of $\hat{Y}_{knn}$ underestimates the correct empirical variance, particularly in the range of interesting $k$ values. For the estimator $\hat{Y}_{pred,knn}$ the optimal values of $k$ were obtained from restricted simulations with 500 runs for each scenario.\\
 Tables \ref{simul1} and \ref{simul2} give the results for global estimation when the fitted model is indeed the true model and Tables \ref{simul3} and \ref{simul4} when the fitted model is a subset of the true model. All simulations are based on $10'000$ runs. The $95\%$ bootstrap confidence intervals are obtained via the well-known formula
 $[2\hat{Y}-q^{*}_{0.975},2\hat{Y}-q^{*}_{0.025}]$, where $q^{*}_{\alpha}$ is the $\alpha$-bootstrap quantile of the bootstrap estimate $\hat{Y}^{*}$, defined as the mean of the point estimates of the 1000 bootstrap samples obtained in each of the 10'000 runs.
 \clearpage\newpage

\begin{table}[h]
\small
\centering
\caption{$n_1=400,n_2=100$: fitted model is the true model }\label{simul1}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline\hline

$\hat{Y}_{reg}$             & 39.17 & 0.19 & 0.19 &0.20  &0.95  & 0.95  \\ \hline
$\hat{Y}^{(1)}_{glkern}$    & 39.15 & 0.20 & 0.19 &0.20  &0.94  & 0.95  \\
$\hat{Y}^{(1)}_{npreg}$     & 39.15 & 0.20 & 0.18 &0.20  &0.94  & 0.94   \\ \hline
$\hat{Y}^{(2)}_{glkern}$    & 39.15 & 0.20 & 0.19 &0.20  &0.94  & 0.95  \\
$\hat{Y}^{(2)}_{npreg}$     & 39.17 & 0.20 & 0.18 &0.20  &0.94  & 0.95  \\ \hline
$\hat{Y}_{pred,knn,k=11}$   & 39.16 & 0.20 & 0.18 &0.20  &0.93  & 0.94  \\
$\hat{Y}_{knn,k=3}$         & 39.15 & 0.18 & 0.14 &0.19  &0.92  & 0.95  \\ \hline \hline
\end{tabular}
\end{table}

\begin{table}[h]
\small
\centering
\caption{$n_1=200,n_2=50$: fitted model is the true model }\label{simul2}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator  & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$
& $\EX^*(\hat{P})$ &  $\EX^*(\hat{P}_{boot})$\\
  \hline\hline

$\hat{Y}_{reg}$             & 39.18 & 0.41 & 0.38 & 0.42       &0.95  & 0.95 \\ \hline
$\hat{Y}^{(1)}_{glkern}$    & 39.17 & 0.42 & 0.36 & 0.45       &0.92  & 0.95\\
$\hat{Y}^{(1)}_{npreg}$     & 39.15 & 0.42 & 0.35 & 0.43       &0.92  & 0.94   \\ \hline
$\hat{Y}^{(2)}_{glkern}$    & 39.17 & 0.42 & 0.36 & 0.45       &0.92  & 0.95 \\
$\hat{Y}^{(2)}_{npreg}$     & 39.17 & 0.43 & 0.36 & 0.42       &0.93  & 0.94   \\ \hline
$\hat{Y}_{pred,knn,k=6}$    & 39.15 & 0.43 & 0.35 & 0.43       &0.92  & 0.94\\
$\hat{Y}_{knn,k=3}$         & 39.15 & 0.43 & 0.31 & 0.43       &0.91  & 0.94  \\ \hline \hline
\end{tabular}
\end{table}

\normalsize
\noindent\textbf{Legend}: \\
The true mean value is $\bar{Y}=39.17$ and the true coefficient of determination is $R^2=0.83$.\\
$\EX^*(\hat{P})$ is the empirical coverage probability of the $95\%$ confidence interval based on the estimated variance and the normal distribution.\\
$\EX^*(\hat{P}_{boot})$ is the empirical coverage probabilities based on the bootstrap confidence intervals.
$\EX^*(\cdot)$ and $\VAR^*(\cdot)$ denote the empirical mean and variances over 10'000 runs.\\
$\hat{\VAR}(\hat{Y}_{estimator})$ is the estimated variance under the external model assumption.\\
$\hat{\VAR}_{boot}(\cdot)$ is the variance obtained from 1000 bootstrap replicates.
\clearpage\newpage
\begin{table}[h]
\small
\centering
\caption{$n_1=400,n_2=100$: fitted model is not the true model }\label{simul3}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline\hline
 $\hat{Y}_{reg}$              & 39.18 & 0.28 & 0.26 &0.27 &0.95 & 0.95 \\ \hline
 $\hat{Y}^{(1)}_{glkern}$     & 39.19 & 0.27 & 0.25 &0.28 &0.93 & 0.94\\
 $\hat{Y}^{(1)}_{npreg}$      & 39.18 & 0.28 & 0.24 &0.27 &0.92 & 0.94   \\ \hline
 $\hat{Y}^{(2)}_{glkern}$     & 39.19 & 0.27 & 0.25 &0.28 &0.93 & 0.94\\
 $\hat{Y}^{(2)}_{npreg}$      & 39.18 & 0.27 & 0.25 &0.27 &0.93 & 0.94  \\ \hline
 $\hat{Y}_{knn,pred,k=15}$    & 39.17 & 0.28 & 0.24 &0.27 &0.93 & 0.94\\
 $\hat{Y}_{knn,k=3}$          & 39.12 & 0.21 & 0.12 &0.22 &0.91 & 0.96  \\ \hline \hline
\end{tabular}
\end{table}

\begin{table}[h]
\small
\centering
\caption{$n_1=200,n_2=50$: fitted model is not the true model }\label{simul4}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline\hline
$\hat{Y}_{reg}$              & 39.21 & 0.56 & 0.50 & 0.56      & 0.94 &  0.95\\ \hline
 $\hat{Y}^{(1)}_{glkern}$    & 39.21 & 0.57 & 0.48 & 0.62      & 0.92 &  0.94\\
 $\hat{Y}^{(1)}_{npreg}$     & 39.18 & 0.57 & 0.46 & 0.56      & 0.92 &  0.93   \\ \hline
 $\hat{Y}^{(2)}_{glkern}$    & 39.21 & 0.57 & 0.48 & 0.62      & 0.92 &  0.94\\
 $\hat{Y}^{(2)}_{npreg}$     & 39.19 & 0.57 & 0.48 & 0.56      & 0.92 &  0.94  \\ \hline
 $\hat{Y}_{knn,pred,k=9}$    & 39.17 & 0.56 & 0.46 & 0.56      & 0.92 &  0.94\\
 $\hat{Y}_{knn,k=3}$         & 39.12 & 0.50 & 0.32 & 0.51      & 0.89 &  0.94  \\ \hline \hline
\end{tabular}
\end{table}
\normalsize
\noindent\textbf{Legend}: \\
The columns headers are the same as in Tables \ref{simul1} and \ref{simul2}. The true mean value is the same, i.e. $\bar{Y}=39.17$. The fitted model did not include the explanatory variables $x_1$ and $x_1x_2$. The true coefficient of determination decreases from $R^2=0.83$ to $R^2=0.66$.
\newpage

\noindent\textbf{Discussion}:
\begin{enumerate}
\item All estimators are practically design-unbiased, whether the fitted model is true or not, which is the main advantage of the design-based approach. The empirical (i.e.'true') variances are smaller when the fitted model is the true model, as expected.
\item
 The external variances always underestimate the empirical, i.e. true, variance. The underestimation can be substantial, except for $\hat{Y}_{reg}$.
 \item
The bootstrap variance estimates are close to the empirical variances.
\item
With respect to the nominal $95\%$ confidence intervals the classical regression estimator $\hat{Y}_{reg}$ performs very well in all investigated cases, and can be even further improved by using the g-weight variance instead of the external variance (\cite{mandallaz3}). For all other estimators one must use the bootstrap confidence intervals  because their external variances clearly underestimates the empirical variances. Slightly higher coverage probabilities can be achieved with the external variance by using the Student's t-distribution on $n_2-p$ degrees of freedom, which is recommended for small $n_2$, particularly for small area estimation.
\item
On the whole the sophisticated kernel-based or \textbf{knn} estimators do not perform better than than the classical regression estimator $\hat{Y}_{reg}$, which is disappointing but worth knowing. The nearest neighbor estimator $\hat{Y}_{knn}$ with the optimal choice $k=3$ yields a slightly smaller empirical variance (correctly estimated with the bootstrap) when the fitted model is not the true model.
\item
The excellent performances of the classical regression estimator $\hat{Y}_{reg}$ is probably due to the fact that the coefficients of determination are rather high under the true and under the working models, even if the value $R^2=0.66$ in the later case is not outstanding in the context e.g. of forest inventories with timber volume as response variable and with LiDAR canopy measurements as explanatory variables.
\item
Great care is needed with respect to the tuning options and the implementation of the bootstrap procedures.
\item
A further theoretical advantage of $\hat{Y}_{reg}$ is that it allows for correct analytical expressions of the asymptotic variances, which seems to be very difficult to obtain for the other estimators.

\end{enumerate}

\section{Conclusions}
These preliminary comparisons between the various estimators show that the simple classical two-phase regression estimator $\hat{Y}_{reg}$ with the external or g-weight variance estimates performs on the whole at least as well as the more sophisticated kernel-based or \textbf{knn} regression estimators with the bootstrap variance estimates. Also, the choice of the tuning options and re-sampling procedures of the later estimators is not trivial. The \textbf{knn} estimators share with some kernel-based estimators the disadvantage that they cannot extrapolate beyond the range of the observations, which is not the case for $\hat{Y}_{reg}$. The resulting coverage probabilities of the confidence intervals based on the bootstrap are satisfactory, likewise for the simpler external and g-weight variance estimates of $\hat{Y}_{reg}$. A definite advantage of the regression estimator $\hat{Y}_{reg}$ is that it is also available for more complex situations such as three-phase cluster sampling with two-types of auxiliary information, or with two-stage sampling of trees at the plot level, as well as their extension to small-area estimation (\cite{mandallaz3,mandallaz4,mandallaz5}). It can reasonably be expected that these conclusions will hold in general if the model is adequate, i.e. if it incorporates the most important explanatory variables (say with an $R^2 \ge 0.6$), which is certainly the case for timber volume with the standard LiDAR explanatory variable (mean canopy height being the most important). \\
In this context the main advantage of $\hat{Y}_{reg}$ is its simplicity and the uniqueness of the results, independently of the choice of kernels, metrics and further tuning options. It might be wise to use also $\hat{Y}_{knn}$, with well chosen options and bootstrap variance, to have a second opinion.

\newpage
\bibliography{biblio1}
\newpage

\begin{appendix}
\section{The Nadaraya-Watson regression estimator}\label{appendixnadaraya}
This estimator goes back to \cite{nadaraya1}, but the original idea is even older, going back to \cite{rosenblatt} and \cite{parzen}. The third kernel given in [\ref{threekernels}] was suggested by \cite{epanechnikov}.
 There is an immense literature on the subject of non-parametric kernel based estimation (for densities as well as for regression) in the mainstream model-dependent literature, but not so many in the design-based approach to survey sampling, and even less so in forest inventory.\\
In this appendix we summarize the main ideas and results and give informal proofs without going into the details of the regularity assumptions.\\
In the classical model-dependent approach to non-parametric kernel-based regression (see e.g. \cite{nonparametricreg1})
one considers $n_2$ i.i.d random vectors $(X_i,Y_i)$ with bivariate joint density $p(x,y)$ and marginal density $g(x)$ for the $X_i$. One wants to estimate the conditional expectation $m(x_0)=\EX(Y \mid X=x_0)$, which is the best predictor of $Y$ given $X=x_0$. The Nadaraya-Watson estimate of the marginal density $g(x)$ at the point $x_0$ is defined by
\begin{equation}\label{nadarayamarginal}
\hat{g}_{\varepsilon}(x_0)=\frac{1}{n_2\varepsilon(n_2,x_0)}\sum^{n_2}_{i=1}K\big(\frac{x_0-X_i}{\varepsilon(n_2,x_0)}\big)
\end{equation}
Under the conditions $\lim_{n_2\to\infty}\varepsilon(n_2,x_0))=0$ and $\lim_{n_2\to\infty}n_2\varepsilon(n_2,x_0)=\infty$ it is shown below that $\hat{g}_{\varepsilon}(x_0)$ is a consistent estimate (convergence in probability since the variance decreases as $\frac{1}{n_2\varepsilon(n_2,x_0)}$).
 Under the  condition $\lim_{n_2\to\infty}n_2(\varepsilon^2(n_2,x_0))=\infty$ one has also uniform convergence in probability,  i.e. $\lim_{n_2\to\infty}\PR(\sup_{x_0}\mid \hat{g}(x_0)-g(x_0)\mid > \eta)=0\;\forall \eta>0$, and even almost surely under further regularity conditions.\\
From a practical point of view we have good consistent estimates of the density functions, albeit in large samples. With respect to the mean square error $\EX_x(\hat{g}_{\varepsilon}(x_0)-g(x_0))^2$ the Epanechnikov kernel is optimal and the optimal bandwidth should go to zero as $n_2^{-\frac{1}{5}}$ so that squared bias and variance are of order $n_2^{-\frac{4}{5}}$ and and not as usual $n_2^{-1}$ (see \cite{lehmann}, p. 413).\\
The Nadaraya-Watson estimate of the bivariate density $p(x,y)$ with the bivariate product kernel $K(u,v)=K_1(u)K_2(u)$ ($K_1(u)$ and $K_2(v)$ are arbitrary kernels, e.g. such as those given in [\ref{threekernels}]) at the point $(x,y)$ is defined as
\begin{equation}\label{nadarayabivariate}
\hat{p}_{\varepsilon}(x,y)=\frac{1}{n_2\epsilon_{X}(n_2,x)\epsilon_{Y}(n_2,y)}\sum_{i=1}^{n_2}
 K_1\big(\frac{x-X_i}{\epsilon_X(n_2,x)}\big)
 K_2\big(\frac{y-Y_i}{\epsilon_{Y}(n_2,y)}\big)
 \end{equation}
One has consistency under the conditions
$$\lim_{n_2\to\infty}\varepsilon_X(n_2,x))=0 \;, \lim_{n_2\to\infty}\varepsilon_Y(n_2,y))=0$$ and $$\lim_{n_2\to\infty}n_2\varepsilon_X(n_2,x)\varepsilon_Y(n_2,y)=\infty$$
From the above a consistent estimate of the conditional density  $p_{Y\mid X}(y,x_0)$ of $Y$ given $X=x_0$ at the point $(y,x_0)$ is
\begin{equation}\label{conditionaldensityYgivenX}
\hat{p}_{Y\mid X}(y,x_0)=\frac{\hat{p}_{\varepsilon}(x_0,y)}{\hat{g}_{\varepsilon}(x_0)}
\end{equation}
To obtain a consistent estimate of the conditional expectation of $Y$ given $X=x_0$ it suffices to calculate the integral
$$\int_{\R}y\hat{p}_{Y\mid X}(y,x_0)dy$$
This integration is feasible only with a constant bandwidth $\varepsilon_Y(n_2,y)\equiv \varepsilon_Y(n_2)$ for $Y$. After changing the variable to $u=\frac{y-Y_i}{\varepsilon_Y(n_2)}$ and using $\int_{\R}K_2(u)=1$, $\int_{\R}uK_2(u)=0$ we get (after replacing $K_2(u)$ with $K(u)$)
\begin{equation}\label{estcondexpectationofYgivenX}
\hat{\EX}(Y \mid X=x_0)=\hat{m}(x_0)=
\frac{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum^{n_2}_{i=1}Y_iK\big(\frac{x_0-X_i}{\varepsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum^{n_2}_{i=1}K\big(\frac{x_0-X_i}{\varepsilon(n_2,x_0)}\big)}
\end{equation}
which is the celebrated Nadaraya-Watson non-parametric regression estimator.\\
\textbf{Remark}\\
For $p$-dimensional densities the the squared bias and variance term (with the optimal bandwidths) tend to $0$ at the rate $n_2^{-\frac{4}{4+p}}$, which can be viewed as the curse of dimensionality (see \cite{lehmann}, p.419-420, for  a dramatic example). This is a further motivation for minimizing the dimension of the feature space, i.e. $p=1$, by choosing the metric based on the $\hat{Y}(x)$ alone.\\
We now proceed to calculate the overall expectation and variance of [\ref{estcondexpectationofYgivenX}], i.e. under both random mechanisms generating the $X_i$ and $Y_i$. For the expectation we get
\begin{equation}\label{overallexpectation1}
\EX_{Y_i,X_i}(\hat{m}(x_0))=\EX_{X_i}\EX_{Y_i\mid X_i}(\hat{m}(x_0))=\EX_x\Big(\frac{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}m(x)K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}\Big)
\end{equation}
It is shown in Appendix \ref{appendixdesignbased} that one has
\begin{equation}\label{overallexpectation2}
\EX(\hat{m}(x_0))=m(x_0)+O(\varepsilon^2(n_2,x_0))
\end{equation}

For the variance we use the decomposition
\begin{equation}\label{overallvar1}
\VAR_{X_i,Y_i}(\hat{m}(x_0))=\VAR_{X_i}\EX_{Y_i\mid X_i}(\hat{m}(x_0))+\EX_{X_i}\VAR_{Y_i\mid X_i}(\hat{m}(x_0))
\end{equation}
With $m(x)=\EX(Y\mid X=x)$ and $\sigma^2(x)=\VAR(Y\mid X=x)$ we see that the first term can be written as
\begin{equation}\label{overallvar2}
\VAR_{X_i}\EX_{Y_i\mid X_i}(\hat{m}(x_0))=\VAR_x\Big(\frac{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}m(x)K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}\Big)
\end{equation}
and the second as
\begin{equation}\label{overallvar3}
\EX_x\Big(\frac{\frac{1}{n^2_2\varepsilon^2(n_2,x_0)}\sum_{x\in{s_2}}\sigma^2(x)K^2\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
{(\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big))^2}\Big)
\end{equation}
The denominator in [\ref{overallvar3}] tends to $g^2(x_0)$ and it is shown below that the numerator is asymptotically equivalent to
$$\frac{1}{n_2\varepsilon(n_2,x_0)}\sigma^2(x_0)g(x_0)\int_{\R}K^2(u)du$$
so that we finally have under the conditions
$$\lim_{n_2\to\infty}\varepsilon(n_2,x_0)=0\quad, \lim_{n_2\to\infty}n_2\varepsilon(n_2,x_0)=\infty$$
\begin{eqnarray}\label{overallvar4}
\EX_{Y_i,X_i}(\hat{m}(x_0))&=&m(x_0)+ O\big(\varepsilon^2(n_2,x_0)\big) \nonumber\\
\VAR_{Y_i,X_i}(\hat{m}(x_0))&=&\frac{1}{n_2\varepsilon(n_2,x_0)}\frac{\sigma^2(x_0)}{g(x_0)}\int_{\R}K^2(u)du
+O\big(\frac{\varepsilon(n_2,x_0)}{n_2}\big)
\end{eqnarray}
See also \cite{glad1} for more details and further developments.\\
 In the next section we shall analyze in more details the case when the $Y_i$ in the original Nadaraya-Watson estimator [\ref{estcondexpectationofYgivenX}] are replaced by a deterministic function $f(x_i)$ at random locations $X_i=x_i$ with probability density function $g(x)$, which is sometimes called the random design context.

\section{Kernel estimators in the design-based approach}\label{appendixdesignbased}
In the design-based approach and we consider instead the following kernel estimator of a deterministic function  $f:x\in{\R}\rightarrow f(x)\in{\R}$ at a point $x_0$
\begin{equation}\label{kerneldetermfofx}
\hat{f}_{\varepsilon}(x_0)=\frac{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}f(x)K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
:=\frac{\hat{e}_{\varepsilon}(x_0)}{\hat{g}_{\varepsilon}(x_0)}
\end{equation}
where $x_0$ is given and \textbf{the $n_2$ points $x\in{s_2}$ are i.i.d on $\R$ with density distribution function $g(x)$} satisfying some regularity conditions (e.g. twice differentiable). $f(x)$ is known for all $x\in{s_2}$. The expected value of the numerator (under the random location of the observations at points $x\in{s_2}$) is
$$\EX_x(\hat{e}_{\varepsilon}(x_0))=
\frac{1}{\varepsilon(n_2,x_0)}\int_{\R}f(x)K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)g(x)dx$$
Changing the integration variable to $u:=\frac{x_0-x}{\varepsilon(n_2,x_0)}$ we get
$$\EX_x(\hat{e}_{\varepsilon}(x_0))=\int_{\R}h(x_0-\varepsilon(n_2,x_0)u)K(u)du$$
where $h(x)=f(x)g(x)$. The gist of many arguments (see \cite{lehmann}, section 6.4 for details) is to use the Taylor expansion of $h(x)$ at $x_0$. Because by assumption $\int_{\R}K(u)du=1$, $\int_ {\R}uK(u)du=0$,$\int_{\R}u^2K(u)du<\infty$, we obtain the following result for the expected value
 \begin{equation}\label{nadaraya1}
\EX_x(\hat{e}_{\varepsilon}(x_0))=h(x_0)=f(x_0)g(x_0)+ O(\varepsilon^2(n_2,x_0))
\end{equation}
With similar arguments one obtains for the variance
\begin{equation}\label{nadaraya2}
\VAR_x(\hat{e}_{\varepsilon}(x_0))=\frac{1}{n_2\varepsilon(n_2,x_0)}f^2(x_0)g(x_0)\int_{\R}K^2(u)du
-\frac{1}{n_2}f^2(x_0)g^2(x_0)+O(\frac{\varepsilon(n_2,x_0)}{n_2})
\end{equation}
The Nadaraya-Watson estimator of the probability density function $g(x)$ at the point $x_0$ is by definition
\begin{equation}\label{nadaraya3}
\hat{g}_{\varepsilon}(x_0)=\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)
\end{equation}
Using the previous result in the special case $f(x)\equiv 1$ we obtain
\begin{eqnarray}\label{nadaraya4}
\EX(\hat{g}_{\varepsilon}(x_0))&=& g(x_0)+O(\varepsilon^2(n_2,x_0))\nonumber\\
\VAR_x(\hat{g}_{\varepsilon}(x_0))&=& \frac{g(x_0)}{n_2\varepsilon(n_2,x_0)}\int_{\R}K^2(u)du-\frac{1}{n_2}g^2(x_0)+O(\frac{\varepsilon(n_2,x_0)}{n_2})\nonumber\\
\end{eqnarray}
Hence, we have consistent estimates of $f(x_0)g(x_0)$. $g(x_0)$ and therefore of $f(x_0)=\frac{f(x_0)g(x_0)}{g(x_0)}$ provided that $\lim_{n_2\to\infty}\varepsilon(n_2,x_0)=0$ and $\lim_{n_2\to\infty}n_2\varepsilon(n_2,x_0)=\infty$. \\ \textbf{Recall that the band-width must go to zero but not so fast in order to ensure} $\lim_{n_2\to\infty}n_2\varepsilon(n_2,x_0)=\infty$.\\
With the same techniques and after some tedious but simple algebra one obtains for the covariance
\begin{equation}\label{nadaraya5}
\COV(\hat{e}_{\varepsilon}(x_0),\hat{g}_{\varepsilon}(x_0))=
\frac{f(x_0)g(x_0)}{n_2\varepsilon(n_2,x_0)}\int_{\R}K^2(u)du-\frac{1}{n_2}f(x_0)g^2(x_0)+O(\frac{\varepsilon(n_2,x_0)}{n_2})
\end{equation}

Using the well-known approximation formula for the variance of a ratio of two random variables $R$ and $S$, we get
\begin{equation}\label{varianceratio}
\VAR(\frac{R}{S})\approx
\frac{\EX^2(R)}{\EX^2(S)}\Big(\frac{\VAR(R)}{\EX^2(R)}-2\frac{COV(R,S)}{\EX(R)\EX(S)}+\frac{\VAR(S)}{\EX^2(S)}\Big)
\end{equation}
and after some algebra we see that the Nadaraya-Watson non-parametric regression estimator
\begin{equation}\label{nadarayareg1}
\hat{f}_{\varepsilon}(x_0)=\frac{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}f(x)K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{x_0-x}{\varepsilon(n_2,x_0)}\big)}
\end{equation}
under the conditions
$$\lim_{n_2\to\infty}\varepsilon(n_2,x_0)=0\quad \lim_{n_2\to\infty}n_2\varepsilon(n_2,x_0)=\infty$$
is a consistent estimator of $f(x_0)$ satisfying
\begin{eqnarray}\label{nadarayareg2}
\EX_(\hat{f}_{\varepsilon}(x_0))&=& f(x_0)+O(\varepsilon^2(n_2,x_0)) \nonumber\\
\VAR_x(\hat{f}_{\varepsilon}(x_0))&=& O(\frac{\varepsilon(n_2,x_0)}{n_2})
\end{eqnarray}
It is interesting to note that the terms in $\frac{1}{n_2(\varepsilon(n_2,x_0))}$ and $\frac{1}{n_2}$ occurring in [\ref{nadaraya4}] and [\ref{nadaraya5}] cancel out and that the design-based variance goes to zero as $O(\frac{\varepsilon(n_2,x_0)}{n_2})$, whereas in the model-dependent approaches the term of order $\frac{1}{n_2(\varepsilon(n_2,x_0))}$ remains as in [\ref{overallvar4}] above. The above design-based result [\ref{nadarayareg2}] corresponds formally, as expected, to the model-dependent case [\ref{overallvar4}] with $\sigma^2(x_0)=0$.\\
In the forest inventory context we estimate the probability density function of the predictions with
$$\hat{f}_{\hat{Y}}(\hat{y}_0)=
\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)$$
We see from the above arguments, after the substitution $x\rightarrow \hat{Y}(x)$ and $dx\rightarrow f_{\hat{Y}}(y)dy$,  that $\hat{f}_{\hat{Y}}(\hat{y}_0)$ is an asymptotically design unbiased estimate of the density function $f_{\hat{Y}}(y)$ at the point $y_0=\hat{Y}(x_0)$. Likewise, we see that
$$\frac{1}{n_2\varepsilon(n_2,x_0)}\sum_{x\in{s_2}}\hat{Y}(x)K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)$$
converges towards $\hat{y}_0f_{\hat{Y}}(y_0)$ (as usual convergence in the mean, in probability or almost surely depending on the regularity assumptions). Likewise, we have also shown that
\begin{equation}\label{consistency2}
\hat{\hat{Y}}_{\epsilon}(x_0):=\frac{ \frac{1}{n_2\epsilon(n_2,x_0)}
\sum_{x\in{s_2}}\hat{Y}(x)K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\epsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
\end{equation}
will converge towards $\hat{Y}(x_0)$.\\
\noindent However, we are here primarily interested in estimators of the form
\begin{equation}
\hat{Y}^{(1)}_{\epsilon}(x_0):=\frac{ \frac{1}{n_2\epsilon(n_2,x_0)}
\sum_{x\in{s_2}}Y(x)K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\epsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
\end{equation}
and
\begin{equation}
\hat{R}^{(2)}_{\epsilon}(x_0):=\frac{ \frac{1}{n_2\epsilon(n_2,x_0)}
\sum_{x\in{s_2}}\hat{R}(x)K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
{\frac{1}{n_2\epsilon(n_2,x_0)}\sum_{x\in{s_2}}K\big(\frac{\hat{Y}(x_0)-\hat{Y}(x)}{\epsilon(n_2,x_0)}\big)}
\end{equation}
This is only meaningful if we can write the local density $Y(x)$ as a function of its prediction $\hat{Y}(x)$, i.e. $Y(x)=Y(\hat{Y}(x))$ (and likewise $\hat{R}(x)=\hat{R}(\hat{Y}(x))$). This is not always possible, e.g. if the $\{\hat{Y}(x)\mid x\in{F}\}$ is a discrete set of points as with post-stratification, where for one $\hat{Y}(x)$ one has infinitely many $Y(x)$. In the continuous case the exact regression line (i.e. with the true optimal $\pmb{\beta}$) of $Y(x)$ (ordinate) on $\hat{Y}(x)$ (abscissa) is the diagonal (because $Y(x)=\hat{Y}(x)+R(x)$ and to each $\hat{Y}(x)$ one has one $Y(x)$, with probability $1$ ).\\
Neglecting the correlation between predictions (or using an independent training data set to fit the model) the mean and variance of $\hat{Y}^{(1)}_{\epsilon}(x_0)$ will be given by [\ref{nadarayareg2}]. The convergence seems to be fast, but recall that the recommended  optimal bandwidth is of order $n_2^{-\frac{1}{5}}$.

\noindent \textbf{Remark:}\\
Since $\hat{Y}^{(1)}_{\epsilon}(x_0)$ is asymptotically unbiased for $Y(x_0)$ it is tempting to consider the following estimator
\begin{equation}\label{badproposal}
\hat{Y}^{(1)}_{\varepsilon,reg}=\frac{1}{n_1}\sum_{x_0\in{s_1}}\hat{Y}^{(1)}_{\epsilon}(x_0)
\end{equation}
The covariance $\COV\big(\hat{Y}^{(1)}_{\epsilon}(x_1),\hat{Y}^{(1)}_{\epsilon}(x_2)\big)$ is of order $O\big(\frac{\varepsilon(n_2)}{n_2}\Big)$, so one may hope that the standard variance estimate
$$\frac{1}{n_1(n_1-1)}\sum_{x_0\in{s_1}}(\hat{Y}^{(1)}_{\epsilon}(x_0)-\bar{\hat{Y}}^{(1)}_{\epsilon})^2$$ is consistent. Let us denote with $R_{(1)}^2$ the R-squared of the kernel-based predictions $\hat{Y}^{(1)}_{\epsilon}(x_0)$ (regression line of the $Y(x_0)$ on the $\hat{Y}^{(1)}_{\epsilon}(x_0)$). The variance under the external model-approach is (see e.g. \cite{mandallaz}, eqn. 5.4)
\begin{equation}\label{extvar}
\VAR(\hat{Y}^{(1)}_{\varepsilon,reg})=\VAR_x(Y(x))\Big(\frac{1-R_{(1)}^2}{n_2}+ \frac{R_{(1)}^2}{n_1}\Big)
\end{equation}
whereas the above proposal yields
\begin{equation}\label{wrongvar1}
\VAR(\hat{Y}^{(1)}_{\varepsilon,reg})=\VAR_x(Y(x))\frac{R_{(1)}^2}{n_1}
\end{equation}
The same difficulty occurs with the classical two-phase regression estimator. In this case, the mean of the empirical residuals is zero and the empirical design-based covariance between predictions and residuals is zero by the orthogonality properties of ordinary least squares, so that one could naively think that the estimated variance is
\begin{equation}\label{wrongvar2}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\hat{\VAR}_x(\hat{Y})
\end{equation}
However, $\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t_{s_1}\hat{\pmb{\beta}}_{s_2}$ and one obtains, using Taylor expansion and the estimated covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ given in [\ref{estvarmatrix}], the correct answer, i.e. the g-weight based variance
\begin{equation}\label{correctvar}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\hat{\VAR}_x(\hat{Y})+
\hat{\pmb{\beta}}^t_{s_2}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\pmb{\beta}}^t_{s_2}
\end{equation}
which can be shown to be asymptotically equivalent (see \cite{mandallaz3} for examples and \cite{mandallazreport1} for detailed proofs) to the classical external variance estimate
\begin{eqnarray}\label{classicalexternalvar}
\hat{\VAR}(\hat{Y}_{reg})&=&\frac{1}{n_1}\hat{\VAR}_x(\hat{Y}(x))+ \frac{1}{n_2}\hat{\VAR}_x(\hat{R }(x)) \nonumber\\
&=& \frac{1}{n_1}\hat{\VAR}_x(Y(x))+ (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\VAR}_x(\hat{R}(x))
\end{eqnarray}
The second version can be expected to be better for kernel estimators since predictions and residuals do not have an exact zero design-based empirical covariance. Obviously, the same holds for the other estimators . Intuitively the result is obvious: there is no free lunch; one cannot expect to have with $n_2<<n_1$ costly terrestrial observations $Y(x)$ (plus the comparatively low cost of the remote sensing data) the same accuracy as with $n_1$!
\clearpage \newpage
\section{Calibration model}\label{calibration}
\subsection{Background}
In the recent literature on finite population surveys the concepts of calibration (calage in French, see \cite{tille1}) and balanced sampling have played, and still play an important role. The former does not deal with the given inclusion probabilities, i.e. the design, but determines weights $w_i$ in estimators of the type $\sum_{i\in{s}}w_iY_i$ such that the total distance $\sum_{i=1}^N d(w_i,d_i)$ to the original weights $d_i=\pi_i^{-1}$, with respect to some metric $d(\cdot,\cdot)$, is minimized under the constraint for the auxiliary variables $\sum_{i=1}^N \pmb{Z}_i=\sum_{i\in{s}}w_i\pmb{Z}_i$. As the original weights (i.e. $d_i=\pi_i^{-1}$) lead to the Horwitz-Thompson estimators (HT), the bias of the resulting estimator, i.e. calibration estimator, is small because it is close to HT. It can be expected to have good properties because it is exact for the auxiliary variables, assumed to be strongly related to the response variable. It can be shown that the calibration estimator is essentially asymptotically equivalent to linear regression estimators (i.e. linear prediction models, see \cite{sarndaldeville1}). In balanced sampling, on the other hand, one constructs a design with given inclusion probabilities $\pi_i$ such that the constraint $\sum_{i\in{s}}\frac{\pmb{Z}_i}{\pi_i}=\sum_{i=1}^N\pmb{Z}_i$ is satisfied, which can be done by the the so-called cube algorithm (\cite{tille2}). Hence, the two techniques achieve the same objective, calibration via the choice of the estimator and balanced sampling via the choice of the design. \\
  The paper \cite{wusitter1} is an important contribution as it generalizes the original calibration techniques to non-linear and generalized linear models used for the predictions. The inference is design-based but the construction of the calibration estimators is based on considerations coming from a super-population.
  \subsection{Calibration estimators in the Monte-Carlo approach}\label{calibrationmontecarlo}
  So far, the calibration estimators are only available in the finite population framework with exhaustive auxiliary information. We shall give below an informal proof of the main result given in \cite{wusitter1} in the infinite population framework germane to the Monte-Carlo approach. The only technical difficulty is that the inclusion probabilities of a given point is constant but $0$.m\\
   Recall that under the external model assumption the predictions are totally arbitrary and can be obtained from a non-linear model. Mathematically they must depend only on the auxiliary information $\pmb{Z}(x)$ at point $x$, at least asymptotically. More formally one can consider non-linear models of the form
 $$Y(x)=\mu(\pmb{Z}(x),\pmb{\beta}) + R(x)=:\mu(x,\pmb{\beta})+R(x)$$
 or generalized linear models $g(\mu(x,\pmb{\beta})=\pmb{Z}^t(x)\pmb{\beta}$
  with a given link function $g(\cdot)$. The ordinary non-linear least square uses the identity function $g(u)=u$. The estimating equation [7] in Wu-Sitter (with the common choice $q_i \equiv 1$ ) reads
 \begin{equation}
 \int_F\pmb{Z}(x)\phi(\mu(x,\pmb{\beta}))v^{-1}(\mu(x,\pmb{\beta})(Y(x)-\mu(x,\pmb{\beta})=0
 \end{equation}
 with $\phi(u)=(\frac{dg(u)}{du})^{-1}$ and an arbitrary weight function $v(\cdot)$. $\hat{\pmb{\beta}}$ can be obtained via sample copy version. The predictions are $\hat{Y}(x)=\mu(\pmb{Z}(x),\hat{\pmb{\beta}})$.\\
 In non-linear least-squares one has to minimize
 \begin{equation}
 Q(\pmb{\beta})=\frac{1}{\lambda(F)}\int_F (Y(x)-\mu(x,\pmb{\beta}))^2dx
 \end{equation}
  $\hat{\pmb{\beta}}$ can be obtained via the sample copy version
  \begin{equation}
  Q_{n_2}(\pmb{\beta})=\frac{1}{n_2}\sum_{x\in{s_2}} (Y(x)-\mu(x,\pmb{\beta}))^2
  \end{equation}
  $Q_{n_2}(\pmb{\beta})$ is a random function of $\pmb{\beta}}$ which converges almost surely point-wise and, under regularity conditions, also uniformly against the non-random function  $Q(\pmb{\beta})$. It is intuitively plausible that the maximum $\hat{\pmb{\beta}}$ of the random function $Q_{n_2}(\pmb{\beta})$ will also converge to the maximum of $Q(\pmb{\beta})$, which is by definition the true value $\pmb{\beta}_0$. Indeed, by the Law of Large Numbers we know that $Q_{n_2}(\pmb{\beta}$ converges point-wise almost surely against $\Q(\pmb{\beta)$. We can slightly strengthen this by requiring uniform convergence in a neighborhood $U_0$ of $\pmb{\beta}_0$. We can also reasonably assume that the functions $Q_{n_2}(\pmb{\beta}$ and $\Q(\pmb{\beta)$ twice differentiable in $U_0$. The derivatives (gradients in the multidimensional case) of $Q_{n_2}(\pmb{\beta}$ and $\Q(\pmb{\beta)$ are $0$ at $\pmb{\beta}_0$ and $\hat{\pmb{\beta}_{s_2}$ since they are minima. Using now the second order Taylor expansion at both minima we get (with 1-d notation) the asymptotic relation

  \begin{eqnarray*}
  Q_{n_2}(\pmb{\beta}&=&-Q(\pmb{\beta}&=& Q_{n_2}(\hat{\pmb{\beta}}_{s_2}-Q(\pmb{\beta}_0)\\
  &+& \frac{(\pmb{\beta}-\hat{\pmb{\beta}}_{s_2})^2{2}Q^{''}(\hat{\pmb{\beta}}_{s_2})+
  \frac{(\pmb{\beta}-\pmb{\beta}}_0)^2{2}Q^{''}(\pmb{\beta}_0)
  \end{eqnarray*}

  It can be proved by standard arguments (see e.g. \cite{cox} for the consistency of the Maximum Likelihoodthat $\hat{\pmb{\beta}}=\pmb{\beta}_0+O_p(n_2^{-\frac{1}{2}})$. Note that proofs of similar results for finite populations are more intricate because both the population and the sample have to tend to infinity. In the Monte-Carlo approach the population is infinite from scratch. We can obtain formally the results given in \cite{wusitter1} (specialized to the case of equal inclusions probabilities $\pi_i$ and weights $q_i$ ) with the following correspondences
  \begin{eqnarray*}
  N & \rightarrow & \lambda(F) \\
  Y_i &\rightarrow & Y(x) \\
  \sum_{i=1}^N Y_i &\rightarrow & \int_{F} Y(x)dx \\
  \pi_i \equiv \pi &\rightarrow & \pi(x)\equiv \frac{n_2}{\lambda(F)} \\
  d_i=\pi_i^{-1}\equiv \pi^{-1} &\rightarrow & d(x)\equiv \pi(x)^{-1}=\frac{\lambda(F)}{n_2} \\
  w_i &\rightarrow & w(x)             \\
  \end{eqnarray*}
  \begin{eqnarray*}
  q_i \equiv q &\rightarrow & q(x)\equiv q \\
  \sum_{i\in{s}}w_i=N &\rightarrow & \sum_{x\in{s_2}}w(x)=\lambda(F) \\
  \sum_{i\in{s}}\mu(\pmb{x}_i,\hat{\pmb{\beta}}) = \sum_{i=1}^N\mu(\pmb{x}_i,\hat{\pmb{\beta}})
  &\rightarrow &\sum_{x\in{s_2}}w(x)\hat{Y}(x) =  \int_F \hat{Y}(x)dx \\
  \min \sum_{i\in{s}}\frac{(d_i-w_i)^2}{q_i d_i} & \rightarrow & \min \sum_{x\in_{s_2}}(w(x)-d(x))^2\\
  \hat{\bar{Y}}_{HT} &\rightarrow & \frac{1}{n_2}\sum_{x\in{s_2}}Y(x) \\
  \hat{B}_N &\rightarrow & \frac{\sum_{x\in{s_2}}(\hat{Y}(x)-\bar{\hat{Y}}_{s_2}))(Y(x)-\bar{Y}_{s_2})}
  {\sum_{x\in{s_2}}(Y(x)-\bar{Y}_{s_2})^2}
  \end{eqnarray*}
  Note that with this notation one has $\EX_x(\sum_{x\in{s_2}}\frac{Y(x)}{\pi(x)})=\int_{F}Y(x)dx$, which is the standard Horwitz-Thompson form.\\
  The optimal weights $w_{opt}(x)$ minimize $\sum_{x\in_{s_2}}(w(x)-d(x))^2$ subject to the constraints
  $\sum_{x\in{s_2}}w(x)=\lambda(F) $ and $\sum_{x\in{s_2}}w(x)\hat{Y}(x)= \int_F \hat{Y}(x)dx$. The resulting \textbf{calibration estimator} $\hat{Y}_{MC}=\sum_{x\in{s_2}}w_{opt}(x)Y(x)$ can be rewritten, according to equation [10] in \cite{wusitter1}, as
 \begin{equation}
 \hat{Y}_{MC}=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)+\Big(\frac{1}{\lambda(F)}\int_F \hat{Y}(x)
 -\frac{1}{n_2}\sum_{x\in{s_2}}\hat{Y}(x)\Big)\hat{B}
 \end{equation}
 If  the first phase is not exhaustive the obvious choice is to set
 \begin{equation}
 \hat{Y}_{MC}=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)+\Big(\frac{1}{n_1}\sum_{x\in{s_1}} \hat{Y}(x)
 -\frac{1}{n_2}\sum_{x\in{s_2}}\hat{Y}(x)\Big)\hat{B}
 \end{equation}
 For linear prediction with intercept term the regression line of $Y(x)$ on $\hat{Y}(x)$ has intercept $0$ and slope $1$ so that $\hat{B}=1$ so that the calibration estimator is the same as the regression estimator.  Note that with linear models the regression estimator, because of zero mean residuals, can also be written  as a weighted sum of the observations, namely $\sum_{x\in{s_2}}\omega(x)Y(x)$ with $\omega(x)=\frac{1}{n_2}g(x)$ and $g(x)=\bar{\pmb{Z}}^t\pmb{A}_2\pmb{Z}(x)$. $g(x)$ are the so-called g-weights. One has $\sum_{x\in{s_2}}\omega(x)\pmb{Z}(x)=\bar{\pmb{Z}}$, which in this case is equivalent to the calibration constraint $\sum_{x\in{s_2}}\omega(x)\hat{Y}(x)=\int_F \hat{Y}(x)$. \\
 In the general case one has
 \begin{equation*}
 \hat{B} = \frac{\hat{\COV}(Y(x),\hat{Y}(x))}{\hat{\VAR}(\hat{Y}(x))}
 =\frac{\hat{\rho} \sqrt{\hat{\VAR}(Y(x)\hat{\VAR}(\hat{Y}(x))}}{\hat{\VAR}(\hat{Y}(x))}=\frac{\hat{\rho}}{\hat{R}}
 \end{equation*}
 where $\hat{\rho}$ is the empirical correlation between $Y(x)$ and
 $\hat{Y}(x)$ and $\hat{R}^2=\frac{\hat{\VAR}(\hat{Y}(x))}{\hat{\VAR}(\hat{Y}(x))}}$ is the empirical coefficient of determination. If the model is adequate one will also have $\sqrt{\hat{R}^2}=\hat{R}=\hat{rho}$ so that in general $\hat{B}$ will be very close to one.
 With the coefficient of determination $R^2=\frac{\VAR(\hat{Y}(x))}{\VAR(Y(x))}$ we get, as the regression line of $Y(x)$ on $\hat{Y}(x)$ should ideally have slope $1$ and intercept $0$ also with non linear predictions, the result
 $$\hat{B} \approx \frac{\rho}{R}\approx 1$$.
Hence, we have shown that the regression estimator can also be based on non-linear predictions models and that the external variance formula remains asymptotically valid. Furthermore, the regression estimator is asymptotically equivalent to the calibration estimator, but it does not require the first-phase to be exhaustive.
\clearpage\newpage

\section{Tables}\label{appendixtables}

 \subsection{Case study}\label{appendixtablescasestudy}
 Results when the optimal bandwidth is recalculated for each bootstrap sample.

 \begin{table}[h]
\centering
\caption{ \label{appcasestudy1}\textbf{Global estimation for the case study in Canton Grisons}}\\[0.5cm]
\small
 \begin{tabular}{lccc}\hline
  Estimator & point estimate & standard error   & s.e. bootstrap \\ \hline\hline
 $\hat{Y}^{(1)}_{lokern}$  & 396.30  & 15.98    & 35.82 \\ \hline
 $\hat{Y}^{(1)}_{glkern}$  & 389.88  & 16.08    & 24.76 \\ \hline\hline
 $\hat{Y}^{(2)}_{lokern}$  & 395.71  & 15.82    & 35.26 \\ \hline
 $\hat{Y}^{(2)}_{glkern}$  & 389.77  & 16.08    & 24.78 \\ \hline\hline
 \end{tabular}
 \end{table}
\normalsize
\begin{table}[ht]
\centering
\caption{ \label{appcasestudy2}\textbf{Small area estimation for the case study in Canton Grisons}}\\[0.5cm]
\small
\begin{tabular}{rrrrr}
  \hline
small area               & $G_1$ & $G_2$ & $G_3$ & $G_4$ \\
$n_1:n_2$                & 94:19   & 81:17   & 66:15   & 65:16 \\ \hline\hline
$\hat{Y}^{(1)}_{lokern}$ & 403.36  & 436.21  & 350.43  & 374.74 \\
                         & (27.89) & (33.93) & (29.94) & (33.66) \\
                         & [51.75] & [41.04] & [42.67] & [44.22] \\ \hline
$\hat{Y}^{(1)}_{glkern}$ & 405.60  & 427.16  & 340.50  & 364.63 \\
                         & (28.82) & (34.63) & (29.65) & (34.65) \\
                         & [48.97] & [38.34] & [35.79] & [37.48] \\  \hline
$\hat{Y}^{(2)}_{lokern}$ & 402.82  & 435.84  & 350.42  & 374.50 \\
                         & (27.89) & (33.93) & (29.95) & (33.67) \\
                         & [51.56] & [37.72] & [42.81] & [44.40] \\ \hline
$\hat{Y}^{(2)}_{glkern}$ & 405.45  & 426.94  & 340.54  & 364.58 \\
                         & (28.81) & (34.62) & (29.67) & (34.64) \\
\end{tabular}
\end{table}
The standard error based on the external model assumption are given in $()$ and the standard errors based on the bootstrap variance estimates are given in $[]$.\newpage

\clearpage \newpage

 \subsection{Simulations}\label{appendixtablessimul}
 \normalsize
 Results when the optimal bandwidth is recalculated for each bootstrap sample.\\
 The legends are the same as in Tables \ref{simul1}, \ref{simul2},\ref{simul3}, \ref{simul4}. The ? mark  in the column $\EX^*(\hat{\VAR}_{boot}())$ indicates that the value is corrupted by some extreme outliers as illustrated in Figures \ref{boot1} and \ref{boot2}.
\small
\begin{table}[h]
\centering
\caption{$n_1=400,n_2=100$: fitted model is the true model }\label{appsimul1}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline
$\hat{Y}^{(1)}_{lokern}$    & 39.16 & 0.20 & 0.18 &0.24  &0.94  & 0.96  \\
$\hat{Y}^{(1)}_{glkern}$    & 39.16 & 0.19 & 0.19 &0.22  &0.94  & 0.95  \\
$\hat{Y}^{(2)}_{lokern}$    & 39.16 & 0.20 & 0.18 &0.24  &0.94  & 0.96   \\
$\hat{Y}^{(2)}_{glkern}$    & 39.16 & 0.19 & 0.19 &0.22  &0.94  & 0.95  \\ \hline
\end{tabular}
\end{table}
\begin{table}[h]
\centering
\caption{$n_1=200,n_2=50$: fitted model is the true model }\label{appsimul2}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator  & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$
& $\EX^*(\hat{P})$ &  $\EX^*(\hat{P}_{boot})$\\ \hline
$\hat{Y}^{(1)}_{lokern}$    & 39.17 & 0.43 & 0.36 & 3.24?      &0.93 & 0.96 \\
$\hat{Y}^{(1)}_{glkern}$    & 39.18 & 0.42 & 0.36 & 0.52       &0.94 & 0.96\\
$\hat{Y}^{(2)}_{lokern}$    & 39.17 & 0.43 & 0.36 & 0.60       &0.93 & 0.96 \\
$\hat{Y}^{(2)}_{glkern}$    & 39.18 & 0.43 & 0.36 & 0.51       &0.94 & 0.96 \\ \hline

\end{tabular}
\end{table}
\clearpage\newpage
\small
\begin{table}[h]
\centering
\caption{$n_1=400,n_2=100$: fitted model is not the true model }\label{appsimul3}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline
 $\hat{Y}^{(1)}_{lokern}$     & 39.17 & 0.28 & 0.27 &0.39 &0.94 & 0.96\\
 $\hat{Y}^{(1)}_{glkern}$     & 39.18 & 0.28 & 0.26 &0.33 &0.94 & 0.96\\
 $\hat{Y}^{(2)}_{lokern}$     & 39.18 & 0.28 & 0.26 &0.38 &0.94 & 0.96\\
 $\hat{Y}^{(2)}_{glkern}$     & 39.18 & 0.28 & 0.26 &0.33 &0.94 & 0.96\\ \hline
\end{tabular}
\end{table}
\begin{table}[h]
\centering
\caption{$n_1=200,n_2=50$: fitted model is not the true model }\label{appsimul3}\\[0.5cm]
\begin{tabular}{lrrrrrr}
  \hline
Estimator & $\EX^*()$ & $\VAR^*()$ & $\EX^*(\hat{\VAR}())$ & $\EX^*(\hat{\VAR}_{boot}())$ & $\EX^*(\hat{P})$ & $\EX^*(\hat{P}_{boot})$ \\ \hline
$\hat{Y}^{(1)}_{lokern}$     & 39.20 & 0.61 & 0.47 & 1.78?     & 0.92 &  0.96\\
 $\hat{Y}^{(1)}_{glkern}$    & 39.20 & 0.58 & 0.48 & 0.76      & 0.92 &  0.95\\
 $\hat{Y}^{(2)}_{lokern}$    & 39.20 & 0.61 & 0.47 & 1.02?     & 0.91 &  0.96\\
 $\hat{Y}^{(2)}_{glkern}$    & 39.20 & 0.58 & 0.48 & 0.77      & 0.92 &  0.95\\ \hline
\end{tabular}
\end{table}

\clearpage\newpage

\section{Figures}\label{appendixfigures}

\subsection{Case study}\label{appendixfigurescasestudy}

\begin{figure}[h]
\begin{center}
\caption{\label{predict1} Predictions $\hat{Y}^{(1,2)}_{gl-lokern}(x_0)$ versus $\hat{Y}(x_0)$}
\end{center}
\begin{center}
\includegraphics[width=16cm,height=14cm]{predictionplot.pdf}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\caption{\label{choiceofkcasestudy} Variance as a function of $k$}
\end{center}
\begin{center}
\includegraphics[width=16cm,height=14cm]{kcasestudy.pdf}
\end{center}
\end{figure}
\clearpage\newpage

\subsection{Simulation}\label{appendixfiguressimulation}
 \begin{figure}[h]
\begin{center}
\caption{\label{plotdensity2}\textbf{Local density} $Y(x)$}
\end{center}
\begin{center}
\includegraphics[width=12cm,height=14cm]{plottrue.jpg}
\end{center}
\end{figure}
\clearpage \newpage

\begin{figure}[h]
\begin{center}
\caption{\label{boot1} \textbf{Bootstrap variances for} $n_1=400,n_2=100$}
\end{center}
\begin{center}
\includegraphics[width=14cm,height=18cm]{scatter400_100.pdf}
\end{center}
\end{figure}

\clearpage\newpage
\begin{figure}[h]
\begin{center}
\caption{\label{boot2} \textbf{Bootstrap variances for} $n_1=200,n_2=50$}
\end{center}
\begin{center}
\includegraphics[width=14cm,height=18cm]{scatter200_50.pdf}
\end{center}
\end{figure}
\clearpage\newpage
\begin{figure}[h]
\begin{center}
\caption{\label{choiceofksim} \textbf{Variance as a function of} $k$}
\end{center}
\begin{center}
\includegraphics[width=12cm,height=14cm]{ksim.pdf}
\end{center}
\end{figure}
\newpage
\subsection{R-Code}\label{appendix_R_code}
Below we give the R-code to run a single iteration of the simulation to produce an artificial data set and to calculate with it all the estimators discussed in this report, in particular the bootstrap variance estimates. Potential users can easily change the variables names to run the program on their own data sets.
\tiny{
\begin{verbatim}

######################
### SAMPLE R CODE  ###
######################

#LOAD LIBRARIES
library(kknn)
library(lokern)
library(np)
library(boot)

##########
#CREATE SINGLE EXAMPLE DATA (CALLED "s1") SET FROM SIMULATION SURFACE
##########
set.seed(7)
n1 <- 200
n2<- 50
s1 <- data.frame(s2.indicator = c(rep(1, times=n2), rep(0, times=n1-n2)), x=2*runif(n1), y=3*runif(n1))
s1$xx <- s1$x^2
s1$yy <- s1$y^2
s1$response <- ifelse(s1$s2.indicator==1, 30 + 13*s1$x - 6*s1$y - 4*s1$x^2 + 3*s1$x*s1$y + 2*s1$y^2 + 6*cos(pi*s1$x)*sin(2*pi*s1$y), NA)
head(s1, 60)

##########
#CREATE TRAINING SET
##########
#      - if true external model is used we need to select independent training set from s1
training.set <- data.frame(x=2*runif(n2), y=3*runif(n2))
training.set$xx <- training.set$x^2
training.set$yy <- training.set$y^2
training.set$response <- 30 + 13*training.set$x - 6*training.set$y - 4*training.set$x^2 + 3*training.set$x*training.set$y + 2*training.set$y^2 +
                         6*cos(pi*training.set$x)*sin(2*pi*training.set$y)
training.set <- training.set[training.set$s2.indicator==1,]

#      - if internal model is used... comment out for true external model from indep. training set chosen above
training.set <- s1[s1$s2.indicator==1,]

########################
##   DEFINE FUNCTIONS ##
########################
## CREATE GENERAL TWO-PHASE R-FUNCTION FOR DIFFERENCE ESTIMATOR
#     - ground.truth.s2 and residuals.s2 are length n2 (without missing values)
#     - predictions.s1 is length n1 (without missing values)
twophase_est <- function(ground.truth.s2, predictions.s1, residuals.s2){
  n1<- length(predictions.s1)
  n2<- length(residuals.s2)
  Est<- mean(predictions.s1) + mean(residuals.s2)
  Var<- (1/n1)*var(ground.truth.s2) + (1-(n2/n1))*(1/n2)*var(residuals.s2)
  Err<- sqrt(Var)

  out<- list(Est=Est, Err=Err)
  return(out)
}

#special leave-one-out-cross validation function for npreg(2) bandwidth selection
cv_bws_npreg <- function(x, y, bandwidths=c(1,2,4,6,10,15,20,30,40,50)) {
  n <- length(x)
  fold_MSEs <- matrix(0, nrow=n, ncol=length(bandwidths))
  colnames(fold_MSEs) = bandwidths
  case.folds <- 1:n
  for (loo in 1:n) {
    x.train = x[-loo]
    y.train = y[-loo]
    x.test = x[loo]
    y.test = y[loo]
    x.df <- as.data.frame(x.test)
    names(x.df) <- "x.train"
    reg <- lm(y ~ x)
    pred <- reg$fitted.values
    res <- reg$residuals
    hat <- pred + npreg(res ~ pred, regtype = "lc", bws = 1)$mean
    reg <- lm(y.train ~ x.train)
    pred.train <- reg$fitted.values
    pred.test <- predict(object = reg, newdata = x.df)
    pred.test.df <- as.data.frame(pred.test)
    names(pred.test.df) <- "pred.train"
    resid <- y.train - pred.train
    for (bw in bandwidths) {
      fit <- pred.test + npreg(resid ~ pred.train, regtype = "lc", newdata = pred.test.df, bws = bw)$mean
      fold_MSEs[loo,paste(bw)] <- (y.test - fit)^2
    }
  }
  CV_MSEs = colMeans(fold_MSEs)
  best.bw = bandwidths[which.min(CV_MSEs)]
  return(best.bw)
}

##########
#CHOOSE DESIRED MODEL
##########
model.formula <- response ~ y + yy + xx  # a.k.a "the working model"

##########
#FIT LINEAR MODEL AND ADD REGRESSION PREDICTIONS TO s1
##########
model <- lm(model.formula, data=training.set, y=TRUE) #note: the training set is used here
summary(model)
s1$lm.predictions <- predict(object=model, newdata=s1)
s1 <- s1[sort.list(s1$lm.predictions),]  # SOME PROCEDURES REQUIRE SORTING
s1$lm.residuals <- s1$response - s1$lm.predictions

##########
# MAKE PREDICTIONS FOR GLKERNS(1), NPREG(1), GLKERNS(2), NPREG(2), KNN(multidim), and KNN(pred)
##########

glkerns1 <- glkerns(x=s1[s1$s2.indicator==1, "lm.predictions"], y=s1[s1$s2.indicator==1, "response"],
                    x.out=s1[, "lm.predictions"])
glkerns1_band <- glkerns1$bandwidth  #save bandwidth
s1$glkerns1 <- glkerns1$est

###

npreg1 <- npreg(response ~ lm.predictions, data=s1[s1$s2.indicator==1,],
                newdata=s1, bwmethod="cv.ls")
npreg1_band <- npreg1$bw  #note: npreg's internally produced bandwidths have different scaling than cv_bws_npreg()
s1$npreg1 <- npreg1$mean

###

glkerns2 <- glkerns(x=s1[s1$s2.indicator==1, "lm.predictions"], y=s1[s1$s2.indicator==1, "lm.residuals"],
                    x.out=s1[, "lm.predictions"])
glkerns2_band <- glkerns2$bandwidth  #save bandwidth
s1$glkerns2 <- glkerns2$est + s1$lm.predictions

###

npreg2_band <- cv_bws_npreg(x=s1$lm.predictions[s1$s2.indicator==1], y=s1$response[s1$s2.indicator==1], bandwidths=seq(0.5, 10, 0.5))  # 2
s1$npreg2 <- npreg(lm.residuals ~ lm.predictions, data=s1[s1$s2.indicator==1,], newdata=s1, bws=npreg2_band)$mean + s1$lm.predictions

###

# leave one out cross validation to choose optimal k (k=3 in this run of the code)
knn_multi_band <- train.kknn(model.formula, data = s1[s1$s2.indicator==1,], kmax=15, kernel = c("gaussian"))$best.parameters$k
knn_multi <- kknn(formula=model.formula, train=s1[s1$s2.indicator==1,], test=s1, kernel="gaussian", k=knn_multi_band)
s1$knn_multi <- knn_multi$fitted.values

###

# leave one out cross validation to choose optimal k (k=12 in this run of the code)
knn_pred_band <- train.kknn(response ~ lm.predictions, data = s1[s1$s2.indicator==1,], kmax=15, kernel = c("gaussian"))$best.parameters$k
knn_pred <- kknn(formula=response ~ lm.predictions, train=s1[s1$s2.indicator==1,], test=s1, kernel="gaussian", k=knn_pred_band)
s1$knn_pred <- knn_pred$fitted.values

######################################################
##  Use difference estimator on above predictions   ##
######################################################
# CALCULATE SAMPLE MEAN AND ITS STANDARD ERROR
mean(s1$response, na.rm=TRUE)         # 40.52217
sqrt(var(s1$response, na.rm=TRUE)/n2) # 1.130878

# CALCULATE REG ESTIMATOR
REG <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
             predictions.s1 = s1[, "lm.predictions"],
             residuals.s2 = s1[s1$s2.indicator==1, "lm.residuals"])

REG$Est  # 39.75979
REG$Err  # 0.7247854

# CALCULATE GLKERN1 ESTIMATOR
GLKERN1 <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                    predictions.s1 = s1[, "glkerns1"],
                    residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "glkerns1"])

GLKERN1$Est  # 39.43728
GLKERN1$Err  # 0.6943061

# CALCULATE NPREG1 ESTIMATOR
NPREG1 <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                        predictions.s1 = s1[, "npreg1"],
                        residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "npreg1"])

NPREG1$Est  # 39.55139
NPREG1$Err  # 0.7038791

# CALCULATE GLKERN2 ESTIMATOR
GLKERN2 <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                       predictions.s1 = s1[, "glkerns2"],
                       residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "glkerns2"])

GLKERN2$Est  # 39.43422
GLKERN2$Err  # 0.6943259

# CALCULATE NPREG2 ESTIMATOR
NPREG2 <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                       predictions.s1 = s1[, "npreg2"],
                       residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "npreg2"])

NPREG2$Est  # 39.56989
NPREG2$Err  # 0.7031609

# CALCULATE KNN(multidim) ESTIMATOR
KNN_multidim <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                       predictions.s1 = s1[, "knn_multi"],
                       residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "knn_multi"])

KNN_multidim$Est  # 39.60094
KNN_multidim$Err  # 0.6325992
#Note: when expect this to be underestimation because we treated an internal model as external

# CALCULATE NPREG2 ESTIMATOR
KNN_pred <- twophase_est(ground.truth.s2 = s1[s1$s2.indicator==1, "response"],
                       predictions.s1 = s1[, "knn_pred"],
                       residuals.s2 = s1[s1$s2.indicator==1, "response"]- s1[s1$s2.indicator==1, "knn_pred"])

KNN_pred$Est  # 39.63425
KNN_pred$Err  # 0.698619

########################
##  BOOTSTRAP ROUTINE ##
########################

#EXAMPLE OF BOOTSTRAP ROUTINE

# Define function that inputs a data.frame with index vector and outputs estimates for above estimators.
# When bandwidths are allowed to be reoptimized within the following routine instability can result.
# All bandwidths obtained from the original sample are thus fixed throughout the routine.
# One must also pass the desired model formula to this function

boot.vector <- function(data, i, glkerns1_band=glkerns1_band, glkerns2_band=glkerns2_band,
                                 npreg1_band=npreg1_band, npreg2_band=npreg2_band,
                                 knn_multi_band=knn_multi_band, knn_pred_band=knn_pred_band,
                                 model.formula=model.formula){
  df=data[i,]
  training.set <- df[df$s2.indicator==1,]  #you change this for true external model approach (not done here)

  model <- lm(model.formula, data=training.set, y=TRUE) #note: the internal training set is used here
  df$lm.predictions <- predict(object=model, newdata=df)
  df <- df[sort.list(df$lm.predictions),]  # SOME PROCEDURES REQUIRE SORTING
  df$lm.residuals <- df$response - df$lm.predictions
  #----#
  MEAN <- mean(df$response, na.rm=TRUE)
  #----#
  REG <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
                      predictions.s1 = df[, "lm.predictions"],
                      residuals.s2 = df[df$s2.indicator==1, "lm.residuals"])$Est
  #----#
  df$glkerns1 <- glkerns(x=df[df$s2.indicator==1, "lm.predictions"], y=df[df$s2.indicator==1, "response"],
                      x.out=df[, "lm.predictions"], bandwidth=glkerns1_band)$est
  GLKERN1 <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
               predictions.s1 = df[, "glkerns1"],
               residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "glkerns1"])$Est
  #----#
  df$npreg1 <- npreg(response ~ lm.predictions, data=df[df$s2.indicator==1,],
                  newdata=df, bws=npreg1_band)$mean
  NPREG1 <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
                         predictions.s1 = df[, "npreg1"],
                         residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "npreg1"])$Est
  #----#
  glkerns2 <- glkerns(x=df[df$s2.indicator==1, "lm.predictions"], y=df[df$s2.indicator==1, "lm.residuals"],
                      x.out=df[, "lm.predictions"], bandwidth=glkerns2_band)
  df$glkerns2 <- glkerns2$est + df$lm.predictions
  GLKERN2 <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
                          predictions.s1 = df[, "glkerns2"],
                          residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "glkerns2"])$Est
  #----#
  df$npreg2 <- npreg(lm.residuals ~ lm.predictions, data=df[df$s2.indicator==1,], newdata=df, bws=npreg2_band)$mean + df$lm.predictions
  NPREG2 <- twophase_est(ground.truth.s2 = df[s1$s2.indicator==1, "response"],
                         predictions.s1 = df[, "npreg2"],
                         residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "npreg2"])$Est
  #----#
  df$knn_multi <- kknn(formula=model.formula, train=df[df$s2.indicator==1,], test=df, kernel="gaussian", k=knn_multi_band)$fitted.values
  KNN_multidim <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
                               predictions.s1 = df[, "knn_multi"],
                               residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "knn_multi"])$Est
  #----#
  df$knn_pred <- kknn(formula=response ~ lm.predictions, train=df[df$s2.indicator==1,], test=df, kernel="gaussian", k=knn_pred_band)$fitted.values
  KNN_pred <- twophase_est(ground.truth.s2 = df[df$s2.indicator==1, "response"],
                           predictions.s1 = df[, "knn_pred"],
                           residuals.s2 = df[df$s2.indicator==1, "response"]- df[df$s2.indicator==1, "knn_pred"])$Est

  return(c(MEAN, REG, GLKERN1, NPREG1, GLKERN2, NPREG2, KNN_multidim, KNN_pred))
}

#Call bootstrap (uncomment "parallel="multicore", ncpus=4" if parallel processing on linux or Mac OSx)
bootvec <- boot(s1, boot.vector, R = 10000,
                #parallel="multicore", ncpus=4,  # uncomment for multicore processing on linus or Mac OSx
                glkerns1_band=glkerns1_band, glkerns2_band=glkerns2_band,
                npreg1_band=npreg1_band, npreg2_band=npreg2_band,
                knn_multi_band=knn_multi_band, knn_pred_band=knn_pred_band,
                model.formula=model.formula)

bootvec

#Note: These bootstrap results (based on 10000 replicates) may differ from simulation results because the
#      original sample here is a SINGLE sample, not the mean and variance of 10000 samples.  That type of
#      simulation is possible because it is an artificial example where we know the truth.

\end{verbatim}
}
\end{appendix}
\end{document}

### 2017-07-28 texfiles Daniel - orig/reporttwophaserevised.tex
upload all files

\documentclass[a4paper,12pt,leqno, titlepage]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\bibliographystyle{mystyle2}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\VAR}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{RE}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}
\setlength{\topmargin}{0cm}
\setlength{\headheight}{3mm}
\setlength{\headsep}{1cm}
\setlength{\topskip}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{22cm}
\setlength{\oddsidemargin}{0cm}
\makeatletter
\def\tagform@#1{\maketag@@@{[\ignorespaces#1\unskip\@@italiccorr]}}
\makeatother

\begin{document}
\doublespacing
\nolinenumbers
\pagestyle{empty}

\title{Design-based properties of some small-area estimators in forest inventory with two-phase sampling \\ \small{revised version}}
\author{Daniel Mandallaz, Andreas Hill, Alexander Massey \\Deparment of Environmental Systems Science \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
\date{January 2016}

\maketitle
\newpage
$ $
\newpage
\begin{abstract}
We consider the small-area estimation problem in forest inventories with two-phase sampling schemes. We propose an improvement of the synthetic estimator, when the true mean of the auxiliary variables over the small-area is unknown and must be estimated, likewise for the residual corrected small-area estimator. We derive the asymptotic design-based variances of these new estimators, \textbf{the pseudo-synthetic and pseudo small-area estimators}, by incorporating also the design-based variance of the regression coefficients. We then propose a very simple mathematical device that transforms pseudo small-area estimators into pseudo-synthetic estimators, which is very convenient to derive asymptotic variances. The results are extended to cluster and two-stage sampling at the plot level. To illustrate the theory we consider the case of post-stratification and a case study.
\begin{center}
\textbf{R\'{e}sum\'{e}}
\end{center}
\begin{sloppypar}
Nous consid\`{e}rons le probl\`{e}me de l'estimation pour petits domaines dans le contexte d'inventaires forestiers en deux phases. Nous proposons une am\'{e}lioration simple de l'estimateur synth\'{e}tique quand la moyenne des variables auxiliaires dans le petit domaine doit \^{e}tre estim\'{e}e en premier lieu, de m\^{e}me pour l'estimateur pour petit domain bas\'{e} sur les r\'{e}sidus. Nous calculons la variance sous le plan de sondage de ces nouveaux estimateurs  en tenant compte de la variance des coefficients de r\'{e}gression.  De plus, nous proposons un artifice math\'{e}matique qui permet de transformer un estimateur pour petit domaine en un estimateur synth\'{e}tique, ce qui simplifie le calcul de la variance asymptotique. L'extension aux sondages par satellites et à deux degr\'{e}s au niveau de la placette est aussi trait\'{e}e. La th\'{e}orie est illustr\'{e}e par la post-stratification et par une \'{e}tude de cas.
\end{sloppypar}
\end{abstract}
\clearpage
\section*{Foreword}
The first version of this technical report has been published in the e-collection in September 2012. Since then many technical reports on further developments have been published in the e-collection
 (\cite{mandallazreport2,mandallazreport3,mandallazreport4,mandallazreport5,masseyphd}) as well as papers in the Canadian Journal of Forest Research, CJFR, (\cite{mandallaz3,mandallaz4,mandallaz5,massey1,massey2,massey3}. Recently, Dr. Alexander Massey and Andreas Hill, MSc, started writing an R-package implementing the various procedures discussed in the aforementioned publications.  Writing this software revealed some minor notational ambiguities used in cluster sampling as well as small numerical discrepancies between the new case study results produced in R and the original ones obtained with the IML procedure of the SAS package.  Detective work revealed that some SAS programs contained a minor error in a small-area variance formulae that caused the mean number of points per cluster in the small area to be taken from the second-phase instead of from the first-phase sample.  We emphasize that the underlying concept, i.e. "borrowing strength" by estimating regression coefficients over an unrestricted large area to improve the calculation of point and variance estimates in a restricted small area, remains precisely the same and that all the original conclusions of this technical report, including all relevant implications on subsequent works, remain valid.  This revised version improves the notation used for cluster sampling, presents updated tables with the corrected results for the case study and provides an enhanced discussion of the results. Some comments have been added to take into account results obtained since 2012. The mathematical derivations have also been carefully checked and, so far, confirmed.

\newpage
\pagestyle{plain}
\pagenumbering{arabic}
\section{Introduction}\label{introduction}
\pagenumbering{arabic} \setcounter{page}{1}
There is an extensive literature on the problem of small area estimation (or small domain estimation in general sampling). In this paper we shall investigate the properties of some estimators in the \textbf{model-assisted framework}, in which prediction models are used to improve the efficiency but are not assumed to be correct as in the \textbf{model-dependent approach}. The validity of the statistical procedures is ensured by the randomization principle: i.e. we are in the \textbf{design-based} inference framework, which has a definite advantage in official statistics. The reader is referred to (\cite{koehl}, section 3.8) for a good review of small-area estimation in forest inventory that presents alternative techniques, in particular Bayesian. Let us now define the sampling scheme.\par
The \textbf{first phase} draws a large sample $s_1$ of $n_1$ points that are independently and uniformly distributed within the forest area $F$. At each point $x\in{s}_1$ auxiliary
information is collected, very often coding information of qualitative nature
(e.g. following the  interpretation of aerial photographs) or quantitative (e.g. timber volume estimates  based on LIDAR measurements). We shall assume that the auxiliary information at point $x$ is described by the column vector $\mathbf{Z}(x)\in{\Re^{p}}$. \par
 The \textbf{second phase} draws a small sample $s_2\subset{s_1}$ of
$n_2$ points from $s_1$ according to \textbf{equal probability
sampling without replacement}. In the forested area $F$ we consider a well-defined population $ \mathcal{P}$ of $N$ trees with response variable
 $Y_i,\;i=1,2 \ldots$, e.g. the timber volume.  \textbf{The objective is to estimate the overall spatial mean}  $\bar{Y}=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i$, where $\lambda(\cdot)$ denotes the surface area (usually in ha) and \textbf{the mean over a small area} $G\subset F$, defined as
 \begin{equation}\label{small1}
 \bar{Y}_G=\frac{1}{\lambda(G)}\sum_{i=1}^N I_G(i)Y_i=:\frac{1}{\lambda(G)}\sum_{i\in{G}}Y_i
 \end{equation}
 where the indicator variable $I_G(i)$ is $1$ if the $i$-th tree lies in $G$, and $0$ otherwise.\\
 For each point $x\in{s_2}$ trees are drawn from the population $\mathcal{P}$ with probabilities $\pi_i$, for instance with concentric circles or angle count techniques. The
set of trees selected at point $x$ is denoted by $s_{2}(x)$. From each of the
selected trees $i\in{s_{2}(x)}$ one determines $Y_i$. The indicator variable $I_i$ is defined as
\begin{equation}\label{1stage}
 I_i(x)=\begin{cases}&1 \text{ if $i\in s_{2}(x)$}\\
                      &0 \text{ if $i\not\in s_{2}(x)$}
         \end{cases}
\end{equation}
At each point $x\in{s_2}$ the
terrestrial inventory provides the \textbf{local density} $Y(x)$
\begin{equation}\label{truelocaldensity}
 Y(x) =\frac{1}{\lambda(F)}\sum_{i=1}^N \frac{I_i(x)Y_i}{\pi_i}=\frac{1}{\lambda(F)}\sum_{i\in{s}_2(x)} \frac{Y_i}{\pi_i}
 \end{equation}
 The term $\frac{1}{\lambda(F)\pi_i}$ is the tree extrapolation factor $f_i$ with dimension $ha^{-1}$. One must include possible boundary adjustments, $\lambda(F)\pi_i=\lambda(F \cap K_i)$, where $K_i$ is the inclusion circle of the $i$-th tree. In the infinite population or Monte Carlo approach one samples the function $Y(x)$ (\cite{mandallaz}) for which the following important relation holds:
 \begin{equation}\label{montecarlo}
 \EX_{x} (Y(x))=\frac{1}{\lambda(F)}\int_{F} Y(x)dx=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i=\bar{Y}
 \end{equation}
 Where $\EX_x$ denotes the expectation with respect to a random point $x$ uniformly distributed in $F$. This establishes the link between the infinite population (continuum) $\{x\in{F} \mid Y(x)\}$ and the finite population of trees $\{i=1,2 \ldots N \mid Y_i\}$.\par
 Usually boundary adjustments are performed only with respect to $F$ and not with respect to the small area $G$. However, we shall assume that we also have
 \begin{equation}\label{small2}
 \bar{Y}_G=\frac{1}{\lambda(G)}\int_{G} Y(x)dx
 \end{equation}
 The afore mentioned randomization principle assume that we have uniformly independently distributed points or clusters in the forested area $F$, whereas in practice systematic grids are used. There is reasonable theoretical and empirical evidence that treating systematic grids as simple random samples is acceptable for point estimation and also for variance estimation (which will be in most instances slightly overestimated) for extensive forest inventories. From a mathematical point of view the only correct, and also most efficient, procedure, is the geostatistical Kriging technique (see \cite{mandallaz}, chapter 7 for a brief introduction and further references), which, however, is difficult to use and not uncontroversial in some aspects (e.g. choice of spatial correlation models and stationarity assumptions).
 \section{The model}
 We consider the linear model (the upper script on vector or matrices denotes thereafter the transposition operator)
 \begin{equation}\label{linearmodel1}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R(x)
 \end{equation}
 In the \textbf{model-dependent approach} the point $x$ is fixed and $R(x)$ is a random variable with zero mean and a given covariance structure. In the \textbf{design-based approach}
  $Y(x), \pmb{Z}(x), R(x)$ are random variables  because $x$ is random. The true regression coefficient $\pmb{\beta}$ is by definition the least squares estimate minimizing
 $$\int_F R^2(x)dx=\int_F(Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2 dx$$
 It satisfies the normal equation
 \begin{equation}\label{normaleq1}
 \Big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\Big)\pmb{\beta}=\int_F Y(x)\pmb{Z}(x)dx
 \end{equation}
 and the orthogonality relationship
 \begin{equation}\label{linearmodel2}
 \int_F R(x)\pmb{Z}(x)dx=\pmb{0}
 \end{equation}
  We shall assume that $\pmb{Z}(x)$ contains the intercept term $1$, or, more generally, that the intercept can be expressed as a linear combination of the component of $\pmb{Z}(x)$, which then insures that the mean residual is zero, i.e.
 $$\int_F R(x)dx=0$$
   The important case of stratification amounts to taking $\pmb{Z}^t(x)=(I_{F_1}(x),I_{F_2}(x),\ldots I_{F_L}(x))$, where $F=\cup_{k=1}^L F_k$ and $I_{F_k}(x)$ is the zero-one indicator variable of the $k$-th stratum $F_k$.\\
    We emphasize the fact that in the design-based model-assisted approach the model [\ref{linearmodel1}] is not viewed as an adequate description of the complex stochastic process generating the $Y(x)$, but, more pragmatically, simply as a tool to reduce the variance of estimators of $\bar{Y}, \bar{Y}_G$. Of course, ideally, the model should capture qualitatively the main features of the underlying natural phenomenon. \\To simplify the notation let us
 set $\pmb{A}=\EX_x\pmb{Z}(x)\pmb{Z}^t(x)$, $\pmb{U}(x)=Y(x)\pmb{Z}(x)$.
 The normal equation then reads
 $$\pmb{A}\pmb{\beta}=\EX_x\pmb{U}(x):=\pmb{U}$$
 Of course, only a sample-based normal equation is available, i.e.
 $$\pmb{A}_{s_2}\hat{\pmb{\beta}}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{U}(x)=\pmb{U}_{s_2}$$
 where we have set
 $$\pmb{A}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x)$$ and
 $$\pmb{U}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)$$
 The theoretical and empirical regression vector parameters are
 \begin{eqnarray}\label{linearmodel3}
 \pmb{\beta}&=&\pmb{A}^{-1}\pmb{U} \nonumber \\
 \hat{\pmb{\beta}}_{s_2}&=&\pmb{A}_{s_2}^{-1}\pmb{U}_{s_2}
 \end{eqnarray}
$\hat{\pmb{\beta}}_{s_2}$ is asymptotically design-unbiased for
$\pmb{\beta}$.
 To calculate the design-based variance-covariance matrix of the regression coefficients we need
 $$\EX(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta})(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta})^t$$
 we shall use the Taylor linearization technique. Let us consider the function
 $f(\cdot,\cdot)$ of an arbitrary $(p,p)$ matrix $\pmb{A}$ and an arbitrary $(p,1)$
vector $\pmb{U}$ defined by
 $f(\pmb{A},\pmb{U})=\pmb{A}^{-1}\pmb{U}$. We can write
 $$\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}=f(\pmb{A}_{s_2},\pmb{U}_{s_2})-f(\pmb{A},\pmb{U})$$
 which can be viewed as the differential of the function $f()$ at
 the point $\pmb{P}_0=(\pmb{A},\pmb{U})$, which is the expected value of
 the random point $\pmb{P}_{s_2}=(\pmb{A}_{s_2},\pmb{U}_{s_2})$. The distances between the fixed and the random point are of the order
 $n_2^{-\frac{1}{2}}$ in design-probability (by the law of large numbers for $\pmb{U}_{s_2}$ and $\pmb{A}_{s_2}$ and the continuity of the inverse operation). The differential of $f(\cdot,\cdot)$ at $\pmb{P}_0$ is, by the derivation
 rule for product
 $$df=d(\pmb{A}^{-1})\pmb{U}+\pmb{A}^{-1}d\pmb{U}$$
 Differentiating the identity $\pmb{A}^{-1}\pmb{A}=\pmb{I}$ one gets
 $$d(\pmb{A}^{-1})=-\pmb{A}^{-1}(d\pmb{A})\pmb{A}^{-1}$$
 and the following first-order Taylor expansion:
 $$\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}\approx
 -\pmb{A}^{-1}(\pmb{A}_{s_2}-\pmb{A})\pmb{A}^{-1}\pmb{U}+\pmb{A}^{-1}(\pmb{U}_{s_2}-\pmb{U})$$
 Expanding this expression and substituting
 $\pmb{A}^{-1}\pmb{U}=\pmb{\beta}$ we obtain the Taylor
 linearization \index{Linearization!Taylor}
 $$\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}\approx \pmb{A}^{-1}\Big(-\pmb{A}_{s_2}\pmb{\beta}+\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{U}(x)\Big)$$
 which is, by definition, equal to
 $$\pmb{A}^{-1}\Big(-\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}(x)^t\pmb{\beta}+\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)$$
 and consequently also to
 $$\pmb{A}^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\big(Y(x)-\pmb{Z}(x)^t\pmb{\beta}\big)\pmb{Z}(x)\Big)
 =\pmb{A}^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x)\Big)$$
 Thus, we finally arrive at
 \begin{equation}\label{Taylor}
 \hat{\pmb{\beta}}_{s_2}-\pmb{\beta}\approx \pmb{A}^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x)\Big)
 \end{equation}
 Using [\ref{linearmodel2}] and the independence of the $R(x)\pmb{Z}(x)$ one obtains the design-based variance-covariance matrix of
$\hat{\pmb{\beta}}_{s_2}$
\begin{equation}\label{varmatrix}
\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\approx
\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX_xR^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}^{-1}
\end{equation}
which can be estimated by replacing the theoretical residual $R(x)$
with their empirical counterparts
$\hat{R}(x)=Y(x)-\hat{Y}(x)$, with $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$, and $\pmb{A}$
with $\pmb{A}_{s_2}$. We then get the \textbf{estimated design-based
variance-covariance matrix} as
\begin{equation}\label{estvarmatrix}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\pmb{A}_{s_2}^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}_{s_2}^{-1}
\end{equation}
Interestingly this is precisely the \textbf{robust estimate of the model-dependent covariance matrix} given in \cite{gregoire2} (see also \cite{mandallaz} p. 107). Setting
$\hat{\sigma}^2=\frac{\sum_{x\in{s_2}}\hat{R}^2(x)}{n_2}$ we get $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\approx \frac{\hat{\sigma^2}}{n_2}\pmb{A}_{s_2}^{-1}$ whereas the model-dependent ordinary least squares theory gives the unbiased estimate of the covariance matrix as $(\frac{n_2}{n_2-p}\hat{\sigma}^2)\frac{1}{n_2}\pmb{A}_{s_2}^{-1}$.\\
The empirical residuals satisfy the sample orthogonality relation
\begin{equation}\label{sampleortho}
\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)\pmb{Z}(x)=\pmb{0}
\end{equation}
Theoretically one may use the exact matrix $\pmb{A}$ if it is available or its estimate $\pmb{A}_{s_1}=\frac{1}{n_1}\sum_{x\in{s_1}}\pmb{Z}(x)\pmb{Z}^t(x)$ based on the large sample. However, the resulting point estimates are not always intuitively convincing and not optimal in the model-dependent framework. Besides, they are not available from the usual statistical software packages. For these reasons we shall only work with $\pmb{A}_{s_2}$.
\section{The estimators}
\subsection{External models}
If the prediction model is \textbf{external}, i.e. not fitted with the inventory data at hand, the regression estimate is defined as
\begin{equation}\label{external1}
\hat{Y}_{reg}=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_0(x)+ \frac{1}{n_2}\sum_{x\in{s_2}}R_0(x)
\end{equation}
with the predictions $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta_0}$ and the residuals $R_0(x)=Y(x)-\hat{Y}_0(x)$, where $\pmb{\beta_0}$ is the given external regression coefficient, ideally obtained from another similar inventory whose sample points are independent of the current inventory's sample points. Note that in this case the mean residual will not necessarily be zero.
To calculate the variance one uses the decomposition
\begin{equation}\label{external2}
\VAR_{1,2}(\hat{Y}_{reg})=\VAR_1\EX_{2\mid 1}(\hat{Y}_{reg})+\EX_1\VAR_{2 \mid 1}(\hat{Y}_{reg})
\end{equation} to obtain
\begin{equation}\label{external3}
\VAR(\hat{Y}_{reg})=\frac{1}{n_1}\VAR(Y(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R_0(x))
 \end{equation}
 which can be unbiasedly estimated with
\begin{equation}\label{varexternalsimple}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R_0(x)-\bar{R}_{0,2})^2
\end{equation}
where $\bar{Y}_2=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ and $\bar{R}_{0,2}=\frac{1}{n_2}\sum_{x\in{s_2}}R_0(x)$.\\
The estimation for any small area $G\subset F$ is straightforward, indeed one simply restricts the samples of $n_1$ and $n_2$ points in $F$ to the $n_{1,G}$ and $n_{2,G}$ points in $G$ and apply the above formulae to obtain an unbiased estimate of the conditional variance (i.e. given $n_{1,G}$ and $n_{2,G}$, which are realizations random variables because in our set-up only $n_1$ and $n_2$ are fixed).
\subsection{Internal models}
In most applications the model has to be fitted with the data provided by the current inventory. In this case, the model is said to be \textbf{internal}. In very large samples one can treat an internal model as external and apply again the formulae given above, which obviously neglects the error in the regression coefficients. This is essentially the framework presented in (\cite{mandallaz}, chapter 5 and section 6.3). We shall show in the present paper how one can take the design-based variance of the regression coefficients into account, albeit still in large samples, and incorporate the mean residual directly in the model.\\
 The model-dependent estimator for the small area $G$ is called the
\textbf{synthetic estimator} and is given by
\begin{eqnarray}\label{synthetic1}
\hat{Y}_{G,synth}&=&\frac{1}{\lambda(G)}\int_G\hat{Y}_{s_2}(x)dx \\ \nonumber
&=& \frac{1}{\lambda(G)}\int_G \pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}dx=\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2}
\end{eqnarray}
where $\bar{\pmb{Z}}_G=\frac{1}{\lambda(G)}\int_G\pmb{Z}(x)dx$ is the true mean of the auxiliary vector over the small area $G$, which is available only if the first phase is exhaustive. $\hat{Y}_{G,synth}$ is unbiased under the model, but not optimal as it does not the take the model-dependent spatial correlation of the $Y(x)$ into account. Let us emphasize the fact that the model, i.e. $\hat{\pmb{\beta}}_{s_2}$, is fitted with the full data set and not only with $\{Y(x), \pmb{Z}(x)\mid x\in G\}$. \\
\textbf{In this paper we shall investigate the properties of $\hat{Y}_{G,synth}$ in the design-based inference framework}. \\
First, let us note that $\hat{Y}_{G,synth}$ is a design-based consistent sample copy of
 $$\frac{1}{\lambda(G)}\int_G \hat{Y}(x)dx=\frac{1}{\lambda(G)}\int_G(Y(x)-R(x))dx=\bar{Y}_G-\frac{1}{\lambda(G)}\int_G R(x)dx$$
 Consequently, the synthetic estimator $\hat{Y}_{G,synth}$ has a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_GR(x)$, which is not zero unless $G=F$ (we have zero mean residual over the entire domain, see [\ref{linearmodel2}]) or, which is unlikely, zero mean residual over the small area of interest. Using [\ref{synthetic1}] and [\ref{estvarmatrix}] \textbf{the estimated design-based variance of the synthetic estimator} is
\begin{equation}\label{synthetic2}
\hat{\VAR}(\hat{Y}_{G,synth})=\bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
\end{equation}
We define the g-weights as
\begin{equation}\label{gweights1}
g_G(x)=\bar{\pmb{Z}}_G^t\pmb{A}^{-1}_{s_2}\pmb{Z}(x)
\end{equation}
It is easily checked that one can rewrite the point estimate and its estimated variance as
\begin{eqnarray}\label{gweights2}
\hat{Y}_{G,synth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}g_G(x)Y(x) \nonumber\\
\hat{\VAR}(\hat{Y}_{G,synth})&=&\frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_G(x)\hat{R}^2(x)
\end{eqnarray}
where the $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ are the empirical residuals.
In the above the special case $G=F$ is possible. The g-weights enjoy  several attractive statistical properties (see \cite{sarndal} for the aspects in general sampling theory and \cite{mandallaz} for their Monte-Carlo counterparts in forest inventory).\\
To compensate for the bias due to the non vanishing mean residual over $G$ one considers the \textbf{small-area estimator} (\cite{mandallaz} p.120)
\begin{equation}\label{smallareaest}
\hat{Y}_{G,small}=\hat{Y}_{G,synth}+
\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)
\end{equation}
\noindent where $s_{2,G}=s_2\cap G$ and $n_{2,G}=\sum_{x\in{s_2}}I_G(x)$ is the number of
points of $s_2$ falling within $G$. It can be shown (\cite{mandallaz}) that $\hat{Y}_{G,small}$ is asymptotically design-unbiased with estimated design-based variance given by
\begin{equation}\label{smallareadesignvariance1}
\hat{\VAR}(\hat{Y}_{G,small})=
\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_{2,G})^2
\end{equation}
\noindent where
$$\bar{\hat{R}}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)$$
is the estimated mean residual over a small area. The above variance estimate neglects the variance of $\hat{\pmb{\beta}}_{s_2}$ and is therefore valid only if $n_2$ is very large and $n_2>> n_{2,G}$. To have better insight we use the expansion [\ref{Taylor}] to obtain
\begin{equation}\label{smallexp1}
\hat{Y}_{G,small}-\bar{Y}_G=
\bar{\pmb{Z}}^t_G\pmb{A}^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x)\Big)
+ \frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R(x)-\bar{R}_G
\end{equation}
which leads to the variance
\begin{equation}\label{smallvar1}
\EX(\hat{Y}_{G,small}-\bar{Y}_G)^2=
\bar{\pmb{Z}}^t_G\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX R^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}^{-1}\bar{\pmb{Z}}_G + \VAR \big(\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R(x)\big )+ C
\end{equation}
\noindent where the cross-product term $C$ is given by
$$2\bar{\pmb{Z}}^t_G\pmb{A}^{-1}\EX\big( (\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x))
(\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R(x)-\bar{R}_G)\big)$$
Both terms of the product tend to zero at rate $(n_2^{-\frac{1}{2}})$, which unfortunately is of the same order as the first two terms. However, using the fact that the $R(x)$, $\pmb{Z}(x)$ are independent of $R(y)$, $\pmb{Z}(y)$ for $x\ne y$ we obtain after tedious but simple calculations
$$C=\frac{1}{n_2}\bar{\pmb{Z}}^t_G\pmb{A}^{-1}\big(\EX_{x\in{G}}R^2(x)\pmb{Z}(x)-\bar{R}_G\EX_{x\in{G}}R(x)\pmb{Z}(x)\Big)$$
which we can reasonably assume to be negligible. The above arguments suggest therefore the following estimate of the design-based variance of the small-area estimator with exhaustive first phase
\begin{eqnarray}\label{smallareadesignvariance2}
\hat{\VAR}(\hat{Y}_{G,small})&=&
\bar{\pmb{Z}}^t_G\pmb{A}_{s_2}^{-1}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}} \hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}_{s_2}^{-1}\bar{\pmb{Z}}_G \\ \nonumber &+&
\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_{2,G})^2
\end{eqnarray}
Comparing with [\ref{varexternalsimple}] (after restricting the samples to $G$) we see that \textbf{treating an internal model as an external model} (i.e. ignoring the variability of $\hat{\pmb{\beta}}_{s_2}$) \textbf{ will underestimate the variance of the small area estimate}. The first term in [\ref{smallareadesignvariance2}] reflects the uncertainty in the regression coefficients.\\
If the first-phase is non-exhaustive, i.e. $n_1 \ne \infty$, then one can replace the true mean $\bar{\pmb{Z}}_G$ by its estimate in the large sample
$$\hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\pmb{Z}(x)$$ where $s_{1,G}$ is the set $s_1\cap G$ of the $n_{1,G}=\sum_{x\in{s_1}}I_g(x)$ points of the large sample falling into the small area $G$. This gives the \textbf{pseudo-synthetic estimator}
\begin{equation}\label{pseudosynth1}
\hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
=\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
\end{equation}
which is clearly asymptotically equivalent to $\hat{Y}_{G,synth}$ as $n_1 \to \infty$ and its design-based expected value tends to  $\bar{\pmb{Z}}^t_G\pmb{\beta}$. To calculate the asymptotic variance we use the decomposition (actually the first order Taylor expansion)
$$\Delta=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}-\bar{\pmb{Z}}_G^t\pmb{\beta}=
\hat{\bar{\pmb{Z}}}_{1,G}^t(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta})+ (\hat{\bar{\pmb{Z}}}_{1,G}^t-\bar{\pmb{Z}}_G)^t\pmb{\beta}$$
Asymptotically we get
$$\EX \Delta^2=\bar{\pmb{Z}}_G ^t\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G+
\pmb{\beta}^t\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\pmb{\beta}+
2\EX\Big((\hat{\pmb{\beta}}_{s_2}-\pmb{\beta})^t\hat{\bar{\pmb{Z}}}_{1,G}(\hat{\bar{\pmb{Z}}}_{1,G}
-\bar{\pmb{Z}}_G)^t\pmb{\beta}\Big)$$
\noindent where $\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}$ is the covariance matrix of
$\hat{\bar{\pmb{Z}}}_{1,G}  $. The first two terms are of order $n_2^{-1}$ and $n_1^{-1}$ respectively. For the third term we note that
$\EX (\hat{\bar{\pmb{Z}}}_{1,G}(\hat{\bar{\pmb{Z}}}_{1,G}-\bar{\pmb{Z}}_G)^t)$ is equal to the covariance matrix $\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}$ and therefore of order $n_1^{-1}$ and that
$\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}$ is of order $n_2^{-\frac{1}{2}}$. The last term is therefore of smaller order than the first two which leads to the following asymptotic design-based estimate of variance
\begin{equation}\label{estvarpseudosynth1}
\hat{\VAR}(\hat{Y}_{G,psynth}): =
\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
+ \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
\end{equation}
The variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}=
\frac{1}{n_{1,G}(n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1,G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1,G})^t
\end{equation}
Usually $\hat{Y}_{G,psynt}$ will have a small variance but at the cost of a potential bias.
We can rewrite [\ref{pseudosynth1}] and [\ref{estvarpseudosynth1}] with the g-weights
\begin{eqnarray}\label{gweightpseudo}
g_{G,1}(x)&=&\hat{\bar{\pmb{Z}}}^t_{1,G}\pmb{A}^{-1}_{s_2}\pmb{Z}(x) \nonumber\\
\hat{Y}_{G,psynth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}g_{G,1}(x)Y(x)
\end{eqnarray}
and after some algebra we get
\begin{eqnarray}\label{gweightpseudo2}
\hat{\VAR}(\hat{Y}_{G,psynth})&=&\frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{G,1}(x)\hat{R}^2(x) + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}  \nonumber \\
&=& \frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{G,1}(x)\hat{R}^2(x)+ \frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}(\hat{Y}(x)-\bar{\hat{Y}}_{1,G})^2
\end{eqnarray}
where $\bar{\hat{Y}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)$. The second term in the last equation is the variance of the predictions over $G$.\\
The \textbf{pseudo small-area estimator}
\begin{equation}\label{pseudosmall}
\hat{Y}_{G,psmall}=\hat{Y}_{G,psynth}+\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)
\end{equation}
is asymptotically design-unbiased and intuitively its variance can be expected to be well approximated by
\begin{equation}\label{pseudosmallvar}
\hat{\VAR}(\hat{Y}_{G,psmall})=
\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
+ \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
+\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_{2,G})^2
\end{equation}
A tedious formal proof can be given by using [\ref{external2}], [\ref{smallvar1}] and
[\ref{estvarpseudosynth1}].\\
Using the same arguments as in [\ref{gweightpseudo2}] we also have
\begin{eqnarray}\label{pseudosmallvar2}
\hat{\VAR}(\hat{Y}_{G,psmall})&=&\frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{G,1}(x)\hat{R}^2(x)
+ \frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}(\hat{Y}(x)-\bar{\hat{Y}}_{1,G})^2 \nonumber \\
&+& \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_{2,G})^2
\end{eqnarray}
This should be compared with the external version [\ref{varexternalsimple}]
\begin{eqnarray}\label{varexternalsimpleG}
\hat{\VAR}(\hat{Y}_{G,psmall})&=&\frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}
\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\
&+& (1-\frac{n_{2,G}}{n_{1,G}})
\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_{2,G})^2
\end{eqnarray}
For very large $n_{1,G}$ it is clear that the external version will underestimate the variance
as it neglects the first term in [\ref{pseudosmallvar2}], which albeit is also small for large $n_2$. \\
The special case $G=F$ deserves special attention: because of the zero mean residual we have $\hat{Y}_{F,psmall}=\hat{Y}_{F,psynt}=\hat{Y}_{reg}$ and [\ref{estvarpseudosynth1},\ref{pseudosmallvar2}] lead to the estimated variance
\begin{equation}\label{varyregsynt1}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{F,1}(x)\hat{R}^2(x)
+ \frac{1}{n_1(n_1-1)}
\sum_{x\in{s_1}}(\hat{Y}(x)-\bar{\hat{Y}}_1)^2
\end{equation}
with $\bar{\hat{Y}}_1=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}(x)$. The external version
[\ref{varexternalsimpleG}] gives
\begin{equation}\label{varexternalsimpleG2}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_{2}}{n_{1}})
\frac{1}{n_{2}}\frac{1}{n_{2}-1}\sum_{x\in{s_{2}}}\hat{R}^2(x)
 \end{equation}
 Writing $Y(x)=\hat{Y}(x)+\hat{R}(x)$ an using [\ref{sampleortho}] we can rewrite [\ref{varexternalsimpleG2}] as
 \begin{equation}\label{varexternalsimpleG3}
 \hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{Y}(x)-\bar{\hat{Y}}_2)^2+
\frac{1}{n_{2}}\frac{1}{n_{2}-1}\sum_{x\in{s_{2}}}\hat{R}^2(x)
\end{equation}
For $G=F$ the $g^2_{F,s_1}(x)$ are asymptotically equal to $1$ (see \cite{mandallaz} p. 113 and the properties of the g-weights discussed below) so that both versions [\ref{varyregsynt1}] and [\ref{varexternalsimpleG3}] are asymptotically equivalent. However, version [\ref{varyregsynt1}] estimates the variance of the predictions in the large sample, which is better, and it rests upon the g-weights for the residual part, which is known to have better conditional properties (see \cite{mandallaz} p. 84 and section \ref{poststrat} for the important special case of stratification).\\
In the next section we present a simple reformulation of the problem that allows one to transform small-area estimators into synthetic estimators, which offers a great mathematical advantage.
\subsection{Alternative estimators in extended model}
 The main difficulty stems from the fact that
$\int_{G}R(x)dx \ne 0$. If we now extend the auxiliary information vector  $\pmb{Z}(x)$ to $\pmb{\mathcal{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))\in{\mathcal{R}}^{(p+1)}$, the corresponding model reads
\begin{equation}\label{extendemodel1}
Y(x)=\pmb{\mathcal{Z}}^t(x)\pmb{\theta} +\mathcal{R}(x)
 \end{equation}
 which leads to the normal equation for the extended parameter vector $\pmb{\theta}\in{\mathcal{R}}^{(p+1)}$
$$\Big(\int_F\pmb{\mathcal{Z}}(x)\pmb{\mathcal{Z}}^t(x)dx\Big)\pmb{\theta}=:
\pmb{\mathcal{A}}\pmb{\theta}=\int_{F}Y(x)\pmb{\mathcal{Z}}(x)dx$$ and the orthogonality relationship
$$\int_{F}\mathcal{R}(x)\pmb{\mathcal{Z}}(x)dx=\pmb{0}$$
 Since $I_F(x)\equiv 1$ is the intercept term (or linear combination of the components of $\pmb{Z}(x)$) and $\pmb{\mathcal{Z}}(x)$ contains $I_G(x)$ \textbf{we have the two zero mean residual properties}
$$\int_F \mathcal{R}(x)dx=\int_G \mathcal{R}(x)dx=0$$
Hence, by including the $0,1$ indicator variable of the small area $G$ into the model, we enforce zero mean residual over $F$ and $G$. Note also that $G$ must be a proper subset of $F$, otherwise $\pmb{\mathcal{A}}$ and $\pmb{\mathcal{A}}_{s_2}$ are singular. In practice near-singularity could cause numerical problems, so that the small area $G$ must indeed be small with respect to $F$. Simple calculations yield the following block structure for $\pmb{\mathcal{A}}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathcal{Z}}(x)\pmb{\mathcal{Z}}^t(x)$
\begin{equation}\label{tildeAs2}
 \pmb{\mathcal{A}}_{s_2}=
\left[ \begin {array}{ll}
\pmb{A}_{s_2}& \hat{p}_{2,G}\hat{\bar{\pmb{Z}}}_{2,G} \\
 \hat{p}_{2,G}\hat{\bar{\pmb{Z}}}_{2,G}^t & \hat{p}_{2,G} \\ \end {array} \right]
\end{equation}
where we have set $\hat{p}_{2,G}=\frac{n_{2,G}}{n_2}$, $\hat{\bar{\pmb{Z}}}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\pmb{Z}(x)$. Using formulae for the inversion of partitioned matrices (see e.g. \cite{searle} p. 27 and \cite{tian} for useful generalizations) one obtains
\begin{equation}\label{inversetildeAs2}
 \pmb{\mathcal{A}}^{-1}_{s_2}=
\left[ \begin {array}{ll}
\pmb{A}^{-1}_{s_2}& \pmb{0} \\
 \pmb{0} & \pmb{0} \\ \end {array} \right]
 + \frac{1}{\gamma}\left[ \begin {array}{ll}
\hat{p}_{2,G}^2\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}\hat{\bar{\pmb{Z}}}^t_{2,G}\pmb{A}^{-1}_{s_2}
& - \hat{p}_{2,G}\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}\\
 -\hat{p}_{2,G}\hat{\bar{\pmb{Z}}}^t_{2,G}\pmb{A}^{-1}_{s_2} & 1 \\ \end {array} \right]
\end{equation}
with $\gamma=\hat{p}_{2,G}-\hat{p}^2_{2,G}\hat{\bar{\pmb{Z}}}^t_{2,G}\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}$.\\
We need
$$\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathcal{Z}}(x)=
(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}^t(x),\hat{p}_{2,G}\hat{\bar{Y}}_G)^t
=((\pmb{A}^{-1}_{s_2}\hat{\pmb{\beta}}_{s_2})^t,\hat{p}_{2,G}\hat{\bar{Y}}_G)^t$$
where $\hat{\bar{Y}}_G=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}Y(x)$. This leads after some algebra to the following relationship between the regressions coefficients
\begin{equation}\label{relationship}
 \hat{\pmb{\theta}}_{s_2}=
\left[ \begin {array}{l}
\hat{\pmb{\beta}}_{s_2} \\
 0 \\ \end {array} \right]
 + \frac{1}{\gamma}\left[ \begin {array}{l}
-\hat{p}_{2,G}^2(\hat{\bar{Y}}_G-\hat{\bar{\pmb{Z}}}^t_{2,G}\hat{\pmb{\beta}}_{s_2})\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}
\\
 \hat{p}_{2,G}(\hat{\bar{Y}}_G-\hat{\bar{\pmb{Z}}}^t_{2,G}\hat{\pmb{\beta}}_{s_2})\end {array} \right]
\end{equation}
Note that the term
$$\hat{\bar{Y}}_G-\hat{\bar{\pmb{Z}}}^t_{2,G}\hat{\pmb{\beta}}_{s_2}=\frac{1}{n_{2,G}}
\sum_{x\in{s_{2,G}}}(Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2})$$
is precisely the mean residual over the small area. Hence, the last component of $\hat{\pmb{\theta}}_{s_2}$ is essentially the residual term. We see that the original regression coefficient $\hat{\pmb{\beta}}_{s_2}$ is corrected in the extended model by the residual term and that the impact of this correction tends to zero as the small area gets smaller with respect to $F$, a very intuitive result indeed. One obtains a very similar but not identical result by least squares minimization under the constraint of zero mean residual over the small area (see \cite{searle}, pp 112-113).\\
In perfect analogy with [\ref{estvarmatrix}] the estimated covariance matrix is given by
\begin{equation}\label{estvarmatrixext}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}=\pmb{\mathcal{A}}_{s_2}^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathcal{R}}^2(x)\pmb{\mathcal{Z}}(x)
\pmb{\mathcal{Z}}(x)^t\Big)\pmb{\mathcal{A}}_{s_2}^{-1}
\end{equation}
where we have set $\hat{\mathcal{R}}(x)=Y(x)-\pmb{\mathcal{Z}}^t(x)\hat{\pmb{\theta}}_{s_2}$.
If the first phase is exhaustive we calculate the synthetic estimator in the extended model
\begin{equation}\label{extsynthethic1}
\hat{\tilde{Y}}_{G,synth}=
\frac{1}{\lambda(G)}\int_G\pmb{\mathcal{Z}}^t(x)\hat{\pmb{\theta}}_{s_2}dx
=\bar{\pmb{\mathcal{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2}
\end{equation}
With $\bar{\pmb{\mathcal{Z}}}^t_{G}=(\bar{\pmb{Z}}^t_{G},1)$ and some algebra one finally obtains
\begin{equation}\label{extsynthethic2}
\hat{\tilde{Y}}_{G,synth}=\bar{\pmb{Z}}_{G}^t\hat{\pmb{\beta}}_{s_2}+
\frac{\alpha}{n_{2,G}}\sum_{x\in{s_{2,G}}}\big(Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}\big)
\end{equation}
where we have set
$$\alpha=\frac{\hat{p}_{2,G}-\hat{p}^2_{2,G}\bar{\pmb{Z}}_{G}^t\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}
{\hat{p}_{2,G}-\hat{p}^2_{2,G}\hat{\bar{\pmb{Z}}}_{2,G}^t\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$$
Clearly $\hat{\tilde{Y}}_{G,synth}$ and
$\hat{{Y}}_{G,small}$ are asymptotically equivalent because $\alpha$ tends to $1$ in large samples. Note that $\alpha=1$ if the sample is exactly balanced, i.e. if $\hat{\bar{\pmb{Z}}}_{2,G}= \bar{\pmb{Z}}_{G}$.\\
By using [\ref{synthetic2}] and replacing $\pmb{Z}(x)$ with $\pmb{\mathcal{Z}}(x)$ we obtain at once the asymptotic variance
\begin{equation}\label{extsynthetic1var}
\hat{\VAR}(\hat{\tilde{Y}}_{G,synth})=
\bar{\pmb{\mathcal{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathcal{Z}}}_{G}
\end{equation}
One can rewrite $\hat{\tilde{Y}}_{G,synth}$ in terms of g-weights in the extended model as in [\ref{gweights1}] and [\ref{gweights2}] with
$$\tilde{g}_{G}(x)=\bar{\pmb{\mathcal{Z}}}_G^t\pmb{\mathcal{A}}^{-1}_{s_2}\pmb{\mathcal{Z}}(x)$$
 \begin{eqnarray}\label{gweightsynthextended}
 \hat{\tilde{Y}}_{G,synth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G}(x)Y(x) \nonumber \\
 \hat{\VAR}(\hat{\tilde{Y}}_{G,synth})&=&\frac{1}{n^2_2}\sum_{x\in{s_2}}\tilde{g}^2_{G}(x)\hat{\mathcal{R}}^2(x)
 \end{eqnarray}
If the first phase is not exhaustive we estimate the true mean of the extended auxiliary variables
\begin{equation}
\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}
\pmb{\mathcal{Z}}(x)
\end{equation}
to get the pseudo-synthetic estimate in the extended model
\begin{equation}\label{pseudosyntheticextended1}
\hat{\tilde{Y}}_{G,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^t\hat{\pmb{\theta}}_{s_2}
\end{equation}
As in [\ref{extsynthethic2}] we have
\begin{equation}\label{extsynthethic3}
\hat{\tilde{Y}}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}+
\frac{\alpha_1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\big(Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}\big)
\end{equation}
where we have set
$\alpha_1=\frac{\hat{p}_{2,G}-\hat{p}^2_{2,G}\hat{\bar{\pmb{Z}}}_{1,G}^t
\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}
{\hat{p}_{2,G}-\hat{p}^2_{2,G}\hat{\bar{\pmb{Z}}}_{2,G}^t
\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$.\\
By [\ref{estvarpseudosynth1}] we get immediately the following consistent estimate of the design-based variance
\begin{equation}\label{estvarpsynth1extended}
\hat{\VAR}(\hat{\tilde{Y}}_{G,psynth})=
\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}
+ \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}}\hat{\pmb{\theta}}_{s_2}
\end{equation}
The variance-covariance matrix of $\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}$ can be estimated as usual by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}}=
\frac{1}{n_{1,G} (n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
(\pmb{\mathcal{Z}}(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G})(\pmb{\mathcal{Z}}(x)-
\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G})^t
\end{equation}
Again, one can rewrite the above expression with the g-weights
$\tilde{g}_{G,1}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^t
\pmb{\mathcal{A}}^{-1}_{s_2}\pmb{\mathcal{Z}}(x)$
namely
\begin{eqnarray}\label{gweightextended}
\hat{\tilde{Y}}_{G,psynth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,1}(x)Y(x)\\
\hat{\VAR}(\hat{\tilde{Y}}_{G,psynth})&=&\frac{1}{n^2_2}
\sum_{x\in{s_2}}\tilde{g}^2_{G,1}(x)\hat{\mathcal{R}}^2(x)+
\hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}}\hat{\pmb{\theta}}_{s_2}
\end{eqnarray}
\textbf{Properties of the g-weights:}\label{gweights}
 \begin{enumerate}
 \item
 The g-weights enjoy the calibration properties
 $\frac{1}{n_2}\sum_{x\in{s_2}}g_{G,1}(x)\pmb{Z}(x)=
 \hat{\bar{\pmb{Z}}}_{1,G}$  and  $\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,1}(x)\pmb{\mathcal{Z}}(x)=
 \hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}$. The proof is immediate by transposing the equalities and by the very definition of the g-weights.
 \item
  The fact that one can assume the g-weights depend only on the point $x$ and not on the whole sample $s_2$ when calculating variances is fully justified by the Taylor expansion leading to the robust design-based covariances.
 \item
 By considering formally the trivial constant local density $Y(x)\equiv 1$ and solving the normal equations one sees that for any $G$ (i.e. also for $G=F$) one has
 $\frac{1}{n_2}\sum_{x\in{s_2}}g_{G,1}(x)=
  \frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,1}(x)=1$, i.e. the g-weights have means equal to 1.
 \item
  When $G=F$ the estimator $\hat{Y}_{reg}$ and the sample mean $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ both converge towards the true value for an arbitrary $Y(x)$ and therefore one gets $\lim_{n_2\to\infty}g_{F,1}(x)=1>0$. This is not true for a proper subset $G\subset F$. In this case, $\hat{\tilde{Y}}_{G,synth}$ and the sample mean $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}Y(x)$ converge towards the true value for an arbitrary $Y(x)$, which implies that $\tilde{g}_{G,1}(x)$ will tend to $0$ for $x\not\in{G}$ (negative values are possible) and to $\frac{n_2}{n_{2,G}} $ for $x\in{G}$.
 \end{enumerate}
In simple random sampling one can derive non-parametric versions of the above procedures, i.e. kernel-based regression estimators or $k$ nearest neighbors estimators (\textbf{knn}). However, they are more difficult to implement than the regression estimators described above, primarily because on must use re-sampling (bootstrap) techniques to obtain reliable variance estimators because analytical results are beyond reach and the classical external variance formula severely underestimate the true variance as shown by simulations. Also, the performances of the non-parametric estimators are not better than the regression estimators (for global and small-area estimation) if the coefficient of determination $R^2$ is sufficient large, say above $0.7$, which is the case for timber volume assessment using modern remote sensing data as auxiliary information (e.g. canopy height parameters obtain via e.g. LiDAR). Details can be found in \cite{mandallazreport4}, \cite{massey2} and the PhD thesis \cite{masseyphd}.
In the next section we generalize the previous results on regression estimators to cluster sampling, widely used in national inventories. The main ideas remain the same but the formulae are slightly more cumbersome due to the random cluster size. The non-parametric estimators are not yet available for cluster sampling.
\newpage
\section{Generalization to cluster sampling}
We follow the description of cluster sampling as defined in \cite{mandallaz} (especially section 5.5). A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$. The geometry of the cluster is given by $M$ vectors $e_1,\ldots e_M$ defining the random cluster $x_l=x+e_l$. $M(x)=\sum_{l=1}^MI_{F}(x_l)$ is the random number of points of the cluster falling into the forest area $F$. We define the local density at the cluster level by $Y_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)Y(x_l)}{M(x)}$, likewise we set $\pmb{Z}_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)\pmb{Z}(x_l)}{M(x)}$. The set $\tilde{F}$ above can be mathematically defined as the smallest set $\{x\in{\mathcal{R}}^2 \mid M(x) \ne 0 \}$. In the first phase we have $n_1$ clusters identified by $x\in{s_1}$ and in the second phase $n_2$ clusters with $x\in{s_2}$, obtained by simple random sampling from $s_1$.\\
We shall use the model-based approach, in which the regression coefficient $\pmb{\beta}_c$ at the cluster level minimizes
$$\EX_{x\in{\tilde{F}}}M(x)(Y_c(x)-\pmb{\beta}^{t}\pmb{Z}_c(x))^2$$
In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (though close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details). The choice of $M(x)$ rather than $M^2(x)$ as weights is suggested by the model-dependent approach. When $Y_c(x)$ is the mean of the
$M(x)$ observations, its variance can be expected to be inversely
proportional to $M(x)$.  This procedure  leads to the normal equation
$$\Big(\EX_{x\in{\tilde{F}}}M(x)\pmb{Z_c}(x)\pmb{Z_c}(x)^{t}\Big)\pmb{\beta}_c=\EX_{x\in{\tilde{F}}}M(x)Y_c(x)\pmb{Z_c}(x)$$
and to $\EX_{x\in{\tilde{F}}}M(x)R_c(x)=0$. An asymptotically design-unbiased estimate $\hat{\pmb{\beta}}_{c,s_2}$  for $\pmb{\beta}_c$ can be obtained  by taking a sample copy of the above equation, i.e.
 \begin{eqnarray}\label{estcluster1}
 \hat{\pmb{\beta}}_{c,s_2}&=&
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}_c(x)\pmb{Z}_c^t(x)\Big)^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \nonumber\\
 &:=&\pmb{A}^{-1}_{c,s_2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \end{eqnarray}
 The empirical residuals at the cluster level are
 $$\hat{R}_{c}(x)=Y_c(x)-\pmb{Z}^t_c(x)\hat{\pmb{\beta}}_{c,s_2}$$
 which satisfy the orthogonality relation
 $$\sum_{x\in{s_2}}M(x)\hat{R}_c(x)\pmb{Z}_c(x)=0$$
 and in particular the zero mean residual property
$$\frac{\sum_{x\in{s_2}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_2}}M(x)}=0$$
Using mutatis mutandis exactly the same arguments as in simple random sampling we get the asymptotic robust design-based estimated variance-covariance matrix
\begin{equation}\label{estvarcluster1}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}=\pmb{A}^{-1}_{c,s_2}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}^2_c(x)
\pmb{Z}_c(x)\pmb{Z}^t_c(x)\Big)\pmb{A}^{-1}_{c,s_2}
\end{equation}
The small-area estimation set up is slightly more complicated because the points of a cluster with $M(x) \ge 0, \;x\in{\tilde{F}}$ can be spread over more than one small area. The notation used in the 2012 version of this technical report was not sufficiently explicit in this respect. This does not affect the regression coefficients, always based on the entire sample in $F$ (the borrowing strength philosophy), but the formulation "after restricting the samples to the small area of interest" is now more explicit in the notation. We define the following quantities:
\begin{eqnarray} \label{extranotation}
M_G(x) &=& \sum_{l=1}^L  I_G(x+e_l) \nonumber \\
Y_{c,G}(x)&=&\frac{\sum_{l=1}^L I_G(x+e_l)Y(x+e_l)}{M_G(x)} \nonumber \\
\hat{\bar{Y}}_{c,2,G} &=& \frac{\sum_{x\in{s_{2,G}}}M_G(x)Y_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}\nonumber \\
\pmb{Z}_{c,G}(x)&=& \frac{\sum_{l=1}^L I_G(x+e_l)\pmb{Z}(x+e_l)}{M_G(x)} \nonumber \\
\hat{Y}_{c,G}(x)&=&\pmb{Z}^t_{c,G}(x)\hat{\pmb{\beta}}_{c,s_2} \nonumber \\
\hat{R}_{c,G}&=& Y_{c,G}(x)-\hat{Y}_{c,G}(x) \nonumber \\
\hat{\bar{\pmb{Z}}}_{c,1,G}&=&\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{Z}_{c,G}(x)}{\sum_{x\in{s_{1,G}}}M_G(x)}
\end{eqnarray}

The estimated variance of the one-phase estimator is given by (\cite{mandallaz}, section 4.3).
\begin{equation}\label{onephaseclustervariance}
\hat{\VAR}(\hat{\bar{Y}}_{c,2,G})=\frac{1}{n_{2,G}(n_{2,G}-1))}
\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2(Y_{c,G}(x)-\hat{\bar{Y}}_{c,2,G})^2
\end{equation}
\noindent where $\bar{M}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)}{n_{2,G}}$.\\
The estimate of the covariance matrix of the auxiliary variables is given by
\begin{equation}\label{estcovauxiliarycluster}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})
(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})^t
\end{equation}
\noindent where $\bar{M}_{1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)}{n_{1,G}}$.\\
The \textbf{pseudo-synthetic estimate} is then
\begin{eqnarray}\label{psynthclusterest}
\hat{Y}_{c,G,psynth}&=&\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}\nonumber \\
&=& \frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)Y_c(x)
\end{eqnarray}
with the g-weights $g_{c,1,G}(x)=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\pmb{A}^{-1}_{c,s_2}M(x)\pmb{Z}_c(x)$. The estimated variance is as in [\ref{estvarpseudosynth1}]

\begin{equation}\label{estvarpseudosynth1cluster}
\hat{\VAR}(\hat{Y}_{c,G,psynth})=
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\end{equation}
The pseudo-synthetic estimate is generally design-biased. Adjusting for the residuals we get the  \textbf{pseudo small-area estimator}
\begin{equation}\label{psmallclusterest}
\hat{Y}_{c,G,psmall}=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}
+ \frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}
\end{equation}
It is asymptotically design-unbiased and, by analogy with simple random sample (see [\ref{pseudosmallvar}]) its variance can be expected to be well approximated by

\begin{eqnarray}\label{estvarpsmallcluster}
\hat{\VAR}(\hat{Y}_{c,G,psmall})&=&
\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}\hat{\bar{\pmb{Z}}}_{c,1,G}
+ \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
\nonumber \\ &+& \frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2
(\hat{R}_{c,G}(x)-\bar{\hat{R}}_{2,G})^2
\end{eqnarray}
where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M_G(x)$ and
$\bar{\hat{R}}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}$.
One can rewrite [\ref{estvarpsmallcluster}] with g-weights and predictions as in [\ref{pseudosmallvar2}].
\par

As in simple two-phase sampling we can transform the above estimator into a synthetic estimator by considering the extended model $\pmb{\mathcal{Z}}_c^t(x)=(\pmb{Z}_c^t(x),I_{c,G}(x))\in{\mathcal{R}}^{(p+1)}$ with
$I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}=\frac{M_G(x)}{M(x)}$. We shall assume that $M_G(x)\equiv M(x)$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$. If we have a partition of the forest in small areas $G_k, k=1,2,\ldots L$, then $\tilde{F} \subset \cup_{k=1}^L \tilde{G}_k$ but it is unlikely in general that the $\tilde{G}_k$ are disjoined. We can reasonably hope that in extensive forest inventory the surface area of points where $M_G(x)< M(x)$ is negligible. The theoretical normal equation
reads
\begin{equation}
\Big(\int_{\tilde{F}}M(x)\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}_c^t(x)dx\Big)\pmb{\theta}_c=:
\pmb{\mathcal{A}}_c\pmb{\theta}_c=\int_{\tilde{F}}M(x)Y_c(x)\pmb{\mathcal{Z}}_c(x)dx
\end{equation}
 which satisfy by construction the two zero mean residuals properties
$\int_{\tilde{F}} M(x)\mathcal{R}_c(x)dx=\int_{\tilde{G}} M(x)\mathcal{R}_c(x)dx=0$.
The second equality will only hold approximately if $I_{c,G}(x)<1$ for some $x$ in $\tilde{G}$.\\
With
$\pmb{\mathcal{A}}_{c,s_2}
=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}_c^t(x)$
we obtain the estimate of the regression coefficients at the cluster level
\begin{equation}\label{regcoeffclusterext}
\hat{\pmb{\theta}}_{c,s_2}=\pmb{\mathcal{A}}^{-1}_{c,s_2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}_c(x)Y(x)\Big)
\end{equation}
with estimated design-based covariance matrix
\begin{equation}\label{regcoeffclusterextestvar}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}=
\pmb{\mathcal{A}}^{-1}_{c,s_2}\Big(\frac{1}{n^2_2}\sum_{\in{s_2}}M^2(x)\hat{\mathcal{R}}_c^2(x)
\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)\Big)\pmb{\mathcal{A}}^{-1}_{c,s_2}
\end{equation}
where the $\hat{\mathcal{R}}_c(x)=Y_c(x)-\pmb{\mathcal{Z}}^t_c(x)\hat{\pmb{\theta}}_{c,s_2}$ are the empirical residuals at the cluster level with respect to the extended model. We define the pseudo-synthetic estimator in the extended model according to
\begin{equation}\label{estimatorclusterextended}
\hat{\tilde{Y}}_{c,G,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,s_2}
\end{equation}
where
$\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{\mathcal{Z}}_{c,G}(x)}
{\sum_{x\in{s_{1,G}}}M_G(x)}$
is the mean of the extended auxiliary vector over the small area.\\
As in [\ref{estvarpseudosynth1cluster}] the estimated variance is given by
\begin{equation}\label{estvarpsynth1clusterextended}
\hat{\VAR}(\hat{\tilde{Y}}_{c,G,psynth})=
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}
+ \hat{\pmb{\theta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}\hat{\pmb{\theta}}_{c,s_2}
\end{equation}
\noindent where
\begin{equation}\label{estcovauxiliaryclusterextended}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{\mathcal{Z}}_{c,G}(x)-
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})
(\pmb{\mathcal{Z}}_{c,G}(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})^t
\end{equation}
\textbf{Remark}:\\
Even if the condition $M_G(x)\equiv M(x)$ is fulfilled in $s_{2,G}$ we cannot expect $\hat{\tilde{Y}}_{c,G,psynth}$ to be asymptotically unbiased (although the sample copy of the normal equation converges towards the theoretical normal equation this does not guarantee zero mean residuals over $G$ as we have seen). However the bias can be expected to be negligible.\\[0.5cm]
Obviously a decomposition similar to [\ref{extsynthethic3}] will hold so that $\hat{\tilde{Y}}_{c,psynth}$ and $\hat{Y}_{c,G,psmall}$ in [\ref{psmallclusterest},\ref{estimatorclusterextended}]
will be close to each other in large samples.\\
Defining the g-weights at the cluster level as
\begin{equation}
\tilde{g}_{c,1,G}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}^t_{c,1,G}\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
\pmb{\mathcal{Z}}_c(x)
\end{equation}
we obtain as usual
\begin{equation}\label{gweightclusterpointestextended}
\hat{\tilde{Y}}_{c,G,psynth}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)Y_c(x)
\end{equation}
and
\begin{equation}\label{gweightclustervarestextended}
\hat{\VAR}(\hat{\tilde{Y}}_{c,G,psynth})=\frac{1}{n_2^2}\sum_{x\in{s_2}}
\tilde{g}^2_{c,1,G}(x)\hat{\mathcal{R}}^2(x)+
\hat{\pmb{\theta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}\hat{\pmb{\theta}}_{c,s_2}
\end{equation}
The synthetic estimator in the extended model corresponds formally to $n_1=\infty$, i.e.
\begin{equation}
\hat{\tilde{Y}}_{c,G,synth}=\bar{\pmb{\mathcal{Z}}}_G^t\hat{\pmb{\theta}}_{c,s_2}
 \end{equation}
 with estimated variance
 \begin{equation}
\hat{\VAR}(\hat{\tilde{Y}}_{c,G,synth})=
\bar{\pmb{\mathcal{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}\bar{\pmb{\mathcal{Z}}}_{G}
\end{equation}
with obvious modification for the g-weights
\begin{eqnarray}
\tilde{g}_{c,G}(x)&=&\bar{\pmb{\mathcal{Z}}}_G^t\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
\pmb{\mathcal{Z}}_c(x)\nonumber \\
\hat{\tilde{Y}}_{c,G,synth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,G}(x)Y_c(x)\nonumber \\
\hat{\VAR}(\hat{\tilde{Y}}_{c,G,synth})&=&\frac{1}{n_2^2}\sum_{x\in{s_2}}
\tilde{g}^2_{c,G}(x)\hat{\mathcal{R}}^2(x)
\end{eqnarray}
\noindent
\textbf{Properties of the g-weights}:
\begin{enumerate}
\item
 We have
 $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)\pmb{Z}_c(x)=
 \hat{\bar{\pmb{Z}}}_{c,1,G}$  and  $\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)\pmb{\mathcal{Z}}_c(x)=
 \hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}$.
 \item
 By considering $Y_c(x)\equiv 1$ one gets
 $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)=
  \frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)=1$, i.e. the g-weights have means equal to 1.
 \item
  When $G=F$ the estimator $\hat{Y}_{c,F,psynth}$ and the sample mean $\frac{\sum_{x\in{s_2}}M(x)Y(x)}{\sum_{x\in{s_2}}M(x)}$ both converge towards the true value over $F$. Hence, for large $n_2$, $g_{c,1,F}(x)\approx \frac{M(x)}{\bar{M}_2}$ with $\bar{M}_2=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)$. Likewise for $\hat{\tilde{Y}}_{G,psynth}$ and the sample mean over $G$, $\hat{\bar{Y}}_{c,2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x) Y_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}$. Hence, for large $n_2$ we get $\tilde{g}_{c,1,G}(x)\approx 0$ for $x\not\in{\tilde{G}}$ (negative values are possible) and to $\tilde{g}_{c,1,G}(x)\approx \frac{M_G(x)}{\bar{M}_{2,G}}$ for $x\in{\tilde{G}}$, where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M_G(x)$.
 \end{enumerate}
\textbf{Remarks:}\\
With simple random sampling the construction of design-unbiased small-area estimators as synthetic or pseudo-synthetic estimators in the extended model containing the indicator variable of the small area of interest is mathematically more convenient than the classical approach. The mathematical approximation of the variances is also more satisfactory than simply treating the internal model as an external one and it can be formulated within the g-weight technique, which is known to offer several theoretical advantages. The same essentially holds in cluster sampling, with a slight advantage maybe for $\hat{Y}_{c,G,psmall}$,with the external variance, in terms of simplicity and validity, since one cannot guarantee in general zero mean residuals over $G$ for the pseudo-synthetic estimator $\hat{\tilde{Y}}_{c,G,psynth}$ in the extended model. It is certainly good practice to calculate the different small-area estimators and their estimated variances, particularly for cluster sampling. Large differences could reveal unsuspected patterns in the data (or eventually even programming errors).\\
It is mathematically clear that one can generalize all the previous results to the simultaneous estimation of $q \ge 2$ small areas by extending the model with $q$ indicator variables (combined extended model). One can conjecture that the combined model will be less efficient, for any given small area, than the individual estimation and, on the other hand, that it will smooth out the residual pattern.\\
To calculate confidence intervals we recommend to use the Student's $T$ distribution on $ n_2$ d.f. for $F$ and $n_{2,G}-1$ d.f. for any small area $G$. If $n_{2,G}$ is to small one can use the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, which will yield short confidence intervals, at the cost of a potential bias. In this case it might instructive to embed $G$ in a larger domain $G^*$ with $n_{2,G^*}$ large enough and to calculate the pseudo-synthetic estimate for $G$ in the model extended with the indicator variable of $G^*$.
\newpage
\section{Generalization to two-stage sampling}
In many applications costs to measure the response variable $Y_i$ are high. For instance,
a good determination of the volume may require that one records
$DBH$, as well as the diameter at $7m$ above ground and total height
in order to utilize a three-way volume function. However, one could rely
on a coarser, but cheaper, approximation of the volume based only on
$DBH$. Nonetheless, it may be most sensible to assess those three
parameters only on a sub-sample of trees. We now briefly formalize this
simple idea, which is used in the Swiss National Forest Inventory. The reader is referred to (\cite{mandallaz}, section 4.4, 4.5, 5.4 and 9.5) for details. For each point $x\in{s_2}$ trees are drawn with
probabilities $\pi_i$. The set of selected trees is denoted by
$s_{2}(x)$. From each of the selected trees $i\in{s_{2}(x)}$ one
gets an approximation $Y_i^*$ of the exact value $Y_i$. From the
finite set $s_{2}(x)$ one draws a sub-sample
$s_{3}(x)\subset{s_{2}(x)}$ of trees by Poisson sampling. For each tree $i\in{s_{3}(x)}$
one then measures the exact variable $Y_i$. Let us now define the
second stage indicator variable
\begin{equation}
 J_i(x)=\begin{cases}&1 \text{ if $i\in s_{3}(x)$}\\
                      &0 \text{ if $i\not\in s_{3}(x)$}
         \end{cases}
\end{equation}

 \par To construct a good point estimate, we must have
 \textbf{the residual} $R_i=Y_i-Y_i^*$ which is known only for trees
 $i\in{s_{3}(x)}$. The \textbf{generalized local density} $Y^*(x)$ is defined
 according to    \index{Generalized local density}
 \begin{eqnarray}\label{gdens}
 Y^*(x)&=&\frac{1}{\LF}\left( \sum_{i=1}^N \frac{I_{i}(x)Y_i^*}{\pi_i} +
 \sum_{i=1}^N \frac{I_{i}(x)J_{i}(x)R_i}{\pi_{i}p_i}\right)\nonumber \\
 &=&\frac{1}{\LF}\left( \sum_{i\in{s_{2}(x)}} \frac{Y_i^*}{\pi_i}
 +\sum_{i\in{s_{3}(x)}} \frac{R_i}{\pi_{i}p_i}\right)
 \end{eqnarray}
 where the $p_i$ are the conditional inclusion probabilities for the the second stage sampling, i.e. $p_i=\PR(J_i(x)=1 \mid I_i(x)=1)$.
 It follows from general principles presented in (\cite{mandallaz}, sections 4.4 and 4.5) that one can use all the previous results by replacing everywhere the exact local densities $Y(x)$, or $Y(x_l)$ in cluster sampling, by the corresponding generalized local densities $Y^*(x)$ or $Y^*(x_l)$. The second-stage variance is automatically taken into account. More mathematical details are given in \cite{mandallazreport5}.
 \newpage
\section{Examples}
\subsection{Post-stratification}\label{poststrat}
We consider the important special case of post-stratification, which illustrates the main issues. We consider a forested area $F$ partitioned in $L$ strata $F_k$, i.e. $F=\cup_{k=1}^L F_k$ and a small area $G\subset F$, we set $G_k=G \cap F_k$. Note some $G_k$ might be the empty set. The auxiliary vector is defined by the indicator variables of the $L$ strata, i.e.
$$\pmb{Z}^t(x)=(I_{F_1}(x),I_{F_2}(x),\ldots I_{F_L}(x))$$
 where $I_{F_k}(x)=1$ if $x\in{F}_k$ otherwise $I_{F_k}(x)=0$. Note that condition [\ref{linearmodel2}] is fulfilled. Straightforward calculations lead to the $(L,L)$ diagonal matrix $\pmb{A}_{s_2}=\frac{1}{n_2}diag(n_{2,k})$ where $n_{2,k}=\sum_{x\in{s_2}}I_{F_k}(x)$. This leads to the obvious regression estimate
 $$\hat{\pmb{\beta}}_{s_2}=(\hat{\beta}_1,\hat{\beta}_2,\ldots \hat{\beta}_L)^t$$
 with the empirical strata means $\hat{\beta}_k=\frac{1}{n_{2,k}}\sum_{x\in{s_2}\cap F_k}Y(x)=\hat{\bar{Y}}_k$.
 After some elementary algebra the estimated variance-covariance matrix is found to be the diagonal $(L,L)$ matrix
 $$\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}=diag(\frac{s_k^2}{n_{2,k}})$$
 where
 $s_k^2=\frac{1}{n_{2,k}}\sum_{x\in{F_k}}\hat{R}^2(x)$ with $\hat{R}(x)=Y(x)-\hat{\beta}_k$ for $x\in{F}_k$.\\
 One obtains for the empirical mean of the auxiliary vector over the small area
 $$\hat{\bar{\pmb{Z}}}_{1,G}=(\hat{p}_{1,G},\hat{p}_{2,G},\ldots, \hat{p}_{L,G})^t=\hat{\pmb{p}}_{1,G}$$
 where
 $$\hat{p}_{k,G}=\frac{\sum_{x\in{s_1}}I_{G_k}(x)}{\sum_{x\in{s_1}}I_G(x)}:=\frac{n_{1k,G}}{n_{1,G}}$$
 \noindent are the proportions of the strata surfaces areas within the small area as estimated from the large sample. Conditionally on $n_{1,G}$ the $n_{1k,G}$ follow the multinomial distribution with cell probabilities given by the vector
 $\pmb{p}^t_{G}=(\frac{\lambda(G_1)}{\lambda(G)},\frac{\lambda(G_2)}{\lambda(G)},\ldots \frac{\lambda(G_L)}{\lambda(G)})$. In this case the estimated variance-covariance matrix is known to be given by
 \begin{equation}\label{covmatrixcellfreq}
 \hat{\pmb{\Sigma}}_{\hat{\pmb{p}}_{1,G}}=
\frac{1}{n_{1,G}}\left[ \begin {array}{cccc}
\hat{p}_{1,G}(1-\hat{p}_{1,G})& \hat{p}_{1,G}\hat{p}_{2,G} &\ldots& \hat{p}_{1,G}\hat{p}_{1,L}\\
\hat{p}_{1,G}\hat{p}_{2,G}& \hat{p}_{2,G}(1-\hat{p}_{2,G}) &\ldots& \hat{p}_{2,G}\hat{p}_{1,L}\\
\ldots & \ldots & \ldots & \ldots \\
\hat{p}_{1,G}\hat{p}_{1,L} & \hat{p}_{2,G}\hat{p}_{1,L}& \ldots & \hat{p}_{1,L}(1-\hat{p}_{1,L}) \end {array} \right]
\end{equation}
Note that the same is obtained by using [\ref{estvarcovaux}] after replacing $n_{1,G}-1$ by $n_{1,G}$. Simple algebra leads then to the pseudo-synthetic estimate
\begin{equation}\label{postsynth1}
\hat{Y}_{G,psynth}=\hat{\pmb{p}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}=\sum_{k=1}^L\hat{p}_{k,G}\hat{\beta}_k
\end{equation}
with estimated asymptotic design-based variance
\begin{equation}\label{postsynth1}
\hat{\VAR}(\hat{Y}_{G,psynth})=\sum_{k=1}^L\hat{p}^2_{k,G}\frac{s_k^2}{n_{2,k}} +
\frac{1}{n_{1,G}}\sum_{k=1}^L \hat{p}_{k,G}\big(\hat{\beta}_k-\hat{Y}_{G,psynt}\big)^2
\end{equation}
When $n_1=\infty$ and $G=F$ this is precisely the exact conditional variance estimate, i.e. given the $n_k$. The g-weights are $g_{s_2}(x)=p_k\frac{n_2}{n_{2,k}}$ for $x\in{F}_k$, where $p_k=\frac{\lambda(F_k)}{\lambda(F)}$. Thus, for $n_1 < \infty$ the overall variance will depend on the variances within strata and on the variance between strata, which is given by the second term. Note also that for $G=F$ in [\ref{postsynth1}] the strata weights are estimated from the large sample whereas this is not the case for the external model approach [\ref{varexternalsimple}], which illustrates perfectly the better conditional properties of the g-weights technique (see \cite{mandallaz}, p.84).\\
\noindent\textbf{Remarks}\\
\noindent If we assume the $\lambda(F_k)$ and therefore $\pmb{A}$ to be known, the estimator $$\hat{\pmb{\beta}}_0=\pmb{A}^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)$$ is easily found to be $(\frac{n_{21}}{n_2p_1}\hat{\bar{Y}}_1, \frac{n_{22}}{n_2p_2}\hat{\bar{Y}}_2,\ldots \frac{n_{2L}}{n_2p_L}\hat{\bar{Y}}_L)^t$ and yields $\hat{Y}_{synth}=\frac{1}{n_2}\sum_{\in{s_2}}Y(x)$, which is unbiased but useless. If we use $\pmb{A}_{s_2} $ to estimate $\pmb{\beta}$ we get as shown above $\hat{\beta}_k=\hat{\bar{Y}}_k$, which is very intuitive, and if we use $\pmb{A}^{-1}$ instead of $\pmb{A}^{-1}_{s_2}$ in the Taylor approximation for the variance, we obtain $\hat{\VAR}(\hat{Y}_{synth})=\sum_{k=1}^L\frac{n_{2k}s_k^2}{n^2_2}$ instead of $\sum_{k=1}^L p_k^2 \frac{s_k^2}{n_{2k}}$, the later is of course much better from a conditional point of view (even if both estimates are asymptotically equivalent). This examples illustrates why it is better to work with $\pmb{A}^{-1}_{s_2}$ throughout.
\\
It can be easily checked that the original small-area estimator is given by
$$\hat{Y}_{G,psmall}=\sum_{k=1}^L \hat{p}_{k,G}\hat{\bar{Y}}_k+\sum_{k=1}^L\frac{n_{2k,G}}{n_{2,G}}
(\hat{\bar{Y}}_{k,G}-\hat{\bar{Y}}_k)$$
where $n_{2k,G}=\sum_{x\in{s_2}}I_{G_k}(x)$ and $\hat{\bar{Y}}_{k,G}=\frac{1}{n_{2k,G}}\sum_{x\in{s_2 \cap G_k}}Y(x)$. Thus, the residual term will have an impact if the strata means within the small area differ from the strata means within the entire domain, which is intuitively clear.
\\ The formulae for the variances are very cumbersome and not really informative, likewise for
$\hat{\tilde{Y}}_{G,psynth}$ in the extended model.
\newpage
\subsection{Case study}
We reanalyze the case study described with in Chapter 9 of \cite{mandallaz}. The inventoried area covered $218ha$.  The auxiliary information is based on 16 stands defined by the following qualitative variables obtained from the manual interpretation of aerial photographs:
\begin{enumerate}
\item
\textbf{Developmental stage} \newline This entails four categories
``pole stage=3," ``young timber tree=4," ``middle age timber
tree=5," and ``old timber tree=6." These were assigned according to
the dominant diameter.
\item
\textbf{Degree of mixture} \newline This variable was simplified to
the categories of ``predominantly conifers=1" and ``predominantly
broadleaves=2."
\item \textbf{Crown closure} \newline
This variable was based on canopy density, defined as the proportion
of the entire ground surface within the stand that was covered by
the tree crowns. It was simplified to the categories of ``dense=1"
and ``close=2."
\end{enumerate}
These factors produced $4 \times 2 \times 2=16$ possible stands, all
of which were found on the study site. \\
The inventory utilized systematic cluster sampling. The cluster
comprises five points: central point, two points each established
$30\;m$ east or west of the central point; two other points each
established $40\;m$ either north or south of the central point.\\
The first phase sets the
central cluster point on a $120\;m$ W-E by $75\;m$ N-S rectangular
grid (note that the clusters partially overlapped in the N-S
direction). The second, terrestrial phase, place the central point on a 1:4
sub-grid of the first phase, i.e. on a $240\;m$ W-E by $150\;m$ N-S
systematic rectangular grid. The terrestrial inventory was purely one-stage
with simple circular plots of $300m^2$ horizontal surface area, and
an inventory threshold set at 12cm DBH.\\
We use the following linear model with the
vector $\pmb{Z}(x)$:

\begin{itemize}
\item
$Z_1(x) \equiv 1$ intercept term
\item
$Z_2(x)=1$ if $x$ lies in Development Stage 3 and $Z_2(x)=0$
otherwise
\newline $Z_3(x)=1$ if $x$ lies in Development Stage 4 and $Z_3(x)=0$
otherwise
\newline $Z_4(x)=1$ if $x$ lies in Development Stage 5 and $Z_4(x)=0$
otherwise
\newline $Z_2(x)=Z_3(x)=Z_4(x)=-1$ if $x$ lies in Development Stage
6
\item
$Z_5(x)=1$ if $x$ lies in a coniferous stand and $Z_5(x)=-1$
otherwise
\item
$Z_6(x)=1$ if $x$ lies in a dense stand and $Z_6(x)=-1$ otherwise
\end{itemize}
Hence, we have an additive ANOVA model with 7 parameters, as compared with 16 parameters for the full stratification model. The coefficient of determination $R^2$ (in the classical model-dependent approach, i.e. without taking the cluster structure into account) was satisfactory for the stem density ($0.53$) but not for the basal area $(0.19)$. This case study was performed in 1989 and is the only data available for cluster sampling in Switzerland (the reason for using cluster sampling was to provide good estimates of the variograms for the geostatistical Kriging procedures, see \cite{mandallaz} and \cite{dmhabil} for details). It should be emphasized that with modern remote sensing techniques (e.g. LiDAR or digital aerial photographs) one can obtain much better $R^2$'s, e.g.  around $0.70$ or more for timber volume (mean canopy height being the most important explanatory variable).\\
We shall consider 5 small areas:
\begin{itemize}
\item
Small area $G_1$ ($\approx 17ha$) was used for a full census. The condition that a cluster hitting the small area has all its points in $F$ within the small area is occasionally  violated (i.e. $I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)} <1$ for some $x$), so that the extended model for $G_1$ is only approximately correct. The mean residual over the small area is not exactly zero. The true values for basal area and stem densities are known.
\item
Small area $G_2$ ($\approx 33ha$) is the most eastern part of the forest.
\item
Small area $G_3$ ($\approx 46ha$) is the most southern part of the forest.
\item
Small area $G_4$ ($\approx 55ha$) is the central part north of $G_3$.
\item
Small area $G_5$ ($\approx 84ha$) is the most western part north of of $G_3$.
\end{itemize}
We have $F=G_2\cup G_3\cup G_4\cup G_5$ and $I_{c,G_k}(x) \equiv 1$ for $k=2,3,4,5$.
Fig. \ref{zbergterrestrial} displays the terrestrial plots according to the domains $G_2-G_5$. Stand map of $F$ and detailed maps of $G_1$ are given in \cite{mandallaz} Chapter 8.
\begin{figure}[h]
\begin{center}
\caption{\label{zbergterrestrial}\textbf{Location of terrestrial plots in $G_2-G_5$}}
\end{center}
\begin{center}
\includegraphics[scale=0.50]{domain2.jpg}
\end{center}
\end{figure}

Tables \ref{result1} and \ref{result3} display the results for the basal area and the stem density for the small areas $G_1$, $G_2$, $G_3$, $G_4$ and $G_5$.\\
The external variances for $\hat{Y}_{c,G,psmall}$ are given by the equations
\begin{equation*}
\hat{Y}_{c,G,psmall}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)\hat{Y}_{c,G}(x)}{\sum_{x\in{s_{1,G}}}M_G(x)}
+\frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}
\end{equation*}
and
\begin{eqnarray*}
\widehat{\VAR}_{ext}(\widehat{Y}_{c,G,psmall})&=&\left(1-\frac{n_{2,G}}{n_{1,G}}\right)
\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}
\sum_{x\in{s_{2,G}}}\left(\frac{M_G(x)}{\bar{M}_{2,G}}\right)^{2}(\hat{R}_{c,G}(x)-\hat{\bar{R}}_{c,2,G})^2\\
&+&
\frac{1}{n_{1,G}}\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\left(\frac{M_G(x)}{\bar{M}_{2,G}}
\right)^{2}(Y_{c,G}(x)-\hat{Y}_{c,2,G})^2
\end{eqnarray*}
The resulting standard errors are given in $(-)$.\\
Tables \ref{result1} (basal area) and \ref{result3} (density of stems) summarize the results for the sample mean $\hat{\bar{Y}}_G$, the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, the pseudo small-area estimator $\hat{Y}_{c,G,psmall}$ and the pseudo-synthetic estimator in the extended model  $\hat{\tilde{Y}}_{c,G,psynth}$ for $F$ and the small areas $G_1$, $G_2$, $G_3$, $G_4$, $G_5$. \\
The extended model for the $\hat{\tilde{Y}}_{c,G_k,psynth}$ contains only the indicator variable of the corresponding small area $G_k$. For this reason the corresponding estimates for the entire domain $F$ depend on the selected $G_k$ and they are given separately in Tables \ref{result2} and \ref{result4}.\\
 We also consider the joint estimation of $F$ and the small areas $G_2$, $G_3$, $G_4$ and $G_5$, which form a partition of $F$. The corresponding model $\pmb{\mathcal{Z}}(x)$ contains the $4$ indicator variables $I_{G_k}(x)$ ($k=2,3,4,5$), the previous components $Z_l(x), l=2,3,4,5,6$ but no longer the intercept term $Z_1(x) \equiv 1$ (otherwise $\pmb{\mathcal{A}}_{s_2}$ would be singular because $Z_1(x)$ is a linear combination of the $I_{G_k}(x)$). The results for this estimator, denoted by $ \hat{\tilde{Y}}_{c,G,cpsynth}$, are displayed in Table  \ref{result5}.\\
 All the calculations were performed with the linear algebra procedure \textbf{proc IML} of the statistical software package SAS.

\section{Discussion and conclusions}
 The results for the basal area are disappointing, except maybe for global estimation, in the sense that the regression estimators, with the exception of the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, frequently have a larger error than the sample mean. This is of course due to the very low coefficient of determination $R^2=0.19$ and we shall therefore not pursue the discussion for basal area, simply noting that all point estimates were close to each other, with the exception of the small area $G_1$ with the full census, for which the pseudo-synthetic estimator was closer to the true values (also for the density of stems). As confirmed by simulations this was due to the fact that the plots within this small area were in the lower tail of the distribution for both basal area and the density of stem. \\
As expected for the stem density (with an acceptable $R^2=0.53$) the two-phase estimators perform much better than the sample mean. For the classical small-area estimator $\hat{Y}_{c,G,psmall}$ the standard errors based on the external model assumption were smaller than their counterparts based on the g-weights (i.e. on the design-based covariance matrix of the regression coefficients) for $F,G_3,G_4,G_5$ and slightly larger for $G_1,G_2$, but the differences were small, a reassuring result. The point estimates of  $\hat{Y}_{c,G,psmall}$ were very close to the point estimates of the pseudo-synthetic estimator $\hat{\tilde{Y}}_{c,G,psynth}$ (in the extended model with the single indicator variable of the small area of interest), which has slightly smaller g-weights standard errors, except for $F$ and $G_1$. From a practical point of view $\hat{Y}_{c,G,psmall}$ and $\hat{\tilde{Y}}_{c,G,psynth}$ are essentially equivalent. The estimators $\hat{\tilde{Y}}_{c,G,psynth}$, $\hat{Y}_{c,G,psmall}$ and the combined estimator $\hat{\tilde{Y}}_{c,G,cpsynth}$ (model without intercept but extended with the 4 indicator variables of $G_2,G_3,G_4,G_5$) were also practically equivalent for $F,G_2,G_3,G_4,G_5$, the combined estimator having slightly smaller g-weights standard errors.\\
The pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$ yields point estimates close to those of
$\hat{Y}_{c,G,psmall}$, $\hat{\tilde{Y}}_{c,G,psynth}$, $\hat{\tilde{Y}}_{c,G,cpsynth}$ (recall that
$\hat{Y}_{c,F,psmall}=\hat{Y}_{c,F,psynth}$) for $F, G_2,G_3,G_4,G_5$ but, of course, with much smaller g-weights based standard errors, at a cost of a potential bias. For $G_1$ the point estimate $\hat{Y}_{c,G_1,psynth}$ was very close to the true value (for the density of stem and the basal area), precisely because the clusters hitting $G_1$ were, as afore mentioned, not "representative" for $G_1$. This can happen!\\
From a mathematical point of view the g-weights technique in the models extended by the indicator variables of the small areas is without any doubts the most elegant approach in simple random sampling: it bypasses the residuals terms and allows for a straightforward calculation of the asymptotic variance that takes into account the errors of the regression coefficients. The same is likely to hold in large sample also with cluster sampling. The results from the case study show that the differences between the asymptotically unbiased estimators are very small from a practical point of the view, and that a potentially biased estimator can by chance be closer to the truth.

\newpage

\bibliography{biblio1}
\newpage
\begin{table}[h]
\begin{center}
\caption{ \label{result1}\textbf{Results for basal area}}
\end{center}
\begin{center}
 \begin{tabular}{l|c|c|c|c|c|}\hline
  \textbf{Domain}     & $n_1:n_2$  & $\hat{\bar{Y}}_G$ & $\hat{Y}_{c,G,psynth}$  & $\hat{Y}_{c,G,psmall}$ & $ \hat{\tilde{Y}}_{c,G,psynth}$ \\[1.0ex]
  & $n_1\bar{M}_1:n_2\bar{M}_2$   & &   & & \\[1.0ex]
  $ $ & $ $ & $ $ &  $ $ & $ $ &  \\ \hline\hline
 $F$     & $ 298:73$     & $31.90$   &  $31.34$     &  $31.34$       & see Table \ref{result2}    \\ [1.0ex]
         & $1203:298$    & $(1.08)$  &  $[0.94]$    &  $[0.94]$      & $ $     \\ [1.0ex]
         &               &           &  $(0.89)$    &  $(0.89)$      & $ $     \\ \hline\hline
 $G_1$   & $29:8$        &  $24.54$  & $30.28$      &  $23.99$       & $25.55$            \\ [1.0ex]
 true=29.60  & $92:19$   &  $(3.59)$ & $[1.37]$      &  $[3.96]$      & $[3.76]$          \\ [1.0ex]
         & $ $           &           &              &  $(3.68)$      &  $ $             \\ \hline \hline
 $G_2$ & $49:9$         & $30.69$     &  $28.27$    &  $29.32 $      & $29.31$     \\ [1.0ex]
       & $185:41$       & $(2.05)$    & $[1.47]$    &  $[2.55]$      & $[2.33]$     \\ [1.0ex]
       & $ $            &   $ $   &   $ $           &  $(2.08)$      &  $ $         \\ \hline \hline
 $G_3$ & $73:18$   &  $32.32$         & $31.62$     &  $31.46$       & $31.46$      \\ [1.0ex]
       &$250:66$   &  $(2.05)$        & $[1.54]$    &  $[2.37]$      & $[2.18] $    \\[1.0ex]
       &           &   $ $   &                      &  $(1.87)$      & $ $               \\ \hline \hline
 $G_4$ & $81:17$   &  $27.63$         & $29.72$     &  $27.78$       & $27.77$          \\ [1.0ex]
       & $306:69$  &  $(2.15)$        & $[1.15] $   &  $[2.27]$      & $[1.98]$        \\ [1.0ex]
       & $ $       &  $ $    &  $ $                 &  $(2.00)$      & $ $            \\ \hline \hline
 $G_5$ &  $125:29$ &  $34.49$         &  $33.50$    & $34.33$        & $34.33$           \\ [1.0ex]
       & $462:122$ &  $(1.78)$        &$[1.00]$     & $[1.55]$       & $[1.30]$           \\ [1.0ex]
       &           &  $ $             & $$          & $(1.35)$       & $ $    \\    \hline \hline

\end{tabular}
\end{center}
\textbf{Design-based standard errors:} $[-]$ (g-weights based) and $(-)$ (based on external or sample variance)\\
For $G_1$ the mean residual of $ \hat{\tilde{Y}}_{c,G,psynth}$ was $-1.59$ and $0$, as expected, for all other domains.
\end{table}

\begin{table}[h]
\begin{center}
\caption{ \label{result2}\textbf{Results basal area for $F$ with extended models}}
\end{center}
\begin{center}
 \begin{tabular}{|c|c|c|c|c|}\hline
  $\hat{\tilde{Y}}^{(1)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(2)}_{c,F,psynth}$& $\hat{\tilde{Y}}^{(3)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(4)}_{c,F,psynth}$ &  $\hat{\tilde{Y}}^{(5)}_{c,F,psynth}$ \\[1.0ex] \hline \hline
  $31.30$    & $31.36$   & $31.35$   & $31.30$  & $31.33$ \\[1.0ex]
  $[0.92] $  & $[0.93]$  & $[0.94]$  & $[0.92]$ & $[0.94]$ \\[1.0ex] \hline \hline
\end{tabular}
\end{center}
$\hat{\tilde{Y}}^{(k)}_{c,F,psynth}$ is the pseudo-synthetic estimate for $F$ in the model extended with the indicator variable of domain $G_k$, $k=1,2,3,4,5$. The g-weights based standard errors are given in $[-]$.
\end{table}

\newpage

\begin{table}[h]
\begin{center}
\caption{ \label{result3}\textbf{Results for stem density}}
\end{center}
\begin{center}
 \begin{tabular}{l|c|c|c|c|c|}\hline
  \textbf{Domain}     & $n_1:n_2$  & $\hat{\bar{Y}}_G$ & $\hat{Y}_{c,G,psynth}$  & $\hat{Y}_{c,G,psmall}$ & $ \hat{\tilde{Y}}_{c,G,psynth}$ \\[1.0ex]
  & $n_1\bar{M}_1:n_2\bar{M}_2$   & &   & & \\[1.0ex]
  $ $ & $ $  & $ $ & $ $   & $ $  &  \\ \hline\hline
 $F$  &    $ 298:73$ & $321.03$  &  $325.79$   &  $325.79 $       & see Table \ref{result4}     \\ [1.0ex]
      & $1203:298$   & $(18.44)$ & $[12.80 ]$  &  $[12.80 ]$      & $ $     \\ [1.0ex]
      &              &           & $(12.09 )$  &  $(12.09)$      & $ $     \\ \hline\hline
 $G_1$ & $29:8$      & $245.61$  & $279.54 $   &  $257.34$       & $258.20$            \\ [1.0ex]
 true=$280.23$ & $92:19$ & $(65.51)$ & $[27.26]$   &  $[48.26]$       & $[49.56]$          \\ [1.0ex]
       & $ $         &           &             &  $(48.29 )$      &  $ $         \\ \hline \hline
 $G_2$ & $49:9$      & $393.50$  & $400.49 $   &  $406.47 $       & $406.41$     \\ [1.0ex]
       & $185:41$    & $(79.69)$ &$ [30.95 ]$  &  $[42.96 ]$      & $[39.36 ]$     \\ [1.0ex]
       & $ $       &   $ $       &   $ $       &  $(43.49 )$      &   $ $         \\ \hline \hline
 $G_3$ & $73:18$   &  $275.25$   & $279.75 $   &  $282.46 $       & $282.40$      \\ [1.0ex]
       &$250:66$   &  $(20.98)$  & $[17.16 ]$  &  $[22.68 ]$      & $[20.77]$    \\[1.0ex]
       &           &   $ $       &             &  $(16.56)$       & $ $       \\ \hline \hline
 $G_4$ & $81:17$   &  $284.06$   &  $307.85$   &  $274.79$        & $274.53 $          \\ [1.0ex]
       & $306:69$  &  $(39.51)$  &  $[19.69]$  &  $[26.64]$       & $[23.55]$        \\ [1.0ex]
       & $ $       &  $ $        &  $ $        &  $(24.12)$       & $ $            \\ \hline \hline
 $G_5$ &  $125:29$ &  $342.35$   &  $332.66$   &  $347.89$        & $347.91 $           \\ [1.0ex]
       & $462:122$ &  $(24.50)$  &  $[15.33]$  &  $[21.26]$       & $[18.65]$           \\ [1.0ex]
       &           &  $ $     & $$             &  $(17.49)$       & $ $       \\    \hline \hline

\end{tabular}
\end{center}
\textbf{Design-based standard errors:} $[-]$ (g-weights based) and $(-)$ (based on external or sample variance)\\
For $G_1$ the mean residual of $ \hat{\tilde{Y}}_{c,G,psynth}$ was $-0.99$ and $0$, as expected, for all other domains.
\end{table}

\begin{table}[h]
\begin{center}
\caption{ \label{result4}\textbf{Results for stem density for $F$ with extended models}}
\end{center}
\begin{center}
 \begin{tabular}{|c|c|c|c|c|}\hline
  $\hat{\tilde{Y}}^{(1)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(2)}_{c,F,psynth}$& $\hat{\tilde{Y}}^{(3)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(4)}_{c,F,psynth}$ &  $\hat{\tilde{Y}}^{(5)}_{c,F,psynth}$ \\[1.0ex] \hline \hline
  $325.62$    & $325.88$  & $325.72$   & $325.15$   & $325.56$ \\[1.0ex]
  $[12.81]$   & $[12.84]$ & $[12.85]$  & $[12.67]$  & $[12.62]$ \\[1.0ex] \hline \hline
\end{tabular}
\end{center}
$\hat{\tilde{Y}}^{(k)}_{c,F,psynth}$ is the pseudo-synthetic estimate for $F$ in the model extended with the indicator variable of domain $G_k$, $k=1,2,3,4,5$. The g-weights based standard errors are given in $[-]$.
\end{table}

\begin{table}[h]
\begin{center}
\caption{ \label{result5}\textbf{Two-phase combined estimates}}
\end{center}
\begin{center}
 \begin{tabular}{|l|c|c|c|}
   \textbf{Domain} & \textbf{sample sizes} & \textbf{basal area}  & \textbf{stem density}\\[1.0ex]
 $ $   & $n_1:n_2$    &  & \ \\[1.0ex]
 $ $   & $n_1\bar{M}_1:n_2\bar{M}_2$  & $ \hat{\tilde{Y}}_{c,G,cpsynth}$ &$\hat{\tilde{Y}}_{c,G,cpsynth}$ \\ \hline
 $F$            & $ 298:73$   &  $31.32 $    &  $325.17$       \\ [1.0ex]
                & $1203:298$  &  $[0.93]$    &  $[12.62]$       \\ \hline \hline
 $G_2$          &  $49:9$     &  $29.39 $    &  $407.84$        \\ [1.0ex]
                &  $185:41$   &  $[2.23]$    &  $[39.04]$       \\ \hline \hline
 $G_3$          &  $73:18$     &  $31.57 $    &  $284.17$         \\ [1.0ex]
                & $250:66$    &  $[2.09]$    &  $[18.09]$       \\ \hline \hline
 $G_4$          &  $81:17$     &  $27.77 $    &  $274.59$        \\ [1.0ex]
                & $306:69$    &  $[1.99]$   &  $[23.49]$        \\ \hline \hline
 $G_5$          &  $125:29$    &  $34.31$    &  $347.76$          \\ [1.0ex]
                & $462:122$   &  $[1.24]$   &  $[16.61]$         \\ \hline \hline
\end{tabular}
\end{center}
The g-weights based standard errors are given in $[-]$. As expected on mathematical grounds all extended models yielded empirical means of the residuals over the entire domain and over one or many small areas which were equal to zero ($<10^{-12})$.
\end{table}

\end{document}

\begin{table}[h]
\begin{center}
\caption{ \label{newresult1}\textbf{Two-phase estimates for basal area}}
\end{center}
\begin{center}
 \begin{tabular}{|l|c|c|c|c|}\hline
                      & \textbf{sample sizes} & $\hat{Y}_{c,G,psynth}$   &$ \hat{Y}_{c,G,psmall}$ & $\hat{\tilde{Y}}_{c,G,psynth}$\\[1.0ex]
                & $n_1:n_2$  &   &  &  \\[1.0ex]
 \textbf{Domain}& $n_1\bar{M}_1:n_2\bar{M}_2$ &   &  &  \\ \hline \hline
 $F$            &  $ 298:73$ &  $31.34$    &  $31.34$    & $31.30[0.92]$   \\ [1.0ex]
                & $1203:298$ &  [0.94]     &  $[0.94]$   & $31.35[0.93]$ \\ [1.0ex]
                &            &             &  $(0.91)$   & $31.35[0.94]$  \\ \hline \hline
 $G_1$          &  $29:8$    &  30.28    &  $23.99$              & $25.55$     \\ [1.0ex]
  true=$29.60$  &  $92:19$   &  [1.34]   &  $[3.90]$             & $[3.79]$   \\ [1.0ex]
                &            &           &  $(3.68)$             &             \\ \hline\hline
 $G_2$          &  $49:9$    &  28.27    &   $29.32$             & $29.31$     \\ [1.0ex]
                &  $185:41$  &  [1.40]   &  $[2.52]$             & $[2.23]$  \\ [1.0ex]
                &            &           &  $(2.08)$             &            \\ \hline
 $G_{21}\subset G_2$ &$17:3$ & $25.55 $  &   $29.52 $            & $ 29.61$     \\ [1.0ex]
                 &  $39:15$   &  $[2.16]$     &  $[ 4.13]$        & $[2.95]$  \\ [1.0ex]
                &            &               &  $(3.53)$         &            \\ \hline \hline
  $G_3$         &  $73:18$   &  31.62    &  $31.46$              & $31.46$  \\ [1.0ex]
                & $250:66$   &  [1.47]   &   $[2.33]$            & $[2.16] $ \\[1.0ex]
                &             &            &  $(1.87)$            & \\ \hline \hline
\end{tabular}
\end{center}
\textbf{Standard errors:} $[-]$ (Taylor, g-weights) and $(-)$ (external model)
\end{table}
\noindent \textbf{Remark:}\\
The mean residual for small area $G_1$ was $-1.59$ for the extended model instead of $0$ because $I_{c,G}(x) \not\equiv 1$.

\newpage
\begin{table}[h]
\begin{center}
\caption{ \label{newresult2}\textbf{Two-phase estimates for stem density}}
\end{center}
\begin{center}
 \begin{tabular}{|l|c|c|c|c|}\hline
        & \textbf{sample sizes} & $\hat{Y}_{c,G,psynth}$  & $\hat{Y}_{c,G,psmall}$ & $ \hat{\tilde{Y}}_{c,G,psynt}$ \\[1.0ex]
        & $n_1:n_2$   & &   &  \\[1.0ex]
    \textbf{Domain} & $n_1\bar{M}_1:n_2\bar{M}_2$ &   &  &  \\ \hline\hline
 $F$            & $ 298:73$  &  $325.79$    &  $325.79$       & $325.62[12.81]$     \\ [1.0ex]
                & $1203:298$ &  $[12.80]$   &  $[12.80]$      & $325.88[12.84]$     \\ [1.0ex]
                &            &              &  $(12.39)$      & $325.72[12.85]$ \\ \hline\hline
 $G_1$          &  $29:8$    &  $279.54$    &  $257.34$              & $258.20$     \\ [1.0ex]
 true=$280.23$  &  $92:19$   &  $[22.65]$   &  $[45.81]$             & $[54.07]$   \\ [1.0ex]
                &            &              &  $(48.29)$             &    \\ \hline\hline
 $G_2$          &  $49:9$    &  $400.49$    &  $406.47$              & $406.41 $     \\ [1.0ex]
                & $185:41$   &  $[23.36]$   &  $[41.83]$             & $[36.22 ]$  \\ [1.0ex]
                &            &              &  $(43.49)$             &            \\ \hline
  $G_{21}\subset G_2$ &$17:3$    &  $578.90$    &  $589.51$           & $589.74 $ \\ [1.0ex]
                    & $39:15$   &  $[35.48]$   &  $[85.68]$          & $[67.16]$  \\ [1.0ex]
                &            &                 &  $(94.31)$          &            \\ \hline\hline
 $G_3$          &  $73:18$   &  $279.75$    &  $282.46$              & $282.40$  \\ [1.0ex]
                & $250:66$   &  $[15.41]$   &  $[21.38]$             & $[20.14 ] $ \\[1.0ex]
                &            &              &  $(16.56)$             & \\ \hline\hline
\end{tabular}
\end{center}
\textbf{Standard errors:} $[-]$ (Taylor, g-weights) and $(-)$ (external model)
\end{table}
\noindent \textbf{Remark:}\\
\noindent The mean residual for small area $G_1$ was $-1.00$ for the extended model instead of $0$ because $I_{c,G}(x) \not\equiv 1$.

\newpage
\begin{table}[h]
\begin{center}
\caption{ \label{newresult3}\textbf{Two-phase combined estimates}}
\end{center}
\begin{center}
 \begin{tabular}{|l|c|c|c|}
   \textbf{Domain} & \textbf{sample sizes} & \textbf{basal area}  & \textbf{stem density}\\[1.0ex]
 $ $   & $n_1:n_2$    &  & \ \\[1.0ex]
 $ $   & $n_1\bar{M}_1:n_2\bar{M}_2$  & $ \hat{\tilde{Y}}_{c,G,cpsynth}$ &$\hat{\tilde{Y}}_{c,G,cpsynth}$ \\ \hline
 $F$            & $ 298:73$   &  $31.32 $    &  $325.17$       \\ [1.0ex]
                & $1203:298$  &  $[0.93]$    &  $[12.62]$       \\ \hline \hline
 $G_2$          &  $49:9$     &  $29.39 $    &  $407.84$        \\ [1.0ex]
                &  $185:41$   &  $[2.23]$    &  $[39.04]$       \\ \hline \hline
 $G_3$          &  $73:18$     &  $31.57 $    &  $284.17$         \\ [1.0ex]
                & $250:66$    &  $[2.09]$    &  $[18.09]$       \\ \hline \hline
 $G_4$          &  $81:17$     &  $27.77 $    &  $274.59$        \\ [1.0ex]
                & $306:69$    &  $[1.99]$   &  $[23.49]$        \\ \hline \hline
 $G_5$          &  $125:29$    &  $34.31$    &  $347.76$          \\ [1.0ex]
                & $462:122$   &  $[1.24]$   &  $[16.61]$         \\ \hline \hline
\end{tabular}
\end{center}
\textbf{Standard errors:} $[-]$ (Taylor, g-weights).
\end{table}
\noindent \textbf{Remarks}:\\
The classical estimates $\hat{Y}_{c,G,psmall}$ for $G_4$ were: $27.78(2.00)$ for basal area and $274.79(24.12)$ for stem density. For $G_5$  the corresponding results were $34.33(1.35)$ and $347.89(17.49)$.\\
As expected on mathematical grounds all extended models yielded empirical means of the residuals over the entire domain and over one or many small areas which were equal to zero ($<10^{-12})$.

The standard error for $\hat{Y}_{c,G,small}$ are given within $(-)$ when considering the internal model as an external one, i.e. by using the formulae (see \cite{mandallaz} p. 87):
\begin{equation*}
\hat{Y}_{c,G,psmall}=\frac{\sum_{x\in{s_{1,G}}}M(x)\hat{Y}_{c}(x)}{\sum_{x\in{s_{1,G}}}M(x)}
+\frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_{c}(x)}{\sum_{x\in{s_{2,G}}}M(x)}
\end{equation*}
\begin{eqnarray*}
\widehat{\VAR}(\widehat{Y}_{c,G,psmall})&=&\left(1-\frac{n_{2,G}}{n_{1,G}}\right)
\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}
\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}\right)^{2}(\hat{R}_c(x)-\hat{\bar{R}}_{2,G})^2\\
&+&
\frac{1}{n_{1,G}}\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}
\right)^{2}(Y_c(x)-\hat{Y}_{2,G})^2
\end{eqnarray*}
The standard errors given in $[-]$ refer to the various Taylor expansions given in [\ref{estvarpseudosynth1cluster}], [\ref{estvarpsmallcluster}], [\ref{estvarpsynth1clusterextended}] or their equivalent g-weights versions.\\

The extended model for the
$\hat{\tilde{Y}}_{c,G_k,psynt}$ contain only the indicator variable of the corresponding small area $G_k$. For this reason the corresponding estimates for the entire domain $F$ are given for $G_1$ $G_2$ and $G_3$ separately (in this order).\\

\end{document}

### 2017-07-28 JStatSoft_forestinventory_master.tex
second change
graphics modified

@@
 		\centering
-    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
+    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(2).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_a}
@@
 		\centering
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
+   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(2).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_b}

### 2017-07-30 JStatSoft_forestinventory_master.tex
update Andreas 30.07.2017

@@
 The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
+The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
 These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
 The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ such as the timber volume density per hectare.
-The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function.
+the local density $Y(x)$ (e.g.
+the timber volume density per hectare).
+The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase}-function.
 Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
+Estimations solely based on the terrestrial sample $s_2$ can be applied by the \code{onephase} function in \pkg{forestinventory} and are frequently used to access the gain in efficiency of the mutliphase estimation procedures.
 
@@
 
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
+The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (figure \ref{fig:concmphase_and_sae_a}).
 The basic assumption is that auxiliary information is available in two different frequencies, i.e.
 sample sizes.
-The phase $s_0$ provides
+The phase $s_0$ provides a large number $n_0$ of auxiliary information, whereas the phase $s_1$ provides additional auxiliary information that is only available at $n_1$ subsamples of $s_0$.
+The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and $s_0$.
+The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(2)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
+From that it follows that we can define \textit{two nested} regression models.
+The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(1)t}(x),\pmb{Z}^{(2)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$.
+The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
+Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(2)}(x)}$ at the sample locations $s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{2}(x)$.
+If the gain in model precision from the reduced to the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
+In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase}-function.
 
-A first phase of auxiliary information (e.g.
-taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares.
-A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions.
-A further subsample produces a third final phase based on terrestrial observations (i.e.
-the local densities of the ground truth) and is used to correct for bias in the design-based sense.
-The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e.
-a census).
-Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.
+% NOTE:
+% We still have to explain the difference between "non-exhausitve" and "exhaustive" auxiliary information (evtl. tricky)
+% and the difference between cluster and no-cluster sampling
+% --> I would not mentioned it here (general methods), but rather go into detail when describing the estimtors (next section)
 
-Explain three phase sampling, giving refrences to existing literature and studies
+\subsection{Small Area Estimation}
+
+We enter the topic of small area estimation if we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (figure \ref{fig:concmphase_and_sae_b}).
+If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be to small to provide sufficient estimation accuracy.
+In this case, the application of multiphase estimation procedures can be efficient.
+However, the terrestrial sample size is too small to fit a separate regression model to them.
+The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and then to apply this  model to the small area.
+A potential model bias within $G$ is then corrected for by using the model residuals derived in the small area $G$.
+If no terrestrial samples fall into a small area $G$ (i.e.
+$n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
+Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase} and \code{threephase}-function in the \proglang{R} package \pkg{forestinventory}.
+
+% NOTES:
+% - borrowing strength
+% - both, two-and three-phase can be extended to sae estimators
+% - explain difference between synthetic and non-sysnthetic estimators
+% - soundly derive reasoning, starting with onephase, then twophase
 
@@
 
+\subsection{Design-based vs. model-dependent approach}
+
+The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories.
+However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest).
+As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e.
+the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable.
+Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process \cite{maltamo2014}.
+In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e.
+the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and $not$ the result of a stochastic process.
+The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism.
+A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
+identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.
+% NOTES: 
+% - we can still describe it a bit better, also with some citations of mcroberts and breidenbach (their md-description)
+
 % Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
@@
 
-\subsection{Small Area Estimation}
+\subsection{Package Structure}
 
-Explain:
-\begin{itemize}
-  \item What is the difference between \textit{global} and \textit{small area} estimation?
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{theephase()} function.
+From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (fgure \ref{fig:struct_package}).
+The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.
+The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global} estimations) or only for subregions of the entire inventory area (\textbf{small area} estimations).
+The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area.
+The first case allows for computing the exact mean $\bar{\pmb{Z}}$ of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$).
+In the second case, the mean of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$) can only be estimated by $\hat{\bar{\pmb{Z}}}$ which adds another component of uncertainty to the respective estimators.
+The package offers three small area estimators for both the exhaustive and non-exhaustive setup.
+The \textit{extended small area estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
+The resulting point estimates and variances are asymptotically design-unbiased.
+If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
+All estimators are also available for \textit{cluster} sampling, where a sample unit conists of multiple, spatially agglomerated samples.
+The implementation and application of the estimators are described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
+
+% NOTES: 
+% - how is the package structured in general? --> likewise a decision tree
+% - for simple and cluster sampling (briefly describe cluster sampling and its specialties)
+% - for non-exhausitve and exhaustive auxiliary information
 
@@
 
-Hallo. Here comes a first try:
-
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
-
-This was the result
+% Hallo. Here comes a first try:
+% 
+% <<>>=
+% library(forestinventory)
+% op <- onephase(formula = tvol~1 ,data = grisons,
+%               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
+% summary(op)
+% @
+% 
+% This was the result
+% 
+% 
+%

### 2017-08-03 JStatSoft_forestinventory_master.rnw
update Andreas 30.07.2017
update Andreas 03.08.2017

@@
 %% almost as usual
-\author{Andreas Hill \\ETH Z\"urich \And 
-        Alexander Massey}
+\author{Andreas Hill \And Alexander Massey}
 \title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
@@
 
-\SweaveInput{rnw/Global Estimators and their Application in R (2).Rnw}
+\SweaveInput{rnw/Global Estimators and their Application(2).Rnw}
 
-\SweaveInput{rnw/SAE Estimators and their Application in R.Rnw}
+\SweaveInput{rnw/SAE Estimators and their Application.Rnw}
 
@@
 % NOTES:
-% - Error handling and adjustments: Sampling designs not nested
+% - Error handling and adjustments: Sampling designs not nested (always give background and then illustrate handling in package)
 % - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
@@
 
-Hallo. Here comes a first try:
-
-<<>>=
-library(forestinventory)
-op <- onephase(formula = tvol~1 ,data = grisons,
-              phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-summary(op)
-@
-
-This was the result
+% Hallo. Here comes a first try:
+% 
+% <<>>=
+% library(forestinventory)
+% op <- onephase(formula = tvol~1 ,data = grisons,
+%               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
+% summary(op)
+% @
+% 
+% This was the result
+% 
+% 
+%

### 2017-08-03 rnw/Global Estimators and their Application(2).Rnw
update Andreas 03.08.2017

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section[Global Estimators and their Application]{Global Estimators and their Application}
\label{sec:globest_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Double Sampling (Two-Phase) Estimators}

% NOTES:
% - The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation})
%   by the \code{onephase} function in \pkg{forestinventory}. This function can also be used to access the gain in efficiency of the mutliphase estimation procedures. 
%
% - we only use the non-cluster sampling here to illustrate:
% a) reg-coef calculation, point estimates and two variances, important components
% b) we introduce the boundary adjustments: explain why, give mathematical formula, mention that this is optional and can be left out, in which case the simple means of Z are used
% c) exhaustive and non-exhaustive

% Here, we give the \textbf{mathematical background} of:
% \begin{itemize}
%   \item the classical two-phase estimator, including:
%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
%   \item ...
% \end{itemize}
% 
%

% 
% Here, we give the \textbf{application} examples:
% \begin{itemize}
%   \item example for non-exhaustive case: use grison dataset
%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
%   \item example for exhaustive case: use cluster set
% \end{itemize}

\subsubsection{Mathematical Background}

The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.

\begin{equation}\label{normequ_simple}
  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
\end{equation}

This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables. Based on the regression coefficients, the prediction for the sample location $x$ is calcualted as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations. This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data. \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g. from previous studies. However, a respective function in the package is up to now non existing and will be subject to future extentions.\par

The design-based variance-covariance matrix of the regression coefficients $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ is then calculated as

\begin{equation}\label{eq:estvarmatrix}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
\end{equation}

The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables. Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information. Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios. In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$. The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}. In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}). This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators. In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}). Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}). As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par

An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g. by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}). This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy. To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}. The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.

\begin{equation}\label{eq:wmeanZ}
  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
\end{equation}

The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively. Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e. \textbf{internal} model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.

\begin{subequations}\label{eq:pointest_2p_reg}
\begin{align}
  \hat{Y}_{2preg,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
  \hat{Y}_{2preg,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
\end{align}
\end{subequations}

The package calculates two kind of variances for the point estimate. The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}). Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.

\begin{subequations}\label{eq:gw_var_2p_reg}
\begin{align}
  \hat{\var}(\hat{Y}_{2preg,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
  \hat{\var}(\hat{Y}_{2preg,nex}) & :=
  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
\end{align}
\end{subequations}

where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
\end{equation}

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}(\hat{Y}_{2preg,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}(\hat{Y}_{2preg,nex}) & =
 \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
\end{suequations}

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of using exhaustive (\textbf{(a)}) and non-exhaustive (\textbf{(b)}) auxiliary information including boundary adjustment. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (underlying polygon). w: weights used for calculating the weighted mean of each explanatory variable. Dotted circles indicate terrestrial sample plots.}
\label{fig:exh_nexh_bw}
\end{figure}

\subsubsection{Application}

To demonstrate the use of the global two-phase estimators, we can use the \code{grisons} data frame that comes with installing the package from CRAN. The dataset contains data from a two-phase forest inventory under simple sampling design and comprises observations from 306 sample locations. Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}). For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume (\code{tvol}) is provided from a terrestrial survey. We can load the data frame in the \proglang{R} environment by executing:

\begin{small}
<<>>=
data("grisons", package = "forestinventory")
@
\end{small}

The non-exhaustive estimatior with boundary correction can be applied as follows:

\begin{small}
<<>>=
reg_nex<- twophase(formula=tvol ~ mean + stddev + max + q75,
                   data=grisons,
                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
                   boundary_weights = "boundary_weights")
@
\end{small}

We can then get a summary of the estimation results by the \code{summary.twophase} function provided by the package:

\begin{small}
<<>>=
summary(reg_nex)
@
\end{small}

\subsection{Triple Sampling (Three-Phase) Estimators}

\subsubsection{Mathematical Background}

\subsubsection{Application}

% The psynth estimator:
% 
% \begin{equation}\label{pseudosynth1}
% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
% \end{equation}
% 
% \begin{equation}\label{estvarpseudosynth1}
% \hat{\var}(\hat{Y}_{G,psynth}): =
% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
% \end{equation}

### 2017-08-03 rnw/Global Estimators and their Application(2)_backup.Rnw
update Andreas 03.08.2017

--- 2017-08-03 rnw/Global Estimators and their Application(2).Rnw
@@
 \label{sec:globest_and_appl}
+
+This section describes the mathematical details and the application of the global multiphase estimators that are implemented in the \proglang{R} package \pkg{forestinventory}.
+While \cite{mandallaz2008, mandallaz2013c} provides an extensive derivation of all estimators, we will here concentrate on the mathematical formulas that are actually implemented and applied in the package, as well as on some modifications.
+For the sake of brevity, we will limited to simple sampling.
+Special cases occuring under cluster sampling are described in section XXX.
 
@@
 
-The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.
+The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients is derived as
 
@@
 Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios.
-In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}.
-In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
-This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators.
-In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
-Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}).
-As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par
+In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$ (points).
+The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in \ref{fig:exh_nexh_bw}.
+In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area \ref{fig:exh_nexh_and_boundweights_a}.
+This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}$ and for using this information in the estimators.
+In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations \ref{fig:exh_nexh_and_boundweights_b}.
+Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators.
+As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations and the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par
 
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
-In forest inventories, the sample is often restricted to those sample locations located within the forest area.
-In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
-by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}).
-This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
-To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}.
-The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.
+An extension to the so-far published estimators by Mandallaz is to consider a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
+...
 
-\begin{equation}\label{eq:wmeanZ}
-  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
+The \textbf{point estimate} for is then calculated according to equation \ref{eq:pointest_simple}.
+Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
+In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
+
+\begin{equation}\label{eq:pointest_simple}
+\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
 \end{equation}
 
-The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e.
-\textbf{internal} model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.
+The package offers the calculation of two kind of variances for the point estimate.
+The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_simple}.
+Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
+The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
+Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
 
-\begin{subequations}\label{eq:pointest_2p_reg}
-\begin{align}
-  \hat{Y}_{2preg,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
-  \hat{Y}_{2preg,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
-\end{align}
-\end{subequations}
-
-The package calculates two kind of variances for the point estimate.
-The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}).
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.
-
-\begin{subequations}\label{eq:gw_var_2p_reg}
-\begin{align}
-  \hat{\var}(\hat{Y}_{2preg,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
-  \hat{\var}(\hat{Y}_{2preg,nex}) & :=
-  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
-\end{align}
-\end{subequations}
-
-where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
-\begin{equation}\label{estvarcovaux}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
-\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
+\begin{equation}\label{eq:gw_var_simple}
+\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
 \end{equation}
 
-The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{subequations}\label{eq:varexternal_2p_reg}
-\begin{align}
-  \hat{\var}(\hat{Y}_{2preg,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
-  \hat{\var}(\hat{Y}_{2preg,nex}) & =
+\begin{equation}\label{eq:varexternalsimple}
+\hat{\var}(\hat{Y}_{reg})=
 \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
-\end{suequations}
+(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
+\end{equation}
 
@@
 
-To demonstrate the use of the global two-phase estimators, we can use the \code{grisons} data frame that comes with installing the package from CRAN.
-The dataset contains data from a two-phase forest inventory under simple sampling design and comprises observations from 306 sample locations.
-Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
-For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume (\code{tvol}) is provided from a terrestrial survey.
-We can load the data frame in the \proglang{R} environment by executing:
-
-\begin{small}
-<<>>=
-data("grisons", package = "forestinventory")
-@
-\end{small}
-
-The non-exhaustive estimatior with boundary correction can be applied as follows:
-
-\begin{small}
-<<>>=
-reg_nex<- twophase(formula=tvol ~ mean + stddev + max + q75,
-                   data=grisons,
-                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                   boundary_weights = "boundary_weights")
-@
-\end{small}
-
-We can then get a summary of the estimation results by the \code{summary.twophase} function provided by the package:
-
-\begin{small}
-<<>>=
-summary(reg_nex)
-@
-\end{small}
-
 \subsection{Triple Sampling (Three-Phase) Estimators}

### 2017-08-03 rnw/Main structure and modules of the Package.Rnw
update Andreas 03.08.2017

@@
 
-%\section{Structure of the Package}
+%\section{Methods and Structure of the Package}
 \section{Methods and Structure of the Package}
@@
 
-%Here comes the general structure of the package, or at least of the two and three-phase functions
+% NOTE: 
+% The aim of this section is to build the basic understanding of two and threephase sampling, and the extention to small area estimation. We introduce the general terminology and show
+% how the package is structured to apply the methods
 
@@
 Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
-Estimations solely based on the terrestrial sample $s_2$ can be applied by the \code{onephase} function in \pkg{forestinventory} and are frequently used to access the gain in efficiency of the mutliphase estimation procedures.
 
@@
 % and the difference between cluster and no-cluster sampling
-% --> I would not mentioned it here (general methods), but rather go into detail when describing the estimtors (next section)
+% --> I would not mentioned it here (keep  it general), but rather go into detail when describing the estimtors (next sections)
 
@@
 In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{theephase()} function.
-From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (fgure \ref{fig:struct_package}).
+From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}).
 The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.
@@
 The package offers three small area estimators for both the exhaustive and non-exhaustive setup.
-The \textit{extended small area estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
+The \textit{extended small area estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
 The resulting point estimates and variances are asymptotically design-unbiased.
 If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
-All estimators are also available for \textit{cluster} sampling, where a sample unit conists of multiple, spatially agglomerated samples.
-The implementation and application of the estimators are described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
+The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation}) by the \code{onephase} function in \pkg{forestinventory}.
+This function can also be used to access the gain in efficiency of the mutliphase estimation procedures.
+All estimators are also available for \textit{cluster} sampling, in which case a sample unit conists of multiple, spatially agglomerated samples.
+The implementation and application of the estimators are described in detail in the following sections.
 
@@
 \centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
+\resizebox{1\hsize}{!}{\includegraphics{fig/packagestructure.PNG}}
 \caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
@@
 \end{figure}
+
+The following sections describe the mathematical details and the application of the multiphase estimators that are implemented in the \proglang{R} package \pkg{forestinventory}.
+While \cite{mandallaz2008, mandallaz2013c} provides an extensive derivation of all estimators, we will concentrate on the mathematical formulas that are actually implemented and applied in the package functions.
+For the sake of brevity, we will limited to simple sampling.
+Special cases occuring under cluster sampling are described in section X.
+
+\newpage

### 2017-08-08 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 09.08.2017

--- 2017-07-28 JStatSoft_forestinventory_master.tex
@@
-\documentclass[article]{jss}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% almost as usual
-\author{Andreas Hill \\ETH Z\"urich \And 
-        Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{
-Forest inventories provide reliable and evidence based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient.
-The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}
-
-\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-  Andreas Hill\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{andreas.hill@usys.ethz.ch}\\
-  Telephone: +41/44/632 32 36\\
-  URL: \url{http://www.lue.ethz.ch/people/hilla}
-
-  Alexander Massey\\
-  Department of Environmental Systems Science\\
-  Chair of Landuse Engineering\\
-  ETH Z\"urich\\
-  Universit\"atstrasse 16\\
-  8092 Z\"urich, Switzerland\\
-  E-mail: \email{afmass@gmail.com}\\}
-}
-
-%% for those who use Sweave please include the following line (with % symbols):
-
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Tex Settings ------------------------------------------ %
-
-% used packages
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathptmx} 
-\usepackage{latexsym}
-\usepackage{a4}
-\usepackage{graphicx}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{flafter}
-\usepackage{bm}
-\usepackage{setspace}
-\usepackage{lineno}
-\usepackage{natbib}
-\usepackage{geometry}
- \geometry{
- a4paper,
- left=28mm,
- bottom=28mm
- }
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage[font=footnotesize]{subcaption}
-
-% new commands
-\newcommand{\bwi}{BWI3}
-\newcommand{\LF}{\ensuremath{\lambda(F)}}
-\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
-\newcommand{\EX}{\mathbb{E}}
-\newcommand{\PR}{\mathbb{P}}
-\newcommand{\var}{\mathbb{V}}
-\newcommand{\COV}{\mathbb{COV}}
-\newcommand{\MAV}{\mathbb{MAV}}
-\newcommand{\MRAV}{\mathbb{MRAV}}
-\newcommand{\POP}{\mathcal{P}}
-\newcommand{\SAMP}{\mathcal{S}}
-\newcommand{\RE}{\mathbb{RE}}
-\newcommand{\PLAN}{\Re^2}
-\newcommand{\SUR}{\mathbb{S}}
-\newcommand{\ING}{\mathbb{I}}
-\newcommand{\DEP}{\mathbb{D}}
-
-%\setlength{\parindent}{1em}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- Main Document------------------------------------------ %
-
-\begin{document}
-
-%------------------------------------------------------------------------------------------------%
-% -------------------------------------- R Settings -------------------------------------------- %
-
-% set global options for R-output:
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Introduction ---------------------------------------------- %
-
 % !Rnw root = JStatSoft_forestinventory_master.Rnw
 
-\section{Introdution}
-\label{sec:intro}
-
-In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods.
-They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
-information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
-However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par
-
-Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multiphase sampling}.
-The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
-This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
-Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
-
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
-The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
-The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
-
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
-
-% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-% Explain the problem that leads to sae estimations (in generell and in forestry)
-%
-%
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-%------------------------------------------------------------------------------------------------%
-% --------------------------- Structure of the Package ----------------------------------------- %
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-%\section{Structure of the Package}
-\section{Methods and Structure of the Package}
-\label{sec:str_and_mod}
-
-%Here comes the general structure of the package, or at least of the two and three-phase functions
-
-\subsection{Two-Phase Sampling}
-
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
-These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ such as the timber volume density per hectare.
-The set of explanatory variables at each sample location $x \in s_1$ is now transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied by the \code{twophase}-function.
-Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
-
-\subsection{Three-Phase Sampling}
-
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}).
-The basic assumption is that auxiliary information is available in two different sample sizes.
-
-A first phase of auxiliary information (e.g.
-taken from remote sensing data) is used to generate model predictions based on multiple linear regression using the method of ordinary least squares.
-A subsample of the first phase comprises a second phase which contains further auxiliary information that produces another set of model predictions.
-A further subsample produces a third final phase based on terrestrial observations (i.e.
-the local densities of the ground truth) and is used to correct for bias in the design-based sense.
-The estimation method is available for simple and cluster sampling and includes the special case where the first phase is based on an exhaustive sample (i.e.
-a census).
-Small-area applications are supported for synthetic estimation as well as two varieties of bias-corrected estimators: the traditional small-area estimator and an asymptotically equivalent version derived under Mandallaz's extended model approach.
-
-Explain three phase sampling, giving refrences to existing literature and studies
-
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-    \resizebox{1.1\hsize}{!}{\includegraphics{fig/multiphase_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth} 
-		\centering
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/sae_graphic.eps}}%
-		\caption{} \label{fig:concmphase_and_sae_b}
-	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
-The square represents the forest area for which an inventory is being conducted.
-The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
-\label{fig:concmphase_and_sae}
-\end{figure}
-
-% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
-The auxiliary information can be any information that exhibits a correlation to the local density, i.e.
-target variable, $Y(x)$.
-% 
-% e.g. past inventory data \citep{massey2014a}
-
-\subsection{Small Area Estimation}
-
-Explain:
-\begin{itemize}
-  \item What is the difference between \textit{global} and \textit{small area} estimation?
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-\begin{figure}[htb]
-\centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure.eps}}
-\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
-\label{fig:struct_package}
-\end{figure}
-
-% NOTES:
-% - Pyramide / decision tree Graphik
-% - Confidence Interval calculation
-% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
-%
-% --> explain each point briefly without to much (mathematical) detail
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Estimators and their Application -------------------------- %
-
-%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Global Estimators and their Application in R]{Global Estimators and their Application}
-\label{sec:globest_and_appl}
+\section{Threephase Estimators and their Application}
+\label{sec:threephase_and_appl}
 
 % ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
-
-% Note: As the mathematical form of estimators is still quite simple and short, we can here
-%       focus on presenting and explaining important 'components' of the estimators, such as Z, C, ...)
-% 
-% \subsection{Mathematical Background}
-% 
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-% 
-% 
-% \subsubsection{Application}
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
-
-\subsubsection{Mathematical Background}
-
-Mention:
-\begin{itemize}
-  \item What is the difference between model-dependent and design-based? (bias-correction, dont have to believe in model predictions.
-  \item explain the structure of the package (graphic) and state that we will only concentrate on specific cases ...
-\end{itemize}
-
-The regression coefficients of the OLS regression model are found by solving the sample-based normal equation. In case of \textbf{simple sampling}, the vector of regression coefficients are derived as
-
-\begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
-\end{equation}
-
-The design-based variance-covariance matrix of the regression coefficients is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} 
-\end{equation}
-
-with the empirical residuals, i.e. the regression model residuals, available at all sample location $x \in s_2$ being
-\begin{equation}\label{eq:resids}
-  \hat{R}(x)=Y(x)-\hat{Y}(x)
-\end{equation}
-
-The \textbf{point estimate} for simpe sampling is calculated according to equation \ref{eq:pointest_simple}.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (\textbf{internal} regression model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by definition and does not have to be added as is necessary for \textbf{external} models.
-Since in the package \pkg{forestinventory} only allows for internal models, the equation simplifies to
-
-\begin{equation}\label{eq:pointest_simple}
-\hat{Y}_{reg}=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2}
-\end{equation}
-
-The estimation precision of the point estimate is specified by the estimated \textbf{design-based variance} as given in equation \ref{eq:gw_var_simple}.
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variance given in \citep{mandallaz2016}.
-The package \pkg{forestinventory} additionally provides the \textbf{external} variance (equation \ref{eq:varexternalsimple}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{equation}\label{eq:gw_var_simple}
-\hat{\var}(\hat{Y}_{reg})=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-\end{equation}
-
-\begin{equation}\label{eq:varexternalsimple}
-\hat{\var}(\hat{Y}_{reg})=
-\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2
-\end{equation}
-
-\subsubsection{Application}
-
-\subsection{Triple Sampling (Three-Phase) Estimators}
-
-\subsubsection{Mathematical Background}
-
-\subsubsection{Application}
-
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
-
-% !Rnw root = JStatSoft_forestinventory_master.Rnw
-
-\section[Small Area Estimators and their Application in R]{Small Area Estimators and their Application}
-\label{sec:saeest_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Double Sampling (Two-Phase) Estimators}
+\subsection{Global Estimators}
 
@@
 
-\subsection{Triple Sampling (Three-Phase) Estimators}
+\subsection{Small Area Estimators}
 
@@
 % \end{equation}
-
-% NOTES:
-% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
-% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
-%
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Special Cases and Scenarios ------------------------------- %
-
-% NOTES:
-% - Error handling and adjustments: Sampling designs not nested
-% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
-%                   or s2 or/and s1 sample in 3-phase case
-% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
-
-% NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
-
-%------------------------------------------------------------------------------------------------%
-% ----------------------------- Visualization and Analysis of Estimation Results --------------- %
-
-% NOTES:
-% - demonstrate visualization of errors and point-ests + CIs
-% - demonstrate a) releff-calculation and gain-analysis
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------------- Literature ------------------------------------------ %
-
-\bibliography{bib/literature}\newpage\cleardoublepage
-
-% -------------------------------------------------------------------------- %
-\end{document}
-
-Hallo. Here comes a first try:
-
-\begin{Schunk}
-\begin{Sinput}
-R> library(forestinventory)
-R> op <- onephase(formula = tvol~1 ,data = grisons,
- +               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-R> summary(op)
-\end{Sinput}
-\begin{Soutput}
-One-phase estimation
-
-Call: 
-onephase(formula = tvol ~ 1, data = grisons, phase_id = list(phase.col = "phase_id_2p", 
-    terrgrid.id = 2))
-
-Method used:
-One-phase estimator
-
-Estimation results:
- estimate variance n2
- 399.4321 567.2001 67
-\end{Soutput}
-\end{Schunk}
-
-This was the result

### 2017-08-08 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 09.08.2017

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Twophase Estimators and their Application}
\label{sec:twophase_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Global Estimators}

% NOTES:
% - The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation})
%   by the \code{onephase} function in \pkg{forestinventory}. This function can also be used to access the gain in efficiency of the mutliphase estimation procedures. 
%
% - we only use the non-cluster sampling here to illustrate:
% a) reg-coef calculation, point estimates and two variances, important components
% b) we introduce the boundary adjustments: explain why, give mathematical formula, mention that this is optional and can be left out, in which case the simple means of Z are used
% c) exhaustive and non-exhaustive

% Here, we give the \textbf{mathematical background} of:
% \begin{itemize}
%   \item the classical two-phase estimator, including:
%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
%   \item ...
% \end{itemize}
% 
%

% 
% Here, we give the \textbf{application} examples:
% \begin{itemize}
%   \item example for non-exhaustive case: use grison dataset
%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
%   \item example for exhaustive case: use cluster set
% \end{itemize}

\subsubsection{Mathematical Background}

The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.

\begin{equation}\label{normequ_simple}
  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
\end{equation}

This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables. Based on the regression coefficients, the prediction for the sample location $x$ is calcualted as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations. This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data. \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g. from previous studies. However, a respective function in the package is up to now non existing and will be subject to future extentions.\par

The design-based variance-covariance matrix of the regression coefficients $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ is then calculated as

\begin{equation}\label{eq:estvarmatrix}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
\end{equation}

The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables. Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information. Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios. In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$. The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}. In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}). This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators. In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}). Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}). As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par

An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g. by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}). This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy. To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}. The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.

\begin{equation}\label{eq:wmeanZ}
  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
\end{equation}

The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively. Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e. \textbf{internal} model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.

\begin{subequations}\label{eq:pointest_2p_reg}
\begin{align}
  \hat{Y}_{reg2p,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
  \hat{Y}_{reg2p,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
\end{align}
\end{subequations}

The package calculates two kind of variances for the point estimate. The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}). Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.

\begin{subequations}\label{eq:gw_var_2p_reg}
\begin{align}
  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
  \hat{\var}(\hat{Y}_{reg2p,nex}) & :=
  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
\end{align}
\end{subequations}

where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
\end{equation}

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & =
 \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
\end{align}
\end{subequations}

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of using exhaustive (\textbf{(a)}) and non-exhaustive (\textbf{(b)}) auxiliary information including boundary adjustment. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (underlying polygon). w: weights used for calculating the weighted mean of each explanatory variable. Dotted circles indicate terrestrial sample plots.}
\label{fig:exh_nexh_bw}
\end{figure}

\subsubsection{Application}

To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository. The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample). Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}). For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey ($s_2$ sample). We can load the data frame in the \proglang{R} environment by executing:

\begin{small}
<<>>=
data("grisons", package = "forestinventory")
@
\end{small}

Note that the structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package. Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location. Note that for global two-phase estimation, the data frame has to comprise a column that indicates whether a row corresponds to a $s_1$ or $s_2$ sample location (in \code{grisons}, this is the column \code{phase_id_2p}).\par

In order to apply the \textbf{non-exhaustive two-phase estimation with boundary weight adjustment} to the \code{grisons} data, we have to specify a) the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}, b) the data frame containing the inventory information (\code{data}), c) the name of the column specifying to which sample ($s_1$, $s_2$) an observation belongs (\code{phase.col}) as well as the indicator for the $s_2$, i.e. terrestrial sample (\code{terrgrid.id}), and d) the columname of the column containing the weights $w(x)$ of the boundary adjustments. The \textbf{non-exhaustive estimator with boundary weight adjustment} can thus be applied as follows:
\begin{small}
<<>>=
reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
                   data=grisons,
                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
                   boundary_weights = "boundary_weights")
@
\end{small}
The \code{twophase}-function creates an object of class \code{twophase}. A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object. It first provides a summary of the function call, the applied regression formula and the estimation method that was applied. The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}). Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model.
\begin{small}
<<>>=
summary(reg2p_nex)
@
\end{small}
If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}. The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1. We assume that in this case, the boundary adjustment has been applied when calculating the exact mean of $\pmb{Z}$.

The right order can easily be established by extracting the columnames of the respective \code{lm} output. This step is particularly recommended if categorical variables are used. In our case this is:
\begin{small}
<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
@
\end{small}
The \textbf{exhaustive estimator} can thus be applied after creating the vector of exact means $\bar{\pmb{Z}}$:
\begin{small}
<<>>=
true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)

reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
                    data = grisons,
                    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                    exhaustive = true.means.Z)
@
\end{small}

We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
\begin{small}
<<>>=
reg2p_ex$estimation
@
\end{small}
Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is because we took out one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

%--------------------------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------------------------%

\subsection{Small Area Estimators}

\subsubsection{Mathematical Background}

The package \pkg{forestinventory} offers 3 small area estimators. A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible. An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which is always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par

%----------------------------------------------- %
% \textbf{Small and Pseudo Small Area Estimator}\par

We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}). Their point estimates are given in equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}. As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables available at the sample location $s_{1,G}$ in the small area. A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term. We call this the \textit{bias} or \textit{residual correction} term.

\begin{subequations}\label{eq:pest_2p_small_psmall}
\begin{align}
  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_reg_small} \\
  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_reg_psmall}
\end{align}
\end{subequations}

The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textbf{small area estimator} and the \textbf{pseudo small area estimator}. Again note that all components are restricted those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.

\begin{subequations}\label{eq:var_2p_small_psmall}
\begin{align}
  \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G 
    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
  \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2} 
  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
\end{align}
\end{subequations}

\begin{subequations}\label{eq:varext_2p_small_psmall}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\ 
 &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
\end{align}
\end{subequations}

The the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
\begin{equation}\label{estvarcovaux_G}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
\end{equation}

%------------------------------------------------------------- %
% \textbf{Synthetic and Pseudo Synthetic Estimator}\par

The \textbf{synthetic} and \textbf{pseudo synthetic estimator} is applied if no terrestrial samples are available within the small area $G$ (i.e. $n_{2,G}=0$). In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$. However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}) is no longer possible. Thus, the synthetic / pseudo synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e. $\int_G R(x)dx=0$. If this assumption is not fullfilled, the synthetic point estimate $\hat{Y}_{G,synth,2p}$ and the pseudo synthetic point estimate $\hat{Y}_{G,psynth,2p}$ will have a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$. Also note that residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}). Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias. Due to the absence of available residuals in $G$, there is also no external variance form for synthetic estimations.

\begin{subequations}\label{eq:pest_2p_synth_psynth}
\begin{align}
  \hat{Y}_{G,synth,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth}
\end{align}
\end{subequations}

\begin{subequations}\label{eq:var_2p_synth_psynth}
\begin{align}
  \hat{\var}(\hat{Y}_{G,synth,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
  \hat{\var}(\hat{Y}_{G,psynth,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
\end{align}
\end{subequations}

%------------------------------------------------------------- %
% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par

The \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator}

% - explain extended estimator
% - mention boundary weight adjustment likewise global estimation

\newpage

%------------------------------------------------------------- %
\subsubsection{Application}

Small area estimations in the package \pkg{forestinventory} can be applied by \textit{additionally} specifying the optional argument \code{small_area}. The input dataset has therefore to be extended by an additional column that specifies for each each row (i.e. observation of a sample location) the respective small area in which it is located. The argument \code{small_area} requires a \code{list}-object that comprises 1) the name of the column specifiying the small area of each observation (\code{sa.col}), 2) a vector specifying the small area(s) for which estimations are desired (\code{areas}), and 3) the argument \code{unbiased} that controls which of the three available estimators should be applied. In addition to global estimation, the \code{summary} now also displays the sample sizes in the respective small area (\code{n1G} and \code{n2G}).

In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}), we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
\begin{small}
<<>>=
psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      boundary_weights = "boundary_weights",
                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                        unbiased = TRUE),
                      psmall = TRUE)
summary(psmall_2p)
@
\end{small}

The \textbf{extended pseudo small area estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
\begin{small}
<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                         phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                         boundary_weights = "boundary_weights",
                         small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                           unbiased = TRUE))
extpsynth_2p$estimation
@
\end{small}

Note that the point estimates and the design-based and external variances of both estimators are very similiar. However, the R$^2$s (\code{r.squared}) under the \textbf{extended pseudo small area estimator} are changing with the small area, while they are identical under the \textbf{pseudo small area estimator}. This is because under the \textbf{extended pseudo small area estimator}, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices. In case of the \textbf{pseudo small area estimator}, the regression model stays the same for each small area estimation. Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling.\par

Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas. Note that in the \code{grisons} dataset, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible. Applying the pseudo synthetic estimator equals treating those small areas as if $n_{2,G}=0$ by ignoring the terrestrial information.
\begin{small}
<<>>=
psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      boundary_weights = "boundary_weights",
                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                        unbiased = FALSE))
psynth_2p$estimation
@
\end{small}

In this case, the variance of the pseudo synthetic estimator is almost only half as high as under the \textit{psmall} and \textit{extended psynth} estimator. The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation  \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and XXX). Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.

\newpage

### 2017-08-09 JStatSoft_forestinventory_master.rnw
update Andreas 03.08.2017(2)
update Andreas 09.08.2017
update Andreas 09.08.2017(4)

@@
 
+% execute everything that goes without echo:
+<<echo=FALSE>>=
+library(forestinventory)
+@
+
 %------------------------------------------------------------------------------------------------%
@@
 
-\SweaveInput{rnw/Global Estimators and their Application(2).Rnw}
+\SweaveInput{rnw/Twophase_Estimators and their Application.Rnw}
 
-\SweaveInput{rnw/SAE Estimators and their Application.Rnw}
+\SweaveInput{rnw/Threephase_Estimators and their Application.Rnw}
 
@@
 %------------------------------------------------------------------------------------------------%
+% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
+
+\SweaveInput{rnw/Calculation of Confidence Intervals.Rnw}
+
+% NOTES:
+% - demonstrate calculation of confidence intervals (give the formulas)
+
+%------------------------------------------------------------------------------------------------%
 % ---------------------------------- Special Cases and Scenarios ------------------------------- %
+
+\SweaveInput{rnw/Special Cases and Scenarios.Rnw}
 
@@
 % - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Calculation of Confidence Intervals ----------------------- %
-
-% NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
 

### 2017-08-09 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 09.08.2017(4)

--- 2017-08-08 rnw/Twophase_Estimators and their Application.Rnw
@@
 
-\section{Twophase Estimators and their Application}
-\label{sec:twophase_and_appl}
-
-% ---------------------------------------------------------------------------- %
-\subsection{Global Estimators}
-
-% NOTES:
-% - The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation})
-%   by the \code{onephase} function in \pkg{forestinventory}. This function can also be used to access the gain in efficiency of the mutliphase estimation procedures. 
-%
-% - we only use the non-cluster sampling here to illustrate:
-% a) reg-coef calculation, point estimates and two variances, important components
-% b) we introduce the boundary adjustments: explain why, give mathematical formula, mention that this is optional and can be left out, in which case the simple means of Z are used
-% c) exhaustive and non-exhaustive
-
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-%
-
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
-
-\subsubsection{Mathematical Background}
-
-The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.
-
-\begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
-\end{equation}
-
-This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables.
-Based on the regression coefficients, the prediction for the sample location $x$ is calcualted as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
-Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
-This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data.
-\citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g.
-from previous studies.
-However, a respective function in the package is up to now non existing and will be subject to future extentions.\par
-
-The design-based variance-covariance matrix of the regression coefficients $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-\end{equation}
-
-The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables.
-Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
-Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios.
-In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}.
-In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
-This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators.
-In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
-Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}).
-As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par
-
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
-In forest inventories, the sample is often restricted to those sample locations located within the forest area.
-In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
-by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}).
-This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
-To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}.
-The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.
-
-\begin{equation}\label{eq:wmeanZ}
-  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
-\end{equation}
-
-The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e.
-\textbf{internal} model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.
-
-\begin{subequations}\label{eq:pointest_2p_reg}
-\begin{align}
-  \hat{Y}_{reg2p,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
-  \hat{Y}_{reg2p,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
-\end{align}
-\end{subequations}
-
-The package calculates two kind of variances for the point estimate.
-The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}).
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.
-
-\begin{subequations}\label{eq:gw_var_2p_reg}
-\begin{align}
-  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
-  \hat{\var}(\hat{Y}_{reg2p,nex}) & :=
-  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
-\end{align}
-\end{subequations}
-
-where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
-\begin{equation}\label{estvarcovaux}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
-\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
-\end{equation}
-
-The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
-
-\begin{subequations}\label{eq:varexternal_2p_reg}
-\begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & =
- \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
-\end{align}
-\end{subequations}
-
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-    \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
-		\caption{} \label{fig:exh_nexh_and_boundweights_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-   \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
-		\caption{} \label{fig:exh_nexh_and_boundweights_b}
-	\end{subfigure}
-\caption{Concept of using exhaustive (\textbf{(a)}) and non-exhaustive (\textbf{(b)}) auxiliary information including boundary adjustment.
-Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (underlying polygon).
-w: weights used for calculating the weighted mean of each explanatory variable.
-Dotted circles indicate terrestrial sample plots.}
-\label{fig:exh_nexh_bw}
-\end{figure}
-
-\subsubsection{Application}
-
-To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository.
-The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample).
-Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
-For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey ($s_2$ sample).
-We can load the data frame in the \proglang{R} environment by executing:
-
-\begin{small}
-<<>>=
-data("grisons", package = "forestinventory")
-@
-\end{small}
-
-Note that the structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package.
-Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location.
-Note that for global two-phase estimation, the data frame has to comprise a column that indicates whether a row corresponds to a $s_1$ or $s_2$ sample location (in \code{grisons}, this is the column \code{phase_id_2p}).\par
-
-In order to apply the \textbf{non-exhaustive two-phase estimation with boundary weight adjustment} to the \code{grisons} data, we have to specify a) the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}, b) the data frame containing the inventory information (\code{data}), c) the name of the column specifying to which sample ($s_1$, $s_2$) an observation belongs (\code{phase.col}) as well as the indicator for the $s_2$, i.e.
-terrestrial sample (\code{terrgrid.id}), and d) the columname of the column containing the weights $w(x)$ of the boundary adjustments.
-The \textbf{non-exhaustive estimator with boundary weight adjustment} can thus be applied as follows:
-\begin{small}
-<<>>=
-reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
-                   data=grisons,
-                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                   boundary_weights = "boundary_weights")
-@
-\end{small}
-The \code{twophase}-function creates an object of class \code{twophase}.
-A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object.
-It first provides a summary of the function call, the applied regression formula and the estimation method that was applied.
-The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}).
-Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model.
-\begin{small}
-<<>>=
-summary(reg2p_nex)
-@
-\end{small}
-If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}.
-The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
-We assume that in this case, the boundary adjustment has been applied when calculating the exact mean of $\pmb{Z}$.
-
-The right order can easily be established by extracting the columnames of the respective \code{lm} output. This step is particularly recommended if categorical variables are used. In our case this is:
-\begin{small}
-<<eval=FALSE>>=
-colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
-@
-\end{small}
-The \textbf{exhaustive estimator} can thus be applied after creating the vector of exact means $\bar{\pmb{Z}}$:
-\begin{small}
-<<>>=
-true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
-
-reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
-                    data = grisons,
-                    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                    exhaustive = true.means.Z)
-@
-\end{small}
-
-We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
-\begin{small}
-<<>>=
-reg2p_ex$estimation
-@
-\end{small}
-Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation.
-This is because we took out one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
-
-%--------------------------------------------------------------------------------------------------%
-%--------------------------------------------------------------------------------------------------%
-
-\subsection{Small Area Estimators}
-
-\subsubsection{Mathematical Background}
-
-The package \pkg{forestinventory} offers 3 small area estimators.
-A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible.
-An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which is always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par
-
-%----------------------------------------------- %
-% \textbf{Small and Pseudo Small Area Estimator}\par
-
-We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
-Their point estimates are given in equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}.
-As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables available at the sample location $s_{1,G}$ in the small area.
-A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term.
-We call this the \textit{bias} or \textit{residual correction} term.
-
-\begin{subequations}\label{eq:pest_2p_small_psmall}
-\begin{align}
-  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_reg_small} \\
-  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_reg_psmall}
-\end{align}
-\end{subequations}
-
-The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textbf{small area estimator} and the \textbf{pseudo small area estimator}.
-Again note that all components are restricted those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
-
-\begin{subequations}\label{eq:var_2p_small_psmall}
-\begin{align}
-  \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G 
-    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
-  \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2} 
-  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
-\end{align}
-\end{subequations}
-
-\begin{subequations}\label{eq:varext_2p_small_psmall}
-\begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\ 
- &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
-\end{align}
-\end{subequations}
-
-The the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
-\begin{equation}\label{estvarcovaux_G}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
-\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
-\end{equation}
-
-%------------------------------------------------------------- %
-% \textbf{Synthetic and Pseudo Synthetic Estimator}\par
-
-The \textbf{synthetic} and \textbf{pseudo synthetic estimator} is applied if no terrestrial samples are available within the small area $G$ (i.e.
-$n_{2,G}=0$).
-In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
-However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}) is no longer possible.
-Thus, the synthetic / pseudo synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
-$\int_G R(x)dx=0$.
-If this assumption is not fullfilled, the synthetic point estimate $\hat{Y}_{G,synth,2p}$ and the pseudo synthetic point estimate $\hat{Y}_{G,psynth,2p}$ will have a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
-Also note that residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
-Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias.
-Due to the absence of available residuals in $G$, there is also no external variance form for synthetic estimations.
-
-\begin{subequations}\label{eq:pest_2p_synth_psynth}
-\begin{align}
-  \hat{Y}_{G,synth,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
-  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth}
-\end{align}
-\end{subequations}
-
-\begin{subequations}\label{eq:var_2p_synth_psynth}
-\begin{align}
-  \hat{\var}(\hat{Y}_{G,synth,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
-  \hat{\var}(\hat{Y}_{G,psynth,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
-\end{align}
-\end{subequations}
-
-%------------------------------------------------------------- %
-% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par
-
-The \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator}
-
-% - explain extended estimator
-% - mention boundary weight adjustment likewise global estimation
+\section{Calculation of Confidence Intervals}
+\label{sec:confint_calc}
 
 \newpage
-
-%------------------------------------------------------------- %
-\subsubsection{Application}
-
-Small area estimations in the package \pkg{forestinventory} can be applied by \textit{additionally} specifying the optional argument \code{small_area}.
-The input dataset has therefore to be extended by an additional column that specifies for each each row (i.e.
-observation of a sample location) the respective small area in which it is located.
-The argument \code{small_area} requires a \code{list}-object that comprises 1) the name of the column specifiying the small area of each observation (\code{sa.col}), 2) a vector specifying the small area(s) for which estimations are desired (\code{areas}), and 3) the argument \code{unbiased} that controls which of the three available estimators should be applied.
-In addition to global estimation, the \code{summary} now also displays the sample sizes in the respective small area (\code{n1G} and \code{n2G}).
-
-In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}), we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
-\begin{small}
-<<>>=
-psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights",
-                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                        unbiased = TRUE),
-                      psmall = TRUE)
-summary(psmall_2p)
-@
-\end{small}
-
-The \textbf{extended pseudo small area estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
-\begin{small}
-<<>>=
-extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                         phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                         boundary_weights = "boundary_weights",
-                         small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                           unbiased = TRUE))
-extpsynth_2p$estimation
-@
-\end{small}
-
-Note that the point estimates and the design-based and external variances of both estimators are very similiar.
-However, the R$^2$s (\code{r.squared}) under the \textbf{extended pseudo small area estimator} are changing with the small area, while they are identical under the \textbf{pseudo small area estimator}.
-This is because under the \textbf{extended pseudo small area estimator}, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
-In case of the \textbf{pseudo small area estimator}, the regression model stays the same for each small area estimation.
-Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling.\par
-
-Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas.
-Note that in the \code{grisons} dataset, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible.
-Applying the pseudo synthetic estimator equals treating those small areas as if $n_{2,G}=0$ by ignoring the terrestrial information.
-\begin{small}
-<<>>=
-psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights",
-                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                        unbiased = FALSE))
-psynth_2p$estimation
-@
-\end{small}
-
-In this case, the variance of the pseudo synthetic estimator is almost only half as high as under the \textit{psmall} and \textit{extended psynth} estimator.
-The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation  \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and XXX).
-Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.
-
-\newpage

### 2017-08-09 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 09.08.2017(4)

@@
 \end{small}
-The \code{twophase}-function creates an object of class \code{twophase}.
+The \code{twophase}-function creates an \code{S3} object of class \code{"global" "twophase"}.
 A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object.
@@
 
-Small area estimations in the package \pkg{forestinventory} can be applied by \textit{additionally} specifying the optional argument \code{small_area}.
+Small area estimations in the package \pkg{forestinventory} can be applied by additionally specifying the optional argument \code{small_area}.
 The input dataset has therefore to be extended by an additional column that specifies for each each row (i.e.
 observation of a sample location) the respective small area in which it is located.
-The argument \code{small_area} requires a \code{list}-object that comprises 1) the name of the column specifiying the small area of each observation (\code{sa.col}), 2) a vector specifying the small area(s) for which estimations are desired (\code{areas}), and 3) the argument \code{unbiased} that controls which of the three available estimators should be applied.
-In addition to global estimation, the \code{summary} now also displays the sample sizes in the respective small area (\code{n1G} and \code{n2G}).
+The argument \code{small_area} requires a \code{list}-object that comprises
+
+\begin{itemize}
+  \itemsep0em
+  \item the name of the column specifiying the small area of each observation (\code{sa.col})
+  \item a vector specifying the small area(s) for which estimations are desired (\code{areas})
+  \item the argument \code{unbiased} that controls which of the three available estimators is applied
+\end{itemize}
 
@@
 summary(psmall_2p)
+@
+\end{small}
+
+The small area functions all return an \code{S3} object of class \code{"smallarea" "twophase"}.
+In addition to global estimation, the \code{summary} output now also contains the sample sizes of each small area (\code{n1G} and \code{n2G}).
+Note that this information is also particularly provided in the \code{samplesizes} object returned by the function call.
+\begin{small}
+<<>>=
+psmall_2p$samplesizes
 @
@@
 Note that the point estimates and the design-based and external variances of both estimators are very similiar.
-However, the R$^2$s (\code{r.squared}) under the \textbf{extended pseudo small area estimator} are changing with the small area, while they are identical under the \textbf{pseudo small area estimator}.
-This is because under the \textbf{extended pseudo small area estimator}, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
-In case of the \textbf{pseudo small area estimator}, the regression model stays the same for each small area estimation.
+However, the R$^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are changing with the small area, while they are identical under the \textit{psmall} estimator.
+This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
+In case of the  \textit{psmall} estimator, the regression model stays the same for each small area estimation.
 Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling.\par
@@
 The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation  \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and XXX).
-Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.
+Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.\par
+
+The \textbf{exhaustive versions of the small area estimators} (equations \ref{eq:pointest_2p_reg_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
+Its application requires that we know the exact means of all explanatory variables within the small area of interest.
+In contrast to the \textit{global} estimators, the exact means have now to be delivered in the form of a \code{dataframe}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
+Note that likewise the case of global estimation, the order of the explanatory variables in the dataframe has to match the order in which they appear in the design matrix defined by the \code{lm} function in \proglang{R}.
+In order to tell \proglang{R} which row describes which small area, the row names have to match the respective names of the small areas specified in the \code{areas} argument.
+
+For the \code{grisons} dataset, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by
+\begin{small}
+<<eval=FALSE>>=
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+@
+\end{small}
+\begin{small}
+<<>>=
+true.means.Z.G <- data.frame(Intercept = rep(1, 4),
+                         mean = c(12.85, 12.21, 9.33, 10.45),
+                         stddev = c(9.31, 9.47, 7.90, 8.36),
+                         max = c(34.92, 35.36, 28.81, 30.22),
+                         q75 = c(19.77, 19.16, 15.40, 16.91))
+rownames(true.means.Z.G) <- c("A", "B", "C", "D")
+@
+\end{small}
+\begin{small}
+<<>>=
+true.means.Z.G
+@
+\end{small}
+
+For example, the \textbf{extended synthetic estimator} (\textit{extsynth}) can then be applied by 
+\begin{small}
+<<>>=
+extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
+                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                        small_area = list(sa.col ="smallarea", areas = c("A", "B"),
+                                          unbiased = TRUE),
+                        exhaustive = true.means.Z.G)
+extsynth_2p$estimation
+@
+\end{small}
+
+Comparing this result with the non-exhaustive counterpart \textit{extpsynth}, we see that the estimation uncertainty (defined by the variance) has again be significantly increased by substituting the estimated by the exact auxiliary means.
+In the function output, exhautive phases are indicated by a sample size of infinity \code{Inf}.
+Also note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set \code{true.means.Z.G}.
 

### 2017-08-10 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 10.08.2017(1)

@@
 \end{small}
-The \code{twophase}-function creates an \code{S3} object of class \code{"global" "twophase"}.
+The \code{twophase}-function creates an \code{S3} object of class \code{"twophase"} and subclass \code{"global"}.
 A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object.
@@
 The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}).
-Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model.
+Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model (\code{r.squared}).
 \begin{small}
@@
 Their point estimates are given in equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}.
-As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables available at the sample location $s_{1,G}$ in the small area.
+As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) available at the sample location $s_{1,G}$ in the small area.
 A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term.
@@
 
-The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textbf{small area estimator} and the \textbf{pseudo small area estimator}.
-Again note that all components are restricted those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
+The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textit{small} and the \textit{psmall} estimator.
+Again note that all components are restricted to those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
 
@@
 
+If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
+
 %------------------------------------------------------------- %
@@
 Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias.
-Due to the absence of available residuals in $G$, there is also no external variance form for synthetic estimations.
+Due to the absence of available residuals in $G$, there is also no external variance form for the synthetic and pseudo synthetic estimator.
 
@@
 
-The \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator}
+In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_reg_small}, \ref{eq:pointest_2p_reg_psmall}).
+However, the equations would considerably simplify if we could mathematically insure that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
+$\int_F R(x)dx=\int_G R(x)dx=0$.
+This is idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable which has the value \code{1} if the observation lies inside the small area $G$ and otherwise set to \code{0}.
+The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))\in{\mathbb{R}}^{(p+1)}$ and corresponds to the classical ANOVA case.
+Note that this adjusts the intercept of the regression model for all data points of small area $G$ and thus insures that the mean of the residuals $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\mathbb{R}(x)$ in $G$ are always zero by construction.\par
 
-% - explain extended estimator
-% - mention boundary weight adjustment likewise global estimation
+Under the extended auxiliary vector, the regression coefficient .. is calculated as
 
-\newpage
+\begin{equation}\label{ext_normequ_simple}
+  \hat{\pmb{\theta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
+\end{equation}
+
+with its variance-covariance matrix
+
+\begin{equation}\label{eq:ext_estvarmatrix}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+\end{equation}
+
+The point estimates are defined as
+
+\begin{subequations}\label{eq:pest_2p_extsynth_extpsynth}
+\begin{align}
+\hat{Y}_{G,extsynth}= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{Y}_{G,extpsynth}=\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth}
+\end{align}
+\end{subequations}
+
+The estimators are called \textit{synthetic} since no residual term is present such as in the small and psmall estimator (equation ...) and thus look similar to the synthetic estimators (equation ...).
+Note that it can however be shown by a decomposition of $\hat{\theta}_{s2}$ that the residual correction is indeed included in the regression coefficient $\hat{\theta}_{s2}$ \citep{mandallaz2016}.
 
@@
 
-In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}), we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
+In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
 \begin{small}
@@
 
-The small area functions all return an \code{S3} object of class \code{"smallarea" "twophase"}.
-In addition to global estimation, the \code{summary} output now also contains the sample sizes of each small area (\code{n1G} and \code{n2G}).
-Note that this information is also particularly provided in the \code{samplesizes} object returned by the function call.
+The small area functions all return an \code{S3} object of class \code{"twophase"} and subclass \code{"smallarea"}.
+In addition to global estimation, the \code{summary} output and the \code{estimation} object comprise the estimates and the sample sizes (\code{n1G} and \code{n2G}) for all calculated small areas (column \code{area}).
+Note that the sample size information is also particularly provided in the \code{samplesizes} object returned by the function call.
 \begin{small}
@@
 Comparing this result with the non-exhaustive counterpart \textit{extpsynth}, we see that the estimation uncertainty (defined by the variance) has again be significantly increased by substituting the estimated by the exact auxiliary means.
-In the function output, exhautive phases are indicated by a sample size of infinity \code{Inf}.
-Also note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set \code{true.means.Z.G}.
+In the function output, exhautive phases are indicated by a sample size of infinity (\code{Inf}).
+Also note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.
 

### 2017-08-10 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 10.08.2017(2)

@@
 This is idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable which has the value \code{1} if the observation lies inside the small area $G$ and otherwise set to \code{0}.
-The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))\in{\mathbb{R}}^{(p+1)}$ and corresponds to the classical ANOVA case.
+The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and corresponds to the classical ANOVA case.
 Note that this adjusts the intercept of the regression model for all data points of small area $G$ and thus insures that the mean of the residuals $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\mathbb{R}(x)$ in $G$ are always zero by construction.\par
@@
   \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathbb{R}}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
@@
 \begin{align}
-\hat{Y}_{G,extsynth}= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
-\hat{Y}_{G,extpsynth}=\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth}
+\hat{Y}_{G,extsynth,2p}= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{Y}_{G,extpsynth,2p}=\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth}
 \end{align}
@@
 
-The estimators are called \textit{synthetic} since no residual term is present such as in the small and psmall estimator (equation ...) and thus look similar to the synthetic estimators (equation ...).
+and the design-based variances are
+
+\begin{subequations}\label{eq:pest_2p_extsynth_extpsynth}
+\begin{align}
+\hat{\var}(\hat{Y}_{G,extsynth,2p})= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,extpsynth,2p})=
+\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}
++ \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
+\end{align}
+\end{subequations}
+
+with the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{\mathbb{Z}}}}_G$ estimated by
+
+\begin{equation}\label{ext_estvarcovaux_G}
+\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}}=
+\frac{1}{n_{1,G} (n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
+(\pmb{\mathbb{Z}}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G})(\pmb{\mathbb{Z}}(x)-
+\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G})^t
+\end{equation}
+
+The package also provides the external variance which is calculated as
+
+\begin{subequations}\label{eq:ext_varexternal_2p_extsynth}
+\begin{align}
+  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\mathbb{R}(x)-\bar{\mathbb{R}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & =
+ \frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\ 
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\mathbb{R}(x)-\bar{\mathbb{R}})^2 \label{eq:ext_varexternal_2p_extpsynth}
+\end{align}
+\end{subequations}
+
+The estimators are called \textit{synthetic} since no residual term is present such as in the small and psmall estimator (equation \ref{eq:pointest_2p_reg_small} and \label{eq:pointest_2p_reg_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
 Note that it can however be shown by a decomposition of $\hat{\theta}_{s2}$ that the residual correction is indeed included in the regression coefficient $\hat{\theta}_{s2}$ \citep{mandallaz2016}.

### 2017-08-11 rnw/Main structure and modules of the Package.Rnw
update Andreas 09.08.2017(4)
update Andreas 10.08.2017(1)
update Andreas 11.08.2017

@@
 The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase}-function.
-Their implementation and application is described in detail in section \ref{sec:globest_and_appl} and \ref{sec:saeest_and_appl}.
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase} function.
 
@@
 If the gain in model precision from the reduced to the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
-In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase}-function.
+In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase} function.
 
@@
 $n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
-Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase} and \code{threephase}-function in the \proglang{R} package \pkg{forestinventory}.
+Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase} and \code{threephase} functions in the \proglang{R} package \pkg{forestinventory}.
 
@@
 the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable.
-Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process \cite{maltamo2014}.
+Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process.
 In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e.
@@
 
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{theephase()} function.
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase} and \code{theephase} function.
 From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}).
 The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.
-The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global} estimations) or only for subregions of the entire inventory area (\textbf{small area} estimations).
+The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global estimators}) or only for subregions of the entire inventory area (\textbf{small area estimators}).
 The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area.
@@
 The package offers three small area estimators for both the exhaustive and non-exhaustive setup.
-The \textit{extended small area estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
+The \textit{extended synthetic estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
 The resulting point estimates and variances are asymptotically design-unbiased.
 If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
-The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation}) by the \code{onephase} function in \pkg{forestinventory}.
-This function can also be used to access the gain in efficiency of the mutliphase estimation procedures.
+The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation}) by the \code{onephase} function.
 All estimators are also available for \textit{cluster} sampling, in which case a sample unit conists of multiple, spatially agglomerated samples.
-The implementation and application of the estimators are described in detail in the following sections.
 
@@
 \centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/packagestructure.PNG}}
+\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_2.eps}}
 \caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
@@
 The following sections describe the mathematical details and the application of the multiphase estimators that are implemented in the \proglang{R} package \pkg{forestinventory}.
-While \cite{mandallaz2008, mandallaz2013c} provides an extensive derivation of all estimators, we will concentrate on the mathematical formulas that are actually implemented and applied in the package functions.
-For the sake of brevity, we will limited to simple sampling.
-Special cases occuring under cluster sampling are described in section X.
+While \cite{mandallaz2008, mandallaz2013c} provides an extensive derivation of all estimators, we will only provide the mathematical formulas that are actually implemented and applied in the functions of the package.
+For the sake of brevity, we will restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
+Special cases occuring under cluster sampling are described in section \ref{sec:speccas_and_scen}.
 

### 2017-08-11 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 11.08.2017

@@
 In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
-This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators.
-In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
-Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}).
-As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par
+This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$.
+In the \textit{non-exhaustive} estimator case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
+Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}).
+As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
 
@@
 The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively.
-Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e.
+Note that this form results under the particular case where the regression coefficients are derived using the data from the current inventory (i.e.
 \textbf{internal} model).
@@
 The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}).
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.
+Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citet{mandallaz2016}.
 
@@
 
-Note that the structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package.
+The structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package.
 Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location.
-Note that for global two-phase estimation, the data frame has to comprise a column that indicates whether a row corresponds to a $s_1$ or $s_2$ sample location (in \code{grisons}, this is the column \code{phase_id_2p}).\par
+For global two-phase estimation, we have to specify
 
-In order to apply the \textbf{non-exhaustive two-phase estimation with boundary weight adjustment} to the \code{grisons} data, we have to specify a) the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}, b) the data frame containing the inventory information (\code{data}), c) the name of the column specifying to which sample ($s_1$, $s_2$) an observation belongs (\code{phase.col}) as well as the indicator for the $s_2$, i.e.
-terrestrial sample (\code{terrgrid.id}), and d) the columname of the column containing the weights $w(x)$ of the boundary adjustments.
+\begin{itemize}
+  \itemsep0em
+  \item the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}
+  \item the \code{data.frame} containing the inventory information (\code{data})
+  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ (\code{terrgrid.id})
+  \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional)
+\end{itemize}
+
 The \textbf{non-exhaustive estimator with boundary weight adjustment} can thus be applied as follows:
+
 \begin{small}
@@
 \end{small}
-The \code{twophase}-function creates an \code{S3} object of class \code{"twophase"} and subclass \code{"global"}.
+The \code{twophase}-function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
 A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object.
@@
 The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
-We assume that in this case, the boundary adjustment has been applied when calculating the exact mean of $\pmb{Z}$.
+We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.
 
@@
 %--------------------------------------------------------------------------------------------------%
+% ################################################################################################ %
 %--------------------------------------------------------------------------------------------------%
@@
 We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
-Their point estimates are given in equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}.
+Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pointest_2p_psmall}.
 As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) available at the sample location $s_{1,G}$ in the small area.
@@
 \begin{align}
-  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_reg_small} \\
-  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_reg_psmall}
+  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
+  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_psmall}
 \end{align}
@@
 In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
-However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}) is no longer possible.
+However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pointest_2p_psmall}) is no longer possible.
 Thus, the synthetic / pseudo synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
@@
 
-In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_reg_small}, \ref{eq:pointest_2p_reg_psmall}).
+In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pointest_2p_psmall}).
 However, the equations would considerably simplify if we could mathematically insure that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This is idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable which has the value \code{1} if the observation lies inside the small area $G$ and otherwise set to \code{0}.
+This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable which has the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
 The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and corresponds to the classical ANOVA case.
@@
 
-Under the extended auxiliary vector, the regression coefficient .. is calculated as
+Under the extended auxiliary vector, the regression coefficient is calculated as
 
@@
 
-The point estimates are defined as
+The point estimate is defined as
 
@@
 
-and the design-based variances are
+and the design-based variances are derived as
 
@@
 
-The estimators are called \textit{synthetic} since no residual term is present such as in the small and psmall estimator (equation \ref{eq:pointest_2p_reg_small} and \label{eq:pointest_2p_reg_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
-Note that it can however be shown by a decomposition of $\hat{\theta}_{s2}$ that the residual correction is indeed included in the regression coefficient $\hat{\theta}_{s2}$ \citep{mandallaz2016}.
+The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \label{eq:pointest_2p_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
+However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown  that the residual correction is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is design-unbiased.
 
@@
 
-The small area functions all return an \code{S3} object of class \code{"twophase"} and subclass \code{"smallarea"}.
+The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}.
 In addition to global estimation, the \code{summary} output and the \code{estimation} object comprise the estimates and the sample sizes (\code{n1G} and \code{n2G}) for all calculated small areas (column \code{area}).
@@
 In case of the  \textit{psmall} estimator, the regression model stays the same for each small area estimation.
-Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling.\par
+Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling (see section \ref{sec:speccas_and_scen}).\par
 
@@
 In this case, the variance of the pseudo synthetic estimator is almost only half as high as under the \textit{psmall} and \textit{extended psynth} estimator.
-The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation  \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and XXX).
+The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and \ref{eq:var_2p_extpsynth}).
 Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.\par
 
-The \textbf{exhaustive versions of the small area estimators} (equations \ref{eq:pointest_2p_reg_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
+The \textbf{exhaustive versions of the small area estimators} (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area of interest.
-In contrast to the \textit{global} estimators, the exact means have now to be delivered in the form of a \code{dataframe}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
-Note that likewise the case of global estimation, the order of the explanatory variables in the dataframe has to match the order in which they appear in the design matrix defined by the \code{lm} function in \proglang{R}.
+In contrast to the \textit{global} estimators, the exact means have now to be delivered in the form of a \code{data.frame}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
+Note that likewise the case of global estimation, the order of the explanatory variables in the data frame has to match the order in which they appear in the design matrix defined by the \code{lm} function in \proglang{R}.
 In order to tell \proglang{R} which row describes which small area, the row names have to match the respective names of the small areas specified in the \code{areas} argument.
@@
 
-Comparing this result with the non-exhaustive counterpart \textit{extpsynth}, we see that the estimation uncertainty (defined by the variance) has again be significantly increased by substituting the estimated by the exact auxiliary means.
-In the function output, exhautive phases are indicated by a sample size of infinity (\code{Inf}).
+Comparing this result with the non-exhaustive counterpart \textit{extpsynth}, we see that the estimation uncertainty defined by the variance has again be significantly increased by substituting the estimated by the exact auxiliary means.
+In the function output, exhaustive phases are indicated by a sample size of infinity (\code{Inf}).
 Also note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.
 
-\newpage
+%\newpage

### 2017-08-11 JStatSoft_forestinventory_master.rnw
update Andreas 10.08.2017(1)
update Andreas 11.08.2017(2)

@@
 \Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations} %% a short title (if necessary)
+\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations in \proglang{R}} %% a short title (if necessary)
 
@@
 <<echo=FALSE>>=
-options(prompt = "R> ", continue = " + ")
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
 @
@@
 %------------------------------------------------------------------------------------------------%
+% ---------------------------------- Special Cases and Scenarios ------------------------------- %
+
+\SweaveInput{rnw/Analysis and Visualization.Rnw}
+
+%------------------------------------------------------------------------------------------------%
 % ----------------------------- Visualization and Analysis of Estimation Results --------------- %
@@
 
-\bibliography{bib/literature}\newpage\cleardoublepage
+\bibliography{bib/literature}\newpage
 
@@
 \end{document}
-
-% Hallo. Here comes a first try:
-% 
-% <<>>=
-% library(forestinventory)
-% op <- onephase(formula = tvol~1 ,data = grisons,
-%               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-% summary(op)
-% @
-% 
-% This was the result
-% 
-% 
-%

### 2017-08-11 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 11.08.2017(2)

@@
 
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
+An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
 In forest inventories, the sample is often restricted to those sample locations located within the forest area.
@@
 \textbf{internal} model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.
+In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.
 
@@
  \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
+(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:varexternal_2p_reg_nex}
 \end{align}
@@
   \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G 
-    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
+    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
   \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
   + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2} 
-  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
+  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
 \end{align}
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
   \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\ 
- &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
+ &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
 \end{align}
@@
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable which has the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
-The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and corresponds to the classical ANOVA case.
-Note that this adjusts the intercept of the regression model for all data points of small area $G$ and thus insures that the mean of the residuals $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\mathbb{R}(x)$ in $G$ are always zero by construction.\par
+This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
+The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and leads to a classical ANOVA regression model.
+Note that the intercept of the regression model is thus adjusted for all data points belonging to small area $G$ which insures that the mean of the residuals $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R(x)$ in $G$ are always zero by construction.\par
 
@@
   \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathbb{R}}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\mathbb{R}(x)-\bar{\mathbb{R}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
   \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & =
  \frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\ 
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\mathbb{R}(x)-\bar{\mathbb{R}})^2 \label{eq:ext_varexternal_2p_extpsynth}
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:ext_varexternal_2p_extpsynth}
 \end{align}
@@
 
-The \textbf{exhaustive versions of the small area estimators} (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
+The \textbf{exhaustive versions} of the small area estimators (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area of interest.

### 2017-08-15 JStatSoft_forestinventory_master.rnw
Corrected version that compiles on my macbook.  Contains a few latex error corrections and applies in the preamble of the master usepackage{Sweave}.

@@
   E-mail: \email{afmass@gmail.com}\\}
-}
 
@@
 
-%% need no \usepackage{Sweave.sty}
+\usepackage{Sweave}
 
@@
 \newcommand{\var}{\mathbb{V}}
-\newcommand{\COV}{\mathbb{COV}}
+% \newcommand{\COV}{\mathbb{COV}}
 \newcommand{\MAV}{\mathbb{MAV}}
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 
-%\SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
+% \SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
 

### 2017-08-15 rnw/Main structure and modules of the Package.Rnw
Corrected version that compiles on my macbook.  Contains a few latex error corrections and applies in the preamble of the master usepackage{Sweave}.

@@
 The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
-Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(2)}(x)}$ at the sample locations $s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{2}(x)$.
+Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(2)}(x)$ at the sample locations $s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{2}(x)$.
 If the gain in model precision from the reduced to the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.

### 2017-08-15 rnw/Twophase_Estimators and their Application.Rnw
Corrected version that compiles on my macbook.  Contains a few latex error corrections and applies in the preamble of the master usepackage{Sweave}.

@@
 \begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
+  \hat{\pmb{\beta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
 \end{equation}
@@
 \begin{equation}\label{eq:wmeanZ}
-  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
+  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_1}}w(x)}
 \end{equation}
@@
 
-If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
+If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
 
@@
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
+This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
 The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and leads to a classical ANOVA regression model.
@@
 \begin{equation}\label{ext_normequ_simple}
-  \hat{\pmb{\theta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
+  \hat{\pmb{\theta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
 \end{equation}
@@
 Note that the point estimates and the design-based and external variances of both estimators are very similiar.
-However, the R$^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are changing with the small area, while they are identical under the \textit{psmall} estimator.
+However, the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are changing with the small area, while they are identical under the \textit{psmall} estimator.
 This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.

### 2017-08-17 JStatSoft_forestinventory_master-concordance.tex
update Andreas 17.08.2017

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 121 1 1 4 1 1 1 4 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 128:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 224:%
136 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 259:%
1 84 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 344:%
172 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 360:%
1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 156 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 888:%
189}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 889:%
1 92 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
191 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 993:%
1 13 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1007:%
203 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1016:%
1 20 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1037:%
213 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1047:%
1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 24 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 79 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
224 20 1}

### 2017-08-17 rnw/Analysis and Visualization.Rnw
update Andreas 11.08.2017(2)
update Andreas 17.08.2017

--- 2017-08-09 rnw/Calculation of Confidence Intervals.Rnw
@@
 
-\section{Calculation of Confidence Intervals}
-\label{sec:confint_calc}
+\section{Analysis and Visualization}
+\label{sec:anal_and_vis}
+
+% include:
+% - estTable()-function demo
+% - mphase.gain()-function demo
+% - plot.esttable()-function demo
+% - 3-4 pages max
+% - pronounce that "it is crucial to quantify the potential benefit of the mphase procedures
+%   compared to the onephase procedure --> allows to calculate the financial benefit --> rel.eff
+
+The package \pkg{forestinventory} also provides methods for \textit{analysing} and \textit{visualizing} the estimation results created by the \code{onephase}, \code{twophase} and \code{threephase} functions.
+
+\subsection{Analysis}
+
+If we want to compare the results and performance of different estimation methods and estimators for a given global or small area inventory data set, the function to start with is \code{estTable()}.
+This function restructures the results of \code{onephase}, \code{twophase} and \code{threephase} estimations and merges them into one single data set that provides the basis for further analysis.
+For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} dataset:
+
+\begin{small}
+<<>>=
+op <- onephase(formula = tvol~1, data = grisons,
+               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+               area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
+                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                        small_area = list(sa.col = "smallarea", 
+                                          areas = c("A", "B","C", "D"),
+                                          unbiased = TRUE))
+@
+\end{small}
+
+\begin{small}
+<<>>=
+psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights",
+                      small_area = list(sa.col = "smallarea", 
+                                        areas = c("A", "B", "C", "D"),
+                                        unbiased = FALSE))
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
+                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+                                unbiased = TRUE))
+@
+\end{small}
+
+\begin{small}
+<<>>=
+psynth_3p<- threephase(formula.s0 = tvol ~ mean,
+                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                small_area = list(sa.col = "smallarea", 
+                                  areas = c("A", "B", "C", "D"),
+                                  unbiased = FALSE))
+@
+\end{small}
+
+We can then aggregate all estimation objects in a \code{list} and pass it to the \code{estTable}-function:
+
+\begin{small}
+<<>>=
+grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
+                                             extpsynth_3p, psynth_3p),
+                             sae = TRUE, 
+                             vartypes = c("variance", "g_variance",  "ext_variance"))
+@
+\end{small}
+
+The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}.
+If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}.
+Note however that the package \pkg{forestinventory} provides several \code{S3} generic functions that particularly respond to the class \code{"esttable"}.
+Note that the \code{vartypes} argument can be used to restrict the \code{estTable} output to certain types of variances.
+
+The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
+However, the point estimates and variances from all estimation objects passed to \code{estTable} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance that was applied.
+
+\begin{small}
+<<eval=FALSE>>=
+str(grisons.sae.table)
+@
+\end{small}
+
+\begin{footnotesize}
+<<echo=FALSE>>=
+str(grisons.sae.table)
+@
+\end{footnotesize}
+
+Note that \code{estTable} also returns the \textbf{\textit{estimation error}} (\code{error}) that is defined as the \textit{standard error} devided by the \textit{point estimate}:
+
+\begin{equation}\label{eq:esterror}
+  error_{[\%]}=\frac{\sqrt{\hat{\var}(\hat{Y})}}{\hat{Y}}\cdot100
+\end{equation}
+
+As mentioned in the introduction (section \ref{sec:intro}), the application of multi-phase estimation techniques in forest practice have two main \textbf{goals}:  
+\begin{enumerate}
+  \item \textit{\textbf{maintaining}} the one-phase estimation precision under reduced costs, i.e. reduced $s_2$ sample size
+  \item \textit{\textbf{increasing}} the estimation precision at identical costs, i.e. without increasing the $s_2$ sample size
+\end{enumerate}
+
+The function \code{mphase.gain()} was particularly written to quantify the potential benefit of a multi-phase global or small area estimation compared to its respective one-phase estimation.
+The function takes an \code{esttable} object as input and returns a validation of which multi-phase method and estimator performed best in comparison to the onephase estimation (\textit{baseline}) in terms of estimation precision.
+If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
+The argument \code{pref.vartype} can be used to define what type of variance (\textit{g-weight} or \textit{external}) should be used for the comparison.
+Per default (\code{exclude.synth = TRUE}), synthetic estimations (\textit{synth} and \textit{psynth} estimator) are not considered for the comparison since they normally have a much smaller variance but to the cost of a potential bias.
+
+<<echo=FALSE>>=
+options(width=10000) 
+@
+
+\begin{small}
+<<>>=
+mphase.gain(grisons.sae.table)
+@
+\end{small}
+
+% quantify the potential benefit of the mphase compared to the onephase estimation method
+% - pronounce that "it is crucial to quantify the potential benefit of the mphase procedures
+%   compared to the onephase procedure --> allows to quantify the financial benefit --> rel.eff
 

### 2017-08-17 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 17.08.2017

@@
 This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables.
-Based on the regression coefficients, the prediction for the sample location $x$ is calcualted as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
+Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
 Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
@@
 A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible.
-An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which is always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par
+An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which are always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par
 
@@
 We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
-Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pointest_2p_psmall}.
+Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pest_2p_psmall}.
 As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) available at the sample location $s_{1,G}$ in the small area.
@@
   \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
-  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_psmall}
+  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pest_2p_psmall}
 \end{align}
@@
 The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textit{small} and the \textit{psmall} estimator.
-Again note that all components are restricted to those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
+Again note that all components are restricted to those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
 
@@
 In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
-However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pointest_2p_psmall}) is no longer possible.
-Thus, the synthetic / pseudo synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
+However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pest_2p_psmall}) is no longer possible.
+Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
 $\int_G R(x)dx=0$.
 If this assumption is not fullfilled, the synthetic point estimate $\hat{Y}_{G,synth,2p}$ and the pseudo synthetic point estimate $\hat{Y}_{G,psynth,2p}$ will have a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
-Also note that residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
+Also note that the residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
 Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias.
@@
 
-In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pointest_2p_psmall}).
+In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pest_2p_psmall}).
 However, the equations would considerably simplify if we could mathematically insure that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
@@
 \frac{1}{n_{1,G} (n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
-(\pmb{\mathbb{Z}}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G})(\pmb{\mathbb{Z}}(x)-
-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G})^t
+(\pmb{\mathbb{Z}}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}_{G})(\pmb{\mathbb{Z}}(x)-
+\hat{\bar{\pmb{\mathbb{Z}}}}_{G})^t
 \end{equation}
@@
 
-The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \label{eq:pointest_2p_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
-However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown  that the residual correction is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is design-unbiased.
+The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \label{eq:pest_2p_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
+However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown that the residual correction term is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is design-unbiased.
 

### 2017-08-18 rnw/Analysis and Visualization.Rnw
update Andreas 18.08.2017

@@
 
-We can then aggregate all estimation objects in a \code{list} and pass it to the \code{estTable}-function:
+We can then aggregate all estimation objects in a \code{list} and pass it to the \code{estTable()}-function:
 
@@
 Note however that the package \pkg{forestinventory} provides several \code{S3} generic functions that particularly respond to the class \code{"esttable"}.
-Note that the \code{vartypes} argument can be used to restrict the \code{estTable} output to certain types of variances.
+Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.
 
 The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
-However, the point estimates and variances from all estimation objects passed to \code{estTable} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance that was applied.
+However, the point estimates, variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
+Note that per default, also the confidence intervals are added.
 
@@
 
-Note that \code{estTable} also returns the \textbf{\textit{estimation error}} (\code{error}) that is defined as the \textit{standard error} devided by the \textit{point estimate}:
+Note that \code{estTable()} also returns the \textbf{\textit{estimation error}} (\code{error}) that is defined as the \textit{standard error} devided by the \textit{point estimate}:
 
@@
 The function takes an \code{esttable} object as input and returns a validation of which multi-phase method and estimator performed best in comparison to the onephase estimation (\textit{baseline}) in terms of estimation precision.
-If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
+If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
 The argument \code{pref.vartype} can be used to define what type of variance (\textit{g-weight} or \textit{external}) should be used for the comparison.
@@
 <<>>=
-mphase.gain(grisons.sae.table)
+mphase.gain(grisons.sae.table, pref.vartype = "g_variance")
 @
@@
 
-% quantify the potential benefit of the mphase compared to the onephase estimation method
-% - pronounce that "it is crucial to quantify the potential benefit of the mphase procedures
-%   compared to the onephase procedure --> allows to quantify the financial benefit --> rel.eff
+% Interpretation: (gain and rel.eff)
+The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multi-phase estimator (\code{var_multiphase}).
+The multi-phase estimation procedure is again specified in the \code{method} and \code{estimator} column.
+The last two columns then actually quantify the potential benefit of the multi-phase estimation.
+The \code{gain} is the reduction (if its value is positive) or possibly also the increase (if its value is negative) in variance when applying the multi-phase as alternative to the onephase estimation.
+For small area \code{"B"}, it can be read as "the \textit{twophase extended psynth} estimation procedure leads to a 67.9 \% reduction in variance compared to the onephase procedure".
+The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:
+
+\begin{equation}\label{eq:releff}
+  rel.eff_{[\%]}=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}\cdot100
+  \end{equation}
+
+The \textit{relative efficiency} can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
+For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in order to get the same estimation precision with the one-phase approach as using the \textit{twophase extended psynth} estimator.
+If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple mean of quantifying the financial benefit of multi-phase forest inventories.
+
+\subsection{Visualization}
+
+The package \pkg{forestinventory} also provides a \code{S3} generic plotting function based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
+Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function.
+The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.
+
+If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case for the default settings of \code{estTable()}:
+
+\begin{small}
+<<>>=
+grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
+                                             extpsynth_3p, psynth_3p),
+                             sae = TRUE, 
+                             vartypes = c("variance", "g_variance",  "ext_variance"))
+@
+\end{small}
+
+The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic plot function and setting the argument \code{yvar = "estimate"}:
+
+\begin{small}
+<<eval=FALSE>>=
+plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
+  ylab("Timber Volume [m3/ha]")
+@
+\end{small}
+
+\begin{figure}[htb]
+\centering
+\resizebox{0.9\hsize}{!}{\includegraphics{fig/plot_est_ci.png}}
+\end{figure}
+
+If we want the estimation errors to be plotted, we simply leave the \code{yvar} argument to its default (\code{"error"}):
+
+\begin{small}
+<<eval=FALSE>>=
+plot(grisons.sae.table, ncol = 2)
+@
+\end{small}
+
+\begin{figure}[htb]
+\centering
+\resizebox{0.86\hsize}{!}{\includegraphics{fig/plot_error.png}}
+\end{figure}
+
+Note that the \code{plot}-function call can arbitrarily be extended by additional \pkg{ggplot2} functions and specifications.
 

### 2017-08-18 rnw/Analysis and Visualization.Rnw
update Andreas 18.08.2017(2)

@@
 The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
-However, the point estimates, variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
-Note that per default, also the confidence intervals are added.
+However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
+Per default, also the confidence intervals are added.
 
@@
 The last two columns then actually quantify the potential benefit of the multi-phase estimation.
-The \code{gain} is the reduction (if its value is positive) or possibly also the increase (if its value is negative) in variance when applying the multi-phase as alternative to the onephase estimation.
-For small area \code{"B"}, it can be read as "the \textit{twophase extended psynth} estimation procedure leads to a 67.9 \% reduction in variance compared to the onephase procedure".
+The \code{gain} is the reduction (if its value is positive) or possibly also the increase (if its value is negative) in variance when applying the multi-phase as alternative to the one-phase estimation.
+For small area \code{"B"}, it can be read as "the \textit{twophase extended psynth} estimation procedure leads to a 67.9 \% reduction in variance compared to the one-phase procedure".
 The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:
@@
 The \textit{relative efficiency} can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
-For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in order to get the same estimation precision with the one-phase approach as using the \textit{twophase extended psynth} estimator.
+For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{twophase extended psynth} estimator.
 If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple mean of quantifying the financial benefit of multi-phase forest inventories.

### 2017-08-18 JStatSoft_forestinventory_master-concordance.tex

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1037:%
-213 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1047:%
-1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 24 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 79 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
-224 20 1}
+213 15 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1053:%
+1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1308:%
+230 3 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1312:%
+1 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1317:%
+235 11 1}

### 2017-08-18 rnw/Acknowledgements.Rnw

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section*{Acknowledgments}

We want to express our gratitude to Prof. H. Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study and providing the possibility of working on the package. We also want to thank Daniel Mandallaz for his support in completing the range of the allready published estimators in the frame of the three-phase small area estimators, many helpful discussions and advice throughout the implementation of our package, and for proofreading the manuscript. Our thanks also go to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of grisons for providing the example data.

### 2017-08-19 JStatSoft_forestinventory_master-concordance.tex
update Andreas 19.08.2017

--- 2017-08-17 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 259:%
-1 84 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 344:%
+1 78 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 338:%
 172 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 360:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 354:%
 1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 156 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 882:%
 189}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 889:%
-1 92 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 883:%
+1 213 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1097:%
 191 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 993:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1108:%
 1 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1007:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1122:%
 203 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1016:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1131:%
 1 20 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1037:%
-213 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1047:%
-1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 24 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 79 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
-224 20 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1152:%
+213 15 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1168:%
+1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1423:%
+230 3 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1427:%
+1 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1432:%
+235 11 1}

### 2017-08-19 rnw/Main structure and modules of the Package.Rnw
update Andreas 17.08.2017
update Andreas 19.08.2017

@@
 The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase} function.
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
 
@@
 The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and $s_0$.
-The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(2)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
+The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
 From that it follows that we can define \textit{two nested} regression models.
-The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(1)t}(x),\pmb{Z}^{(2)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$.
+The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
 The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
-Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(2)}(x)$ at the sample locations $s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{2}(x)$.
-If the gain in model precision from the reduced to the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
-In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase} function.
-
-% NOTE:
-% We still have to explain the difference between "non-exhausitve" and "exhaustive" auxiliary information (evtl. tricky)
-% and the difference between cluster and no-cluster sampling
-% --> I would not mentioned it here (keep  it general), but rather go into detail when describing the estimtors (next sections)
+Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(0)}(x)$ in the so called \textit{reduced model} at the sample locations $x \in s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{0}(x)$.
+If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
+In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
 
@@
 $n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
-Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase} and \code{threephase} functions in the \proglang{R} package \pkg{forestinventory}.
+Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.
 
@@
 
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase} and \code{theephase} function.
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
 From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}).
@@
 If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
-The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation}) by the \code{onephase} function.
-All estimators are also available for \textit{cluster} sampling, in which case a sample unit conists of multiple, spatially agglomerated samples.
+The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{one-phase estimation}) by the \code{onephase()} function.
+All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
 

### 2017-08-19 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 09.08.2017(4)
update Andreas 19.08.2017

@@
 
+% ---------------------------------------------------------------------------- %
 % ---------------------------------------------------------------------------- %
@@
 %
-% - 
 % 
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
-
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
+% NOTES:
+% - mention that the exhaustive 3phase was originally called "2phase with partially exhaustive info"
+% - clarify terminology "full" and "reduced"
 %
 
+% ------------------------------------- %
+\subsubsection{Mathematical Background}
+
+The the vector of regression coefficients for the \textbf{\textit{reduced model}} using the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$ is derived by solving the normal equation
+
+\begin{eqnarray}\label{eq:normequ_redmod}
+\hat{\pmb{\alpha}}_{s_2} &=& \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)
+\end{eqnarray}
+
+and likewise the vector of regression coefficients for the \textbf{\textit{full model}} using the full set of explanatory variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$:
+
+\begin{eqnarray}\label{eq:normequ_redmod}
+\hat{\pmb{\beta}}_{{s_2}}&=&\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^t(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x)
+\end{eqnarray}
+
+The variance-covariance matrices of the regression coefficients are thus defined as
+
+\begin{subequations}\label{eq:covar3p}
+\begin{align}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \label{eq:covar_beta}
+\end{align}
+\end{subequations}
+
+with the residuals $R_0(x)$ and $R(x)$ derived under the reduced and the full regression model respectively.
+
+Likewise the two-phase estimation (section \ref{sec:twophase_and_appl}), also the three-phase estimation can be applied under \textit{non-exhaustive} and \textit{exhaustive} derivation of the explanatory variables at the $s_0$ phase.
+
+In the \textbf{non-exhaustive case}, we use the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over both the $s_1$ phase (equation \ref{meanvalues3pnex}).
+If the explanatory variables used in the reduced model are exhaustively derived (\textbf{exhaustive case}), the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ is simply replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
+Note that in case of applied boundary adjustment (section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to equation \ref{eq:wmeanZ}.
+
+\begin{equation}\label{meanvalues3pnex}
+\hat{\bar{\pmb{Z}}}^{(0)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(0)}(x) ,
+\quad \hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}(x)
+\end{equation}
+
+Under the internal model prerequisite which insures that the zero mean residual property over $F$, \textbf{point estimates} are then calculated as
+
+\begin{subequations}\label{reg3p}
+\begin{align}
+\hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_ex} \\
+\hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_nex}
+\end{align}
+\end{subequations}
+
+Equations \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} nicely demonstrate the philosophy of the three-phase estimation approach.
+Without the first term, the equations exactly equal the one of two-phase estimator where the estimated means in $\hat{\bar{\pmb{Z}}}_1$ are based on the $s_1$ sample and are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$.
+The difference however is that we can now derive a more precise mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ (or even the exact mean $\bar{\pmb{Z}}^{(0)}_0$) over $s_0$ for a subset $\pmb{Z}^{(0)}(x) \in \pmb{Z}(x)$.
+The first term of equation \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} can thus be interpreted as a correction term that transforms the difference to the more precise / exact mean of $\pmb{Z}^{(0)}(x)$ into a residual of $Y(x)$ and adds it to the second term, thus making the point estimate more precise.
+
+The \textbf{design-based variance} is then calculated as
+
+\begin{subequations}\label{eq:dbvar_reg3p}
+\begin{align}
+\hat{\var}(\hat{Y}_{reg3p,ex})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_ex} \\
+\hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
++\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
+\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
+\end{align}
+\end{subequations}
+
+with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}$:
+
+\begin{equation}\label{estvarcovaux3p}
+\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}=
+\frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
+(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t
+\end{equation}
+
+We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
+This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.
+
+The package also gives the \textbf{external variance} defined as
+
+\begin{subequations}\label{eq:extvar_reg3p}
+\begin{align}
+\hat{\var}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x))
++ (1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s_2}}\hat{R}^2(x) \label{eq:extvar_reg3p_ex} \\
+\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_0(x)-\hat{\bar{Y}}_0)^2}{n_0-1} 
++\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_0(x)-\hat{\bar{R_0}})^2 \nonumber\\
+&+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
+\end{align}
+\end{subequations}
+
+with ${R}(x)$ being the full model residuals calculated as $Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ and ${R}_{0}(x)$ being the reduced model residuals calculated as $Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$.
+
+% \begin{subequations}\label{eq:var_reg3p_nex}
+% \begin{align}
+% \hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
+% +\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
+% \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}^{t}_1 \label{eq:dbvar_reg3p_nex} \\
+% \hat{\var}_{ext}(\hat{Y}_{reg3p,nex})& = \frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1} \nonumber\\
+% &+ \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_1(x)-\hat{\bar{R_1}})^2+
+% (1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
+% \end{align}
+% \end{subequations}
+
+% ------------------------------------- %
+\subsubsection{Application}
+
+%--------------------------------------------------------------------------------------------------%
+% ################################################################################################ %
+%--------------------------------------------------------------------------------------------------%
+
+% ---------------------------------------------------------------------------- %
 \subsection{Small Area Estimators}
 
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
+% ------------------------------------- %
+\subsubsection{Mathematical Background}
+
+%------------------------------------------------------------- %
+\subsubsection{Application}
+
+\newpage

### 2017-08-22 JStatSoft_forestinventory_master-concordance.tex
update Andreas 22.08.2017

--- 2017-08-17 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 121 1 1 4 1 1 1 4 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 128:%
+1 123 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
 1 95 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 224:%
-136 34 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 259:%
-1 84 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 344:%
-172 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 360:%
-1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 156 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 888:%
-189}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 889:%
-1 92 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
-191 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 993:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
+139 34 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
+1 78 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 341:%
+175 13 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Onephase_Estimators and their Application.Rnw:ofs 355:%
+1 25 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 381:%
+190}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 382:%
+1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 911:%
+192}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 912:%
+1 118 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 179 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1302:%
+194 10 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1313:%
 1 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1007:%
-203 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1016:%
-1 20 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1037:%
-213 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1047:%
-1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 24 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 79 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
-224 20 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1327:%
+206 8 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1336:%
+1 21 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1358:%
+216 15 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1374:%
+1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1629:%
+233 3 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1633:%
+1 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1638:%
+238 11 1}

### 2017-08-22 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 22.08.2017

@@
 
-\section{Threephase Estimators and their Application}
+\section{Three-phase Estimators and their Application}
 \label{sec:threephase_and_appl}
@@
 
-The the vector of regression coefficients for the \textbf{\textit{reduced model}} using the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$ is derived by solving the normal equation
+The the vector of regression coefficients for the \textbf{\textit{reduced model}} using the \textit{reduced set of explanatory variables} $\pmb{Z}^{(0)}(x)$ available at all sample locations $x \in s_0$ is derived by solving the normal equation
 
@@
 
-and likewise the vector of regression coefficients for the \textbf{\textit{full model}} using the full set of explanatory variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$:
+and likewise the vector of regression coefficients for the \textbf{\textit{full model}} using the \textit{full set of explanatory variables} $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ available only at a subset of all sample locations $x \in s_1 \subset s_0$:
 
@@
 
-with the residuals $R_0(x)$ and $R(x)$ derived under the reduced and the full regression model respectively.
+with the full model residuals ${R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ and the reduced model residuals ${R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ derived under the reduced and the full regression model respectively.
 
@@
 
-In the \textbf{non-exhaustive case}, we use the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over both the $s_1$ phase (equation \ref{meanvalues3pnex}).
+In the \textbf{non-exhaustive case}, we use the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase (equation \ref{meanvalues3pnex}).
 If the explanatory variables used in the reduced model are exhaustively derived (\textbf{exhaustive case}), the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ is simply replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
@@
 
-Under the internal model prerequisite which insures that the zero mean residual property over $F$, \textbf{point estimates} are then calculated as
+The internal model case again insures that the zero mean residual property over $F$, and the \textbf{point estimates} are thus calculated as
 
@@
 
-with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}$:
+with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}$:
 
@@
 \frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
-(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t
+(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t
 \end{equation}
@@
 \hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_0(x)-\hat{\bar{Y}}_0)^2}{n_0-1} 
-+\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_0(x)-\hat{\bar{R_0}})^2 \nonumber\\
-&+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
++\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_0(x)-\bar{\hat{R_0}})^2 \nonumber\\
+&+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:extvar_reg3p_nex}
 \end{align}
@@
 
-with ${R}(x)$ being the full model residuals calculated as $Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ and ${R}_{0}(x)$ being the reduced model residuals calculated as $Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$.
-
-% \begin{subequations}\label{eq:var_reg3p_nex}
-% \begin{align}
-% \hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
-% +\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
-% \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}^{t}_1 \label{eq:dbvar_reg3p_nex} \\
-% \hat{\var}_{ext}(\hat{Y}_{reg3p,nex})& = \frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1} \nonumber\\
-% &+ \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_1(x)-\hat{\bar{R_1}})^2+
-% (1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
-% \end{align}
-% \end{subequations}
-
 % ------------------------------------- %
 \subsubsection{Application}
+
+Note that the \code{grisons} inventory dataset originally corresponds to a classical two-phase sample design, i.e.
+the full set of explanatory variables is available at all sample locations in the inventory area.
+To demonstrate the three-phase estimation procedure, we provide an artificial three-phase scenario by recoding the phase indicators in the data set (column \code{phase_id_3p}).
+The values for indicating the phases (\code{0,1,2}) follow the terminology used in this article ($s_0$, $s_1$, $s_2$).
+
+We can use the \code{table()} function to see the sample sizes of all phases for the three-phase secenario:
+
+\begin{small}
+<<>>=
+table(grisons$phase_id_3p)
+@
+\end{small}
+
+We see that in the three-phase example, we now have 40 terrestrial sample locations providing the timber volume per hectare (\code{tvol}).
+The large phase $s_0$ comprises $178 + 88 + 40 = 306$ sample locations, since $s_2 \subset s_1 \subset s_0$.
+We now assume that at sample locations $x \in s_0$, only the mean canopy height (\code{mean}) is available, whereas at sample locations $x \in s_1$ we have also have the explanatory variables \code{stddev}, \code{max} and \code{q75} at hand.
+We can now define the reduced and full regression model formulas that we will use in the three-phase estimation.
+Note that the models are nested.
+
+\begin{small}
+<<>>=
+formula.rm <- tvol ~ mean # reduced model applied to s0 phase
+formula.fm <- tvol ~ mean + stddev + max + q75 # full model applied to s1 phase
+@
+\end{small}
+
+Note that compared to the \code{twophase()} function, we now have to specify the two regression models, i.e.
+the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
+In addition, we also have to specify the indication of the $s_1$ phase (\code{s1.id}) in the argument \code{phase_id}.
+The global three-phase estimation can thus be applied by
+
+<<echo=FALSE>>=
+options(width=10000) 
+@
+
+\begin{small}
+<<>>=
+reg3p_nex<- threephase(formula.s0 = formula.rm, 
+                  formula.s1 = formula.fm, data = grisons,
+                  phase_id = list(phase.col="phase_id_3p", s1.id = 1, 
+                                  terrgrid.id = 2),
+                  boundary_weights = "boundary_weights")
+@
+\end{small}
+
+\begin{small}
+<<>>=
+summary(reg3p_nex)
+@
+\end{small}
+
+The summary of a \code{threephase} function recalls both regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
+The set of sample sizes have also been extented by that of the $s_0$ phase (\code{n0}).
+In order to compare the results to the respective two-phase estimation, we assume that we have the mean canopy height as the only explanatory variable at all 306 sample locations:
+
+\begin{small}
+<<>>=
+reg2p_nex<- twophase(formula = formula.rm, data = grisons,
+                       phase_id = list(phase.col="phase_id_3p", terrgrid.id = 2),
+                       boundary_weights = "boundary_weights")
+@
+\end{small}
+
+\begin{small}
+<<>>=
+reg2p_nex$estimation
+@
+\end{small}
+
+Comparing the $R^2$s of the three and two-phase result, it becomes obvious that the incorporation of the additional explanatory variables \code{stddev}, \code{max} and \code{q75} led to a 20\% increase in explained variance in the regression model.
+While the more precise regression model was only applied at 40\% of all 306 sample locations (\textit{sampling fraction}), it was possible to considerably reduce the two-phase variance and thus achieve a higher estimation precision.
+
+Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
+In the first scenario, more precise auxiliary information becomes available at a subset of the allready existing and used sample locations.
+In this case, one simply reclassifies the existing sample locations $x \in s_1$ into $s_0$ sample locations and assigns the $s_1$ indicator to those points where the new variables are available.
+In the second scenario, the auxiliary information is available at all sample locations $x \in s_1$, but the computation of some of the explanatory variables is computationally expensive.
+In this case, a considerable time saving could be realized by restricing their computation to a subset of all $s_1$ sample locations, which consequently leads to a three-phase scenario that is handeled likewise the first scenario.
+A respective example is given in \citet{mandallaz2013b} and \citet{mandallaz2013c}.
+In both the first and second scenario, the original sample size of the two-phase inventory remains unchanged, but the $s_1$ sample locations are splitted into $s_0$ and $s_1$ locations.
+In the third and last scenario, one has already used a set of explanatory variables at $s_1$ sample locations for a time before additional, less precise information becomes available in a much larger quantity over the inventory area.
+It still might yield an increase in estimation precision if this information is considered in the estimation in a much larger sample size than the more precise information (sample fraction).
+In this case, one creates a new phase $s_0$ by densifying the allready existing $s_1$ sample locations (which in practice are often arranged in regular grids) such that $n_0 >> n_1$ and $s_1 \in s_0$.
 
@@
 
+The 3 small area estimators under three-phase sampling provided by the package extend the two-phase small area estimator of section \ref{sec:twophase_sae} to the three-phase case.
+Note that the general principle stays the same, i.e.
+all information used is restricted to the small area $G$, with exception of retrieving the regession coefficients of the reduced and full model and their variance-covariance matrices.
+
+We first define the estimated means of the explanatory variables within the small area $G$:
+
+\begin{equation}\label{meanvalues3psaenext}
+\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x) ,
+\quad \hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)
+\end{equation}
+
+%----------------------------------------------- %
+% \textbf{Small and Pseudo Small Area Estimator}\par
+
+The \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) follow the same principle as in the two-phase case, i.e.
+applying the globally derived reduced and full regression model to the small area means of the explanatory variables, and then correcting for a potential model bias by adding the residual mean of the full model, which are not zero in general, to the point estimate (equations \ref{eq:pe_3p_small} and \ref{eq:pe_3p_psmall}).
+Note that the difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ derived in the $s_1$ phase and the (more) exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ or $\bar{\pmb{Z}}^{(0)}_{0,G}$ derived in the $s_0$ phase is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
+
+The \textbf{point estimates} are unbiased and given by:
+
+\begin{subequations}\label{eq:pest_3p_small_psmall}
+\begin{align}
+\hat{Y}_{G,small,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_small} \\
+\hat{Y}_{G,psmall,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_psmall}
+\end{align}
+\end{subequations}
+
+with ${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$ being the full model residuals in small area $G$. The \textbf{design-based variance} is then calculated as
+
+\begin{subequations}\label{eq:var_3p_small_psmall}
+\begin{flalign}
+\hat{\var}(\hat{Y}_{G,small,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\bar{\pmb{Z}}}_{1,G} + 
+\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}_{G}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:var_3p_reg_small}&\\
+\hat{\var}(\hat{Y}_{G,psmall,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\bar{\pmb{Z}}}_{1,G} \nonumber &\\
+&+ \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}_{G}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:var_3p_reg_psmall}
+\end{flalign}
+\end{subequations}
+
+with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}$ as
+
+\begin{equation}\label{estvarcovaux3pG}
+\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}=
+\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
+(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})^t
+\end{equation}
+
+The \textbf{external variance} is defined as:
+
+\begin{subequations}\label{eq:pest_3p_small_psmall}
+\begin{flalign}
+\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}
+ (\hat{R}_0(x)-\hat{\bar{R}}_{0,G})^2 + 
+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_G)^2 \label{eq:var_3p_reg_small}&\\
+\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}(\hat{Y}_0(x)-\hat{\bar{Y}}_{0,G})^2 + 
+{\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
+ (\hat{R}_0(x)-\bar{\hat{R}}_{0,G})^2 \nonumber &\\
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_G)^2 \label{eq:var_3p_reg_psmall}
+\end{flalign}
+\end{subequations}
+
+with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$ being the reduced model predictions and $\hat{R}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
+Likewise the design-based variance, the difference between the exhaustive and non-exhaustive form is the variance of the predictions $\hat{Y}_{0}(x)$ within the $s_{0,G}$ phase.
+
+%------------------------------------------------------------- %
+% \textbf{Synthetic and Pseudo Synthetic Estimator}\par
+
+The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} ca be applied if no terrestrial samples are available in the small area, i.e.
+$n_{2,G}=0$.
+Consequently, the residual correction and the residual variation term of the full model can no longer be applied an vanishes from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and design-based variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
+The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$.
+Also the variance will be small but to the cost of ignoring the model uncertainties.
+Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.
+
+\begin{subequations}\label{eq:pest_3p_synth_psynth}
+\begin{align}
+\hat{Y}_{G,synth,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2} \label{eq:pe_3p_synth} \\
+\hat{Y}_{G,psynth,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}\label{eq:pe_3p_psynth}
+\end{align}
+\end{subequations}
+
+\begin{subequations}\label{eq:var_3p_synth_psynth}
+\begin{align}
+\hat{\var}(\hat{Y}_{G,synth,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
+\hat{\var}(\hat{Y}_{G,psynth,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
+\end{align}
+\end{subequations}
+
+%------------------------------------------------------------- %
+% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par
+
 %------------------------------------------------------------- %

### 2017-08-22 r/application_threephase.R

@@ -1,0 +1,106 @@
+#---------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                  #
+#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                  #
+#                                                                                   #
+#  Code for running all examples given in                                           #
+#  Section 4.1  "Application of three-phase global and small area estimators"       #
+#                                                                                   #
+#---------------------------------------------------------------------------------- #
+
+library(forestinventory)
+
+
+#####################################
+# GLOBAL THREE-PHASE ESTIMATIONS    #
+#####################################
+
+## get sample sizes of three-phase secenario:
+table(grisons$phase_id_3p)
+
+## define reduced and full regression model:
+formula.s0 <- tvol ~ mean # reduced model:
+formula.s1 <- tvol ~ mean + stddev + max + q75 # full model
+
+## calculate one-phase estimation:
+op <- onephase(formula = tvol~1 ,data = grisons,
+               phase_id =list(phase.col = "phase_id_3p",terrgrid.id = 2))
+
+## calculate two-phase estimation:
+reg2p_nex<- twophase(formula=formula.s0,
+                     data=grisons,
+                     phase_id=list(phase.col = "phase_id_3p", terrgrid.id = 2),
+                     boundary_weights = "boundary_weights")
+
+## calculate three-phase estimation:
+reg3p_nex<- threephase(formula.s0, formula.s1, data = grisons,
+                       phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                       boundary_weights = "boundary_weights")
+
+## get summaries:
+summary(reg2p_nex)
+summary(reg3p_nex)
+
+# merge to esttable-object:
+etab.2p.glob<- estTable(est.list = list(op, reg2p_nex))
+etab.3p.glob<- estTable(est.list = list(op, reg2p_nex, reg3p_nex))
+
+# visualize estimation errors:
+plot(etab.3p.glob)
+
+# quantify gain:
+etab.2p.glob<- estTable(est.list = list(op, reg2p_nex))
+mphase.gain(etab.2p.glob)
+mphase.gain(etab.3p.glob)
+
+
+
+#########################################
+# SMALL AREA THREE-PHASE ESTIMATIONS    #
[…]

### 2017-08-22 JStatSoft_forestinventory_master-concordance.tex
update Andreas 22.08.2017(2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 912:%
-1 118 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 179 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1302:%
+1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 2 1 1 2 9 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 9 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 9 0 1 2 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1421:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1313:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1432:%
 1 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1327:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1446:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1336:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1455:%
 1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1358:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1477:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1374:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1493:%
 1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1629:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1633:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1752:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1638:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1757:%
 238 11 1}

### 2017-08-22 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 22.08.2017(2)

@@
 \subsection{Global Estimators}
+\label{sec:glob_est_3p}
 
@@
 \begin{align}
-\hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_ex} \\
-\hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_nex}
+\hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
+\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
+\hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
+\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
 \end{align}
@@
 \hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_ex} \\
-\hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
+\hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_{s_2} ^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_{s_2}
 +\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
-\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
+\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
 \end{align}
@@
 
-We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
+We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
 This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.
@@
 We now assume that at sample locations $x \in s_0$, only the mean canopy height (\code{mean}) is available, whereas at sample locations $x \in s_1$ we have also have the explanatory variables \code{stddev}, \code{max} and \code{q75} at hand.
-We can now define the reduced and full regression model formulas that we will use in the three-phase estimation.
+We can now define the reduced and full regression model formulas that we will use in the \code{threephase()} function.
 Note that the models are nested.
@@
 
-Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
-In the first scenario, more precise auxiliary information becomes available at a subset of the allready existing and used sample locations.
-In this case, one simply reclassifies the existing sample locations $x \in s_1$ into $s_0$ sample locations and assigns the $s_1$ indicator to those points where the new variables are available.
-In the second scenario, the auxiliary information is available at all sample locations $x \in s_1$, but the computation of some of the explanatory variables is computationally expensive.
-In this case, a considerable time saving could be realized by restricing their computation to a subset of all $s_1$ sample locations, which consequently leads to a three-phase scenario that is handeled likewise the first scenario.
-A respective example is given in \citet{mandallaz2013b} and \citet{mandallaz2013c}.
-In both the first and second scenario, the original sample size of the two-phase inventory remains unchanged, but the $s_1$ sample locations are splitted into $s_0$ and $s_1$ locations.
-In the third and last scenario, one has already used a set of explanatory variables at $s_1$ sample locations for a time before additional, less precise information becomes available in a much larger quantity over the inventory area.
-It still might yield an increase in estimation precision if this information is considered in the estimation in a much larger sample size than the more precise information (sample fraction).
-In this case, one creates a new phase $s_0$ by densifying the allready existing $s_1$ sample locations (which in practice are often arranged in regular grids) such that $n_0 >> n_1$ and $s_1 \in s_0$.
-
 %--------------------------------------------------------------------------------------------------%
@@
 \begin{align}
-\hat{Y}_{G,small,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_small} \\
-\hat{Y}_{G,psmall,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_psmall}
+\hat{Y}_{G,small,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_small} \\
+\hat{Y}_{G,psmall,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_psmall}
 \end{align}
@@
 \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}_{G}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:var_3p_reg_small}&\\
-\hat{\var}(\hat{Y}_{G,psmall,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
+\hat{\var}(\hat{Y}_{G,psmall,3p})& = \hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_{s_2} +
 \frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
@@
 
-with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}$ as
+with the variance-covariance matrix
 
@@
 
-The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} ca be applied if no terrestrial samples are available in the small area, i.e.
+The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) ca be applied if no terrestrial samples are available in the small area, i.e.
 $n_{2,G}=0$.
@@
 \hat{Y}_{G,synth,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2} \label{eq:pe_3p_synth} \\
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
 \hat{Y}_{G,psynth,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{2}\label{eq:pe_3p_psynth}
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth}
 \end{align}
@@
 
+The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.
+$\int_F R(x)dx=\int_G R(x)dx=0$.
+This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
+The extended auxiliary vector thus is defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{\mathbb{Z}}^{t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathbb{Z}}^{t}(x)=\pmb{Z}^{t}(x)$.
+This practically means that we have to specify the binary indicator variable for all observations in the data frame that is passed to the package-functions.
+
+We first recalculate the regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ as
+
+\begin{subequations}\label{eq:normequ_redmod_ext}
+\begin{align}
+\hat{\pmb{\gamma}}_{s_2} &= \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}^{(0)}(x) \\
+\hat{\pmb{\theta}}_{{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x)
+\end{align}
+\end{subequations}
+
+and
+
+\begin{subequations}\label{eq:covar3p_ext}
+\begin{align}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+\end{align}
+\end{subequations}
+
+After computing the auxiliary means $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x)$, $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x)$, and $\hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)$, we can calculate the \textbf{point estimate} as
+
+\begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
+\begin{align}
+\hat{Y}_{G,extsynth,3p}&=(\bar{\pmb{\mathbb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
+\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2} \label{eq:pe_3p_extsynth} \\
+\hat{Y}_{G,extpsynth,3p}&=(\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
+\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth}
+\end{align}
+\end{subequations}
+
+and the \textbf{design-based variance} defined as
+
+\begin{subequations}\label{eq:var_synth_psynth_ext}
+\begin{align}
+\hat{\var}(\hat{Y}_{G,extsynth,3p})& =\frac{n_2}{n_1}\bar{\pmb{\mathbb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extsynth}\\
+\hat{\var}(\hat{Y}_{G,extpsynth,3p})& = \hat{\pmb{\gamma}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}\hat{\pmb{\gamma}}_{s_2} + \frac{n_2}{n_1}\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
+\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extpsynth}
+\end{align}
+\end{subequations}
+
+with
+
+\begin{equation}\label{estvarcovaux3pG_ext}
+\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}=
+\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
+(\pmb{\mathbb{Z}}^{(0)}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G})(\pmb{\mathbb{Z}}^{(0)}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G})^t
+\end{equation}
+
+The package also returns the \textbf{external variance} which is defined as
+
+\begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
+\begin{flalign}
+\hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}
+ (\hat{\mathbb{R}}_0(x)-\hat{\bar{\mathbb{R}}}_{0,G})^2 + 
+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}}_G)^2 \label{eq:var_3p_reg_extsynth}&\\
+\hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}(\hat{Y}_0(x)-\hat{\bar{Y}}_{0,G})^2 + 
+{\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
+ (\hat{\mathbb{R}}_0(x)-\bar{\hat{\mathbb{R}}}_{0,G})^2 \nonumber &\\
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}}_G)^2 \label{eq:var_3p_reg_extpsynth}
+\end{flalign}
+\end{subequations}
+
+with $\hat{Y}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$ being the reduced model predictions and $\hat{\mathbb{R}}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
+The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in equation \ref{eq:pe_3p_synth}, \ref{eq:pe_3p_psynth}, \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}.
+Note that they are however design-unbiased since the residual correction term is included in the regression coefficient vector $\hat{\pmb{\theta}}_{s_2}$.
+
 %------------------------------------------------------------- %
@@
 
+We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the \textbf{extended synthetic} and the \textbf{synthetic} estimator to the \code{grisons} dataset.
+The setup is thus exactly the same as in the example for global three-phase estimation (section \ref{sec:glob_est_3p}).
+However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}.
+We thus first define the true means for each small area likewise we did in the \code{twophase()} example (section \ref{sec:twophase_sae}):
+
+\begin{small}
+<<>>=
+truemeans.G <- data.frame(Intercept = rep(1, 4),
+                         mean = c(12.85, 12.21, 9.33, 10.45))
+rownames(truemeans.G) <- c("A", "B", "C", "D")
+@
+\end{small}
+
+Three-phase small area estimation in the package can in general be applied by additionally specifying the \code{small_area} list argument.
+The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument.
+The \textbf{extended synthetic estimator} can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):
+
+\begin{small}
+<<>>=
+extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+                  phase_id = list(phase.col = "phase_id_3p", 
+                                  s1.id = 1, terrgrid.id = 2),
+                  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+                                   unbiased = TRUE),
+                  exhaustive = truemeans.G, 
+                  boundary_weights = "boundary_weights")
+
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extsynth_3p$estimation
+@
+\end{small}
+
+The \textbf{synthetic estimator} can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to ignore any terrestrial data available in the respective small area.
+
+\begin{small}
+<<>>=
+synth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+              phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+              small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+                               unbiased = FALSE),
+              exhaustive = truemeans.G,
+              boundary_weights = "boundary_weights")
+@
+\end{small}
+
+\begin{small}
+<<>>=
+synth_3p$estimation
+@
+\end{small}
+
+We see that the \code{threephase()} function returns the sample sizes in the entire inventory area as well as within each small area.
+The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively.
+If we compare the two results, we see that the synthetic estimation again yields a much lower variance than the extended synthetic estimation, but to the cost of ignoring the model uncertainties.
+
+We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is not non-exhaustively available but only at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$.
+To do this, we additionally compute the \textbf{extended pseudo synthetic} estimates.
+As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.
+
+\begin{small}
+<<>>=
+extpsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+                  phase_id = list(phase.col = "phase_id_3p", 
+                                  s1.id = 1, terrgrid.id = 2),
+                  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+                                   unbiased = TRUE),
+                  boundary_weights = "boundary_weights")
+
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsynth_3p$estimation
+@
+\end{small}
+
+Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
+In the first scenario, more precise auxiliary information becomes available at a subset of the allready existing and used sample locations.
+In this case, one simply reclassifies the existing sample locations $x \in s_1$ into $s_0$ sample locations and assigns the $s_1$ indicator to those points where the new variables are available.
+In the second scenario, the auxiliary information is available at all sample locations $x \in s_1$, but the computation of some of the explanatory variables is computationally expensive.
+In this case, a considerable time saving could be realized by restricing their computation to a subset of all $s_1$ sample locations, which consequently leads to a three-phase scenario that is handeled likewise the first scenario.
+A respective example is given in \citet{mandallaz2013b} and \citet{mandallaz2013c}.
+In both the first and second scenario, the original sample size of the two-phase inventory remains unchanged, but the $s_1$ sample locations are splitted into $s_0$ and $s_1$ locations.
+In the third and last scenario, one has already used a set of explanatory variables at $s_1$ sample locations for a time before additional, less precise information becomes available in a much larger quantity over the inventory area.
+It still might yield an increase in estimation precision if this information is considered in the estimation in a much larger sample size than the more precise information (sample fraction).
+In this case, one creates a new phase $s_0$ by densifying the allready existing $s_1$ sample locations (which in practice are often arranged in regular grids) such that $n_0 >> n_1$ and $s_1 \in s_0$.
+
 \newpage

### 2017-08-22 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 22.08.2017
update Andreas 22.08.2017(2)

@@
 
-\section{Twophase Estimators and their Application}
+\section{Two-phase Estimators and their Application}
 \label{sec:twophase_and_appl}
@@
 \subsection{Small Area Estimators}
+\label{sec:twophase_sae}
 
@@
 
-The package \pkg{forestinventory} offers 3 small area estimators.
+The package \pkg{forestinventory} provides 3 small area estimators.
 A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible.
@@
 A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term.
-We call this the \textit{bias} or \textit{residual correction} term.
+This is called the \textit{bias} or \textit{residual correction} term.
 
@@
 
-%\newpage
+\newpage

### 2017-08-23 JStatSoft_forestinventory_master-concordance.tex
update Andreas 23.08.2017

--- 2017-08-22 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
-1 78 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 341:%
+1 83 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 346:%
 175 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Onephase_Estimators and their Application.Rnw:ofs 355:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Onephase_Estimators and their Application.Rnw:ofs 360:%
 1 25 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 381:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 386:%
 190}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 382:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 387:%
 1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 911:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 916:%
 192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 912:%
-1 118 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 179 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1302:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 917:%
+1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1436:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1313:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1447:%
 1 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1327:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1461:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1336:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1470:%
 1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1358:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1492:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1374:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1508:%
 1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1629:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1763:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1633:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1767:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1638:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1772:%
 238 11 1}

### 2017-08-23 rnw/Main structure and modules of the Package.Rnw
update Andreas 23.08.2017

@@
 The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
+%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
 
@@
 If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
-In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
+%In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
 
@@
 $n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
-Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.
+% Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.
 
@@
 In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e.
-the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and $not$ the result of a stochastic process.
+the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process.
 The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism.
@@
 identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.
+The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.
+They are therefore also referred to as \textit{model-assisted}.
+The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
+However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.
+
 % NOTES:
@@
 The following sections describe the mathematical details and the application of the multiphase estimators that are implemented in the \proglang{R} package \pkg{forestinventory}.
-While \cite{mandallaz2008, mandallaz2013c} provides an extensive derivation of all estimators, we will only provide the mathematical formulas that are actually implemented and applied in the functions of the package.
+While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech}
+provides an extensive derivation of all estimators, we will restrict to the mathematical formulas that are actually implemented in the package.
 For the sake of brevity, we will restrict to simple sampling and provide the equations for cluster sampling in the Appendix.

### 2017-08-24 JStatSoft_forestinventory_master-concordance.tex
update Andreas 24.08.2017

--- 2017-08-22 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
-1 78 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 341:%
-175 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Onephase_Estimators and their Application.Rnw:ofs 355:%
-1 25 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 381:%
-190}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 382:%
-1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 911:%
+1 77 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
+175 15 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
+1 125 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 886:%
 192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 912:%
-1 118 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 179 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1302:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 887:%
+1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1406:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1313:%
-1 13 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1327:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1417:%
+1 54 1 1 2 11 0 1 2 32 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1517:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1336:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1526:%
 1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1358:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1374:%
-1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1629:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1564:%
+1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1822:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1633:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1826:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1638:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1831:%
 238 11 1}

### 2017-08-24 rnw/Analysis and Visualization.Rnw
update Andreas 24.08.2017

@@
 If we want to compare the results and performance of different estimation methods and estimators for a given global or small area inventory data set, the function to start with is \code{estTable()}.
-This function restructures the results of \code{onephase}, \code{twophase} and \code{threephase} estimations and merges them into one single data set that provides the basis for further analysis.
+This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis.
 For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} dataset:
@@
                 small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                unbiased = TRUE))
+                                unbiased = TRUE),
+                boundary_weights = "boundary_weights")
 @
@@
                                   areas = c("A", "B", "C", "D"),
-                                  unbiased = FALSE))
+                                  unbiased = FALSE),
+                boundary_weights = "boundary_weights")
 @

### 2017-08-24 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 24.08.2017

@@
 
+% NOTE:
+% - Provide all confidence equations for the estimators: onephase, twophase, threephase
+% - Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
+% - ... and asymptotic of estimators
+% - Interpretation of Confidence Intervals
+%
+
+The estimated variance provides a measure of uncertainty for a given point estimate.
+
+Confidence intervals thus allow for a much more intuitive interpretation of the uncertainties associated to a point estimate, and are also more suited to be incorporated into planning forest management operations.
+
+% Frequentisitsche Interpretation streng genommen die einzig richtige: unter rein hypothetischer Wiederholung der Inventur (Gleicher Schaetzer, gleicher Wald, gleiche Hilfvariablen / Aux.info) werden 95\% der berechneten Intervalle den wahren Wert enthalten.
+Man weiss nun aber - selbst wenn man das machen wuerde (was praktisch ja nie der Fall sein wird) - nie, auf welche der berechneten Intervalle das zutreffen wird.
+Die klassische statistische Interpration ist hier also nicht zulaessig (ungleich: der wahre wert befindet sich mit 95\% Sicherheit in meinem berechneten Intervall).
+% 
+% Da diese strenge Interpretation dem Praktiker nicht viel nuetzen wird (ausser evtl einem Gefuhl der Sicherheit), kann man sich aufgrund der CI-Breite (abhaengig von der Varianz der SChaetzung) an folgender Frage orientieren: "Wuerde ich innerhalb des Vertrauensintervalles zu der selben Entscheidung (z.b.
+Eingriff ja/nein, Vorrat um x\% reduzieren) kommen, oder nicht?" Wenn ja, dann ist die Genauigkeit der SChaetzung bereits ausreichend, wenn nein, dann ist dies die Motivation nach effizien
+% Stichwort: signifikanz vs. relevanz
+
+The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are defined by
+
+\textit{One-Phase Estimation}
+
+\begin{equation}\label{ci_1phase}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
+                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+\end{equation}
+
+% ------------------------------------------ %
+\textit{Two-Phase and Three-Phase Global Estimation}
+
+\begin{equation}\label{ci_2phase_global}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
+                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+\end{equation}
+
+% ------------------------------------------ %
+\textit{Two-Phase and Three-Phase Small Area Estimation}
+
+\begin{equation}\label{ci_2phase_sae}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
+                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+\end{equation}
+
+where $\hat{Y}$ is the respective point estimate, $\var(\hat{Y})$ is the respective estimated variance, and $1-\alpha$ is the confidence level.
+Note that in the three-phase case, $p$ is the number of parameters used in the \textit{full} regression model.
+In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
+
+In the package \pkg{forestinventory}, the confidence levels for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}.
+The function requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
+Optionally, the confidence level $1-\alpha$ can be specified by the argument \code{level} and has a default value of 0.95, which will cause the function to return the 95\% confidence intervall.
+
+The 95\% confidence intervall for the small area estimations by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:
+
+\begin{small}
+<<>>=
+confint(extpsynth_2p)
+@
+\end{small}
+
 \newpage

### 2017-08-24 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 23.08.2017
update Andreas 24.08.2017

@@
 
-We first recalculate the regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ as
+We first recalculate the regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ under the extended auxiliary vector as
 
@@
 
+<<echo=FALSE>>=
+options(width=90) 
+@
+
 \begin{small}

### 2017-08-24 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 24.08.2017

@@
 \subsection{Global Estimators}
+\label{sec:twophase_globest}
 
@@
   \item the \code{data.frame} containing the inventory information (\code{data})
-  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ (\code{terrgrid.id})
+  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ phase (\code{terrgrid.id})
   \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional)
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
   \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & =
  \frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:ext_varexternal_2p_extpsynth}
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extpsynth}
 \end{align}

### 2017-08-24 r/create figures.R

@@ -21,23 +21,25 @@ psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons
 
 
 extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                          formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                          phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                          small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                          unbiased = TRUE))
+                         formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+                         phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                         small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+                                         unbiased = TRUE),
+                         boundary_weights = "boundary_weights")
 
 
 psynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                       formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                       phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                       small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                         unbiased = FALSE))
+                      formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+                      phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                      small_area = list(sa.col = "smallarea", 
+                                        areas = c("A", "B", "C", "D"),
+                                        unbiased = FALSE),
+                      boundary_weights = "boundary_weights")
           
 
 
 grisons.sae.table.CI<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                                 extpsynth_3p, psynth_3p),
-                                key.var = "ci",
                                 sae = TRUE, 
                                 vartypes = c("variance", "g_variance",  "ext_variance"))
 
@@ -80,12 +82,4 @@ ggsave("fig\\plot_error.png",
                legend.title=element_text(size=rel(7)),
                legend.text=element_text(size=rel(6))),
        width = 27, height = 17, dpi = 300)
-
-
-
-
-
-
-
-
 

### 2017-08-24 rnw/Main structure and modules of the Package.Rnw

@@
 If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
-The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{one-phase estimation}) by the \code{onephase()} function.
+The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (\textit{one-phase estimation}) by the \code{onephase()} function.
 All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
-
-% NOTES:
-% - how is the package structured in general? --> likewise a decision tree
-% - for simple and cluster sampling (briefly describe cluster sampling and its specialties)
-% - for non-exhausitve and exhaustive auxiliary information
 
@@
 
-The following sections describe the mathematical details and the application of the multiphase estimators that are implemented in the \proglang{R} package \pkg{forestinventory}.
-While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech}
-provides an extensive derivation of all estimators, we will restrict to the mathematical formulas that are actually implemented in the package.
-For the sake of brevity, we will restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
-Special cases occuring under cluster sampling are described in section \ref{sec:speccas_and_scen}.
+The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
+While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
+We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
+Special cases under cluster sampling are described in section \ref{sec:speccas_and_scen}.
 

### 2017-08-28 JStatSoft_forestinventory_master-concordance.tex
update Andreas 28.08.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
-1 125 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 886:%
+1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
 192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 887:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1406:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1417:%
-1 54 1 1 2 11 0 1 2 32 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1517:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
+1 46 1 1 2 11 0 1 2 35 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1513:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1526:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1522:%
 1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1544:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1564:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1560:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1822:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1818:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1826:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1822:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1831:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1827:%
 238 11 1}

### 2017-08-28 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 28.08.2017(1)

@@
 
-The estimated variance provides a measure of uncertainty for a given point estimate.
+The estimated variance of a point estimate provides us with a measure of estimation uncertainty.
+However, its interpretation is rather diffult to use by practitioners.
+Converting the estimated variance into a 95\% \textit{confidence interval} allows for a much more intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate.
+In the design-based framework, it denotes that if we would repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information but each time using a new independently and randomly selected sample, 95\% of the calculated confidence intervals will contain the true but unknown quantity.
+The confidence level $1-\alpha$ (e.g.
+95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown quantity and is therefore often also referred to as \textit{coverage rate}.
 
-Confidence intervals thus allow for a much more intuitive interpretation of the uncertainties associated to a point estimate, and are also more suited to be incorporated into planning forest management operations.
-
-% Frequentisitsche Interpretation streng genommen die einzig richtige: unter rein hypothetischer Wiederholung der Inventur (Gleicher Schaetzer, gleicher Wald, gleiche Hilfvariablen / Aux.info) werden 95\% der berechneten Intervalle den wahren Wert enthalten.
-Man weiss nun aber - selbst wenn man das machen wuerde (was praktisch ja nie der Fall sein wird) - nie, auf welche der berechneten Intervalle das zutreffen wird.
-Die klassische statistische Interpration ist hier also nicht zulaessig (ungleich: der wahre wert befindet sich mit 95\% Sicherheit in meinem berechneten Intervall).
-% 
-% Da diese strenge Interpretation dem Praktiker nicht viel nuetzen wird (ausser evtl einem Gefuhl der Sicherheit), kann man sich aufgrund der CI-Breite (abhaengig von der Varianz der SChaetzung) an folgender Frage orientieren: "Wuerde ich innerhalb des Vertrauensintervalles zu der selben Entscheidung (z.b.
-Eingriff ja/nein, Vorrat um x\% reduzieren) kommen, oder nicht?" Wenn ja, dann ist die Genauigkeit der SChaetzung bereits ausreichend, wenn nein, dann ist dies die Motivation nach effizien
-% Stichwort: signifikanz vs. relevanz
-
-The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are defined by
+Based on the central limit theorem, it can be assumed that under hypothetical repeated sampling, the point estimates will asymptotically follow a normal distribution.
+\citet{mandallaz2013a} recommended to better base the confidence intervals on the Student's \textit{t} distribution.
+The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates calculated by the package \pkg{forestinventory} are thus defined as \\
 
@@
 
-In the package \pkg{forestinventory}, the confidence levels for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}.
-The function requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-Optionally, the confidence level $1-\alpha$ can be specified by the argument \code{level} and has a default value of 0.95, which will cause the function to return the 95\% confidence intervall.
-
-The 95\% confidence intervall for the small area estimations by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:
+In the package, the confidence levels for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function, and the confidence level $1-\alpha$ specified by the argument \code{level} (default value of 0.95, i.e.
+95\%).
+The 95\% confidence intervall of the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:
 
@@
 
+% Frequentisitsche Interpretation streng genommen die einzig richtige: unter rein hypothetischer Wiederholung der Inventur (Gleicher Schaetzer, gleicher Wald, gleiche Hilfvariablen / Aux.info) werden 95\% der berechneten Intervalle den wahren Wert enthalten.
+Man weiss nun aber - selbst wenn man das machen wuerde (was praktisch ja nie der Fall sein wird) - nie, auf welche der berechneten Intervalle das zutreffen wird.
+Die klassische statistische Interpration ist hier also nicht zulaessig (ungleich: der wahre wert befindet sich mit 95\% Sicherheit in meinem berechneten Intervall).
+% 
+% Da diese strenge Interpretation dem Praktiker nicht viel nuetzen wird (ausser evtl einem Gefuhl der Sicherheit), kann man sich aufgrund der CI-Breite (abhaengig von der Varianz der SChaetzung) an folgender Frage orientieren: "Wuerde ich innerhalb des Vertrauensintervalles zu der selben Entscheidung (z.b.
+Eingriff ja/nein, Vorrat um x\% reduzieren) kommen, oder nicht?" Wenn ja, dann ist die Genauigkeit der SChaetzung bereits ausreichend, wenn nein, dann ist dies die Motivation nach effizien
+% Stichwort: signifikanz vs. relevanz
+
 \newpage

### 2017-08-28 rnw/Main structure and modules of the Package.Rnw
update Andreas 28.08.2017(1)

@@
 A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
-identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.
+identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}.
 The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.

### 2017-08-28 JStatSoft_forestinventory_master-concordance.tex
update Andreas 28.08.2017(2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
-1 46 1 1 2 11 0 1 2 35 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1513:%
+1 44 1 1 2 11 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1485:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1522:%
-1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1544:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1494:%
+1 16 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 62 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1639:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1560:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1655:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1818:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1913:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1822:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1917:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1827:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1922:%
 238 11 1}

### 2017-08-28 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 28.08.2017(2)

@@
 
-The estimated variance of a point estimate provides us with a measure of estimation uncertainty.
-However, its interpretation is rather diffult to use by practitioners.
+Whereas the estimated variance of a point estimate provides a measure of estimation uncertainty, its interpretation is rather diffult to use by practitioners.
 Converting the estimated variance into a 95\% \textit{confidence interval} allows for a much more intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate.
-In the design-based framework, it denotes that if we would repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information but each time using a new independently and randomly selected sample, 95\% of the calculated confidence intervals will contain the true but unknown quantity.
+The correct interpretation of a confidence interval is not that there is a 95\% probability that it contains the true value.
+In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sampling design.
+Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\% confidence intervals from each sample, then 95\% of them are expected to contain the true population parameter.
 The confidence level $1-\alpha$ (e.g.
-95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown quantity and is therefore often also referred to as \textit{coverage rate}.
+95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown  population parameter and is therefore often also referred to as \textit{coverage rate}.
 
@@
 \citet{mandallaz2013a} recommended to better base the confidence intervals on the Student's \textit{t} distribution.
-The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates calculated by the package \pkg{forestinventory} are thus defined as \\
+The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as \\
 
@@
 
-where $\hat{Y}$ is the respective point estimate, $\var(\hat{Y})$ is the respective estimated variance, and $1-\alpha$ is the confidence level.
-Note that in the three-phase case, $p$ is the number of parameters used in the \textit{full} regression model.
+where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level.
+Note that for three-phase sampling, $p$ constitutes the number of parameters used in the \textit{full} regression model.
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
 
-In the package, the confidence levels for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function, and the confidence level $1-\alpha$ specified by the argument \code{level} (default value of 0.95, i.e.
-95\%).
+In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
 The 95\% confidence intervall of the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:
@@
 \end{small}
+
+%  NOTE:
+% - note that the CIs can be used for hypothesis testing, i.e. it can give us an idea wether
+%   an estimate is statistically different from a given value
+% - 
+%
+%
 

### 2017-08-28 rnw/Special Cases and Scenarios.Rnw
update Andreas 09.08.2017(4)
update Andreas 22.08.2017
update Andreas 28.08.2017(2)

--- 2017-08-09 rnw/Calculation of Confidence Intervals.Rnw
@@
 
-\section{Calculation of Confidence Intervals}
-\label{sec:confint_calc}
+\section{Special Cases and Scenarios}
+\label{sec:speccas_and_scen}
+
+% include:
+% - categorical variables (zberg) and interaction terms (not so clear to everyone)
+% - !!! cluster sampling and differences between (p)small and extended (p)small if cluster is not completely inside small area
+% - !!! violation of nested sample design + error handling / data adjustments / printed warnings
+% - !!! missing factor levels in s2-sample that occur in s1 / s0 sample + error warning
+% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
+%
+
+\subsection{Small Area Estimation under Cluster Sampling}
+
+<<echo=FALSE>>=
+extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      cluster = "cluster",
+                      small_area = list(sa.col = "ismallold", areas = c("1"),
+                                        unbiased = TRUE))
+@
+
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+R> extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
++                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
++                      cluster = "cluster",
++                      small_area = list(sa.col = "ismallold", areas = c("1"),
++                                        unbiased = TRUE))
+Warning message:
+At least one terrestrial cluster not entirely included within the small area 1.
+Zero mean residual assumption for small area maybe violated.
+Check mean_Rc_x_hat_G and consider alternative estimator 'psmall'
+\end{Sinput}
+\end{Schunk}
+\end{small}
+
+\begin{small}
+<<>>=
+extpsynth.clust$mean_Rc_x_hat_G
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsmall.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      cluster = "cluster",
+                      small_area = list(sa.col = "ismallold", areas = c("1"),
+                                        unbiased = TRUE),
+                     psmall = TRUE)
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsynth.clust$estimation
+@
+\end{small}
+
+\begin{small}
+<<>>=
+extpsmall.clust$estimation
+@
+\end{small}
+
+% --------------------------------------------------------------------- %
+\subsection{Nesting Violation in Sample Design}
+
+% explain
+% 
+% # Checking the nesting of the sample-design:
+% # 1) --> each s2-point muss have the complete set of s1-auxvars (s1-info) available
+% # 2) --> each s2-point muss have the complete set of s0-auxvars (s0-info) available
+% # 3) --> each s1-point muss have the complete set of s0-auxvars (s0-info) available
 

### 2017-08-28 r/nesting_violation.R

@@ -1,0 +1,237 @@
+
+## Demonstrate error handling in package (examples)
+
+
+# --------------------------------------------------------------------------- #
+# TWO-PHASE
+# --------------------------------------------------------------------------- #
+
+## save grisons as new dataset:
+grisons.n<- grisons
+
+
+## delete explanatory variables from an s2-sample point:
+grisons.n[which(grisons.n$phase_id_2p==2)[1],c(4,5,6,7)]<- NA
+
+tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
+                 data=grisons.n,
+                 phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                 boundary_weights = "boundary_weights")
+
+# Violation:
+# s2-point MUST have all explanatory variables available
+
+# --> error message due to nesting violation: s2 not nested in s1
+
+# error handling:
+# s2 point with missing auxiliary information is deleted from the dataset
+
+# --------------------------------------------------------------------------- #
+
+## save grisons as new dataset:
+grisons.n<- grisons
+
+## delete explanatory variables from an s1-sample point:
+grisons.n[which(grisons.n$phase_id_2p==1)[1],"mean"]<- NA
+
+tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
+              data=grisons.n,
+              phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+              boundary_weights = "boundary_weights")
+
+# Violation:
+# all s1 point must have all aux.variables available
+
+# warning message:
+# --> error message due to missing explanatory variable(s) in s1
+
+# error handling:
+# s1 point with missing auxiliary information is deleted from the dataset
+
+
+# --------------------------------------------------------------------------- #
+
+## save grisons as new dataset:
+grisons.n<- grisons
+
+## delete response value from an s2-sample point:
+grisons.n[which(grisons.n$phase_id_2p==2)[1],"tvol"]<- NA
+
+tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
+              data=grisons.n,
+              phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+              boundary_weights = "boundary_weights")
+
+# Violation:
+# every s2 point must provide a response value
+
+# warning message:
+# --> 
+
+# error handling:
+# s2 point with missing response value recoded to s1-point
+
[…]

### 2017-08-29 JStatSoft_forestinventory_master-concordance.tex
update Andreas 29.08.2017(1)

--- 2017-08-28 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
-1 46 1 1 2 11 0 1 2 35 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1513:%
+1 42 1 1 2 11 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1483:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1522:%
-1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1544:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1492:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 21 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 64 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1698:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1560:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1714:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1818:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1972:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1822:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1976:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1827:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
 238 11 1}

### 2017-08-29 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 29.08.2017(1)

@@
 % - Provide all confidence equations for the estimators: onephase, twophase, threephase
-% - Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
-% - ... and asymptotic of estimators
+% - !!! Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
+% - ... and asymptotics of estimators
 % - Interpretation of Confidence Intervals
@@
 
-Whereas the estimated variance of a point estimate provides a measure of estimation uncertainty, its interpretation is rather diffult to use by practitioners.
-Converting the estimated variance into a 95\% \textit{confidence interval} allows for a much more intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate.
-The correct interpretation of a confidence interval is not that there is a 95\% probability that it contains the true value.
+Converting the estimated variance into a 95\% \textit{confidence interval} (CI) allows for an intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate.
+The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value.
 In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sampling design.
-Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\% confidence intervals from each sample, then 95\% of them are expected to contain the true population parameter.
+Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\%  CIs from each sample, then 95\% of them are expected to contain the true population parameter.
 The confidence level $1-\alpha$ (e.g.
 95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown population parameter and is therefore often also referred to as \textit{coverage rate}.
+For a user, the CI is a measure of precision that can also be used for hypothesis testing, i.e.
+providing evidence that an estimate is statistically different from a given value if that value lies outside the CI boundaries.
 
 Based on the central limit theorem, it can be assumed that under hypothetical repeated sampling, the point estimates will asymptotically follow a normal distribution.
-\citet{mandallaz2013a} recommended to better base the confidence intervals on the Student's \textit{t} distribution.
-The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as \\
+\citet{mandallaz2013a} recommends to better base the confidence intervals on the Student's \textit{t} distribution.
+The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as
 
@@
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
-
 In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-The 95\% confidence intervall of the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:
+The 95\% confidence intervals for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) for example can  be applied by:
 
@@
 \end{small}
-
-%  NOTE:
-% - note that the CIs can be used for hypothesis testing, i.e. it can give us an idea wether
-%   an estimate is statistically different from a given value
-% - 
-%
-%
 

### 2017-08-29 rnw/Main structure and modules of the Package.Rnw
update Andreas 29.08.2017(1)

@@
 The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
-However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.
+However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2013a}.
 

### 2017-08-29 rnw/Special Cases and Scenarios.Rnw
update Andreas 29.08.2017(1)

@@
 % - categorical variables (zberg) and interaction terms (not so clear to everyone)
+% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
 % - !!! cluster sampling and differences between (p)small and extended (p)small if cluster is not completely inside small area
@@
 % - !!! missing factor levels in s2-sample that occur in s1 / s0 sample + error warning
-% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
 %
 
+% --------------------------------------------------------- %
+\subsection{Poststratification}
+
+A special case of multi-phase regression estimators is \textit{poststratification}, which can further be divided into the case of \textit{multi-phase sampling for stratification} and \textit{multi-phase sampling for regression within strata}.
+Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.
+
+To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} dataset into 3 classes:
+
+\begin{small}
+<<>>=
+grisons$stage<- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
+@
+\end{small}
+
+\textit{Two-phase sampling for stratification} is now applied if we only use categorical variables in the model, in this case the factor variable \code{stage} (i.e.
+we have an ANOVA model).
+It basically means that the model predictions are the means of terrestrial response values within each development stage (within-strata means).
+
+\begin{small}
+<<eval=FALSE>>=
+# two-phase sampling for stratification
+twophase(formula=tvol ~ stage,
+         data=grisons,
+         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+         boundary_weights = "boundary_weights")
+@
+\end{small}
+
+\textit{Two-phase sampling for regression within strata} implies the combination of continuous and categorical variables within the model (i.e.
+we have an ANCOVA model).
+If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts.
+If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
+Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators.
+
+\begin{small}
+<<eval=FALSE>>=
+# two-phase sampling for regression within strata:
+twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
+         data=grisons,
+         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+         boundary_weights = "boundary_weights")
+@
+\end{small}
+
+The overall variance of estimators under poststratification will mainly depend on the between and within-strata variation of the model residuals and can therefore be reduced by minimizing the within strata residual square sum.
+In case of poststratification, the g-weight variance estimator has the advantage that the strata weights are estimated from the large sample, whereas the external variance calculates them solely based on the terrestrial sample $s_2$.
+
+% --------------------------------------------------------- %
 \subsection{Small Area Estimation under Cluster Sampling}
+
+% NOTE:
+% 
+%
+%
 

### 2017-08-29 rnw/Twophase_Estimators and their Application.Rnw
test on new branch
test2 newbranch
update Andreas 29.08.2017(1)

@@
 The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
+Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly smaller than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
 

### 2017-08-29 r/special_cases_and_scenarios.R

@@ -1,9 +1,88 @@
+#---------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                  #
+#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                  #
+#                                                                                   #
+#  Code for running all examples given in                                           #
+#  Section 6  "Special Cases and Scenarios"                                         #
+#                                                                                   #
+#---------------------------------------------------------------------------------- #
 
-## Demonstrate error handling in package (examples)
+library(forestinventory)
+
+#####################################################
+#          Poststratification                       #
+#####################################################
+
+## save grisons as new dataset:
+grisons.n<- grisons
+
+## create artificial development stages as categorical (factor) variables
+#  used for poststratification:
+grisons.n$stage<- as.factor(kmeans(grisons.n$mean, centers = 3)$cluster)
+
+## plot:
+ggplot(data=grisons.n, aes(x = mean, y = tvol)) +
+  geom_point(aes(colour = factor(stage)))
+
+
+## apply double sampling for (post)stratification:
+twophase(formula=tvol ~ stage,
+         data=grisons.n,
+         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+         boundary_weights = "boundary_weights")
+
+
+## apply double sampling for regression within (post)strata:
+twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
+         data=grisons.n,
+         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+         boundary_weights = "boundary_weights")
+
+
+
+#####################################################
+#  SMALL AREA ESTIMATION UNDER CLUSTER SAMPLING     #
+#####################################################
+
+
+## apply extended pseudo synthetic estimator under cluster sampling:
+extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+                           phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                           cluster = "cluster",
+                           small_area = list(sa.col = "ismallold", areas = c("1"),
+                                             unbiased = TRUE))
+
+# --> creates warning message: At least one cluster not entirely included within small area
+
+## check mean of residuals in small area:
+extpsynth.clust$mean_Rc_x_hat_G
+
+
+## alternatively apply pseudo small area estimator:
[…]

### 2017-08-29 JStatSoft_forestinventory_master-concordance.tex
update Andreas 29.08.2017(2)

--- 2017-08-28 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
-1 77 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
+1 76 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 339:%
 175 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 355:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 886:%
 192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 887:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1406:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
-1 46 1 1 2 11 0 1 2 35 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1513:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1417:%
+1 42 1 1 2 11 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1482:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1522:%
-1 21 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1544:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1491:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1788:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1560:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1804:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1818:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2062:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1822:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2066:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1827:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2071:%
 238 11 1}

### 2017-08-29 rnw/Special Cases and Scenarios.Rnw
update Andreas 29.08.2017(2)

@@
 % NOTE:
-% 
-%
+% - not going into too much detail
+% - rather briefly describe the important difference to simple sampling needed
+%   to understand the problem
 %
 
+As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample location, i.e.
+sample unit, does no longer consist of one, but multiple spatially agglomerated sample points.
+Whereas the maximal possible number of sample points per cluster $M$ is defined by the user, the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random.
+Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix).
+The package functions allow to consider cluster-structures by storing aspecifying the argument \code{cluster}
+
+For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e.
+only a subset $M_{G}(x) < M(x)$ in included in the small area of interest.
+In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured.
+
+In oder to keep track of such cases, the package \pkg{forestinventory} tells the user to do so by returning a warning message:
 <<echo=FALSE>>=
@@
 @
-
 \begin{small}
@@
 \end{Schunk}
-\end{small}
-
-\begin{small}
-<<>>=
-extpsynth.clust$mean_Rc_x_hat_G
-@
 \end{small}
@@
 
+Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that the point estimates are close and more important, the external as well as the design-based variances only differ marginally.
+This can be taken as a hint that the violation of the zero mean residual property can expected to be negligible in this case.
+We can also check the calculated mean of the residuals for the small area by
+
+\begin{small}
+<<>>=
+extpsynth.clust$mean_Rc_x_hat_G
+@
+\end{small}
+
+The deviation to zero is in this case rather small, especially compared to the magnitude of the point estimate.
+Note however that the zero mean residual property is a mathematical property of an estimator and can not be proven by analyzing the empirical model residuals.
+
 % --------------------------------------------------------------------- %
-\subsection{Nesting Violation in Sample Design}
+\subsection{Violation of Nesting in Sample Design}
 
@@
 
+As explained in section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
+The correct nesting thereby concerns the spatial arrangement of the sample phases (figure \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location.
+For the latter, the package \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing.
+We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data set.
+
+\subsubsection*{Violation 1}
+
+Based on the first nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e.
+terrestrially surveyed) and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model.
+If e.g.
+an $s_2$ and /or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
+
+\begin{small}
+<<eval=FALSE>>=
+## delete "mean" value from an s2-sample point:
+grisons[which(grisons$phase_id_3p==2)[1],"mean"]<- NA
+@
+\end{small}
+
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+R> threephase(formula.s0 = tvol ~ mean, 
++             formula.s1 = tvol ~  mean + stddev + max + q75, 
++             data = grisons,
++             phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
++             boundary_weights = "boundary_weights")
+1 rows deleted due to missingness in the set of auxiliary parameters for the first phase (s0) 
+ (1 terrestrial plots affected by deletion)
+Warning messages:
+1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
+  of the second phase (s1) is missing
+2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
+  of the first phase (s0) is missing
+\end{Sinput}
+\end{Schunk}
+\end{small}
+
+\subsubsection*{Violation 2} % 2) and 3)
+
+However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (here \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
+If this concerns an $s_2$ sample loaction, the associated response variable can no longer be used.
+
+\begin{small}
+<<eval=FALSE>>=
+## delete "q75" value from an s2-sample point:
+grisons[which(grisons$phase_id_3p==2)[1],"q75"]<- NA
+@
+\end{small}
+
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+R> threephase(formula.s0 = tvol ~ mean, 
++             formula.s1 = tvol ~  mean + stddev + max + q75, 
++             data = grisons,
++             phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
++             boundary_weights = "boundary_weights")
+Changed the phase_id for 1 rows to the first phase (s0) due to missingness in the set of 
+  auxiliary parameters for the second phase (s1) (1 terrestrial information no longer usable
+  by this change)
+Warning message:
+In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
+  of the second phase (s1) is missing
+\end{Sinput}
+\end{Schunk}
+\end{small}
+
+% # Violation:
+% # each s2-point muss have the complete set of s1-auxvars (s1-info) available
+% 
+% # error message:
+% # missing expl.variable in s1-aux.set at 1 s2-plot.
+% 
+% # error handling:
+% # since there is still the complete set of s0 available, the phase-id has been changed
+% # to s0 (by this, we could keep the plot); but we thereby can no longer use the available
+% # terrestrial info from this plot --> could hamper sae estimation)
+% 
+% # NOTE: This could violate the "randomness" assumption.
+% #       Be sure that we you have a "completely missing at random" case!!!
+
+\subsubsection*{Violation 3}
+
+If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations is \textbf{deleted} from the data set.
+
+\begin{small}
+<<eval=FALSE>>=
+## delete "mean" value from an s0-sample point:
+grisons[which(grisons$phase_id_3p==2)[0],"mean"]<- NA
+@
+\end{small}
+
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+R> threephase(formula.s0 = tvol ~ mean, 
++             formula.s1 = tvol ~  mean + stddev + max + q75, 
++             data = grisons,
++             phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
++             boundary_weights = "boundary_weights")
+1 rows deleted due to missingness in the set of auxiliary parameters for the first phase (s0) 
+ (0 terrestrial plots affected by deletion)
+\end{Sinput}
+\end{Schunk}
+\end{small}
+
+Note that that all the automatic data adjustments (deletion, recoding) have to be accepted with caution.
+Recapitulating, the unbiasedness of estimators in the design-based framework is based on the randomization of the sample locations.
+This means that every possible location within the forest area $F$ has a inclusion probability greater zero.
+Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.
+If any information should be missing at the sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.
+
+\subsubsection*{Violation 4}
+
+If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \not\in s_2$ sample, and thus missing in the $s_2$ sample.
+Conclusively, an internal regression model can not be calculated and the function \textbf{stops} with the following error message:
+
+\begin{small}
+<<eval=FALSE>>=
+## delete s2-points with "stade"-level '300'
+zberg<- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
+@
+\end{small}
+
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+R> twophase(formula = basal ~ stade + couver + melange,
++           data = zberg,
++           phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
++           cluster = "cluster")
+Error in check.mandatoryInputs(formula, data, phase_id) : 
+ Level '300' of factor variable 'stade' existing in s1(s0)- but not in s2 sample. 
+ Calculation of coefficient not feasible.
+\end{Sinput}
+\end{Schunk}
+\end{small}
+
 \newpage

### 2017-08-29 r/special_cases_and_scenarios.R

@@ -183,7 +183,7 @@ tp<- threephase(formula.s0 = tvol ~ mean,
                 boundary_weights = "boundary_weights")
 
 # Violation:
-# s2 point misses expl.variable used in s1 and s0
+# s2 point misses expl.variable used in s0
 
 # warning message:
 # s2 point misses expl.variable used in s1
@@ -192,6 +192,32 @@ tp<- threephase(formula.s0 = tvol ~ mean,
 # error handling:
 # plot deleted
 
+
+# 1) & 2) ---------------------------------
+
+## save grisons as new dataset:
+grisons.n<- grisons
+
+
+## delete "mean" value from an s2-sample point:
+grisons.n[which(grisons.n$phase_id_3p==0)[1],"mean"]<- NA
+
+
+tp<- threephase(formula.s0 = tvol ~ mean, 
+                formula.s1 = tvol ~  mean + stddev + max + q75, 
+                data = grisons.n,
+                phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+                boundary_weights = "boundary_weights")
+
+# Violation:
+# s2 point misses expl.variable used in s0
+
+# warning message:
+# s2 point misses expl.variable used in s1
+# since s1 nested in s0, this expl.variables is also missing in s0
+
+# error handling:
+# plot deleted
 
 
 # 1) & 2) ---------------------------------
@@ -282,6 +308,7 @@ tp<- threephase(formula.s0 = tvol ~ mean,
 #------------------------------------------------------------------------------#
 
 # Missing factor level in s2-sample:
+zberg.n<- zberg
 
 levels(zberg.n$stade)
 

### 2017-08-30 JStatSoft_forestinventory_master-concordance.tex
update Andreas 30.08.2017(2)

--- 2017-08-28 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
-1 77 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
+1 75 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 338:%
 175 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 354:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 885:%
 192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 886:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1405:%
 194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
-1 44 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1485:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1416:%
+1 42 1 1 2 11 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1481:%
 206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1494:%
-1 16 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 62 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1639:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1490:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1787:%
 216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1655:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1803:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1913:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2061:%
 233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1917:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2065:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1922:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2070:%
 238 11 1}

### 2017-08-30 JStatSoft_forestinventory_master.rnw
update Andreas 17.08.2017
update Andreas 22.08.2017
update Andreas 30.08.2017(2)

@@
 Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double- and triple-sampling regression estimators (so-called \textit{multiphase} forest inventory methods) has proved to be efficient.
-The core concept of multiphase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
+With respect to this objective, the application of double and triple-sampling regression estimators (so-called \textit{multi-phase} forest inventory methods) has proved to be efficient.
+The core concept of multi-phase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
 Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double- and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from the mathematical summary of the estimators to their implementation and application in \proglang{R}.}
+The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling.
+The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double and triple sampling for stratification, regression and regression within strata.
+This article provides a bridge from a mathematical summary of the estimators to their implementation and application in \proglang{R}.}
 
-\Keywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three phase sampling, regression estimators, small area estimation}
+\Keywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
+\Plainkeywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
 
@@
 %\setlength{\parindent}{1em}
+%\setlength{\mathindent}{0pt}
 
@@
 
-\begin{document}
+\begin{document}\sloppy
 \SweaveOpts{concordance=TRUE}
@@
 
-% \SweaveInput{rnw/One-Phase Estimators and their Application.Rnw}
+% \SweaveInput{rnw/Onephase_Estimators and their Application.Rnw}
 
@@
 
-\SweaveInput{rnw/Analysis and Visualization.Rnw}
+% \SweaveInput{rnw/.Rnw}
 
@@
 
-% NOTES:
-% - demonstrate visualization of errors and point-ests + CIs
-% - demonstrate a) releff-calculation and gain-analysis
+\SweaveInput{rnw/Analysis and Visualization.Rnw}
+
+%------------------------------------------------------------------------------------------------%
+
+\SweaveInput{rnw/Acknowledgements.Rnw}
 
@@
 
-\bibliography{bib/literature}\newpage
+\bibliography{bib/literature}
 

### 2017-08-30 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 30.08.2017(2)

@@
 In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-The 95\% confidence intervals for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) for example can  be applied by:
+For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) be applied by:
 

### 2017-08-30 rnw/Introduction.Rnw
fixed spelling error
update Andreas 30.08.2017(2)

@@
 Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multiphase sampling}.
+A method which has become particularly attractive is so called \textit{multi-phase sampling}.
 The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
@@
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling design and thus cover 32 inventory scenarios in total.
+The package comprises two and cover 32 inventory scenarios in total.
 The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.

### 2017-08-30 rnw/Main structure and modules of the Package.Rnw
update Andreas 29.08.2017(2)
update Andreas 30.08.2017(2)

@@
 The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
+The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each location $x \in s_1$.
 These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
@@
 The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
-%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
 
@@
 The phase $s_0$ provides a large number $n_0$ of auxiliary information, whereas the phase $s_1$ provides additional auxiliary information that is only available at $n_1$ subsamples of $s_0$.
-The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and $s_0$.
-The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
+The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and thus $s_0$.
+The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ available in $s_1$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
 From that it follows that we can define \textit{two nested} regression models.
@@
 If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
-%In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
 
@@
 If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be to small to provide sufficient estimation accuracy.
-In this case, the application of multiphase estimation procedures can be efficient.
-However, the terrestrial sample size is too small to fit a separate regression model to them.
-The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and then to apply this  model to the small area.
-A potential model bias within $G$ is then corrected for by using the model residuals derived in the small area $G$.
-If no terrestrial samples fall into a small area $G$ (i.e.
-$n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
-% Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.
+In this case, the application of multi-phase estimation procedures can be particularly efficient.
+However, the terrestrial sample size in $G$ is often too small to fit a separate regression model.
+The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ for model fitting, and then to apply this model in the small area.
+If $G$ provides terrestrial samples, a model bias in $G$ is corrected using the model residuals in $G$, leading to unbiased estimates.
+Otherwise, the resulting estimates can be biased.
 
@@
 They are therefore also referred to as \textit{model-assisted}.
-The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
-However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2013a}.
+The random distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids in order to minimize travelling cost of the terrestrial survey.
+However, there is reasonable and empirical evidence that softening this assumption is acceptable for point as well as variance estimation.
+The latter will in most cases be slightly overestimated \citep{mandallaz2013a}.
 
@@
 \subsection{Package Structure}
+\label{sec:packstruc}
 
@@
 The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (\textit{one-phase estimation}) by the \code{onephase()} function.
-All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
+All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated sample points.
 
@@
 
-The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
+The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
 While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
 We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
-Special cases under cluster sampling are described in section \ref{sec:speccas_and_scen}.
+A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
 

### 2017-08-30 rnw/Special Cases and Scenarios.Rnw
update Andreas 30.08.2017(2)

@@
 For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e.
-only a subset $M_{G}(x) < M(x)$ in included in the small area of interest.
+only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
 In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
@@
 
-Based on the first nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e.
+Based on the nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e.
 terrestrially surveyed) and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model.
@@
 Recapitulating, the unbiasedness of estimators in the design-based framework is based on the randomization of the sample locations.
-This means that every possible location within the forest area $F$ has a inclusion probability greater zero.
+This means that every possible location within the forest area $F$ has an inclusion probability greater zero.
 Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.

### 2017-08-30 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 30.08.2017(2)

--- 2017-08-24 rnw/Twophase_Estimators and their Application.Rnw
@@
 
-The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.
+The vector of regression coefficients of the OLS regression model is found by solving the sample-based normal equation
 
@@
 Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
-This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data.
+This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} function as input data.
 \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g.
@@
 In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}.
+The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area (figure \ref{fig:exh_nexh_bw}).
 In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
 This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$.
-In the \textit{non-exhaustive} estimator case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
+In the \textit{non-exhaustive} case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
 Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}).
@@
 
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package.
+An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{theephase()} function of the package.
 In forest inventories, the sample is often restricted to those sample locations located within the forest area.
@@
 Note that this form results under the particular case where the regression coefficients are derived using the data from the current inventory (i.e.
-\textbf{internal} model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.
+\textit{internal} model).
+In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textit{external} models \citep{mandallaz2016}.
 
@@
 The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}).
-Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
+Note that the external variance assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
+It is usually slightly smaller than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
 
@@
 
-The structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package.
+The structure of the data frame has to meet certain requirements to be usable by the \code{onephase}, \code{twophase()} and \code{threephase()} function of the package.
 Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location.
@@
 \end{small}
-The \code{twophase}-function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
-A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object.
+The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
+A \textbf{summary of the estimation results} can be obtained by passing this object to the \code{S3} generic \code{summary()} function.
 It first provides a summary of the function call, the applied regression formula and the estimation method that was applied.
@@
 If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}.
-The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
+The vector must be input in the same order that the \code{lm()} function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
 We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.
@@
 Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation.
-This is because we took out one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
+This is because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
 

### 2017-08-31 JStatSoft_forestinventory_master-concordance.tex
update Andreas 31.08.2017

--- 2017-08-29 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 123 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
-1 95 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
-139 34 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
-1 77 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
-175 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
+1 104 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 112:%
+1 94 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 207:%
+120 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 217:%
+1 75 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 293:%
+131 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 305:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
-192}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 836:%
+144}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 837:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
-194 10 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1356:%
+146 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1368:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1483:%
-206 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1492:%
-1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 21 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 64 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1698:%
-216 15 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1714:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1433:%
+159 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1443:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1740:%
+170 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1750:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1972:%
-233 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1976:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2008:%
+181 8 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2017:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
-238 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2022:%
+191 11 1}

### 2017-08-31 JStatSoft_forestinventory_master.rnw
update Andreas 31.08.2017

@@
   Telephone: +41/44/632 32 36\\
-  URL: \url{http://www.lue.ethz.ch/people/hilla}
 
@@
 %% for those who use Sweave please include the following line (with % symbols):
-
 \usepackage{Sweave}
@@
 \newcommand{\bwi}{BWI3}
-\newcommand{\LF}{\ensuremath{\lambda(F)}}
-\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
-\newcommand{\EX}{\mathbb{E}}
-\newcommand{\PR}{\mathbb{P}}
 \newcommand{\var}{\mathbb{V}}
-% \newcommand{\COV}{\mathbb{COV}}
-\newcommand{\MAV}{\mathbb{MAV}}
-\newcommand{\MRAV}{\mathbb{MRAV}}
-\newcommand{\POP}{\mathcal{P}}
-\newcommand{\SAMP}{\mathcal{S}}
-\newcommand{\RE}{\mathbb{RE}}
-\newcommand{\PLAN}{\Re^2}
-\newcommand{\SUR}{\mathbb{S}}
-\newcommand{\ING}{\mathbb{I}}
-\newcommand{\DEP}{\mathbb{D}}
-
-%\setlength{\parindent}{1em}
-%\setlength{\mathindent}{0pt}
 
@@
 % NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
 %
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
+%
+%
+%
 
 %------------------------------------------------------------------------------------------------%
-% --------------------------- Structure of the Package ----------------------------------------- %
+% --------------------------- Methods and structure of the Package ----------------------------- %
 
@@
 % NOTES:
-% - Pyramide / decision tree Graphik
-% - Confidence Interval calculation
-% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
+% - mention difference between finite and infinite population apporach ?
 %
-% --> explain each point briefly without to much (mathematical) detail
+%
+%
 
@@
 % NOTES:
-% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
-% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
+%
+%
+%
 %
@@
 % NOTES:
-% - demonstrate calculation of confidence intervals (give the formulas)
+%
+%
 
@@
 % NOTES:
-% - Error handling and adjustments: Sampling designs not nested (always give background and then illustrate handling in package)
-% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
-%                   or s2 or/and s1 sample in 3-phase case
-% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation
-
-%------------------------------------------------------------------------------------------------%
-% ---------------------------------- Special Cases and Scenarios ------------------------------- %
-
-% \SweaveInput{rnw/.Rnw}
+%
+%
+%
 
@@
 
+% NOTES:
+% 
+%
+%
+
 %------------------------------------------------------------------------------------------------%
+% ---------------------------------------- Acknowledgements ------------------------------------ %
 

### 2017-08-31 rnw/Analysis and Visualization.Rnw
update Andreas 31.08.2017

@@
 
-The package \pkg{forestinventory} also provides methods for \textit{analysing} and \textit{visualizing} the estimation results created by the \code{onephase}, \code{twophase} and \code{threephase} functions.
+The package \pkg{forestinventory} also provides methods for \textit{analyzing} and \textit{visualizing} the estimation results created by the \code{onephase()}, \code{twophase()} and \code{threephase()} functions.
 
@@
 
-If we want to compare the results and performance of different estimation methods and estimators for a given global or small area inventory data set, the function to start with is \code{estTable()}.
+If we want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, the function to start with is \code{estTable()}.
 This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis.
-For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} dataset:
+For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} data set:
 
@@
 If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}.
-Note however that the package \pkg{forestinventory} provides several \code{S3} generic functions that particularly respond to the class \code{"esttable"}.
+Note however that the package \pkg{forestinventory} provides several \code{S3} generic methods that particularly respond to the class \code{"esttable"}.
 Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.
@@
 The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
-However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
+However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
 Per default, also the confidence intervals are added.
@@
 
-The package \pkg{forestinventory} also provides a \code{S3} generic plotting function based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
+The package \pkg{forestinventory} also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
 Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function.
@@
 
-If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case for the default settings of \code{estTable()}:
+If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case using the default setting of \code{estTable()}:
 
@@
 
-The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic plot function and setting the argument \code{yvar = "estimate"}:
+The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic \code{plot} method and setting the argument \code{yvar = "estimate"}:
 
@@
 
-If we want the estimation errors to be plotted, we simply leave the \code{yvar} argument to its default (\code{"error"}):
+If we want the estimation errors to be plotted, we simply set the \code{yvar} argument \code{"error"} (default):
 
@@
 
-Note that the \code{plot}-function call can arbitrarily be extended by additional \pkg{ggplot2} functions and specifications.
+Note that the \code{plot()}-function call can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.
 

### 2017-08-31 rnw/Introduction.Rnw
update Andreas 31.08.2017

@@
 They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
+As a full census of all trees within the target forest area only remains feasible for spatially small areas, forest inventories commonly gather their information of interest by means of statistical sampling methods, i.e.
 information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
+This sample is then used to provide a \textit{point estimate} of the true, albeit unknown population parameter of the forest area, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
+The information of terrestrial samples is assumed to be very precise, and increasing the precision of the point estimate - characterized by its \textit{variance} - could primarily be achieved by increasing the terrestrial sample size.
 However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
@@
 
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
+Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
+However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available.
 Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
@@
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two and cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
+The implemented estimators cover 32 inventory scenarios and can be used for stratification, regression and regression within strata \citep{massey2015}.
 The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
 
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
+The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle such situations.
 

### 2017-08-31 rnw/Special Cases and Scenarios.Rnw
update Andreas 31.08.2017

@@
 
-To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} dataset into 3 classes:
+To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 classes:
 
@@
 Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix).
-The package functions allow to consider cluster-structures by storing aspecifying the argument \code{cluster}
+The package functions allow to define cluster-structures by specifying the argument \code{cluster}.
 
@@
 only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
-In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
-In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured.
+In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured under this scenario.
 
@@
 
-Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that the point estimates are close and more important, the external as well as the design-based variances only differ marginally.
-This can be taken as a hint that the violation of the zero mean residual property can expected to be negligible in this case.
+Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the design-based variances only differ marginally.
+This can be taken as evidence that the violation of the zero mean residual property can here expected to be negligible.
 We can also check the calculated mean of the residuals for the small area by
@@
 For the latter, the package \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing.
-We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data set.
+We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data sets.
 
@@
 However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (here \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
-If this concerns an $s_2$ sample loaction, the associated response variable can no longer be used.
+If this concerns an $s_2$ sample location, the associated response variable can no longer be used.
 

### 2017-08-31 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 29.08.2017(2)
update Andreas 31.08.2017

@@
 
-The internal model case again insures that the zero mean residual property over $F$, and the \textbf{point estimates} are thus calculated as
+The internal model case again insures the zero mean residual property over $F$, and the \textbf{point estimates} are thus calculated as
 
@@
 
-Note that the \code{grisons} inventory dataset originally corresponds to a classical two-phase sample design, i.e.
+Note that the \code{grisons} inventory data set originally corresponds to a classical two-phase sample design, i.e.
 the full set of explanatory variables is available at all sample locations in the inventory area.
@@
 
-We can use the \code{table()} function to see the sample sizes of all phases for the three-phase secenario:
+We can use the \code{table()} function to see the sample sizes of all phases for the three-phase scenario:
 
@@
 The large phase $s_0$ comprises $178 + 88 + 40 = 306$ sample locations, since $s_2 \subset s_1 \subset s_0$.
-We now assume that at sample locations $x \in s_0$, only the mean canopy height (\code{mean}) is available, whereas at sample locations $x \in s_1$ we have also have the explanatory variables \code{stddev}, \code{max} and \code{q75} at hand.
+We now assume that at sample locations $x \in s_0$, only the mean canopy height (\code{mean}) is available, whereas at sample locations $x \in s_1$ we additionally have the explanatory variables \code{stddev}, \code{max} and \code{q75} at hand.
 We can now define the reduced and full regression model formulas that we will use in the \code{threephase()} function.
@@
 
-Note that compared to the \code{twophase()} function, we now have to specify the two regression models, i.e.
+Note that compared to the \code{twophase()}-function, we now have to specify two regression models, i.e.
 the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
@@
 
-The summary of a \code{threephase} function recalls both regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
-The set of sample sizes have also been extented by that of the $s_0$ phase (\code{n0}).
+The \code{summary()} of a \code{threephase()}-function recalls both regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
+The set of sample sizes have also been extented by the $s_0$ phase (\code{n0}).
 In order to compare the results to the respective two-phase estimation, we assume that we have the mean canopy height as the only explanatory variable at all 306 sample locations:
@@
 \subsection{Small Area Estimators}
+\label{sec:threephase_sae}
 
@@
 The \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) follow the same principle as in the two-phase case, i.e.
-applying the globally derived reduced and full regression model to the small area means of the explanatory variables, and then correcting for a potential model bias by adding the residual mean of the full model, which are not zero in general, to the point estimate (equations \ref{eq:pe_3p_small} and \ref{eq:pe_3p_psmall}).
-Note that the difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ derived in the $s_1$ phase and the (more) exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ or $\bar{\pmb{Z}}^{(0)}_{0,G}$ derived in the $s_0$ phase is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
+applying the globally derived reduced and full regression model to the small area means of the explanatory variables, and then correcting for a potential model bias by adding the residual mean of the full model, which is not zero in general, to the point estimate (equations \ref{eq:pe_3p_small} and \ref{eq:pe_3p_psmall}).
+Note that the difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ derived in the $s_1$ phase and the more precise mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ or exact mean $\bar{\pmb{Z}}^{(0)}_{0,G}$ derived in the $s_0$ phase is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
 
@@
 
-The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) ca be applied if no terrestrial samples are available in the small area, i.e.
+The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) can be applied if no terrestrial samples are available in the small area, i.e.
 $n_{2,G}=0$.
-Consequently, the residual correction and the residual variation term of the full model can no longer be applied an vanishes from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and design-based variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
+Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and design-based variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
 The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$.
@@
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
-The extended auxiliary vector thus is defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{\mathbb{Z}}^{t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathbb{Z}}^{t}(x)=\pmb{Z}^{t}(x)$.
-This practically means that we have to specify the binary indicator variable for all observations in the data frame that is passed to the package-functions.
+This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
+The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{\mathbb{Z}}^{t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathbb{Z}}^{t}(x)=\pmb{Z}^{t}(x)$.
+This practically means that we have to specify the binary indicator variable for \textit{all} observations in the data frame that is passed to the package-functions.
 
@@
 
-We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the \textbf{extended synthetic} and the \textbf{synthetic} estimator to the \code{grisons} dataset.
+We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the \textbf{extended synthetic} and the \textbf{synthetic} estimator to the \code{grisons} data set.
 The setup is thus exactly the same as in the example for global three-phase estimation (section \ref{sec:glob_est_3p}).
@@
 
-We see that the \code{threephase()} function returns the sample sizes in the entire inventory area as well as within each small area.
+We see that the \code{threephase()}-function returns the sample sizes in the entire inventory area as well as within each small area.
 The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively.
-If we compare the two results, we see that the synthetic estimation again yields a much lower variance than the extended synthetic estimation, but to the cost of ignoring the model uncertainties.
+If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but to the cost of ignoring the model uncertainties.
 
-We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is not non-exhaustively available but only at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$.
+We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$.
 To do this, we additionally compute the \textbf{extended pseudo synthetic} estimates.

### 2017-08-31 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 31.08.2017

@@
 This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables.
-Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
+Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
 Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
@@
   \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
@@
 
-The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables.
-Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
+\textbf{Point estimates} and their associated \textbf{variances} can now be calculated under \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables.
+Note that exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
 Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios.
 In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area (figure \ref{fig:exh_nexh_bw}).
-In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
+The design-based regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area (figure \ref{fig:exh_nexh_bw}).
+In the \textit{exhaustive} case, the explanatory variables are actucally calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
 This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$.
 In the \textit{non-exhaustive} case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
-Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}).
+Consequently, the true mean $\bar{\pmb{Z}}$ has thus to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}).
 As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
@@
 by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}).
-This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
+This method was suggested in \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
 To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}.
@@
 \begin{align}
-  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
+  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\bar{\pmb{Z}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}} \label{eq:gw_var_2p_reg_ex}\\
   \hat{\var}(\hat{Y}_{reg2p,nex}) & :=
@@
 
-The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}).
-Note that the external variance assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
-It is usually slightly smaller than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
+The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}) that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
+Applied under internal models, the external variance is asymptotically unbiases and usually slightly smaller than the design-based variance, where the uncertainty of the regression coefficients is considered in the variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
 
@@
 	\end{subfigure}
-\caption{Concept of using exhaustive (\textbf{(a)}) and non-exhaustive (\textbf{(b)}) auxiliary information including boundary adjustment.
+\caption{Concept of exhaustive \textbf{(a)} and non-exhaustive \textbf{(b)} calculation of explanatory variables including boundary adjustment at the support level.
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (underlying polygon).
+A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
 w: weights used for calculating the weighted mean of each explanatory variable.
@@
 
-To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository.
+To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository.
 The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample).
@@
 The structure of the data frame has to meet certain requirements to be usable by the \code{onephase}, \code{twophase()} and \code{threephase()} function of the package.
-Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location.
+Firstly, a row in the data frame corresponds to a sample location while the columns specify the attributes associated to the respective sample location.
 For global two-phase estimation, we have to specify
@@
   \itemsep0em
-  \item the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}
+  \item the regression model (\code{formula}) as specified in the \code{lm}()-function \citep{R}
   \item the \code{data.frame} containing the inventory information (\code{data})
@@
 The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
-A \textbf{summary of the estimation results} can be obtained by passing this object to the \code{S3} generic \code{summary()} function.
-It first provides a summary of the function call, the applied regression formula and the estimation method that was applied.
+A \textbf{summary of the estimation results} can be obtained by passing this object to the \code{S3} generic \code{summary()}-function.
+The summary informs about the function call, the applied regression formula and the estimation method that was applied.
 The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}).
@@
 If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}.
-The vector must be input in the same order that the \code{lm()} function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
+The vector must be input in the same order that the \code{lm()}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
 We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.
 
-The right order can easily be established by extracting the columnames of the respective \code{lm} output. This step is particularly recommended if categorical variables are used. In our case this is:
+The right order can easily be established by extracting the columnames of the respective \code{lm}-output. This step is particularly recommended if categorical variables are used. In our case this is:
 \begin{small}
@@
 We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
-Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pest_2p_psmall}.
+Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}.
 As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) available at the sample location $s_{1,G}$ in the small area.
-A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term.
+A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the model residuals in $G$ to the first term.
 This is called the \textit{bias} or \textit{residual correction} term.
@@
   \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
-  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pest_2p_psmall}
+  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pepsmall}
 \end{align}
@@
 
-The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textit{small} and the \textit{psmall} estimator.
+The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external variance} (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textit{small} and the \textit{psmall} estimator.
 Again note that all components are restricted to those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
@@
 
-The \textbf{synthetic} and \textbf{pseudo synthetic estimator} is applied if no terrestrial samples are available within the small area $G$ (i.e.
+The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are applied if no terrestrial samples are available within the small area $G$ (i.e.
 $n_{2,G}=0$).
 In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
-However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pest_2p_psmall}) is no longer possible.
+However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}) is no longer possible.
 Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
@@
 
-In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pest_2p_psmall}).
+In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pepsmall}).
 However, the equations would considerably simplify if we could mathematically insure that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value \code{1} if the observation lies inside the small area $G$ and is otherwise set to \code{0}.
+This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the small area $G$ and is otherwise set to 0.
 The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and leads to a classical ANOVA regression model.
@@
 \hat{\var}(\hat{Y}_{G,extpsynth,2p})=
-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}
+\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+\hat{\bar{\pmb{\mathbb{Z}}}}_{G}
 + \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
@@
 
-The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \label{eq:pest_2p_psmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
-However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown that the residual correction term is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is design-unbiased.
+The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
+However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown that the residual correction term is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is asymptotically design-unbiased.
 
@@
 Small area estimations in the package \pkg{forestinventory} can be applied by additionally specifying the optional argument \code{small_area}.
-The input dataset has therefore to be extended by an additional column that specifies for each each row (i.e.
+The input data set has therefore to be extended by an additional column that specifies for each each row (i.e.
 observation of a sample location) the respective small area in which it is located.
@@
 The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}.
-In addition to global estimation, the \code{summary} output and the \code{estimation} object comprise the estimates and the sample sizes (\code{n1G} and \code{n2G}) for all calculated small areas (column \code{area}).
+In addition to global estimation, the \code{summary()}-output and the \code{estimation} object comprise the estimates and the sample sizes (\code{n1G} and \code{n2G}) for all calculated small areas (column \code{area}).
 Note that the sample size information is also particularly provided in the \code{samplesizes} object returned by the function call.
@@
 Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas.
-Note that in the \code{grisons} dataset, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible.
+Note that in the \code{grisons} data set, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible.
 Applying the pseudo synthetic estimator equals treating those small areas as if $n_{2,G}=0$ by ignoring the terrestrial information.
@@
 
-In this case, the variance of the pseudo synthetic estimator is almost only half as high as under the \textit{psmall} and \textit{extended psynth} estimator.
+In this case, the variances of the pseudo synthetic estimator is almost only half the variances of the \textit{psmall} and \textit{extended psynth} estimator.
 The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and \ref{eq:var_2p_extpsynth}).
@@
 The \textbf{exhaustive versions} of the small area estimators (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
-Its application requires that we know the exact means of all explanatory variables within the small area of interest.
+Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest.
 In contrast to the \textit{global} estimators, the exact means have now to be delivered in the form of a \code{data.frame}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
-Note that likewise the case of global estimation, the order of the explanatory variables in the data frame has to match the order in which they appear in the design matrix defined by the \code{lm} function in \proglang{R}.
+Note that likewise the case of global estimation, the order of the explanatory variables in the data frame has to match the order in which they appear in the design matrix defined by the \code{lm()}-function in \proglang{R}.
 In order to tell \proglang{R} which row describes which small area, the row names have to match the respective names of the small areas specified in the \code{areas} argument.
 
-For the \code{grisons} dataset, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by
+For the \code{grisons} data set, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by
 \begin{small}

### 2017-08-31 JStatSoft_forestinventory_master.rnw
Changes to abstract to make more concise.

--- 2017-08-30 JStatSoft_forestinventory_master.rnw
@@
 Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double and triple-sampling regression estimators (so-called \textit{multi-phase} forest inventory methods) has proved to be efficient.
-The core concept of multi-phase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from a mathematical summary of the estimators to their implementation and application in \proglang{R}.}
+They typically consist of a random sample of plot locations in the forest that are assessed individually by hand.
+Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs (e.g.
+LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
+With respect to this objective, the application of \textit{multiphase} forest inventory methods (e.g.
+double- and triple-sampling regression estimators) has proved to be efficient.
+While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent.
+The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
+The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
+This article summarizes the mathematical theory of this family of design-based estimators and demostrates how they can be easily implemented in \proglang{R}.}
 
-\Keywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
+\Keywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
+\Plainkeywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
 

### 2017-08-31 rnw/Introduction.Rnw
Edits to Introduction.

--- 2017-08-30 rnw/Introduction.Rnw
@@
 
-\section{Introdution}
+\section{Introduction}
 \label{sec:intro}
 
-In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods.
-They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
-Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions.
-For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e.
-information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}.
-This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
-The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size.
-However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par
+In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time.
+They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
+As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods.
+Typically this means that discrete sample locations (\textit{sample plots}) are randomly chosen in the forest, making up the framework of a \textit{terrestrial inventory}.
+This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals.
+There is a broad range of literature describing the concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
 
-Developments in recent years thus showed an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is so called \textit{multi-phase sampling}.
+Terrestrial inventories have the benefit of being very flexible in the sense that they can be used to produce high quality estimates for a wide-variety of different forest attributes.
+However, they have the downside of being very expensive.
+In order to increase the precision of the estimates, the number of sample plots must be increased, which essentially means travel costs will ballon as trained inventorists are sent to more and more random plot locations.
+This is why the number of terrestrial samples is usually limited.
+Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause undesirably high estimation errors.
+As a result, in recent years there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
+A method which has become particularly attractive is called \textit{multiphase sampling}.
 The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
@@
 % section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
+Despite these promising methods, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
 One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
+However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing, which was the motivation for developing the \proglang{R} package \pkg{forestinventory}.
 The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two and cover 32 inventory scenarios in total.
-The estimators can be used for stratification, regression and regression within strata \citep{massey2015}.
+The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling designs and cover 32 inventory scenarios in total.
+The estimators can be used for post-stratification, regression and regression within strata \citep{massey2015}.
 The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
 
-The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
+The objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
 rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).

### 2017-08-31 rnw/Special Cases and Scenarios.Rnw
update Andreas 31.08.2017(2)

@@
 <<eval=FALSE>>=
-## delete "mean" value from an s2-sample point:
+## delete "mean" value from an s2- (i.e. s1- and s0-) sample point:
 grisons[which(grisons$phase_id_3p==2)[1],"mean"]<- NA
@@
 ## delete "mean" value from an s0-sample point:
-grisons[which(grisons$phase_id_3p==2)[0],"mean"]<- NA
+grisons[which(grisons$phase_id_3p==0)[1],"mean"]<- NA
 @

### 2017-08-31 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 31.08.2017(2)

@@
 
-The \textbf{extended pseudo small area estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
+The \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
 \begin{small}

### 2017-08-31 r/confidence_intervals.R

@@ -1,0 +1,92 @@
+#---------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                  #
+#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                  #
+#                                                                                   #
+#  Code for running all examples given in                                           #
+#  Section 5  "Calculation of Confidence Intervals"                                 #
+#                                                                                   #
+#---------------------------------------------------------------------------------- #
+
+library(forestinventory)
+
+
+#####################################################
+# ----- CI for One-Phase Estimation --------------- #
+#####################################################
+
+## load grisons dataset:
+data(grisons)
+
+## 1) calculate onephase-estimation for entire dataset:
+op <- onephase(formula = tvol~1 ,data = grisons,
+               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
+summary(op)
+confint(op)
+
+## 2) calculate onephase-estimation for given domains (areas) in dataset:
+op.a <- onephase(formula = tvol~1,
+                 data = grisons,
+                 phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                 area = list(sa.col = "smallarea", areas = c("A", "B")))
+summary(op.a)
+confint(op.a)
+
+
+#####################################################
+# ----- CI for Two-Phase Estimation --------------- #
+#####################################################
+
+
+extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
+                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                        small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
+                                          unbiased = TRUE))
+
+confint(extpsynth_2p)
+
+## CIs with bonferroni correction
+confint(extpsynth_2p, adjust.method = "bonferroni")
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+#######
+# END #
+#######
+
+

### 2017-08-31 r/special_cases_and_scenarios.R

@@ -26,14 +26,21 @@ ggplot(data=grisons.n, aes(x = mean, y = tvol)) +
   geom_point(aes(colour = factor(stage)))
 
 
-## apply double sampling for (post)stratification:
+## apply 'double sampling for (post)stratification':
 twophase(formula=tvol ~ stage,
          data=grisons.n,
          phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
          boundary_weights = "boundary_weights")
 
 
-## apply double sampling for regression within (post)strata:
+## apply 'double sampling for regression':
+twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
+         data=grisons.n,
+         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+         boundary_weights = "boundary_weights")
+
+
+## apply 'double sampling for regression within (post)strata'":
 twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
          data=grisons.n,
          phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@ -88,8 +95,7 @@ extpsmall.clust$estimation
 ## save grisons as new dataset:
 grisons.n<- grisons
 
-
-## delete explanatory variables from an s2-sample point:
+## delete explanatory variables from an s2-(i.e. s1-) sample point:
 grisons.n[which(grisons.n$phase_id_2p==2)[1],c(4,5,6,7)]<- NA
 
 tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
@@ -100,7 +106,7 @@ tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
 # Violation:
 # s2-point MUST have all explanatory variables available
 
-# --> error message due to nesting violation: s2 not nested in s1
+# --> error message due to nesting violation: s2 not nested
 
 # error handling:
 # s2 point with missing auxiliary information is deleted from the dataset
@@ -144,9 +150,6 @@ tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
 # Violation:
 # every s2 point must provide a response value
 
-# warning message:
-# --> 
-
 # error handling:
 # s2 point with missing response value recoded to s1-point
 
@@ -171,37 +174,8 @@ tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
 ## save grisons as new dataset:
 grisons.n<- grisons
 
-
[…]

### 2017-08-31 r/two_phase_estimators.R

@@ -1,0 +1,152 @@
+#---------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                  #
+#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                  #
+#                                                                                   #
+#  Code for running all examples given in                                           #
+#  Section 3  "Two-phase Estimators and their Application"                          #
+#                                                                                   #
+#---------------------------------------------------------------------------------- #
+
+library(forestinventory)
+
+#####################################################
+# -------- Global Estimations --------------------- #
+#####################################################
+
+
+## load grisons data set:
+data("grisons", package = "forestinventory")
+
+## see structure of grisons:
+str(grisons)
+
+## see first 12 rows of grisons:
+head(grisons, n = 12)
+
+
+# --------------------------------- #
+## non-exhaustive estimator 
+#  with boundary weight adjustment
+
+## apply estimator:
+reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
+                      data=grisons,
+                      phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights")
+
+## get summary of estimation results:
+summary(reg2p_nex)
+
+
+
+# --------------------------------- #
+## exhaustive estimator 
+#  with boundary weight adjustment
+
+## retrieve order of columnames that the lm()-function processes a formula object:
+#  --> actually the order they appear in the design-matrix ...
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+
+## store true means of explanatory variables:
+true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
+
+## apply estimator:
+reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                     data = grisons,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                     exhaustive = true.means.Z)
+
+## get summary of estimation results:
+summary(reg2p_ex)
+
+## get data frame with estimation results
+reg2p_ex$estimation
+
+
+
+#####################################################
+# ------ Small Area Estimations ------------------- #
[…]

### 2017-09-01 r/confidence_intervals.R
update Andreas 01.09.2017(1)

@@ -18,18 +18,20 @@ library(forestinventory)
 ## load grisons dataset:
 data(grisons)
 
-## 1) calculate onephase-estimation for entire dataset:
+## calculate onephase-estimation for entire dataset:
 op <- onephase(formula = tvol~1 ,data = grisons,
                phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-summary(op)
+
+## Confidence Intervals:
 confint(op)
 
-## 2) calculate onephase-estimation for given domains (areas) in dataset:
+## 2) calculate one-phase-estimation for given domains (areas) in dataset:
 op.a <- onephase(formula = tvol~1,
                  data = grisons,
                  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                  area = list(sa.col = "smallarea", areas = c("A", "B")))
-summary(op.a)
+
+## Confidence Intervals:
 confint(op.a)
 
 
@@ -37,56 +39,22 @@ confint(op.a)
 # ----- CI for Two-Phase Estimation --------------- #
 #####################################################
 
-
+## calculate two-phase-estimation for entire dataset:
 extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                         phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                         small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
                                           unbiased = TRUE))
-
+## Confidence Intervals:
 confint(extpsynth_2p)
 
-## CIs with bonferroni correction
+## change confidence level:
+confint(extpsynth_2p, level = 0.9)
+
+## CIs with bonferroni correction to better meet overall coverage probability
+#  (multiple testing problem if a large number of CIs are compared to each other)
 confint(extpsynth_2p, adjust.method = "bonferroni")
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
 
 
 #######
 # END #
 #######
-
-

### 2017-09-01 rnw/Introduction.Rnw
More changes to improve the flow in first 3 sections.

@@
 In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time.
-They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on national- and international level.
+They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on the national and international levels.
 As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods.
@@
 This is why the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause undesirably high estimation errors.
+Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, areas with smaller sample sizes, such as forest management units, often are associated with undesirably high estimation errors.
 As a result, in recent years there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
@@
 These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
+Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
 Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
@@
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three phase sampling regression estimators for global and small area estimations under simple and cluster sampling designs and cover 32 inventory scenarios in total.
+The package comprises two- and three-phase sampling regression estimators for global and small area estimations under simple and cluster sampling designs and cover 32 inventory scenarios in total.
 The estimators can be used for post-stratification, regression and regression within strata \citep{massey2015}.
@@
 
-The objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
+The objective of this article is a) to establish the link between the mathematical formulation of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
 rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).

### 2017-09-01 rnw/Main structure and modules of the Package.Rnw
More changes to improve the flow in first 3 sections.

--- 2017-08-28 rnw/Main structure and modules of the Package.Rnw
@@
 
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}).
+The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}).
 The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
@@
 the timber volume density per hectare).
-The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model.
-The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.
+The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS).
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases (i.e.
+$\EX{(Y(x)-\hat{Y}(x))}$) using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed.
+In the design-based context, the two-phase estimator is unbiased regardless of the model you use to produce the predictions.
+This property comes from the assumption that each phase's sample is selected via simple random sampling.
+It should be noted that most forest inventories use systematic grids rather than simple random sampling, as is depicted in Figure \ref{fig:concmphase_and_sae_a}.
+However, this is generally considered to be acceptable as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
 %In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
@@
 
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (figure \ref{fig:concmphase_and_sae_a}).
+The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Figure \ref{fig:concmphase_and_sae_a}).
 The basic assumption is that auxiliary information is available in two different frequencies, i.e.
@@
 Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(0)}(x)$ in the so called \textit{reduced model} at the sample locations $x \in s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{0}(x)$.
-If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
+If there is a significant gain in model precision between the reduced and the full model and the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
 %In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
@@
 
-We enter the topic of small area estimation if we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (figure \ref{fig:concmphase_and_sae_b}).
-If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be to small to provide sufficient estimation accuracy.
+Small area estimation does not necessarily refer to small spatial areas but rather areas that contain little or no terrestrial sample.
+In the design-based approach, if there are no terrestrial plots in the small area then there is nothing you can do and you will be forced to depend on the model you choose and accept its potential bias.
+To formulate this mathematically, we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (Figure \ref{fig:concmphase_and_sae_b}).
+If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be too small to provide sufficient estimation precision (i.e.
+the confidence intervals will be very large).
 In this case, the application of multiphase estimation procedures can be efficient.
-However, the terrestrial sample size is too small to fit a separate regression model to them.
+However, the terrestrial sample size is too small to justify fitting a separate regression model just that area.
 The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and then to apply this  model to the small area.
-A potential model bias within $G$ is then corrected for by using the model residuals derived in the small area $G$.
+The potential bias of applying that model in $G$ is then corrected for by using the model residuals derived from the small area $G$.
 If no terrestrial samples fall into a small area $G$ (i.e.
@@
 
-The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories.
-However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest).
-As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e.
-the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable.
-Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process.
-In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e.
-the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process.
+The selection of a suitable predictive model is an essential part in the application of multiphase inventories.
+If the model is chosen poorly it could lead to inefficient use of the auxiliary information or even biased estimates.
+This brings us to the two schools of thought concerning estimation: \textit{model-dependent} (also known as \textit{model-based}) and design-based.
+In the \textit{model-dependent} framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process.
+The model does not need to be fit from a probability sample, but it does need to be trusted in the sense that all of its model assumptions need to be verified.
+For example, if we select a linear model using OLS then there is an implicit mathematical assumption that the residuals are \textit{iid}.
+This model assumption needs to be empirically verified in order to avoid misinterpreting the regression coefficients and even the predictions.
+The unbiasedness of model-dependent estimators depend on the correct specification of the model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process.
+In contrast, the \textit{design-based} (also called \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$.
+At any given location $x_0$, the ground truth, $Y(x_0)$, is assumed to be fixed and \textit{not} the result of a stochastic process.
+$Y(x)$ is random only because of the random selection of the plot locations $x$.
+but the selection of     forest itself is fixed .
 The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism.
 A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
-identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}.
+identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.
 The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.
@@
 In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
-From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}).
+From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
 The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.

### 2017-09-04 rnw/Introduction.Rnw
update Andreas .4.09.2017(1)

--- 2017-08-31 rnw/Introduction.Rnw
@@
 
-\section{Introdution}
+\section{Introduction}
 \label{sec:intro}
@@
 
-Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare.
+Despite the broad range of methodological literature and case studies, freely available software supporting the application of the well-documented multi-phase sampling methods in forest practice has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
@@
 
-The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle such situations.
+The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle them.
 

### 2017-09-04 JStatSoft_forestinventory_master.rnw
update Andreas 31.08.2017(2)
some changes

@@
 The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
-This article summarizes the mathematical theory of this family of design-based estimators and demostrates how they can be easily implemented in \proglang{R}.}
+This article summarizes the mathematical theory of this family of design-based estimators and demonstrates how they can be easily implemented in \proglang{R}.}
 

### 2017-09-04 rnw/Main structure and modules of the Package.Rnw
update Andreas .4.09.2017(1)
update Andreas 04.09.2017(3)
some changes

@@
 At any given location $x_0$, the ground truth, $Y(x_0)$, is assumed to be fixed and \textit{not} the result of a stochastic process.
-$Y(x)$ is random only because of the random selection of the plot locations $x$.
-but the selection of     forest itself is fixed .
-The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism.
-A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
+$Y(x)$ is random only because of the random selection of the plot locations $x$, which is determined by the sampling design.
+A consequence of this approach is that the models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
 identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.

### 2017-09-04 JStatSoft_forestinventory_master-concordance.tex

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1750:%
-1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2008:%
+1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
 181 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2017:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1990:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2022:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1995:%
 191 11 1}

### 2017-09-04 rnw/Calculation of Confidence Intervals.Rnw

@@
 The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value.
-In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sampling design.
+In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sample design.
 Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\%  CIs from each sample, then 95\% of them are expected to contain the true population parameter.
@@
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
-In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) be applied by:
+In the package \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
+For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) are calculated by:
 

### 2017-09-04 rnw/Twophase_Estimators and their Application.Rnw

@@
 
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{theephase()} function of the package.
+An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{threephase()} function of the package.
 In forest inventories, the sample is often restricted to those sample locations located within the forest area.

### 2017-09-04 JStatSoft_forestinventory_master.rnw

--- 2017-08-31 JStatSoft_forestinventory_master.rnw
@@
 \author{Andreas Hill \And Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
+\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories}
 
@@
 Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.
-The information is collected at discrete sample locations distributed over the forest area by means of statistical sampling methods.
-This sample is then used to provide estimates of a target variable for a given spatial unit.
-Due to the high costs of the terrestrial campaigns, there is a need for alternative methods that can provide a) the same estimation precision at lower costs or b) higher estimation precision at identical costs.
-With respect to this objective, the application of double and triple-sampling regression estimators (so-called \textit{multi-phase} forest inventory methods) has proved to be efficient.
-The core concept of multi-phase methods is to combine the terrestrial sample with a much larger sample of target variable \textit{predictions} based on \textit{auxiliary information} that is available in high quantity and low costs.
-Whereas these methods have been successfully applied in practice, the availability of repective open-source software has been rare up to non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling.
-The implemented methods have been demonstrated in various scientific studies covering small to large scale forest inventories, and can be used for double and triple sampling for stratification, regression and regression within strata.
-This article provides a bridge from a mathematical summary of the estimators to their implementation and application in \proglang{R}.}
+They typically consist of a random sample of plot locations in the forest that are assessed individually by hand.
+Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs (e.g.
+LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
+With respect to this objective, the application of \textit{multiphase} forest inventory methods (e.g.
+double- and triple-sampling regression estimators) has proved to be efficient.
+While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent.
+The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
+The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
+This article summarizes the mathematical theory of this family of design-based estimators and demonstrates how they can be easily implemented in \proglang{R}.}
 
-\Keywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
-\Plainkeywords{forest inventory, design-based, infinite population approach, two and three-phase sampling, regression estimators, small area estimation}
+\Keywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
+\Plainkeywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
 
@@
 % NOTES:
-% - mention difference between finite and infinite population apporach ?
+% - mention difference between finite and infinite population approach ?
 %

### 2017-09-04 rnw/Main structure and modules of the Package.Rnw

@@
 		\centering
-    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(2).eps}}%
+    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_a}
@@
 		\centering
-   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(2).eps}}%
+   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_b}
@@
 \centering
-\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_2.eps}}
-\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}}
+\resizebox{0.97\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
+\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
 \label{fig:struct_package}

### 2017-09-04 JStatSoft_forestinventory_master-concordance.tex
update Andreas 04.09.2017(4)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 104 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 112:%
-1 94 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 207:%
-120 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 217:%
-1 75 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 293:%
-131 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 305:%
+1 105 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
+1 97 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 211:%
+121 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
+1 88 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 310:%
+132 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 322:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 836:%
-144}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 837:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 853:%
+145}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 854:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1356:%
-146 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1368:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1373:%
+147 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1385:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1433:%
-159 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1443:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1450:%
+160 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1460:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1740:%
-170 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1750:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1757:%
+171 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1767:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
-181 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1990:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1998:%
+182 8 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2007:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1995:%
-191 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2012:%
+192 11 1}

### 2017-09-04 JStatSoft_forestinventory_master-concordance.tex
update Andreas 04.09.2017(5)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
-1 88 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 310:%
+1 87 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 309:%
 132 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 322:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 321:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 853:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 852:%
 145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 854:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 853:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1373:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1372:%
 147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1385:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1384:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1450:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1449:%
 160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1460:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1459:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1757:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1756:%
 171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1767:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1766:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1998:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1997:%
 182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2007:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2006:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2012:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2011:%
 192 11 1}

### 2017-09-04 JStatSoft_forestinventory_master.rnw
update Andreas for alex_version 2
update Andreas 04.09.2017(5)

@@
 The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
-This article summarizes the mathematical theory of this family of design-based estimators and demonstrates how they can be easily implemented in \proglang{R}.}
+This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in \proglang{R}.}
 
@@
 \newcommand{\var}{\mathbb{V}}
+\newcommand{\EX}{\mathbb{E}}
 

### 2017-09-04 rnw/Introduction.Rnw
update Andreas 04.09.2017(5)

--- 2017-09-01 rnw/Introduction.Rnw
@@
 % section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite these promising methods, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software.
-One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach.
-However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing, which was the motivation for developing the \proglang{R} package \pkg{forestinventory}.
+Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
+However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available.
+Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
 The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The package comprises two- and three-phase sampling regression estimators for global and small area estimations under simple and cluster sampling designs and cover 32 inventory scenarios in total.
-The estimators can be used for post-stratification, regression and regression within strata \citep{massey2015}.
+The implemented estimators cover 32 inventory scenarios and can be used for stratification, regression and regression within strata \citep{massey2015}.
 The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
 
-The objective of this article is a) to establish the link between the mathematical formulation of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e.
-rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).
+The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle them.
 

### 2017-09-04 rnw/Main structure and modules of the Package.Rnw
update Andreas for alex_version 2
update Andreas 04.09.2017(5)

@@
 The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS).
-The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases (i.e.
-$\EX{(Y(x)-\hat{Y}(x))}$) using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed.
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e.
+$\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed.
 In the design-based context, the two-phase estimator is unbiased regardless of the model you use to produce the predictions.
@@
 The unbiasedness of model-dependent estimators depend on the correct specification of the model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process.
-In contrast, the \textit{design-based} (also called \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$.
-At any given location $x_0$, the ground truth, $Y(x_0)$, is assumed to be fixed and \textit{not} the result of a stochastic process.
+In contrast, the \textit{design-based} approach rests upon the randomization of the sample locations $x$.
+At any given location $x$, the ground truth, $Y(x)$, is assumed to be fixed and \textit{not} the result of a stochastic process.
 $Y(x)$ is random only because of the random selection of the plot locations $x$, which is determined by the sampling design.
@@
 \centering
-\resizebox{0.97\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
+\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
 \caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}

### 2017-09-06 JStatSoft_forestinventory_master-concordance.tex
update Andreas 06.09.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
-1 87 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 309:%
+1 82 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 304:%
 132 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 321:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 316:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 852:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 847:%
 145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 853:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 848:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1372:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1367:%
 147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1384:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1379:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1449:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1444:%
 160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1459:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1454:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1756:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
 171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1766:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1761:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1997:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1992:%
 182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2006:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2001:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2011:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2006:%
 192 11 1}

### 2017-09-06 rnw/Acknowledgements.Rnw
update Andreas 22.08.2017
update Andreas 06.09.2017(1)

@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
@@
 Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study and providing the possibility of working on the package.
-We also want to thank Daniel Mandallaz for his support in completing the range of the allready published estimators in the frame of the three-phase small area estimators, many helpful discussions and advice throughout the implementation of our package, and for proofreading the manuscript.
-Our thanks also go to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of grisons for providing the example data.
+We also want to thank Daniel Mandallaz for his support in completing the range of the allready published estimators in the frame of the three-phase small area estimators, many helpful discussions and advice throughout the implementation of our package.
+Our thanks also go to Meinrad Abegg for proofreading the manuscript, and to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of grisons for providing the example data.

### 2017-09-06 rnw/Main structure and modules of the Package.Rnw
update Andreas 06.09.2017(1)

@@
 the timber volume density per hectare).
-The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS).
+The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS) technique.
 The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e.
-$\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed.
-In the design-based context, the two-phase estimator is unbiased regardless of the model you use to produce the predictions.
-This property comes from the assumption that each phase's sample is selected via simple random sampling.
-It should be noted that most forest inventories use systematic grids rather than simple random sampling, as is depicted in Figure \ref{fig:concmphase_and_sae_a}.
-However, this is generally considered to be acceptable as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
+$\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.
+In the design-based context, the two-phase estimator is unbiased regardless of the model used to produce the predictions.
+This property comes from the assumption that each phase's sample is selected via simple random sampling (see section \ref{sec:db_vs_md}).
 %In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
@@
 \subsection{Design-based vs. model-dependent approach}
+\label{sec:db_vs_md}
 
-The selection of a suitable predictive model is an essential part in the application of multiphase inventories.
-If the model is chosen poorly it could lead to inefficient use of the auxiliary information or even biased estimates.
-This brings us to the two schools of thought concerning estimation: \textit{model-dependent} (also known as \textit{model-based}) and design-based.
-In the \textit{model-dependent} framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process.
-The model does not need to be fit from a probability sample, but it does need to be trusted in the sense that all of its model assumptions need to be verified.
-For example, if we select a linear model using OLS then there is an implicit mathematical assumption that the residuals are \textit{iid}.
-This model assumption needs to be empirically verified in order to avoid misinterpreting the regression coefficients and even the predictions.
-The unbiasedness of model-dependent estimators depend on the correct specification of the model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process.
-In contrast, the \textit{design-based} approach rests upon the randomization of the sample locations $x$.
-At any given location $x$, the ground truth, $Y(x)$, is assumed to be fixed and \textit{not} the result of a stochastic process.
-$Y(x)$ is random only because of the random selection of the plot locations $x$, which is determined by the sampling design.
+The selection of a suitable prediction model is an essential part in the application of multiphase inventories.
+However, the requirements on the model differ fundamentally between two major mathematical approaches: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
+One consequence is that the model does not need to be fit from a probability sample, i.e.
+the sample locations could arbitrarily be chosen.
+However, another consequence is that for unbiasedness estimates it has to be insured that the model adequately describes the underlying stochastic process, in other words: the model assumptions have to be met and empirically verified.
+For example, if we use an ordinary least square regression model, implicit mathematical assumptions are (i) normality and homogeneity of the  target variable \textit{Y}(\textit{x}) at each value of the explanatory variables, and (ii) independence of the target variable \textit{Y}(\textit{x}) or the model residuals $R(x)$, i.e.
+\textit{Y}(\textit{x}) and $R(x)$ are assumed to be identically distributed, independent random variables and not to be correlated in any way.
+Whereas verifying the validity of model assumptions can be seen as evidence that the model adequately describes the underlying stochastic process, a proof remains infeasible.
+This leads to the situation that in the model-dependent framework, one has to \textit{trust} the model.
+In contrast to the model-dependent approach, the \textit{design-based} approach rests upon the randomization of the sample locations $x$.
+While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the observation $Y(x)$ at any location $x$ is fixed and \textit{not} the result of a stochastic process.
+Conclusively, the observation \textit{Y}(\textit{x}) still remains a random variable, but solemnly due to the random sample mechanism.
 A consequence of this approach is that the models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
-identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}.
+identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-dependent framework \citep{mandallaz2008}.
 The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.
 They are therefore also referred to as \textit{model-assisted}.
-The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
-However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.
-
-% NOTES:
-% - we can still describe it a bit better, also with some citations of mcroberts and breidenbach (their md-description)
+It should be noted that the random distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
+However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
+However, the variance will in most cases be slightly overestimated and thus lead to wider and thus more conservative confidence intervals \citep{mandallaz2013a}.
 
@@
 The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.
-The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global estimators}) or only for subregions of the entire inventory area (\textbf{small area estimators}).
+The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}).
 The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area.
@@
 All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
+The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
+While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
+We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
+A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
+
+\newpage
 
@@
 
-The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
-While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
-We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
-A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
-
 \newpage

### 2017-09-07 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 07.09.2017(1)

@@
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
+A set $\pmb{Z}$ of $p$ explanatory variables $z_{1},...,z_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
 w: weights used for calculating the weighted mean of each explanatory variable.

### 2017-09-07 r/special_cases_and_scenarios02.R
update Andreas07.09.2017(2)

@@ -10,77 +10,6 @@
 
 library(forestinventory)
 
-#####################################################
-#          Poststratification                       #
-#####################################################
-
-## save grisons as new dataset:
-grisons.n<- grisons
-
-## create artificial development stages as categorical (factor) variables
-#  used for poststratification:
-grisons.n$stage<- as.factor(kmeans(grisons.n$mean, centers = 3)$cluster)
-
-## plot:
-ggplot(data=grisons.n, aes(x = mean, y = tvol)) +
-  geom_point(aes(colour = factor(stage)))
-
-
-## apply 'double sampling for (post)stratification':
-twophase(formula=tvol ~ stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
-
-
-## apply 'double sampling for regression':
-twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
-
-
-## apply 'double sampling for regression within (post)strata'":
-twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
-
-
-
-#####################################################
-#  SMALL AREA ESTIMATION UNDER CLUSTER SAMPLING     #
-#####################################################
-
-
-## apply extended pseudo synthetic estimator under cluster sampling:
-extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
-                           phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                           cluster = "cluster",
-                           small_area = list(sa.col = "ismallold", areas = c("1"),
-                                             unbiased = TRUE))
-
-# --> creates warning message: At least one cluster not entirely included within small area
-
[…]

### 2017-09-07 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 07.09.2017(3)

@@
 reg2p_nex<- twophase(formula = formula.rm, data = grisons,
-                       phase_id = list(phase.col="phase_id_3p", terrgrid.id = 2),
+                       phase_id = list(phase.col="phase_id_2p", terrgrid.id = 2),
                        boundary_weights = "boundary_weights")

### 2017-09-08 rnw/Special Cases and Scenarios.Rnw
update Andreas 08092017

@@
 
-To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 classes:
+To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
 
@@
 If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
-Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators.
+Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables.
 
@@
 
-As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample location, i.e.
-sample unit, does no longer consist of one, but multiple spatially agglomerated sample points.
-Whereas the maximal possible number of sample points per cluster $M$ is defined by the user, the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random.
-Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix).
-The package functions allow to define cluster-structures by specifying the argument \code{cluster}.
+As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample unit does no longer consist of one, but multiple spatially agglomerated sample points.
+One still randomly places the sample location $x$ in the inventory area likewise in the simple sampling design, but then creates $M-1$ additional sample locations $x_2,...,x_M$ close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
+The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign.
+Not however that the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and that this average value is again referred to exactly one point, i.e.
+the cluster origin $x$, in the frame of the infinite population approach.
+Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix).
+Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e defined by the user), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
+The package functions of \pkg{forstinventory} also allow to consider cluster-structures by specifying a unique cluster identification (cluster ID) for each cluster in an additional column whose column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.
 

### 2017-09-08 r/application_threephase.R

@@ -1,106 +1,0 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 4.1  "Application of three-phase global and small area estimators"       #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-
-#####################################
-# GLOBAL THREE-PHASE ESTIMATIONS    #
-#####################################
-
-## get sample sizes of three-phase secenario:
-table(grisons$phase_id_3p)
-
-## define reduced and full regression model:
-formula.s0 <- tvol ~ mean # reduced model:
-formula.s1 <- tvol ~ mean + stddev + max + q75 # full model
-
-## calculate one-phase estimation:
-op <- onephase(formula = tvol~1 ,data = grisons,
-               phase_id =list(phase.col = "phase_id_3p",terrgrid.id = 2))
-
-## calculate two-phase estimation:
-reg2p_nex<- twophase(formula=formula.s0,
-                     data=grisons,
-                     phase_id=list(phase.col = "phase_id_3p", terrgrid.id = 2),
-                     boundary_weights = "boundary_weights")
-
-## calculate three-phase estimation:
-reg3p_nex<- threephase(formula.s0, formula.s1, data = grisons,
-                       phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                       boundary_weights = "boundary_weights")
-
-## get summaries:
-summary(reg2p_nex)
-summary(reg3p_nex)
-
-# merge to esttable-object:
-etab.2p.glob<- estTable(est.list = list(op, reg2p_nex))
-etab.3p.glob<- estTable(est.list = list(op, reg2p_nex, reg3p_nex))
-
-# visualize estimation errors:
-plot(etab.3p.glob)
-
-# quantify gain:
-etab.2p.glob<- estTable(est.list = list(op, reg2p_nex))
-mphase.gain(etab.2p.glob)
-mphase.gain(etab.3p.glob)
-
-
-
-#########################################
-# SMALL AREA THREE-PHASE ESTIMATIONS    #
[…]

### 2017-09-08 r/create figures.R

@@ -54,7 +54,7 @@ grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
 plot(grisons.sae.table.CI, ncol = 2, yvar = "estimate")
 
 ## save graphic for article:
-ggsave("fig\\plot_est_ci.png", 
+ggsave("plot_est_ci.png", 
        plot(grisons.sae.table.CI, ncol = 2, yvar = "estimate") +
          ylab("Timber Volume [m3/ha]") +
          theme(axis.title.x =element_text(size=rel(6)), # size of axis-labels text
@@ -72,7 +72,7 @@ plot(grisons.sae.table, ncol = 2, yvar = "error")
 
 
 ## save graphic for article:
-ggsave("fig\\plot_error.png", 
+ggsave("plot_error.png", 
        plot(grisons.sae.table, ncol = 2, yvar = "error") +
          theme(axis.title.x =element_text(size=rel(6)), # size of axis-labels text
                axis.title.y =element_text(size=rel(6)),

### 2017-09-08 r/special_cases_and_scenarios02.R

@@ -1,269 +1,0 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 6  "Special Cases and Scenarios"                                         #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-
-
-#####################################################
-# Demonstrate error handling in package (examples)  #
-#####################################################
-
-
-# --------------------------------------------------------------------------- #
-# TWO-PHASE
-# --------------------------------------------------------------------------- #
-
-## save grisons as new dataset:
-grisons.n<- grisons
-
-grisons.n[grisons$phase_id_2p==2, "phase_id_2p"]<- 3
-grisons.n[grisons$phase_id_2p==1, "phase_id_2p"]<- 2
-table(grisons.n$phase_id_2p)
-
-## delete explanatory variables from an s2-(i.e. s1-) sample point:
-grisons.n[which(grisons.n$phase_id_2p==3)[1],c(4,5,6,7)]<- NA
-
-tp<- twophase(formula=tvol ~ mean + stddev + max + q75,
-                 data=grisons.n,
-                 phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 3),
-                 boundary_weights = "boundary_weights")
-
-# Violation:
-# s2-point MUST have all explanatory variables available
-
-# --> error message due to nesting violation: s2 not nested
-
-# error handling:
-# s2 point with missing auxiliary information is deleted from the dataset
-
-# --------------------------------------------------------------------------- #
-
-## save grisons as new dataset:
-grisons.n<- grisons
-
-grisons.n[grisons$phase_id_2p==2, "phase_id_2p"]<- 3
-grisons.n[grisons$phase_id_2p==1, "phase_id_2p"]<- 2
-table(grisons.n$phase_id_2p)
-
-
[…]

### 2017-09-11 JStatSoft_forestinventory_master-concordance.tex
update Andreas 11.09.2017

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
-1 82 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 304:%
+1 131 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 353:%
 132 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 316:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 365:%
 1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 847:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 896:%
 145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 848:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 897:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1367:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1416:%
 147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1379:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1428:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1444:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1493:%
 160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1454:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1503:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1800:%
 171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1761:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1810:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1992:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2041:%
 182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2001:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2050:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2006:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2055:%
 192 11 1}

### 2017-09-11 rnw/Main structure and modules of the Package.Rnw
update Andreas 11.09.2017

@@
 
-% NOTE:
-% The aim of this section is to build the basic understanding of two and threephase sampling, and the extention to small area estimation. We introduce the general terminology and show
-% how the package is structured to apply the methods
+\subsection{Forest Inventory in the Infinite Population Approach}
+
+Lets assume that the spatial distribution of a target variable (e.g.
+the timber volume) over the forest area is determined by a mathematical continuous function.
+The total amount of the target variable $Y$ (e.g.
+the total timber volume of the forest) consequently is the integral of that function, in other words: the volume that expands under the surface of that function (Figure \ref{fig:inf_pop_apr}).
+We can thus reconsider our task of estimating e.g.
+the true timber volume of a forest as "estimating the integral of the unknown function that determines the timber volume distribution over our forest".
+The challenging part of an inventory is that the form of this function, and thus $Y$ is unknown.
+Theoretically, we still could get $Y$ by observing the function value $Y(x)$, called the \textit{local density}, at each possible point $x$ over the function plain (i.e.
+the forest area) and sum those obervations up.
+However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
+So the only thing we can do is to get a sample of $Y(x)$ to \textit{estimate} $Y$ by $\hat{Y}$.
+Since we thereby sample from an infinite population of sample points, and thus function values, this approach is consequently called the \textit{infinite population approach}.
+All estimators included in the \proglang{R} package \pkg{forestinventory} basically rest upon these theoretical thoughts.
+In practice, the value of $Y(x)$ is determined by measuring sample trees that lie within a circle (called sample plot) around the sample point $x$ located in the forest area, and aggregating their individual characteristics (e.g.
+their individual timber volumes).
+However, a key point to understand is that in the infinite population approach, a local density value $Y(x)$ has \textit{no spatial association}.
+Even if we use a spatial unit to calculate $Y(x)$, its value solely corresponds to the sample point $x$, which constitutes the sample unit.
+This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
+One is the impossibility of a perfect tesselation of an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not defined.
+The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
+
+\begin{figure}[htb]
+\centering
+\resizebox{0.5\hsize}{!}{\includegraphics{fig/OnePhase.png}}
+\caption{Visualization of the infinite population approach. The spatial distribution of the target variable $Y$ is determined by a continous function (green plain) from which 
+only a sample of function values $Y(x)$ (dots on the plain) is known from a terrestrial survey.}
+\label{fig:inf_pop_apr}
+\end{figure}
+
+% NOTES:
+% - Use onephase sampling to explain the difference between finite and infinite population approach
+% - infinite population approach: 
+%   - Lets assume that the spatial distribution of the target variable (e.g. the timber volume) over the forest area is determined by a fixed mathematical continuous function,
+%     and the total amount of the target variable (e.g. the total timber volume of the forest), lets call it $Y$, is the integral of that function, 
+%     in other words: the volume that expands under the surface of that function (picture). The thing is that in the inventory context,
+%     the form of this function is unknown. Theoretically, we still could get Y by observing the
+%     function value Y(x), called the local density, at each possible location $x$ over the function "plane" and sum those obervations up. However, we see at the first spot 
+%     that this is impossible, because
+%     the number of function values $Y(x)$ and thus the number of possible sample locations $x$ is $infinite$. So the only thing we can do is to get ourself a sample of Y(x)s to 
+%     $estimate$ $Y$ by $\hat{Y}$. Since we thereby sample from an infinite population of possible sample locations / function values, this approach is consequently called the 
+%     \textit{infinite population approach}.
+%   - Transferring this approach to forest inventory problems is indeed straightforward when we reconsider our task of estimating e.g. the true timber volume of the forest 
+%     as "estimating the integral of an unknown function that determines the timber volume distribution over our forest". What we do in practice is to visit each sample 
+%     location $x$ in the field and determine the respective timber volume value, i.e. local density, by measuring sample trees around each location. 
+%     The selection of sample trees is mostly done constructing one or multiple nested circles (sample plots) around a sample location, and include each tree that lies 
+%     within the plot boundaries. The local density value is then determined by aggregating the sample tree characteristics (e.g. their indiv. timber volume).
+%     Doing so, we get a sample of the infinite number of (sample locations and thus) timber volume values we could theoretically observe in our forest. 
+%  -  A key point to understand is that in the infinite population approach, a local density value Y(x) has no spatial association. Even if we used a spatial unit, i.e. 
+%    the plot, extended around the sample point to calculate Y(x), its value only corresponds to the sample point / plot center. The best example of non-existing 
+%    spatial association of Y(x) is indeed angle count sampling (ref). Whereas the lack of spatial association seems rather distracting in the first place, it actually 
+%    has some obvious practical advantages over the finite population approach where the sample units are indeed cirular or rectangular plots.
+%
 
@@
 
+\begin{figure}[htb]
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
+		\caption{} \label{fig:concmphase_and_sae_a}
+		\end{subfigure}
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
+		\caption{} \label{fig:concmphase_and_sae_b}
+	\end{subfigure}
+\caption{\textbf{(a)} Concept of multiphase sampling.
+The square represents the forest area for which an inventory is being conducted.
+The points denote the sample locations $x$.
+Filled points indicate \textit{available} information.
+\textbf{(b)} Illustration of the Small Area Estimation problem}
+\label{fig:concmphase_and_sae}
+\end{figure}
+
 \subsection{Small Area Estimation}
@@
 
-\begin{figure}[htb]
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
-		\caption{} \label{fig:concmphase_and_sae_a}
-		\end{subfigure}
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
-		\caption{} \label{fig:concmphase_and_sae_b}
-	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
+% \begin{figure}[htb]
+% 	\begin{subfigure}[t]{0.5\textwidth}
+% 		\centering
+%     \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
+% 		\caption{} \label{fig:concmphase_and_sae_a}
+% 		\end{subfigure}
+% 	\begin{subfigure}[t]{0.5\textwidth}
+% 		\centering
+%    \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
+% 		\caption{} \label{fig:concmphase_and_sae_b}
+% 	\end{subfigure}
+% \caption{\textbf{(a)} Concept of multiphase sampling.
 The square represents the forest area for which an inventory is being conducted.
@@
 \textbf{(b)} Illustration of the Small Area Estimation problem}
-\label{fig:concmphase_and_sae}
-\end{figure}
+% \label{fig:concmphase_and_sae}
+% \end{figure}
 
@@
 
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
+% In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
 From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
@@
 
-\newpage
+In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
+From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
+The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications.
+The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}).
+In the second case, the package offers three small area estimators that will in detail be described in the following sections.
+The estimators are available under \textit{exhaustive} and \textit{non-exhaustive} use of the auxiliary data.
+Additionally, the package provides the option of calculating one-phase estimates solely based on terrestrial samples.
+All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
+The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
+While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
+We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
+A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
 

### 2017-09-12 JStatSoft_forestinventory_master-concordance.tex
update Andreas 12.09.2017

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 365:%
-1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 896:%
+1 169 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 939:%
 145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 897:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 940:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1416:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1459:%
 147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1428:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1471:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1493:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1536:%
 160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1503:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1546:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1800:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1843:%
 171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1810:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1853:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2041:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2084:%
 182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2050:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2093:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2055:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2098:%
 192 11 1}

### 2017-09-12 rnw/Twophase_Estimators and their Application.Rnw
update Andreas07.09.2017(2)
update Andreas 12.09.2017

--- 2017-09-04 rnw/Twophase_Estimators and their Application.Rnw
@@
 
-The vector of regression coefficients of the OLS regression model is found by solving the sample-based normal equation
+The vector of regression coefficients of the OLS regression model is found by using the following solution to the sample-based normal equation:
 
@@
 
-This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables.
-Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
-Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
-This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} function as input data.
-\citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g.
-from previous studies.
-However, a respective function in the package is up to now non existing and will be subject to future extentions.\par
+The individual predictions can then be calculated as $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
+In the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
+This means that the vector of regression coefficients are always derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} function.
+While virtually all inventorists fit their models using the current inventory data, \citet{massey2015a} also consider scenarios using external models where the sample used to train the model is assumed to be taken from an independent source.
+However, such functions have yet to be implemented in the package at this time.\par
 
-The design-based variance-covariance matrix of the regression coefficients $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ is then calculated as
-
-\begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-\end{equation}
-
-\textbf{Point estimates} and their associated \textbf{variances} can now be calculated under \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables.
-Note that exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
-Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios.
-In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-The design-based regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area (figure \ref{fig:exh_nexh_bw}).
-In the \textit{exhaustive} case, the explanatory variables are actucally calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}).
-This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$.
-In the \textit{non-exhaustive} case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}).
-Consequently, the true mean $\bar{\pmb{Z}}$ has thus to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}).
-As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
-
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{threephase()} function of the package.
-In forest inventories, the sample is often restricted to those sample locations located within the forest area.
-In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
-by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}).
-This method was suggested in \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
-To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}.
-The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.
-
-\begin{equation}\label{eq:wmeanZ}
-  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_1}}w(x)}
-\end{equation}
-
-The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively.
-Note that this form results under the particular case where the regression coefficients are derived using the data from the current inventory (i.e.
-\textit{internal} model).
-In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textit{external} models \citep{mandallaz2016}.
+The package provides the calculation of \textbf{point estimates} under exhaustive ($ex$) and non-exhaustive ($nex$) use of the auxiliary information, which means to respectively apply $\hat{\pmb{\beta}}_{s_2}$ to $\bar{\pmb{Z}}$, i.e.
+the \textit{exact} spatial mean of $\pmb{Z}(x)$, or to $\hat{\bar{\pmb{Z}}}$, i.e.
+an \textit{estimate} of the spatial mean of $\pmb{Z}(x)$:
 
@@
 
-The package calculates two kind of variances for the point estimate.
-The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}).
-Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citet{mandallaz2016}.
+Note that for \textit{internal} linear models the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction which is why it does not appear in the point estimate.
 
+The package \pkg{forestinventory} implements two kinds of variances for each of these point estimates: the \textit{g-weight} formulation, also referred to as \textit{design-based variance}, (see \ref{eq:gw_var_2p_reg_ex} and \ref{eq:gw_var_2p_reg_nex}) given in \citet{mandallaz2016}, and the \textit{external variance} formulation (see \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}) that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
+
+The \textit{g-weight} formulation is
 \begin{subequations}\label{eq:gw_var_2p_reg}
@@
   \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
+  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
 \end{align}
@@
 
-where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
-\begin{equation}\label{estvarcovaux}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
-\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
+where the design-based variance-covariance matrix of $\beta_{s_2}$ is calculated as
+
+\begin{equation}\label{eq:estvarmatrix}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
 \end{equation}
 
-The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}) that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
-Applied under internal models, the external variance is asymptotically unbiases and usually slightly smaller than the design-based variance, where the uncertainty of the regression coefficients is considered in the variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
+and the uncertainty caused by using the $s_1$ sample to estimate $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is accounted for by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$
+\begin{equation}\label{estvarcovaux}
+\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}=
+\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
+(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})^t
+\end{equation}
+
+The \textit{external variance} formulation for linear regression models is
 
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & =
- \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
-(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:varexternal_2p_reg_nex}
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}(R(x))=\frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(\hat{R}(x)-\bar{\hat{R}}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}(R(x)) \nonumber \\ & =
+ \frac{1}{n_1}\frac{1}{n_1-1}\sum_{x\in{s_1}}(\hat{Y}(x)-\bar{\hat{Y}}_{s_1})^2 + \frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:varexternal_2p_reg_nex}
 \end{align}
 \end{subequations}
+
+Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (\ref{eq:estvarmatrix}).
+The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formation is possible, as is the case with non-parametric models like kNN.
+
+%% ------------ in progress -------------------- %%
 
@@
 	\end{subfigure}
-\caption{Concept of exhaustive \textbf{(a)} and non-exhaustive \textbf{(b)} calculation of explanatory variables including boundary adjustment at the support level.
+\caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level.
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
@@
 
+At this point, an inquisitive reader should notice that the calculation of $\bar{\pmb{Z}}$ and $\hat{\bar{\pmb{Z}}}$ has not been explicitly defined yet.
+
+We
+
+For the non-exhaustive context, assuming there is no boundary adjustment necessary, we can simply use $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ and then the mathematical theory is perfectly aligned with its implementation in the \pkg{forestinventory} package.
+However, in the exhaustive context under the infinite population approach, $n_1=\infty$ technically, which means that mathematically $\bar{\pmb{Z}}=\frac{1}{\lambda{(F)}}\int_{x\in{F}}\pmb{Z}(x)$.
+This integral is practically impossible to calculate \textit{exactly} because $\pmb{Z}(x)$ would have to be calculated for an infinite number of $x$'s in $s_1$.
+What we want to do to get around this is to find a way to exhaustively use all available \textbf{auxiliary information} in such a way that $n_1$ is so large that we can safely assume that $\bar{\pmb{Z}}$ is known.
+Practically speaking, since very large $n_1$ translates into very large input datasets which may conflict with the memory limitations in \proglang{R}, the \pkg{forestinventory} package allows the user to calculate vector $\bar{\pmb{Z}}$ on their own and input it directly into the \code{twophase()} and \code{threephase()}.
+
+% In the \pkg{forestinventory} package, exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
+
+The key to calculating your own $\bar{\pmb{Z}}$ is to understand the difference between the explanatory variables contained in vector $\pmb{Z}(x)$ and spatial extent of the auxiliary information around the point $x$ (the so-called support) from which $\pmb{Z}(x)$ is derived.
+Figure \ref{fig:exh_nexh_bw} illustrates the practical difference between the exhaustive and non-exhaustive set-ups.
+In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports should be tesselated to form a partition over the inventory area as is the case with the dark green polygons (squares tiles in this case) in Figure \ref{fig:exh_nexh_and_boundweights_a}.
+As long as the size of the supports are not unreasonably large, $n_1$ is expected to be reasonably large to assume that the difference between $\hat{\bar{\pmb{Z}}}$ and $\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ is neglible.
+This is in contrast to the non-exhaustive set-up in Figure \ref{fig:exh_nexh_and_boundweights_b} where the supports leave much of the available auxiliary information unused (represented by the light green tiles).
+In both cases, the vector of explanatory variables, $\pmb{Z}(x)$, can be calculated for each point $x$ using the support around that point.
+It should be noted that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
+
+%% NOTES:
+% - first, mention what Z(x) means in practice, and how we calculate a variable stored in Z(x), i.e. use supports (maybe mention the choice of support size,
+%   and that it can have an influence on model accuracy). 
+% - Figure 4(b) shows how the derivation of the Z(x) in the non-exhaustive case often looks like in practice, i.e. a regular grid is given by a 
+%   terrestrial inventory, densified to an s1-sample, and then supports are used to calculate the set of explanatory variables at sample location x.
+%   This is in perfect aggreement with just the presented theory of the non-exhaustive estimators (except for using regular grids rather than randomly
+%   placed sample points) and the forestinventory package will just use the emprirical mean of the Z(x) values given in the input dataframe, i.e. 
+%   $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
+% - The exhaustive case however leaves some space for the implementation:
+%   This is because likewise the local density $Y(x)$ gathered in the second phase $s_2$, also the calculation of Z(x) only referrs to 
+%   the sample point x, and has no spatial association (see section 2.1). Likewise the mean of the local density, the only way to determine the exact mean 
+%   of $\bar{Z}$ would be to calculate the integral, i.e. in practice calculate Z(x) at an infinite number ($n_1=\infty$) of sample points $x$. In practice,
+%   this is obviously impossible.
+% - Figure 4(a) shows the implemented solution of an "exhaustive use" of the auxiliary information as described in Mandallaz, Breschan, Hill (2013). Is has to be noted
+%   that this would perfectly enable for calculating the exact mean $\bar{Z}$ in the \textit{finite} population approach (i.e. we calculate Z(x) for the entire 
+%   population of supports, and can thus calculate the exact mean $\bar{Z}$). However, this is not equal to the exact mean $\bar{Z}$ in the infinite population
+%   approach for the mentioned reasons.
+% - What we can do in practice is to get a reasonable approximation of $\bar{Z}$. 
+%
+%
+%   as an approximation, i.e. if we cover the
+%
+%
+% - mention that in the strict inf.pop.appr. frame, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample
+%   n1 of Z(x) to get a reasonable approximation of $\bar{Z}$. This is possible because likewise the local density $Y(x)$ gathered in the second phase $s_2$,
+%   also the calculation of Z(x) only referrs to the sample point, and has no spatial association.
+%
+
+%% ------------ in progress -------------------- %%
+
+An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{threephase()} function of the package.
+In forest inventories, the sample is often restricted to those sample locations located within the forest area.
+In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
+by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (Figure \ref{fig:exh_nexh_bw}).
+This method was suggested in \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
+To insure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}(x)$ or $\bar{\pmb{Z}}(x)$, the respective means have then to be calculated as the \textit{weighted} mean according to \ref{eq:wmeanZ}.
+The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.
+
+\begin{equation}\label{eq:wmeanZ}
+  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_1}}w(x)}
+\end{equation}
+
+% --------------------------- %
 \subsubsection{Application}

### 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
update Andreas 18.09.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 365:%
-1 169 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 939:%
+1 189 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 959:%
 145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 940:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 960:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1459:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1479:%
 147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1471:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1491:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1536:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1556:%
 160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1546:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1566:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1843:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1863:%
 171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1853:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1873:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2084:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2104:%
 182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2093:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2113:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2098:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2118:%
 192 11 1}

### 2017-09-18 rnw/Main structure and modules of the Package.Rnw
update Andreas 12.09.2017
update Andreas 18.09.2017(1)

@@
 the timber volume) over the forest area is determined by a mathematical continuous function.
-The total amount of the target variable $Y$ (e.g.
+The \textit{total} of the target variable $Y$ (e.g.
 the total timber volume of the forest) consequently is the integral of that function, in other words: the volume that expands under the surface of that function (Figure \ref{fig:inf_pop_apr}).
@@
 the true timber volume of a forest as "estimating the integral of the unknown function that determines the timber volume distribution over our forest".
-The challenging part of an inventory is that the form of this function, and thus $Y$ is unknown.
+From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
 Theoretically, we still could get $Y$ by observing the function value $Y(x)$, called the \textit{local density}, at each possible point $x$ over the function plain (i.e.
@@
 \centering
-\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
+\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
 \caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}

### 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
update Andreas 18.09.2017(2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 105 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
+1 107 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 115:%
 1 97 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 211:%
-121 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
+123 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
 1 131 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 353:%
-132 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 365:%
-1 189 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 959:%
-145}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 960:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 355:%
+134 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
+1 203 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 975:%
+147}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 976:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1479:%
-147 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1491:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1495:%
+149 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1507:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1556:%
-160 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1566:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1572:%
+162 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1582:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1863:%
-171 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1873:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1879:%
+173 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1889:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2104:%
-182 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2113:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2120:%
+184 8 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2129:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2118:%
-192 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2134:%
+194 11 1}

### 2017-09-18 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 15.09.2017
update Andreas 18.09.2017(1)
update Andreas 18.09.2017(2)

@@
 
+% \begin{figure}[htb]
+% 	\begin{subfigure}[t]{0.5\textwidth}
+% 		\centering
+%     \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
+% 		\caption{} \label{fig:exh_nexh_and_boundweights_a}
+% 		\end{subfigure}
+% 	\begin{subfigure}[t]{0.5\textwidth}
+% 		\centering
+%    \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
+% 		\caption{} \label{fig:exh_nexh_and_boundweights_b}
+% 	\end{subfigure}
+% \caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level.
+Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
+A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
+w: weights used for calculating the weighted mean of each explanatory variable.
+Dotted circles indicate terrestrial sample plots.}
+% \label{fig:exh_nexh_bw}
+% \end{figure}
+
+% At this point, an inquisitive reader should notice that the calculation of $\bar{\pmb{Z}}$ and $\hat{\bar{\pmb{Z}}}$ has not been explicitly defined yet.  
+% 
+% We 
+% 
+% For the non-exhaustive context, assuming there is no boundary adjustment necessary, we can simply use $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ and then the mathematical theory is perfectly aligned with its implementation in the \pkg{forestinventory} package.
+However, in the exhaustive context under the infinite population approach, $n_1=\infty$ technically, which means that mathematically $\bar{\pmb{Z}}=\frac{1}{\lambda{(F)}}\int_{x\in{F}}\pmb{Z}(x)$.
+This integral is practically impossible to calculate \textit{exactly} because $\pmb{Z}(x)$ would have to be calculated for an infinite number of $x$'s in $s_1$.
+What we want to do to get around this is to find a way to exhaustively use all available \textbf{auxiliary information} in such a way that $n_1$ is so large that we can safely assume that $\bar{\pmb{Z}}$ is known.
+Practically speaking, since very large $n_1$ translates into very large input datasets which may conflict with the memory limitations in \proglang{R}, the \pkg{forestinventory} package allows the user to calculate vector $\bar{\pmb{Z}}$ on their own and input it directly into the \code{twophase()} and \code{threephase()}.
+% 
+% % In the \pkg{forestinventory} package, exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
+% 
+% The key to calculating your own $\bar{\pmb{Z}}$ is to understand the difference between the explanatory variables contained in vector $\pmb{Z}(x)$ and spatial extent of the auxiliary information around the point $x$ (the so-called support) from which $\pmb{Z}(x)$ is derived.
+Figure \ref{fig:exh_nexh_bw} illustrates the practical difference between the exhaustive and non-exhaustive set-ups.
+In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports should be tesselated to form a partition over the inventory area as is the case with the dark green polygons (squares tiles in this case) in Figure \ref{fig:exh_nexh_and_boundweights_a}.
+As long as the size of the supports are not unreasonably large, $n_1$ is expected to be reasonably large to assume that the difference between $\hat{\bar{\pmb{Z}}}$ and $\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ is neglible.
+This is in contrast to the non-exhaustive set-up in Figure \ref{fig:exh_nexh_and_boundweights_b} where the supports leave much of the available auxiliary information unused (represented by the light green tiles).
+In both cases, the vector of explanatory variables, $\pmb{Z}(x)$, can be calculated for each point $x$ using the support around that point.
+It should be noted that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
+
+\subsubsection{Calculation of Explanatory Variables}
+
+We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
+In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_1,...,x_p$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
+Figure \ref{fig:exh_nexh_and_boundweights_b} shows how the derivation of $\pmb{Z}(x)$ in the \textbf{non-exhaustive} case often looks like in practice, i.e.
+a regular terrestrial grid $s_2$ is given by a terrestrial inventory, densified to a larger sample $s_1$, and then supports are used to calculate the set of explanatory variables at the sample locations.
+This is in perfect aggreement with the just presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
+In this case, the \pkg{forestinventory} package calculates the emprirical mean of the $\pmb{Z}(x)$ values given in the input dataframe, i.e.
+$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
+The \textbf{exhaustive} case however leaves some space of discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only referrs to the sample point $x$ and has no spatial association.
+Likewise the mean of the local density (i.e.
+the total), the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number ($n_1=\infty$) of sample points $x$.
+Since this is obviously impossible, we can only try get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+One possibility is shown in Figure \ref{fig:exh_nexh_and_boundweights_a} which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
+Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
+deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
+While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
+Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
+One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
+Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a prefect tesselation might not always be feasible.
+In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
+This is again because $\pmb{Z}(x)$ only referrs to the sample point and has no spatial association.
+With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
+One could even consider the use of different support sizes for an explanatory variable as long as they are consistently used at all sample locations.
+
+Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
+The first case is when the explanatory variables are provided by polygon layers (e.g.
+map of development stages).
+In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
+The second case is when the exact mean can be calculated in one step, e.g.
+taking the mean of all height pixels of a raster canopy height model will equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+
 \begin{figure}[htb]
@@
 
-At this point, an inquisitive reader should notice that the calculation of $\bar{\pmb{Z}}$ and $\hat{\bar{\pmb{Z}}}$ has not been explicitly defined yet.
-
-We
-
-For the non-exhaustive context, assuming there is no boundary adjustment necessary, we can simply use $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ and then the mathematical theory is perfectly aligned with its implementation in the \pkg{forestinventory} package.
-However, in the exhaustive context under the infinite population approach, $n_1=\infty$ technically, which means that mathematically $\bar{\pmb{Z}}=\frac{1}{\lambda{(F)}}\int_{x\in{F}}\pmb{Z}(x)$.
-This integral is practically impossible to calculate \textit{exactly} because $\pmb{Z}(x)$ would have to be calculated for an infinite number of $x$'s in $s_1$.
-What we want to do to get around this is to find a way to exhaustively use all available \textbf{auxiliary information} in such a way that $n_1$ is so large that we can safely assume that $\bar{\pmb{Z}}$ is known.
-Practically speaking, since very large $n_1$ translates into very large input datasets which may conflict with the memory limitations in \proglang{R}, the \pkg{forestinventory} package allows the user to calculate vector $\bar{\pmb{Z}}$ on their own and input it directly into the \code{twophase()} and \code{threephase()}.
-
-% In the \pkg{forestinventory} package, exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
-
-The key to calculating your own $\bar{\pmb{Z}}$ is to understand the difference between the explanatory variables contained in vector $\pmb{Z}(x)$ and spatial extent of the auxiliary information around the point $x$ (the so-called support) from which $\pmb{Z}(x)$ is derived.
-Figure \ref{fig:exh_nexh_bw} illustrates the practical difference between the exhaustive and non-exhaustive set-ups.
-In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports should be tesselated to form a partition over the inventory area as is the case with the dark green polygons (squares tiles in this case) in Figure \ref{fig:exh_nexh_and_boundweights_a}.
-As long as the size of the supports are not unreasonably large, $n_1$ is expected to be reasonably large to assume that the difference between $\hat{\bar{\pmb{Z}}}$ and $\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ is neglible.
-This is in contrast to the non-exhaustive set-up in Figure \ref{fig:exh_nexh_and_boundweights_b} where the supports leave much of the available auxiliary information unused (represented by the light green tiles).
-In both cases, the vector of explanatory variables, $\pmb{Z}(x)$, can be calculated for each point $x$ using the support around that point.
-It should be noted that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
-
 %% NOTES:
+% We will now quickly draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
+In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_1,...,x_p$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
+Figure 4(b) shows how the derivation of Z(x) in the non-exhaustive case often looks like in practice, i.e.
+a regular terrestrial grid $s_2$ is given by a terrestrial inventory, densified to a larger sample $s_1$, and then supports are used to calculate the set of explanatory variables at the sample location x.
+This is in perfect aggreement with the just presented theory of the non-exhaustive estimators, except for using regular grids rather than randomly placed sample points, i.e.
+the forestinventory package will calculate the emprirical mean of the Z(x) values given in the input dataframe, i.e.
+$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
+% The exhaustive case however leaves some space of discussion concerning the calculation of $Z(x)$:
+%
+%
 % - first, mention what Z(x) means in practice, and how we calculate a variable stored in Z(x), i.e. use supports (maybe mention the choice of support size,
@@
 %   terrestrial inventory, densified to an s1-sample, and then supports are used to calculate the set of explanatory variables at sample location x.
-%   This is in perfect aggreement with just the presented theory of the non-exhaustive estimators (except for using regular grids rather than randomly
+%   This is in perfect aggreement with the just presented theory of the non-exhaustive estimators (except for using regular grids rather than randomly
 %   placed sample points) and the forestinventory package will just use the emprirical mean of the Z(x) values given in the input dataframe, i.e. 
@@
 %
-%   as an approximation, i.e. if we cover the
-%
-%
 % - mention that in the strict inf.pop.appr. frame, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample
@@
 %   also the calculation of Z(x) only referrs to the sample point, and has no spatial association.
+%
+% - mention that the support size thereby does not have to match the plotsize of the terrestrial inventory.
+In the strict inf.pop.appr.
+frame, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample n1 of Z(x) to get a reasonable approximation of $\bar{Z}$.
+This is possible because likewise the local density $Y(x)$ gathered in the second phase $s_2$, also the calculation of Z(x) only referrs to the sample point, and has no spatial association.
+In fact, the support size will in practice be chosen in order to achieve a best possible explanatory power of the regression model, thus minimzing the residual variation.
+mention that a change of support size between sample points for a given expl.
+variable is not allowed (because....
+Daniel ?)
+%
 %

### 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
update Andreas 18.092017(3)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
-1 203 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 975:%
+1 202 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 974:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 976:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 975:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1495:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1494:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1507:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1506:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1572:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1571:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1582:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1581:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1879:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1878:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1889:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1888:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2120:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2119:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2129:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2128:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2134:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2133:%
 194 11 1}

### 2017-09-18 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 18.092017(3)

@@
 This is in perfect aggreement with the just presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
-In this case, the \pkg{forestinventory} package calculates the emprirical mean of the $\pmb{Z}(x)$ values given in the input dataframe, i.e.
-$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
+In this case, the \pkg{forestinventory} package calculates the emprirical mean of the $\pmb{Z}(x)$ values given in the input data frame, i.e.
+$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
 The \textbf{exhaustive} case however leaves some space of discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only referrs to the sample point $x$ and has no spatial association.
@@
 Since this is obviously impossible, we can only try get a reasonable approximation of $\bar{\pmb{Z}}$.\par
-One possibility is shown in Figure \ref{fig:exh_nexh_and_boundweights_a} which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
 In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
@@
 One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
-Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a prefect tesselation might not always be feasible.
+Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
 In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.

### 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
update Andreas 18.09.2017(4)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
-1 202 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 974:%
+1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1003:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 975:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 1004:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1494:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1523:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1506:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1535:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1571:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1600:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1581:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1610:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1878:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1907:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1917:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2119:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2148:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2128:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2157:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2133:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2162:%
 194 11 1}

### 2017-09-18 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 18.09.2017(4)

@@
 One could even consider the use of different support sizes for an explanatory variable as long as they are consistently used at all sample locations.
+% or: as long as they are not changed between sample locations (change of support)
 
@@
 To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository.
-The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample).
+The dataset contains data from a two-phase forest inventory under simple (i.e.
+non-cluster) sampling design conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
+The $s_1$ same is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
 Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
-For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey ($s_2$ sample).
-We can load the data frame in the \proglang{R} environment by executing:
+For a systematic subsample of 67 ($s_2$ sample), terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey.
+We can load \pkg{forestinventory} and examine the \code{grisons} data set in the \proglang{R} environment as follows:
+
+<<echo=FALSE>>=
+options(width=10000) 
+@
 
 \begin{small}
-<<>>=
+<<eval=FALSE>>=
+library(forestinventory)
 data("grisons", package = "forestinventory")
+head(grisons)
 @
@@
 
-The structure of the data frame has to meet certain requirements to be usable by the \code{onephase}, \code{twophase()} and \code{threephase()} function of the package.
-Firstly, a row in the data frame corresponds to a sample location while the columns specify the attributes associated to the respective sample location.
+\begin{small}
+<<eval=TRUE, echo=FALSE>>=
+library(forestinventory)
+set.seed(666);head(grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)], 10)
+@
+\end{small}
+
+Estimates can be made using the \code{onephase()}, \code{twophase()} or \code{threephase()} functions.
+The data frame inputted to these functions must have the structure where each row corresponds to a unique sample location and the columns specify the attributes associated to that respective sample location.
+Attributes that are missing, e.g.
+because they are associated with sample locations that were not selected in the subsample for the subsequent phase, should be designated as \code{NA} and the phase membership is encoded as numeric.
+
 For global two-phase estimation, we have to specify
@@
   \item the regression model (\code{formula}) as specified in the \code{lm}()-function \citep{R}
-  \item the \code{data.frame} containing the inventory information (\code{data})
-  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ phase (\code{terrgrid.id})
+  \item the inputted \code{data.frame} containing the inventory information (\code{data})
+  \item the \code{list}-object \code{phase_id} containing: the \code{phase.col} argument identifying the name of the column specifying membership to $s_1$ or $s_2$, and the \code{terrgrid.id} argument specifying which numeric value indicates $s_2$ membership in that column
   \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional)
@@
 \end{small}
+
 The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
-A \textbf{summary of the estimation results} can be obtained by passing this object to the \code{S3} generic \code{summary()}-function.
-The summary informs about the function call, the applied regression formula and the estimation method that was applied.
-The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}).
-Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model (\code{r.squared}).
+A readable \textbf{summary of the estimation results} can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the design-based and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:
+
 \begin{small}
@@
 \end{small}
-If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}.
-The vector must be input in the same order that the \code{lm()}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
-We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.
 
-The right order can easily be established by extracting the columnames of the respective \code{lm}-output. This step is particularly recommended if categorical variables are used. In our case this is:
+For practical use, one should normally always prefer the design-based variance over the external variance.
+This is because when we use internal models, the regression coefficients actually dependent on our sample.
+In contrast to the external variance, the design-based variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical properties (g-weights).
+The external and design-based variances are asymptotically equivalent but the external variance is really only included here in case the user wants to compare to another estimator where no design-based variance exists.
+
+The \textbf{exhaustive estimator} can be applied by additionally passing a vector containing the \textit{exact} means of the explanatory variables, i.e.
+$\bar{\pmb{Z}}$, to the optional argument \code{exhaustive}.
+This vector must be calculated beforehand in such a way that any desired boundary adjustment has already been applied.
+Note that the vector input to \code{exhaustive} must be in the same order that the \code{lm()}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
+This step is particularly recommended if categorical variables are present because the \code{lm()}-function, which is internally used to set up the design-matrix, automatically creates dummy variables with one of the categories used as a reference.
+The correct order for our \code{grisons} example can easily be extracted using:
+
 \begin{small}
@@
 \end{small}
-The \textbf{exhaustive estimator} can thus be applied after creating the vector of exact means $\bar{\pmb{Z}}$:
+
+The \textbf{exhaustive estimator} can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b} and denoted as \code{true.means.Z}:
+
 \begin{small}
@@
 
-We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
+An alternative way to look at the estimation results without using the \code{summary()} is to query \code{reg2p_ex} directly:
 \begin{small}
@@
 Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation.
-This is because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
+This is essentially because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
 

### 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
update Andreas 19.09.2017(1)

--- 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
-1 202 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 974:%
+1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 979:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 975:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 980:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1494:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1499:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1506:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1511:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1571:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1576:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1581:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1586:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1878:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1883:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1893:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2119:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2124:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2128:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2133:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2133:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2138:%
 194 11 1}

### 2017-09-19 rnw/Introduction.Rnw
update Andreas 19.09.2017(1)

@@
 
-The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle them.
+The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} address special cases and demonstrate how the package-functions handle them.
 

### 2017-09-19 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 19.09.2017(1)

@@
 $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
-The \textbf{exhaustive} case however leaves some space of discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only referrs to the sample point $x$ and has no spatial association.
+The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association.
 Likewise the mean of the local density (i.e.
-the total), the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number ($n_1=\infty$) of sample points $x$.
-Since this is obviously impossible, we can only try get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+the total), the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number of sample points $x$ ($n_1=\infty$).
+Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
 One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
@@
 In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
-This is again because $\pmb{Z}(x)$ only referrs to the sample point and has no spatial association.
+This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
 With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
@@
 library(forestinventory)
-set.seed(666);head(grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)], 10)
+set.seed(666);print(head(grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)], 10),row.names=c(1:10))
 @
@@
 This is because when we use internal models, the regression coefficients actually dependent on our sample.
-In contrast to the external variance, the design-based variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical properties (g-weights).
+In contrast to the external variance, the design-based variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical calibration properties (g-weights).
 The external and design-based variances are asymptotically equivalent but the external variance is really only included here in case the user wants to compare to another estimator where no design-based variance exists.
@@
 
-The \textbf{exhaustive estimator} can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b} and denoted as \code{true.means.Z}:
+The \textbf{exhaustive estimator} can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b}, denoted as \code{true.means.Z}:
 
@@
 
-The package \pkg{forestinventory} provides 3 small area estimators.
-A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible.
-An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which are always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par
+The package \pkg{forestinventory} provides three broad types of small area estimators each of which has an exhaustive and non-exhaustive form.
+We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label \textit{pseudo} to indicate that the mean of the explanatory variables within the small area was based on a finite sample.
+The main idea for all these small area estimators is to calculate the regression coefficient vector $\hat{\pmb{\beta}}_{s_2}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ on the entire $s_2$ sample according to equations \ref{normequ_simple} and \ref{eq:estvarmatrix}, and subsequently use that to make predictions for sample locations restricted to small area $G$.\par
 
@@
 We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
-Their point estimates are given in equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}.
-As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) available at the sample location $s_{1,G}$ in the small area.
-A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the model residuals in $G$ to the first term.
-This is called the \textit{bias} or \textit{residual correction} term.
 
@@
 
-The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external variance} (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textit{small} and the \textit{psmall} estimator.
-Again note that all components are restricted to those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
+In the equations for the point estimates (\ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
+A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$ to the first term.
+This is called the \textit{bias} or \textit{residual correction} term.
+
+The package provides the \textit{design-based} variance for \textit{small} and \textit{psmall} respectively (Equations \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
+Again note that all components are restricted to those available at the sample locations in the small area ($s_{1,G}$ and $s_{2,G}$), with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
 
@@
 
-The the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
-\begin{equation}\label{estvarcovaux_G}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
-\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
-\end{equation}
-
 If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
@@
 
-The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are applied if no terrestrial samples are available within the small area $G$ (i.e.
+The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are commonly applied when no terrestrial sample is available within the small area $G$ (i.e.
 $n_{2,G}=0$).
-In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
-However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}) is no longer possible.
+In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
+However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Equations \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
 Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
 $\int_G R(x)dx=0$.
-If this assumption is not fullfilled, the synthetic point estimate $\hat{Y}_{G,synth,2p}$ and the pseudo synthetic point estimate $\hat{Y}_{G,psynth,2p}$ will have a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
+However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
 Also note that the residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
@@
   \hat{Y}_{G,synth,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
-  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth}
-\end{align}
-\end{subequations}
-
-\begin{subequations}\label{eq:var_2p_synth_psynth}
-\begin{align}
-  \hat{\var}(\hat{Y}_{G,synth,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
-  \hat{\var}(\hat{Y}_{G,psynth,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
+  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth} \\
+% \end{align}
+% \end{subequations}
+% \begin{subequations}\label{eq:var_2p_synth_psynth}
+% \begin{align}
+  \hat{\var}(\hat{Y}_{G,synth,2p}) & = \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
+  \hat{\var}(\hat{Y}_{G,psynth,2p}) & = \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
   + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
@@
 
+where the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
+\begin{equation}\label{estvarcovaux_G}
+\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
+\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
+(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
+\end{equation}
+
 %------------------------------------------------------------- %
@@
 
-In the \textit{small} and \textit{psmall} estimator, the crucial zero mean residual property for the small area $G$ is forced by the bias adjustment terms (equation \ref{eq:pointest_2p_small}, \ref{eq:pepsmall}).
-However, the equations would considerably simplify if we could mathematically insure that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
+The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates as well as unreliably small variances.
+The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e.
+$\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
+The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
 $\int_F R(x)dx=\int_G R(x)dx=0$.
-This idea is realized in the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the small area $G$ and is otherwise set to 0.
-The \textit{extended} auxiliary vector thus reads as $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and leads to a classical ANOVA regression model.
-Note that the intercept of the regression model is thus adjusted for all data points belonging to small area $G$ which insures that the mean of the residuals $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R(x)$ in $G$ are always zero by construction.\par
+This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
+Recalling that linear models fit using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
+The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Equations \ref{eq:pest_2p_synth_psynth}, \ref{eq:var_2p_synth_psynth} and  \ref{estvarcovaux_G}.
+Note that the package functions internally extent the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
 
-Under the extended auxiliary vector, the regression coefficient is calculated as
+Not every equation needs to be re-written here, but to give an example of the notational change, the regression coefficient under extended model approach becomes
 
@@
 
-with its variance-covariance matrix
-
-\begin{equation}\label{eq:ext_estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{Z}(x)^t\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-\end{equation}
-
-The point estimate is defined as
+The point estimates and their g-weight variances can then be re-written as
 
@@
 \begin{align}
-\hat{Y}_{G,extsynth,2p}= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
-\hat{Y}_{G,extpsynth,2p}=\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth}
-\end{align}
-\end{subequations}
-
-and the design-based variances are derived as
-
-\begin{subequations}\label{eq:pest_2p_extsynth_extpsynth}
-\begin{align}
-\hat{\var}(\hat{Y}_{G,extsynth,2p})= \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
-\hat{\var}(\hat{Y}_{G,extpsynth,2p})=
+\hat{Y}_{G,extsynth,2p} & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{Y}_{G,extpsynth,2p} & =\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,extsynth,2p}) & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,extpsynth,2p})& =
 \hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
@@
 
-with the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{\mathbb{Z}}}}_G$ estimated by
+While the formulas look similar to the synthetic estimators, note that a decomposition of $\hat{\pmb{\theta}}_{s_2}$ reveals that the residual correction term is now included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimates are asymptotically design-unbiased.
 
-\begin{equation}\label{ext_estvarcovaux_G}
-\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}}=
-\frac{1}{n_{1,G} (n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
-(\pmb{\mathbb{Z}}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}_{G})(\pmb{\mathbb{Z}}(x)-
-\hat{\bar{\pmb{\mathbb{Z}}}}_{G})^t
-\end{equation}
-
-The package also provides the external variance which is calculated as
+The package also provides the external variance for both the \textit{extended} synthetic and \textit{extended} pseudo synthetic estimator.
+Note that both neither the extended model approach nor external variance estimates are possible in the absence of terrestrial samples und thus model residuals in $G$, which is precisely when one must rely on the (pseudo) synthetic estimates.
+The external variance forms of \textit{extsynth} and \textit{extpsynth} are
 
@@
 
-The estimators are called \textit{synthetic} since no residual term is present such as in the \textit{small} and \textit{psmall} estimator (equation \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}) and thus look similar to the synthetic estimators (equation \ref{eq:pointest_2p_reg_synth}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}, \ref{eq:var_2p_reg_psynth}).
-However, note that by a decomposition of $\hat{\pmb{\theta}}_{s_2}$ it can be shown that the residual correction term is indeed included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimator is asymptotically design-unbiased.
+To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small design-based variance due to its potential bias.
+When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias directly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply uses the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
 

### 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
update Andreas 19.09.2017(2)

--- 2017-09-18 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
-1 202 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 974:%
+1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 981:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 975:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 982:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1494:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1501:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1506:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1513:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1571:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1578:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1581:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1588:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1878:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1885:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1888:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1895:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2119:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2126:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2128:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2135:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2133:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2140:%
 194 11 1}

### 2017-09-19 rnw/Main structure and modules of the Package.Rnw
update Andreas 19.09.2017(1)
Infinite population approach.

@@
 
-Lets assume that the spatial distribution of a target variable (e.g.
-the timber volume) over the forest area is determined by a mathematical continuous function.
-The \textit{total} of the target variable $Y$ (e.g.
-the total timber volume of the forest) consequently is the integral of that function, in other words: the volume that expands under the surface of that function (Figure \ref{fig:inf_pop_apr}).
-We can thus reconsider our task of estimating e.g.
-the true timber volume of a forest as "estimating the integral of the unknown function that determines the timber volume distribution over our forest".
-From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
-Theoretically, we still could get $Y$ by observing the function value $Y(x)$, called the \textit{local density}, at each possible point $x$ over the function plain (i.e.
-the forest area) and sum those obervations up.
-However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
-So the only thing we can do is to get a sample of $Y(x)$ to \textit{estimate} $Y$ by $\hat{Y}$.
-Since we thereby sample from an infinite population of sample points, and thus function values, this approach is consequently called the \textit{infinite population approach}.
-All estimators included in the \proglang{R} package \pkg{forestinventory} basically rest upon these theoretical thoughts.
-In practice, the value of $Y(x)$ is determined by measuring sample trees that lie within a circle (called sample plot) around the sample point $x$ located in the forest area, and aggregating their individual characteristics (e.g.
-their individual timber volumes).
-However, a key point to understand is that in the infinite population approach, a local density value $Y(x)$ has \textit{no spatial association}.
-Even if we use a spatial unit to calculate $Y(x)$, its value solely corresponds to the sample point $x$, which constitutes the sample unit.
-This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
-One is the impossibility of a perfect tesselation of an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not defined.
-The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
+Most forest inventories typically follow a procedure that involves randomly selecting points in the forest area defined by GPS coordinates to which inventorists are sent.
+At these points, the inventorist demarcates a sample plot, possibly with the help of an inclusion circle that indicates which trees above a certain threshold are to be selected for measurement.
+These measurements are then converted into local plot densities (e.g.
+the timber density in $\frac{m^3}{ha}$) before the desired estimator is applied.
+The \pkg{forestinventory} package bridges this common inventory paradigm to the mathematics of the estimation process using what is called the \textit{infinite population approach}.
+In this approach the sampling units are \textit{points} each of which has a local density that exists but might remain unobserved if that point was not selected in our inventory's sample.
+For the sake of argument, imagine that the local density for every point in the forested area was somehow known and we could visualize the spatial distribution of the local density function of a target variable (e.g.
+the timber volume) for the entire forest area (Figure \ref{fig:inf_pop_apr}).
+The \textit{population total} of the target variable (e.g.
+the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in (Figure \ref{fig:inf_pop_apr}).
+Clearly, evaluating the local densities for every one of the infinite point so that we can add them up is impossible.
+Thus, in the infinite population approach we are essentially approximating the integral of this unobserved density surface numerically by sampling directly from the surface with our terrestrial inventory sample.
+This is why the infinite population approach is also called the \textit{Monte-Carlo approach}.
+
+All estimators included in the \pkg{forestinventory} package basically rest upon these theoretical thoughts.
+The motivation for this decision was that no bothersome spatial correlations need to be considered as would be the case with geostatisical methods since we assume that the inventory points are selected uniformly and independently, i.e.
+the infinite analogy for simple random sampling.
+Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite populatoin approach} where the sampling plot defined around the point is considered the sampling unit.
+However, this has the disadvatage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.
+In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests.
+We also want to emphasize that the consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
 
@@
 \resizebox{0.5\hsize}{!}{\includegraphics{fig/OnePhase.png}}
-\caption{Visualization of the infinite population approach. The spatial distribution of the target variable $Y$ is determined by a continous function (green plain) from which 
-only a sample of function values $Y(x)$ (dots on the plain) is known from a terrestrial survey.}
+\caption{Visualization of the infinite population approach.
+The spatial distribution of a hypothetical density function for every point in a forested area is represented as wavy green surface.
+Sample plots (grean circles) inform the inventorist of the value of the density function at that point.}
 \label{fig:inf_pop_apr}

### 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
Infinite pop approach

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
-1 131 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 355:%
+1 132 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 356:%
 134 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
-1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 981:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 368:%
+1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1004:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 982:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 1005:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1501:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1524:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1513:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1536:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1578:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1601:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1588:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1611:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1885:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1908:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1895:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1918:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2126:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2149:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2135:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2158:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2140:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2163:%
 194 11 1}

### 2017-09-19 rnw/Main structure and modules of the Package.Rnw
Infinite pop approach

@@
 The \textit{population total} of the target variable (e.g.
-the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in (Figure \ref{fig:inf_pop_apr}).
-Clearly, evaluating the local densities for every one of the infinite point so that we can add them up is impossible.
+the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in Figure \ref{fig:inf_pop_apr}.
+Practically speaking, attempting to evaluate the local densities for every one of the infinite points so that we can add them up is impossible.
 Thus, in the infinite population approach we are essentially approximating the integral of this unobserved density surface numerically by sampling directly from the surface with our terrestrial inventory sample.
@@
 
-All estimators included in the \pkg{forestinventory} package basically rest upon these theoretical thoughts.
-The motivation for this decision was that no bothersome spatial correlations need to be considered as would be the case with geostatisical methods since we assume that the inventory points are selected uniformly and independently, i.e.
+All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach.
+The motivation for this decision was that no bothersome spatial correlations need to be considered, as would be the case with geostatisical methods, since we assume that the inventory points are selected uniformly and independently, i.e.
 the infinite analogy for simple random sampling.
-Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite populatoin approach} where the sampling plot defined around the point is considered the sampling unit.
-However, this has the disadvatage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.
+Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite population approach} where the sampling plot defined around the point is considered the sampling unit.
+The finite population approach has the disadvantage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.
 In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests.
-We also want to emphasize that the consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
+The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
 
@@
 the local density $Y(x)$ (e.g.
-the timber volume density per hectare).
+the timber volume per hectare).
 The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS) technique.

### 2017-09-20 rnw/Twophase_Estimators and their Application.Rnw

@@
 
-Small area estimations in the package \pkg{forestinventory} can be applied by additionally specifying the optional argument \code{small_area}.
-The input data set has therefore to be extended by an additional column that specifies for each each row (i.e.
-observation of a sample location) the respective small area in which it is located.
+Small area estimates in the \pkg{forestinventory} package can be applied by specifying the optional argument \code{small_area}.
+The input data set has to include an additional column of class factor that describes the small area membership of the sample location represented by that row.
 The argument \code{small_area} requires a \code{list}-object that comprises
@@
 The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}.
-In addition to global estimation, the \code{summary()}-output and the \code{estimation} object comprise the estimates and the sample sizes (\code{n1G} and \code{n2G}) for all calculated small areas (column \code{area}).
-Note that the sample size information is also particularly provided in the \code{samplesizes} object returned by the function call.
+In addition to global estimation, the \code{estimation} object now comprises the estimates and variances for all small areas (column \code{area}).
+We can view the sample sizes by looking into the object itself
 \begin{small}
@@
 
-Note that the point estimates and the design-based and external variances of both estimators are very similiar.
-However, the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are changing with the small area, while they are identical under the \textit{psmall} estimator.
+The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need to implement it yourself.
+Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are vary between small areas, while they are identical under the \textit{psmall} estimator.
 This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
@@
 Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas.
-Note that in the \code{grisons} data set, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible.
-Applying the pseudo synthetic estimator equals treating those small areas as if $n_{2,G}=0$ by ignoring the terrestrial information.
+Note that in the \code{grisons} data set, all small areas possess much more than the suggested minimum number of terrestrial observations (a rule of thumb is that $n_{2,G} \geq 6$) required to produce reliable design-unbiased estimates.
+Hence, choosing to use \textit{psynth} is probably not desireable and is just applied here for demonstration purposes.
+The function will in this case ignore the terrestrial information in the data set.
+
 \begin{small}
@@
 
-In this case, the variances of the pseudo synthetic estimator is almost only half the variances of the \textit{psmall} and \textit{extended psynth} estimator.
-The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and \ref{eq:var_2p_extpsynth}).
-Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.\par
+We see here that the \textit{psynth} variances are almost only half the variances of the \textit{psmall} and \textit{extended psynth} estimators.
+However, \textit{psmall} and \textit{extended psynth} are design unbiased and their variances reflect the fact that they account for the uncertainty of the regression model predictions.
+The design-based variance of \textit{psynth} completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
 
-The \textbf{exhaustive versions} of the small area estimators (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) can again be applied by specifying the optional argument \code{exhaustive}.
+The \textbf{exhaustive versions} of the small area estimators (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest.
@@
 
-Comparing this result with the non-exhaustive counterpart \textit{extpsynth}, we see that the estimation uncertainty defined by the variance has again be significantly increased by substituting the estimated by the exact auxiliary means.
-In the function output, exhaustive phases are indicated by a sample size of infinity (\code{Inf}).
-Also note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.
+Just as in the global case, we see that the variance has again been significantly decreased by substituting the \textit{exact} auxiliary means and both first phase sample sizes are now infinity.
+Note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.
 

### 2017-09-21 rnw/Main structure and modules of the Package.Rnw
Still making errors...
update Andreas 21.09.2017(1)

--- 2017-09-18 rnw/Main structure and modules of the Package.Rnw
@@
 
+Most forest inventories gather the field information by randomly or systematically selecting points in the forest area defined by coordinates to which inventorists are sent.
+At these points, the inventorist measures sample trees that commonly lie within a circle (called sample plot) around the sample point $x$, and aggregates their individual characteristics (e.g.
+their individual timber volumes) to local plot densities (e.g.
+the timber density in $m^3/ha$).
+The estimators implemented in the \pkg{forestinventory} package use the so called \textit{infinite population approach} in order to bridge this sampling process to the mathematics behind the estimators.
+The infinite population approach assumes that the spatial distribution of the local density $Y(x)$ over the forest area is determined by a mathematical continuous function, as visualized in Figure \ref{fig:inf_pop_apr}.
+The \textit{population total} of the target variable (e.g.
+the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in Figure \ref{fig:inf_pop_apr}.
+From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
+Theoretically, we still could get $Y$ by observing the function value,i.e.
+the local density $Y(x)$ at each possible point $x$ over the function plain (i.e.
+the forest area) and sum those obervations up.
+However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
+So we are left with only a sample of $Y(x)$ from an infinite population of sample points in order to \textit{estimate} $Y$ by $\hat{Y}$.
+
+All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach.
+
+a key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ has \textit{no spatial association}.
+Even if we use a spatial unit to calculate $Y(x)$, its value solely corresponds to the sample point $x$, which constitutes the sample unit.
+
+This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
+One is the impossibility of a perfect tesselation of an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not defined.
+The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
+
+Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite population approach} where the sampling plot defined around the point is considered the sampling unit.
+The finite population approach has the disadvantage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.
+In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests.
+The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
+
+%% Andi
 Lets assume that the spatial distribution of a target variable (e.g.
@@
 the true timber volume of a forest as "estimating the integral of the unknown function that determines the timber volume distribution over our forest".
+
 From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
@@
 Since we thereby sample from an infinite population of sample points, and thus function values, this approach is consequently called the \textit{infinite population approach}.
+
 All estimators included in the \proglang{R} package \pkg{forestinventory} basically rest upon these theoretical thoughts.
@@
 
-\subsection{Design-based vs. model-dependent approach}
+\subsection{Design-Based vs. Model-Dependent Approach}
 \label{sec:db_vs_md}

### 2017-09-21 JStatSoft_forestinventory_master-concordance.tex
update Andreas 21.09.2017

--- 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
-1 131 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 355:%
+1 130 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 354:%
 134 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 367:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 366:%
 1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 981:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 980:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 982:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 981:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1501:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1500:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1513:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1512:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1578:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1577:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1588:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1587:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1885:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1884:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1895:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1894:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2126:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2125:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2135:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2134:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2140:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2139:%
 194 11 1}

### 2017-09-21 rnw/Main structure and modules of the Package.Rnw
update Andreas 21.09.2017

@@
 Most forest inventories gather the field information by randomly or systematically selecting points in the forest area defined by coordinates to which inventorists are sent.
-At these points, the inventorist measures sample trees that commonly lie within a circle (called sample plot) around the sample point $x$, and aggregates their individual characteristics (e.g.
+At these points, the inventorist measures sample trees located within a constructed circle (called sample plot) around the sample point $x$, and aggregates their individual characteristics (e.g.
 their individual timber volumes) to local plot densities (e.g.
-the timber density in $m^3/ha$).
+the timber volume density in $m^3/ha$).
 The estimators implemented in the \pkg{forestinventory} package use the so called \textit{infinite population approach} in order to bridge this sampling process to the mathematics behind the estimators.
-The infinite population approach assumes that the spatial distribution of the local density $Y(x)$ over the forest area is determined by a mathematical continuous function, as visualized in Figure \ref{fig:inf_pop_apr}.
+The infinite population approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a mathematical continuous function, as visualized in Figure \ref{fig:inf_pop_apr}.
 The \textit{population total} of the target variable (e.g.
@@
 From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
-Theoretically, we still could get $Y$ by observing the function value,i.e.
+Theoretically, we still could get $Y$ by observing the function value, i.e.
 the local density $Y(x)$ at each possible point $x$ over the function plain (i.e.
@@
 However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
-So we are left with only a sample of $Y(x)$ from an infinite population of sample points in order to \textit{estimate} $Y$ by $\hat{Y}$.
-
+So we are left with only a sample of $Y(x)$ from an infinite population of sample points in order to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ by $\hat{Y}$.
 All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach.
-
-a key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ has \textit{no spatial association}.
-Even if we use a spatial unit to calculate $Y(x)$, its value solely corresponds to the sample point $x$, which constitutes the sample unit.
-
-This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
-One is the impossibility of a perfect tesselation of an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not defined.
-The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
-
-Furthermore, points can perfectly fill up any forest area with any amorphous shape, which is not possible if one were to use the \textit{finite population approach} where the sampling plot defined around the point is considered the sampling unit.
-The finite population approach has the disadvantage that circles don't tesselate perfectly and polygons will inevitably distort and pixelate the forest boundary.
-In practice, the population of sampling units in the finite approach will almost never perfectly cover the forests.
-The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
-
-%% Andi
-Lets assume that the spatial distribution of a target variable (e.g.
-the timber volume) over the forest area is determined by a mathematical continuous function.
-The \textit{total} of the target variable $Y$ (e.g.
-the total timber volume of the forest) consequently is the integral of that function, in other words: the volume that expands under the surface of that function (Figure \ref{fig:inf_pop_apr}).
-We can thus reconsider our task of estimating e.g.
-the true timber volume of a forest as "estimating the integral of the unknown function that determines the timber volume distribution over our forest".
-
-From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
-Theoretically, we still could get $Y$ by observing the function value $Y(x)$, called the \textit{local density}, at each possible point $x$ over the function plain (i.e.
-the forest area) and sum those obervations up.
-However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
-So the only thing we can do is to get a sample of $Y(x)$ to \textit{estimate} $Y$ by $\hat{Y}$.
-Since we thereby sample from an infinite population of sample points, and thus function values, this approach is consequently called the \textit{infinite population approach}.
-
-All estimators included in the \proglang{R} package \pkg{forestinventory} basically rest upon these theoretical thoughts.
-In practice, the value of $Y(x)$ is determined by measuring sample trees that lie within a circle (called sample plot) around the sample point $x$ located in the forest area, and aggregating their individual characteristics (e.g.
-their individual timber volumes).
-However, a key point to understand is that in the infinite population approach, a local density value $Y(x)$ has \textit{no spatial association}.
-Even if we use a spatial unit to calculate $Y(x)$, its value solely corresponds to the sample point $x$, which constitutes the sample unit.
+A key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ is a point information associated to the sample point $x$, which constitutes the sample unit, but not to the sample plot area.
 This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
@@
 \resizebox{0.5\hsize}{!}{\includegraphics{fig/OnePhase.png}}
-\caption{Visualization of the infinite population approach.
+\caption{Viisualization of the infinite population approach.
 The spatial distribution of a hypothetical density function for every point in a forested area is represented as wavy green surface.

### 2017-09-21 JStatSoft_forestinventory_master-concordance.tex
update AndreasAlex 21.09.2017

--- 2017-09-19 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 368:%
-1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1004:%
+1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 1005:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 983:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1524:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1502:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1536:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1514:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1601:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1579:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1611:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1589:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1908:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1886:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1918:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1896:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2149:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2127:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2158:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2136:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2163:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2141:%
 194 11 1}

### 2017-09-21 rnw/Main structure and modules of the Package.Rnw
update AndreasAlex 21.09.2017

@@
 their individual timber volumes) to local plot densities (e.g.
-the timber volume density in $m^3/ha$).
-The estimators implemented in the \pkg{forestinventory} package use the so called \textit{infinite population approach} in order to bridge this sampling process to the mathematics behind the estimators.
-The infinite population approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a mathematical continuous function, as visualized in Figure \ref{fig:inf_pop_apr}.
+the timber density in $m^3/ha$).
+The estimators implemented in the \pkg{forestinventory} package use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators.
+The infinite population approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a piecewise constant function, as visualized in Figure \ref{fig:inf_pop_apr}.
 The \textit{population total} of the target variable (e.g.
-the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which defines the volume under the green surface in Figure \ref{fig:inf_pop_apr}.
-From this perspective, the challenging part of an inventory is that the form of this function, and thus $Y$, is unknown.
-Theoretically, we still could get $Y$ by observing the function value, i.e.
-the local density $Y(x)$ at each possible point $x$ over the function plain (i.e.
-the forest area) and sum those obervations up.
-However, we see at the first spot that this is impossible, because the number of function values $Y(x)$ and thus the number of possible sample points $x$ is $infinite$.
-So we are left with only a sample of $Y(x)$ from an infinite population of sample points in order to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ by $\hat{Y}$.
+the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Figure \ref{fig:inf_pop_apr} as the volume under the density surface.
+From this perspective, the challenging part of an inventory is that the form of this function is unknown.
+Theoretically, we still could get the total timber volume by observing the function value, i.e.
+the local density $Y(x)$, at each possible point $x$ over the forest area and sum those obervations up.
+However, this is impossible because there is an infinite number of points in the forest area.
+Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e.
+$Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{s_2}Y(x)$.
+The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
 All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach.
-A key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ is a point information associated to the sample point $x$, which constitutes the sample unit, but not to the sample plot area.
-This has some obvious practical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
-One is the impossibility of a perfect tesselation of an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not defined.
+A key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not to the sample plot area.
+This has some obvious theoretical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
+One is the impossibility of a perfect tesselation over an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not perfectly defined with respect to the forest area.
 The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
@@
 \centering
-\resizebox{0.5\hsize}{!}{\includegraphics{fig/OnePhase.png}}
-\caption{Viisualization of the infinite population approach.
-The spatial distribution of a hypothetical density function for every point in a forested area is represented as wavy green surface.
-Sample plots (grean circles) inform the inventorist of the value of the density function at that point.}
+\resizebox{0.42\hsize}{!}{\includegraphics{fig/localdensity_final-01.eps}}
+\caption{Artificial representation of local density surface.
+The spatial distribution of a hypotheticaldensity function for every point in a forested area is represented as wavy stepwise constant green surface.
+Sample plots (white dots) inform the inventorist of the value of the density function at that point.}
 \label{fig:inf_pop_apr}
 \end{figure}
+
+% "we are sampling values of a function that is defined by the spatial arrangement of the trees"
 
@@
 \centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
+\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
 \caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}

### 2017-09-21 JStatSoft_forestinventory_master-concordance.tex
update Andreas 21.09.2017(3)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 368:%
-1 201 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
+1 172 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 953:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 983:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 954:%
 1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1502:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1473:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1514:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1485:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1579:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1550:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1589:%
-1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1886:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1560:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1846:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1896:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1856:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2127:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2087:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2136:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2096:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2141:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2101:%
 194 11 1}

### 2017-09-21 rnw/Main structure and modules of the Package.Rnw
update Andreas 21.09.2017(3)

@@
 \resizebox{0.42\hsize}{!}{\includegraphics{fig/localdensity_final-01.eps}}
-\caption{Artificial representation of local density surface.
-The spatial distribution of a hypotheticaldensity function for every point in a forested area is represented as wavy stepwise constant green surface.
+\caption{Artificial representation of a local density surface.
+The spatial distribution of a hypothetical density function for every point in a forested area is represented as wavy piecewise constant green surface.
 Sample plots (white dots) inform the inventorist of the value of the density function at that point.}
@@
 the sample locations could arbitrarily be chosen.
-However, another consequence is that for unbiasedness estimates it has to be insured that the model adequately describes the underlying stochastic process, in other words: the model assumptions have to be met and empirically verified.
+However, another consequence is that for unbiased estimates it has to be insured that the model adequately describes the underlying stochastic process, in other words: the model assumptions have to be met and empirically verified.
 For example, if we use an ordinary least square regression model, implicit mathematical assumptions are (i) normality and homogeneity of the  target variable \textit{Y}(\textit{x}) at each value of the explanatory variables, and (ii) independence of the target variable \textit{Y}(\textit{x}) or the model residuals $R(x)$, i.e.

### 2017-09-21 rnw/Special Cases and Scenarios.Rnw
update Andreas 21.09.2017(3)

@@
 This can be taken as evidence that the violation of the zero mean residual property can here expected to be negligible.
-We can also check the calculated mean of the residuals for the small area by
-
-\begin{small}
-<<>>=
-extpsynth.clust$mean_Rc_x_hat_G
-@
-\end{small}
-
-The deviation to zero is in this case rather small, especially compared to the magnitude of the point estimate.
-Note however that the zero mean residual property is a mathematical property of an estimator and can not be proven by analyzing the empirical model residuals.
 
@@
 +             boundary_weights = "boundary_weights")
-1 rows deleted due to missingness in the set of auxiliary parameters for the first phase (s0) 
- (1 terrestrial plots affected by deletion)
 Warning messages:
@@
   Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the second phase (s1) is missing
+  of the first phase (s1) is missing
 2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
   Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the first phase (s0) is missing
+  of the zero phase (s0) is missing
+3: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  1 rows deleted due to missingness in the set of auxiliary parameters for the zero phase (s0)
+  (1 terrestrial plots affected by deletion)
 \end{Sinput}
@@
 +             boundary_weights = "boundary_weights")
-Changed the phase_id for 1 rows to the first phase (s0) due to missingness in the set of 
-  auxiliary parameters for the second phase (s1) (1 terrestrial information no longer usable
+Warning messages:
+1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
+  of the first phase (s1) is missing
+2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in the set of
+  auxiliary parameters for the first phase (s1) (1 terrestrial information no longer usable 
   by this change)
-Warning message:
-In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the second phase (s1) is missing
 \end{Sinput}
@@
 +             boundary_weights = "boundary_weights")
-1 rows deleted due to missingness in the set of auxiliary parameters for the first phase (s0) 
+Warning message:
+In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
+  1 rows deleted due to missingness in the set of auxiliary parameters for the zero phase (s0) 
   (0 terrestrial plots affected by deletion)

### 2017-09-21 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 21.09.2017(3)

@@
 
-%% ------------ in progress -------------------- %%
-
-% \begin{figure}[htb]
-% 	\begin{subfigure}[t]{0.5\textwidth}
-% 		\centering
-%     \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
-% 		\caption{} \label{fig:exh_nexh_and_boundweights_a}
-% 		\end{subfigure}
-% 	\begin{subfigure}[t]{0.5\textwidth}
-% 		\centering
-%    \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
-% 		\caption{} \label{fig:exh_nexh_and_boundweights_b}
-% 	\end{subfigure}
-% \caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level.
-Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
-w: weights used for calculating the weighted mean of each explanatory variable.
-Dotted circles indicate terrestrial sample plots.}
-% \label{fig:exh_nexh_bw}
-% \end{figure}
-
-% At this point, an inquisitive reader should notice that the calculation of $\bar{\pmb{Z}}$ and $\hat{\bar{\pmb{Z}}}$ has not been explicitly defined yet.  
-% 
-% We 
-% 
-% For the non-exhaustive context, assuming there is no boundary adjustment necessary, we can simply use $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ and then the mathematical theory is perfectly aligned with its implementation in the \pkg{forestinventory} package.
-However, in the exhaustive context under the infinite population approach, $n_1=\infty$ technically, which means that mathematically $\bar{\pmb{Z}}=\frac{1}{\lambda{(F)}}\int_{x\in{F}}\pmb{Z}(x)$.
-This integral is practically impossible to calculate \textit{exactly} because $\pmb{Z}(x)$ would have to be calculated for an infinite number of $x$'s in $s_1$.
-What we want to do to get around this is to find a way to exhaustively use all available \textbf{auxiliary information} in such a way that $n_1$ is so large that we can safely assume that $\bar{\pmb{Z}}$ is known.
-Practically speaking, since very large $n_1$ translates into very large input datasets which may conflict with the memory limitations in \proglang{R}, the \pkg{forestinventory} package allows the user to calculate vector $\bar{\pmb{Z}}$ on their own and input it directly into the \code{twophase()} and \code{threephase()}.
-% 
-% % In the \pkg{forestinventory} package, exhaustively available auxiliary information over the entire inventory area does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information.
-% 
-% The key to calculating your own $\bar{\pmb{Z}}$ is to understand the difference between the explanatory variables contained in vector $\pmb{Z}(x)$ and spatial extent of the auxiliary information around the point $x$ (the so-called support) from which $\pmb{Z}(x)$ is derived.
-Figure \ref{fig:exh_nexh_bw} illustrates the practical difference between the exhaustive and non-exhaustive set-ups.
-In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports should be tesselated to form a partition over the inventory area as is the case with the dark green polygons (squares tiles in this case) in Figure \ref{fig:exh_nexh_and_boundweights_a}.
-As long as the size of the supports are not unreasonably large, $n_1$ is expected to be reasonably large to assume that the difference between $\hat{\bar{\pmb{Z}}}$ and $\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$ is neglible.
-This is in contrast to the non-exhaustive set-up in Figure \ref{fig:exh_nexh_and_boundweights_b} where the supports leave much of the available auxiliary information unused (represented by the light green tiles).
-In both cases, the vector of explanatory variables, $\pmb{Z}(x)$, can be calculated for each point $x$ using the support around that point.
-It should be noted that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par
-
 \subsubsection{Calculation of Explanatory Variables}
@@
 With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could even consider the use of different support sizes for an explanatory variable as long as they are consistently used at all sample locations.
-% or: as long as they are not changed between sample locations (change of support)
-
+One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
 Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
@@
 The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+taking the mean of all height pixels of a raster canopy height model will almost perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
 

### 2017-10-02 JStatSoft_forestinventory_master-concordance.tex
update Andreas 02.10.2017

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 954:%
-1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1473:%
+1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1403:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1485:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1415:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1550:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1480:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1560:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1490:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1846:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1776:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1856:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1786:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2087:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2017:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2096:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2026:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2101:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2031:%
 194 11 1}

### 2017-10-02 rnw/Main structure and modules of the Package.Rnw
update Andreas 02.10.2017

@@
 \caption{Artificial representation of a local density surface.
-The spatial distribution of a hypothetical density function for every point in a forested area is represented as wavy piecewise constant green surface.
+The spatial distribution of a hypothetical density function for every point in a forested area is represented as a wavy piecewise constant green surface.
 Sample plots (white dots) inform the inventorist of the value of the density function at that point.}

### 2017-10-02 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 02.10.2017

@@
 
-The the vector of regression coefficients for the \textbf{\textit{reduced model}} using the \textit{reduced set of explanatory variables} $\pmb{Z}^{(0)}(x)$ available at all sample locations $x \in s_0$ is derived by solving the normal equation
+Solving the sample based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the \textbf{\textit{reduced model}}, i.e.
+using the \textit{reduced set of explanatory variables} $\pmb{Z}^{(0)}(x)$ available at $x \in s_0$, and likewise the vector of regression coefficients $\hat{\pmb{\beta}}_{s_2}$ for the \textbf{\textit{full model}}, i.e.
+using the \textit{full set of explanatory variables} $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ available only at a subset $x \in s_1 \subset s_0$, are derived as
 
-\begin{eqnarray}\label{eq:normequ_redmod}
-\hat{\pmb{\alpha}}_{s_2} &=& \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)
-\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)
-\end{eqnarray}
-
-and likewise the vector of regression coefficients for the \textbf{\textit{full model}} using the \textit{full set of explanatory variables} $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ available only at a subset of all sample locations $x \in s_1 \subset s_0$:
-
-\begin{eqnarray}\label{eq:normequ_redmod}
-\hat{\pmb{\beta}}_{{s_2}}&=&\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^t(x)
-\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x)
-\end{eqnarray}
-
-The variance-covariance matrices of the regression coefficients are thus defined as
-
-\begin{subequations}\label{eq:covar3p}
+\begin{subequations}\label{eq:normequ_3p}
 \begin{align}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \label{eq:covar_beta}
+\hat{\pmb{\alpha}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)  \label{eq:normequ_redmod} \\
+\hat{\pmb{\beta}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^t(x)
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x) \label{eq:normequ_fullmod}
 \end{align}
@@
 
-with the full model residuals ${R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ and the reduced model residuals ${R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ derived under the reduced and the full regression model respectively.
-
-Likewise the two-phase estimation (section \ref{sec:twophase_and_appl}), also the three-phase estimation can be applied under \textit{non-exhaustive} and \textit{exhaustive} derivation of the explanatory variables at the $s_0$ phase.
-
-In the \textbf{non-exhaustive case}, we use the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase (equation \ref{meanvalues3pnex}).
-If the explanatory variables used in the reduced model are exhaustively derived (\textbf{exhaustive case}), the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ is simply replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
-Note that in case of applied boundary adjustment (section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to equation \ref{eq:wmeanZ}.
-
-\begin{equation}\label{meanvalues3pnex}
-\hat{\bar{\pmb{Z}}}^{(0)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(0)}(x) ,
-\quad \hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}(x)
-\end{equation}
-
-The internal model case again insures the zero mean residual property over $F$, and the \textbf{point estimates} are thus calculated as
+The package allows for the calculation of \textbf{point estimates} under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase.
+Staying in the internal model approach again insures the zero mean residual property over $F$.
 
@@
 
-Equations \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} nicely demonstrate the philosophy of the three-phase estimation approach.
-Without the first term, the equations exactly equal the one of two-phase estimator where the estimated means in $\hat{\bar{\pmb{Z}}}_1$ are based on the $s_1$ sample and are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$.
-The difference however is that we can now derive a more precise mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ (or even the exact mean $\bar{\pmb{Z}}^{(0)}_0$) over $s_0$ for a subset $\pmb{Z}^{(0)}(x) \in \pmb{Z}(x)$.
-The first term of equation \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} can thus be interpreted as a correction term that transforms the difference to the more precise / exact mean of $\pmb{Z}^{(0)}(x)$ into a residual of $Y(x)$ and adds it to the second term, thus making the point estimate more precise.
+In the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to \ref{meanvalues3pnex}.
+If the exact mean over $s_0$ is known, the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ can simply be replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
+Note that in case of applied boundary adjustment (section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to equation \ref{eq:wmeanZ}.
 
-The \textbf{design-based variance} is then calculated as
+\begin{equation}\label{meanvalues3pnex}
+\hat{\bar{\pmb{Z}}}^{(0)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(0)}(x) ,
+\quad \hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}(x)
+\end{equation}
+
+An intuitive interpretation of \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex} is a follows: Likewise in the twophase approach, the estimated means in $\hat{\bar{\pmb{Z}}}_1$ based on the $s_1$ sample are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$.
+The difference between the estimated means $\hat{\bar{\pmb{Z}}}^{(0)}_1$ and their more precise or even exact counterparts $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and $\bar{\pmb{Z}}^{(0)}_0$ based on the larger or infinite $s_0$ sample is then added as a residual term that expresses the lack of model precision due to varying sample sizes behind the estimated auxilary means.
+
+The package again provides the g-weight and external variances. The \textbf{g-weight variance} formulation is
 
@@
 
-with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}$:
+with the variance-covariance matrix of $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and the variance-covariance matrices of the regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ and $\hat{\pmb{\beta}}_{s_2}$:
 
-\begin{equation}\label{estvarcovaux3p}
-\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}=
+\begin{subequations}\label{eq:covar3p}
+\begin{align}
+  \hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}&=
   \frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
-(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t
-\end{equation}
+  (\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \label{eq:covar_beta}
+\end{align}
+\end{subequations}
 
-We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
-This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.
-
-The package also gives the \textbf{external variance} defined as
+Note that ${R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the residuals of the full model, whereas ${R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the residuals of the reduced model.
+The \textbf{external variance} form under linear regression models is defined as
 
@@
 
+% We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
+This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.
+
 % ------------------------------------- %
@@
 
-Note that the \code{grisons} inventory data set originally corresponds to a classical two-phase sample design, i.e.
-the full set of explanatory variables is available at all sample locations in the inventory area.
-To demonstrate the three-phase estimation procedure, we provide an artificial three-phase scenario by recoding the phase indicators in the data set (column \code{phase_id_3p}).
-The values for indicating the phases (\code{0,1,2}) follow the terminology used in this article ($s_0$, $s_1$, $s_2$).
-
-We can use the \code{table()} function to see the sample sizes of all phases for the three-phase scenario:
-
-\begin{small}
-<<>>=
-table(grisons$phase_id_3p)
-@
-\end{small}
-
-We see that in the three-phase example, we now have 40 terrestrial sample locations providing the timber volume per hectare (\code{tvol}).
-The large phase $s_0$ comprises $178 + 88 + 40 = 306$ sample locations, since $s_2 \subset s_1 \subset s_0$.
-We now assume that at sample locations $x \in s_0$, only the mean canopy height (\code{mean}) is available, whereas at sample locations $x \in s_1$ we additionally have the explanatory variables \code{stddev}, \code{max} and \code{q75} at hand.
-We can now define the reduced and full regression model formulas that we will use in the \code{threephase()} function.
-Note that the models are nested.
+In order to demonstrate the three-phase estimators in the package, we created an artificial three-phase scenario by recoding the phase indicators in the \code{grisons} data set (column \code{phase_id_3p}) according to the terminology used in this article (\code{0} for $s_0$, \code{1} for $s_1$, \code{2} for $s_2$).
+We now assume that the mean canopy height (\code{mean}) is available at all 306 sample locations $x \in s_0$, whereas we have the explanatory variables \code{stddev}, \code{max} and \code{q75} only at 128 subsamples $s_1$ of $s_0$.
+At 40 further subsamples $s_2$ we have the observations $Y(x)$ from the field inventory.
+Based on this setup, we can now define the reduced and full regression model formulas to be used in the \code{threephase()} function (note that the models are nested).
 
@@
 
-Note that compared to the \code{twophase()}-function, we now have to specify two regression models, i.e.
+Compared to the \code{twophase()}-function, we now have to specify two regression models, i.e.
 the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
@@
 
-The \code{summary()} of a \code{threephase()}-function recalls both regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
-The set of sample sizes have also been extented by the $s_0$ phase (\code{n0}).
-In order to compare the results to the respective two-phase estimation, we assume that we have the mean canopy height as the only explanatory variable at all 306 sample locations:
-
-\begin{small}
-<<>>=
-reg2p_nex<- twophase(formula = formula.rm, data = grisons,
-                       phase_id = list(phase.col="phase_id_2p", terrgrid.id = 2),
-                       boundary_weights = "boundary_weights")
-@
-\end{small}
-
-\begin{small}
-<<>>=
-reg2p_nex$estimation
-@
-\end{small}
-
-Comparing the $R^2$s of the three and two-phase result, it becomes obvious that the incorporation of the additional explanatory variables \code{stddev}, \code{max} and \code{q75} led to a 20\% increase in explained variance in the regression model.
-While the more precise regression model was only applied at 40\% of all 306 sample locations (\textit{sampling fraction}), it was possible to considerably reduce the two-phase variance and thus achieve a higher estimation precision.
+The \code{summary()} of a \code{threephase()}-function now recalls \textit{both} regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
+We are told that including \code{stddev}, \code{max} and \code{q75} besides \code{mean} yields a 20 \% improvement in R$^2$.
+If we used \code{mean} as the only auxiliary variable at all sample locations (two-phase approach), we would see that the three-phase extention led to a considerable reduction in variance.
 
@@
 
-The 3 small area estimators under three-phase sampling provided by the package extend the two-phase small area estimator of section \ref{sec:twophase_sae} to the three-phase case.
-Note that the general principle stays the same, i.e.
-all information used is restricted to the small area $G$, with exception of retrieving the regession coefficients of the reduced and full model and their variance-covariance matrices.
+The three two-phase small area estimators described in section \ref{sec:twophase_sae} can also be extended to the three-phase scenario.
+The general principle thereby stays the same, i.e.
+the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to equations \ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.
 
-We first define the estimated means of the explanatory variables within the small area $G$:
-
-\begin{equation}\label{meanvalues3psaenext}
-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x) ,
-\quad \hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)
-\end{equation}
+% We first define the estimated means of the explanatory variables within the small area $G$:
+% 
+% \begin{equation}\label{meanvalues3psaenext}
+% \hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x) ,
+% \quad \hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)
+% \end{equation}
 
@@
 
-The \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) follow the same principle as in the two-phase case, i.e.
-applying the globally derived reduced and full regression model to the small area means of the explanatory variables, and then correcting for a potential model bias by adding the residual mean of the full model, which is not zero in general, to the point estimate (equations \ref{eq:pe_3p_small} and \ref{eq:pe_3p_psmall}).
-Note that the difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ derived in the $s_1$ phase and the more precise mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ or exact mean $\bar{\pmb{Z}}^{(0)}_{0,G}$ derived in the $s_0$ phase is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
-
-The \textbf{point estimates} are unbiased and given by:
+The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
+${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate.
+The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
 
@@
 \hat{Y}_{G,small,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_small} \\
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_small} \\
 \hat{Y}_{G,psmall,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\bar{\hat{R}}_{G}(x) \label{eq:pe_3p_psmall}
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_psmall}
 \end{align}
@@
 
-with ${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$ being the full model residuals in small area $G$. The \textbf{design-based variance} is then calculated as
+The \textbf{g-weight variance} is then calculated as
 
@@
 
-\begin{subequations}\label{eq:pest_3p_synth_psynth}
+\begin{subequations}\label{eq:3p_synth_psynth}
 \begin{align}
@@
 \hat{Y}_{G,psynth,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth}
-\end{align}
-\end{subequations}
-
-\begin{subequations}\label{eq:var_3p_synth_psynth}
-\begin{align}
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
 \hat{\var}(\hat{Y}_{G,synth,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
@@
 
+% \begin{subequations}\label{eq:var_3p_synth_psynth}
+% \begin{align}
+% \hat{\var}(\hat{Y}_{G,synth,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+% \bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+% \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
+% \hat{\var}(\hat{Y}_{G,psynth,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
+% \frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+% \hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+% \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
+% \end{align}
+% \end{subequations}
+
 %------------------------------------------------------------- %
@@
 This practically means that we have to specify the binary indicator variable for \textit{all} observations in the data frame that is passed to the package-functions.
-
-We first recalculate the regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ under the extended auxiliary vector as
+The regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ under the extended auxiliary vector are thus recalculated as
 
@@
 \hat{\pmb{\theta}}_{{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)
-\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x)
-\end{align}
-\end{subequations}
-
-and
-
-\begin{subequations}\label{eq:covar3p_ext}
-\begin{align}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
+\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x) \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)\Big)
   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1} \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)\Big)
@@
 
-After computing the auxiliary means $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x)$, $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x)$, and $\hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)$, we can calculate the \textbf{point estimate} as
+With the auxliary means and their variance-covariance matrices following the same notational change, the \textbf{point estimate} and \textbf{g-weight variance} are derived as
 
@@
 \hat{Y}_{G,extpsynth,3p}&=(\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
-\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth}
-\end{align}
-\end{subequations}
-
-and the \textbf{design-based variance} defined as
-
-\begin{subequations}\label{eq:var_synth_psynth_ext}
-\begin{align}
+\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth} \\
 \hat{\var}(\hat{Y}_{G,extsynth,3p})& =\frac{n_2}{n_1}\bar{\pmb{\mathbb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
@@
 \end{subequations}
-
-with
-
-\begin{equation}\label{estvarcovaux3pG_ext}
-\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}=
-\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
-(\pmb{\mathbb{Z}}^{(0)}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G})(\pmb{\mathbb{Z}}^{(0)}(x)-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G})^t
-\end{equation}
 
@@
 
-Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
+% Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
 In the first scenario, more precise auxiliary information becomes available at a subset of the allready existing and used sample locations.

### 2017-10-03 JStatSoft_forestinventory_master-concordance.tex
update Andreas 03.10.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
-1 132 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 356:%
+1 134 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 358:%
 134 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 368:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 370:%
 1 172 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 953:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 955:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 954:%
-1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 956:%
+1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1403:%

### 2017-10-03 rnw/Special Cases and Scenarios.Rnw
update Andreas 03.10.2017(1)

@@
 
-Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the design-based variances only differ marginally.
+Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the g-weight variances only differ marginally.
 This can be taken as evidence that the violation of the zero mean residual property can here expected to be negligible.

### 2017-10-03 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 03.10.2017(1)

@@
 with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$ being the reduced model predictions and $\hat{R}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
-Likewise the design-based variance, the difference between the exhaustive and non-exhaustive form is the variance of the predictions $\hat{Y}_{0}(x)$ within the $s_{0,G}$ phase.
+Likewise the g-weight variance, the difference between the exhaustive and non-exhaustive form is the variance of the predictions $\hat{Y}_{0}(x)$ within the $s_{0,G}$ phase.
 
@@
 $n_{2,G}=0$.
-Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and design-based variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
+Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
 The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$.
@@
                   boundary_weights = "boundary_weights")
-
 @

### 2017-10-03 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 03.10.2017(1)

@@
 
-where the design-based variance-covariance matrix of $\beta_{s_2}$ is calculated as
+where the g-weight variance-covariance matrix of $\beta_{s_2}$ is calculated as
 
@@
 The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
-A readable \textbf{summary of the estimation results} can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the design-based and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:
+A readable \textbf{summary of the estimation results} can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the g-weight and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:
 
@@
 
-For practical use, one should normally always prefer the design-based variance over the external variance.
+For practical use, one should normally always prefer the g-weight variance over the external variance.
 This is because when we use internal models, the regression coefficients actually dependent on our sample.
-In contrast to the external variance, the design-based variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical calibration properties (g-weights).
-The external and design-based variances are asymptotically equivalent but the external variance is really only included here in case the user wants to compare to another estimator where no design-based variance exists.
+In contrast to the external variance, the g-weight variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical calibration properties (g-weights).
+The external and g-weight variances are asymptotically equivalent but the external variance is really only included here in case the user wants to compare to another estimator where no g-weight variance exists.
 
@@
 
-The package provides the \textit{design-based} variance for \textit{small} and \textit{psmall} respectively (Equations \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
+The package provides the \textit{g-weight} variance for \textit{small} and \textit{psmall} respectively (Equations \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
 Again note that all components are restricted to those available at the sample locations in the small area ($s_{1,G}$ and $s_{2,G}$), with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
@@
 However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
-Also note that the residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
+Also note that the residual variation can no longer be considered in the g-weight variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
 Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias.
@@
 
-To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small design-based variance due to its potential bias.
+To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
 When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias directly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply uses the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
@@
 However, \textit{psmall} and \textit{extended psynth} are design unbiased and their variances reflect the fact that they account for the uncertainty of the regression model predictions.
-The design-based variance of \textit{psynth} completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
+The g-weight variance of \textit{psynth} completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
 

### 2017-10-04 JStatSoft_forestinventory_master-concordance.tex
update Andreas 04.10.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 370:%
-1 172 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 133 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 955:%
+1 174 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 953:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 956:%
-1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1403:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 954:%
+1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1398:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1415:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1410:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1480:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1475:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1490:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1485:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1776:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1771:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1786:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1781:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2017:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2012:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2026:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2021:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2031:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2026:%
 194 11 1}

### 2017-10-04 rnw/Introduction.Rnw
update Andreas 04.10.2017(1)

@@
 However, they have the downside of being very expensive.
-In order to increase the precision of the estimates, the number of sample plots must be increased, which essentially means travel costs will ballon as trained inventorists are sent to more and more random plot locations.
+In order to increase the precision of the estimates, the number of sample plots must be increased, which essentially means travel costs will rise as trained inventorists are sent to more and more random plot locations.
 This is why the number of terrestrial samples is usually limited.
@@
 This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
+These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of remote sensing data available in the inventory area in large sample or even wall-to-wall.
 Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.

### 2017-10-04 rnw/Special Cases and Scenarios.Rnw
update Andreas 04.10.2017(1)

@@
 
-If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \not\in s_2$ sample, and thus missing in the $s_2$ sample.
+If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category level is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample.
 Conclusively, an internal regression model can not be calculated and the function \textbf{stops} with the following error message:

### 2017-10-04 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 04.10.2017(1)

@@
 The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively.
-If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but to the cost of ignoring the model uncertainties.
+If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but at the cost of ignoring the model uncertainties.
 
@@
 
-% Finally note that in practice, there are basically \textit{three scenarios} how a two-phase sample design might be extended to a three-phase estimation procedure.
-In the first scenario, more precise auxiliary information becomes available at a subset of the allready existing and used sample locations.
-In this case, one simply reclassifies the existing sample locations $x \in s_1$ into $s_0$ sample locations and assigns the $s_1$ indicator to those points where the new variables are available.
-In the second scenario, the auxiliary information is available at all sample locations $x \in s_1$, but the computation of some of the explanatory variables is computationally expensive.
-In this case, a considerable time saving could be realized by restricing their computation to a subset of all $s_1$ sample locations, which consequently leads to a three-phase scenario that is handeled likewise the first scenario.
-A respective example is given in \citet{mandallaz2013b} and \citet{mandallaz2013c}.
-In both the first and second scenario, the original sample size of the two-phase inventory remains unchanged, but the $s_1$ sample locations are splitted into $s_0$ and $s_1$ locations.
-In the third and last scenario, one has already used a set of explanatory variables at $s_1$ sample locations for a time before additional, less precise information becomes available in a much larger quantity over the inventory area.
-It still might yield an increase in estimation precision if this information is considered in the estimation in a much larger sample size than the more precise information (sample fraction).
-In this case, one creates a new phase $s_0$ by densifying the allready existing $s_1$ sample locations (which in practice are often arranged in regular grids) such that $n_0 >> n_1$ and $s_1 \in s_0$.
-
 \newpage

### 2017-10-04 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 04.10.2017(1)

@@
 		\centering
-    \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
+    %\resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
+    \resizebox{0.85\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
 		\caption{} \label{fig:exh_nexh_and_boundweights_a}
@@
 		\centering
-   \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
+   %\resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
+   \resizebox{0.85\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
 		\caption{} \label{fig:exh_nexh_and_boundweights_b}
@@
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
+A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
 w: weights used for calculating the weighted mean of each explanatory variable.
@@
   \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth} \\
-% \end{align}
-% \end{subequations}
-% \begin{subequations}\label{eq:var_2p_synth_psynth}
-% \begin{align}
   \hat{\var}(\hat{Y}_{G,synth,2p}) & = \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
@@
 Recalling that linear models fit using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
-The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Equations \ref{eq:pest_2p_synth_psynth}, \ref{eq:var_2p_synth_psynth} and  \ref{estvarcovaux_G}.
+The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Equations \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
 Note that the package functions internally extent the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.

### 2017-10-04 rnw/Analysis and Visualization.Rnw
update Andreas 04.10.2017(2)

@@
 <<>>=
-extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
+extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                         phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 <<>>=
-extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
+extpsynth_3p <- threephase(formula.s0 = tvol ~ mean,
                 formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
@@
 <<>>=
-psynth_3p<- threephase(formula.s0 = tvol ~ mean,
+psynth_3p <- threephase(formula.s0 = tvol ~ mean,
                 formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
@@
 <<>>=
-grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
+grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                              extpsynth_3p, psynth_3p),

### 2017-10-04 rnw/Special Cases and Scenarios.Rnw
update Andreas 04.10.2017(2)

@@
 <<>>=
-grisons$stage<- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
+grisons$stage <- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
 @
@@
 <<echo=FALSE>>=
-extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 \begin{Sinput}
-R> extpsynth.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
 +                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 <<>>=
-extpsmall.clust<- twophase(formula = basal ~ stade + couver + melange, data=zberg,
+extpsmall.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 ## delete "mean" value from an s2- (i.e. s1- and s0-) sample point:
-grisons[which(grisons$phase_id_3p==2)[1],"mean"]<- NA
+grisons[which(grisons$phase_id_3p==2)[1],"mean"] <- NA
 @
@@
 ## delete "q75" value from an s2-sample point:
-grisons[which(grisons$phase_id_3p==2)[1],"q75"]<- NA
+grisons[which(grisons$phase_id_3p==2)[1],"q75"] <- NA
 @
@@
 ## delete "mean" value from an s0-sample point:
-grisons[which(grisons$phase_id_3p==0)[1],"mean"]<- NA
+grisons[which(grisons$phase_id_3p==0)[1],"mean"] <- NA
 @
@@
 ## delete s2-points with "stade"-level '300'
-zberg<- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
+zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
 @

### 2017-10-04 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 04.10.2017(2)

@@
 <<>>=
-reg3p_nex<- threephase(formula.s0 = formula.rm, 
+reg3p_nex <- threephase(formula.s0 = formula.rm, 
                   formula.s1 = formula.fm, data = grisons,

### 2017-10-04 r/create figures.R

@@ -37,13 +37,6 @@ psynth_3p<- threephase(formula.s0 = tvol ~ mean,
                       boundary_weights = "boundary_weights")
           
 
-
-grisons.sae.table.CI<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
-                                                extpsynth_3p, psynth_3p),
-                                sae = TRUE, 
-                                vartypes = c("variance", "g_variance",  "ext_variance"))
-
-
 grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                                 extpsynth_3p, psynth_3p),
                                 sae = TRUE, 
@@ -51,7 +44,7 @@ grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
 
 
 
-plot(grisons.sae.table.CI, ncol = 2, yvar = "estimate")
+plot(grisons.sae.table, ncol = 2, yvar = "estimate")
 
 ## save graphic for article:
 ggsave("plot_est_ci.png", 
@@ -69,7 +62,6 @@ ggsave("plot_est_ci.png",
 
 
 plot(grisons.sae.table, ncol = 2, yvar = "error")
-
 
 ## save graphic for article:
 ggsave("plot_error.png", 

### 2017-10-05 rnw/Introduction.Rnw
Introduction merged 05.10.2017

@@
 However, they have the downside of being very expensive.
-In order to increase the precision of the estimates, the number of sample plots must be increased, which essentially means travel costs will rise as trained inventorists are sent to more and more random plot locations.
-This is why the number of terrestrial samples is usually limited.
-Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, areas with smaller sample sizes, such as forest management units, often are associated with undesirably high estimation errors.
-As a result, in recent years there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
+Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more random plot locations.
+This is why the number of terrestrial samples is often limited.
+Although national inventories usually possess sufficient the terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.
+As a result, there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
 A method which has become particularly attractive is called \textit{multiphase sampling}.
@@
 These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of remote sensing data available in the inventory area in large sample or even wall-to-wall.
-Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples of auxiliary information available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
+Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
 Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
@@
 % section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare.
-Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
-However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available.
-Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
+Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the \textit{finite population approach}, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
+However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}.
+The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the \textit{infinite population approach} by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
-The implemented estimators cover 32 inventory scenarios and can be used for stratification, regression and regression within strata \citep{massey2015}.
+The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}.
 The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
 
-The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} address special cases and demonstrate how the package-functions handle them.
+The objectives of this article are to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} address special cases and demonstrate how the package-functions handle them.
 

### 2017-10-05 JStatSoft_forestinventory_master-concordance.tex
merged section 2

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
-1 134 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 358:%
+1 71 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 295:%
 134 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 370:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 307:%
 1 174 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 953:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 890:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 954:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 891:%
 1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1398:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1335:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1410:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1347:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1475:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1412:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1485:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1422:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1771:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1708:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1781:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1718:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2012:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1949:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2021:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1958:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2026:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1963:%
 194 11 1}

### 2017-10-05 rnw/Main structure and modules of the Package.Rnw
update Andreas 03.10.2017(1)
merged section 2

@@
 
-Most forest inventories gather the field information by randomly or systematically selecting points in the forest area defined by coordinates to which inventorists are sent.
-At these points, the inventorist measures sample trees located within a constructed circle (called sample plot) around the sample point $x$, and aggregates their individual characteristics (e.g.
-their individual timber volumes) to local plot densities (e.g.
+Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by GPS coordinates.
+The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g.
+timber volumes) to local plot densities (e.g.
 the timber density in $m^3/ha$).
-The estimators implemented in the \pkg{forestinventory} package use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators.
-The infinite population approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a piecewise constant function, as visualized in Figure \ref{fig:inf_pop_apr}.
+The estimators implemented in \pkg{forestinventory} use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators.
+This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Figure \ref{fig:inf_pop_apr}.
 The \textit{population total} of the target variable (e.g.
 the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Figure \ref{fig:inf_pop_apr} as the volume under the density surface.
-From this perspective, the challenging part of an inventory is that the form of this function is unknown.
-Theoretically, we still could get the total timber volume by observing the function value, i.e.
-the local density $Y(x)$, at each possible point $x$ over the forest area and sum those obervations up.
+From this perspective, the practical challenge is that the form of this function is unknown.
+Theoretically, we could get the total timber volume by observing the function value, i.e.
+the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum.
 However, this is impossible because there is an infinite number of points in the forest area.
@@
 The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
-All estimators included in the \pkg{forestinventory} package basically rest upon this theoretical approach.
-A key point for practical implementation is that in the infinite population approach, a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not to the sample plot area.
-This has some obvious theoretical advantages over the finite population approach, where the sample units actually are cirular or rectangular plots.
-One is the impossibility of a perfect tesselation over an amorphous forest area by any kind of plot-shape, and hence, the population in the finite approach is, strictly speaking, not perfectly defined with respect to the forest area.
+All estimators included in \pkg{forestinventory} rest upon this theoretical perspective.
+The key point to understand in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
+This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
+This is mainly due to the impossibility of a perfect tesselation over an amorphous forest area by any choice of plot shape.
+Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area.
 The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
@@
 
-% "we are sampling values of a function that is defined by the spatial arrangement of the trees"
-
-% NOTES:
-% - Use onephase sampling to explain the difference between finite and infinite population approach
-% - infinite population approach: 
-%   - Lets assume that the spatial distribution of the target variable (e.g. the timber volume) over the forest area is determined by a fixed mathematical continuous function,
-%     and the total amount of the target variable (e.g. the total timber volume of the forest), lets call it $Y$, is the integral of that function, 
-%     in other words: the volume that expands under the surface of that function (picture). The thing is that in the inventory context,
-%     the form of this function is unknown. Theoretically, we still could get Y by observing the
-%     function value Y(x), called the local density, at each possible location $x$ over the function "plane" and sum those obervations up. However, we see at the first spot 
-%     that this is impossible, because
-%     the number of function values $Y(x)$ and thus the number of possible sample locations $x$ is $infinite$. So the only thing we can do is to get ourself a sample of Y(x)s to 
-%     $estimate$ $Y$ by $\hat{Y}$. Since we thereby sample from an infinite population of possible sample locations / function values, this approach is consequently called the 
-%     \textit{infinite population approach}.
-%   - Transferring this approach to forest inventory problems is indeed straightforward when we reconsider our task of estimating e.g. the true timber volume of the forest 
-%     as "estimating the integral of an unknown function that determines the timber volume distribution over our forest". What we do in practice is to visit each sample 
-%     location $x$ in the field and determine the respective timber volume value, i.e. local density, by measuring sample trees around each location. 
-%     The selection of sample trees is mostly done constructing one or multiple nested circles (sample plots) around a sample location, and include each tree that lies 
-%     within the plot boundaries. The local density value is then determined by aggregating the sample tree characteristics (e.g. their indiv. timber volume).
-%     Doing so, we get a sample of the infinite number of (sample locations and thus) timber volume values we could theoretically observe in our forest. 
-%  -  A key point to understand is that in the infinite population approach, a local density value Y(x) has no spatial association. Even if we used a spatial unit, i.e. 
-%    the plot, extended around the sample point to calculate Y(x), its value only corresponds to the sample point / plot center. The best example of non-existing 
-%    spatial association of Y(x) is indeed angle count sampling (ref). Whereas the lack of spatial association seems rather distracting in the first place, it actually 
-%    has some obvious practical advantages over the finite population approach where the sample units are indeed cirular or rectangular plots.
-%
-
 \subsection{Two-Phase Sampling}
@@
 The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$.
+The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ at each point $x \in s_1$.
 These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ (e.g.
+The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value, the local density $Y(x)$, of the target variable (e.g.
 the timber volume per hectare).
-The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS) technique.
-The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e.
+For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some model, which in \pkg{forestinventory} is always a linear model fit in $s_2$ using ordinary least squares (OLS).
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model biases, i.e.
 $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.
-In the design-based context, the two-phase estimator is unbiased regardless of the model used to produce the predictions.
+In the design-based context, the two-phase estimator is typically unbiased regardless of the model used to produce the predictions.
 This property comes from the assumption that each phase's sample is selected via simple random sampling (see section \ref{sec:db_vs_md}).
-%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.
 
@@
 
-The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Figure \ref{fig:concmphase_and_sae_a}).
-The basic assumption is that auxiliary information is available in two different frequencies, i.e.
-sample sizes.
-The phase $s_0$ provides a large number $n_0$ of auxiliary information, whereas the phase $s_1$ provides additional auxiliary information that is only available at $n_1$ subsamples of $s_0$.
-The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and $s_0$.
-The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$.
+Three-phase estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Figure \ref{fig:concmphase_and_sae_a}).
+The basic setup is that the explanatory variables calculated from the auxiliary information are available in two different frequencies.
+The phase $s_0$ provides the largest number $n_0$ of auxiliary sample locations, whereas the phase $s_1$ provides fewer auxiliary locations that are only available at $n_1$ subsamples of $s_0$.
+The terrestrial information is then collected at a further subsample $s_2$ of $s_1$.
+The motivation for three-phase sampling is that the additional set of explanatory variables available at $s_1$, now denoted $\pmb{Z}^{(1)}(x)$, adds considerable explanatory power to the set of variables available at all sample locations $x \in s_0$, denoted $\pmb{Z}^{(0)}(x)$.
 From that it follows that we can define \textit{two nested} regression models.
-The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
+The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
 The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
-Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(0)}(x)$ in the so called \textit{reduced model} at the sample locations $x \in s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{0}(x)$.
+Less accurate predictions, $\hat{Y}_{0}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
 If there is a significant gain in model precision between the reduced and the full model and the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
-%In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.
 
@@
 		\centering
-    \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
+    %\resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
+    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_a}
@@
 		\centering
-   \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
+   %\resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
+   \resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_b}
@@
 
-Small area estimation does not necessarily refer to small spatial areas but rather areas that contain little or no terrestrial sample.
-In the design-based approach, if there are no terrestrial plots in the small area then there is nothing you can do and you will be forced to depend on the model you choose and accept its potential bias.
-To formulate this mathematically, we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (Figure \ref{fig:concmphase_and_sae_b}).
-If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be too small to provide sufficient estimation precision (i.e.
-the confidence intervals will be very large).
-In this case, the application of multiphase estimation procedures can be efficient.
-However, the terrestrial sample size is too small to justify fitting a separate regression model just that area.
-The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and then to apply this  model to the small area.
-The potential bias of applying that model in $G$ is then corrected for by using the model residuals derived from the small area $G$.
-If no terrestrial samples fall into a small area $G$ (i.e.
-$n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased.
-% Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.
-
-% NOTES:
-% - borrowing strength
-% - both, two-and three-phase can be extended to sae estimators
-% - explain difference between synthetic and non-sysnthetic estimators
-% - soundly derive reasoning, starting with onephase, then twophase
-
-% \begin{figure}[htb]
-% 	\begin{subfigure}[t]{0.5\textwidth}
-% 		\centering
-%     \resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
-% 		\caption{} \label{fig:concmphase_and_sae_a}
-% 		\end{subfigure}
-% 	\begin{subfigure}[t]{0.5\textwidth}
-% 		\centering
-%    \resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
-% 		\caption{} \label{fig:concmphase_and_sae_b}
-% 	\end{subfigure}
-% \caption{\textbf{(a)} Concept of multiphase sampling.
-The square represents the forest area for which an inventory is being conducted.
-The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
-% \label{fig:concmphase_and_sae}
-% \end{figure}
+Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample.
+To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Figure \ref{fig:concmphase_and_sae_b}).
+As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient.
+However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals.
+The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and to apply this model to the small area.
+The potential bias of applying that model in $G$ is then corrected for by using the model residuals derived from that small area.
+If there are no terrestrial plots in $G$ (i.e.
+$n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator.
+These are called \textit{synthetic} estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.
 
@@
 
-The selection of a suitable prediction model is an essential part in the application of multiphase inventories.
-However, the requirements on the model differ fundamentally between two major mathematical approaches: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
-One consequence is that the model does not need to be fit from a probability sample, i.e.
-the sample locations could arbitrarily be chosen.
-However, another consequence is that for unbiased estimates it has to be insured that the model adequately describes the underlying stochastic process, in other words: the model assumptions have to be met and empirically verified.
-For example, if we use an ordinary least square regression model, implicit mathematical assumptions are (i) normality and homogeneity of the  target variable \textit{Y}(\textit{x}) at each value of the explanatory variables, and (ii) independence of the target variable \textit{Y}(\textit{x}) or the model residuals $R(x)$, i.e.
-\textit{Y}(\textit{x}) and $R(x)$ are assumed to be identically distributed, independent random variables and not to be correlated in any way.
-Whereas verifying the validity of model assumptions can be seen as evidence that the model adequately describes the underlying stochastic process, a proof remains infeasible.
-This leads to the situation that in the model-dependent framework, one has to \textit{trust} the model.
-In contrast to the model-dependent approach, the \textit{design-based} approach rests upon the randomization of the sample locations $x$.
+The subject of model selection gets a lot of attention in the field of forest inventory.
+This is why it is important to understand that the mathematical interpretation of how a model is used to produce estimates is fundamentally different between the design-based and model-dependent approaches.
+In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
+Although the model does not need to be fit from a probability sample, i.e.
+the sample locations could arbitrarily be chosen, the model should adequately describe the underlying stochastic process in order to efficiently ensure unbiased results.
+In practice this means that special attention must be made to ensure that the variable selection is appropriate to avoid overfitting, important variables are not omitted and all model assumptions are reasonably met through empirical verification.
+If a model is misspecified then estimation based on inference from that model may not be reliable.
+To put it simply, in the model-dependent framework one has to \textit{trust} the model.
+In contrast, the \textit{design-based} approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
 While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the observation $Y(x)$ at any location $x$ is fixed and \textit{not} the result of a stochastic process.
 Conclusively, the observation \textit{Y}(\textit{x}) still remains a random variable, but solemnly due to the random sample mechanism.
-A consequence of this approach is that the models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g.
-identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-dependent framework \citep{mandallaz2008}.
+A consequence of this approach is that the estimation properties of design-based regression estimators (i.e.
+unbiasedness) typically hold regardless of the model that is chosen.
 The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.
 They are therefore also referred to as \textit{model-assisted}.
-It should be noted that the random distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey.
-However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
-However, the variance will in most cases be slightly overestimated and thus lead to wider and thus more conservative confidence intervals \citep{mandallaz2013a}.
+It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to minimize travelling costs in the terrestrial survey.
+However, there is reasonable evidence that softening this assumption is acceptable for point and variance estimation as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
+The variance will in most cases be slightly overestimated and lead to wider, more conservative confidence intervals \citep{mandallaz2013a}.
 
-% Note that while the auxiliary information is most often provided by remote sensing data, that does not necessarily have to be the case.
-The auxiliary information can be any information that exhibits a correlation to the local density, i.e.
-target variable, $Y(x)$.
-%
-% e.g. past inventory data \citep{massey2014a}
+\newpage
 
@@
 
-% In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
-From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
-The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline.
-The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}).
-The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area.
-The first case allows for computing the exact mean $\bar{\pmb{Z}}$ of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$).
-In the second case, the mean of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$) can only be estimated by $\hat{\bar{\pmb{Z}}}$ which adds another component of uncertainty to the respective estimators.
-The package offers three small area estimators for both the exhaustive and non-exhaustive setup.
-The \textit{extended synthetic estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$.
-The resulting point estimates and variances are asymptotically design-unbiased.
-If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased.
-The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (\textit{one-phase estimation}) by the \code{onephase()} function.
-All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
-The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
-While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
-We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
-A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
-
-In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function.
+In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions.
 From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
@@
 The estimators are available under \textit{exhaustive} and \textit{non-exhaustive} use of the auxiliary data.
-Additionally, the package provides the option of calculating one-phase estimates solely based on terrestrial samples.
+Additionally, the package also can calculate one-phase estimates solely based on terrestrial samples.
 All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
@@
 While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
-We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix.
+We will also restrict discussion to simple sampling and provide the equations for cluster sampling in the Appendix.
 A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.

### 2017-10-05 JStatSoft_forestinventory_master-concordance.tex
twophase merged (1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 307:%
-1 174 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 890:%
+1 117 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 833:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 891:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 834:%
 1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1335:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1278:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1347:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1290:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1412:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1355:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1422:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1365:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1708:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1651:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1718:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1661:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1949:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1892:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1958:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1901:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1963:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1906:%
 194 11 1}

### 2017-10-05 rnw/Twophase_Estimators and their Application.Rnw
merged section 2
twophase merged (1)

@@
 \label{sec:twophase_globest}
-
-% NOTES:
-% - The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation})
-%   by the \code{onephase} function in \pkg{forestinventory}. This function can also be used to access the gain in efficiency of the mutliphase estimation procedures. 
-%
-% - we only use the non-cluster sampling here to illustrate:
-% a) reg-coef calculation, point estimates and two variances, important components
-% b) we introduce the boundary adjustments: explain why, give mathematical formula, mention that this is optional and can be left out, in which case the simple means of Z are used
-% c) exhaustive and non-exhaustive
-
-% Here, we give the \textbf{mathematical background} of:
-% \begin{itemize}
-%   \item the classical two-phase estimator, including:
-%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
-%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
-%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
-%   \item ...
-% \end{itemize}
-% 
-%
-
-% 
-% Here, we give the \textbf{application} examples:
-% \begin{itemize}
-%   \item example for non-exhaustive case: use grison dataset
-%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
-%   \item example for exhaustive case: use cluster set
-% \end{itemize}
 
@@
 The individual predictions can then be calculated as $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
-In the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations.
-This means that the vector of regression coefficients are always derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} function.
-While virtually all inventorists fit their models using the current inventory data, \citet{massey2015a} also consider scenarios using external models where the sample used to train the model is assumed to be taken from an independent source.
-However, such functions have yet to be implemented in the package at this time.\par
+Unless stated otherwise, \pkg{forestinventory} only uses \textit{internal models} to calculate estimates.
+This means that the model fit, i.e.
+$\hat{\pmb{\beta}}_{s_2}$, is derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} functions.
+While virtually all inventorists fit their models using the current inventory data, sometimes there is reason to use formulas derived from external models where the sample used to train the model is assumed to be taken from an independent source \citep{massey2015a}.
+However, this usually occurs when using a model other than the OLS regression model and is beyond the scope of the package at this time.\par
 
@@
 Note that for \textit{internal} linear models the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction which is why it does not appear in the point estimate.
+More explanation about how to obain the auxiliary means is given in the next subsection.
 
-The package \pkg{forestinventory} implements two kinds of variances for each of these point estimates: the \textit{g-weight} formulation, also referred to as \textit{design-based variance}, (see \ref{eq:gw_var_2p_reg_ex} and \ref{eq:gw_var_2p_reg_nex}) given in \citet{mandallaz2016}, and the \textit{external variance} formulation (see \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}) that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients.
+The \pkg{forestinventory} package implements two kinds of variances for each of these point estimates: the \textit{g-weight} formulation that accounts for the fact that our model is in fact \textit{internal}, and the \textit{external variance} formulation that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients \citep{mandallaz2016}.
 
@@
 
-Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (\ref{eq:estvarmatrix}).
+Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (Eq \ref{eq:estvarmatrix}).
 The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formation is possible, as is the case with non-parametric models like kNN.
@@
 \subsubsection{Calculation of Explanatory Variables}
+
+%% ALEX VERSION:
+% We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
+Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \underline{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in the Fig.
+\ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points).
+For a every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
+We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the point whereas the support is the spatial extent of the auxiliary information used to calculate those values.
+So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
+The \pkg{forestinventory} package calculates the empirical mean of the $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
+% 
+% The \underline{exhaustive} case requires more explanation.
+Since $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it, the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx$, involves the integral of a function of unkown form that determines Z(x) over $F$.
+% 
+% the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number of sample points $x$ ($n_1=\infty$).
+Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+% 
+% This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
+$n_1=\infty$).
+However, this is actually possible for some specific types of variables, such as when the auxiliary information is pixelated and the desired explanatory variable is the mean of pixel values in the support.
+However, for most types of explanatory variables we are going to try to get an approximation of $\bar{\pmb{Z}}$ that is only neglibly different.
+To do this we want \textit{all} of the wall-to-wall auxiliary information to be exploited.
+Figure \ref{fig:exh_nexh_and_boundweights_a} depicts such an implementation where the spatial arrangement of the supports (the dark green tiles) are tesselated to form a perfect partition over the inventory area.
+It has to be noted that this setup would allow for perfect calculation of the exact mean in the \textit{finite population approach} where the supports themselves would be considered the sampling units.
+While in the \textit{infinite population approach} this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large.
+% 
+% The perfect tesselation implementation may make it easier to calculate the exact auxiliary mean in the \textit{finite approach}, but it can also impose drawbacks.
+For example, the support size in practice should ideally be chosen in order to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) not to merely conform to the geometric constraints of a perfect tesselation.
+Furthermore, the size of these supports are assumed to be the same for each explanatory variable.
+This is not the case in the \textit{infinite population approach} where $\pmb{Z}(x)$ only refers to the sample point, not its spatial support.
+Thus, the supports are allowed to overlap if necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$ and each explanatory variable is allowed to have a different sized support as long as they are consistently used at all sample locations.
+In this respect, the \textit{infinite population approach} is more flexible.
+% 
+% It was mentioned that there are a few situations where the exact mean can be precisely calculated in the infinite population framework.
+The first common situation is when the explanatory variables are provided by polygon layers (e.g.
+map of development stages).
+In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
+The second case is when the exact mean can be calculated in one step, e.g.
+taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
 
@@
 In this case, the \pkg{forestinventory} package calculates the emprirical mean of the $\pmb{Z}(x)$ values given in the input data frame, i.e.
-$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
+$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
 The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association.
@@
 The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will almost perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
 
@@
 		\centering
-    %\resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
-    \resizebox{0.85\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
+    \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
 		\caption{} \label{fig:exh_nexh_and_boundweights_a}
@@
 		\centering
-   %\resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
-   \resizebox{0.85\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
+   \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
 		\caption{} \label{fig:exh_nexh_and_boundweights_b}
@@
 
-%% NOTES:
-% We will now quickly draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
-In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_1,...,x_p$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-Figure 4(b) shows how the derivation of Z(x) in the non-exhaustive case often looks like in practice, i.e.
-a regular terrestrial grid $s_2$ is given by a terrestrial inventory, densified to a larger sample $s_1$, and then supports are used to calculate the set of explanatory variables at the sample location x.
-This is in perfect aggreement with the just presented theory of the non-exhaustive estimators, except for using regular grids rather than randomly placed sample points, i.e.
-the forestinventory package will calculate the emprirical mean of the Z(x) values given in the input dataframe, i.e.
-$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
-% The exhaustive case however leaves some space of discussion concerning the calculation of $Z(x)$:
-%
-%
-% - first, mention what Z(x) means in practice, and how we calculate a variable stored in Z(x), i.e. use supports (maybe mention the choice of support size,
-%   and that it can have an influence on model accuracy). 
-% - Figure 4(b) shows how the derivation of the Z(x) in the non-exhaustive case often looks like in practice, i.e. a regular grid is given by a 
-%   terrestrial inventory, densified to an s1-sample, and then supports are used to calculate the set of explanatory variables at sample location x.
-%   This is in perfect aggreement with the just presented theory of the non-exhaustive estimators (except for using regular grids rather than randomly
-%   placed sample points) and the forestinventory package will just use the emprirical mean of the Z(x) values given in the input dataframe, i.e. 
-%   $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.
-% - The exhaustive case however leaves some space for the implementation:
-%   This is because likewise the local density $Y(x)$ gathered in the second phase $s_2$, also the calculation of Z(x) only referrs to 
-%   the sample point x, and has no spatial association (see section 2.1). Likewise the mean of the local density, the only way to determine the exact mean 
-%   of $\bar{Z}$ would be to calculate the integral, i.e. in practice calculate Z(x) at an infinite number ($n_1=\infty$) of sample points $x$. In practice,
-%   this is obviously impossible.
-% - Figure 4(a) shows the implemented solution of an "exhaustive use" of the auxiliary information as described in Mandallaz, Breschan, Hill (2013). Is has to be noted
-%   that this would perfectly enable for calculating the exact mean $\bar{Z}$ in the \textit{finite} population approach (i.e. we calculate Z(x) for the entire 
-%   population of supports, and can thus calculate the exact mean $\bar{Z}$). However, this is not equal to the exact mean $\bar{Z}$ in the infinite population
-%   approach for the mentioned reasons.
-% - What we can do in practice is to get a reasonable approximation of $\bar{Z}$. 
-%
-%
-% - mention that in the strict inf.pop.appr. frame, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample
-%   n1 of Z(x) to get a reasonable approximation of $\bar{Z}$. This is possible because likewise the local density $Y(x)$ gathered in the second phase $s_2$,
-%   also the calculation of Z(x) only referrs to the sample point, and has no spatial association.
-%
-% - mention that the support size thereby does not have to match the plotsize of the terrestrial inventory.
-In the strict inf.pop.appr.
-frame, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample n1 of Z(x) to get a reasonable approximation of $\bar{Z}$.
-This is possible because likewise the local density $Y(x)$ gathered in the second phase $s_2$, also the calculation of Z(x) only referrs to the sample point, and has no spatial association.
-In fact, the support size will in practice be chosen in order to achieve a best possible explanatory power of the regression model, thus minimzing the residual variation.
-mention that a change of support size between sample points for a given expl.
-variable is not allowed (because....
-Daniel ?)
-%
-%
-
-%% ------------ in progress -------------------- %%
-
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{threephase()} function of the package.
+An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}}.
 In forest inventories, the sample is often restricted to those sample locations located within the forest area.
 In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
-by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (Figure \ref{fig:exh_nexh_bw}).
-This method was suggested in \citet{mandallaz2013b} and led to an improvement of the regression model accuracy.
-To insure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}(x)$ or $\bar{\pmb{Z}}(x)$, the respective means have then to be calculated as the \textit{weighted} mean according to \ref{eq:wmeanZ}.
-The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.
+by a polygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (see Fig.
+\ref{fig:exh_nexh_bw}).
+This method was suggested in \citet{mandallaz2013b} and led to an improvement estimation precision.
+In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the \textit{weighted} mean where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.
 
@@
 To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository.
-The dataset contains data from a two-phase forest inventory under simple (i.e.
-non-cluster) sampling design conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
-The $s_1$ same is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
+The dataset contains data from a simple (i.e.
+non-cluster) two-phase forest inventory conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
+The $s_1$ sample is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
 Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
@@
 For practical use, one should normally always prefer the g-weight variance over the external variance.
-This is because when we use internal models, the regression coefficients actually dependent on our sample.
+This is because when we use internal models, the regression coefficients actually depend on the terrestrial sample realized by the sampling design.
 In contrast to the external variance, the g-weight variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical calibration properties (g-weights).
@@
 
-The package \pkg{forestinventory} provides three broad types of small area estimators each of which has an exhaustive and non-exhaustive form.
+The \pkg{forestinventory} package provides three broad types of small area estimators each of which has an exhaustive and non-exhaustive form.
 We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label \textit{pseudo} to indicate that the mean of the explanatory variables within the small area was based on a finite sample.
-The main idea for all these small area estimators is to calculate the regression coefficient vector $\hat{\pmb{\beta}}_{s_2}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ on the entire $s_2$ sample according to equations \ref{normequ_simple} and \ref{eq:estvarmatrix}, and subsequently use that to make predictions for sample locations restricted to small area $G$.\par
+The main idea for all these small area estimators is to calculate the regression coefficient vector $\hat{\pmb{\beta}}_{s_2}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ on the entire $s_2$ sample according to Eq.
+\ref{normequ_simple} and \ref{eq:estvarmatrix}, and subsequently use that to make predictions for sample locations restricted to small area $G$.\par
 
@@
 In the equations for the point estimates (\ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
-A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$ to the first term.
+A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$.
 This is called the \textit{bias} or \textit{residual correction} term.
@@
 $n_{2,G}=0$).
-In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$.
+In this case, the point estimates (Equations \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
 However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Equations \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
@@
 However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
-Also note that the residual variation can no longer be considered in the g-weight variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
+Also note that the residual variation can no longer be considered in the g-weight variance (Eq.
+\ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}).
 Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias.
@@
 
-The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates as well as unreliably small variances.
+The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively.
 The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e.
@@
 The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Equations \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
-Note that the package functions internally extent the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
+Note that the package functions internally extend the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
 
@@
 The package also provides the external variance for both the \textit{extended} synthetic and \textit{extended} pseudo synthetic estimator.
-Note that both neither the extended model approach nor external variance estimates are possible in the absence of terrestrial samples und thus model residuals in $G$, which is precisely when one must rely on the (pseudo) synthetic estimates.
+Note that neither the extended model approach nor external variance estimates are possible in the absence of terrestrial samples and thus model residuals in $G$, which is precisely when one must rely on the (pseudo) synthetic estimates.
 The external variance forms of \textit{extsynth} and \textit{extpsynth} are
@@
 The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need to implement it yourself.
-Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator are vary between small areas, while they are identical under the \textit{psmall} estimator.
+Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator vary between small areas, while they are identical under the \textit{psmall} estimator.
 This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
 In case of the  \textit{psmall} estimator, the regression model stays the same for each small area estimation.
-Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling (see section \ref{sec:speccas_and_scen}).\par
+Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling (see Section \ref{sec:speccas_and_scen}).\par
 
@@
 We see here that the \textit{psynth} variances are almost only half the variances of the \textit{psmall} and \textit{extended psynth} estimators.
-However, \textit{psmall} and \textit{extended psynth} are design unbiased and their variances reflect the fact that they account for the uncertainty of the regression model predictions.
+However, \textit{psmall} and \textit{extended psynth} are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions.
 The g-weight variance of \textit{psynth} completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par

### 2017-10-05 JStatSoft_forestinventory_master-concordance.tex
merge twophase (2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 307:%
-1 117 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 833:%
+1 132 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 848:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 834:%
-1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 175 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1278:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 849:%
+1 102 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 169 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1284:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1290:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1296:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1355:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1361:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1365:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1371:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1651:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1657:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1661:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1667:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1892:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1898:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1901:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1907:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1906:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1912:%
 194 11 1}

### 2017-10-05 rnw/Threephase_Estimators and their Application.Rnw
merge twophase (2)

@@
 The package allows for the calculation of \textbf{point estimates} under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase.
-Staying in the internal model approach again insures the zero mean residual property over $F$.
+Staying in the internal model approach again ensures the zero mean residual property over $F$.
 
@@
 
-% We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase.
-This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.
-
 % ------------------------------------- %
@@
 
-The \code{summary()} of a \code{threephase()}-function now recalls \textit{both} regression model formulas and also gives the coefficients of determination $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) regression model.
-We are told that including \code{stddev}, \code{max} and \code{q75} besides \code{mean} yields a 20 \% improvement in R$^2$.
+The \code{summary()} of a \code{threephase()}-function now recalls \textit{both} regression model formulas and also gives the $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) models.
+We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$.
 If we used \code{mean} as the only auxiliary variable at all sample locations (two-phase approach), we would see that the three-phase extention led to a considerable reduction in variance.
@@
 
-The three two-phase small area estimators described in section \ref{sec:twophase_sae} can also be extended to the three-phase scenario.
+The three two-phase small area estimators described in Section \ref{sec:twophase_sae} can also be extended to the three-phase scenario.
 The general principle thereby stays the same, i.e.
 the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to equations \ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.
-
-% We first define the estimated means of the explanatory variables within the small area $G$:
-% 
-% \begin{equation}\label{meanvalues3psaenext}
-% \hat{\bar{\pmb{Z}}}^{(0)}_{0,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}^{(0)}(x) ,
-% \quad \hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)
-% \end{equation}
 
@@
 with $\hat{Y}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$ being the reduced model predictions and $\hat{\mathbb{R}}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
-The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in equation \ref{eq:pe_3p_synth}, \ref{eq:pe_3p_psynth}, \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}.
+The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in Eq.
+\ref{eq:pe_3p_synth}, \ref{eq:pe_3p_psynth}, \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}.
 Note that they are however design-unbiased since the residual correction term is included in the regression coefficient vector $\hat{\pmb{\theta}}_{s_2}$.

### 2017-10-05 rnw/Twophase_Estimators and their Application.Rnw
merge twophase (2)

@@
 
+% -----------------------
 %% ALEX VERSION:
-% We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
-Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \underline{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in the Fig.
+
+% spatial means of $p$ functions that respectively determine the function values $z_1(x),...,z_p(x) \in Z(x)$ at all possible locations in $F$.
+
+We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
+Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in the Fig.
 \ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points).
@@
 So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
-The \pkg{forestinventory} package calculates the empirical mean of the $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
+The \pkg{forestinventory} package calculates the empirical mean of the $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par %% ok!
+
+% The \textbf{exhaustive} case requires a closer look.
+Since being in the infinite population approach, $\pmb{Z}(x)$ again refers to the sample point $x$ and not the area around it.
+That implies that deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx$, involves computing the spatial means of all component, i.e.
+$z_1(x),...,z_p(x)$, by calculating those at all possible locations in $F$.
+This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
+$n_1=\infty$).
+Of course this is practically infeasible, but there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
+The first case is when the explanatory variables are provided by polygon layers (e.g.
+map of development stages).
+In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
+The second case is when the exact mean can be calculated in one step, e.g.
+taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+
+The \textbf{exhaustive} case requires a closer look.
+Since being in the infinite population approach, $\pmb{Z}(x)$ again refers to the sample point $x$ and not the area around it.
+Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{x \in F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{x \in F} z_p(x)dx)^t$, implies that we need to calculate the spatial means of each component of $\pmb{Z}(x)$ using all possible points in $F$.
+This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
+$n_1=\infty$).
+Of course this is practically infeasible, but there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
+The first case is when the explanatory variables are provided by polygon layers (e.g.
+map of development stages).
+In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
+The second case is when the exact mean can be calculated in one step, e.g.
+taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+
+One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
+Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
+deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
+While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
+Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
+One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
+Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
+In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
+This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
+With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
+One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
+
 % 
-% The \underline{exhaustive} case requires more explanation.
-Since $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it, the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx$, involves the integral of a function of unkown form that determines Z(x) over $F$.
-% 
-% the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number of sample points $x$ ($n_1=\infty$).
-Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
-% 
-% This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
-$n_1=\infty$).
-However, this is actually possible for some specific types of variables, such as when the auxiliary information is pixelated and the desired explanatory variable is the mean of pixel values in the support.
+% However, this is actually possible for some specific types of variables, such as when the auxiliary information is pixelated and the desired explanatory variable is the mean of pixel values in the support.
 However, for most types of explanatory variables we are going to try to get an approximation of $\bar{\pmb{Z}}$ that is only neglibly different.
@@
 While in the \textit{infinite population approach} this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large.
+In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+% 
+% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
+Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
+deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
+While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
+Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
+One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
+Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
+In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
+This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
+With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
+One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
+
+% 
 % 
@@
 taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
-
-We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
-In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_1,...,x_p$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$.
-Figure \ref{fig:exh_nexh_and_boundweights_b} shows how the derivation of $\pmb{Z}(x)$ in the \textbf{non-exhaustive} case often looks like in practice, i.e.
-a regular terrestrial grid $s_2$ is given by a terrestrial inventory, densified to a larger sample $s_1$, and then supports are used to calculate the set of explanatory variables at the sample locations.
-This is in perfect aggreement with the just presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
-In this case, the \pkg{forestinventory} package calculates the emprirical mean of the $\pmb{Z}(x)$ values given in the input data frame, i.e.
-$\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
-The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association.
+% 
+% 
+% % -----------------------
+% %% my version:
+% The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association.
 Likewise the mean of the local density (i.e.
@@
 Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
-One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+% 
+% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
 In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
@@
 One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
-Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
+% Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
 The first case is when the explanatory variables are provided by polygon layers (e.g.

### 2017-10-06 JStatSoft_forestinventory_master-concordance.tex
merge twophase 3

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 307:%
-1 132 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 129 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 848:%
+1 110 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 823:%
 147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 849:%
-1 102 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 169 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1284:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 824:%
+1 101 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 162 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1251:%
 149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1296:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1263:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1361:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1328:%
 162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1371:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1338:%
 1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1657:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1624:%
 173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1667:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1634:%
 1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1898:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1865:%
 184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1907:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1874:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1912:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1879:%
 194 11 1}

### 2017-10-06 rnw/Threephase_Estimators and their Application.Rnw
merge twophase 3

@@
 
-Solving the sample based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the \textbf{\textit{reduced model}}, i.e.
+Solving the sample-based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the \textbf{\textit{reduced model}}, i.e.
 using the \textit{reduced set of explanatory variables} $\pmb{Z}^{(0)}(x)$ available at $x \in s_0$, and likewise the vector of regression coefficients $\hat{\pmb{\beta}}_{s_2}$ for the \textbf{\textit{full model}}, i.e.
@@
 The package allows for the calculation of \textbf{point estimates} under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase.
-Staying in the internal model approach again ensures the zero mean residual property over $F$.
+Fitting the model using $s_2$ (i.e.
+\textit{internally}) ensures the zero mean residual property over $F$.
 
@@
 
-In the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to \ref{meanvalues3pnex}.
+In the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
+\ref{meanvalues3pnex}.
 If the exact mean over $s_0$ is known, the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ can simply be replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
-Note that in case of applied boundary adjustment (section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to equation \ref{eq:wmeanZ}.
+Note that in case of applied boundary adjustment (Section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to Eq.
+\ref{eq:wmeanZ}.
 
@@
 
-Note that ${R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the residuals of the full model, whereas ${R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the residuals of the reduced model.
+Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the residuals of the reduced model.
 The \textbf{external variance} form under linear regression models is defined as
@@
 \begin{align}
-\hat{\var}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x))
-+ (1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s_2}}\hat{R}^2(x) \label{eq:extvar_reg3p_ex} \\
-\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_0(x)-\hat{\bar{Y}}_0)^2}{n_0-1} 
-+\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_0(x)-\bar{\hat{R_0}})^2 \nonumber\\
-&+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:extvar_reg3p_nex}
+\hat{\var}_{ext}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}_{0}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
+\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}_{0}(x))
++\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}_{0}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:extvar_reg3p_nex}
 \end{align}
 \end{subequations}
+where $\hat{\var}_{s_0}$ indicates taking the sample variance over $s_0$.
 
@@
 The general principle thereby stays the same, i.e.
-the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to equations \ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.
+the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to Eq.
+\ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.
 
@@
 ${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate.
-The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (equation \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
+The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (Eq.
+\ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
 
@@
 \begin{flalign}
-\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}
- (\hat{R}_0(x)-\hat{\bar{R}}_{0,G})^2 + 
- (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_G)^2 \label{eq:var_3p_reg_small}&\\
-\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}(\hat{Y}_0(x)-\hat{\bar{Y}}_{0,G})^2 + 
-{\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
- (\hat{R}_0(x)-\bar{\hat{R}}_{0,G})^2 \nonumber &\\
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\bar{\hat{R}}_G)^2 \label{eq:var_3p_reg_psmall}
+\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}_{0}(x)) + 
+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
+\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}_{0}(x)) \nonumber &\\
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
 \end{flalign}
 \end{subequations}
-
-with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$ being the reduced model predictions and $\hat{R}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
+where $\hat{R}_{0}(x)=Y(x)-\hat{Y}_{0}(x)$ with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$.
 Likewise the g-weight variance, the difference between the exhaustive and non-exhaustive form is the variance of the predictions $\hat{Y}_{0}(x)$ within the $s_{0,G}$ phase.
@@
 \begin{flalign}
-\hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}
- (\hat{\mathbb{R}}_0(x)-\hat{\bar{\mathbb{R}}}_{0,G})^2 + 
- (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}}_G)^2 \label{eq:var_3p_reg_extsynth}&\\
-\hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}(\hat{Y}_0(x)-\hat{\bar{Y}}_{0,G})^2 + 
-{\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
- (\hat{\mathbb{R}}_0(x)-\bar{\hat{\mathbb{R}}}_{0,G})^2 \nonumber &\\
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}}_G)^2 \label{eq:var_3p_reg_extpsynth}
+\hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) + 
+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extsynth}&\\
+\hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) \nonumber & \\
+ &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extpsynth}
 \end{flalign}
 \end{subequations}
-
-with $\hat{Y}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$ being the reduced model predictions and $\hat{\mathbb{R}}_{0}=Y(x)-\hat{Y}_{0}$ being the reduced model residuals.
+where $\hat{\mathbb{R}}_{0}=Y(x)-\hat{\mathbb{Y}}_{0}(x)$ with $\hat{\mathbb{Y}}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$.
 The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in Eq.
 \ref{eq:pe_3p_synth}, \ref{eq:pe_3p_psynth}, \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}.
-Note that they are however design-unbiased since the residual correction term is included in the regression coefficient vector $\hat{\pmb{\theta}}_{s_2}$.
+Just as in the two-phase context, they are design-unbiased since the residual correction term is included in the regression coefficient vector $\hat{\pmb{\theta}}_{s_2}$.
 

### 2017-10-06 rnw/Twophase_Estimators and their Application.Rnw
merge twophase 3

@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}(R(x))=\frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(\hat{R}(x)-\bar{\hat{R}}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}(R(x)) \nonumber \\ & =
- \frac{1}{n_1}\frac{1}{n_1-1}\sum_{x\in{s_1}}(\hat{Y}(x)-\bar{\hat{Y}}_{s_1})^2 + \frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:varexternal_2p_reg_nex}
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:varexternal_2p_reg_ex} \\
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}_{s_1}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \nonumber  \label{eq:varexternal_2p_reg_nex}
 \end{align}
 \end{subequations}
+where $\hat{\var}_{s_2}$ and $\hat{\var}_{s_1}$ indicate taking the sample variance over $s_2$ and $s_1$ respectively.
 
@@
 With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
-
-% 
-% However, this is actually possible for some specific types of variables, such as when the auxiliary information is pixelated and the desired explanatory variable is the mean of pixel values in the support.
-However, for most types of explanatory variables we are going to try to get an approximation of $\bar{\pmb{Z}}$ that is only neglibly different.
-To do this we want \textit{all} of the wall-to-wall auxiliary information to be exploited.
-Figure \ref{fig:exh_nexh_and_boundweights_a} depicts such an implementation where the spatial arrangement of the supports (the dark green tiles) are tesselated to form a perfect partition over the inventory area.
-It has to be noted that this setup would allow for perfect calculation of the exact mean in the \textit{finite population approach} where the supports themselves would be considered the sampling units.
-While in the \textit{infinite population approach} this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large.
-In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
-% 
-% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
-In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
-Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
-deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
-While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
-Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
-One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
-Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
-In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
-This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
-With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
-
-% 
-% 
-% The perfect tesselation implementation may make it easier to calculate the exact auxiliary mean in the \textit{finite approach}, but it can also impose drawbacks.
-For example, the support size in practice should ideally be chosen in order to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) not to merely conform to the geometric constraints of a perfect tesselation.
-Furthermore, the size of these supports are assumed to be the same for each explanatory variable.
-This is not the case in the \textit{infinite population approach} where $\pmb{Z}(x)$ only refers to the sample point, not its spatial support.
-Thus, the supports are allowed to overlap if necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$ and each explanatory variable is allowed to have a different sized support as long as they are consistently used at all sample locations.
-In this respect, the \textit{infinite population approach} is more flexible.
-% 
-% It was mentioned that there are a few situations where the exact mean can be precisely calculated in the infinite population framework.
-The first common situation is when the explanatory variables are provided by polygon layers (e.g.
-map of development stages).
-In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
-The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
-% 
-% 
-% % -----------------------
-% %% my version:
-% The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association.
-Likewise the mean of the local density (i.e.
-the total), the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number of sample points $x$ ($n_1=\infty$).
-Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
-% 
-% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
-In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
-Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
-deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
-While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
-Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
-One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
-Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
-In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
-This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
-With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
-% Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
-The first case is when the explanatory variables are provided by polygon layers (e.g.
-map of development stages).
-In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
-The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.
 
@@
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
-A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon).
-w: weights used for calculating the weighted mean of each explanatory variable.
-Dotted circles indicate terrestrial sample plots.}
+A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon).}
 \label{fig:exh_nexh_bw}
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\
- &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
+  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
 \end{align}
 \end{subequations}
-
+where $\hat{\var}_{s_{2,G}}$ indicates taking the sample variance over $s_{2,G}$.
 If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & =
- \frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extpsynth}
+  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extsynth} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & = \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extpsynth}
 \end{align}
 \end{subequations}
+where $\hat{\mathbb{R}}(x)=Y(x)-\bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2}$.
 

### 2017-10-09 JStatSoft_forestinventory_master-concordance.tex
update Andreas 09.10.2017(spelling errors)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
-123 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 223:%
-1 71 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 295:%
-134 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 307:%
-1 110 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 823:%
-147}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 824:%
-1 101 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 162 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1251:%
-149 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1263:%
+123 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 218:%
+1 69 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
+129 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
+1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 803:%
+135}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
+1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1210:%
+137 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1215:%
 1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1328:%
-162 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1338:%
-1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 21 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1624:%
-173 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1634:%
-1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1865:%
-184 8 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1874:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1280:%
+143 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1285:%
+1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 12 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1557:%
+149 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1562:%
+1 11 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1784:%
+155 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1789:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1879:%
-194 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1794:%
+161 11 1}

### 2017-10-09 JStatSoft_forestinventory_master.rnw
update Andreas 18.09.2017(2)
update Andreas 05.10.2017(2)
update Andreas 09.10.2017(spelling errors)

@@
 \Abstract{
-Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.
+Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.
 They typically consist of a random sample of plot locations in the forest that are assessed individually by hand.
@@
 LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
-With respect to this objective, the application of \textit{multiphase} forest inventory methods (e.g.
+With respect to this objective, the application of \textit{multi-phase} forest inventory methods (e.g.
 double- and triple-sampling regression estimators) has proved to be efficient.
@@
 
+\setlength{\parindent}{1em} % Einzug bei neuen Absätzen
+
 %------------------------------------------------------------------------------------------------%
@@
 
-% NOTES:
-%
-%
-%
-%
-
 %------------------------------------------------------------------------------------------------%
@@
 
-% NOTES:
-% - mention difference between finite and infinite population approach ?
-%
-%
-%
-
 %------------------------------------------------------------------------------------------------%
 % ---------------------------------- Estimators and their Application -------------------------- %
-
-% \SweaveInput{rnw/Onephase_Estimators and their Application.Rnw}
 
@@
 \SweaveInput{rnw/Threephase_Estimators and their Application.Rnw}
-
-% NOTES:
-%
-%
-%
-%
 
@@
 
-% NOTES:
-%
-%
-
 %------------------------------------------------------------------------------------------------%
@@
 
-% NOTES:
-%
-%
-%
-
 %------------------------------------------------------------------------------------------------%
@@
 \SweaveInput{rnw/Analysis and Visualization.Rnw}
-
-% NOTES:
-% 
-%
-%
 
@@
 
-% -------------------------------------------------------------------------- %
+%------------------------------------------------------------------------------------------------%
 \end{document}

### 2017-10-09 rnw/Analysis and Visualization.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 \label{sec:anal_and_vis}
-
-% include:
-% - estTable()-function demo
-% - mphase.gain()-function demo
-% - plot.esttable()-function demo
-% - 3-4 pages max
-% - pronounce that "it is crucial to quantify the potential benefit of the mphase procedures
-%   compared to the onephase procedure --> allows to calculate the financial benefit --> rel.eff
 
@@
 
-As mentioned in the introduction (section \ref{sec:intro}), the application of multi-phase estimation techniques in forest practice have two main \textbf{goals}:  
+As mentioned in the introduction (Section \ref{sec:intro}), the application of multi-phase estimation techniques in forest practice have two main \textbf{goals}:  
 \begin{enumerate}

### 2017-10-09 rnw/Introduction.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 This is why the number of terrestrial samples is often limited.
-Although national inventories usually possess sufficient the terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.
+Although national inventories usually possess a sufficiently large terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.
 As a result, there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.

### 2017-10-09 rnw/Main structure and modules of the Package.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 The estimators implemented in \pkg{forestinventory} use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators.
-This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Figure \ref{fig:inf_pop_apr}.
+This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig.
+\ref{fig:inf_pop_apr}.
 The \textit{population total} of the target variable (e.g.
-the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Figure \ref{fig:inf_pop_apr} as the volume under the density surface.
+the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig.
+\ref{fig:inf_pop_apr} as the volume under the density surface.
 From this perspective, the practical challenge is that the form of this function is unknown.
@@
 All estimators included in \pkg{forestinventory} rest upon this theoretical perspective.
-The key point to understand in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
+The key point in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
 This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
@@
 
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}).
+The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Fig.
+\ref{fig:concmphase_and_sae_a}).
 The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ at each point $x \in s_1$.
 These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value, the local density $Y(x)$, of the target variable (e.g.
+The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
+the local density $Y(x)$ (e.g.
 the timber volume per hectare).
@@
 In the design-based context, the two-phase estimator is typically unbiased regardless of the model used to produce the predictions.
-This property comes from the assumption that each phase's sample is selected via simple random sampling (see section \ref{sec:db_vs_md}).
+This property comes from the assumption that each phase's sample is selected via simple random sampling (see Section \ref{sec:db_vs_md}).
 
@@
 
-Three-phase estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Figure \ref{fig:concmphase_and_sae_a}).
+Three-phase estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Fig.
+\ref{fig:concmphase_and_sae_a}).
 The basic setup is that the explanatory variables calculated from the auxiliary information are available in two different frequencies.
@@
 		\centering
-    %\resizebox{1.12\hsize}{!}{\includegraphics{fig/phases_graphic(3).eps}}%
     \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
@@
 		\centering
-   %\resizebox{0.8\hsize}{!}{\includegraphics{fig/sae_graphic(3).eps}}%
    \resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
@@
 Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample.
-To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Figure \ref{fig:concmphase_and_sae_b}).
+To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig.
+\ref{fig:concmphase_and_sae_b}).
 As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient.
@@
 If a model is misspecified then estimation based on inference from that model may not be reliable.
-To put it simply, in the model-dependent framework one has to \textit{trust} the model.
+In the model-dependent framework one thus has to \textit{trust} the model.
 In contrast, the \textit{design-based} approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
@@
 It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to minimize travelling costs in the terrestrial survey.
-However, there is reasonable evidence that softening this assumption is acceptable for point and variance estimation as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
+However, there is reasonable evidence that softening this assumption is acceptable for point and variance estimation as long as the grid does not interact with periodic features in the forest structure \citep{mandallaz2008}.
 The variance will in most cases be slightly overestimated and lead to wider, more conservative confidence intervals \citep{mandallaz2013a}.
@@
 In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions.
-From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Figure \ref{fig:struct_package}).
+From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig.
+\ref{fig:struct_package}).
 The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications.
@@
 While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
-We will also restrict discussion to simple sampling and provide the equations for cluster sampling in the Appendix.
-A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.
+We will also restrict discussion to simple sampling.
+A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.
 

### 2017-10-09 rnw/Special Cases and Scenarios.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
-Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables.
+Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations.
 
@@
 
-% NOTE:
-% - not going into too much detail
-% - rather briefly describe the important difference to simple sampling needed
-%   to understand the problem
-%
-
-As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample unit does no longer consist of one, but multiple spatially agglomerated sample points.
+As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample unit does no longer consist of one, but multiple spatially agglomerated sample points.
 One still randomly places the sample location $x$ in the inventory area likewise in the simple sampling design, but then creates $M-1$ additional sample locations $x_2,...,x_M$ close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
 The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign.
-Not however that the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and that this average value is again referred to exactly one point, i.e.
+Note however that the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and that this average value is again referred to exactly one point, i.e.
 the cluster origin $x$, in the frame of the infinite population approach.
-Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix).
-Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e defined by the user), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
+Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights.
+Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e.
+defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
 The package functions of \pkg{forstinventory} also allow to consider cluster-structures by specifying a unique cluster identification (cluster ID) for each cluster in an additional column whose column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.
@@
 only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
-In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see Section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
 In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured under this scenario.
@@
 Warning message:
-At least one terrestrial cluster not entirely included within the small area 1.
+At least one terrestrial cluster not entirely included within small area 1.
 Zero mean residual assumption for small area maybe violated.
@@
 
-% explain
-% 
-% # Checking the nesting of the sample-design:
-% # 1) --> each s2-point muss have the complete set of s1-auxvars (s1-info) available
-% # 2) --> each s2-point muss have the complete set of s0-auxvars (s0-info) available
-% # 3) --> each s1-point muss have the complete set of s0-auxvars (s0-info) available
-
-As explained in section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
-The correct nesting thereby concerns the spatial arrangement of the sample phases (figure \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location.
-For the latter, the package \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing.
+As explained in Section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
+The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig.
+\ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location.
+For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing.
 We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data sets.
@@
 If e.g.
-an $s_2$ and /or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
+an $s_2$ and/or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
 
@@
 Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.
-If any information should be missing at the sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.
+If any information should be missing at any sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.
 

### 2017-10-09 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 \label{sec:glob_est_3p}
-
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% 
-% NOTES:
-% - mention that the exhaustive 3phase was originally called "2phase with partially exhaustive info"
-% - clarify terminology "full" and "reduced"
-%
 
@@
 An intuitive interpretation of \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex} is a follows: Likewise in the twophase approach, the estimated means in $\hat{\bar{\pmb{Z}}}_1$ based on the $s_1$ sample are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$.
-The difference between the estimated means $\hat{\bar{\pmb{Z}}}^{(0)}_1$ and their more precise or even exact counterparts $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and $\bar{\pmb{Z}}^{(0)}_0$ based on the larger or infinite $s_0$ sample is then added as a residual term that expresses the lack of model precision due to varying sample sizes behind the estimated auxilary means.
+The difference between the estimated means $\hat{\bar{\pmb{Z}}}^{(0)}_1$ and their more precise or even exact counterparts $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and $\bar{\pmb{Z}}^{(0)}_0$ based on the larger or infinite $s_0$ sample is then added as a residual term that expresses the lack of precision in the point estimate due to varying sample sizes behind the estimated auxilary means.
 
@@
 At 40 further subsamples $s_2$ we have the observations $Y(x)$ from the field inventory.
-Based on this setup, we can now define the reduced and full regression model formulas to be used in the \code{threephase()} function (note that the models are nested).
+Based on this setup, we can now define the reduced and full regression model formulas to be used in the \code{threephase()} function (note that the models are nested):
 
@@
 We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$.
-If we used \code{mean} as the only auxiliary variable at all sample locations (two-phase approach), we would see that the three-phase extention led to a considerable reduction in variance.
+If we used \code{mean} as the only auxiliary variable at all sample locations (two-phase approach), we would see that the three-phase extention here leads to a considerable reduction in variance.
 
@@
 $n_{2,G}=0$.
-Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (equation \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (equation \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
+Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (Eq.
+\ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq.
+\ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
 The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$.
@@
 
-% \begin{subequations}\label{eq:var_3p_synth_psynth}
-% \begin{align}
-% \hat{\var}(\hat{Y}_{G,synth,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-% \bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
-% \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
-% \hat{\var}(\hat{Y}_{G,psynth,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
-% \frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-% \hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
-% \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
-% \end{align}
-% \end{subequations}
-
 %------------------------------------------------------------- %
@@
 This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
-The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{\mathbb{Z}}^{t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathbb{Z}}^{t}(x)=\pmb{Z}^{t}(x)$.
+The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
 This practically means that we have to specify the binary indicator variable for \textit{all} observations in the data frame that is passed to the package-functions.
@@
 We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the \textbf{extended synthetic} and the \textbf{synthetic} estimator to the \code{grisons} data set.
-The setup is thus exactly the same as in the example for global three-phase estimation (section \ref{sec:glob_est_3p}).
+The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}).
 However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}.
-We thus first define the true means for each small area likewise we did in the \code{twophase()} example (section \ref{sec:twophase_sae}):
+We thus first define the true means for each small area likewise we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):
 

### 2017-10-09 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 09.10.2017(spelling errors)

@@
 
-where the g-weight variance-covariance matrix of $\beta_{s_2}$ is calculated as
+where the g-weight variance-covariance matrix of $\hat{\beta}_{s_2}$ is calculated as
 
@@
 
-Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (Eq \ref{eq:estvarmatrix}).
+Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (Eq.
+\ref{eq:estvarmatrix}).
 The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formation is possible, as is the case with non-parametric models like kNN.
@@
 
-% -----------------------
-%% ALEX VERSION:
-
-% spatial means of $p$ functions that respectively determine the function values $z_1(x),...,z_p(x) \in Z(x)$ at all possible locations in $F$.
-
 We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
-Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in the Fig.
+Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in Fig.
 \ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points).
-For a every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
+For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
 We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the point whereas the support is the spatial extent of the auxiliary information used to calculate those values.
 So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
-The \pkg{forestinventory} package calculates the empirical mean of the $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par %% ok!
-
-% The \textbf{exhaustive} case requires a closer look.
-Since being in the infinite population approach, $\pmb{Z}(x)$ again refers to the sample point $x$ and not the area around it.
-That implies that deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx$, involves computing the spatial means of all component, i.e.
-$z_1(x),...,z_p(x)$, by calculating those at all possible locations in $F$.
-This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
-$n_1=\infty$).
-Of course this is practically infeasible, but there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
-The first case is when the explanatory variables are provided by polygon layers (e.g.
-map of development stages).
-In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
-The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
-In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+The \pkg{forestinventory} package calculates the empirical mean of $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par
 
@@
 
-One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
+One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
+\ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
 In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
 Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
-deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$.
+deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequently calculating the exact mean $\bar{\pmb{Z}}$.
 While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
@@
 	\end{subfigure}
-\caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level.
+\caption{Concept of \textbf{(a)} exhaustive and \textbf{(b)} non-exhaustive calculation of explanatory variables including boundary adjustment at the support level.
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
@@
 \ref{fig:exh_nexh_bw}).
-This method was suggested in \citet{mandallaz2013b} and led to an improvement estimation precision.
+This method was suggested in \citet{mandallaz2013b} and led to an improvement in estimation precision.
 In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the \textit{weighted} mean where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.
@@
 Note that the vector input to \code{exhaustive} must be in the same order that the \code{lm()}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1.
-This step is particularly recommended if categorical variables are present because the \code{lm()}-function, which is internally used to set up the design-matrix, automatically creates dummy variables with one of the categories used as a reference.
+Particular caution must be taken if categorical variables are present because the \code{lm()}-function, which is internally used to set up the design-matrix, automatically creates dummy variables with one of the categories used as a reference.
 The correct order for our \code{grisons} example can easily be extracted using:
@@
 
-The \pkg{forestinventory} package provides three broad types of small area estimators each of which has an exhaustive and non-exhaustive form.
+The \pkg{forestinventory} package provides three types of small area estimators each of which has an exhaustive and non-exhaustive form.
 We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label \textit{pseudo} to indicate that the mean of the explanatory variables within the small area was based on a finite sample.
@@
 
-In the equations for the point estimates (\ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
+In the equations for the point estimates (Eq.
+\ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
 A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$.
@@
 
-The package provides the \textit{g-weight} variance for \textit{small} and \textit{psmall} respectively (Equations \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
+The package provides the \textit{g-weight} variance for \textit{small} and \textit{psmall} respectively (Eq.
+\ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Eq.
+\ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
 Again note that all components are restricted to those available at the sample locations in the small area ($s_{1,G}$ and $s_{2,G}$), with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.
@@
 $n_{2,G}=0$).
-In this case, the point estimates (Equations \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
-However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Equations \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
+In this case, the point estimates (Eq.
+\ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
+However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Eq.
+\ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
 Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
@@
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
-Recalling that linear models fit using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
-The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Equations \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
+Recalling that linear models fitted using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
+The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq.
+\ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
 Note that the package functions internally extend the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
@@
 Small area estimates in the \pkg{forestinventory} package can be applied by specifying the optional argument \code{small_area}.
-The input data set has to include an additional column of class factor that describes the small area membership of the sample location represented by that row.
+The input data set has to include an additional column of class \code{factor} that describes the small area membership of the sample location represented by that row.
 The argument \code{small_area} requires a \code{list}-object that comprises
@@
 
-The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need to implement it yourself.
+The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need for the user to implement it.
 Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator vary between small areas, while they are identical under the \textit{psmall} estimator.
@@
 
-The \textbf{exhaustive versions} of the small area estimators (equations \ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
+The \textbf{exhaustive versions} of the small area estimators (Eq.
+\ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest.

### 2017-10-10 JStatSoft_forestinventory_master-concordance.tex
update Andreas 10.10.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
-1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1210:%
+1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1211:%
 137 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1215:%
-1 42 1 1 2 11 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1280:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1216:%
+1 35 1 1 2 11 0 1 2 3 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1268:%
 143 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1285:%
-1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 12 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 40 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 19 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1557:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1273:%
+1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1522:%
 149 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1562:%
-1 11 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1784:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1527:%
+1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 18 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1724:%
 155 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1789:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1729:%
+1 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1741:%
+161 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1746:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1794:%
-161 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
+167 11 1}

### 2017-10-10 JStatSoft_forestinventory_master.rnw
update Andreas 10.10.2017(1)

@@
 
-  Alexander Massey\\
+  \noindent Alexander Massey\\
   Department of Environmental Systems Science\\
@@
 %------------------------------------------------------------------------------------------------%
+% ---------------------------------------- Future Plans---- ------------------------------------ %
+
+\SweaveInput{rnw/Future Plans.Rnw}
+
+%------------------------------------------------------------------------------------------------%
 % ---------------------------------------- Acknowledgements ------------------------------------ %

### 2017-10-10 rnw/Analysis and Visualization.Rnw
update Andreas 10.10.2017(1)

@@
 
-The package \pkg{forestinventory} also provides methods for \textit{analyzing} and \textit{visualizing} the estimation results created by the \code{onephase()}, \code{twophase()} and \code{threephase()} functions.
-
 \subsection{Analysis}
 
-If we want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, the function to start with is \code{estTable()}.
-This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis.
-For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} data set:
+We often want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, which can be easily accomplished in \pkg{forestinventory} using the \code{estTable()} function.
+This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis.
+For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimators for the \code{grisons} data set:
 
@@
 The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}.
+The \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.
 If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}.
-Note however that the package \pkg{forestinventory} provides several \code{S3} generic methods that particularly respond to the class \code{"esttable"}.
-Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.
+Note however that \pkg{forestinventory} provides several \code{S3} generic methods specifically for the class \code{"esttable"}.
 
 The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
-However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied.
-Per default, also the confidence intervals are added.
+However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance that was applied (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance).
+By default, the confidence intervals are also added.
 
@@
 
-As mentioned in the introduction (Section \ref{sec:intro}), the application of multi-phase estimation techniques in forest practice have two main \textbf{goals}:  
-\begin{enumerate}
-  \item \textit{\textbf{maintaining}} the one-phase estimation precision under reduced costs, i.e. reduced $s_2$ sample size
-  \item \textit{\textbf{increasing}} the estimation precision at identical costs, i.e. without increasing the $s_2$ sample size
-\end{enumerate}
-
-The function \code{mphase.gain()} was particularly written to quantify the potential benefit of a multi-phase global or small area estimation compared to its respective one-phase estimation.
-The function takes an \code{esttable} object as input and returns a validation of which multi-phase method and estimator performed best in comparison to the onephase estimation (\textit{baseline}) in terms of estimation precision.
+As multi-phase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multi-phase global or small area estimate compared to its respective one-phase estimate.
+The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a \textit{baseline}).
 If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
 The argument \code{pref.vartype} can be used to define what type of variance (\textit{g-weight} or \textit{external}) should be used for the comparison.
-Per default (\code{exclude.synth = TRUE}), synthetic estimations (\textit{synth} and \textit{psynth} estimator) are not considered for the comparison since they normally have a much smaller variance but to the cost of a potential bias.
+Synthetic estimates (\textit{synth} and \textit{psynth} estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.
 
@@
 The multi-phase estimation procedure is again specified in the \code{method} and \code{estimator} column.
-The last two columns then actually quantify the potential benefit of the multi-phase estimation.
-The \code{gain} is the reduction (if its value is positive) or possibly also the increase (if its value is negative) in variance when applying the multi-phase as alternative to the one-phase estimation.
-For small area \code{"B"}, it can be read as "the \textit{twophase extended psynth} estimation procedure leads to a 67.9 \% reduction in variance compared to the one-phase procedure".
+The last two columns quantify the potential benefit of the multi-phase estimation.
+The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation.For example, it is indicated that the \textit{twophase extended psynth} estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
 The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:
@@
 For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{twophase extended psynth} estimator.
-If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple mean of quantifying the financial benefit of multi-phase forest inventories.
+If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multi-phase estimation for forest inventories.
 
@@
 
-The package \pkg{forestinventory} also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
-Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function.
+% The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
+Like \code{mphase.gain()}, also the plot method also requires an \code{esttable} object created by the \code{estTable()}-function.
 The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.
+Both plots can be obtained by passing the \code{esstable} object to the \code{S3} generic \code{plot} method.
+Whereas the estimation errors are plotted by default, the graphic of the point estimates and confidence intervals is returned when setting the argument \code{yvar = "estimate"}.
+Note that the graphics can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.
 
-If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case using the default setting of \code{estTable()}:
+The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.
+Both plots can be obtained by passing the \code{esstable} object to the \code{plot()} function.
 
 \begin{small}
-<<>>=
-grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
-                                             extpsynth_3p, psynth_3p),
-                             sae = TRUE, 
-                             vartypes = c("variance", "g_variance",  "ext_variance"))
+<<eval=FALSE>>=
+plot(grisons.sae.table, ncol = 2)
 @
@@
 
-The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic \code{plot} method and setting the argument \code{yvar = "estimate"}:
+\begin{figure}[h]
+\centering
+\resizebox{0.85\hsize}{!}{\includegraphics{fig/plot_error.png}}
+\end{figure}
+
+\newpage
+
+Whereas the estimation errors are plotted by default, the point estimates and confidence intervals are returned when setting the argument \code{yvar = "estimate"}.
+Note that the graphics can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[h]
 \centering
-\resizebox{0.9\hsize}{!}{\includegraphics{fig/plot_est_ci.png}}
+\resizebox{0.85\hsize}{!}{\includegraphics{fig/plot_est_ci.png}}
 \end{figure}
-
-If we want the estimation errors to be plotted, we simply set the \code{yvar} argument \code{"error"} (default):
-
-\begin{small}
-<<eval=FALSE>>=
-plot(grisons.sae.table, ncol = 2)
-@
-\end{small}
-
-\begin{figure}[htb]
-\centering
-\resizebox{0.86\hsize}{!}{\includegraphics{fig/plot_error.png}}
-\end{figure}
-
-Note that the \code{plot()}-function call can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.
-
-\newpage

### 2017-10-10 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 10.10.2017(1)

@@
 
-% NOTE:
-% - Provide all confidence equations for the estimators: onephase, twophase, threephase
-% - !!! Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
-% - ... and asymptotics of estimators
-% - Interpretation of Confidence Intervals
-%
-
-Converting the estimated variance into a 95\% \textit{confidence interval} (CI) allows for an intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate.
+Converting the estimated variance into a 95\% \textit{confidence interval} (CI) allows for a more practical interpretation of a point estimate's precision.
 The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value.
 In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sample design.
-Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\%  CIs from each sample, then 95\% of them are expected to contain the true population parameter.
+Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples and calculate the 95\%  CI from each sample, then 95\% of the CIs are expected to contain the true population parameter.
 The confidence level $1-\alpha$ (e.g.
-95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown population parameter and is therefore often also referred to as \textit{coverage rate}.
-For a user, the CI is a measure of precision that can also be used for hypothesis testing, i.e.
-providing evidence that an estimate is statistically different from a given value if that value lies outside the CI boundaries.
+95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as \textit{coverage rate}.
+The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries.
 
-Based on the central limit theorem, it can be assumed that under hypothetical repeated sampling, the point estimates will asymptotically follow a normal distribution.
-\citet{mandallaz2013a} recommends to better base the confidence intervals on the Student's \textit{t} distribution.
-The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as
+Based on the central limit theorem it can be assumed that under hypothetical repeated sampling the point estimates will asymptotically follow a normal distribution.
+However, on the recommendation of \citet{mandallaz2013a}, better confidence intervals can obtained using the Student's \textit{t} distribution (\citet{mandallaz2013a}) defined as
 
@@
 
-where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level.
-Note that for three-phase sampling, $p$ constitutes the number of parameters used in the \textit{full} regression model.
+where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
-In the package \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) are calculated by:
+In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
+For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (Section \ref{sec:twophase_sae}) are calculated by:
 
@@
 
-% Frequentisitsche Interpretation streng genommen die einzig richtige: unter rein hypothetischer Wiederholung der Inventur (Gleicher Schaetzer, gleicher Wald, gleiche Hilfvariablen / Aux.info) werden 95\% der berechneten Intervalle den wahren Wert enthalten.
-Man weiss nun aber - selbst wenn man das machen wuerde (was praktisch ja nie der Fall sein wird) - nie, auf welche der berechneten Intervalle das zutreffen wird.
-Die klassische statistische Interpration ist hier also nicht zulaessig (ungleich: der wahre wert befindet sich mit 95\% Sicherheit in meinem berechneten Intervall).
-% 
-% Da diese strenge Interpretation dem Praktiker nicht viel nuetzen wird (ausser evtl einem Gefuhl der Sicherheit), kann man sich aufgrund der CI-Breite (abhaengig von der Varianz der SChaetzung) an folgender Frage orientieren: "Wuerde ich innerhalb des Vertrauensintervalles zu der selben Entscheidung (z.b.
-Eingriff ja/nein, Vorrat um x\% reduzieren) kommen, oder nicht?" Wenn ja, dann ist die Genauigkeit der SChaetzung bereits ausreichend, wenn nein, dann ist dies die Motivation nach effizien
-% Stichwort: signifikanz vs. relevanz
-
 \newpage

### 2017-10-10 rnw/Future Plans.Rnw
update Andreas 10.10.2017(1)

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Future Plans}
\label{sec:future}

The \pkg{forestinventory} package currently provides a fairly well-rounded toolkit for forestry inventorists to integrate auxiliary information into their estimates using the model-assisted methods under the design-based approach.  Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future. As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub. Currently planned extensions include:

\begin{itemize}
\item Implement parallel procedures for efficiently calculating many small areas.
\item Allow functions to accept objects of class \code{data.table} from the \pkg{data.table} package to improve memory efficiency.
\item Enable the user to choose other types of models than linear regressions fit with OLS.
\end{itemize}

### 2017-10-10 rnw/Special Cases and Scenarios.Rnw
update Andreas 10.10.2017(1)

@@
 
-% include:
-% - categorical variables (zberg) and interaction terms (not so clear to everyone)
-% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
-% - !!! cluster sampling and differences between (p)small and extended (p)small if cluster is not completely inside small area
-% - !!! violation of nested sample design + error handling / data adjustments / printed warnings
-% - !!! missing factor levels in s2-sample that occur in s1 / s0 sample + error warning
-%
+% --------------------------------------------------------- %
+\subsection{Post-Stratification}
 
-% --------------------------------------------------------- %
-\subsection{Poststratification}
-
-A special case of multi-phase regression estimators is \textit{poststratification}, which can further be divided into the case of \textit{multi-phase sampling for stratification} and \textit{multi-phase sampling for regression within strata}.
+A special case of multi-phase regression estimators is \textit{post-stratification}, which can further be divided into the cases of \textit{multi-phase sampling for stratification} and \textit{multi-phase sampling for regression within strata}.
 Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.
 
-To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
+To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
 
@@
 
-\textit{Two-phase sampling for stratification} is now applied if we only use categorical variables in the model, in this case the factor variable \code{stage} (i.e.
-we have an ANOVA model).
-It basically means that the model predictions are the means of terrestrial response values within each development stage (within-strata means).
+\textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
+Linear regression models only fit with categorical variables produce ANOVA models, which when used in multi-phase regression estimators is equivalent to post-stratification.
+For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).
 
@@
 If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
-Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations.
+Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables, which makes then very flexible.
 
@@
 
-The overall variance of estimators under poststratification will mainly depend on the between and within-strata variation of the model residuals and can therefore be reduced by minimizing the within strata residual square sum.
-In case of poststratification, the g-weight variance estimator has the advantage that the strata weights are estimated from the large sample, whereas the external variance calculates them solely based on the terrestrial sample $s_2$.
+The variance of all model-assisted estimators included in \pkg{forestinventory} can be decreased by reducing the sum of square residuals of the regression model.
+In case of post-stratification, this particularly implies minimizing the within strata residual square sum.
+Also, for post-stratification, the g-weight variance should be trusted over the external variance because it has the advantage that the strata weights are estimated from the large sample rather than the terrestrial sample $s_2$.
 
@@
 
-As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample unit does no longer consist of one, but multiple spatially agglomerated sample points.
-One still randomly places the sample location $x$ in the inventory area likewise in the simple sampling design, but then creates $M-1$ additional sample locations $x_2,...,x_M$ close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
+As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points.
+One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
 The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign.
-Note however that the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and that this average value is again referred to exactly one point, i.e.
-the cluster origin $x$, in the frame of the infinite population approach.
+However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e.
+the cluster origin $x$.
 Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights.
@@
 defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
-The package functions of \pkg{forstinventory} also allow to consider cluster-structures by specifying a unique cluster identification (cluster ID) for each cluster in an additional column whose column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.
+The \pkg{forstinventory} package identifies clusters via a unique cluster ID that is assigned to a column in the input data set.
+Its column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.
 
@@
 only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
-In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see Section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
-In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured under this scenario.
+In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, it is adviseable to use the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still ensured.
 
-In oder to keep track of such cases, the package \pkg{forestinventory} tells the user to do so by returning a warning message:
+In order to keep track of such cases, \pkg{forestinventory} tells the user to do so by returning a warning message:
 <<echo=FALSE>>=
@@
 Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the g-weight variances only differ marginally.
-This can be taken as evidence that the violation of the zero mean residual property can here expected to be negligible.
+This can be taken as evidence that the violation of the zero mean residual property can here be expected to have negligible consequences.
 
@@
 
-As explained in Section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
+As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multi-phase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
 The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig.
 \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location.
-For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing.
-We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data sets.
+For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first-aid adjustments to the inventory data set to prevent the calculations from failing.
+We will demonstrate possible nesting violations by applying the global three-phase estimator to the \code{grisons} and \code{zberg} data sets.
 
@@
 
-Based on the nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e.
-terrestrially surveyed) and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model.
+Based on the nesting rule, $s_2 \in s_1 \in s_0$, each $s_2$ and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model.
 If e.g.
-an $s_2$ and/or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
+an $s_2$ and/or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (in this case \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
 
@@
 
-\subsubsection*{Violation 2} % 2) and 3)
+\subsubsection*{Violation 2}
 
-However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (here \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
+However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
 If this concerns an $s_2$ sample location, the associated response variable can no longer be used.
@@
 
-% # Violation:
-% # each s2-point muss have the complete set of s1-auxvars (s1-info) available
-% 
-% # error message:
-% # missing expl.variable in s1-aux.set at 1 s2-plot.
-% 
-% # error handling:
-% # since there is still the complete set of s0 available, the phase-id has been changed
-% # to s0 (by this, we could keep the plot); but we thereby can no longer use the available
-% # terrestrial info from this plot --> could hamper sae estimation)
-% 
-% # NOTE: This could violate the "randomness" assumption.
-% #       Be sure that we you have a "completely missing at random" case!!!
-
 \subsubsection*{Violation 3}
 
-If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations is \textbf{deleted} from the data set.
+If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations are \textbf{deleted} from the data set.
 
@@
 
-Note that that all the automatic data adjustments (deletion, recoding) have to be accepted with caution.
-Recapitulating, the unbiasedness of estimators in the design-based framework is based on the randomization of the sample locations.
-This means that every possible location within the forest area $F$ has an inclusion probability greater zero.
+Note that all the automatic data adjustments (deletion, recoding) have to be accepted with caution.
+Recapitulating, the unbiasedness of estimators in the design-based framework is based on the uniform and independent randomization of the sample locations.
+This means that every possible location within the forest area $F$, as well as pair of locations, have inclusion and joint inclusion probabilities greater than zero.
 Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.
-If any information should be missing at any sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.
+If any information should be missing at the sample locations, one should clarify the reason for this and make sure that the information can reasonably be assumed to be \textit{completely missing at random}.
 
@@
 
-If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category level is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample.
-Conclusively, an internal regression model can not be calculated and the function \textbf{stops} with the following error message:
+If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample.
+Conclusively, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:
 

### 2017-10-10 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 10.10.2017(1)

@@
 the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
-In addition, we also have to specify the indication of the $s_1$ phase (\code{s1.id}) in the argument \code{phase_id}.
+In addition, we also have to specify the indication of the $s_1$ phase (\code{s1.id}) in the argument \code{phase_id} (note that \pkg{forestinventory} implicitly assumes that all other rows in the input data set belong to $s_0$).
 The global three-phase estimation can thus be applied by
@@
 We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$.
-If we used \code{mean} as the only auxiliary variable at all sample locations (two-phase approach), we would see that the three-phase extention here leads to a considerable reduction in variance.
+Compared to using only \code{mean} under a two-phase approach, we would see a considerable reduction in variance by the three-phase extension.
 
@@
 The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
+In other words, when extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e.
+$s_0$).
+The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq.
+\ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}.
+Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficents resulting from the random sample $s_2$.
+
+% The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.
+$\int_F R(x)dx=\int_G R(x)dx=0$.
+This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
+The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
 This practically means that we have to specify the binary indicator variable for \textit{all} observations in the data frame that is passed to the package-functions.
 The regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ under the extended auxiliary vector are thus recalculated as
-
-\begin{subequations}\label{eq:normequ_redmod_ext}
-\begin{align}
-\hat{\pmb{\gamma}}_{s_2} &= \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)
-\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}^{(0)}(x) \\
-\hat{\pmb{\theta}}_{{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)
-\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x) \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1} \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-\end{align}
-\end{subequations}
-
-With the auxliary means and their variance-covariance matrices following the same notational change, the \textbf{point estimate} and \textbf{g-weight variance} are derived as
-
-\begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
-\begin{align}
-\hat{Y}_{G,extsynth,3p}&=(\bar{\pmb{\mathbb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
-\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2} \label{eq:pe_3p_extsynth} \\
-\hat{Y}_{G,extpsynth,3p}&=(\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
-\hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth} \\
-\hat{\var}(\hat{Y}_{G,extsynth,3p})& =\frac{n_2}{n_1}\bar{\pmb{\mathbb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
-\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extsynth}\\
-\hat{\var}(\hat{Y}_{G,extpsynth,3p})& = \hat{\pmb{\gamma}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}\hat{\pmb{\gamma}}_{s_2} + \frac{n_2}{n_1}\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
-\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extpsynth}
-\end{align}
-\end{subequations}
-
-The package also returns the \textbf{external variance} which is defined as
-
-\begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
-\begin{flalign}
-\hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) + 
- (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extsynth}&\\
-\hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) \nonumber & \\
- &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extpsynth}
-\end{flalign}
-\end{subequations}
-where $\hat{\mathbb{R}}_{0}=Y(x)-\hat{\mathbb{Y}}_{0}(x)$ with $\hat{\mathbb{Y}}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$.
+% 
+% \begin{subequations}\label{eq:normequ_redmod_ext}
+% \begin{align}
+% \hat{\pmb{\gamma}}_{s_2} &= \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)
+% \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}^{(0)}(x) \\
+% \hat{\pmb{\theta}}_{{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)
+% \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x) \\
+%   \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
+%   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)\Big)
+%   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1} \\
+%   \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+%   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)\Big)
+%   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
+% \end{align}
+% \end{subequations}
+% 
+% With the auxliary means and their variance-covariance matrices following the same notational change, the \textbf{point estimate} and \textbf{g-weight variance} are derived as
+% 
+% \begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
+% \begin{align}
+% \hat{Y}_{G,extsynth,3p}&=(\bar{\pmb{\mathbb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
+% \hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2} \label{eq:pe_3p_extsynth} \\
+% \hat{Y}_{G,extpsynth,3p}&=(\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
+% \hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth} \\
+% \hat{\var}(\hat{Y}_{G,extsynth,3p})& =\frac{n_2}{n_1}\bar{\pmb{\mathbb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
+% \bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+% \hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extsynth}\\
+% \hat{\var}(\hat{Y}_{G,extpsynth,3p})& = \hat{\pmb{\gamma}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}\hat{\pmb{\gamma}}_{s_2} + \frac{n_2}{n_1}\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
+% \hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+% \hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extpsynth}
+% \end{align}
+% \end{subequations}
+% 
+% 
+% The package also returns the \textbf{external variance} which is defined as
+% 
+% \begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
+% \begin{flalign}
+% \hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) +
+%  (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extsynth}&\\
+% \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) \nonumber & \\
+%  &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extpsynth}
+% \end{flalign}
+% \end{subequations}
+% where $\hat{\mathbb{R}}_{0}=Y(x)-\hat{\mathbb{Y}}_{0}(x)$ with $\hat{\mathbb{Y}}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$.
 The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in Eq.

### 2017-10-10 r/create figures.R

@@ -48,7 +48,7 @@ plot(grisons.sae.table, ncol = 2, yvar = "estimate")
 
 ## save graphic for article:
 ggsave("plot_est_ci.png", 
-       plot(grisons.sae.table.CI, ncol = 2, yvar = "estimate") +
+       plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
          ylab("Timber Volume [m3/ha]") +
          theme(axis.title.x =element_text(size=rel(6)), # size of axis-labels text
                axis.title.y =element_text(size=rel(6)),
@@ -57,7 +57,7 @@ ggsave("plot_est_ci.png",
                axis.text.y = element_text(size=rel(6)),
                legend.title=element_text(size=rel(7)),
                legend.text=element_text(size=rel(6))),
-       width = 27, height = 17, dpi = 300)
+       width = 27, height = 14, dpi = 300)
 
 
 
@@ -73,5 +73,4 @@ ggsave("plot_error.png",
                axis.text.y = element_text(size=rel(6)),
                legend.title=element_text(size=rel(7)),
                legend.text=element_text(size=rel(6))),
-       width = 27, height = 17, dpi = 300)
-
+       width = 27, height = 14, dpi = 300)

### 2017-10-10 JStatSoft_forestinventory_master-concordance.tex
update Andreas 10.10.2017(2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 107 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 115:%
+1 108 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
 1 97 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
-123 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 218:%
-1 69 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 214:%
+124 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 219:%
+1 68 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
-129 4 1}
+130 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 803:%
-135}
+136}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1211:%
-137 4 1}
+138 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1216:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1268:%
-143 4 1}
+144 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1273:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1522:%
-149 4 1}
+150 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1527:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1724:%
-155 4 1}
+156 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1729:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1741:%
-161 4 1}
+162 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1746:%
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
-167 11 1}
+168 11 1}

### 2017-10-10 rnw/Main structure and modules of the Package.Rnw
update Andreas 10.10.2017(2)

@@
 
-%\section{Methods and Structure of the Package}
 \section{Methods and Structure of the Package}
@@
 The spatial distribution of a hypothetical density function for every point in a forested area is represented as a wavy piecewise constant green surface.
-Sample plots (white dots) inform the inventorist of the value of the density function at that point.}
+Sample plots (white dots) inform the inventorist of the value of the density function at that point.
+Note that the plateaus of constant $Y(x)$ values here have the shape of squares whereas in reality they are likely to be formed by the intersection of circles around trees.}
 \label{fig:inf_pop_apr}
@@
 The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and to apply this model to the small area.
-The potential bias of applying that model in $G$ is then corrected for by using the model residuals derived from that small area.
+The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area.
 If there are no terrestrial plots in $G$ (i.e.

### 2017-10-10 rnw/Special Cases and Scenarios.Rnw
update Andreas 10.10.2017(2)

@@
 only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
-In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
 In this case, it is adviseable to use the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still ensured.

### 2017-10-10 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 10.10.2017(2)

@@
 
-Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the residuals of the reduced model.
+Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the empirical residuals of the reduced model.
 The \textbf{external variance} form under linear regression models is defined as
@@
 The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
-${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate.
+$\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate.
 The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (Eq.
@@
 \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
-The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$.
+The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be insured within small area $G$.
 Also the variance will be small but to the cost of ignoring the model uncertainties.
@@
 
-The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.
-$\int_F R(x)dx=\int_G R(x)dx=0$.
+The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.$\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
 This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
 The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
-In other words, when extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e.
+In other words, when the \textit{extended} option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e.
 $s_0$).

### 2017-10-10 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 10.10.2017(2)

@@
 
-Note that for \textit{internal} linear models the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction which is why it does not appear in the point estimate.
+Note that for \textit{internal} linear models the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (\textit{zero mean residual property}) which is why it does not appear in the point estimate.
 More explanation about how to obain the auxiliary means is given in the next subsection.
@@
 \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
-A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$.
+A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$.
 This is called the \textit{bias} or \textit{residual correction} term.
@@
 \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
-However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Eq.
+However, the bias correction using the observed residuals $\hat{R}(x)$ is not applied as was the case in the small and pseudo small area estimator (Eq.
 \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
-Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e.
-$\int_G R(x)dx=0$.
+Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
 However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
@@
 $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
-The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e.
-$\int_F R(x)dx=\int_G R(x)dx=0$.
+The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
-Recalling that linear models fitted using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
+Recalling that linear models fitted using OLS have zero mean residual property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
 The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq.

### 2017-10-11 JStatSoft_forestinventory_master-concordance.tex
update Andreas 11.10.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 219:%
-1 68 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
+1 69 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 289:%
 130 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 294:%
 1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 803:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 804:%
 136}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 805:%
 1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1211:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1212:%
 138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1216:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1217:%
 1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1268:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1269:%
 144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1273:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1274:%
 1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1522:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1523:%
 150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1527:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1528:%
 1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 18 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1724:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1725:%
 156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1729:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1730:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1741:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1742:%
 162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1746:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1747:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1752:%
 168 11 1}

### 2017-10-11 JStatSoft_forestinventory_master-concordance.tex
update Andreas 11.10.2017(3)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
-1 97 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 214:%
+1 98 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 215:%
 124 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 219:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
 1 69 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 289:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 290:%
 130 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 294:%
-1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 804:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 295:%
+1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 806:%
 136}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 805:%
-1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1212:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 807:%
+1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1227:%
 138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1217:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1232:%
 1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1269:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1284:%
 144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1274:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1289:%
 1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1523:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1538:%
 150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1528:%
-1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 18 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1725:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1543:%
+1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1738:%
 156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1730:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1743:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1742:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1755:%
 162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1747:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1760:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1752:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
 168 11 1}

### 2017-10-11 JStatSoft_forestinventory_master.rnw
update Andreas 10.10.2017(2)
update Andreas 11.10.2017(3)

@@
 \documentclass[article]{jss}
+% \documentclass[nojss]{jss} % for creating the vignette
 
@@
 The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
-This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in \proglang{R}.}
+This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}
 

### 2017-10-11 rnw/Acknowledgements.Rnw
update Andreas 11.10.2017(3)

@@
 Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study and providing the possibility of working on the package.
-We also want to thank Daniel Mandallaz for his support in completing the range of the allready published estimators in the frame of the three-phase small area estimators, many helpful discussions and advice throughout the implementation of our package.
+We also want to thank Daniel Mandallaz for his support in completing the range of the already published estimators in the frame of the three-phase small area estimators, as well as many helpful discussions and advice throughout the implementation of our package.
 Our thanks also go to Meinrad Abegg for proofreading the manuscript, and to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of grisons for providing the example data.

### 2017-10-11 rnw/Analysis and Visualization.Rnw
update Andreas 11.10.2017(3)

@@
 As multi-phase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multi-phase global or small area estimate compared to its respective one-phase estimate.
-The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a \textit{baseline}).
+The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a \textit{baseline}.
 If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
@@
 The last two columns quantify the potential benefit of the multi-phase estimation.
-The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation.For example, it is indicated that the \textit{twophase extended psynth} estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
+The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation.
+For example, it is indicated that the \textit{two-phase extended psynth} estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
 The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:
@@
 The \textit{relative efficiency} can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
-For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{twophase extended psynth} estimator.
+For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{two-phase extended psynth} estimator.
 If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multi-phase estimation for forest inventories.
@@
 
-% The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results.
-Like \code{mphase.gain()}, also the plot method also requires an \code{esttable} object created by the \code{estTable()}-function.
-The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.
-Both plots can be obtained by passing the \code{esstable} object to the \code{S3} generic \code{plot} method.
-Whereas the estimation errors are plotted by default, the graphic of the point estimates and confidence intervals is returned when setting the argument \code{yvar = "estimate"}.
-Note that the graphics can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.
-
-The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.
+The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: 1) the \textit{\textbf{point estimates}} with overlayed confidence intervals, and 2) the \textit{\textbf{estimation errors}}.
 Both plots can be obtained by passing the \code{esstable} object to the \code{plot()} function.

### 2017-10-11 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 11.10.2017(3)

@@
 
-where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
+where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).

### 2017-10-11 rnw/Introduction.Rnw
update Andreas 10.10.2017(2)
update Andreas 11.10.2017(3)

@@
 In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time.
-They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on the national and international levels.
+They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on national and international levels.
 As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods.
@@
 This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals.
-There is a broad range of literature describing the concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
+There is a broad range of literature describing the concepts and methods regarding the choice of different estimators under various sample designs \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.
 
@@
 However, they have the downside of being very expensive.
-Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more random plot locations.
+Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more plot locations.
 This is why the number of terrestrial samples is often limited.
 Although national inventories usually possess a sufficiently large terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.
-As a result, there has been increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is called \textit{multiphase sampling}.
+As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
+A method which has become particularly attractive is called \textit{multi-phase sampling}.
 The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
 This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
-These predictions are produced by regression models that use explanatory variables derived from auxiliary data, commonly in the form of remote sensing data available in the inventory area in large sample or even wall-to-wall.
+These predictions are produced by regression models that use explanatory variables derived from auxiliary
+data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
 Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
@@
 The objectives of this article are to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} address special cases and demonstrate how the package-functions handle them.
+
+\newpage
 

### 2017-10-11 rnw/Main structure and modules of the Package.Rnw
update Andreas 11.10.2017(1)
update Andreas 11.10.2017(3)

@@
 Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e.
-$Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{s_2}Y(x)$.
+$Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
 The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
@@
 For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some model, which in \pkg{forestinventory} is always a linear model fit in $s_2$ using ordinary least squares (OLS).
-The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model biases, i.e.
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e.
 $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.
@@
 The basic setup is that the explanatory variables calculated from the auxiliary information are available in two different frequencies.
-The phase $s_0$ provides the largest number $n_0$ of auxiliary sample locations, whereas the phase $s_1$ provides fewer auxiliary locations that are only available at $n_1$ subsamples of $s_0$.
+The phase $s_0$ provides a large number $n_0$ of auxiliary data, whereas the phase $s_1$ provides additional auxiliary data that are only available at $n_1$ subsamples of $s_0$.
 The terrestrial information is then collected at a further subsample $s_2$ of $s_1$.
@@
 From that it follows that we can define \textit{two nested} regression models.
-The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
+The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}^{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
 The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
-Less accurate predictions, $\hat{Y}_{0}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
+Less accurate predictions, $\hat{Y}^{0}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
 If there is a significant gain in model precision between the reduced and the full model and the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
@@
 		\centering
-   \resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
+	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
+   %\resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_b}
@@
 However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals.
-The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and to apply this model to the small area.
+The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area.
 The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area.
@@
 The subject of model selection gets a lot of attention in the field of forest inventory.
-This is why it is important to understand that the mathematical interpretation of how a model is used to produce estimates is fundamentally different between the design-based and model-dependent approaches.
+This is why it is important to understand that the mathematical interpretation of how a model is used to produce estimates is fundamentally different between the design-based and model-dependent approach.
 In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
@@
 \centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
+\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_4.PNG}}
 \caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}

### 2017-10-11 rnw/Special Cases and Scenarios.Rnw
update Andreas 11.10.2017(3)

@@
 \textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
-Linear regression models only fit with categorical variables produce ANOVA models, which when used in multi-phase regression estimators is equivalent to post-stratification.
+Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators is equivalent to post-stratification.
 For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).
@@
 If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
-Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables, which makes then very flexible.
+Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.
 
@@
 Recapitulating, the unbiasedness of estimators in the design-based framework is based on the uniform and independent randomization of the sample locations.
-This means that every possible location within the forest area $F$, as well as pair of locations, have inclusion and joint inclusion probabilities greater than zero.
+This means that every possible location within the forest area $F$, as well as pairs of locations, have inclusion and joint inclusion probabilities greater than zero.
 Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.

### 2017-10-11 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 11.10.2017(3)

@@
 Fitting the model using $s_2$ (i.e.
-\textit{internally}) ensures the zero mean residual property over $F$.
+\textit{internally}) ensures the zero mean residual property over $s_2$.
+
+% \begin{subequations}\label{reg3p}
+% \begin{align}
+% \hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
+% \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
+% \hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
+% \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
+% \end{align}
+% \end{subequations}
+% 
+% In the non-exhaustive case, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
+\ref{meanvalues3pnex}.
+If the exact mean over $s_0$ is known, the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ can simply be replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
+Note that in case of applied boundary adjustment (Section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to Eq.
+\ref{eq:wmeanZ}.
 
@@
 \begin{align}
-\hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
+\hat{Y}_{reg3p,ex}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2})
+\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
 \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
-\hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
+\hat{Y}_{reg3p,nex}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2})
+\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
 \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
@@
 
-In the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
+Intuitively, the three phase estimator is simply the mean of the weaker predictions using the reduced model, corrected by the mean difference between the reduced model and the stronger full model, corrected by the mean difference between the ground truth and the full model predictions.
+For the compact version of the formula in the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
 \ref{meanvalues3pnex}.
@@
 
-An intuitive interpretation of \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex} is a follows: Likewise in the twophase approach, the estimated means in $\hat{\bar{\pmb{Z}}}_1$ based on the $s_1$ sample are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$.
+% An intuitive interpretation of \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex} is a follows: Likewise in the two-phase approach, the estimated means in $\hat{\bar{\pmb{Z}}}_1$ based on the $s_1$ sample are transformed into predictions $\hat{Y}_{1}(x)$ of the local density $Y(x)$.
 The difference between the estimated means $\hat{\bar{\pmb{Z}}}^{(0)}_1$ and their more precise or even exact counterparts $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and $\bar{\pmb{Z}}^{(0)}_0$ based on the larger or infinite $s_0$ sample is then added as a residual term that expresses the lack of precision in the point estimate due to varying sample sizes behind the estimated auxilary means.
@@
 
-Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the empirical residuals of the reduced model.
+Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model.
 The \textbf{external variance} form under linear regression models is defined as
@@
 We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$.
-Compared to using only \code{mean} under a two-phase approach, we would see a considerable reduction in variance by the three-phase extension.
+When comparing to using only \code{mean} under a two-phase approach, we would see a considerable reduction in variance by the three-phase extension.
 
@@
 The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
-$\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate.
-The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (Eq.
+$\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
+The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a correction term likewise in the global estimation (Eq.
 \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
@@
 where $\hat{R}_{0}(x)=Y(x)-\hat{Y}_{0}(x)$ with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$.
-Likewise the g-weight variance, the difference between the exhaustive and non-exhaustive form is the variance of the predictions $\hat{Y}_{0}(x)$ within the $s_{0,G}$ phase.
 
@@
 \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas.
-The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be insured within small area $G$.
+The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be ensured within small area $G$.
 Also the variance will be small but to the cost of ignoring the model uncertainties.
@@
 \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}.
-Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficents resulting from the random sample $s_2$.
+Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficients resulting from the random sample $s_2$.
 
@@
 The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively.
-If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but at the cost of ignoring the model uncertainties.
+If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but at the cost of a potential bias.
 

### 2017-10-11 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 11.10.2017(3)

@@
   \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
+  + \hat{\pmb{\beta}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
 \end{align}
@@
 
-where the g-weight variance-covariance matrix of $\hat{\beta}_{s_2}$ is calculated as
+where the g-weight variance-covariance matrix of $\hat{\pmb{\beta}}_{s_2}$ is calculated as
 
@@
 \ref{eq:estvarmatrix}).
-The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formation is possible, as is the case with non-parametric models like kNN.
+The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formulation is possible, as is the case with non-parametric models like kNN.
 
@@
 Since being in the infinite population approach, $\pmb{Z}(x)$ again refers to the sample point $x$ and not the area around it.
-Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{x \in F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{x \in F} z_p(x)dx)^t$, implies that we need to calculate the spatial means of each component of $\pmb{Z}(x)$ using all possible points in $F$.
+Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^t$, implies that we need to calculate the spatial means of each component of $\pmb{Z}(x)$ using all possible points in $F$.
 This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
@@
 With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.
+One could also consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.
 
@@
 This method was suggested in \citet{mandallaz2013b} and led to an improvement in estimation precision.
-In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the \textit{weighted} mean where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.
+In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the \textit{weighted} mean (Eq.
+\ref{eq:wmeanZ}) where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.
 
@@
 To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository.
-The dataset contains data from a simple (i.e.
+The data set contains data from a simple (i.e.
 non-cluster) two-phase forest inventory conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
 The $s_1$ sample is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
-Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
 For a systematic subsample of 67 ($s_2$ sample), terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey.
@@
 Particular caution must be taken if categorical variables are present because the \code{lm()}-function, which is internally used to set up the design-matrix, automatically creates dummy variables with one of the categories used as a reference.
-The correct order for our \code{grisons} example can easily be extracted using:
+Using our \code{grisons} example, the correct order can always be extracted by the following \proglang{R}-code:
 
@@
 \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$.
-A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$.
+A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data also outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$.
 This is called the \textit{bias} or \textit{residual correction} term.
@@
 \hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
-\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
+\frac{1}{n_{1,G}(n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
 (\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
@@
 \hat{\bar{\pmb{\mathbb{Z}}}}_{G}
-+ \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
++ \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
 \end{align}
@@
 \end{subequations}
-where $\hat{\mathbb{R}}(x)=Y(x)-\bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2}$.
+where $\hat{\mathbb{R}}(x)$ are the empirical residuals under the extended auxiliary vector.
 
 To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
-When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias directly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply uses the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
+When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias directly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
 
@@
 The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need for the user to implement it.
-Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator vary between small areas, while they are identical under the \textit{psmall} estimator.
-This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices.
+Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator vary between the small areas, while they are identical under the \textit{psmall} estimator.
+This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrix.
 In case of the  \textit{psmall} estimator, the regression model stays the same for each small area estimation.
@@
 The \textbf{exhaustive versions} of the small area estimators (Eq.
-\ref{eq:pointest_2p_small},\ref{eq:var_2p_reg_small},\ref{eq:varext_2p_reg_small},\ref{eq:pointest_2p_reg_synth},\ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
+\ref{eq:pointest_2p_small}, \ref{eq:var_2p_reg_small}, \ref{eq:varext_2p_reg_small}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest.

### 2017-10-12 JStatSoft_forestinventory_master-concordance.tex
update Andreas 12.10.2017(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 295:%
-1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 806:%
+1 109 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 811:%
 136}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 807:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 812:%
 1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1227:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
 138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1232:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1237:%
 1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1284:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
 144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1289:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1294:%
 1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1538:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1543:%
 150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1543:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1548:%
 1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1738:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
 156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1743:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1755:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
 162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1760:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1770:%
 168 11 1}

### 2017-10-12 rnw/Main structure and modules of the Package.Rnw
update Andreas 12.10.2017(1)

@@
 From that it follows that we can define \textit{two nested} regression models.
-The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}^{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
+The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
 The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
-Less accurate predictions, $\hat{Y}^{0}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
+Less accurate predictions, $\hat{Y}^{(0)}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
 If there is a significant gain in model precision between the reduced and the full model and the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.
@@
 In contrast, the \textit{design-based} approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
-While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the observation $Y(x)$ at any location $x$ is fixed and \textit{not} the result of a stochastic process.
-Conclusively, the observation \textit{Y}(\textit{x}) still remains a random variable, but solemnly due to the random sample mechanism.
-A consequence of this approach is that the estimation properties of design-based regression estimators (i.e.
+While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local densities $Y(x)$ at any location $x \in F$ are fixed and \textit{not} the result of a stochastic process.
+A selected observation \textit{Y}(\textit{x}) still remains a random variable, but solely due to the random sample mechanism.
+A consequence of this approach is that the estimation properties of design-based regression estimators (e.g.
 unbiasedness) typically hold regardless of the model that is chosen.

### 2017-10-12 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 12.10.2017(1)

@@
   \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^{(0)2}(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
@@
 
-Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model.
+Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}^{(0)}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model.
 The \textbf{external variance} form under linear regression models is defined as
@@
 \begin{align}
-\hat{\var}_{ext}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}_{0}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
-\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}_{0}(x))
-+\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}_{0}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:extvar_reg3p_nex}
+\hat{\var}_{ext}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
+\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}^{(0)}(x))
++\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:extvar_reg3p_nex}
 \end{align}
@@
 \begin{flalign}
-\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}_{0}(x)) + 
+\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) + 
  (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
-\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}_{0}(x)) \nonumber &\\
+\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) \nonumber &\\
  &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
@@
 \end{subequations}
-where $\hat{R}_{0}(x)=Y(x)-\hat{Y}_{0}(x)$ with $\hat{Y}_{0}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$.
+where $\hat{R}^{(0)}(x)=Y(x)-\hat{Y}^{(0)}(x)$ with $\hat{Y}^{(0)}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$.
 

### 2017-10-12 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 12.10.2017(1)

@@
 
-We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive case.
+We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive cases.
 Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in Fig.
@@
 The \textbf{exhaustive} case requires a closer look.
-Since being in the infinite population approach, $\pmb{Z}(x)$ again refers to the sample point $x$ and not the area around it.
+In the infinite population approach, $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it.
 Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^t$, implies that we need to calculate the spatial means of each component of $\pmb{Z}(x)$ using all possible points in $F$.
@@
 $n_1=\infty$).
-Of course this is practically infeasible, but there are few cases where the exact mean can also be precisely calculated in the infinite population framework.
+Although it is practically infeasible to assess $\pmb{Z}(x)$ for every $x$, there are few cases where the exact mean can in fact be precisely calculated.
 The first case is when the explanatory variables are provided by polygon layers (e.g.
@@
 taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
-In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
+However, for most types of explanatory variables we will try to get an approximation of $\bar{\pmb{Z}}$ that is only negligibly different.
+\par
 
 One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
+\ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited.
+Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
+deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units.
+While in the infinite population approach this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large.
+However, the perfect tessellation implementation can also impose drawbacks.
+One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size.
+Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tesselation might often not be feasible.
+In the infinite population frame however, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$.
+With this respect, the infinite population approach provides more flexibility than the finite approach.
+
+% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
 \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.

### 2017-10-12 JStatSoft_forestinventory_master-concordance.tex
update Andreas 12.10.2017(2)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
-1 69 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 290:%
+1 68 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 289:%
 130 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 295:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 294:%
 1 109 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 811:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 810:%
 136}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 812:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 811:%
 1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1231:%
 138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1237:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1236:%
 1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1288:%
 144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1294:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1293:%
 1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1543:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1542:%
 150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1548:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1547:%
 1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1742:%
 156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1747:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1759:%
 162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1764:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1770:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1769:%
 168 11 1}

### 2017-10-12 rnw/Future Plans.Rnw
update Andreas 10.10.2017(2)
update Andreas 12.10.2017(2)

@@
 \item Implement parallel procedures for efficiently calculating many small areas.
-\item Allow functions to accept objects of class \code{data.table} from the \pkg{data.table} package to improve memory efficiency.
-\item Enable the user to choose other types of models than linear regressions fit with OLS.
+\item Allow functions to accept objects of class \code{data.table} from the \pkg{data.table} package \citep{dt2017} to improve memory efficiency.
+\item Enable the user to choose other types of models than linear regressions fitted with OLS.
 \end{itemize}

### 2017-10-12 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 12.10.2017(2)

@@
 
-Intuitively, the three phase estimator is simply the mean of the weaker predictions using the reduced model, corrected by the mean difference between the reduced model and the stronger full model, corrected by the mean difference between the ground truth and the full model predictions.
+Intuitively, the three phase estimator is simply the mean of the predictions using the reduced model, corrected by the mean difference between the reduced model predictions and the more accurate full model predictions, corrected by the mean difference between the ground truth and the full model predictions.
 For the compact version of the formula in the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.

### 2017-10-12 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 12.10.2017(2)

@@
 In the infinite population approach, $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it.
-Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^t$, implies that we need to calculate the spatial means of each component of $\pmb{Z}(x)$ using all possible points in $F$.
+Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^t$, implies that we need to calculate the spatial mean of each component of $\pmb{Z}(x)$ using all possible points in $F$.
 This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.

### 2017-10-12 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 12.10.2017(3)

@@
 \begin{equation}\label{ci_1phase}
-CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
-                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
+                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
 \end{equation}
@@
 \begin{equation}\label{ci_2phase_global}
-CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
-                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
+                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
 \end{equation}
@@
 \begin{equation}\label{ci_2phase_sae}
-CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
-                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
+CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
+                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
 \end{equation}
 
-where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
+where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).

### 2017-10-12 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 12.10.2017(3)

@@
 For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
-We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the point whereas the support is the spatial extent of the auxiliary information used to calculate those values.
+We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the sample point whereas the support is the spatial extent of the auxiliary information used to calculate those values.
 So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points.
@@
 
-One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
+One implementation to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
 \ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited.
@@
 Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tesselation might often not be feasible.
-In the infinite population frame however, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$.
+In the infinite population frame, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$.
 With this respect, the infinite population approach provides more flexibility than the finite approach.

### 2017-10-12 JStatSoft_forestinventory_master-concordance.tex
readyforsubmission

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
-1 68 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 289:%
+1 67 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
 130 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 294:%
-1 109 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 810:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
+1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 139 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
 136}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 811:%
-1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1231:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
+1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1237:%
 138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1236:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1242:%
 1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1288:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1294:%
 144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1293:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1299:%
 1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1542:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
 150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1547:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1553:%
 1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1742:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
 156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1747:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1759:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
 162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1764:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1769:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1775:%
 168 11 1}

### 2017-10-12 rnw/Future Plans.Rnw
readyforsubmission

@@
 Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future.
-As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub.
+As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub \citep{github_forestinventory}.
 Currently planned extensions include:

### 2017-10-12 rnw/Main structure and modules of the Package.Rnw
update Andreas 12.10.2017(2)
update Andreas 12.10.2017(3)
readyforsubmission

@@
 This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
-This is mainly due to the impossibility of a perfect tesselation over an amorphous forest area by any choice of plot shape.
+This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape.
 Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area.
@@
 	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
-   %\resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
 		\caption{} \label{fig:concmphase_and_sae_b}
@@
 \ref{fig:struct_package}).
-The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications.
+Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications.
 The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}).
@@
 While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
-We will also restrict discussion to simple sampling.
+We will also restrict discussion to simple sampling, while the formulas for cluster sampling are available in the technical reports \citep{mandallaz2016, mandallaz2013techa, mandallaz2013techb}.
 A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.
@@
 \centering
-\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_4.PNG}}
+\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
 \caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}

### 2017-10-12 rnw/Special Cases and Scenarios.Rnw
update Andreas 12.10.2017(3)
readyforsubmission

@@
 
-To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
+To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{development stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
 
@@
 \textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
-Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators is equivalent to post-stratification.
+Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators, is equivalent to post-stratification.
 For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).
@@
 One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
-The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign.
+The idea of cluster sampling is to increase the amount of information without increasing the travel costs of the terrestrial campaign.
 However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e.
@@
 
-However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
-If this concerns an $s_2$ sample location, the associated response variable can no longer be used.
+However, if an $s_2$ and/or $s_1$ point is missing a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
+If this concerns an $s_2$ sample location, the associated value of the response variable can no longer be used.
 
@@
 If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample.
-Conclusively, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:
+Therefore, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:
 

### 2017-10-12 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 12.10.2017(3)
readyforsubmission

@@
 
-\begin{subequations}\label{reg3p}
+\begin{subequations}\label{eq:reg3p}
 \begin{align}
@@
 
-The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
+The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
 $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
 The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a correction term likewise in the global estimation (Eq.
-\ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).
+\ref{eq:reg3p}).
 
@@
 \hat{\bar{\pmb{Z}}}_{1,G} + 
-\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}_{G}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:var_3p_reg_small}&\\
+\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
 \hat{\var}(\hat{Y}_{G,psmall,3p})& = \hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_{s_2} +
@@
 \hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
-\hat{\bar{\pmb{Z}}}_{1,G} \nonumber &\\
-&+ \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}_{G}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:var_3p_reg_psmall}
+\hat{\bar{\pmb{Z}}}_{1,G} + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
 \end{flalign}
@@
 $n_{2,G}=0$.
-Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (Eq.
+Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq.
 \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq.
@@
 
-The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.$\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
+The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.$\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
 This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
@@
 However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}.
-We thus first define the true means for each small area likewise we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):
+We thus first define the true means for each small area just as we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):
 
@@
 
-The \textbf{synthetic estimator} can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to ignore any terrestrial data available in the respective small area.
+The \textbf{synthetic estimator} can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.
 

### 2017-10-12 rnw/Twophase_Estimators and their Application.Rnw
readyforsubmission

@@
 We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive cases.
-Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in Fig.
-\ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points).
+Fig.
+\ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles) and densified to a larger sample $s_1$ (the points).
 For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
@@
 However, the perfect tessellation implementation can also impose drawbacks.
-One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size.
-Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tesselation might often not be feasible.
+One is that a perfect tessellation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size.
+Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tessellation might often not be feasible.
 In the infinite population frame, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$.
 With this respect, the infinite population approach provides more flexibility than the finite approach.
-
-% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig.
-\ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}.
-In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case).
-Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
-deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequently calculating the exact mean $\bar{\pmb{Z}}$.
-While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large.
-Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks.
-One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support.
-Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible.
-In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$.
-This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association.
-With this respect, the infinite population approach can thus provide more flexibility than the finite approach.
-One could also consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.
 
@@
 
+% \begin{subequations}\label{eq:var_2p_small_psmall}
+% \begin{align}
+%   \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
+%     + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
+%   \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
+%   + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
+%   + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
+% \end{align}
+% \end{subequations}
+
 \begin{subequations}\label{eq:var_2p_small_psmall}
@@
   \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
-    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
+    + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_small} \\
   \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
   + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
-  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
+  + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_psmall}
 \end{align}
@@
   \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
+  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
 \end{align}
@@
 \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}).
-Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
-However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
+Thus, the (pseudo) synthetic estimator has a potentially unobservable design-based bias.
 Also note that the residual variation can no longer be considered in the g-weight variance (Eq.
@@
 Due to the absence of available residuals in $G$, there is also no external variance form for the synthetic and pseudo synthetic estimator.
+
+% Left out for simplicity reasons in article, maybe include in Diss:
+% However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.
 
@@
 
-The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively.
+The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively optimistic.
 The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e.
@@
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
-Recalling that linear models fitted using OLS have zero mean residual property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model.
+Recalling that linear models fitted using OLS have zero mean residual property by construction also if categorical variables are used, this leads to unbiased point estimates.
 The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq.
@@
 To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
-When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias directly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
+When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias explicitly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
 
@@
 Hence, choosing to use \textit{psynth} is probably not desireable and is just applied here for demonstration purposes.
-The function will in this case ignore the terrestrial information in the data set.
+In this case the residual correction will not be applied.
 
@@
 
-For example, the \textbf{extended synthetic estimator} (\textit{extsynth}) can then be applied by
+The \textbf{extended synthetic estimator} (\textit{extsynth}) can then be applied by
 \begin{small}

### 2017-10-12 r/analysis_and_visualization.R

@@ -65,6 +65,10 @@ grisons.sae.table
 ## see structure:
 str(grisons.sae.table)
 
+## turn into data.frame:
+grisons.sae.df<- data.frame(grisons.sae.table)
+class(grisons.sae.df)
+
 
 #####################################################
 # ----- mphase.gain() ----------------------------- #
@@ -80,10 +84,11 @@ mphase.gain(grisons.sae.table, pref.vartype = "g_variance")
 #####################################################
 
 ## plot point estimates and confidence intervals:
-plot(grisons.sae.table, ncol = 2, yvar = "estimate") 
+plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
+  ylab("Timber Volume [m3/ha]")
 
 ## plot estimation errors:
-plot(grisons.sae.table, ncol = 2, yvar = "error") 
+plot(grisons.sae.table, ncol = 2) 
 
 
 #######

### 2017-10-12 r/special_cases_and_scenarios.R

@@ -22,30 +22,35 @@ grisons.n<- grisons
 grisons.n$stage<- as.factor(kmeans(grisons.n$mean, centers = 3)$cluster)
 
 ## plot:
+library(ggplot2)
 ggplot(data=grisons.n, aes(x = mean, y = tvol)) +
   geom_point(aes(colour = factor(stage)))
 
 
 ## apply 'double sampling for (post)stratification':
-twophase(formula=tvol ~ stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
+summary(
+  twophase(formula=tvol ~ stage,
+           data=grisons.n,
+           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+           boundary_weights = "boundary_weights")
+)
 
 
 ## apply 'double sampling for regression':
-twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
-
+summary(
+  twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
+           data=grisons.n,
+           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+           boundary_weights = "boundary_weights")
+)
 
 ## apply 'double sampling for regression within (post)strata'":
-twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-         data=grisons.n,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-         boundary_weights = "boundary_weights")
-
+summary(
+  twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
+           data=grisons.n,
+           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+           boundary_weights = "boundary_weights")
+)
 
 
 #####################################################
@@ -211,7 +216,7 @@ tp<- threephase(formula.s0 = tvol ~ mean,
                 boundary_weights = "boundary_weights")
 
 # Violation:
-# s2 point misses expl.variable used (reduced model) at s1 sample points
+# s2 point misses expl.variable used in reduced model at s1 sample points
 
 # warning message:
 # s2 point misses expl.variable used in s1

### 2017-10-12 r/three_phase_estimators.R

@@ -1,12 +1,12 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 4  "Three-phase Estimators and their Application"                        #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples given in                                            #
+#  Section 4  "Three-phase Estimators and their Application"                         #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
 
 library(forestinventory)
 

### 2017-10-12 r/two_phase_estimators.R

@@ -1,12 +1,12 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 3  "Two-phase Estimators and their Application"                          #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples given in                                            #
+#  Section 3  "Two-phase Estimators and their Application"                           #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
 
 library(forestinventory)
 
@@ -117,8 +117,7 @@ psynth_2p$estimation
 
 # --------------------------------- #
 ## extended synthetic estimator (extpsynth)
-#  with boundary weight adjustment
-
+# boundary adjustment was already applied when retrieving the exact means
 
 ## retrieve order of columnames that the lm()-function processes a formula object:
 #  --> actually the order they appear in the design-matrix ...

### 2017-10-24 JStatSoft_forestinventory_master.rnw
update Andreas 12.10.2017(2)
readyforsubmission
formatting revision (1)

@@
 \documentclass[article]{jss}
-% \documentclass[nojss]{jss} % for creating the vignette
+%\documentclass[nojss]{jss} % for creating the vignette
 
@@
 They typically consist of a random sample of plot locations in the forest that are assessed individually by hand.
-Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs (e.g.
-LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
-With respect to this objective, the application of \textit{multi-phase} forest inventory methods (e.g.
-double- and triple-sampling regression estimators) has proved to be efficient.
+Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
+With respect to this objective, the application of multi-phase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient.
 While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent.
@@
 \Plainkeywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
+
+ % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
+ % ... free --> more suited for forest practice (limited financial ressources)
 
@@
  }
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage[font=footnotesize]{subcaption}
 
@@
 \begin{document}\sloppy
-\SweaveOpts{concordance=TRUE}
 
@@
 <<echo=FALSE>>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
 @

### 2017-10-24 rnw/Analysis and Visualization.Rnw
formatting revision (1)

@@
 
-\section{Analysis and Visualization}
+% set global options for R-output:
+<<echo=FALSE>>=
+options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
+@
+
+\section{Analysis and visualization}
 \label{sec:anal_and_vis}
@@
 This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis.
-For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimators for the \code{grisons} data set:
+For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase extended pseudo synthetic and the pseudo synthetic estimators for the \code{grisons} data set:
 
@@
 <<>>=
-extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                        small_area = list(sa.col = "smallarea", 
-                                          areas = c("A", "B","C", "D"),
-                                          unbiased = TRUE))
+extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+  small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
+  unbiased = TRUE), boundary_weights = "boundary_weights")
 @
@@
 <<>>=
-psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights",
-                      small_area = list(sa.col = "smallarea", 
-                                        areas = c("A", "B", "C", "D"),
-                                        unbiased = FALSE))
+psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+  small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+  unbiased = FALSE), boundary_weights = "boundary_weights")
 @
@@
   phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                unbiased = TRUE),
-                boundary_weights = "boundary_weights")
+  small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+  unbiased = TRUE), boundary_weights = "boundary_weights")
 @
@@
  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                small_area = list(sa.col = "smallarea", 
-                                  areas = c("A", "B", "C", "D"),
-                                  unbiased = FALSE),
-                boundary_weights = "boundary_weights")
+ small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+ unbiased = FALSE), boundary_weights = "boundary_weights")
 @
@@
 grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
-                                             extpsynth_3p, psynth_3p),
-                             sae = TRUE, 
-                             vartypes = c("variance", "g_variance",  "ext_variance"))
+  extpsynth_3p, psynth_3p), sae = TRUE, vartypes = c("variance", "g_variance", 
+  "ext_variance"))
 @
@@
 The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package.
-However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance that was applied (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance).
+However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance that was applied (\code{g_} for \textit{g-weight} and \code{ext_} for external variance).
 By default, the confidence intervals are also added.
@@
 
-Note that \code{estTable()} also returns the \textbf{\textit{estimation error}} (\code{error}) that is defined as the \textit{standard error} devided by the \textit{point estimate}:
+Note that \code{estTable()} also returns the estimation error (\code{error}) that is defined as the standard error devided by the point estimate:
 
@@
 As multi-phase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multi-phase global or small area estimate compared to its respective one-phase estimate.
-The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a \textit{baseline}.
+The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a baseline.
 If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
-The argument \code{pref.vartype} can be used to define what type of variance (\textit{g-weight} or \textit{external}) should be used for the comparison.
-Synthetic estimates (\textit{synth} and \textit{psynth} estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.
-
-<<echo=FALSE>>=
-options(width=10000) 
-@
+The argument \code{pref.vartype} can be used to define what type of variance (g-weight or external) should be used for the comparison.
+Synthetic estimates (SYNTH and PSYNTH estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.
 
@@
 The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation.
-For example, it is indicated that the \textit{two-phase extended psynth} estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
-The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:
+For example, it is indicated that the two-phase extended PSYNTH estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
+The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multi-phase variance:
 
@@
 
-The \textit{relative efficiency} can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
-For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{two-phase extended psynth} estimator.
+The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
+For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the two-phase EXTPSYNTH estimator.
 If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multi-phase estimation for forest inventories.
@@
 
-The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: 1) the \textit{\textbf{point estimates}} with overlayed confidence intervals, and 2) the \textit{\textbf{estimation errors}}.
+The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: 1) the point estimates with overlayed confidence intervals, and 2) the estimation errors.
 Both plots can be obtained by passing the \code{esstable} object to the \code{plot()} function.

### 2017-10-24 rnw/Calculation of Confidence Intervals.Rnw
formatting revision (1)

@@
 
-\section{Calculation of Confidence Intervals}
+% set global options for R-output:
+<<echo=FALSE>>=
+options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
+@
+
+\section{Calculation of confidence intervals}
 \label{sec:confint_calc}
 
-Converting the estimated variance into a 95\% \textit{confidence interval} (CI) allows for a more practical interpretation of a point estimate's precision.
+Converting the estimated variance into a 95\% confidence interval (CI) allows for a more practical interpretation of a point estimate's precision.
 The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value.
@@
 The confidence level $1-\alpha$ (e.g.
-95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as \textit{coverage rate}.
+95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate.
 The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries.
@@
 
-\textit{One-Phase Estimation}
+\vspace{8mm}
+
+One-phase estimation:
 
@@
 % ------------------------------------------ %
-\textit{Two-Phase and Three-Phase Global Estimation}
+Two-phase and three-phase global estimation:
 
@@
 % ------------------------------------------ %
-\textit{Two-Phase and Three-Phase Small Area Estimation}
+Two-phase and three-phase small area estimation:
 
@@
 
+\vspace{7mm}
+
 where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model.
@@
 In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (Section \ref{sec:twophase_sae}) are calculated by:
+For example, the 95\% confidence interval for the small area estimates by the extpsynth estimator (Section \ref{sec:twophase_sae}) are calculated by:
 

### 2017-10-24 rnw/Introduction.Rnw
formatting revision (1)

@@
 As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods.
-Typically this means that discrete sample locations (\textit{sample plots}) are randomly chosen in the forest, making up the framework of a \textit{terrestrial inventory}.
+Typically this means that discrete sample locations (sample plots) are randomly chosen in the forest, making up the framework of a terrestrial inventory.
 This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals.
@@
 As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is called \textit{multi-phase sampling}.
-The core concept is to enlarge the sample size in order to gain higher estimation precision \textit{without} enlarging the terrestrial sample size.
+A method which has become particularly attractive is called multi-phase sampling.
+The core concept is to enlarge the sample size in order to gain higher estimation precision without enlarging the terrestrial sample size.
 This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered.
@@
 data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either \textit{one} additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or \textit{two} additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
+Regression estimators using this concept can consider either one additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
 Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
@@
 Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare.
-Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the \textit{finite population approach}, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
 However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}.
-The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the \textit{infinite population approach} by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
+The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
@@
 
-The objectives of this article are to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} address special cases and demonstrate how the package-functions handle them.
+The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.
 
-\newpage
-
-% what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% NOTES:
-% --> take parts of the old introduction version of the diploma-thesis!
-
-% NOTES:
-% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
-%                model-dependent vs. design-based
-% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
-%                 and seperate them from our package (focus, applied areas, approach)
-%              + important reasoning for our package: we are the first that implemented and provide 
-%                a large range of design-based estimators
-% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
-%              available to a large user community, to facilitate their application in science as well as 
-%              operational forest management (why is this important? --> because the methods have demonstrate
-%              the potential of considerably increasing the estimation precision)
-% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
-%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
-%                and their implementation in our package
-%             b) illustrate the application of the various estimators by the respsective functions in our package 
-%                to real-world inventory scenarios
-%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
-%                deal with such situations (error-checking functions and data-adjustments)
-
-% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
-%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
-%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
-%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
-%
-% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)
-
-% Explain the problem that leads to sae estimations (in generell and in forestry)
-%
-%
+%\newpage

### 2017-10-24 rnw/Main structure and modules of the Package.Rnw
formatting revision (1)

@@
 
-\section{Methods and Structure of the Package}
+\section{Methods and structure of the package}
 \label{sec:str_and_mod}
 
-\subsection{Forest Inventory in the Infinite Population Approach}
+\subsection{Forest inventory in the infinite population approach}
 
-Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by GPS coordinates.
-The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g.
-timber volumes) to local plot densities (e.g.
-the timber density in $m^3/ha$).
-The estimators implemented in \pkg{forestinventory} use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators.
+Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by coordinates.
+The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha).
+The estimators implemented in \pkg{forestinventory} use the so called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators.
 This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig.
 \ref{fig:inf_pop_apr}.
-The \textit{population total} of the target variable (e.g.
-the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig.
+The population total of the target variable (e.g., the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig.
 \ref{fig:inf_pop_apr} as the volume under the density surface.
 From this perspective, the practical challenge is that the form of this function is unknown.
-Theoretically, we could get the total timber volume by observing the function value, i.e.
-the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum.
+Theoretically, we could get the total timber volume by observing the function value, i.e., the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum.
 However, this is impossible because there is an infinite number of points in the forest area.
-Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e.
-$Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
+Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e., $Y(x)$, to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
 The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
 All estimators included in \pkg{forestinventory} rest upon this theoretical perspective.
-The key point in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
-This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
+The key point in the infinite population approach is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
+This has some theoretical advantages over the finite population approach, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
 This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape.
@@
 
-\subsection{Two-Phase Sampling}
+\subsection{Two-phase sampling}
 
-The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Fig.
+The two-phase or double-sampling estimators use inventory information from two nested samples which are commonly referred to as phases (Fig.
 \ref{fig:concmphase_and_sae_a}).
-The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ at each point $x \in s_1$.
+The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$.
 These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$.
-The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e.
-the local density $Y(x)$ (e.g.
-the timber volume per hectare).
+The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e., the local density $Y(x)$ (e.g., the timber volume per hectare).
 For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some model, which in \pkg{forestinventory} is always a linear model fit in $s_2$ using ordinary least squares (OLS).
-The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e.
-$\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.
+The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.
 In the design-based context, the two-phase estimator is typically unbiased regardless of the model used to produce the predictions.
@@
 
-\subsection{Three-Phase Sampling}
+\subsection{Three-phase sampling}
 
-Three-phase estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (Fig.
+Three-phase estimators extend the principle of two-phase sampling and use inventory information from three nested samples (phases) (Fig.
 \ref{fig:concmphase_and_sae_a}).
@@
 The motivation for three-phase sampling is that the additional set of explanatory variables available at $s_1$, now denoted $\pmb{Z}^{(1)}(x)$, adds considerable explanatory power to the set of variables available at all sample locations $x \in s_0$, denoted $\pmb{Z}^{(0)}(x)$.
-From that it follows that we can define \textit{two nested} regression models.
-The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ can be used to calculate the predictions $\hat{Y}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
-The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}.
+From that it follows that we can define two nested regression models.
+The full set of predictor variables $\pmb{Z}^{\top}(x)=(\pmb{Z}^{(0){\top}}(x),\pmb{Z}^{(1){\top}}(x))$ can be used to calculate the predictions $\hat{Y}(x)$ of $Y(x)$ at all sample locations $x \in s_1$.
+The regression model applicable to the $s_1$ phase is thus referred to as the full model.
 Less accurate predictions, $\hat{Y}^{(0)}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$.
@@
 	\end{subfigure}
-\caption{\textbf{(a)} Concept of multiphase sampling.
+\caption{(a) Concept of multiphase sampling.
 The square represents the forest area for which an inventory is being conducted.
 The points denote the sample locations $x$.
-Filled points indicate \textit{available} information.
-\textbf{(b)} Illustration of the Small Area Estimation problem}
+Filled points indicate available information.
+(b) Illustration of the small area estimation problem.}
 \label{fig:concmphase_and_sae}
@@
 
-\subsection{Small Area Estimation}
+\subsection{Small area estimation}
 
@@
 However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals.
-The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area.
+The idea is then to borrow strength from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area.
 The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area.
-If there are no terrestrial plots in $G$ (i.e.
-$n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator.
-These are called \textit{synthetic} estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.
+If there are no terrestrial plots in $G$ (i.e., $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator.
+These are called synthetic estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.
 
-\subsection{Design-Based vs. Model-Dependent Approach}
+\subsection{Design-based vs. model-dependent approach}
 \label{sec:db_vs_md}
@@
 This is why it is important to understand that the mathematical interpretation of how a model is used to produce estimates is fundamentally different between the design-based and model-dependent approach.
-In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
-Although the model does not need to be fit from a probability sample, i.e.
-the sample locations could arbitrarily be chosen, the model should adequately describe the underlying stochastic process in order to efficiently ensure unbiased results.
+In the model-dependent (also known as model-based) framework, the sample locations $x$ are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process.
+Although the model does not need to be fit from a probability sample, i.e., the sample locations could arbitrarily be chosen, the model should adequately describe the underlying stochastic process in order to efficiently ensure unbiased results.
 In practice this means that special attention must be made to ensure that the variable selection is appropriate to avoid overfitting, important variables are not omitted and all model assumptions are reasonably met through empirical verification.
 If a model is misspecified then estimation based on inference from that model may not be reliable.
-In the model-dependent framework one thus has to \textit{trust} the model.
-In contrast, the \textit{design-based} approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
-While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local densities $Y(x)$ at any location $x \in F$ are fixed and \textit{not} the result of a stochastic process.
-A selected observation \textit{Y}(\textit{x}) still remains a random variable, but solely due to the random sample mechanism.
-A consequence of this approach is that the estimation properties of design-based regression estimators (e.g.
-unbiasedness) typically hold regardless of the model that is chosen.
-The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics.
-They are therefore also referred to as \textit{model-assisted}.
+In the model-dependent framework one thus has to trust the model.
+In contrast, the design-based approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
+While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local densities $Y(x)$ at any location $x \in F$ are fixed and not the result of a stochastic process.
+A selected observation $Y(x)$ still remains a random variable, but solely due to the random sample mechanism.
+A consequence of this approach is that the estimation properties of design-based regression estimators (e.g., unbiasedness) typically hold regardless of the model that is chosen.
+The philosophy of the design-based approach is thus to use prediction models to improve the efficiency of the estimators without having to rely on their correct specification, which makes them very attractive to be used in official statistics.
+They are therefore also referred to as model-assisted.
 It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to minimize travelling costs in the terrestrial survey.
@@
 
-\newpage
+%\newpage
 
-\subsection{Package Structure}
+\subsection{Package structure}
 \label{sec:packstruc}
@@
 Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications.
-The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}).
+The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (global estimators) or only for subregions of the entire inventory area (small area estimators).
 In the second case, the package offers three small area estimators that will in detail be described in the following sections.
-The estimators are available under \textit{exhaustive} and \textit{non-exhaustive} use of the auxiliary data.
+The estimators are available under exhaustive and non-exhaustive use of the auxiliary data.
 Additionally, the package also can calculate one-phase estimates solely based on terrestrial samples.
-All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
+All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
 The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
@@
 \resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
-\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
+\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}.}
 \label{fig:struct_package}

### 2017-10-24 rnw/Special Cases and Scenarios.Rnw
formatting revision (1)

@@
 
-\section{Special Cases and Scenarios}
+\section{Special cases and scenarios}
 \label{sec:speccas_and_scen}
@@
 % --------------------------------------------------------- %
-\subsection{Post-Stratification}
+\subsection{Post-stratification}
 
-A special case of multi-phase regression estimators is \textit{post-stratification}, which can further be divided into the cases of \textit{multi-phase sampling for stratification} and \textit{multi-phase sampling for regression within strata}.
+A special case of multi-phase regression estimators is post-stratification, which can further be divided into the cases of multi-phase sampling for stratification and multi-phase sampling for regression within strata.
 Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.
 
-To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{development stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
+To demonstrate post-stratification, we first create an artificial categorical variable development stage (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:
 
@@
 
-\textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
+Two-phase sampling for stratification is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
 Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators, is equivalent to post-stratification.
@@
 <<eval=FALSE>>=
-# two-phase sampling for stratification
-twophase(formula=tvol ~ stage,
-         data=grisons,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+twophase(formula = tvol ~ stage, data = grisons, 
+  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   boundary_weights = "boundary_weights")
@@
 
-\textit{Two-phase sampling for regression within strata} implies the combination of continuous and categorical variables within the model (i.e.
-we have an ANCOVA model).
+Two-phase sampling for regression within strata implies the combination of continuous and categorical variables within the model (i.e., we have an ANCOVA model).
 If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts.
@@
 <<eval=FALSE>>=
-# two-phase sampling for regression within strata:
-twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-         data=grisons,
-         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+twophase(formula=tvol ~ mean + stddev + max + q75 + stage, data = grisons,
+  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   boundary_weights = "boundary_weights")
@@
 % --------------------------------------------------------- %
-\subsection{Small Area Estimation under Cluster Sampling}
+\subsection{Small area estimation under cluster sampling}
 
-As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points.
+As mentioned in Section \ref{sec:packstruc}, cluster sampling is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points.
 One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$.
 The idea of cluster sampling is to increase the amount of information without increasing the travel costs of the terrestrial campaign.
-However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e.
-the cluster origin $x$.
+However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e., the cluster origin $x$.
 Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights.
-Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e.
-defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
+Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e., defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random.
 The \pkg{forstinventory} package identifies clusters via a unique cluster ID that is assigned to a column in the input data set.
@@
 
-For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e.
-only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
-In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
-In this case, it is adviseable to use the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still ensured.
+For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e., only a subset $M_{G}(x) < M(x)$ is included in the small area of interest.
+In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the extended and pseudo extended synthetic estimator (see EXTSYNTH and EXTPSYNTH in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
+In this case, it is adviseable to use the (pseudo) small area estimator (PSMALL) where the zero mean residual property is still ensured.
 
@@
 \begin{Sinput}
-R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
-+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-+                      cluster = "cluster",
-+                      small_area = list(sa.col = "ismallold", areas = c("1"),
+R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, 
++  data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
++  cluster = "cluster",  small_area = list(sa.col = "ismallold", areas = c("1"),
 +  unbiased = TRUE))
@@
 <<>>=
-extpsmall.clust <- twophase(formula = basal ~ stade + couver + melange, data=zberg,
-                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      cluster = "cluster",
-                      small_area = list(sa.col = "ismallold", areas = c("1"),
-                                        unbiased = TRUE),
-                     psmall = TRUE)
+psmall.clust <- twophase(formula = basal ~ stade + couver + melange, 
+  data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+  cluster = "cluster", small_area = list(sa.col = "ismallold", areas = c("1"),
+  unbiased = TRUE), psmall = TRUE)
 @
@@
 <<>>=
-extpsmall.clust$estimation
+psmall.clust$estimation
 @
@@
 
-Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the g-weight variances only differ marginally.
+Comparing the EXTPSYNTH and PSMALL estimates, we see that in this particular case the point estimates are close and more important, the external as well as the g-weight variances only differ marginally.
 This can be taken as evidence that the violation of the zero mean residual property can here be expected to have negligible consequences.
@@
 % --------------------------------------------------------------------- %
-\subsection{Violation of Nesting in Sample Design}
+\subsection{Violation of nesting in sample design}
 
-As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multi-phase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other.
+As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multi-phase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other.
 The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig.
@@
 
-Based on the nesting rule, $s_2 \in s_1 \in s_0$, each $s_2$ and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model.
-If e.g.
-an $s_2$ and/or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (in this case \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:
+Based on the nesting rule, $s_2 \in s_1 \in s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model.
+If e.g., an $s_2$ and/or $s_1$ point misses a variable which is used in the full and reduced model (in this case \code{mean}), the function will delete this sample point from the dataset and produce the following messages:
 
@@
 <<eval=FALSE>>=
-## delete "mean" value from an s2- (i.e. s1- and s0-) sample point:
-grisons[which(grisons$phase_id_3p==2)[1],"mean"] <- NA
+grisons[which(grisons$phase_id_3p == 2)[1], "mean"] <- NA
 @
@@
 R> threephase(formula.s0 = tvol ~ mean, 
-+             formula.s1 = tvol ~  mean + stddev + max + q75, 
-+             data = grisons,
-+             phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
++  formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
++  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
 +  boundary_weights = "boundary_weights")
@@
 1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the first phase (s1) is missing
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
+  parameter of the first phase (s1) is missing
 2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the zero phase (s0) is missing
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
+  parameter of the zero phase (s0) is missing
 3: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  1 rows deleted due to missingness in the set of auxiliary parameters for the zero phase (s0)
-  (1 terrestrial plots affected by deletion)
+  1 rows deleted due to missingness in the set of auxiliary parameters for the 
+  zero phase (s0) (1 terrestrial plots affected by deletion)
 \end{Sinput}
@@
 
-However, if an $s_2$ and/or $s_1$ point is missing a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
+However, if an $s_2$ and/or $s_1$ point is missing a variable which is only used in the full regression model (in this example \code{q75}), the function will recode the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model.
 If this concerns an $s_2$ sample location, the associated value of the response variable can no longer be used.
@@
 <<eval=FALSE>>=
-## delete "q75" value from an s2-sample point:
-grisons[which(grisons$phase_id_3p==2)[1],"q75"] <- NA
+grisons[which(grisons$phase_id_3p == 2)[1], "q75"] <- NA
 @
@@
 R> threephase(formula.s0 = tvol ~ mean, 
-+             formula.s1 = tvol ~  mean + stddev + max + q75, 
-+             data = grisons,
++  formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
 +  phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
@@
 1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  Sample design not nested: for 1 terrestrial plots at least one auxiliary parameter 
-  of the first phase (s1) is missing
+  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
+  parameter of the first phase (s1) is missing
 2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in the set of
-  auxiliary parameters for the first phase (s1) (1 terrestrial information no longer usable 
-  by this change)
+  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in 
+  the set of auxiliary parameters for the first phase (s1) (1 terrestrial 
+  information no longer usable by this change)
 \end{Sinput}
@@
 
-If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations are \textbf{deleted} from the data set.
+If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations are deleted from the data set.
 
@@
 <<eval=FALSE>>=
-## delete "mean" value from an s0-sample point:
-grisons[which(grisons$phase_id_3p==0)[1],"mean"] <- NA
+grisons[which(grisons$phase_id_3p == 0)[1], "mean"] <- NA
 @
@@
 R> threephase(formula.s0 = tvol ~ mean, 
-+             formula.s1 = tvol ~  mean + stddev + max + q75, 
-+             data = grisons,
-+             phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
++  formula.s1 = tvol ~  mean + stddev + max + q75,  data = grisons,
++  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
 +  boundary_weights = "boundary_weights")
@@
 In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
-  1 rows deleted due to missingness in the set of auxiliary parameters for the zero phase (s0) 
-  (0 terrestrial plots affected by deletion)
+  1 rows deleted due to missingness in the set of auxiliary parameters for the 
+  zero phase (s0) (0 terrestrial plots affected by deletion)
 \end{Sinput}
@@
 Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures.
-If any information should be missing at the sample locations, one should clarify the reason for this and make sure that the information can reasonably be assumed to be \textit{completely missing at random}.
+If any information should be missing at the sample locations, one should clarify the reason for this and make sure that the information can reasonably be assumed to be completely missing at random.
 
@@
 If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample.
-Therefore, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:
+Therefore, an internal regression model cannot be calculated and the function stops with the following error message:
 
@@
 <<eval=FALSE>>=
-## delete s2-points with "stade"-level '300'
-zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade=="300"), ]
+zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade == "300"), ]
 @
@@
 \begin{Sinput}
-R> twophase(formula = basal ~ stade + couver + melange,
-+           data = zberg,
+R> twophase(formula = basal ~ stade + couver + melange, data = zberg,
 +  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 Error in check.mandatoryInputs(formula, data, phase_id) : 
- Level '300' of factor variable 'stade' existing in s1(s0)- but not in s2 sample.
+ Level '300' of factor variable 'stade' existing in s1(s0)-but not in s2 sample.
  Calculation of coefficient not feasible.

### 2017-10-24 rnw/Threephase_Estimators and their Application.Rnw
formatting revision (1)

@@
 
-\section{Three-phase Estimators and their Application}
+% set global options for R-output:
+<<echo=FALSE>>=
+options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
+@
+
+\section{Three-phase estimators and their application}
 \label{sec:threephase_and_appl}
@@
 % ---------------------------------------------------------------------------- %
-\subsection{Global Estimators}
+\subsection{Global estimators}
 \label{sec:glob_est_3p}
@@
 % ------------------------------------- %
-\subsubsection{Mathematical Background}
+\subsubsection{Mathematical background}
 
-Solving the sample-based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the \textbf{\textit{reduced model}}, i.e.
-using the \textit{reduced set of explanatory variables} $\pmb{Z}^{(0)}(x)$ available at $x \in s_0$, and likewise the vector of regression coefficients $\hat{\pmb{\beta}}_{s_2}$ for the \textbf{\textit{full model}}, i.e.
-using the \textit{full set of explanatory variables} $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$ available only at a subset $x \in s_1 \subset s_0$, are derived as
+Solving the sample-based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the reduced model, i.e., using the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$ available at $x \in s_0$, and likewise the vector of regression coefficients $\hat{\pmb{\beta}}_{s_2}$ for the full model, i.e., using the full set of explanatory variables $\pmb{Z}^{\top}(x)=(\pmb{Z}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$ available only at a subset $x \in s_1 \subset s_0$, are derived as
 
@@
 \begin{align}
-\hat{\pmb{\alpha}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)
+\hat{\pmb{\alpha}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x)
 \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)  \label{eq:normequ_redmod} \\
-\hat{\pmb{\beta}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^t(x)
+\hat{\pmb{\beta}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^{\top}(x)
 \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x) \label{eq:normequ_fullmod}
@@
 
-The package allows for the calculation of \textbf{point estimates} under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase.
-Fitting the model using $s_2$ (i.e.
-\textit{internally}) ensures the zero mean residual property over $s_2$.
-
-% \begin{subequations}\label{reg3p}
-% \begin{align}
-% \hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
-% \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
-% \hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
-% \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
-% \end{align}
-% \end{subequations}
-% 
-% In the non-exhaustive case, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
-\ref{meanvalues3pnex}.
-If the exact mean over $s_0$ is known, the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ can simply be replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$.
-Note that in case of applied boundary adjustment (Section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to Eq.
-\ref{eq:wmeanZ}.
+The package allows for the calculation of point estimates under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase.
+Fitting the model using $s_2$ (i.e., internally) ensures the zero mean residual property over $s_2$.
 
@@
 \begin{align}
-\hat{Y}_{reg3p,ex}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2})
-\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
+\hat{Y}_{reg3p,EX}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
+\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2} +
 \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
-\hat{Y}_{reg3p,nex}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{t}(x)\hat{\pmb{\beta}}_{s_2})
-\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_{s_2}  +
+\hat{Y}_{reg3p,NEX}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
+\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2}  +
 \hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
@@
 Intuitively, the three phase estimator is simply the mean of the predictions using the reduced model, corrected by the mean difference between the reduced model predictions and the more accurate full model predictions, corrected by the mean difference between the ground truth and the full model predictions.
-For the compact version of the formula in the \textit{non-exhaustive case}, the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
+For the compact version of the formula in the non-exhaustive case, the estimated means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the estimated mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.
 \ref{meanvalues3pnex}.
@@
 
-% An intuitive interpretation of \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex} is a follows: Likewise in the two-phase approach, the estimated means in $\hat{\bar{\pmb{Z}}}_1$ based on the $s_1$ sample are transformed into predictions $\hat{Y}_{1}(x)$ of the local density $Y(x)$.
-The difference between the estimated means $\hat{\bar{\pmb{Z}}}^{(0)}_1$ and their more precise or even exact counterparts $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and $\bar{\pmb{Z}}^{(0)}_0$ based on the larger or infinite $s_0$ sample is then added as a residual term that expresses the lack of precision in the point estimate due to varying sample sizes behind the estimated auxilary means.
-
-The package again provides the g-weight and external variances. The \textbf{g-weight variance} formulation is
+The package again provides the g-weight and external variances. The g-weight variance formulation is
 
@@
 \begin{align}
-\hat{\var}(\hat{Y}_{reg3p,ex})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-\bar{\pmb{Z}}^{(0)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\var}(\hat{Y}_{reg3p,EX})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
 \hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_ex} \\
-\hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_{s_2} ^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_{s_2}
-+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
-\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
+\hat{\var}(\hat{Y}_{reg3p,NEX})& =\hat{\pmb{\alpha}}_{s_2} ^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_{s_2}
++\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_0
+\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
 \end{align}
@@
   \frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
-  (\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^{(0)2}(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \label{eq:covar_beta}
+  (\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^{\top} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^{(0)2}(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x) \Big)^{-1} \label{eq:covar_alpha} \\
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^{\top}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1} \label{eq:covar_beta}
 \end{align}
@@
 
-Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}^{(0)}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model.
-The \textbf{external variance} form under linear regression models is defined as
+Note that $\hat{R}(x)=Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}^{(0)}(x)=Y(x)-\pmb{Z}^{(0)\top}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model.
+The external variance form under linear regression models is defined as
 
@@
 \begin{align}
-\hat{\var}_{ext}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
-\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}^{(0)}(x))
+\hat{\var}_{ext}(\hat{Y}_{reg3p,EX})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
+\hat{\var}_{ext}(\hat{Y}_{reg3p,NEX})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}^{(0)}(x))
 +\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:extvar_reg3p_nex}
@@
 <<>>=
-formula.rm <- tvol ~ mean # reduced model applied to s0 phase
-formula.fm <- tvol ~ mean + stddev + max + q75 # full model applied to s1 phase
+formula.rm <- tvol ~ mean
+formula.fm <- tvol ~ mean + stddev + max + q75
 @
@@
 
-Compared to the \code{twophase()}-function, we now have to specify two regression models, i.e.
-the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
+Compared to the \code{twophase()}-function, we now have to specify two regression models, i.e., the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model.
 In addition, we also have to specify the indication of the $s_1$ phase (\code{s1.id}) in the argument \code{phase_id} (note that \pkg{forestinventory} implicitly assumes that all other rows in the input data set belong to $s_0$).
@@
 
-<<echo=FALSE>>=
-options(width=10000) 
-@
-
 \begin{small}
 <<>>=
-reg3p_nex <- threephase(formula.s0 = formula.rm, 
-                  formula.s1 = formula.fm, data = grisons,
-                  phase_id = list(phase.col="phase_id_3p", s1.id = 1, 
-                                  terrgrid.id = 2),
-                  boundary_weights = "boundary_weights")
+reg3p_nex <- threephase(formula.s0 = formula.rm, formula.s1 = formula.fm, 
+  data = grisons, phase_id = list(phase.col = "phase_id_3p", s1.id = 1,  
+  terrgrid.id = 2), boundary_weights = "boundary_weights")
 @
@@
 
-The \code{summary()} of a \code{threephase()}-function now recalls \textit{both} regression model formulas and also gives the $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) models.
+The \code{summary()} of a \code{threephase()}-function now recalls both regression model formulas and also gives the $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) models.
 We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$.
@@
 % ---------------------------------------------------------------------------- %
-\subsection{Small Area Estimators}
+\subsection{Small area estimators}
 \label{sec:threephase_sae}
@@
 % ------------------------------------- %
-\subsubsection{Mathematical Background}
+\subsubsection{Mathematical background}
 
 The three two-phase small area estimators described in Section \ref{sec:twophase_sae} can also be extended to the three-phase scenario.
-The general principle thereby stays the same, i.e.
-the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to Eq.
+The general principle thereby stays the same, i.e., the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to Eq.
 \ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.
@@
 
-The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e.
-$\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
+The unbiased point estimates of the SMALL and PSMALL are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e., $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^{\top}(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
 The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a correction term likewise in the global estimation (Eq.
@@
 \begin{align}
-\hat{Y}_{G,small,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_small} \\
-\hat{Y}_{G,psmall,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_{s_2} +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_psmall}
+\hat{Y}_{G,SMALL,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_{s_2} +
+\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_small} \\
+\hat{Y}_{G,PSMALL,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_{s_2} +
+\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_psmall}
 \end{align}
@@
 
-The \textbf{g-weight variance} is then calculated as
+The g-weight variance is then calculated as
 
 \begin{subequations}\label{eq:var_3p_small_psmall}
-\begin{flalign}
-\hat{\var}(\hat{Y}_{G,small,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\begin{align}
+\hat{\var}(\hat{Y}_{G,SMALL,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
 \hat{\bar{\pmb{Z}}}_{1,G} +
 \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
-\hat{\var}(\hat{Y}_{G,psmall,3p})& = \hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_{s_2} +
-\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\var}(\hat{Y}_{G,PSMALL,3p})& = \hat{\pmb{\alpha}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_{s_2} +
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
 \hat{\bar{\pmb{Z}}}_{1,G} + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
-\end{flalign}
+\end{align}
 \end{subequations}
@@
 \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
-(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})^t
+(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})^{\top}
 \end{equation}
 
-The \textbf{external variance} is defined as:
+The external variance is defined as:
 
@@
 \begin{flalign}
-\hat{\var}_{ext}(\hat{Y}_{G,small,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) + 
+\hat{\var}_{ext}(\hat{Y}_{G,SMALL,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) +
  (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
-\hat{\var}_{ext}(\hat{Y}_{G,psmall,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) \nonumber &\\
+\hat{\var}_{ext}(\hat{Y}_{G,PSMALL,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) \nonumber &\\
  &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
@@
 \end{subequations}
-where $\hat{R}^{(0)}(x)=Y(x)-\hat{Y}^{(0)}(x)$ with $\hat{Y}^{(0)}(x)=\pmb{Z}^{(0)t}(x)\hat{\pmb{\alpha}}_{s_2}$.
+where $\hat{R}^{(0)}(x)=Y(x)-\hat{Y}^{(0)}(x)$ with $\hat{Y}^{(0)}(x)=\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}$.
 
@@
 
-The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) can be applied if no terrestrial samples are available in the small area, i.e.
-$n_{2,G}=0$.
+The synthetic (SYNTH) and pseudo synthetic estimator (PSYNTH) can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$.
 Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq.
@@
 \begin{align}
-\hat{Y}_{G,synth,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
-\hat{Y}_{G,psynth,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^t\hat{\pmb{\alpha}}_2 +
+\hat{Y}_{G,SYNTH,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
+\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
+\hat{Y}_{G,PSYNTH,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
 \hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
-\hat{\var}(\hat{Y}_{G,synth,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\var}(\hat{Y}_{G,SYNTH,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
 \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
-\hat{\var}(\hat{Y}_{G,psynth,3p})& = \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
-\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
+\hat{\var}(\hat{Y}_{G,PSYNTH,3p})& = \hat{\pmb{\alpha}}_2^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
+\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
 \hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
@@
 
-The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.$\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
+The extended synthetic (EXTSYNTH) and extended pseudo synthetic estimator (EXTPSYNTH) ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
 This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
-The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
-In other words, when the \textit{extended} option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e.
-$s_0$).
+The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{\mathbb{Z}}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$, where $\pmb{\mathbb{Z}}^{(0)\top}(x)=(\pmb{Z}^{(0)\top}(x), I_G^{\top}(x))$.
+In other words, when the extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for all observations in the input data frame (i.e., $s_0$).
 The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq.
@@
 
-% The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.
-$\int_F R(x)dx=\int_G R(x)dx=0$.
-This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
-The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$.
-This practically means that we have to specify the binary indicator variable for \textit{all} observations in the data frame that is passed to the package-functions.
-The regression coefficients $\hat{\pmb{\gamma}}_{s_2}$ and $\hat{\pmb{\theta}}_{s_2}$, and their variance-covarianve matrices $\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}$ under the extended auxiliary vector are thus recalculated as
-% 
-% \begin{subequations}\label{eq:normequ_redmod_ext}
-% \begin{align}
-% \hat{\pmb{\gamma}}_{s_2} &= \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)
-% \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}^{(0)}(x) \\
-% \hat{\pmb{\theta}}_{{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)
-% \Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathbb{Z}}(x) \\
-%   \hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1}
-%   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x)\Big)
-%   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}^{(0)}(x)\pmb{\mathbb{Z}}^{(0)t}(x) \Big)^{-1} \\
-%   \hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-%   \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x)\Big)
-%   \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1}
-% \end{align}
-% \end{subequations}
-% 
-% With the auxliary means and their variance-covariance matrices following the same notational change, the \textbf{point estimate} and \textbf{g-weight variance} are derived as
-% 
-% \begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
-% \begin{align}
-% \hat{Y}_{G,extsynth,3p}&=(\bar{\pmb{\mathbb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
-% \hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2} \label{eq:pe_3p_extsynth} \\
-% \hat{Y}_{G,extpsynth,3p}&=(\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}-\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{1,G})^t\hat{\pmb{\gamma}}_2 +
-% \hat{\bar{\pmb{\mathbb{Z}}}}^t_{1,G}\hat{\pmb{\theta}}_{s_2}\label{eq:pe_3p_extpsynth} \\
-% \hat{\var}(\hat{Y}_{G,extsynth,3p})& =\frac{n_2}{n_1}\bar{\pmb{\mathbb{Z}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
-% \bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
-% \hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extsynth}\\
-% \hat{\var}(\hat{Y}_{G,extpsynth,3p})& = \hat{\pmb{\gamma}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}}\hat{\pmb{\gamma}}_{s_2} + \frac{n_2}{n_1}\hat{\bar{\pmb{\mathbb{Z}}}}^{(0)t}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{s_2}}
-% \hat{\bar{\pmb{\mathbb{Z}}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathbb{Z}}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
-% \hat{\bar{\pmb{\mathbb{Z}}}}_{1,G} \label{eq:var_3p_reg_extpsynth}
-% \end{align}
-% \end{subequations}
-% 
-% 
-% The package also returns the \textbf{external variance} which is defined as
-% 
-% \begin{subequations}\label{eq:pest_3p_synth_psynth_ext}
-% \begin{flalign}
-% \hat{\var}_{ext}(\hat{Y}_{G,extsynth,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) +
-%  (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extsynth}&\\
-% \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}_{0}(x)) \nonumber & \\
-%  &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:var_3p_reg_extpsynth}
-% \end{flalign}
-% \end{subequations}
-% where $\hat{\mathbb{R}}_{0}=Y(x)-\hat{\mathbb{Y}}_{0}(x)$ with $\hat{\mathbb{Y}}_{0}(x)=\pmb{\mathbb{Z}}^{(0)t}(x)\hat{\pmb{\gamma}}_{s_2}$.
-The equations for the point estimate and variances again look similar to the synthetic and pseudo synthetic form given in Eq.
-\ref{eq:pe_3p_synth}, \ref{eq:pe_3p_psynth}, \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}.
-Just as in the two-phase context, they are design-unbiased since the residual correction term is included in the regression coefficient vector $\hat{\pmb{\theta}}_{s_2}$.
-
 %------------------------------------------------------------- %
@@
 
-We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the \textbf{extended synthetic} and the \textbf{synthetic} estimator to the \code{grisons} data set.
+We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the extended synthetic and the synthetic estimator to the \code{grisons} data set.
 The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}).
@@
 The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument.
-The \textbf{extended synthetic estimator} can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):
+The extended synthetic estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):
 
@@
 extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
-                  phase_id = list(phase.col = "phase_id_3p", 
-                                  s1.id = 1, terrgrid.id = 2),
-                  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                   unbiased = TRUE),
-                  exhaustive = truemeans.G, 
-                  boundary_weights = "boundary_weights")
+  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE),
+  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
 @
 \end{small}
-
-<<echo=FALSE>>=
-options(width=90) 
-@
 
@@
 
-The \textbf{synthetic estimator} can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.
+The synthetic estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.
 
@@
 synth_3p <- threephase(formula.rm, formula.fm, data = grisons,
-              phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
-              small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                               unbiased = FALSE),
-              exhaustive = truemeans.G,
-              boundary_weights = "boundary_weights")
+  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = FALSE),
+  exhaustive = truemeans.G, boundary_weights = "boundary_weights")
 @
@@
 We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$.
-To do this, we additionally compute the \textbf{extended pseudo synthetic} estimates.
+To do this, we additionally compute the extended pseudo synthetic estimates.
 As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.
@@
 extpsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
-                  phase_id = list(phase.col = "phase_id_3p", 
-                                  s1.id = 1, terrgrid.id = 2),
-                  small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                   unbiased = TRUE),
+  phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+  small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE), 
   boundary_weights = "boundary_weights")

### 2017-10-24 rnw/Twophase_Estimators and their Application.Rnw
formatting revision (1)

@@
 
-\section{Two-phase Estimators and their Application}
+% set global options for R-output:
+<<echo=FALSE>>=
+options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
+@
+
+\section{Two-phase estimators and their application}
 \label{sec:twophase_and_appl}
@@
 % ---------------------------------------------------------------------------- %
-\subsection{Global Estimators}
+\subsection{Global estimators}
 \label{sec:twophase_globest}
 
-\subsubsection{Mathematical Background}
+\subsubsection{Mathematical background}
 
@@
 \begin{equation}\label{normequ_simple}
-  \hat{\pmb{\beta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
+  \hat{\pmb{\beta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
 \end{equation}
 
-The individual predictions can then be calculated as $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
-Unless stated otherwise, \pkg{forestinventory} only uses \textit{internal models} to calculate estimates.
-This means that the model fit, i.e.
-$\hat{\pmb{\beta}}_{s_2}$, is derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} functions.
+The individual predictions can then be calculated as $\hat{Y}(x)=\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$.
+Unless stated otherwise, \pkg{forestinventory} only uses internal models to calculate estimates.
+This means that the model fit, i.e., $\hat{\pmb{\beta}}_{s_2}$, is derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} functions.
 While virtually all inventorists fit their models using the current inventory data, sometimes there is reason to use formulas derived from external models where the sample used to train the model is assumed to be taken from an independent source \citep{massey2015a}.
@@
 
-The package provides the calculation of \textbf{point estimates} under exhaustive ($ex$) and non-exhaustive ($nex$) use of the auxiliary information, which means to respectively apply $\hat{\pmb{\beta}}_{s_2}$ to $\bar{\pmb{Z}}$, i.e.
-the \textit{exact} spatial mean of $\pmb{Z}(x)$, or to $\hat{\bar{\pmb{Z}}}$, i.e.
-an \textit{estimate} of the spatial mean of $\pmb{Z}(x)$:
+The package provides the calculation of point estimates under exhaustive (EX) and non-exhaustive (NEX) use of the auxiliary information, which means to respectively apply $\hat{\pmb{\beta}}_{s_2}$ to $\bar{\pmb{Z}}$, i.e., the exact spatial mean of $\pmb{Z}(x)$, or to $\hat{\bar{\pmb{Z}}}$, i.e., an estimate of the spatial mean of $\pmb{Z}(x)$:
 
@@
 \begin{align}
-  \hat{Y}_{reg2p,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
-  \hat{Y}_{reg2p,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
+  \hat{Y}_{reg2p,EX} & =\bar{\pmb{Z}}^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
+  \hat{Y}_{reg2p,NEX} & =\hat{\bar{\pmb{Z}}}^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
 \end{align}
@@
 
-Note that for \textit{internal} linear models the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (\textit{zero mean residual property}) which is why it does not appear in the point estimate.
+Note that for internal linear models the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (zero mean residual property) which is why it does not appear in the point estimate.
 More explanation about how to obain the auxiliary means is given in the next subsection.
 
-The \pkg{forestinventory} package implements two kinds of variances for each of these point estimates: the \textit{g-weight} formulation that accounts for the fact that our model is in fact \textit{internal}, and the \textit{external variance} formulation that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients \citep{mandallaz2016}.
+The \pkg{forestinventory} package implements two kinds of variances for each of these point estimates: the g-weight formulation that accounts for the fact that our model is in fact internal, and the external variance formulation that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients \citep{mandallaz2016}.
 
-The \textit{g-weight} formulation is
+The g-weight formulation is
+
 \begin{subequations}\label{eq:gw_var_2p_reg}
 \begin{align}
-  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\bar{\pmb{Z}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}} \label{eq:gw_var_2p_reg_ex}\\
-  \hat{\var}(\hat{Y}_{reg2p,nex}) & :=
-  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
+  \hat{\var}(\hat{Y}_{reg2p,EX}) & :=\bar{\pmb{Z}}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}} \label{eq:gw_var_2p_reg_ex}\\
+  \hat{\var}(\hat{Y}_{reg2p,NEX}) & :=
+  \hat{\bar{\pmb{Z}}}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
+  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
 \end{align}
@@
 \begin{equation}\label{eq:estvarmatrix}
-  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
-  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
-  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
+  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
+  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^{\top}(x)\Big)
+  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
 \end{equation}
@@
 \frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})^t
+(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})^{\top}
 \end{equation}
 
-The \textit{external variance} formulation for linear regression models is
+The external variance formulation for linear regression models is
 
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:varexternal_2p_reg_ex} \\
-  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}_{s_1}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \nonumber  \label{eq:varexternal_2p_reg_nex}
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,EX}) & = \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:varexternal_2p_reg_ex} \\
+  \hat{\var}_{ext}(\hat{Y}_{reg2p,NEX}) & = \frac{1}{n_1}\hat{\var}_{s_1}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \nonumber  \label{eq:varexternal_2p_reg_nex}
 \end{align}
@@
 
-\subsubsection{Calculation of Explanatory Variables}
+\subsubsection{Calculation of explanatory variables}
 
@@
 Fig.
-\ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles) and densified to a larger sample $s_1$ (the points).
-For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles).
+\ref{fig:exh_nexh_and_boundweights_b} depicts how the non-exhaustive case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles) and densified to a larger sample $s_1$ (the points).
+For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ is calculated using a defined spatial extent of auxiliary information around that point called the support (the dark green square tiles).
 We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the sample point whereas the support is the spatial extent of the auxiliary information used to calculate those values.
@@
 
-The \textbf{exhaustive} case requires a closer look.
+The exhaustive case requires a closer look.
 In the infinite population approach, $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it.
-Deriving the \textit{exact} spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^t$, implies that we need to calculate the spatial mean of each component of $\pmb{Z}(x)$ using all possible points in $F$.
-This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e.
-$n_1=\infty$).
+Deriving the exact spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^{\top}$, implies that we need to calculate the spatial mean of each component of $\pmb{Z}(x)$ using all possible points in $F$.
+This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e., $n_1=\infty$).
 Although it is practically infeasible to assess $\pmb{Z}(x)$ for every $x$, there are few cases where the exact mean can in fact be precisely calculated.
-The first case is when the explanatory variables are provided by polygon layers (e.g.
-map of development stages).
+The first case is when the explanatory variables are provided by polygon layers (e.g., map of development stages).
 In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
-The second case is when the exact mean can be calculated in one step, e.g.
-taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+The second case is when the exact mean can be calculated in one step, e.g., taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
 However, for most types of explanatory variables we will try to get an approximation of $\bar{\pmb{Z}}$ that is only negligibly different.
@@
 \ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited.
-Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e.
-deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units.
+Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the finite population approach, i.e., deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units.
 While in the infinite population approach this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large.
@@
 	\end{subfigure}
-\caption{Concept of \textbf{(a)} exhaustive and \textbf{(b)} non-exhaustive calculation of explanatory variables including boundary adjustment at the support level.
+\caption{Concept of (a) exhaustive and (b) non-exhaustive calculation of explanatory variables including boundary adjustment at the support level.
 Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle.
@@
 
-An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}}.
+\subsubsection{Boundary adjustment}
+
+An extension to the so-far published estimators by Mandallaz is the consideration of a boundary adjustment.
 In forest inventories, the sample is often restricted to those sample locations located within the forest area.
-In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g.
-by a polygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (see Fig.
+In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g., by a polygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (see Fig.
 \ref{fig:exh_nexh_bw}).
 This method was suggested in \citet{mandallaz2013b} and led to an improvement in estimation precision.
-In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the \textit{weighted} mean (Eq.
+In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the weighted mean (Eq.
 \ref{eq:wmeanZ}) where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.
@@
 To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository.
-The data set contains data from a simple (i.e.
-non-cluster) two-phase forest inventory conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
-The $s_1$ sample is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
+The data set contains data from a simple (i.e., non-cluster) two-phase forest inventory conducted in 2007 that was used in \citet{mandallaz2013b} as a case study.
+The $s_1$ sample is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of airborne laserscanning (LiDAR) canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}).
 For a systematic subsample of 67 ($s_2$ sample), terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey.
@@
 
-<<echo=FALSE>>=
-options(width=10000) 
-@
-
 \begin{small}
 <<eval=FALSE>>=
-library(forestinventory)
+library("forestinventory")
 data("grisons", package = "forestinventory")
@@
 <<eval=TRUE, echo=FALSE>>=
-library(forestinventory)
-set.seed(666);print(head(grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)], 10),row.names=c(1:10))
+library("forestinventory")
+set.seed(666)
+g.temp<- grisons[sample(1:306,10),c(1,3,4,5,6,7,8,9)]
+ g.temp[,c(2,3,4,5,6,8)]<- round(g.temp[,c(2,3,4,5,6,8)], digits = 2)
+print(head(g.temp, n = 10), row.names = seq(1:10))
+rm(g.temp)
 @
@@
 The data frame inputted to these functions must have the structure where each row corresponds to a unique sample location and the columns specify the attributes associated to that respective sample location.
-Attributes that are missing, e.g.
-because they are associated with sample locations that were not selected in the subsample for the subsequent phase, should be designated as \code{NA} and the phase membership is encoded as numeric.
+Attributes that are missing, e.g., because they are associated with sample locations that were not selected in the subsample for the subsequent phase, should be designated as \code{NA} and the phase membership is encoded as numeric.
 
@@
   \itemsep0em
-  \item the regression model (\code{formula}) as specified in the \code{lm}()-function \citep{R}
-  \item the inputted \code{data.frame} containing the inventory information (\code{data})
-  \item the \code{list}-object \code{phase_id} containing: the \code{phase.col} argument identifying the name of the column specifying membership to $s_1$ or $s_2$, and the \code{terrgrid.id} argument specifying which numeric value indicates $s_2$ membership in that column
-  \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional)
+  \item the regression model (\code{formula}) as specified in the \code{lm}()-function \citep{R}.
+  \item the inputted \code{data.frame} containing the inventory information (\code{data}).
+  \item the \code{list}-object \code{phase_id} containing: the \code{phase.col} argument identifying the name of the column specifying membership to $s_1$ or $s_2$, and the \code{terrgrid.id} argument specifying which numeric value indicates $s_2$ membership in that column.
+  \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional).
 \end{itemize}
 
-The \textbf{non-exhaustive estimator with boundary weight adjustment} can thus be applied as follows:
+The non-exhaustive estimator with boundary weight adjustment can thus be applied as follows:
 
 \begin{small}
-<<>>=
-reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
-                   data=grisons,
+<<width=30>>=
+reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75, data=grisons, 
   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
@@
 The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}.
-A readable \textbf{summary of the estimation results} can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the g-weight and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:
+A readable summary of the estimation results can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the g-weight and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:
 
@@
 
+% \begin{Schunk}
+% \begin{Sinput}
+% Two-Phase global estimation
+%  
+% Call: 
+% twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+%     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2), 
+%     boundary_weights = "boundary_weights")
+% 
+% Method used:
+% Non-exhaustive global estimator
+%  
+% Regression Model:
+% tvol ~ mean + stddev + max + q75
+% 
+% Estimation results:
+%  estimate ext_variance g_variance  n1 n2 r.squared
+%  383.5354      279.954   271.5057 306 67 0.6428771
+% 
+% 'boundary_weight'- option was used to calculate weighted means of auxiliary variables
+% \end{Sinput}
+% \end{Schunk}
+
 For practical use, one should normally always prefer the g-weight variance over the external variance.
@@
 
-The \textbf{exhaustive estimator} can be applied by additionally passing a vector containing the \textit{exact} means of the explanatory variables, i.e.
-$\bar{\pmb{Z}}$, to the optional argument \code{exhaustive}.
+The exhaustive estimator can be applied by additionally passing a vector containing the exact means of the explanatory variables, i.e., $\bar{\pmb{Z}}$, to the optional argument \code{exhaustive}.
 This vector must be calculated beforehand in such a way that any desired boundary adjustment has already been applied.
@@
 
-The \textbf{exhaustive estimator} can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b}, denoted as \code{true.means.Z}:
+The exhaustive estimator can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b}, denoted as \code{true.means.Z}:
 
@@
 
-\subsection{Small Area Estimators}
+\subsection{Small area estimators}
 \label{sec:twophase_sae}
 
-\subsubsection{Mathematical Background}
+\subsubsection{Mathematical background}
 
 The \pkg{forestinventory} package provides three types of small area estimators each of which has an exhaustive and non-exhaustive form.
-We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label \textit{pseudo} to indicate that the mean of the explanatory variables within the small area was based on a finite sample.
+We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label pseudo to indicate that the mean of the explanatory variables within the small area was based on a finite sample.
 The main idea for all these small area estimators is to calculate the regression coefficient vector $\hat{\pmb{\beta}}_{s_2}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ on the entire $s_2$ sample according to Eq.
@@
 
-We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}).
+We first introduce the small area estimator (SMALL), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the pseudo small area estimator (PSMALL).
 
@@
 \begin{align}
-  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
-  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pepsmall}
+  \hat{Y}_{G,SMALL,2p} & =\bar{\pmb{Z}}_G^{\top}\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
+  \hat{Y}_{G,PSMALL,2p} & =\hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pepsmall}
 \end{align}
@@
 A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data also outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$.
-This is called the \textit{bias} or \textit{residual correction} term.
+This is called the bias or residual correction term.
 
-The package provides the \textit{g-weight} variance for \textit{small} and \textit{psmall} respectively (Eq.
-\ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  \textit{external variance} (Eq.
+The package provides the g-weight variance for SMALL and PSMALL respectively (Eq.
+\ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  external variance (Eq.
 \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}).
@@
 
-% \begin{subequations}\label{eq:var_2p_small_psmall}
-% \begin{align}
-%   \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
-%     + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
-%   \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-%   + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
-%   + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
-% \end{align}
-% \end{subequations}
-
 \begin{subequations}\label{eq:var_2p_small_psmall}
 \begin{align}
-  \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
+  \hat{\var}(\hat{Y}_{G,SMALL,2p}) & := \bar{\pmb{Z}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
     + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_small} \\
-  \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
+  \hat{\var}(\hat{Y}_{G,PSMALL,2p}) & := \hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
+  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
   + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_psmall}
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
+  \hat{\var}_{ext}(\hat{Y}_{G,SMALL,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,PSMALL,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
 \end{align}
@@
 
-The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are commonly applied when no terrestrial sample is available within the small area $G$ (i.e.
-$n_{2,G}=0$).
+The synthetic estimator (SYNTH) and pseudo synthetic estimator (PSYNTH) are commonly applied when no terrestrial sample is available within the small area $G$ (i.e., $n_{2,G}=0$).
 In this case, the point estimates (Eq.
-\ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
+\ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based only on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively.
 However, the bias correction using the observed residuals $\hat{R}(x)$ is not applied as was the case in the small and pseudo small area estimator (Eq.
@@
 \begin{align}
-  \hat{Y}_{G,synth,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
-  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth} \\
-  \hat{\var}(\hat{Y}_{G,synth,2p}) & = \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
-  \hat{\var}(\hat{Y}_{G,psynth,2p}) & = \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
+  \hat{Y}_{G,SYNTH,2p} & =\bar{\pmb{Z}}_G^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
+  \hat{Y}_{G,PSYNTH,2p} & =\hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth} \\
+  \hat{\var}(\hat{Y}_{G,SYNTH,2p}) & = \bar{\pmb{Z}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
+  \hat{\var}(\hat{Y}_{G,PSYNTH,2p}) & = \hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
+  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
 \end{align}
@@
 \frac{1}{n_{1,G}(n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
-(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
+(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^{\top}
 \end{equation}
@@
 
-The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively optimistic.
-The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e.
-$\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
-The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
+The synthetic estimators, SYNTH and PSYNTH, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively optimistic.
+The SMALL and PSMALL estimators overcome this issue by using a bias correction term, i.e., $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
+The motivation behind the extended synthetic andextended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
 Recalling that linear models fitted using OLS have zero mean residual property by construction also if categorical variables are used, this leads to unbiased point estimates.
-The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq.
+The new extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{Z}^{\top}(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^{\top}(x)$ whereever it is used in Eq.
 \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
-Note that the package functions internally extend the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
+Note that the package functions internally extend the data set by the indicator variable if the EXTSYNTH or EXTPSYNTH estimator is called.
 
@@
 \begin{equation}\label{ext_normequ_simple}
-  \hat{\pmb{\theta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
+  \hat{\pmb{\theta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^{\top}(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
 \end{equation}
@@
 \begin{align}
-\hat{Y}_{G,extsynth,2p} & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
-\hat{Y}_{G,extpsynth,2p} & =\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
-\hat{\var}(\hat{Y}_{G,extsynth,2p}) & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
-\hat{\var}(\hat{Y}_{G,extpsynth,2p})& =
-\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
+\hat{Y}_{G,EXTSYNTH,2p} & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{Y}_{G,EXTPSYNTH,2p} & =\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^{\top}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,EXTSYNTH,2p}) & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,EXTPSYNTH,2p})& =
+\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
 \hat{\bar{\pmb{\mathbb{Z}}}}_{G}
@@
 
-The package also provides the external variance for both the \textit{extended} synthetic and \textit{extended} pseudo synthetic estimator.
+The package also provides the external variance for both the extended synthetic and extended pseudo synthetic estimator.
 Note that neither the extended model approach nor external variance estimates are possible in the absence of terrestrial samples and thus model residuals in $G$, which is precisely when one must rely on the (pseudo) synthetic estimates.
-The external variance forms of \textit{extsynth} and \textit{extpsynth} are
+The external variance forms of EXTSYNTH and EXTPSYNTH are
 
@@
 \begin{align}
-  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extsynth} \\
-  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & = \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extpsynth}
+  \hat{\var}_{ext}(\hat{Y}_{G,EXTSYNTH,2p}) & = \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extsynth} \\
+  \hat{\var}_{ext}(\hat{Y}_{G,EXTPSYNTH,2p}) & = \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extpsynth}
 \end{align}
@@
 To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
-When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either \textit{small} or \textit{psmall} which remove this bias explicitly with a mean residual term, or more elegantly with \textit{extsynth} or \textit{extpsynth} which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
+When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either SMALL or PSMALL which remove this bias explicitly with a mean residual term, or more elegantly with EXTSYNTH or EXTPSYNTH which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.
 
@@
 
-In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
+In order to apply the pseudo small area estimator (PSMALL) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
 \begin{small}
@@
   psmall = TRUE)
+@
+\end{small}
+
+\begin{small}
+<<>>=
 summary(psmall_2p)
@@
 
-The \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
+The extended pseudo synthetic estimator (EXTPSYNTH) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
+
 \begin{small}
-<<>>=
+<<width=77>>=
 extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
@@
 The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need for the user to implement it.
-Notice that the $R^2$s (\code{r.squared}) under the \textit{extpsynth} estimator vary between the small areas, while they are identical under the \textit{psmall} estimator.
-This is because under the \textit{extpsynth} estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrix.
-In case of the  \textit{psmall} estimator, the regression model stays the same for each small area estimation.
+Notice that the $R^2$s (\code{r.squared}) under the EXTPSYNTH estimator vary between the small areas, while they are identical under the PSMALL estimator.
+This is because under the EXTPSYNTH estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrix.
+In case of the PSMALL estimator, the regression model stays the same for each small area estimation.
 Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling (see Section \ref{sec:speccas_and_scen}).\par
 
-Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas.
+Setting the argument \code{unbiased=FALSE} applies the pseudo synthetic estimator to the selected small areas.
 Note that in the \code{grisons} data set, all small areas possess much more than the suggested minimum number of terrestrial observations (a rule of thumb is that $n_{2,G} \geq 6$) required to produce reliable design-unbiased estimates.
-Hence, choosing to use \textit{psynth} is probably not desireable and is just applied here for demonstration purposes.
+Hence, choosing to use PSYNTH is probably not desireable and is just applied here for demonstration purposes.
 In this case the residual correction will not be applied.
@@
 
-We see here that the \textit{psynth} variances are almost only half the variances of the \textit{psmall} and \textit{extended psynth} estimators.
-However, \textit{psmall} and \textit{extended psynth} are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions.
-The g-weight variance of \textit{psynth} completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
+We see here that the PSYNTH variances are almost only half the variances of the PSMALL and extended psynth estimators.
+However, PSMALL and extended psynth are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions.
+The g-weight variance of PSYNTH completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
 
-The \textbf{exhaustive versions} of the small area estimators (Eq.
+The exhaustive versions of the small area estimators (Eq.
 \ref{eq:pointest_2p_small}, \ref{eq:var_2p_reg_small}, \ref{eq:varext_2p_reg_small}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}.
 Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest.
-In contrast to the \textit{global} estimators, the exact means have now to be delivered in the form of a \code{data.frame}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
+In contrast to the global estimators, the exact means have now to be delivered in the form of a \code{data.frame}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable.
 Note that likewise the case of global estimation, the order of the explanatory variables in the data frame has to match the order in which they appear in the design matrix defined by the \code{lm()}-function in \proglang{R}.
@@
 
-The \textbf{extended synthetic estimator} (\textit{extsynth}) can then be applied by
+The extended synthetic estimator (EXTSYNTH) can then be applied by
 \begin{small}
@@
   exhaustive = true.means.Z.G)
+@
+\end{small}
+
+\begin{small}
+<<>>=
 extsynth_2p$estimation
@@
 
-Just as in the global case, we see that the variance has again been significantly decreased by substituting the \textit{exact} auxiliary means and both first phase sample sizes are now infinity.
+Just as in the global case, we see that the variance has again been significantly decreased by substituting the exact auxiliary means and both first phase sample sizes are now infinity.
 Note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.

### 2017-10-24 r/analysis_and_visualization.R

@@ -1,96 +1,0 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 7  "Analysis and Visualization"                                          #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
-
-
-library(forestinventory)
-
-## calculate some estimations (one-phase, two-phase, three-phase)
-
-op <- onephase(formula = tvol~1, data = grisons,
-               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-               area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))
-
-extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                        small_area = list(sa.col = "smallarea", 
-                                          areas = c("A", "B","C", "D"),
-                                          unbiased = TRUE))
-
-psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights",
-                      small_area = list(sa.col = "smallarea", 
-                                        areas = c("A", "B", "C", "D"),
-                                        unbiased = FALSE))
- 
-extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                          formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                          phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                          small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                          unbiased = TRUE),
-                          boundary_weights = "boundary_weights")
-
-psynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                       formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                       phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                       small_area = list(sa.col = "smallarea", 
[…]

### 2017-10-24 r/confidence_intervals.R

@@ -1,60 +1,0 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 5  "Calculation of Confidence Intervals"                                 #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-
-#####################################################
-# ----- CI for One-Phase Estimation --------------- #
-#####################################################
-
-## load grisons dataset:
-data(grisons)
-
-## calculate onephase-estimation for entire dataset:
-op <- onephase(formula = tvol~1 ,data = grisons,
-               phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))
-
-## Confidence Intervals:
-confint(op)
-
-## 2) calculate one-phase-estimation for given domains (areas) in dataset:
-op.a <- onephase(formula = tvol~1,
-                 data = grisons,
-                 phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                 area = list(sa.col = "smallarea", areas = c("A", "B")))
-
-## Confidence Intervals:
-confint(op.a)
-
-
-#####################################################
-# ----- CI for Two-Phase Estimation --------------- #
-#####################################################
-
-## calculate two-phase-estimation for entire dataset:
-extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                        small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
-                                          unbiased = TRUE))
-## Confidence Intervals:
-confint(extpsynth_2p)
-
-## change confidence level:
-confint(extpsynth_2p, level = 0.9)
-
-## CIs with bonferroni correction to better meet overall coverage probability
-#  (multiple testing problem if a large number of CIs are compared to each other)
-confint(extpsynth_2p, adjust.method = "bonferroni")
-
-
-#######
-# END #
-#######

### 2017-10-24 r/create figures.R

@@ -1,76 +1,0 @@
-
-library(forestinventory)
-library(ggplot2)
-
-op <- onephase(formula = tvol~1, data = grisons,
-               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-               area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))
-
-
-extpsynth_2p<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                        small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
-                                          unbiased = TRUE))
-
-
-psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights",
-                      small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                        unbiased = FALSE))
-
-
-extpsynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                         formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                         phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                         small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
-                                         unbiased = TRUE),
-                         boundary_weights = "boundary_weights")
-
-
-psynth_3p<- threephase(formula.s0 = tvol ~ mean,
-                      formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
-                      phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
-                      small_area = list(sa.col = "smallarea", 
-                                        areas = c("A", "B", "C", "D"),
-                                        unbiased = FALSE),
-                      boundary_weights = "boundary_weights")
-          
-
-grisons.sae.table<- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
-                                                extpsynth_3p, psynth_3p),
-                                sae = TRUE, 
-                                vartypes = c("variance", "g_variance",  "ext_variance"))
-
-
-
-plot(grisons.sae.table, ncol = 2, yvar = "estimate")
-
-## save graphic for article:
[…]

### 2017-10-24 r/replication_rcode.R

@@ -1,0 +1,589 @@
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples                                                     #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
+
+library("forestinventory")
+
+# **************************************************************************** #
+# ------ SECTION 3: TWO-PHASE ESTIMATORS -------------------------------------
+# **************************************************************************** #
+
+
+# ----- 3.1 Global Estimations --------
+
+
+## load grisons data set:
+data("grisons", package = "forestinventory")
+
+## see structure of grisons:
+str(grisons)
+
+## see first 12 rows of grisons:
+head(grisons, n = 12)
+
+
+# **********************************************************
+## non-exhaustive (NEX) estimator with boundary weight adjustment
+reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                      data = grisons,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights")
+
+## get summary of estimation results:
+summary(reg2p_nex)
+
+
+# *****************************************************
+## exhaustive (EX) estimator with boundary weight adjustment
+## first, retrieve order of columnames that the lm()-function processes a formula object:
+#  --> actually the order they appear in the design-matrix ...
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+
+## store true means of explanatory variables:
+true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
+
+## apply estimator:
+reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                     data = grisons,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                     exhaustive = true.means.Z)
+
+## get summary of estimation results:
+summary(reg2p_ex)
+
+## get data frame with estimation results
+reg2p_ex$estimation
+
[…]

### 2017-10-24 r/special_cases_and_scenarios.R

@@ -1,319 +1,0 @@
-#---------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                  #
-#  Design-Based Global and Small Area Estimations for Multiphase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                  #
-#                                                                                   #
-#  Code for running all examples given in                                           #
-#  Section 6  "Special Cases and Scenarios"                                         #
-#                                                                                   #
-#---------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-#####################################################
-#          Poststratification                       #
-#####################################################
-
-## save grisons as new dataset:
-grisons.n<- grisons
-
-## create artificial development stages as categorical (factor) variables
-#  used for poststratification:
-grisons.n$stage<- as.factor(kmeans(grisons.n$mean, centers = 3)$cluster)
-
-## plot:
-library(ggplot2)
-ggplot(data=grisons.n, aes(x = mean, y = tvol)) +
-  geom_point(aes(colour = factor(stage)))
-
-
-## apply 'double sampling for (post)stratification':
-summary(
-  twophase(formula=tvol ~ stage,
-           data=grisons.n,
-           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-           boundary_weights = "boundary_weights")
-)
-
-
-## apply 'double sampling for regression':
-summary(
-  twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-           data=grisons.n,
-           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-           boundary_weights = "boundary_weights")
-)
-
-## apply 'double sampling for regression within (post)strata'":
-summary(
-  twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
-           data=grisons.n,
-           phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-           boundary_weights = "boundary_weights")
-)
-
-
-#####################################################
-#  SMALL AREA ESTIMATION UNDER CLUSTER SAMPLING     #
-#####################################################
-
-
-## apply extended pseudo synthetic estimator under cluster sampling:
[…]

### 2017-10-24 r/three_phase_estimators.R

@@ -1,106 +1,0 @@
-#----------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                   #
-#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                   #
-#                                                                                    #
-#  Code for running all examples given in                                            #
-#  Section 4  "Three-phase Estimators and their Application"                         #
-#                                                                                    #
-#----------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-
-#####################################################
-# -------- Global Estimations --------------------- #
-#####################################################
-
-
-## load grisons data set:
-data("grisons", package = "forestinventory")
-
-## see structure of grisons:
-str(grisons)
-
-## see first 12 rows of grisons:
-head(grisons, n = 12)
-
-## define reduced and full regression model formulas:
-formula.rm <- tvol ~ mean # reduced model applied to s0 phase
-formula.fm <- tvol ~ mean + stddev + max + q75 # full model applied to s1 phase
-
-## apply estimator:
-reg3p_nex<- threephase(formula.s0 = formula.rm, 
-                       formula.s1 = formula.fm, data = grisons,
-                       phase_id = list(phase.col="phase_id_3p", s1.id = 1, 
-                                       terrgrid.id = 2),
-                       boundary_weights = "boundary_weights")
-
-## get summary of estimation results:
-summary(reg3p_nex)
-
-
-#####################################################
-# ------ Small Area Estimations ------------------- #
-#####################################################
-
-
-# --------------------------------- #
-## extended synthetic estimator (extsynth)
-#  with boundary weight adjustment
-
-## store true means of explanatory variables of reduced-model in a data frame:
-truemeans.G <- data.frame(Intercept = rep(1, 4),
-                          mean = c(12.85, 12.21, 9.33, 10.45))
-rownames(truemeans.G) <- c("A", "B", "C", "D")
-
-## apply estimator:
-extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
-                          phase_id = list(phase.col = "phase_id_3p", 
-                                          s1.id = 1, terrgrid.id = 2),
-                          small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-                                            unbiased = TRUE),
-                          exhaustive = truemeans.G, 
[…]

### 2017-10-24 r/two_phase_estimators.R

@@ -1,151 +1,0 @@
-#----------------------------------------------------------------------------------- #
-#  Journal of Statistical Software                                                   #
-#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
-#  Andreas Hill & Alexander Massey                                                   #
-#                                                                                    #
-#  Code for running all examples given in                                            #
-#  Section 3  "Two-phase Estimators and their Application"                           #
-#                                                                                    #
-#----------------------------------------------------------------------------------- #
-
-library(forestinventory)
-
-#####################################################
-# -------- Global Estimations --------------------- #
-#####################################################
-
-
-## load grisons data set:
-data("grisons", package = "forestinventory")
-
-## see structure of grisons:
-str(grisons)
-
-## see first 12 rows of grisons:
-head(grisons, n = 12)
-
-
-# --------------------------------- #
-## non-exhaustive estimator 
-#  with boundary weight adjustment
-
-## apply estimator:
-reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
-                      data=grisons,
-                      phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                      boundary_weights = "boundary_weights")
-
-## get summary of estimation results:
-summary(reg2p_nex)
-
-
-
-# --------------------------------- #
-## exhaustive estimator 
-#  with boundary weight adjustment
-
-## retrieve order of columnames that the lm()-function processes a formula object:
-#  --> actually the order they appear in the design-matrix ...
-colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
-
-## store true means of explanatory variables:
-true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
-
-## apply estimator:
-reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
-                     data = grisons,
-                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-                     exhaustive = true.means.Z)
-
-## get summary of estimation results:
-summary(reg2p_ex)
-
-## get data frame with estimation results
-reg2p_ex$estimation
-
-
-
-#####################################################
-# ------ Small Area Estimations ------------------- #
[…]

### 2017-10-25 JStatSoft_forestinventory_master.rnw
formatting revision (2)

@@
 While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
+The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling.
 The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
@@
 \Plainkeywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
-
- % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
- % ... free --> more suited for forest practice (limited financial ressources)
 
@@
 
-\setlength{\parindent}{1em} % Einzug bei neuen Absätzen
+\setlength{\parindent}{0em} % Einzug bei neuen Absätzen
 
@@
 <<echo=FALSE>>=
-library(forestinventory)
+library("forestinventory")
 @

### 2017-10-25 rnw/Analysis and Visualization.Rnw
formatting revision (2)

@@
 This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis.
-For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase extended pseudo synthetic and the pseudo synthetic estimators for the \code{grisons} data set:
+For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase EXTPSYNTH and PSYNTH estimator for the \code{grisons} data set:
 

### 2017-10-25 rnw/Calculation of Confidence Intervals.Rnw
formatting revision (2)

@@
 Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples and calculate the 95\%  CI from each sample, then 95\% of the CIs are expected to contain the true population parameter.
-The confidence level $1-\alpha$ (e.g.
-95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate.
+The confidence level $1-\alpha$ (e.g., 95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate.
 The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries.
@@
 In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
-For example, the 95\% confidence interval for the small area estimates by the extpsynth estimator (Section \ref{sec:twophase_sae}) are calculated by:
+For example, the 95\% confidence interval for the small area estimates by the EXTPSYNTH estimator (Section \ref{sec:twophase_sae}) are calculated by:
 

### 2017-10-25 rnw/Introduction.Rnw
formatting revision (2)

@@
 
-% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
-Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare.
+% section: briefly give a review of existing R-packages for forest inventories (not much around)
+While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}.
+However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available.
+The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multi-phase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
+
+The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
+The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
+The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}.
+The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
+The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.
+
+% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.
@@
 
-The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.
+% Our thoughts:
+ % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
+ % ... free --> more suited for forest practice (limited financial ressources)
+ % ... not up to date: don't provide an extensive set of estimation methods (possibilities)
+ % ... lacks the possibility to choose 
+%  
+% I'll try to keep it very general by emphasizing the pros of open-source software like ours
+% that is: 1) full transparency of the methods, 2) free access also for (small) forest
+% enterprises who maybe cannot afford commercial services, and 3) providing a huge set of estimators in a open source language that provides high flexibility to be incorporated in operational database system of forest services
+% do you think that good?
 
 %\newpage
+
+%% NOTE on commercial software:
+%
+% - SILVIATERRA: they call it "CruiseBoost" (https://silviaterra.com/bark/cruiseboost.html) --> means thinning the terrestrial inventory and
+%                using remote sensing data instead , i.e. two-phase procedures. They say they are "model-assisted estimators". 
+%                Costs "a low, per acre fee". 
+%                Slogans "We guarantee that CruiseBoost will tighten up your confidence intervals or your money back". and 
+%                "Your CruiseBoost estimate cannot be worse than the underlying traditional grid cruise" --> which is not true as we know
+%
+% - ForestMetrix: 
+%
+%
+% Our thoughts:
+ % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
+ % ... free --> more suited for forest practice (limited financial ressources)
+ % ... not up to date: don't provide an extensive set of estimation methods (possibilities)
+ % ... lacks the possibility to choose

### 2017-10-25 rnw/Threephase_Estimators and their Application.Rnw
formatting revision (2)

@@
 \nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2} +
-\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
+\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
 \hat{Y}_{reg3p,NEX}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
 \nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2}  +
-\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
+\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
 \end{align}
@@
 
-The synthetic (SYNTH) and pseudo synthetic estimator (PSYNTH) can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$.
+The synthetic (SYNTH) and pseudo synthetic (PSYNTH) estimator can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$.
 Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq.
@@
 \hat{Y}_{G,PSYNTH,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
-\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
+\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
 \hat{\var}(\hat{Y}_{G,SYNTH,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
@@
 
-The extended synthetic (EXTSYNTH) and extended pseudo synthetic estimator (EXTPSYNTH) ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
+The extended synthetic (EXTSYNTH) and extended pseudo synthetic (EXTPSYNTH) estimator ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$.
 This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0.
@@
 
-We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the extended synthetic and the synthetic estimator to the \code{grisons} data set.
+We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the EXTSYNTH and SYNTH estimator to the \code{grisons} data set.
 The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}).
@@
 The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument.
-The extended synthetic estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):
+The EXTSYNTH estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):
 
@@
 
-The synthetic estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.
+The SYNTH estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.
 
@@
 The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively.
-If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but at the cost of a potential bias.
+If we compare the two results, we see that the SYNTH estimation again yields a much smaller variance than the EXTSYNTH estimation, but at the cost of a potential bias.
 
 We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$.
-To do this, we additionally compute the extended pseudo synthetic estimates.
+To do this, we additionally compute the EXTPSYNTH estimates.
 As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.

### 2017-10-25 rnw/Twophase_Estimators and their Application.Rnw
formatting revision (2)

@@
 \begin{small}
-<<width=30>>=
-reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75, data=grisons, 
-  phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
+<<>>=
+reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   boundary_weights = "boundary_weights")
@@
 <<eval=FALSE>>=
-colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+  x = TRUE)$x)
 @
@@
 Recalling that linear models fitted using OLS have zero mean residual property by construction also if categorical variables are used, this leads to unbiased point estimates.
-The new extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{Z}^{\top}(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^{\top}(x)$ whereever it is used in Eq.
+The new extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{Z}^{\top}(x),I^{\top}_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^{\top}(x)$ whereever it is used in Eq.
 \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}.
@@
 \hat{\bar{\pmb{\mathbb{Z}}}}_{G}
-+ \hat{\pmb{\theta}}^t_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
++ \hat{\pmb{\theta}}^{\top}_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
 \end{align}
@@
   \itemsep0em
-  \item the name of the column specifiying the small area of each observation (\code{sa.col})
-  \item a vector specifying the small area(s) for which estimations are desired (\code{areas})
-  \item the argument \code{unbiased} that controls which of the three available estimators is applied
+  \item the name of the column specifiying the small area of each observation (\code{sa.col}).
+  \item a vector specifying the small area(s) for which estimations are desired (\code{areas}).
+  \item the argument \code{unbiased} that controls which of the three available estimators is applied.
 \end{itemize}
@@
 <<>>=
-psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-  boundary_weights = "boundary_weights",
+psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-  unbiased = TRUE),
-  psmall = TRUE)
+  unbiased = TRUE), psmall = TRUE, boundary_weights = "boundary_weights")
 @
@@
 \begin{small}
-<<width=77>>=
-extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-  boundary_weights = "boundary_weights",
+<<>>=
+extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-  unbiased = TRUE))
+  unbiased = TRUE), boundary_weights = "boundary_weights")
 extpsynth_2p$estimation
@@
 <<>>=
-psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
-  boundary_weights = "boundary_weights",
+psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   small_area = list(sa.col = "smallarea", areas = c("A", "B"),
-  unbiased = FALSE))
+  unbiased = FALSE), boundary_weights = "boundary_weights")
+@
+\end{small}
+
+\begin{small}
+<<>>=
 psynth_2p$estimation
@@
 
-We see here that the PSYNTH variances are almost only half the variances of the PSMALL and extended psynth estimators.
-However, PSMALL and extended psynth are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions.
+We see here that the PSYNTH variances are almost only half the variances of the PSMALL and EXTPSYNTH estimator.
+However, PSMALL and EXTPSYNTH are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions.
 The g-weight variance of PSYNTH completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par
@@
 <<>>=
-extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
-  phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+  data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
   small_area = list(sa.col ="smallarea", areas = c("A", "B"),
-  unbiased = TRUE),
-  exhaustive = true.means.Z.G)
+  unbiased = TRUE), exhaustive = true.means.Z.G)
 @

### 2017-10-25 rnw/Analysis and Visualization.Rnw
formatting revision (3)

@@
 
-\begin{footnotesize}
-<<echo=FALSE>>=
-str(grisons.sae.table)
-@
-\end{footnotesize}
+\begin{small}
+\begin{Schunk}
+\begin{Sinput}
+List of 20
+ $ area             : chr [1:28] "A" "A" "A" "A" ...
+ $ domain           : Factor w/ 2 levels "global","smallarea": 1 2 2 2 2 2 2 1 2  ...
+ $ method           : Factor w/ 3 levels "onephase","twophase",..: 1 3 3 3 2 2 2  ...
+ $ estimator        : Factor w/ 3 levels "onephase","psynth extended",..: 1 2 2 3 ...
+ $ vartype          : Factor w/ 3 levels "ext_variance",..: 3 1 2 2 1 2 2 3 1 2 ...
+ $ estimate         : num [1:28] 410 395 395 422 392 ...
+ $ variance         : num [1:28] 1987 1901 1858 726 996 ...
+ $ std              : num [1:28] 44.6 43.6 43.1 26.9 31.6 ...
+ $ error            : num [1:28] 10.86 11.03 10.91 6.39 8.05 ...
+ $ n2               : num [1:28] 19 40 40 40 67 67 67 17 40 40 ...
+ $ n2G              : num [1:28] NA 12 12 12 19 19 19 NA 11 11 ...
+ $ n1               : num [1:28] NA 128 128 128 306 306 306 NA 128 128 ...
+ $ n1G              : num [1:28] NA 38 38 38 94 94 94 NA 34 34 ...
+ $ n0               : int [1:28] NA 306 306 306 NA NA NA NA 306 306 ...
+ $ n0G              : int [1:28] NA 94 94 94 NA NA NA NA 81 81 ...
+ $ r.squared        : num [1:28] NA NA NA NA 0.653 ...
+ $ r.squared_reduced: num [1:28] NA 0.545 0.545 0.527 NA ...
+ $ r.squared_full   : num [1:28] NA 0.724 0.724 0.717 NA ...
+ $ ci_lower         : num [1:28] 317 299 300 367 326 ...
+ $ ci_upper         : num [1:28] 504 491 490 476 458 ...
+ - attr(*, "row.names")= int [1:28] 1 2 3 4 5 6 7 8 9 10 ...
+ - attr(*, "class")= chr [1:3] "list" "esttable" "smallarea"
+\end{Sinput}
+\end{Schunk}
+\end{small}
 

### 2017-10-25 rnw/Introduction.Rnw
formatting revision (3)

@@
 % section: briefly give a review of existing R-packages for forest inventories (not much around)
-While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare.
+% While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}.
@@
 The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multi-phase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
-
-The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
+% 
+% The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
 The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
@@
 The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
+The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.
+
+Commercial as well as open-source software already implement multi-stage and multi-phase estimation, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}.
+However, both are targeted towards list-sampling as it is applied in official statistics.
+Available software providing multi-phase sampling methods better suited for forest inventories has been rare.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}.
+However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}.
+The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
+The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
+The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}.
+The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.
+
 The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.
@@
 
+%% Alex thoughts:
+% Maybe you could mention that commercial and open-source packages exist that implement design-based estimators, but few are specifically adapted to forest inventory.
+%
+% I would mention them as both having the same issue that they are targeted towards list sampling as it is applied in official statistics and not towards forest inventory.
+%
+%
+
 %\newpage

### 2017-10-25 r/replication_rcode.R

@@ -579,11 +579,13 @@ mphase.gain(grisons.sae.table, pref.vartype = "g_variance")
 
 # ----- 7.3 plot()-method --------
 
-## plot point estimates and confidence intervals:
-plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
-  ylab("Timber Volume [m3/ha]")
-
 ## plot estimation errors:
 plot(grisons.sae.table, ncol = 2) 
+
+## plot point estimates and confidence intervals:
+library("ggplot2")
+plot(grisons.sae.table, ncol = 2, yvar = "estimate") +  
+  ylab("Timber Volume [m3/ha]")
+
 
 

### 2017-10-27 rnw/Calculation of Confidence Intervals.Rnw
formatting revision (4)

@@
 Based on the central limit theorem it can be assumed that under hypothetical repeated sampling the point estimates will asymptotically follow a normal distribution.
-However, on the recommendation of \citet{mandallaz2013a}, better confidence intervals can obtained using the Student's \textit{t} distribution (\citet{mandallaz2013a}) defined as
+However, on the recommendation of \citet{mandallaz2013a}, better confidence intervals can obtained using the Student's \textit{t} distribution defined as
 

### 2017-10-27 rnw/Introduction.Rnw
formatting revision (4)

@@
 data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either one additional sample of plot locations consisting of auxiliary information (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
+Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
 Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par
@@
 
-Commercial as well as open-source software already implement multi-stage and multi-phase estimation, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}.
+Multi-stage and multi-phase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}.
 However, both are targeted towards list-sampling as it is applied in official statistics.
 Available software providing multi-phase sampling methods better suited for forest inventories has been rare.
-Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}.
+Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the \pkg{maSAE} package by \citet{cullmann2016}.
 However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}.
@@
  % ... lacks the possibility to choose
+
+ \newpage

### 2017-10-27 rnw/Main structure and modules of the Package.Rnw
formatting revision (4)

@@
 However, this is impossible because there is an infinite number of points in the forest area.
-Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e., $Y(x)$, to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
+Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
 The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
@@
 	\end{subfigure}
-\caption{(a) Concept of multiphase sampling.
+\caption{(a) Concept of multi-phase sampling.
 The square represents the forest area for which an inventory is being conducted.
@@
 In contrast, the design-based approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$.
-While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local densities $Y(x)$ at any location $x \in F$ are fixed and not the result of a stochastic process.
+While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local density surface at any location $x \in F$ are fixed and not the result of a stochastic process.
 A selected observation $Y(x)$ still remains a random variable, but solely due to the random sample mechanism.
@@
 The estimators are available under exhaustive and non-exhaustive use of the auxiliary data.
-Additionally, the package also can calculate one-phase estimates solely based on terrestrial samples.
+Additionally, the package can also calculate one-phase estimates solely based on terrestrial samples.
 All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.

### 2017-10-27 rnw/Special Cases and Scenarios.Rnw
formatting revision (2)
formatting revision (4)

@@
 
-A special case of multi-phase regression estimators is post-stratification, which can further be divided into the cases of multi-phase sampling for stratification and multi-phase sampling for regression within strata.
+A special case of multi-phase regression estimation is post-stratification, which can further be divided into the cases of multi-phase sampling for stratification and multi-phase sampling for regression within strata.
 Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.
@@
 
-The variance of all model-assisted estimators included in \pkg{forestinventory} can be decreased by reducing the sum of square residuals of the regression model.
+The variance of all design-based estimators included in \pkg{forestinventory} can be decreased by reducing the sum of squared residuals of the regression model.
 In case of post-stratification, this particularly implies minimizing the within strata residual square sum.
@@
 In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the extended and pseudo extended synthetic estimator (see EXTSYNTH and EXTPSYNTH in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}).
-In this case, it is adviseable to use the (pseudo) small area estimator (PSMALL) where the zero mean residual property is still ensured.
+In this case, it is adviseable to use the (pseudo) small area estimator (SMALL or PSMALL) where the zero mean residual property is still ensured.
 
@@
 @
+
 \begin{small}

### 2017-10-27 rnw/Threephase_Estimators and their Application.Rnw
formatting revision (4)

@@
 
-The unbiased point estimates of the SMALL and PSMALL are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e., $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^{\top}(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
+The unbiased point estimates of the SMALL and PSMALL estimator are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e., $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^{\top}(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate.
 The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a correction term likewise in the global estimation (Eq.

### 2017-10-27 rnw/Twophase_Estimators and their Application.Rnw
formatting revision (4)

@@
   \item the \code{list}-object \code{phase_id} containing: the \code{phase.col} argument identifying the name of the column specifying membership to $s_1$ or $s_2$, and the \code{terrgrid.id} argument specifying which numeric value indicates $s_2$ membership in that column.
+Note that \pkg{forestinventory} implicitly assumes that all rows not indcated as $s_2$ belong to the $s_1$ phase.
   \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional).
@@
 where $\hat{\var}_{s_{2,G}}$ indicates taking the sample variance over $s_{2,G}$.
-If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
+If boundary adjustment is applied, the simple mean of the explanatory variable vector over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.
 
@@
 The SMALL and PSMALL estimators overcome this issue by using a bias correction term, i.e., $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
-The motivation behind the extended synthetic andextended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
+The motivation behind the extended synthetic and extended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
@@
 \begin{align}
-\hat{Y}_{G,EXTSYNTH,2p} & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
+\hat{Y}_{G,EXTSYNTH,2p} & = \bar{\pmb{\mathbb{Z}}}^{\top}_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
 \hat{Y}_{G,EXTPSYNTH,2p} & =\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^{\top}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
-\hat{\var}(\hat{Y}_{G,EXTSYNTH,2p}) & = \bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
+\hat{\var}(\hat{Y}_{G,EXTSYNTH,2p}) & = \bar{\pmb{\mathbb{Z}}}^{\top}_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
 \hat{\var}(\hat{Y}_{G,EXTPSYNTH,2p})& =
@@
 
-To summarize, the (pseudo) synthetic estimator can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
+To summarize, the synthetic estimators SYNTH and PSYNTH can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.
 When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either SMALL or PSMALL which remove this bias explicitly with a mean residual term, or more elegantly with EXTSYNTH or EXTPSYNTH which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.

### 2017-10-27 submission/JSS_3280_prescreened/replication_rcode.R
revised version after prescreening

@@ -1,0 +1,591 @@
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples                                                     #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
+
+library("forestinventory")
+
+# **************************************************************************** #
+# ------ SECTION 3: TWO-PHASE ESTIMATORS -------------------------------------
+# **************************************************************************** #
+
+
+# ----- 3.1 Global Estimations --------
+
+
+## load grisons data set:
+data("grisons", package = "forestinventory")
+
+## see structure of grisons:
+str(grisons)
+
+## see first 12 rows of grisons:
+head(grisons, n = 12)
+
+
+# **********************************************************
+## non-exhaustive (NEX) estimator with boundary weight adjustment
+reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                      data = grisons,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights")
+
+## get summary of estimation results:
+summary(reg2p_nex)
+
+
+# *****************************************************
+## exhaustive (EX) estimator with boundary weight adjustment
+## first, retrieve order of columnames that the lm()-function processes a formula object:
+#  --> actually the order they appear in the design-matrix ...
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+
+## store true means of explanatory variables:
+true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
+
+## apply estimator:
+reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                     data = grisons,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                     exhaustive = true.means.Z)
+
+## get summary of estimation results:
+summary(reg2p_ex)
+
+## get data frame with estimation results
+reg2p_ex$estimation
+
[…]

### 2017-12-15 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 15_12_2017

@@
   \hat{\var}(\hat{Y}_{G,PSYNTH,2p}) & = \hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
-  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
+  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
 \end{align}
@@
 The SMALL and PSMALL estimators overcome this issue by using a bias correction term, i.e., $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$.
-The motivation behind the extended synthetic and extended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
+The motivation behind the extended synthetic and extended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while ensuring that the mean of the empirical prediction model residuals in the entire area $F$ and the small area $G$ are by construction both zero at the same time.
 This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0.
@@
 \hat{\bar{\pmb{\mathbb{Z}}}}_{G}
-+ \hat{\pmb{\theta}}^{\top}_{s_2}\hat{\Sigma}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
++ \hat{\pmb{\theta}}^{\top}_{s_2}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
 \end{align}

### 2018-08-06 JStatSoft_forestinventory_master.rnw
update 06_08_2018
update 06_08_2018 - first hands on revision

@@
 \author{Andreas Hill \And Alexander Massey}
-\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories}
+\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}
 
@@
 Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.
-They typically consist of a random sample of plot locations in the forest that are assessed individually by hand.
+They typically consist of a random sample of plot locations in the forest that are assessed individually by field crews.
 Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.

### 2018-08-06 revision/reviews.tex
update 06_08_2018
update 06_08_2018 - first hands on revision

\documentclass{article}
\usepackage{authblk}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[ngerman]{babel}
\usepackage[normalem]{ulem}

%============================================ Define Titlepage & packages =============================================%

\title{Reviews}

\author{The R Package forestinventory: design-based global and small area estimations for multi-phase forest inventories\\
Andreas Hill, Alexander Massey}

\usepackage{fancyhdr}     
\usepackage{amsmath} %Paket für erweiterte math. Formeln
\usepackage[labelfont=bf]{caption}
\usepackage[font=footnotesize]{caption}
\usepackage[font=footnotesize]{subcaption}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[final]{pdfpages}
\usepackage{color}

\usepackage{geometry}
\geometry{
	a4paper,
	left=25mm,
	right=25mm,
	top=30mm,
	bottom=30mm
}

\setlength{\parindent}{0em} % Einzug bei neuen Absätzen

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Main Document------------------------------------------ %

\begin{document}

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Tex Settings ------------------------------------------ %

\maketitle
\thispagestyle{empty}
\newpage

\pagenumbering{arabic}
\setcounter{page}{1}

\pagestyle{fancy} %Kopfzeile und Fusszeile
\fancyfoot[C]{\thepage}
\setlength{\headsep}{15mm}

\definecolor{mybrown}{rgb}{0.6, 0.15, 0.1}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\definecolor{mygreen}{rgb}{0.1, 0.4, 0.4}
\newcommand{\answer}[1]{\small \color{mybrown}{#1} \color{black}}
\newcommand{\note}[1]{\textit{\small \color{amaranth} \textbf{Note:} #1} \color{black}}
\newcommand{\todo}[1]{\color{red}{#1} \color{black}}
\newcommand{\answerfin}[1]{\small \color{mygreen}{#1} \color{black}}

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Reviewer 1 ------------------------------------------------ %

\section*{Reviewer 1:}

The manuscript “The R Package forestinventory: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories” by Hill and Massey presents the statistical background and the application of the R package “forestinventory” (Hill, Massey, Mandallaz). This package provides methods for computing estimates and confidence intervals for relevant quantities (such as total timber volume or volume in a subset) in a typical forest inventory setting where precise information about the response is only available at a small sample of locations while explanatory data is available for a wider area.\\

The package, in connection with the manuscript, is likely to be very useful for a wide range of applications in forest inventories (and possibly other fields), and I strongly recommend publication of the manuscript in JSS once the concerns regarding limitations of the methodology, as described below, have been addressed in a suitable form.\\

The manuscript itself is carefully and well written with only very minor inconsistencies. It has a clear structure, offering a healthy combination of mathematical details and presentation of the various methods. There is no question as to its usefulness. The formalism is based on that developed by D. Mandallaz in a book and a number of papers following the design-based approach (where stochasticity is tied to the choice of the sampling locations).\\

The accompanying R package includes more than 7000 lines of code, which in part rely on other packages. Of course, in the context of this review more than a very cursory glance at the implementation (indeed evaluating correctness) is out of the question, so, obviously, any individual results produced by the package, particularly in edge cases, will require double-checking by the user – or trust in the authors. That being said, the code generally appears to be well-written with meaningful comments included throughout. The placement of the various functions in the different files is logical, although the separate handling of one-phase, two-phase and three-phase methods probably involves some duplication. The installation of the package did not cause any trouble, and the R documentation is sound and follows R conventions. The functions tested by the reviewer worked as advertised, but as usual with R involved some time to become familiar with. Specifically, the separate argument phase\_id in the fitting functions (cf. page 10 in the manuscript) didn't seem intuitive when it would presumably be possible to simply define the phases as those rows where the corresponding variables are present (i. e. not NA). In fact, there is some automatic recoding which in part does just that (page 27). Some related discussion is found on page 28.\\

My main concern is that the limitations of the methodology are nowhere clearly specified in the manuscript. On page 5, it is suggested that a consequence of the design-based approach "is that the estimation properties of design-based regression estimators (e. g. unbiasedness) typically hold regardless of the model that is chosen." While this is basically true for unbiasedness, unbiasedness by itself means little, and, more important, the statement does not extend to relevant properties such as coverage rates of confidence intervals. Indeed, if s2 contains less than 1-$\alpha$ of the entire population, any true 1-$\alpha$ confidence interval for the population mean necessarily involves distributional assumptions about the population because otherwise any fixed Y(x0), which has probability > $\alpha$ of not being included in the sample, can be anything, so any conclusion about the mean would be wrong with probability > $\alpha$. (Similarly, with the infinite population approach, one needs assumptions on the tails.) Of course, it is very plausible that such issues go away once reasonable assumptions (suitable sample sizes etc.) are met. While this will be obvious in some situations, I expect many potential uses of the package where this is less so, either because users are specifically looking at edge cases in search of interesting phenomena or, more commonly, because there really is very little data available - the typical situation for a forestry undergraduate writing a thesis. The vague invocation of the central limit theorem (page 24) won't help here. Below I include three numerical examples where the nominal 95\% confidence intervals never exceed a coverage of 85\%.\\

As I personally regard solid knowledge of the limitations of a statistical method as a prerequisite of its use in science and in addition believe it would help many potential users of the package, I ask the authors to consider adding some clarification regarding typical situations where the methods can be or should not be applied. For example, showing a few artificial test cases highlighting the most important requirements would probably not be difficult.\\

The following is a list of minor issues encountered while reading:\\

\begin{enumerate}

	% XX) ++++++++++++++++++++++++++ %
	\item \textit{On page 3, n2 is not introduced (e. g. say a “a sample of n2 points, s2”)}.

	\answerfin{}
	% --------------------------- %

    % XX) ++++++++++++++++++++++++++ %
	\item \textit{On page 27, the element symbol should be a subset symbol.}.

	\answerfin{}
	% --------------------------- %

    % XX) ++++++++++++++++++++++++++ %
	\item \textit{On page 32, grisons should probably be capitalized.}.

	\answerfin{}
	% --------------------------- %

	% XX) ++++++++++++++++++++++++++ %
	\item \textit{Equations (1), (4), (13), (16), (20) might be better readable without the n2 fractions}.

	\answerfin{}
	% --------------------------- %

    % XX) ++++++++++++++++++++++++++ %
	\item \textit{On page 7, the zero mean residual property requires that the regression model includes a constant term (or equivalent). While the R functions apparently fail when you remove the intercept, this should probably be stated somewhere.}.

	\answerfin{}
	% --------------------------- %

\end{enumerate}

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Reviewer 2 ------------------------------------------------ %

\section*{Reviewer 2:}

The authors presented a well-written and interesting paper that will be an important documentation of the well-designed, timely, and highly relevant R package “forestinventory”. In the package, the multi-phase, infinite
(continuous) population estimators by Prof. Daniel Mandallaz and his group are implemented which will help many analysts to make use of these elegant methods. The package is extremely well documented in the help files with many examples and comes with several data sets. The package also comes with methods for ggplot2 that allow a very elegant comparison of the implemented estimators. To conclude, I’d like to thank the authors for sharing their work -- our group is certainly going to use the package a lot.\\

Comments:\\

The paper is rather long which can be a good or bad thing. I think that the 2 and 3 phase estimators are so similar that it may be sufficient to present, for example, SAE only for the 2 phase estimator and global estimation only for the 3 phase estimator. The same is true for the exhaustive and non-exhaustive estimators. Describing just once that in all cases both types are possible (as done for cluster sampling), would make the text a lot easier to read. Also section 5 on CIs could be considerably shortened. Needs to be a section on its own right? Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\

The current description of how to obtaining the explanatory variables seems not to fit well into this paper. Explanatory variables have to be calculated by the analyst before using the package (except for an optional boundary adjustment). Also in the interest of making the paper shorter and more general (not only focusing on forest inventories), consider omitting or shorten drastically this part of the document.\\

Furthermore, I wonder whether the infinite population estimators in combination with the explanatory variables used in the examples are applicable without modification for the exhaustive (wall-to-wall) estimators such as eq. 17a. It may be just a misunderstanding of how to interpret the integral in the equations and the sentence above eq. 2b. However, it sounds like it is suggested to use the full cover of auxiliary information such as lidar height returns and to calculate “metrics” that include order statistics such as 75th percentile and the maximum globally. While this approach is unproblematic for the mean, it results in biased order statistics as they are scale-dependent. For example, the mean of the maximum values for grid-cells (population units) tessellating the area should be calculated rather than the global maximum over the area. The grid cells should have approximately the same size as the sampled units because of the scale-dependency of the explanatory variables but do not necessarily be of the same shape. I am sure this was done correctly when calculating the explanatory variables for the examples. However, this seems to be a crucial point that should be made much clearer; maybe especially so if the authors have a different point of view.\\

Related to the previous comment is whether the difference between the infinite and finite population approaches is relevant from a practical point of view. At least the estimators may be equivalent (see for example eq. 6a). Furthermore, for some cases as described in the previous comment, the population needs to be tessellated which, in my opinion, practically results in finite populations. I think it would be very helpful to discuss the similarities and differences of the finite and infinite population approach more as most readers familiar with forest inventories are likely to be “indoctrinated” with the finite population approach.\\

When providing two estimators for the same purpose such as the synthetic or extended synthetic, some (extreme) users might just chose what results in the smallest standard error. Some more guidance or discussion on when to use which estimator may be needed.\\

Often when using remotely-sensed auxiliary information in forest inventories, due to temporal or spatial mismatch or other errors, analysts have to exclude “outliers” from the sample that would have large leverage in working models. In this case, the zero mean residual property of internal linear models (after eq. 2) does, strictly speaking, not hold. Consider giving some guidance for these very common cases. In the design-based framework, one could of course ignore the issue. However, a lot of efficiency may be lost just due to few observations.\\

Dealing with sample plots close to borders (split plots) is a common issue in forest inventories. It would be helpful to mention how to handle them in the estimators.\\

Specific comments:\\

\begin{enumerate}

% XX) ++++++++++++++++++++++++++ %
\item \textit{Abstract: “by hand” – change for example to “by field crews”}.

\answerfin{Has been changed accordingly.}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Multiphase vs. multi-phase}.

\note{\textit{Andi}: There is a slight inconsistency in our article here: Our package says 'multiphase', but here we used 'multi-phase'. However, should we then also change 'two-phase' to 'twophase'?}

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Introduction: Consider mentioning the important role of (N)FIs in monitoring carbon change in the context of climate reporting and mitigation.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Section2.1: Forest area is often itself an important parameter to be estimated. For completeness, it should be mentioned that (or how) it is possible to estimate forest area under the infinite population approach.}.

\answerfin{The estimators presented in our article particularly rely on the assumption that the forest area $\lambda$ is known. The reason for this is that the local density $Y(x)$ is actually the Horwitz-Thompson}

\note{\textit{Andi}: Whereas the forest area $\lambda(F)$ appears in the Horwitz-Thompson estimator, we actually only have to know the forest boundaries to compute the intersection of the circle and the forest boundary. The exact forest area is only needed to compute the total by multiplying, e.g., the sample mean with $\lambda(F)$. The author is, however, right by saying that in most NFIs the forest area is rather estimated than known exactly. A solution, as applied in the Swiss NFI, is to estimate the forest area by multiplying the \textit{known} country area by the percentage of (randomly and uniformly distributed) sample points falling within the forest (decision made by field crews) -> see Mandallaz 2014: Note on the estimation of totals. So actually, its no big deal "to do it in the infinite population approach". However, in the case of multiphase sampling, also $s_1$ or $s_0$ points can no longer be restricted to the forest, which can lead to loss of efficiency (see also Mandallaz 2014).}

% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Section2.1: “population in the finite approach … not well defined”. Can this really be generalized in this way? If not, consider deleting. Think about estimating forest area where every finite population unit is covered by a certain proportion of forest.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Section 2.4: Consider mentioning already here that the bias correction is not deemed reliable for $n_2G < 6$.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{S2.5: systematic grids are not used for reducing travel costs but rather because of practicalities around obtaining a spatially balanced sample.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Fig 3: Meaning of the asterisk *?}.

\answerfin{Has been corrected.}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{After eq. 2: Where comes the term g-weight from? No “g” in any of the equations around eq. 2?}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{After eq. 6: “variance … external model … usually slightly smaller than” internal… In the examples for the global estimators further down it is vice versa. Consider rephrasing.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{After eqs. 12, 25 and possibly elsewhere “…unbiased point estimates…” Can an estimate be biased or is it estimators that are biased?}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Section6.1: Ancova model: That means the model of the extended synthetic estimator is also of type ancova. Worth mentioning?}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{To me, the term “estimation error” seems a bit unconventional.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{I think the title of the package is a bit narrow, because it suggests that it could only be used for forest inventories (FIs). FIs were the background and examples for the estimators. However, the methods are generally applicable for all kinds of infinite and finite populations that are sampled by a probability design. The package name may be difficult to change (are package aliases allowed on CRAN?) but at least the function parameters could be general. terrgrid.id in the phase\_id list could for example be replaced by a general term. In this list, consider using the underscore as a consistent separator.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Some of the estimators are also implemented in the well-known “survey” package which could be mentioned.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Around eq. 13. The extended estimator is described as being “elegant”. However, it seems to be necessary to fit the model for each small area with a change in the indicator variable. How does this fit?}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{After eq. 14. How can there be a residual correction term in a synthetic estimator? There may not even be observations available from within small area G. Please describe and define the term $\theta_s2$ in more detail.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Sentence after eq. 17: difficult to follow. Consider rephrasing.}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Acknowledgements: grisons - Grisons}.

\answerfin{}
% --------------------------- %

% XX) ++++++++++++++++++++++++++ %
\item \textit{Value of twophase() (Package forestinventory version 0.3.1): $Z_bar_1G$ is $Z_bar_1$?}.

\answerfin{}
% --------------------------- %

\end{enumerate}

%------------------------------------------------------------------------------------------------%
\end{document}

### 2018-08-30 JStatSoft_forestinventory_master-concordance.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 108 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
-1 98 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 215:%
-124 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
+1 105 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
+1 64 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
+121 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
 1 67 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
-130 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
-1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 139 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
+127 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
+1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
+0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
+1 2 8 0 1 2 130 1 1 5 7 0 1 2 2 1 1 2 27 0 1 2 3 1 1 2 13 0 1 2 5 1 %
+1 5 4 0 1 1 8 0 1 2 7 1 1 5 7 0 1 2 2 1 1 2 9 0 1 2 7 1 1 3 5 0 1 2 %
+1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 5 7 0 1 2 2 1 1 2 9 0 %
+1 2 4 1}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
-136}
+133}
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
-1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1237:%
-138 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1242:%
-1 35 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1294:%
-144 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1299:%
-1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
-150 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1553:%
-1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
-156 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
+1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
+102 1 1 3 2 0 1 1 3 0 1 2 4 1 1 5 7 0 1 2 3 1 1 2 12 0 1 2 4 1 1 5 7 %
+0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
+135 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
+1 2 1 1 4 38 1 1 2 11 0 1 2 3 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
+141 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1237:%
+1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
+7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
+0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1472:%
+147 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1477:%
+1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
+0 1 2 2 1 1 6 8 0 1 2 4 1 1 4 6 0 1 2 6 1 1 2 4 0 1 2 40 1 1 2 11 0 %
+1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1666:%
+153 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1671:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
-162 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1683:%
+159 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1688:%
 1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1775:%
-168 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1693:%
+165 11 1}

### 2018-08-30 JStatSoft_forestinventory_master.rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision.
-With respect to this objective, the application of multi-phase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient.
+With respect to this objective, the application of multiphase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient.
 While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent.
-The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling.
+The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling.
 The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata.
@@
 \begin{document}\sloppy
+\SweaveOpts{concordance=TRUE}
 

### 2018-08-30 revision/reviews.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 
-	\answerfin{}
+	\answerfin{Has been changed accordingly.}
 	% --------------------------- %
@@
 
-	\answerfin{}
+	\answerfin{Whereas we agree that they might be slightly more readable without the n2 fractions, the decision to include them was based on the fact that this is how the formulas were originally presented in Mandallaz's papers.
+We feel that the marginal loss in readability is offset by perfect continuity with the original notation.}
 	% --------------------------- %
@@
 
-	\answerfin{}
+	\answerfin{It has been added that the models require intercept terms.}
 	% --------------------------- %
@@
 Describing just once that in all cases both types are possible (as done for cluster sampling), would make the text a lot easier to read.
-Also section 5 on CIs could be considerably shortened.
-Needs to be a section on its own right?
-Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\
+\\
+\\
+\answerfin{Different versions of this article were written that attempted to shorten it in ways just like the review mentioned.
+However, we ultimately decided that the purpose of this article is to bridge the gap between likely users of the software (e.g.
+forest engineers and technical practitioners) and the mathematical details that might be beyond their skillset.
+For this reason, we decided to ensure that every function call was connected explicitly to the underlying formula that it applies and a citation is given for those interested in further technical details.
+We believe this also enables the reader to write their own routines using the functions which can be used to error check our routines.
+Considering that the current length is not excessively long compared to other JSS papers, we strongly prefer to not to shorten sections merely for the sake of making the article shorter.} \\
+
+Also section 5 on CIs could be considerably shortened. Needs to be a section on its own right? Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\
+\\
+\answerfin{Reviewer 1 made a very valid criticism about designed-based confidence intervals when the sampling fraction is too small that we would like to include in the paper and Section 5 is the most logical place to put it.} \\
 
@@
 
+\answerfin{When the package was orginally beta tested at the Swiss Federal Institute of Forest, Snow and Landscape Management on likely operational users as well as forest engineers, the most contentious area for confusion was the calculation of explanatory variables in the exhaustive case.
+The reviewer appears to have confirmed that confusion in the previous paragraph, which is explained in the section on calculating explanatory variables.
+It is explictly stated that the exact exhaustive mean vector can only be calculated for explanatory variables based on means (note that polygon layers are means of binary variables).
+It has now been explicitly added as an example that order statistics are not included in this category with the reviewers example given.
+It is already explicitly stated that the size of the support should ideally be chosen to achieve a best possible explanatory power of the regression model on page 9 and a citation has been added for the interested reader.} \\
+
 Related to the previous comment is whether the difference between the infinite and finite population approaches is relevant from a practical point of view.
@@
 
+\answerfin{The population doesn't need to be tessellated and can be allowed to overlap or not use all the wall-to-wall map in the infinite population approach.
+This is already mentioned on page 9.
+Section 2.1 discusses the Infinite Population Approach.
+Some brief discussion comparing to the finite is included, but as this package is about software written for the infinite not finite approach, we prefer not to delve too deeply into a topic that makes little practical difference.} \\
+
 When providing two estimators for the same purpose such as the synthetic or extended synthetic, some (extreme) users might just chose what results in the smallest standard error.
 Some more guidance or discussion on when to use which estimator may be needed.\\
+
+\answerfin{This is true, but it is already stated many times that the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias making them seem deceptively optimistic (see pages 5, 13 (twice), 14, 16 (this is explained here using a concrete example), 21, 26 and 30)} \\
 
@@
 
+\answerfin{Removing outliers is more of a common practice in model-dependent estimation, which is beyond the scope of this article.
+You should never remove outliers in design-based estimation unless you can defend the belief that the resulting (and self-inflicted) non-response is missing at random.
+The observed value of the the response/explanatory variable is fixed at each $x$ so deleting an observation effectively means the plot at x has been removed from the forest.
+For temporal mismatches this is pretty much indefensible because the outliers likely represent harvesting after the aerial measurement.
+We initially included a discussion comparing design-based with model-dependent estimation but we ultimately decided that the paper has more focus if we concentrate only on explaining how to use this software under the design-based infinite population approach.
+In any case, many citations are currently included that already discuss these issues.} \\
+
 Dealing with sample plots close to borders (split plots) is a common issue in forest inventories. It would be helpful to mention how to handle them in the estimators.\\
+
+\answerfin{A mention of sample plot boundary adjustments is now included in Section 2.1 with a citation.} \\
 
@@
 
-\note{\textit{Andi}: There is a slight inconsistency in our article here: Our package says 'multiphase', but here we used 'multi-phase'. However, should we then also change 'two-phase' to 'twophase'?}
-
-\answerfin{}
+\answerfin{Multiphase is now used consistently to match package's spelling}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{Good point.  Has been changed accordingly.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{That suggestion only can be done if the plots form a perfect tesselation.
+Circular units, the most common in NFIs, can't do this.
+The rest of the sentence actually says "with respect to the forest area".
+This is true.
+Points can fill any forested area even if it is a fractal whereas a counter example forest area can always be given for any plot shape you put forward.
+Extending your tesselation beyond the forest edge is possible, but this increases the weights for boundary cases which likely increases the variance.
+Many of the cited articles in the infinite population approach already discuss this but the focus for this article is not a comparison of the finite and infinite approaches, but a presentation of a software package written under the infinite approach.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The g-weight is the name of a variance estimator for the point estimate in eq.2 as opposed to another possible variance estimator here named the external estimator.
+The origin of the g-weight name is more obvious in the citation provided in the same paragraph.
+The g-weight is analogous to a hat matrix and can be extracted from the more compact form that was included in the article.
+It the spirit of not being too verbose we chose the compact form.
+}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{In this context "a biased estimate" is an estimate that is produced from a biased estimator.
+It would be a bit tedious to explain every time that an given estimate was produced from a biased estimator.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The "survey" package is in the finite population approach (aka list sampling).  It is mentioned on page 2.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{It is elegant in the sense that the SAE estimator is unbiased merely by modifying the design matrix.
+It is not necessary to refit every small area with an indicator, but it is  how the package implemented the estimator.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{It says that "the formulas look similar to the synthetic estimators", because they are exactly same.  They merely have a modified design matrix.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{Change has been implemented.}
 % --------------------------- %

### 2018-08-30 rnw/Analysis and Visualization.Rnw
formula for rel.effiency corrected
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 
-As multi-phase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multi-phase global or small area estimate compared to its respective one-phase estimate.
-The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a baseline.
-If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
+As multiphase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multiphase global or small area estimate compared to its respective one-phase estimate.
+The function takes an \code{esttable} object as input and returns a summary of which multiphase method and estimator performed best using the precision from the one-phase estimator as a baseline.
+If the \code{esttable} object contains more than one multiphase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation.
 The argument \code{pref.vartype} can be used to define what type of variance (g-weight or external) should be used for the comparison.
@@
 % Interpretation: (gain and rel.eff)
-The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multi-phase estimator (\code{var_multiphase}).
-The multi-phase estimation procedure is again specified in the \code{method} and \code{estimator} column.
-The last two columns quantify the potential benefit of the multi-phase estimation.
-The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation.
+The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multiphase estimator (\code{var_multiphase}).
+The multiphase estimation procedure is again specified in the \code{method} and \code{estimator} column.
+The last two columns quantify the potential benefit of the multiphase estimation.
+The \code{gain} is the reduction (if its value is positive) in variance when applying the multiphase as alternative to the one-phase estimation.
 For example, it is indicated that the two-phase extended PSYNTH estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure.
-The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multi-phase variance:
+The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multiphase variance:
 
 \begin{equation}\label{eq:releff}
-  rel.eff_{[\%]}=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}\cdot100
+  rel.eff=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}
   \end{equation}
 
-The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator.
+The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multiphase estimator.
 For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the two-phase EXTPSYNTH estimator.
-If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multi-phase estimation for forest inventories.
+If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multiphase estimation for forest inventories.
 

### 2018-08-30 rnw/Introduction.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time.
-They provide accurate quantitative information that can be used to define management actions and to adapt forest management strategies according to guidelines on national and international levels.
+They provide accurate quantitative information that can be used to define management actions, to adapt forest management strategies according to guidelines on national and international levels and to support in monitoring carbon change in the context of climate reporting and mitigation.
 As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods.
@@
 As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}.
-A method which has become particularly attractive is called multi-phase sampling.
+A method which has become particularly attractive is called multiphase sampling.
 The core concept is to enlarge the sample size in order to gain higher estimation precision without enlarging the terrestrial sample size.
@@
 % section: briefly give a review of existing R-packages for forest inventories (not much around)
-% While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare.
+% While commercial enterprises have already been offering multiphase inventory analyses, freely available software for application of multiphase estimators has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}.
 However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available.
-The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multi-phase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
+The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multiphase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
 % 
@@
 
-Multi-stage and multi-phase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}.
+Multistage and multiphase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}.
 However, both are targeted towards list-sampling as it is applied in official statistics.
-Available software providing multi-phase sampling methods better suited for forest inventories has been rare.
+Available software providing multiphase sampling methods better suited for forest inventories has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the \pkg{maSAE} package by \citet{cullmann2016}.
@@
 
-% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare.
+% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multiphase sampling methods to forestry has been rare.
 Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}.

### 2018-08-30 rnw/Main structure and modules of the Package.Rnw
update 06_08_2018
update 06_08_2018 - first hands on revision
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by coordinates.
-The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha).
-The estimators implemented in \pkg{forestinventory} use the so called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators.
+The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha) including potential boundary adjustments (see \cite{mandallaz2008} for details).
+The estimators implemented in \pkg{forestinventory} use the so-called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators.
 This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig.
@@
 However, this is impossible because there is an infinite number of points in the forest area.
-Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
+Our strategy is thus to take a sample of $n_2$ points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
 The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
@@
 The key point in the infinite population approach is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area.
-This has some theoretical advantages over the finite population approach, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.
+Sampling the actual plot areas follows the so-called finite population approach, where the sample units are usually assumed to be either circular or rectangular plots.
+Although choosing between infinite and finite population schemas does not likely make much practical difference, the former does has some theoretical advantages.
 This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape.
 Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area.
+Also, cluster sampling is much more easily generalized in the infinite population approach.
 The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.
@@
 	\end{subfigure}
-\caption{(a) Concept of multi-phase sampling.
+\caption{(a) Concept of multiphase sampling.
 The square represents the forest area for which an inventory is being conducted.
@@
 \ref{fig:concmphase_and_sae_b}).
-As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient.
+As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multiphase estimation can be efficient.
 However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals.
@@
 All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.
-The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
+The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}.
 While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package.
@@
 \resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
-\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}.}
+\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}. All estimators are also available for cluster sampling.}
 \label{fig:struct_package}

### 2018-08-30 rnw/Onephase_Estimators and their Application.Rnw
update Andreas 22.08.2017
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

--- 2017-08-08 rnw/Threephase_Estimators and their Application.Rnw
@@
 
-\section{Threephase Estimators and their Application}
-\label{sec:threephase_and_appl}
+\section{One-phase Estimators and their Application}
+\label{sec:onephase_and_appl}
 
 % ---------------------------------------------------------------------------- %
-\subsection{Global Estimators}
+  \subsection{One-phase Estimators}
 
-% - Merge Mathematical Background and Application
-% - Quite Likely, we cannot go through all estimators ...
-%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
-%
-% - 
-%
-% - We could give the function call for all cases, but we should only show the output
-%   for demonstrating particular things
+\subsubsection{Mathematical Background}
 
-% Concentrate on subset of estimators to go through the "tree": --> non-exhaustive case
-% 1) nex, simple sampling, pseudo extended sae estimator
-% 2) nex, simple sampling, pseudo sae estimator
-% 3) nex, simple sampling, pseudo synthetic sae estimator --> give reason when to use (if n2G < 5, because ...)
-% 
-% 4) for the cluster case, show: nex, simple sampling, pseudo extended sae estimator
-% 5) for the exhaustive case, show: ex, simple sampling, pseudo extended sae estimator
-%    (mention that in cluster case, the cluster means shall be used)
-% always use bw-adjustment (?)
-%
+\subsubsection{Application}
 
-\subsection{Small Area Estimators}
+In the package \pkg{forestinventory}, calculating estimates exclusively based on terrestrial observations, i.e.
+one-phase estimations, can be applied with the \code{onephase()} function.
+To demonstrate the function, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository.
+The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample).
+A systematic subsample of 67 out of the 306 observations ($s_2$ sample) provides the local density $Y(x)$, in this case the timber volume per hectare (\code{tvol}) derived from the terrestrial inventory campaign.
+We can load the data frame by executing:
 
-% The psynth estimator:
-% 
-% \begin{equation}\label{pseudosynth1}
-% \hat{Y}_{G,psynth}=\hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\beta}}_{s_2}
-% =\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\hat{Y}(x)
-% \end{equation}
-% 
-% \begin{equation}\label{estvarpseudosynth1}
-% \hat{\var}(\hat{Y}_{G,psynth}): =
-% \hat{\bar{\pmb{Z}}}_{1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1,G}
-% + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1,G}}\hat{\pmb{\beta}}_{s_2}
-% \end{equation}
+\begin{small}
+<<>>=
+data("grisons", package = "forestinventory")
+@
+\end{small}
+
+The inventory data for the \code{onephase} estimation must be provided as a \code{data.frame} that contains a column with the local density values and additionally a column specifying to which sample ($s_0$, $s_1$, $s_2$) the respective observation belongs.
+In the simple one-phase case, all observations in the data frame will of course only belong to one, i.e.
+the terrestrial phase.
+However, the specification is mandatory since in most applications, the input dataset will contain observations from multiple phases, and the indication of the terrestrial sample is thus required to identify the terrestrial sample.
+
+We apply the \code{onephase()} function by specifying the following arguments:
+
+\begin{itemize}
+  \itemsep0em
+  \item the name of the column containing the local density $Y$, which is defined as a formula expression that must be of the form $Y~1$
+  \item the \code{data.frame} containing the inventory data
+  \item the \code{list}-object \code{phase_id} taking the column name (\code{phase.col}) specifying the sample association of an observation ($s_1$ or $s_2$) and the indicator for the $s_2$ phase (\code{terrgrid.id})
+\end{itemize}
+
+\begin{small}
+<<>>=
+op <- onephase(formula = tvol~1 ,data = grisons,
+              phase_id =list(phase.col = "phase_id_2p",terrgrid.id = 2))y
+@
+\end{small}
+
+\newpage

### 2018-08-30 rnw/Special Cases and Scenarios.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 
-A special case of multi-phase regression estimation is post-stratification, which can further be divided into the cases of multi-phase sampling for stratification and multi-phase sampling for regression within strata.
+A special case of multiphase regression estimation is post-stratification, which can further be divided into the cases of multiphase sampling for stratification and multiphase sampling for regression within strata.
 Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.
@@
 Two-phase sampling for stratification is applied if the model only contains categorical variables, in this case the factor variable \code{stage}.
-Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators, is equivalent to post-stratification.
+Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multiphase regression estimators, is equivalent to post-stratification.
 For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).
@@
 If an interaction term is present, both the intercept and the slope are allowed to vary within the strata.
-Note that one can actually use the entire range of OLS regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.
+Note that one can actually use the entire range of OLS regression techniques in the multiphase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.
 
@@
 
-As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multi-phase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other.
+As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multiphase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other.
 The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig.
@@
 
-Based on the nesting rule, $s_2 \in s_1 \in s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model.
+Based on the nesting rule, $s_2 \subset s_1 \subset s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model.
 If e.g., an $s_2$ and/or $s_1$ point misses a variable which is used in the full and reduced model (in this case \code{mean}), the function will delete this sample point from the dataset and produce the following messages:

### 2018-08-30 rnw/Twophase_Estimators and their Application.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@
 
-Note that for internal linear models the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (zero mean residual property) which is why it does not appear in the point estimate.
+Note that for internal linear models containing intercept terms, the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (zero mean residual property) which is why it does not appear in the point estimate.
 More explanation about how to obain the auxiliary means is given in the next subsection.
@@
 In this case, one can calculate the exact mean as the area-weighted average of each categorical variable.
-The second case is when the exact mean can be calculated in one step, e.g., taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
-However, for most types of explanatory variables we will try to get an approximation of $\bar{\pmb{Z}}$ that is only negligibly different.
+The second case is when the exact mean can be calculated in one step, e.g., taking the mean of all height pixels of a raster canopy height model will exactly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
+However, for most types of explanatory variables, e.g.
+order statistics such as maximum or median canopy height in the spatial extent, this approach will fail.
+For example, the mean of the maximum pixel heights across all supports is clearly not equal to the global maximum pixel height across the entire area.
+In such cases, we will try to get an approximation of $\bar{\pmb{Z}}$ that is only negligibly different.
 \par
@@
 One is that a perfect tessellation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size.
-Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tessellation might often not be feasible.
+Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model \citep{hill2018} a perfect tessellation might often not be feasible.
 In the infinite population frame, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$.

### 2018-08-30 texfiles Daniel/techreportpartialexhaustive.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

\documentclass[a4paper,12pt,leqno, titlepage]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\bibliographystyle{mystyle2}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\VAR}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{RE}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}
\newcommand{\R}{\mathbb{R}}
\setlength{\topmargin}{0cm}
\setlength{\headheight}{3mm}
\setlength{\headsep}{1cm}
\setlength{\topskip}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{22cm}
\setlength{\oddsidemargin}{0cm}
\makeatletter
\def\tagform@#1{\maketag@@@{[\ignorespaces#1\unskip\@@italiccorr]}}
\makeatother

\begin{document}
\doublespacing
\pagestyle{plain}
\pagenumbering{arabic}

\title{Regression estimators in forest inventories with two-phase sampling and partially exhaustive information with applications to small-area estimation}
\author{Daniel Mandallaz \thanks{Tel. ++41(0)44 6323186 e-mail
    daniel.mandallaz@env.ethz.ch} \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
\date{}

\maketitle
\newpage
\nolinenumbers
\begin{abstract}
We consider two-phase sampling schemes where one component of the auxiliary information is known in every point ("wall-to wall") and a second component is available only in the large sample of the first phase, whereas the second phase yields a sub-sample with the terrestrial inventory data based on general tree inclusion probabilities. We propose a generalized version of the classical two-phase regression estimator, for global and local estimation and derive its asymptotic design-based variance. Cluster and two-stage sampling procedures are also considered.

\end{abstract}
\clearpage

\section{Introduction}\label{introduction}
\pagenumbering{arabic} \setcounter{page}{1}
The motivation for this work is due to the increasing need of using national or regional inventories for local estimation in order to meet tighter budgetary constraints, which is only feasible under extensive use of auxiliary information, provided e.g. by remote sensing. It is hoped that the proposed estimators will contribute to achieve this objective, particularly because they are easy to implement in software packages like SAS or R.\\
There is an extensive literature on the problem of small area estimation (or small domain estimation in general sampling). In this paper we shall investigate the properties of some estimators in the \textbf{model-assisted framework}, in which prediction models are used to improve the efficiency but are not assumed to be correct as in the \textbf{model-dependent approach}. The validity of the statistical procedures is ensured by the randomization principle: i.e. we are in the \textbf{design-based} inference framework, which has a definite advantage in official statistics. The reader is referred to (\cite{koehl}, section 3.8) for a good review of small-area estimation in forest inventory that presents alternative techniques, in particular Bayesian. Let us now define the sampling scheme.\\
The \textbf{first phase} draws a large sample $s_1$ of $n_1$ points
$x_{i}\in{s_1}$ ($i=1,2\ldots n_1$) that are independently and uniformly distributed
within the forest area $F$. At each of those points auxiliary
information is collected, very often coding information of qualitative nature
(e.g. following the  interpretation of aerial photographs) or quantitative (e.g. timber volume estimates  based on LIDAR measurements). We shall assume that the auxiliary information at point $x$
is described by the row vector $\pmb{Z}^t(x)=(\pmb{Z}^{(1)t}(x),\pmb{Z}^{(2)t}(x))\in{\Re^{p+q}}$ (the upper index $t$ denotes the transposition operator). The first component $\pmb{Z}^{(1)}(x)\in{\Re^{p}}$ of this vector is known at all points $x\in{F}$, it is the \textbf{exhaustive} part of the auxiliary information, e.g. it could be given by thematic maps. The second component $\pmb{Z}^{(2)}(x)\in{\Re}^q$ is known only at points $x\in{s}_1$.\par
 The \textbf{second phase} draws a small sample $s_2\subset{s_1}$ of
$n_2$ points from $s_1$ according to equal probability
sampling without replacement. In the forested area $F$ we consider a well defined population $ \mathcal{P}$ of $N$ trees with response variable
 $Y_i,\;i=1,2 \ldots$, e.g. the timber volume.  The objective is to estimate the spatial mean  $\bar{Y}=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i$, where $\lambda(F)$ denotes the surface area of $F$ (usually in ha). For each point $x\in{s_2}$ trees are drawn from the population $\mathcal{P}$ with probabilities $\pi_i$, for instance with concentric circles or angle count techniques. The
set of trees selected at point $x$ is denoted by $s_{2}(x)$. From each of the
selected trees $i\in{s_{2}(x)}$ one determines $Y_i$. The indicator variable $I_i$ is defined as
\begin{equation}\label{1stage}
 I_i(x)=\begin{cases}&1 \text{ if $i\in s_{2}(x)$}\\
                      &0 \text{ if $i\not\in s_{2}(x)$}
         \end{cases}
\end{equation}
At each point $x\in{s_2}$ the
terrestrial inventory provides the local density $Y(x)$
\begin{equation}\label{truelocaldensity}
 Y(x) =\frac{1}{\lambda(F)}\sum_{i=1}^N \frac{I_i(x)Y_i}{\pi_i}=\frac{1}{\lambda(F)}\sum_{i\in{s}_2(x)} \frac{Y_i}{\pi_i}
 \end{equation}
 The term $\frac{1}{\lambda(F)\pi_i}$ is the tree extrapolation factor $f_i$ with dimension $ha^{-1}$. Because of possible boundary adjustments $\lambda(F)\pi_i=\lambda(F \cap K_i)$, where $K_i$ is the inclusion circle of the $i$-th tree. In the infinite population or Monte Carlo approach one samples the function $Y(x)$ (\cite{mandallaz}) for which the following important relation holds:
 \begin{equation}\label{montecarlo}
 \EX_{x} (Y(x))=\frac{1}{\lambda(F)}\int_{F} Y(x)dx=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i=\bar{Y}
 \end{equation}
 Where $\EX_x$ denotes the expectation with respect to a random point $x$ uniformly distributed in $F$. This establishes the link between the infinite population (continuum) $\{x\in{F} \mid Y(x)\}$ and the finite population of trees $\{i=1,2 \ldots N \mid Y_i\}$.\par
 We shall work with the following linear models (see \cite{mandallaz}, Chapter 6, and \cite{mandallazreport1} for more details)
 \begin{enumerate}
 \item
 The large model $M$
 $$ Y(x)=\pmb{Z}(x)^t\pmb{\beta}+ R(x)=\pmb{Z}^{(1)t}(x)\pmb{\beta}^{(1)}
 +\pmb{Z}^{(2)t}(x)\pmb{\beta}^{(2)}+ R(x)$$
 \noindent with $\pmb{\beta}^t=({\pmb{\beta}^{(1)t}}, {\pmb{\beta}^{(2)t}})$.
 The intercept term is contained in $\pmb{Z}^{(1)}(x)$ or it is a linear combination of its components.\\
 The theoretical regression parameter $\pmb{\beta}$ minimizes
 $$\int_F (Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2dx$$
 It satisfies the normal equation
 $$\big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\big)\pmb{\beta}=\int_F Y(x)\pmb{Z}(x)dx$$ and the orthogonality relationship
 $$\int_F R(x)\pmb{Z}(x)dx=\pmb{0}$$
 \noindent in particular the zero mean residual property
 $$\frac{1}{\lambda(F)}\int_F R(x)dx=0$$
 \item
 The reduced model $M_1$
 $$ Y(x)=\pmb{Z}^{(1)t}(x)\pmb{\alpha} + R_1(x)$$
 The theoretical regression parameter $\pmb{\alpha}$ minimizes
 $$\int_F (Y(x)-\pmb{Z}^{(1)t}(x)\pmb{\alpha})^2dx$$
 It satisfies the normal equation
 $$\big(\int_F \pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)dx\big)\pmb{\alpha}=\int_FY(x)\pmb{Z}^{(1)}(x)dx$$ the orthogonality relationship
 $$\int_F R_1(x)\pmb{Z}^{(1)}(x)dx=\pmb{0}$$
 \noindent in particular the zero mean residual property
 $$ \frac{1}{\lambda(F)}\int_F R_1(x)dx=0$$.
\end{enumerate}
Let us emphasize the fact that we do not assume that the regression models are correct: the inference is based on the sampling design, that is we are doing model-assisted (and not model-dependent or model-based) inference in the sense of S\"{a}rndal (see \cite{sarndal}).
\section{The generalized regression estimator}
We need the following design-based least squares estimators of the regression coefficients of the reduced model, which are essentially solutions of sample copies of the normal equations
\begin{eqnarray}\label{coeff1}
\hat{\pmb{\alpha}}_k &=& \Big(\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)}(x)^t
\Big)^{-1}\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)\nonumber\\
&:=&(\pmb{A}^{(1)}_k)^{-1}
\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)\quad k=1,2
\end{eqnarray}
For the large model we set
\begin{equation}\label{coeff2}
\hat{\pmb{\beta}}_{k}=\Big(\frac{1}{n_2}\sum_{x\in{s}_k}\pmb{Z}(x)\pmb{Z}(x)^t
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_k}Y(x)\pmb{Z}(x):
=\pmb{A}^{-1}_k\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x)
\end{equation}
Note that only $\hat{\pmb{\alpha}}_2$ and $\hat{\pmb{\beta}}_2$ are observable and that
in general the vector consisting of the first $p$ components of $\hat{\pmb{\beta}}_{2}$ is not equal to $\hat{\pmb{\alpha}}_2$ (they are if the corresponding explanatory variables are orthogonal in the classical least squares sense).\\
The large model yields the predictions $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{2}$ and the reduced model the predictions $\hat{Y}_1(x)=\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2$.\\
The \textbf{generalized regression estimate} is defined as
\begin{equation}\label{ygreg1}
\hat{Y}_{greg} = \frac{1}{\lambda(F)}\int_F \hat{Y}_1(x)dx
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}(x)-\hat{Y}_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x))
\end{equation}
This estimator is the Monte Carlo version of S\"{a}rndal's  regression estimator for two-phase sampling in finite population (see \cite{sarndal}, equation 9.7.20). It is clear by the law of large numbers that  $\hat{\pmb{\beta}}_2$ and $\hat{\pmb{\alpha}}_2$ are asymptotically design-unbiased estimators of $\pmb{\beta}$ and $\pmb{\alpha}$. This implies at once that
$$\EX_{1,2}\hat{Y}_{reg}= \EX_1\EX_{2 \mid 1}\hat{Y}_{reg} \approx \bar{Y}$$
$ \EX_{2 \mid 1}$ denotes the conditional expectation of the second phase given the first phase (i.e. simple random sampling without replacement in the population $s_1$) and $\EX_1$ denotes the expectation with respect to uniformly distribution points of the first phase (i.e. to $\EX_x$).\\The generalized regression estimate is therefore asymptotically design-unbiased.\\
To understand the potential usefulness of $\hat{Y}_{greg}$ we shall assume for the time being that the model is \textbf{external}, i.e. not fitted by the inventory data, and that the regression coefficients have given fixed values. Using the well known variance decomposition
\begin{equation}\label{conddecomposition}
\VAR(\hat{Y}_{greg})=\VAR_1\EX_{2 \mid 1}(\hat{Y}_{greg})+\EX_1\VAR_{2 \mid 1}(\hat{Y}_{greg})
\end{equation}we get the design-based variance as
\begin{equation}\label{avar1}
\VAR(\hat{Y}_{greg})=\frac{1}{n_1}\VAR_x(R_1(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_x(R(x))
\end{equation}
The variances $\VAR_x( \cdot)$ are calculated under the uniform distribution in $F$ of the random point $x$. An unbiased estimate of the variance is given by
\begin{equation}\label{estvarclassic}
\hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R_1(x)-\hat{\bar{R}}_1)^2
+ (1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s_2}}(R(x)-\hat{\bar{R}})^2
\end{equation}
where $\hat{\bar{R}}_1=\frac{1}{n_2}\sum_{x\in{s_2}}R_1(x)$ and likewise for $\hat{\bar{R}}$.

This should be compared with the standard result for the variance of the regression estimator $\hat{Y}_{reg}$ under the large model
\begin{equation}\label{classicreg}
\hat{Y}_{reg}=\frac{1}{n_1}\sum_{x\in{s}_1}\hat{Y}(x)+\frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x))
\end{equation}
whose theoretical variance is given by
\begin{equation}\label{avarclassic}
\VAR(\hat{Y}_{reg})=\frac{1}{n_1}\VAR_x(Y(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_x(R(x))
\end{equation}

Thus, by using also the exhaustive information the variance of the observations in [\ref{avarclassic}] is replaced by the variance of the residuals under the reduced model, a very nice and intuitive result indeed. To have better insight consider the design-based coefficients of determination
\begin{equation}\label{coeffdeter}
R_1^2=\frac{\VAR_x(\hat{Y}_1(x))}{\VAR_x(Y(x))} \quad \text{and} \quad R^2=\frac{\VAR_x(\hat{Y}(x))}{\VAR_x(Y(x))}
\end{equation}
According to (\cite{mandallaz}, equation 5.4) the reduction in variance is easily found to be
\begin{equation}\label{varreduction}
\VAR(\hat{Y}_{reg})-\VAR(\hat{Y}_{greg})=\VAR(Y(x))\Big(\frac{R^2-R_1^2}{n_1}+\frac{1-R^2}{n_2}\Big)>0
\end{equation}
We now give an alternative definition of $\hat{Y}_{greg}$ based on the estimated regression coefficients which is essential to derive the design-based variance with \textbf{internal models}, i.e. fitted with the inventory data at hand, and for future generalization to the small-area estimation problem. To this end we need the following mean values
\begin{equation}\label{meanvalues}
\bar{\pmb{Z}}^{(1)}=\frac{1}{\lambda(F)}\int_F \pmb{Z}^{(1)}(x)dx ,\quad \hat{\bar{\pmb{Z}}}^{(1)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(1)}(x) ,
\quad \hat{\bar{\pmb{Z}}}_k=\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}(x),\;k=1,2
\end{equation}
The regression estimate can be rewritten as
\begin{eqnarray}\label{ygreg2}
\hat{Y}_{greg}&=&(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 + (\hat{\bar{\pmb{Z}}}_1-\hat{\bar{\pmb{Z}}}_2)^t\hat{\pmb{\beta}}_{2}+\frac{1}{n_2}
\sum_{x\in{s}_2}Y(x) \nonumber \\
&=& (\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2}
\end{eqnarray}
The last equations follows from the fact that the sum of the residuals is zero by construction.
Note that it suffices to know the integral of $\pmb{Z}^{(1)}(x)$ and not necessarily the values at all points $x\in{F}$.

\section{Variance estimate}
To obtain a first estimate of the variance we can treat the internal model as an external one and replace in [\ref{avar1}] the theoretical residuals by their empirical versions $\hat{R}_1(x)=Y(x)-\hat{Y}_1(x)=Y(x)-\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2$ and $\hat{R}(x)=Y(x)-\hat{Y}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_2$, which have zero means, to obtain
\begin{equation}\label{estavar1}
\hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}\hat{R}_1^2(x)+
\frac{1}{n_2}(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{x\in{s}_2}\hat{R}^2(x)
\end{equation}
To derive better variance estimates we shall use the g-weights technique (for details see \cite{mandallaz}, section 6.2, for the Monte Carlo approach and \cite{sarndal}, sections 6.5 and 6.6 for finite populations). The g-weights are defined by
\begin{eqnarray}\label{gweight1}
g_2(x) &=& 1+(\hat{\bar{\pmb{Z}}}_1-\hat{\bar{\pmb{Z}}}_2)^t\pmb{A}_2 ^{-1}\pmb{Z}(x)  =\hat{\bar{\pmb{Z}}}_1^t\pmb{A}_2}^{-1}{\pmb{Z}(x)\nonumber \\
g^{(1)}_1(x) &=& 1+(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t(\pmb{A}^{(1)}_1)^{-1}\pmb{Z}^{(1)}(x) =\bar{\pmb{Z}}^{(1)t}(\pmb{A}^{(1)}_1)^{-1}\pmb{Z}^{(1)}(x)
\end{eqnarray}
That the two versions of the g-weights are equivalent is a consequence of the zero residual sum for any local density (see \cite{mandallaz}, section 6.2). The g-weights are therefore of order $1+O_p(n_k^{-\frac{1}{2}})$ in probability. Note that they depend not only on the point $x$ but also on the entire sample $s_2$, though weakly. Straightforward algebra leads to the following important calibration properties
\begin{eqnarray}\label{gweight2}
\frac{1}{n_2}\sum_{x\in{s}_2}g_2(x)\pmb{Z}(x)&=&\hat{\bar{\pmb{Z}}}_1 \nonumber \\
\frac{1}{n_1}\sum_{x\in{s}_1}g^{(1)}_1(x)\pmb{Z}^{(1)}(x)&=&\bar{\pmb{Z}}^{(1)}
\end{eqnarray}
and also to
\begin{eqnarray}\label{gweight3}
\frac{1}{n_1}\sum_{x\in{s}_1}g_1^{(1)}(x)Y(x)&=&\frac{1}{n_1}\sum_{x\in{s_1}}Y(x) +
(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)\hat{\pmb{\beta}}^{(1)}_1 \nonumber \\
\frac{1}{n_2}\sum_{x\in{s}_2}g_2(x)Y(x)&=&\frac{1}{n_2}\sum_{x\in{s}_2}Y(x) +
(\hat{\bar{\pmb{Z}}}_1-\hat{\bar{\pmb{Z}}}_2)^t\hat{\pmb{\beta}}_2 \nonumber \\
&=& \hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\beta}}_2=\hat{Y}_{reg}
\end{eqnarray}
$\hat{Y}_{reg}$ is the standard regression estimator based on the large model only (\cite{mandallaz}, section 5.1). The last equation in [\ref{gweight3}] follows again from the fact that the residuals sum up to zero. Also, note that the first quantity in [\ref{gweight3}] is not observable. \\
Intuitively, because the g-weights provide perfect estimates for the means of the auxiliary variables, they must perform well for the response variables if the models are adequate. Let us define
\begin{equation}\label{delta}
\Delta=(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t
(\hat{\pmb{\alpha}}_2-\hat{\pmb{\alpha}}_1)
\end{equation}
Then, using [\ref{gweight3}] and [\ref{ygreg2}] we obtain after some algebra the following formal decomposition of the regression estimate
\begin{equation}\label{yreg3}
\hat{Y}_{greg}=\frac{1}{n_1}\sum_{x\in{s}_1}g_1^{(1)}(x)Y(x)+
\frac{1}{n_2}\sum_{x\in{s}_2}g_2(x)Y(x)-\frac{1}{n_1}\sum_{x\in{s}_1}Y(x) +\Delta
\end{equation}
This is a purely formal identity because only the second term is observable. In the definition
[\ref{delta}] the first factor is of order $O(n_1^{-\frac{1}{2}})$ in design-probability and likewise the second factor of order $O(n_2^{-\frac{1}{2}})$. Thus $\Delta$ is of order $=O(n_2^{-1})$ in design-probability and can be neglected with respect to the first three terms in [\ref{yreg3}], which are of order $O(1)$.\\
In many applications with categorical explanatory variables it can happen that the matrices occurring in
[{\ref{gweight1}] are singular, in which case generalized inverse must be used instead. In this case any particular solution $\hat{\pmb{\beta}}_2^*$ of the consistent normal equations can be used and all statistically relevant quantities like g-weights and predictions are independent of the particular solution chosen and equations [\ref{gweight2}] remain valid (see \cite{Renssen} for details and further references). \\
We introduce the theoretical residuals $R_1(x)$ and $R(x)$ by the relations
\begin{eqnarray}\label{theorres}
Y(x)&=& R_1(x)+\pmb{Z}^{(1)t}(x)\pmb{\beta}^{(1)} \nonumber \\
Y(x) &=& R(x)+\pmb{Z}^t(x)\pmb{\beta}
\end{eqnarray}
Substituting these equalities into [\ref{yreg3}] and using [\ref{gweight2}] we obtain
following expression for the error term
\begin{eqnarray}\label{error1}
\hat{Y}_{greg}-\bar{Y} &=&\Big(\frac{1}{n_1}\sum_{x\in{s_1}}g_1^{(1)}(x)R_1(x)-\frac{1}{\lambda}\int_F R_1(x)dx\Big) \nonumber \\
&+& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)R(x)-\frac{1}{n_1}\sum_{x\in{s_1}}R(x)\Big)+ \Delta
\end{eqnarray}
 This is the Monte Carlo version of equation 9.7.18 given in \cite{sarndal}.\\
According to general heuristic principles (described in \cite{mandallaz}, section 6.2 and \cite{sarndal}, section 6.6) it can be expected that the following variance estimate based on the g-weights has better performances
\begin{equation}\label{estavar2}
\hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(g_1^{(1)}(x))^2\hat{R}_1^2(x)+
\frac{1}{n_2}(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{x\in{s}_2}g_2^2(x)\hat{R}^2(x)
\end{equation}
This is the perfect Monte Carlo analogy of equation 9.7.22 in \cite{sarndal}.\\
We now propose a different technique to obtain the design-based variance of $\hat{Y}_{greg}$ which is better suited for the small-area estimation problem. The starting point is the following important result for the design-based variance of the regression coefficients based on the Taylor linearization technique (for proofs see \cite{mandallaz}, p. 125 and \cite{mandallazreport1}) which leads to the asymptotic covariance matrices
\begin{equation}\label{lemma1}
\pmb{\Sigma}_{\hat{\pmb{\beta}}_k}=\pmb{A}^{-1}
\Big(\frac{1}{n_k}\EX R^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}^{-1}
\end{equation}
where $\pmb{A}=\EX_x \pmb{Z}(x)\pmb{Z}^t(x)$ and
\begin{equation}\label{lemma2}
\pmb{\Sigma}_{\hat{\pmb{\alpha}}_k}=(\pmb{A}^{(1)})^{-1}
\Big(\frac{1}{n_k}\EX R_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)})^{-1}
\end{equation}
We use the variance decomposition [\ref{conddecomposition}] on [\ref{ygreg2}]. One gets
$$\EX_{2 \mid 1}\hat{Y}_{greg}=(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}_1^{(1)})^t\hat{\pmb{\alpha}}_1+
\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\beta}}_1 \approx \bar{\pmb{Z}}^{(1)t}\hat{\pmb{\alpha}}_1$$
because $\hat{\bar{\pmb{Z}}}_1^{(1)t}\hat{\pmb{\alpha}}_1$ and $\hat{\pmb{Z}}^t_1\hat{\pmb{\beta}}_1$ both tend to $\bar{Y}$ asymptotically. Therefore, one has
$$\VAR_1\EX_{2 \mid 1}(\hat{Y}_{greg})=\bar{\pmb{Z}}^{(1)t}\pmb{\Sigma}_{\hat{\pmb{\alpha}}_1}\bar{\pmb{Z}}^{(1)}
=\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}\pmb{\Sigma}_{\hat{\pmb{\alpha}}_2}\bar{\pmb{Z}}^{(1)}$$
To calculate $\VAR_{2 \mid 1}(\hat{Y}_{greg})$ we note that
$$\hat{Y}_{greg}-\EX_{2 \mid 1}Y_{greg}=(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}_1^{(1)})^t(\hat{\pmb{\alpha}}_2-\hat{\pmb{\alpha}}_1)
+\hat{\bar{\pmb{Z}}}_1^t(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1) \approx \hat{\bar{\pmb{Z}}}_1^t(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1) $$
because by the law of large numbers the first term is of order $O(n_2^{-1})$ and the second of order $O(n_2^{-\frac{1}{2}})$. Using the Taylor expansion given in \cite{mandallaz} (at point $(\pmb{A}_1,\;\hat{\pmb{\beta}}_1)$ instead of $(\pmb{A},\;\pmb{\beta}))$ we get
$$\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1\approx \pmb{A}_1^{-1}\big(\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x)\big)$$
\noindent By the properties of simple random sampling without replacement within $s_1$ and the approximations $\frac{n_2-1}{n_1-1} \approx\frac{n_2}{n_1}$, $\sum_{x\in{s_1}}R(x)\pmb{Z}(x)=0$ (orthogonality relationship), we obtain after some algebra
$$\EX_1\VAR_{2 \mid 1}(\hat{Y}_{greg})=(1-\frac{n_2}{n_1})\bar{\pmb{Z}}^t\pmb{\Sigma}_{\hat{\pmb{\beta}}_2}\bar{\pmb{Z}}$$
\noindent To get asymptotic estimates of the covariances matrices one can replace $\pmb{A}$ by $\pmb{A}_1$ or $\pmb{A}_2$, and likewise $\pmb{A}^{(1)}$ by $\pmb{A}^{(1)}_1$ or $\pmb{A}^{(1)}_2$. To ensure the important calibration properties [\ref{gweight2}] we will use
\begin{equation}\label{estvarygregnew}
\hat{\VAR}(\hat{Y}_{greg})=\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}
\bar{\pmb{Z}}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}
\hat{\bar{\pmb{Z}}}_1
\end{equation}
with
\begin{equation}\label{robustvar1}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}=\pmb{A}_2^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}_2^{-1}
\end{equation}
and
\begin{equation}\label{robustvar2}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}=(\pmb{A}^{(1)}_1)^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)}_1)^{-1}
\end{equation}
are the estimated \textbf{design-based covariance matrices} of the regression coefficients under the large and reduced models, which have also been discussed in a totally different context, i.e.  model-dependent least squares theory under non-standard conditions, by \cite{huber1} and \cite{gregoire2}, they are sometimes called \textbf{ robust covariance matrices}. It is straightforward to see that [\ref{estvarygregnew}] and [\ref{estavar2}] are equal.\\
 To get further insight into $\hat{Y}_{greg}$ we note that one can write
 \begin{equation}\label{partition1}
 \pmb{A}_2=
\left[ \begin {array}{ll}
\pmb{A}^{(1)}_2 & \pmb{A}_2^{(12)} \\
 \pmb{A}_2^{(12)t} & \pmb{A}^{(2)}_2 \\ \end {array} \right]
\end{equation}
with $\pmb{A}^{(k)}_2=\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(k)}(x)\pmb{Z}^{(k)t}(x),\;k=1,2$
and $\pmb{A}_2^{(12)}=\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(1)}(x)\pmb{Z}^{(2)t}(x)$. Developing the normal equations accordingly one obtains after some algebra the well-known relation
 \begin{equation}\label{lsreduced1}
 \hat{\pmb{\beta}}^{(1)}_2=\hat{\pmb{\alpha}}_2-(\pmb{A}^{(1)}_2)^{-1}
 \pmb{A}_2^{(12)}\hat{\pmb{\beta}}^{(2)}_2
 \end{equation}
 where $\hat{\pmb{\beta}}^t_2=(\hat{\pmb{\beta}}_2^{(1)t},\hat{\pmb{\beta}}_2^{(2)t})$.
 Substituting in [\ref{ygreg2}] we obtain
 $$\hat{Y}_{greg}=\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\beta}}_2^{(1)}+
 \hat{\bar{\pmb{Z}}}^{(2)t}_1\hat{\pmb{\beta}}^{(2)}_2+\delta$$
 where $\delta=(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)})^t
 (\pmb{A}_2^{(1)})^{-1}\pmb{A}_2^{(1,2)}\hat{\pmb{\beta}}^{(2)}_2$, which is of order
 $O(n_1^{-\frac{1}{2}})$ in design-probability. Hence, we have the asymptotic equivalence
 \begin{equation}\label{equivygreg1}
\hat{Y}_{greg}\doteq \bar{\pmb{Z}}^{(1)t}\hat{\pmb{\beta}}_2^{(1)}+\hat{\bar{\pmb{Z}}}^{(2)t}_1\hat{\pmb{\beta}}^{(2)}_2
\end{equation}
which is intuitively very appealing: the exhaustive component $\pmb{Z}^{(1)}(x)$ occurs with its known true mean and the non-exhaustive component $\pmb{Z}^{(2)}(x)$ with its estimated mean from the large sample, as compared with the classical two-phase estimator
\begin{equation}\label{yregbeta}
\hat{Y}_{reg}= \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_2= \hat{\bar{\pmb{Z}}}_1^{(1)t}\hat{\pmb{\beta}}_2^{(1)}+
\hat{\bar{\pmb{Z}}}^{(2)t}_1\hat{\pmb{\beta}}^{(2)}_2
\end{equation}
One could also consider
\begin{equation}\label{ygregmod}
\hat{Y}_{gregmod}=
\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\beta}}_2^{(1)}+\hat{\bar{\pmb{Z}}}^{(2)t}_1\hat{\pmb{\beta}}^{(2)}_2
\end{equation}
\noindent as a further estimator in its own right. If $\pmb{A}_2^{(1,2)}=\pmb{0}$, i.e. if the exhaustive and non-exhaustive components are orthogonal ("independent") then we have exactly $\hat{Y}_{gregmod}=\hat{Y}_{greg}$. We do not advocate the use of $\hat{Y}_{gregmod}$ when $\pmb{Z}^{(1)}(x)$ and $\pmb{Z}^{(2)}(x)$ are non-orthogonal because, as we found out by simulations, $\hat{Y}_{gregmod}$ can have a larger variance than $\hat{Y}_{reg}$ with moderate sample sizes. Beside, the resulting formulae for the asymptotic covariance are more cumbersome than [\ref{estvarygregnew}].  For these reasons we shall not consider $\hat{Y}_{gregmod}$ any more and we now proceed to adapt the previous results to the important small-area estimation problem.

\section{Generalized small-area estimators}

We consider a small area $G\subset F$ and we want to estimate
$$\bar{Y}_G=\frac{1}{\lambda(G)}\sum_{i=1}^NI_G(i)Y_i=\frac{1}{\lambda(G)}\int_G Y(x)dx$$
where $I_G(i)=1 $ if the $i$-th tree is in $G$, otherwise $I_G(i)=0$. Strictly speaking the last equality holds if boundary adjustments are performed in $G$, whereas they are in most instances only performed with respect to $F$. We shall need the following notation: $s_{1,G}=s_1\cap G$, $s_{2,G}=s_2 \cap G$, $n_{k,G}=\sum_{s\in{s_2}}I_G(x), \;k=1,2$ (restriction of the samples and sample sizes to $G$, note that the $n_{k,G}$ are random variables). The simplest solution is to restrict the generalized regression estimator [\ref{ygreg1}] to $G$, i.e. to consider \textbf{the generalized small-area estimator}
\begin{equation}\label{smallareaest1}
\hat{Y}_{G,greg}=\frac{1}{\lambda(G)}\int_G \hat{Y}_1(x)dx+
\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}(\hat{Y}(x)-\hat{Y}_1(x))+
\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}(Y(x)-\hat{Y}(x))
\end{equation}
 and treat the internal model as an external one to obtain the estimated conditional variance (i.e. given the $n_{k,G}$)
 \begin{equation}\label{extsmallareaestvariance1}
 \hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s}_{2,G}}
 (\hat{R}_1(x)-\hat{\bar{R}}_{1,G})^2
+ (1-\frac{n_{2,G}}{n_{1,G}})
\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_2}}(\hat{R}(x)-\hat{\bar{R}}_G)^2
\end{equation}
where $\hat{\bar{R}}_{1,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}R_1(x)$ and likewise for $\hat{\bar{R}}_G$. This variance estimate neglects the uncertainty of the regression coefficients but there is empirical evidence that this is acceptable in large samples (see \cite{mandallazreport1} for examples with $\hat{Y}_{reg}$).\\
We can rewrite $\hat{Y}_{G,greg}$ as
\begin{equation}\label{smallareaest2}
\hat{Y}_{G,greg}=(\bar{\pmb{Z}}_G^{(1)}-\hat{\bar{\pmb{Z}}}_G^{(1)})^t\hat{\pmb{\alpha}}_2
+\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_2+\frac{1}{n_{2,G}}\sum_{x\in{s}_{2,G}}\hat{R}(x)
\end{equation}
where we have set
$$\bar{\pmb{Z}}_G^{(1)}=\frac{1}{\lambda(G)}\int_G \pmb{Z}^{(1)}(x)dx,\;\;
\hat{\bar{\pmb{Z}}}_{1,G}=\frac{1}{n_{1,G}}\sum_{x\in{s}_{1,G}}\pmb{Z}(x)
=(\hat{\bar{\pmb{Z}}}^{(1)t}_{1,G},\hat{\bar{\pmb{Z}}}^{(2)t}_{1,G})^t$$
The essential difference with $\hat{Y}_{greg}=\hat{Y}_{F,greg}$ is that the mean residual term in [\ref{smallareaest2}] does no longer vanish in general, which makes the calculation of the variance very difficult. To bypass this difficulty we use the technique presented in \cite{mandallazreport1} by extending the model with the indicator variable $I_G(x)$ of the small area $G$, \textbf{which insures zero mean residual over $F$ and $G$}. We can include $I_G(x)$ in $\pmb{Z}^{(1)}(x)$ or $\pmb{Z}^{(2)}(x)$. It seems more natural to include it in the first component so that the zero mean residual properties will hold for both the reduced and the large model. Also, it is reasonable to assume that the perimeter and consequently the surface area of $G$ are known. We consider therefore the following extended models with auxiliary vectors:
$\pmb{\mathcal{Z}}^t(x)=(\pmb{\mathcal{Z}}^{(1)t}(x),\pmb{\mathcal{Z}}^{(2)t}(x))$,
where $\pmb{\mathcal{Z}}^{(1)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathcal{Z}}^{(2)t}(x)=\pmb{Z}^{(2)t}(x)$. To have a uniform notation throughout we also change the notation for the second component, i.e. we will use $\pmb{\mathcal{Z}}^{(2)}(x)$ instead of $\pmb{Z}^{(2)}(x)$ in this section. We have therefore the following set up

\begin{enumerate}
 \item
 The large extended model $M$
 $$ Y(x)=\pmb{\mathcal{Z}}(x)^t\pmb{\theta}+ \mathcal{R(}x)=\pmb{\mathcal{Z}}^{(1)t}(x)\pmb{\theta}^{(1)}
 +\pmb{\mathcal{Z}}^{(2)t}(x)\pmb{\theta}^{(2)}+ \mathcal{R}(x)$$
 \noindent with $\pmb{\theta}^t=({\pmb{\theta}^{(1)t}}, {\pmb{\theta}^{(2)t}})$.
 The intercept term is contained in $\pmb{\mathcal{Z}}^{(1)}(x)$ or it is a linear combination of its components.\\
 The theoretical regression parameter $\pmb{\theta}$ minimizes
 $$\int_F (Y(x)-\pmb{\mathcal{Z}}^t(x)\pmb{\theta})^2dx$$
 It satisfies the normal equation
 $$\big(\int_F \pmb{\mathcal{Z}}(x)\pmb{\mathcal{Z}}^t(x)dx\big)\pmb{\theta}
 =\int_F(Y(x)\pmb{\mathcal{Z}}(x)dx)$$ and the orthogonality relationship
 $$\int_F \mathcal{R}(x)\pmb{\mathcal{Z}}(x)dx=\pmb{0}$$
 \noindent in particular the zero mean residual properties
 $$ \frac{1}{\lambda(F)}\int_F \mathcal{R}(x)dx=\frac{1}{\lambda(G)}\int_G \mathcal{R}(x)dx=0$$

 \item
 The reduced extended model $M_1$
 $$ Y(x)=\pmb{\mathcal{Z}}^{(1)t}(x)\pmb{\gamma} + \mathcal{R}_1(x)$$
 The theoretical regression parameter $\pmb{\gamma}$ minimizes
 $$\int_F (Y(x)-\pmb{\mathcal{Z}}^{(1)t}(x)\pmb{\gamma})^2dx$$
 It satisfies the normal equation
 $$\big(\int_F \pmb{\mathcal{Z}}^{(1)}(x)\pmb{\mathcal{Z}}^{(1)t}(x)dx\big)\pmb{\gamma}
 =\int_F Y(x)\pmb{\mathcal{Z}}^{(1)}(x)dx$$
 and the orthogonality relationship
 $$\int_G R_1(x)\pmb{\mathcal{Z}}^{(1)}(x)dx=\pmb{0}$$
 \noindent in particular the zero mean residual properties
 $$\frac{1}{\lambda(F)}\int_F \mathcal{R}_1(x)dx=\frac{1}{\lambda(G)}\int_G \mathcal{R}_1(x)dx=0$$
\end{enumerate}
We can obviously apply mutatis mutandis all the previous results. The estimated regression coefficients are
\begin{eqnarray}\label{extcoeff1}
\hat{\pmb{\gamma}}_2 &=& \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathcal{Z}}^{(1)}(x)\pmb{\mathcal{Z}}^{(1)t}(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}^{(1)}(x)\nonumber\\
&:=&(\pmb{\mathcal{A}}^{(1)}_2)^{-1}
\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}^{(1)}(x)
\end{eqnarray}
and
\begin{equation}\label{extcoeff2}
\hat{\pmb{\theta}}_{2}=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathcal{Z}}(x)\pmb{\mathcal{Z}}^t(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}(x):
=\pmb{\mathcal{A}}^{-1}_2\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}(x)
\end{equation}
The estimated covariance matrices are according to [{\ref{robustvar1}] and [\ref{robustvar2}]
\begin{eqnarray}\label{robustvar3}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_2}&=&\pmb{\mathcal{A}}_2^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathcal{R}}^2(x)\pmb{\mathcal{Z}}(x)
\pmb{\mathcal{Z}}^t(x)\Big)\pmb{\mathcal{A}}_2^{-1} \nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_2}&=&(\pmb{\mathcal{A}}^{(1)}_1)^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathcal{R}}_1^2(x)\pmb{\mathcal{Z}}(x)
\pmb{\mathcal{Z}}^t(x)\Big)(\pmb{\mathcal{A}}^{(1)}_1)^{-1} \nonumber \\
\end{eqnarray}
\noindent
where $\hat{\mathcal{R}}(x)=Y(x)-\pmb{\mathcal{Z}}^t(x)\hat{\pmb{\theta}}_2$ and
$\hat{\mathcal{R}}_1(x)=Y(x)-\pmb{\mathcal{Z}}^{(1)t}(x)\hat{\pmb{\gamma}}_2$ are the residuals. \\

\noindent Because the sum of the residuals over $s_{2,G}$ is now zero we can write the new small-area estimator $\hat{\tilde{Y}}_{G,greg}$ as in [\ref{ygreg2}]

\begin{equation}\label{extsmallareaest1}
\hat{\tilde{Y}}_{G,greg}
= (\bar{\pmb{\mathcal{Z}}}_G^{(1)}-\hat{\bar{\pmb{\mathcal{Z}}}}_G^{(1)})^t\hat{\pmb{\gamma}}_2 +
\hat{\bar{\pmb{\mathcal{Z}}}}_G^t\hat{\pmb{\theta}}_{2}
\end{equation}
where we have set
 $$\bar{\pmb{\mathcal{Z}}}_G^{(1)}=\frac{1}{\lambda(G)}\int_G \mathcal{Z}(x)dx,\; \hat{\bar{\pmb{\mathcal{Z}}}}_G^{(1)}=
 \frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\pmb{\mathcal{Z}}^{(1)}(x)$$
 To get an estimate of the design-based variance we use mutatis mutandis [\ref{estvarygregnew}]

 \begin{equation}\label{estvarygreg}
\hat{\VAR}(\hat{\tilde{Y}}_{greg})=\frac{n_2}{n_1}\bar{\pmb{\mathcal{Z}}}_G^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_2}
\bar{\pmb{\mathcal{Z}}}_G^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathcal{Z}}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_2}
\hat{\bar{\pmb{\mathcal{Z}}}}_G
\end{equation}

\section{Generalization to cluster sampling}
We follow the description of cluster sampling as defined in \cite{mandallaz} (especially section 5.5) and \cite{mandallazreport1}. A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$. The geometry of the cluster is given by $M$ vectors $e_1,\ldots e_M$ defining the random cluster $x_l=x+e_l$. $M(x)=\sum_{l=1}^MI_{F}(x_l)$ is the random number of points of the cluster falling into the forest area $F$. We define the local density at the cluster level by $Y_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)Y(x_l)}{M(x)}$, likewise we set $\pmb{Z}_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)\pmb{Z}(x_l)}{M(x)}$. The set $\tilde{F}$ above can be mathematically defined as the smallest set $\{x\in{\mathcal{R}}^2 \mid M(x) \ne 0 \}$. In the first phase we have $n_1$ clusters identified by $x\in{s_1}$ and in the second phase $n_2$ clusters with $x\in{s_2}$, obtained by simple random sampling from $s_1$.\\
We shall use the model-based approach, in which the regression coefficient $\pmb{\beta}_c$ at the cluster level, under the large model with $\pmb{Z}^t=(\pmb{Z}^{(1)t}(x), \pmb{Z}^{(2t)}(x))$,  minimizes
$$\int_F M(x)(Y_c(x)-\pmb{\beta}_c^{t}\pmb{Z}_c(x))^2dx$$
In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (thought close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details). The choice of $M(x)$ rather than $M^2(x)$ as weights is suggested by the model-dependent approach. When $Y_c(x)$ is the mean of the
$M(x)$ observations, its variance can be expected to be inversely
proportional to $M(x)$.  This procedure  leads to the normal equation
$$\Big(\int_F M(x)\pmb{Z_c}(x)\pmb{Z_c}^t(x)dx\Big)\pmb{\beta}_c=\int_F M(x)Y_c(x)\pmb{Z_c}(x)dx$$
and to $\int_F M(x)R_c(x)=0$. An asymptotically design-unbiased estimate $\hat{\pmb{\beta}}_{c,2}$  for $\pmb{\beta}_c$ can be obtained  by taking a sample copy of the above equation, i.e.
 \begin{eqnarray}\label{estbetacluster1}
 \hat{\pmb{\beta}}_{c,2}&=&
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}_c(x)\pmb{Z}_c^t(x)\Big)^{-1}
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \nonumber\\
 &:=&\pmb{A}^{-1}_{c,s_2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \end{eqnarray}
 The empirical residuals at the cluster level are
 $$\hat{R}_{c}(x)=Y_c(x)-\pmb{Z}^t_c(x)\hat{\pmb{\beta}}_{c,2}$$
 which satisfy the orthogonality relation
 $$\sum_{x\in{s_2}}M(x)\hat{R}_c(x)\pmb{Z}_c(x)=0$$
 and in particular the zero mean residual property
$$\frac{\sum_{x\in{s_2}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_2}}M(x)}=0$$
With obvious notational changes we get the corresponding results under the reduced model with
$\pmb{Z}^{(1)}$ alone, where the regression coefficient $\pmb{\alpha}_c$ minimizes
$$\int_F M(x)(Y_c(x)-\pmb{\alpha}_c^{t}\pmb{Z}^{(1)}_c(x))^2dx$$
with estimate
\begin{eqnarray}\label{estalphacluster1}
 \hat{\pmb{\alpha}}_{c,2}&=&
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}^{(1)}_c(x)\pmb{Z}_c^{(1)t}(x)\Big)^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \nonumber\\
 &:=&(\pmb{A}^{(1)}_{c,2})^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}^{(1)}_c(x)\Big)
 \end{eqnarray}
 The residuals for the reduced model are $\hat{R}_{1,c}(x)=Y_c(x)-\pmb{Z}^{(1)t}_c(x)\hat{\pmb{\alpha}}_{c,2}$ and enjoy the same properties as
 $\hat{R}_{c}(x)$.\\
 \noindent Using mutatis mutandis exactly the same arguments as in simple random sampling we get the asymptotic robust design-based estimated covariance matrices
\begin{eqnarray}\label{estcovmatrixcluster1}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,2}}&=&\pmb{A}^{-1}_{c,2}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}^2_c(x)
\pmb{Z}_c(x)\pmb{Z}^t_c(x)\Big)\pmb{A}^{-1}_{c,2}\nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{c,2}}&=&(\pmb{A}^{(1)}_{c,1})^{-1}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}_{1,c}^2(x)
\pmb{Z}_c(x)\pmb{Z}^t_c(x)\Big)(\pmb{A}^{(1)}_{c,1})^{-1}
\end{eqnarray}
where $\pmb{A}^{(1)}_{c,1}=\frac{1}{n_1}\sum_{x\in{s_1}}M(x)\pmb{Z}^{(1)}_c(x)\pmb{Z}_c^{(1)t}(x)$.\\

We define the \textbf{generalized regression estimator in cluster sampling} directly by adapting [\ref{ygreg2}]
\begin{equation}\label{ygregcluster}
\hat{Y}_{c,greg}= (\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_{c,1})^t\hat{\pmb{\alpha}}_{c,2} +
\hat{\bar{\pmb{Z}}}^t_{c,1}\hat{\pmb{\beta}}_{c,2}
\end{equation}
where the mean is now defined as
$$\hat{\bar{\pmb{Z}}}^t_{c,1}=\frac{\sum_{x\in{s_1}}M(x)\pmb{Z}_c(x)}{\sum_{x\in{s_1}}M(x)}$$
and similarly for $\hat{\bar{\pmb{Z}}}^{(1)}_{c,1}$.\\
With the same technique as in simple random sampling we obtain the estimated design-based variance
\begin{equation}\label{estvarclusterygereg}
\hat{\VAR}(\hat{Y}_{c,greg})=\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{c,2}}
\bar{\pmb{Z}}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{c,1}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,2}}
\hat{\bar{\pmb{Z}}}_{c,1}
\end{equation}
For small-area estimation we shall work with the extended model
$$\pmb{\mathcal{Z}}^t(x)=(\pmb{\mathcal{Z}}^{(1)t}(x), \pmb{\mathcal{Z}}^{(2)t}(x))$$
 with
$\pmb{\mathcal{Z}}^{(1)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$, $\pmb{\mathcal{Z}}^{(2)t}(x)=\pmb{Z}^{(2)t}(x)$ and $I_G(x)$ is the indicator of the small area $G$. At the cluster level we have $\pmb{\mathcal{Z}}_c^{(1)t}(x)=(\pmb{Z}_c^{(1)t}(x),I^t_{c,G}(x))$ where $I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}$. In extensive inventories we can reasonably assume that all the points of a cluster lying in the forest area $F$ will belong to the same small area $G$ so that in fact $I_{c,G}(x) \equiv 1$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$. This ensures again that we will have zero mean residual over $F$ and $G$.  For the regression estimates in the extended model we have

\begin{equation}\label{regbettaclusterext}
\hat{\pmb{\gamma}}_{c,2}=(\pmb{\mathcal{A}}_{c,2}^{(1)})^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)Y(x)\Big)
\end{equation}
with $\pmb{\mathcal{A}}_{c,2}^{(1)}=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)\pmb{\mathcal{Z}}^{(1)t}_c(x)$. Likewise we get
\begin{equation}\label{regbettaclusterext}
\hat{\pmb{\theta}}_{c,2}=\pmb{\mathcal{A}}^{-1}_{c,2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}_c(x)Y(x)\Big)
\end{equation}
with $\pmb{\mathcal{A}}_{c,2}=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)$.\\
We define as in [\ref{extsmallareaest1}] the generalized small-area estimator by

\begin{equation}\label{extclustersmallareaest1}
\hat{\tilde{Y}}_{c,G,greg}
= (\bar{\pmb{\mathcal{Z}}}_G^{(1)}-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,G}^{(1)})^t\hat{\pmb{\gamma}}_{c,2} +
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,G}^t\hat{\pmb{\theta}}_{c,2}
\end{equation}
where we have set
 $$\bar{\pmb{\mathcal{Z}}}_G^{(1)}=\frac{1}{\lambda(G)}\int_G \mathcal{Z}(x)dx,\; \hat{\bar{\pmb{\mathcal{Z}}}}_{c,G}^{(1)}=
 \frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c^{(1)}(x)}{\sum_{x\in{s_{1,G}}}M(x)},\;
 \hat{\bar{\pmb{\mathcal{Z}}}}_{c,G}=
 \frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c(x)}{\sum_{x\in{s_{1,G}}}M(x)}$$

The estimated design-based covariance matrix are now
\begin{eqnarray}\label{gammathetaclusterextestvar}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{c,2}}&=&
(\pmb{\mathcal{A}}^{(1)}_{c,1})^{-1}\Big(\frac{1}{n^2_2}\sum_{\in{s_2}}M^2(x)\hat{\mathcal{R}}_{1,c}^2(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)\pmb{\mathcal{Z}}^{(1)t}_c(x)\Big)(\pmb{\mathcal{A}}^{(1)}_{c,1})^{-1}\nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,2}}&=&
\pmb{\mathcal{A}}^{-1}_{c,2}\Big(\frac{1}{n^2_2}\sum_{\in{s_2}}M^2(x)\hat{\mathcal{R}}_c^2(x)
\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)\Big)\pmb{\mathcal{A}}^{-1}_{c,2}
\end{eqnarray}
with the residuals in the extended models $\hat{\mathcal{R}}_{1,c}(x)=Y_c(x)-\pmb{\mathcal{Z}}^{(1)t}_c(x)\hat{\pmb{\gamma}}_{c,2}$ and
$\hat{\mathcal{R}}_{c}(x)=Y_c(x)-\pmb{\mathcal{Z}}^{t}_c(x)\hat{\pmb{\theta}}_{c,2}$ and
$\pmb{\mathcal{A}}_{c,1}=\frac{1}{n_1}\sum_{x\in{s_1}}M(x)\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)$.\\
We obtain as in [\ref{estvarygregnew}] the estimated design-based variance

\begin{equation}\label{extendedclustersmallestvariance}
\hat{\VAR}(\hat{Y}_{c,G,greg}) =\frac{n_2}{n_1}\bar{\pmb{\mathcal{Z}}}_G^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{c,2}}
\bar{\pmb{\mathcal{Z}}}_G^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathcal{Z}}}}_{c,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{G,1}
\end{equation}
As shown in \cite{mandallazreport1} it is straightforward to consider simultaneously several
small areas by extending the model with as many small area indicator variables.\\

\section{Generalization to two-stage sampling}
In many applications costs to measure the response variable $Y_i$ are high. For instance,
a good determination of the volume may require that one records
$DBH$, as well as the diameter at $7m$ above ground and total height
in order to utilize a three-way volume function. However, one could rely
on a coarser, but cheaper, approximation of the volume based only on
$DBH$. Nonetheless, it may be most sensible to assess those three
parameters only on a sub-sample of trees. We now briefly formalize this
simple idea, which is used in the Swiss National Forest Inventory. The reader is referred to (\cite{mandallaz}, section 4.4, 4.5, 5.4 and 9.5) for details. For each point $x\in{s_2}$ trees are drawn with
probabilities $\pi_i$. The set of selected trees is denoted by
$s_{2}(x)$. From each of the selected trees $i\in{s_{2}(x)}$ one
gets an approximation $Y_i^*$ of the exact value $Y_i$. From the
finite set $s_{2}(x)$ one draws a sub-sample
$s_{3}(x)\subset{s_{2}(x)}$ of trees by Poisson sampling. For each tree $i\in{s_{3}(x)}$
one then measures the exact variable $Y_i$. Let us now define the
second stage indicator variable
\begin{equation}
 J_i(x)=\begin{cases}&1 \text{ if $i\in s_{3}(x)$}\\
                      &0 \text{ if $i\not\in s_{3}(x)$}
         \end{cases}
\end{equation}

 \par To construct a good point estimate, we must have
 the residual $R_i=Y_i-Y_i^*$ which is known only for trees
 $i\in{s_{3}(x)}$. The generalized local density $Y^*(x)$ is defined
 according to    \index{Generalized local density}
 \begin{eqnarray}\label{gdens}
 Y^*(x)&=&\frac{1}{\LF}\left( \sum_{i=1}^N \frac{I_{i}(x)Y_i^*}{\pi_i} +
 \sum_{i=1}^N \frac{I_{i}(x)J_{i}(x)R_i}{\pi_{i}p_i}\right)\nonumber \\
 &=&\frac{1}{\LF}\left( \sum_{i\in{s_{2}(x)}} \frac{Y_i^*}{\pi_i}
 +\sum_{i\in{s_{3}(x)}} \frac{R_i}{\pi_{i}p_i}\right)
 \end{eqnarray}
 where the $p_i$ are the conditional inclusion probabilities for the the second stage sampling, i.e. $p_i=\PR(J_i(x)=1 \mid I_i(x)=1)$.
 It follows from general principles presented in (\cite{mandallaz}, sections 4.4 and 4.5) that one can use all the previous results by replacing everywhere the exact local densities $Y(x)$, or $Y(x_l)$ in cluster sampling, by the corresponding generalized local densities $Y^*(x)$ or $Y^*(x_l)$. The second-stage variance is automatically taken into account.

\section{Example: Post-stratification}\label{examples}

We consider the forested area $F$ embedded in a subset $\tilde{F}$, which in many national inventories is simply the whole country or part of it. Hence, $\tilde{F}\setminus F=F_0$ is the non-forested area. The forested area itself is partitioned in $L$ strata $F_k$, i.e. $F=\cup_{k=1}^L F_k$. We assume that the surface areas $\lambda(\tilde{F})$ and $\lambda(F_0)$ are known exactly, \textbf{whereas the surface areas of the strata $\lambda(F_k)$ for $k \ge 1$ are not}. The first-phase sampling consists of $n_1$ points, uniformly distributed in $\tilde{F}$ (set $s_1$), out of which $n_2$ points (set $s_2$) are selected by equal probability sampling without replacement (in practice one uses grids and sub-grids). The local density $Y(x)$ is set to zero whenever $x\in{F}_0$ (even if trees in $F$ could be selected from $x\in{F}_0$), and is calculated according to [\ref{truelocaldensity}] if $x\in{F}$ (i.e. with boundary adjustments if necessary). This ensures that $\int_{\tilde{F}}Y(x)dx=\int_{F}Y(x)dx=\sum_{i=1}^NY_i$. We define the (random) sample sizes $n_{1k}=\sum_{x\in{s}_1}I_{F_k}(x)$ ,
$n_{2k}=\sum_{x\in{s}_2}I_{F_k}(x)$ for $k\ge 0$ and $n_{1F}=\sum_{k=1}^Ln_{1k}$, $n_{2F}=\sum_{k=1}^L n_{2k}$.\\
For the reduced model we define the one dimensional explanatory vector according to $\pmb{Z}^{(1)}(x)=I_F(x)$. Trivial calculations lead to
$\hat{\pmb{\alpha}}=\hat{\bar{Y}}_F^t=\frac{1}{n_{2,F}}\sum_{x\in{s}_2}Y(x)$ (i.e. the empirical mean of $Y(x)$ in the forested area $F$). We get $\bar{\pmb{Z}}^{(1)}=p_F=\frac{\lambda(F)}{\lambda(\tilde{F})}$. For the g-weight one obtains
$g^{(1)}(x)=p_F\frac{n_1}{n_{1F}}I_F(x)$. For the predictions one obtains $\hat{Y}_1(x)=\hat{\bar{Y}}_FI_F(x)$. \\
For the large model, we define $\pmb{Z}^{(2)}(x)$ as the $L$ dimensional vector defined by the indicator variables of all strata in the forested area, i.e. $\pmb{Z}^{(2)}(x)=(Z_1(x),Z_2(x), \ldots Z_{L}(x))^t$, $Z_k(x)=I_{F_k}(x)$ for $k=1,2 \ldots L$. Note that all the components of $\pmb{Z}(x)$ are zero for $x\not\in{F}$. The $(L+1,L+1)$ matrix $\pmb{A}=\sum_{x\in{s}_1}\pmb{Z}(x)\pmb{Z}(x)^t$ is almost diagonal
\begin{equation*}
\pmb{A}= \left[ \begin {array}{cccccc}
{\it n_{2F}} &{\it n_{21}}&{\it n_{22}}& {\it n_{23} }&...& {\it n_{2L}}\\
\noalign{\medskip}{\it n_{21}}&{\it n_{21}}& $0$ & $0$  &...& $0$ \\
\noalign{\medskip}{\it n_{22}}& $0$ & {\it n_{22} } & $0$  & ...& $0$ \\
\noalign{\medskip} ...  & & ... & ... & ...& $0$ \\
\noalign{\medskip} {\it n_{2k} } & ... & $0$ & {\it n_{2k} }& ... & $0$ \\
\noalign{\medskip} ...  & ... & ... & ...& $0$ \\
\noalign{\medskip}{\it n_{2L}}  & $0$ & $0$ &... & $0$ & {\it n_{2L}} \end {array} \right]
\end{equation*}
but it is singular because the first column is the sum of the last $L$ columns. We set
$\pmb{\beta}=(\beta_0,\beta_1,\ldots \beta_L)^t$. A particular solution of the singular normal equation is easily found to be $\hat{\beta}_0=\hat{\bar{Y}}_F$ and $\hat{\beta}_k=\hat{\bar{Y}}_k-\hat{\bar{Y}}_F$
where $\hat{\bar{Y}}_k=\frac{1}{n_{2,k}}\sum_{x\in{F_k}}Y(x)$ (this corresponds to the solution of the standard one-way ANOVA). This leads to the intuitively obvious predictions $\hat{Y}(x)=0$ for $x\in{F}_0$ and $\hat{Y}(x)=\hat{\bar{Y}}_k$ for $x\in{F}_k$. Note also that all empirical residuals are zero outside $F$. \\
The standard regression estimate (with respect to $\tilde{F}$) based on the large model alone is easily found to be [\ref{gweight3}]
$$\hat{Y}_{reg}=\hat{Y}_{post,standard}=\sum_{k=1}^L \frac{n_{1k}}{n_1}\hat{\bar{Y}}_k=\frac{1}{n_2}\sum_{x\in{s}_2}g_2(x)Y(x)$$
Since this algebraic identity is true for an arbitrary density $Y(x)$ we have
\begin{equation}\label{gweightpost}
g_2(x)=\frac{n_2}{n_1}\frac{n_{1k}}{n_{2k}} \quad \text{for} \quad x\in{F}_k, \;k=0,1,2,\ldots
\end{equation}
Simple algebra yields then for [\ref{ygreg2}] the result
\begin{equation}\label{post1}
\hat{Y}_{greg}=\hat{Y}_{post,new}=(p_F-\hat{p}_F)\hat{\bar{Y}}_F+\sum_{k=1}^L\hat{p}_k\hat{\bar{Y}}_k
\end{equation}
where we have set $p_F=\frac{\lambda(F)}{\lambda(\tilde{F})}$, $\hat{p}_F=\frac{n_{1F}}{n_1}$ and
$\hat{p}_k=\frac{n_{1k}}{n_1}$. For $n_1,n_2 \rightarrow \infty$ this is obviously a consistent estimate of the density with respect to $\tilde{F}$, i.e. of $\bar{Y}_{\tilde{F}}=\frac{1}{\lambda(\tilde{F})}\sum_{i=1}^N Y_i$.\\
The standard post-stratified estimate, i.e. with respect to $\tilde{F}$ but without knowledge of the exact surface area $\lambda(F)$ and $\lambda(\tilde{F})$, is given by the second term only (see \cite{mandallaz}, section 5.2.1):
\begin{equation}\label{post2}
\hat{Y}_{post,standard}=\sum_{k=1}^L\hat{p}_k\hat{\bar{Y}}_k
\end{equation}
The new estimate [\ref{post1}] is intuitively appealing: if the number of points falling in the forested area $F$ is above or below its expected value, then the classical estimate is corrected accordingly.
Tedious but simple calculations and the approximations $ n_{2k}-1 \approx n_{2k}$ lead to the following estimated variances
\begin{equation}\label{estpostvar1}
\hat{\VAR}(\hat{Y}_{post,new})=\frac{1}{n_1n_2}\frac{p_F^2}{\hat{p}_F^2}(n_{2F}-1)\hat{\sigma}^2_F+
(1-\frac{n_2}{n_1})\sum_{k=1}^L\hat{p}^2_k\frac{\hat{\sigma}_k^2}{n_{2k}}
\end{equation}
where $p_k=\frac{\lambda(F_k)}{\lambda(\tilde{F})},\; \hat{p}_k=\frac{n_{1k}}{n_1}$,
$\hat{\sigma}_F^2=\frac{1}{n_{2F}-1}\sum_{x\in{s_2}\cap F} (Y(x)-\hat{\bar{Y}}_F)^2$.\\
The main advantage of the g-weight variance estimate is that the strata weights $p_k$ are estimated from the large sample and the contribution of the strata to the variance is inversely proportional to the sample sizes $n_{2k}$. Result [\ref{estpostvar1}] is similar to previous findings given in (\cite{mandallaz}, pp 84 and 107-108).\\
The asymptotic variance ($n_1,\; n_2 \rightarrow \infty$ ) is then
\begin{equation}\label{asymvarpostnew}
\hat{\VAR}(\hat{Y}_{post,new})=\frac{1}{n_1}p_F\sigma^2_F +
(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{k=1}^Lp_k\sigma^2_k
\end{equation}
where $\sigma^2_F=\frac{1}{\lambda(F)}\int_F (Y(x)-\bar{Y}_F)^2dx$ is the overall variance within $F$ and the $\sigma^2_k=\frac{1}{\lambda(F_k)}\int_{F_k} (Y(x)-\bar{Y}_k)^2dx$ (with the strata means $\bar{Y}_k=\frac{1}{\lambda(F_k)}$)  are the variances within strata. \\
To calculate the variance of $\hat{Y}_{post,standard}$ we use the external model assumption and equation [5.5.1] in \cite{mandallaz} to get
$$\VAR(\hat{Y}_{post,standard})=\frac{1}{n_1}\VAR_{x\in{\tilde{F}}}(Y(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_{x\in{\tilde{F}}}(R(x))$$
\noindent where $R(x)=0$ for $x\not\in{F}$ and $R(x)=Y(x)-\bar{Y}_k$ for $x\in{F_k}$.\\
Noting that $\bar{Y}_{\tilde{F}}=\bar{Y}=p_F\bar{Y}_F,\;\bar{R}_{\tilde{F}}=\bar{R}_F=0$ and writing $(Y(x)-\bar{Y})=(Y(x)-\bar{Y}_F+\bar{Y}_F-p_F\bar{Y}_F)$ we obtain after some algebra the asymptotic variance of the standard post-stratified estimate with respect to $\tilde{F}$ as:
\begin{equation}\label{postvar2}
\VAR(\hat{Y}_{post,standard})=\frac{1}{n_1}(p_F\sigma^2_F + p_F(1-p_F)\bar{Y}^2_F) +(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{k=1}^Lp_k\sigma_k^2 > \VAR(\hat{Y}_{post,new})
\end{equation}
Hence, as expected, using the exhaustive information $I_F(x)$ reduces the variance.\\
In practice forest inventories occasionally present totals over the entire region $\tilde{F}$ based on post-stratification with the non forested area $F_0=\tilde{F}\setminus F$ viewed as a further ordinary stratum, that is $\lambda(\tilde{F})$ is known, but neither $\lambda(F_0)$, not the $\lambda(F_k)$ are known. We have therefore to consider the following estimates of totals and their asymptotic variances:
\begin{eqnarray}\label{totals}
\hat{T}_{post,standard} &=& \lambda(\tilde{F})\hat{Y}_{post,standard} \nonumber \\
\VAR(\hat{T}_{post,standard}) &=& \lambda^2(\tilde{F})\Big(\frac{1}{n_1}(p_F\sigma^2_F + p_F(1-p_F)\bar{Y}^2_F) +(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{k=1}^Lp_k\sigma_k^2\Big) \nonumber \\
\hat{T}_{post,new} &=& \lambda(\tilde{F})\hat{Y}_{post,new} \nonumber \\
\VAR(\hat{T}_{post,new}) &=& \lambda^2(\tilde{F})\Big(\frac{1}{n_1}p_F\sigma^2_F  +(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{k=1}^Lp_k\sigma_k^2\Big)
\end{eqnarray}
 In the present framework, because $F$ is assumed to be known, we can also consider the conditional estimate based only on the $n_{1F}$ and $n_{2F}$ points falling into $F$, that is
 \begin{eqnarray}\label{postcond}
 \hat{Y}_{post,cond}&=&\sum_{k=L}^L\hat{\tilde{p}}_k\hat{\bar{Y}}_k \nonumber \\
 \VAR(\hat{Y}_{post,cond}) &=& \frac{1}{n_{1F}}\sigma^2_F+(1-\frac{n_{2F}}{n_{1F}})\frac{1}{n_{2F}}\sum_{k=1}^L\tilde{p}_k\sigma_k^2 \nonumber \\
 \hat{T}_{post,cond}&=& \lambda(F) \hat{Y}_{post,cond} \nonumber \\
 \VAR(\hat{T}_{post,cond}) &=& \lambda^2(F)\VAR(\hat{Y}_{post,cond})
 \end{eqnarray}
 where we have set $\hat{\tilde{p}}_k=\frac{n_{1k}}{n_{1F}}$ and $\tilde{p}_k=\frac{\lambda(F_k)}{\lambda(F)}$. Using the facts that $\tilde{p}_k=\frac{p_k}{p_F}$, $\lambda(F)=p_F\lambda(\tilde{F})$, $\EX(n_{1F})=n_1p_F$, $\EX(n_{2F})=p_F n_2)$, we see that in large samples $\hat{Y}_{post,cond}$ and
 $\hat{T}_{post,new}$ are equivalent and better than $\VAR(\hat{T}_{post,standard})$, in the sense that
 $$\VAR(\hat{T}_{post,new}) \approx  \VAR(\hat{T}_{post,cond}) < \VAR(\hat{T}_{post,standard}) $$
 This fact is surprisingly not widely known and many national inventories are using estimators related somehow to $\hat{T}_{post,standard}$. Usually the decision for $x\notin{F}$ is easy (implying necessarily $Y(x)=0$ ), whereas the delineation of the forested area can be more problematic (i.e. to decide for $x\in{F}$, even if $Y(x)=0$ is still possible). With the technological advances in remote sensing it is only a matter of time until we can assume that $F$ and $\lambda(F)$ are known with the same accuracy as $\tilde{F}$ and $\lambda(\tilde{F})$. From a pragmatic point of view and as far as the estimation of total is concerned a coarse delineation by polygons defining a set $F$ containing the "true" forest should suffice (so that $p_F$ is close to $1$ and the extra variance term $p_F(1-p_f)\bar{Y}_F^2$ is small). \\
 We emphasize again the fact that the surface areas of the strata within the forest area need not be known. In any case, the conditional estimator $\hat{Y}_{post,cond}$ and the new estimator
 $\hat{T}_{post,new}$ should be preferred  to the occasionally used practice with $\hat{T}_{post,standard}$, particularly if $\tilde{F}$ is much larger than $F$.
\newpage

\bibliography{biblio1}

\end{document}

### 2018-08-30 texfiles Daniel/techreportthreephaset.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

\documentclass[a4paper,12pt,leqno, titlepage]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\bibliographystyle{mystyle2}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\VAR}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{RE}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}
\newcommand{\R}{\mathbb{R}}
\setlength{\topmargin}{0cm}
\setlength{\headheight}{3mm}
\setlength{\headsep}{1cm}
\setlength{\topskip}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{22cm}
\setlength{\oddsidemargin}{0cm}
\makeatletter
\def\tagform@#1{\maketag@@@{[\ignorespaces#1\unskip\@@italiccorr]}}
\makeatother
\begin{document}
\doublespacing
\pagestyle{plain}
\pagenumbering{arabic}

\title{Regression estimators in forest inventories with three-phase sampling and two multivariate components of auxiliary information}
\author{Daniel Mandallaz,\thanks{Tel. ++41(0)44 6323186 e-mail
    daniel.mandallaz@env.ethz.ch, CHN K74.1, CH 8092 Zurich, Switzerland} \\ETH Zurich\\ Department of Environmental Systems Science\\ Chair of Land Use Engineering}
\date{}

\maketitle
\newpage

\begin{abstract}
We consider three-phase sampling schemes where one multivariate component of the auxiliary information is known in the very large sample of the so-called null phase ("almost exhaustive") and the second multivariate component is available only in the large sample of the first phase, whereas the second phase provides the terrestrial inventory data. We extend the generalized regression estimator with partially exhaustive information to this three-phase set-up, for both global and local estimation and derive its asymptotic design-based variance. We also consider further extensions to cluster and two-stage sampling.
\end{abstract}
\clearpage

\section{Introduction}\label{introduction}
\pagenumbering{arabic} \setcounter{page}{1}

The motivation for this work is due to the increasing need of using national or regional inventories for local estimation in order to meet tighter budgetary constraints, which is only feasible under extensive use of auxiliary information, provided e.g. by remote sensing (aerial photographs or LiDAr data). The small-area estimation problem is in this context of the utmost importance. This paper adapts the so-called generalized regression estimator proposed in \cite{mandallaz4} to the case where the first component of the auxiliary information is no longer exhaustive, but is provided by a very large sample, the null-phase. The second component is available on a sub-sample of the null-phase, the first-phase, and the terrestrial inventory is performed on a sub-sample of the first phase, the second-phase. This set-up is particularly useful for national or regional inventories for two reasons: (1) the first component may not be available exhaustively (2) even if it were it may be computationally prohibitive for some of its variables, particularly those based on sophisticated algorithms requiring single tree identification (as in \cite{mandallaz4}). The methodology and terminology rests upon the design-based Monte-Carlo approach to sampling theory for forest theory. The reader unfamiliar with this topic should first consult \cite{mandallaz3} for a short bibliographical review and a thorough but terse analysis of the regression estimator in a new setting, particularly useful in the small-area estimation context. The book \cite{mandallaz}, chapters 4 and 5, is recommended for a first perusal, as well as the on-line technical reports \cite{mandallazreport1,mandallazreport2} for the mathematical details. \\

\section{Methodology}
The \textbf{null phase} draws a very large sample $s_0$ of $n_0$ points
$x_{i}\in{s_0}$ ($i=1,2\ldots n_0$) that are independently and uniformly distributed
within the forest area $F$. At each of those points auxiliary
information is collected, very often coding information of qualitative nature
(e.g. following the  interpretation of aerial photographs) or quantitative (e.g. timber volume estimates  based on LiDAR measurements). We shall assume that the auxiliary information at point $x$
is described by the column vector $\pmb{Z}^{(1)}(x)\in{\Re^p}$. We shall call this component \textbf{pseudo-exhaustive} by analogy with the terminology used in \cite{mandallaz3}, which deals with the case $n_0=\infty$ (i.e. $\pmb{Z}^{(1)}(x)$ is in this case the \textbf{exhaustive} component). The \textbf{first phase} draws a large sample $s_1 \subset s_0$ of $n_1 <<n_0$ points by simple random sampling in $s_0$. Note that the points $x\in{s_1}$ are also uniformly independently distributed in $F$. For each point in the first phase a further component $\pmb{Z}^{(2)}(x)\in{\Re^q}$ of the auxiliary information is available and hence also the vector $\pmb{Z}^t(x)=(\pmb{Z}^{(1)t}(x),\pmb{Z}^{(2)t}(x))\in{\Re^{p+q}}$ (the upper index $t$ denotes the transposition operator). The \textbf{second phase} draws a small sample $s_2\subset{s_1}$ of
$n_2$ points from $s_1$ by simple random sampling and consists of the terrestrial inventory. Note that we have used the terms null, first and second phases instead of first, second and third phases simply to ensure compatibility of the terminology used in previous work, in which the null-phase was exhaustive.\\
The reason for introducing the null phase sample $s_0$ is that the component $\pmb{Z}^{(1)}(x)$ can be very computing intensive to calculate exhaustively (see \cite{mandallaz3} for a case study with LiDAR data).\\
In the forested area $F$ we consider a well defined population $ \mathcal{P}$ of $N$ trees with response variable
 $Y_i,\;i=1,2 \ldots$, e.g. the timber volume.  The objective is to estimate the spatial mean  $\bar{Y}=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i$, where $\lambda(F)$ denotes the surface area of $F$ (usually in ha). For each point $x\in{s_2}$ trees are drawn from the population $\mathcal{P}$ with probabilities $\pi_i$, for instance with concentric circles or angle count techniques. The
set of trees selected at point $x$ is denoted by $s_{2}(x)$. From each of the
selected trees $i\in{s_{2}(x)}$ one determines $Y_i$. The indicator variable $I_i$ is defined as
\begin{equation}\label{1stage}
 I_i(x)=\begin{cases}&1 \text{ if $i\in s_{2}(x)$}\\
                      &0 \text{ if $i\not\in s_{2}(x)$}
         \end{cases}
\end{equation}
At each point $x\in{s_2}$ the
terrestrial inventory provides the local density $Y(x)$
\begin{equation}\label{truelocaldensity}
 Y(x) =\frac{1}{\lambda(F)}\sum_{i=1}^N \frac{I_i(x)Y_i}{\pi_i}=\frac{1}{\lambda(F)}\sum_{i\in{s}_2(x)} \frac{Y_i}{\pi_i}
 \end{equation}
 The term $\frac{1}{\lambda(F)\pi_i}$ is the tree extrapolation factor $f_i$ with dimension $ha^{-1}$. Because of possible boundary adjustments $\lambda(F)\pi_i=\lambda(F \cap K_i)$, where $K_i$ is the inclusion circle of the $i$-th tree. In the infinite population or Monte Carlo approach one samples the function $Y(x)$ and we have $\EX_{x} (Y(x))=\frac{1}{\lambda(F)}\int_{F} Y(x)dx=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i=\bar{Y}$, where $\EX_x$ denotes the expectation with respect to a random point $x$ uniformly distributed in $F$.
 \section{The models}
 We shall work with the following linear models (see \cite{mandallazreport2} for more details)
 \begin{enumerate}
 \item
 The large model $M$
 \begin{equation}\label{largemodel}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R(x)=\pmb{Z}^{(1)t}(x)\pmb{\beta}^{(1)}
 +\pmb{Z}^{(2)t}(x)\pmb{\beta}^{(2)}+ R(x)=:\hat{Y}_{theo}(x)+R(x)
 \end{equation}
 \noindent with $\pmb{\beta}^t=({\pmb{\beta}^{(1)t}}, {\pmb{\beta}^{(2)t}})$ and the theoretical predictions
 $\hat{Y}_{theo}(x)=\pmb{Z}^t(x)\pmb{\beta}$.\\
 The intercept term is contained in $\pmb{Z}^{(1)}(x)$ or it is a linear combination of its components.\\
 The theoretical regression parameter $\pmb{\beta}$ minimizes
 $\int_F (Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2dx$, it satisfies the normal equation
 $\big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\big)\pmb{\beta}=\int_F Y(x)\pmb{Z}(x)dx$ and the orthogonality relationship
 $\int_F R(x)\pmb{Z}(x)dx=\pmb{0}$, in particular the zero mean residual property
 $\frac{1}{\lambda(F)}\int_F R(x)dx=0$.
 \item
 The reduced model $M_1$
 \begin{equation}\label{reducedmodel}
 Y(x)=\pmb{Z}^{(1)t}(x)\pmb{\alpha} + R_1(x)=:\hat{Y}_{theo,1}(x)+ R_1(x)
 \end{equation}
 The theoretical regression parameter $\pmb{\alpha}$ minimizes
 $\int_F (Y(x)-\pmb{Z}^{(1)t}(x)\pmb{\alpha})^2dx$. It satisfies the normal equation
 $\big(\int_F \pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)dx\big)\pmb{\alpha}=\int_FY(x)\pmb{Z}^{(1)}(x)dx$ and the orthogonality relationship
 $\int_F R_1(x)\pmb{Z}^{(1)}(x)dx=\pmb{0}$, in particular the zero mean residual property $ \frac{1}{\lambda(F)}\int_F R_1(x)dx=0$. $\hat{Y}_{theo,1}(x)=\pmb{Z}^{(1)t}(x)\pmb{\alpha}$ are the theoretical predictions.
\end{enumerate}
Let us emphasize the fact that in this paper we consider only the properties of estimators in the \textbf{design-based} paradigm and that we do not assume the above models to be correct in the sense of \textbf{model-dependent} inference
(see \cite{mandallaz3}).

\section{The generalized regression estimator}
We consider the following design-based least squares estimators of the regression coefficients of the reduced model, which are solutions of sample copies of the normal equations
\begin{eqnarray}\label{coeff1}
\hat{\pmb{\alpha}}_k &=& \Big(\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)
\Big)^{-1}\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)\nonumber\\
&:=&(\pmb{A}^{(1)}_k)^{-1}
\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)=:(\pmb{A}^{(1)}_k)^{-1}\pmb{U}^{(1)}_k,\quad  k=0,1,2
\end{eqnarray}
Likewise for the large large model we set
\begin{eqnarray}\label{coeff2}
\hat{\pmb{\beta}}_{k}&=&\Big(\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}(x)\pmb{Z}^t(x)
\Big)^{-1}\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}(x)\nonumber \\
&=&\pmb{A}^{-1}_k\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}(x)=:\pmb{A}^{-1}_k\pmb{U}_k,\quad k=0,1,2
\end{eqnarray}
Note that only $\hat{\pmb{\alpha}}_2$ and $\hat{\pmb{\beta}}_2$ are observable and that
in general the vector consisting of the first $p$ components of $\hat{\pmb{\beta}}_{2}$ is not equal to $\hat{\pmb{\alpha}}_2$ (they are if the corresponding explanatory variables are orthogonal in the classical least squares sense).\\
The large model yields the predictions $\hat{Y}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{2}$ and the reduced model the predictions $\hat{Y}_1(x)=\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2$.\\
The \textbf{generalized regression estimate} introduced and discussed by \cite{mandallazreport2,mandallaz3} is defined as
\begin{equation}\label{ygreg1}
\hat{Y}_{F,greg} = \frac{1}{\lambda(F)}\int_F \hat{Y}_1(x)dx
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}(x)-\hat{Y}_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x))
\end{equation}
This estimator is the Monte Carlo version of S\"{a}rndal's  regression estimator for two-phase sampling in finite population (see \cite{sarndal}, equation 9.7.20). It is clear by the law of large numbers that  $\hat{\pmb{\beta}}_2$ and $\hat{\pmb{\alpha}}_2$ are asymptotically design-unbiased estimators of $\pmb{\beta}$ and $\pmb{\alpha}$. This implies at once that $\EX_{1,2}\hat{Y}_{greg}= \EX_1\EX_{2 \mid 1}\hat{Y}_{greg} \approx \bar{Y}$ (with $\EX_{2 \mid 1}$ denoting the conditional expectation of the second phase given the first phase, i.e. simple random sampling without replacement in the set $s_1$), and $\EX_1$ denoting the expectation with respect to uniformly distribution points of the first phase (i.e. to $\EX_x$). The generalized regression estimate is therefore asymptotically design-unbiased.\\
Under the \textbf{external model assumption} the design-based variance is given by
\begin{equation}\label{externalvarYgreg}
\VAR(\hat{Y}_{F,greg})=\frac{1}{n_1}\VAR_x(R_1(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_x(R(x))
\end{equation}
This should be compared with the standard result for the variance of the regression estimator $\hat{Y}_{reg}$ under the large model
\begin{equation}\label{classicreg}
\hat{Y}_{F,reg}=\frac{1}{n_1}\sum_{x\in{s}_1}\hat{Y}(x)+\frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x))
\end{equation}
whose theoretical variance is given by
\begin{equation}\label{externalvarYreg}
\VAR(\hat{Y}_{F,reg})=\frac{1}{n_1}\VAR_x(Y(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_x(R(x))
\end{equation}
Thus, by using the exhaustive information the variance of the observations in [\ref{externalvarYreg}] is replaced by the variance of the residuals under the reduced model, a very nice and intuitive result indeed.\\
The generalized regression estimator can be rewritten as
\begin{eqnarray}\label{ygreg2}
\hat{Y}_{F,greg}&=&(\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 + (\hat{\bar{\pmb{Z}}}_1-\hat{\bar{\pmb{Z}}}_2)^t\hat{\pmb{\beta}}_{2}+\frac{1}{n_2}
\sum_{x\in{s}_2}Y(x) \nonumber \\
&=& (\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2}
\end{eqnarray}
Consistent estimates of the design-based variance of the regression coefficients are given by
\begin{equation}\label{robustvar1}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}=\pmb{A}_2^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}_2^{-1}
\end{equation}
and
\begin{equation}\label{robustvar2}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}=(\pmb{A}^{(1)}_1)^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)}_1)^{-1}
\end{equation}
with the empirical residuals $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_{2}$ and $\hat{R_1}(x)=Y(x)-\pmb{Z}^{1)t}(x)\hat{\pmb{\alpha}}_2$ (see section \ref{proofofvariance} in the Appendix for a short review and \cite{mandallaz},pp 124-125 for details).\\
Furthermore
\begin{equation}\label{estvarygregnew}
\hat{\VAR}(\hat{Y}_{F,greg})=\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}
\bar{\pmb{Z}}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}
\hat{\bar{\pmb{Z}}}_1
\end{equation}
\noindent is a consistent estimate of the \textbf{design-based variance} of $\hat{Y}_{F,greg}$ (see \cite{mandallazreport2} for the proof).\\
The g-weights are defined as
\begin{eqnarray}\label{gweightygreg1}
g_2(x) &=& \hat{\bar{\pmb{Z}}}_1^t\pmb{A}_2}^{-1}{\pmb{Z}(x)\nonumber \\
g^{(1)}_1(x) &=& \bar{\pmb{Z}}^{(1)t}(\pmb{A}^{(1)}_1)^{-1}\pmb{Z}^{(1)}(x)
\end{eqnarray}
They enjoy the calibration properties
\begin{eqnarray}\label{gweightygreg2}
\frac{1}{n_2}\sum_{x\in{s}_2}g_2(x)\pmb{Z}(x)&=&\hat{\bar{\pmb{Z}}}_1 \nonumber \\
\frac{1}{n_1}\sum_{x\in{s}_1}g^{(1)}_1(x)\pmb{Z}^{(1)}(x)&=&\bar{\pmb{Z}}^{(1)}
\end{eqnarray}
The variance [\ref{estvarygregnew}] can be rewritten as
\begin{equation}\label{estvarygreggweights}
\hat{\VAR}(\hat{Y}_{F,greg})=\hat{\VAR}(\hat{Y}_{F,greg})=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(g_1^{(1)}(x))^2\hat{R}_1^2(x)+
\frac{1}{n_2}(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{x\in{s}_2}g_2^2(x)\hat{R}^2(x)
\end{equation}
Under the external model assumptions it is straightforward to obtain approximate estimates of the design-based variances of $\hat{Y}_{F,reg}$ and $\hat{Y}_{F,greg}$ by using the standard variance estimates in [\ref{externalvarYgreg}] and [\ref{externalvarYreg}] after replacing the theoretical residuals by the empirical ones.
The generalized regression estimator $\hat{Y}_{G,greg}$ for any small-area $G\subset F$ has been discussed in \cite{mandallazreport2, mandallaz4}. $\hat{Y}_{F,greg}$ and $\hat{Y}_{G,greg}$ can be viewed as limit cases of the estimators presented in the next two sections.

\section{The generalized regression estimator in three-phase sampling}
If $\pmb{Z}^{(1)}(x)$ is no longer exhaustive we replace the first term in [\ref{ygreg1}] by its sample mean in the null phase sample and we define the new three-phase estimator as
\begin{equation}\label{y0greg1}
\hat{Y}_{F,g3reg} = \frac{1}{n_0} \sum_{x\in{s_0}} \hat{Y}_1(x)
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}(x)-\hat{Y}_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x))
\end{equation}
By using the properties of conditional expectations and variances (see \cite{mandallaz}, appendix B) we have
\begin{equation}\label{condexp1}
\EX_{0,1,2}(\hat{Y}_{F,g3reg})=\EX_0 \EX_{1 \mid 0}\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})
\end{equation}
and
\begin{eqnarray}\label{condvar1}
\VAR_{0,1,2}(\hat{Y}_{F,g3reg})&=&\EX_0 \EX_{1 \mid 0}\VAR_{2 \mid 0,1}(\hat{Y}_{F,g3reg})+
\EX_0 \VAR_{1 \mid 0}\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})
+\VAR_0 \EX_{1 \mid 0}\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})\nonumber \\
&=:& T_1 + T_2 + T_3
\end{eqnarray}
Under the external model assumption it follows from [\ref{condvar1}] after some algebra that $\hat{Y}_{F,g3reg}$ is exactly unbiased with variance
\begin{equation}\label{varexterny0greg}
\VAR_{0,1,2}(\hat{Y}_{F,g3reg})=\frac{1}{n_0}\VAR_{x}(Y(x))+(1-\frac{n_1}{n_0})\frac{1}{n_1}\VAR_x(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x))
\end{equation}
If the predictions $\hat{Y}_1(x)$ and the residuals $R_1(x)$ are orthogonal in the design-based sense (this is the case for their theoretical versions) then, because $\VAR_x(Y(x))=\VAR_x(\hat{Y}_1(x))+\VAR_x(R_1(x))$, one obtains
\begin{equation}\label{varexterny0gregbis}
\VAR_{0,1,2}(\hat{Y}_{F,g3reg})=\frac{1}{n_0}\VAR_{x}(\hat{Y}_1(x))+\frac{1}{n_1}\VAR_x(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x))
\end{equation}
Let us now define
\begin{equation}\label{meanvalues}
\hat{\bar{\pmb{Z}}}^{(1)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(1)}(x)dx ,\quad \hat{\bar{\pmb{Z}}}^{(1)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(1)}(x) ,
\quad \hat{\bar{\pmb{Z}}}_k=\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}(x),\;k=1,2
\end{equation}
and rewrite $\hat{Y}_{F,g3reg}$ as

\begin{eqnarray}\label{0,ygreg2}
\hat{Y}_{F,g3reg}&=&(\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 + (\hat{\bar{\pmb{Z}}}_1-\hat{\bar{\pmb{Z}}}_2)^t\hat{\pmb{\beta}}_{2}+\frac{1}{n_2}
\sum_{x\in{s}_2}Y(x) \nonumber \\
&=& (\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2}
\end{eqnarray}
The last equation follows from the fact that the sum of the residuals is zero by construction.\\
We now calculate the variance of [\ref{0,ygreg2}]  by using [\ref{robustvar1},\ref{robustvar2},\ref{condvar1}]. We begin with the easiest part, namely the last term $T_3$ in [\ref{condvar1}]. First, we obtain the asymptotic equivalence
\begin{equation}\label{equivalenz1}
\EX_{2\mid 0,1}(\hat{Y}_{F,g3reg})\approx ( \hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_1 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{1}\approx \hat{\bar{\pmb{Z}}}^{(1)t}_0\hat{\pmb{\alpha}}_{1}
\end{equation}
The last equation results from the fact that both $\hat{\bar{\pmb{Z}}}^{(1)t}_1\hat{\pmb{\alpha}}_{1}$ and $
\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\beta}}_{1}$ tend to $\bar{Y}$.\\
 We then get
$\EX_{1 \mid 0}\EX_{2\mid 0,1}(\hat{Y}_{F,g3reg})\approx \hat{\bar{\pmb{Z}}}^{(1)t}_0\hat{\pmb{\alpha}}_0$.
From [\ref{appendixhatt3}] in the Appendix (section \ref{proofofvariance}) we get the asymptotically consistent estimate of
 $T_3\approx \VAR_0(\hat{\bar{\pmb{Z}}}^{(1)t}_0\hat{\pmb{\alpha}}_0)$
\begin{equation}\label{hatt3}
\hat{T}_3=\frac{n_2}{n_0}\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}\hat{\bar{\pmb{Z}}}_0^{(1)}
+ \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}_2
\end{equation}
where
\begin{equation}\label{estcovarz}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}=
\frac{1}{n_{0}}\frac{\sum_{x\in{s_{0}}}(\pmb{Z}^{(1)}(x)-
\hat{\bar{\pmb{Z}}}^{(1)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(1)}_{0})^t}{n_0-1}
\end{equation}
\noindent The calculations for $T_2$ and $T_1$ are more complicated and rely on Taylor expansions for the regression coefficients. Details are given in section \ref{proofofvariance} of the Appendix.\\
According to [\ref{appendixhatt2}] (section \ref{proofofvariance} of the Appendix) we have the following asymptotically consistent estimate of $T_2$
\begin{equation}\label{hatt2}
\hat{T}_2=\frac{n_2}{n_1}(1-\frac{n_1}{n_0})\hat{\bar{\pmb{Z}}}^{(1)t}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(1)}_0
\end{equation}
\noindent and according to [\ref{appendixhatt1}] the following asymptotically consistent estimate of $T_1$
\begin{equation}\label{hatt1}
\hat{T}_1 \approx (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}^{t}_1
\end{equation}
and consequently the following asymptotically consistent estimate of the overall three-phase variance ([\ref{estvary0greg2}], section \ref{proofofvariance} of the Appendix)
\begin{equation}\label{estvarianceY0greg}
\hat{\VAR}_{0,1,2}(\hat{Y}_{F,g3reg})=\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}_2
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(1)t}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(1)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}^{t}_1
\end{equation}
\noindent By comparing with [\ref{estvarygregnew}] we see that in large samples we have
\begin{equation}\label{comparison1}
\hat{\VAR}(\hat{Y}_{F,g3reg})- \hat{\VAR}(\hat{Y}_{F,greg})=
\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}_2
=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1}
\end{equation}
\noindent where we have set $\hat{Y}_1(x)=\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2$ and
$\hat{\bar{Y}}_1=\frac{1}{n_0}\sum_{x\in{s_0}}\hat{Y}_1(x)$. Hence, the extra term in the variance due to three-phase sampling is given by the variance of the predictions of the reduced model in the very large null-phase sample, which can be made very small while cutting down significantly the computing time.
The g-weights are defined as
\begin{eqnarray}\label{gweight1}
g(x) &=& \hat{\bar{\pmb{Z}}}_1^t\pmb{A}_2^{-1}\pmb{Z}(x)\nonumber \\
g^{(1)}(x) &=& \hat{\bar{\pmb{Z}}}_0^{(1)t}(\pmb{A}^{(1)}_1)^{-1}\pmb{Z}^{(1)}(x)
\end{eqnarray}
They satisfy the calibration properties
\begin{eqnarray}\label{gweight2}
\frac{1}{n_2}\sum_{x\in{s}_2}g(x)\pmb{Z}(x)&=&\hat{\bar{\pmb{Z}}}_1 \nonumber \\
\frac{1}{n_1}\sum_{x\in{s}_1}g^{(1)}(x)\pmb{Z}^{(1)}(x)&=&\hat{\bar{\pmb{Z}}}_0^{(1)}
\end{eqnarray}
\noindent Note the slights differences between [\ref{gweight1}], [\ref{gweight2}] and [\ref{gweightygreg1}],[\ref{gweightygreg2}].\\
It can be shown (see \cite{mandallazreport1, mandallazreport2}) that $g^{(1)}(x)$ and $g(x)$ tend to $1$ for $n_1,n_2 \to \infty$ and that
\begin{eqnarray*}
\frac{1}{n_1}\sum_{x_{s_1}}g^{(1)}(x)&=& 1  \\
\frac{1}{n_2}\sum_{x\in{s_2}}g(x)&=& 1\\
\end{eqnarray*}
\noindent One can rewrite [\ref{estvarianceY0greg}] as
\begin{eqnarray}\label{estvary0greggweights}
\hat{\VAR}(\hat{Y}_{F,g3reg})&=&\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}_2
\nonumber\\
&+&\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(g^{(1)}(x))^2\hat{R}_1^2(x)+
\frac{1}{n_2}(1-\frac{n_2}{n_1})\frac{1}{n_2}\sum_{x\in{s}_2}g^2(x)\hat{R}^2(x)
\end{eqnarray}
Under the external model assumptions one obtains from [\ref{varexterny0gregbis}] the following variance estimate
\begin{eqnarray}\label{extexternvary0greg}
\hat{\VAR}_{ext}(\hat{Y}_{F,g3reg})&=&\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1}
\nonumber\\
&+&\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_1(x)-\hat{\bar{R_1}})^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2
\end{eqnarray}
\noindent where $\hat{\bar{Y}}_1=\frac{1}{n_0}\sum_{x\in{s_0}}\hat{Y}_1(x)$, $\hat{\bar{R}}_1=\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_1(x)=0$ and
$\hat{\bar{R}}=\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)=0$.
Since the g-weights $g^{(1)}(x)$ and $g(x)$ tend to $1$ for $n_1,n_2 \to \infty$ we see that both variance estimates [\ref{estvary0greggweights}] and [\ref{extexternvary0greg}] are asymptotically equivalent. However, the g-weight version offers several statistical advantages from a theoretical point of view.
\section{Application to small-area estimation}

We consider a small area $G\subset F$ and we want to estimate
$$\bar{Y}_G=\frac{1}{\lambda(G)}\sum_{i=1}^NI_G(i)Y_i=\frac{1}{\lambda(G)}\int_G Y(x)dx$$
where $I_G(i)=1 $ if the $i$-th tree is in $G$, otherwise $I_G(i)=0$. Strictly speaking the last equality holds if boundary adjustments are performed in $G$, whereas they are in most instances only performed with respect to $F$. We shall need the following notation: $s_{0,G}=s_o \cap G$, $s_{1,G}=s_1\cap G$, $s_{2,G}=s_2 \cap G$, $n_{k,G}=\sum_{s\in{s_k}}I_G(x), \;k=0,1,2$. \\
The simplest procedure is to restrict the samples to $G$, i.e. to consider \textbf{the generalized small-area estimator}
\begin{equation}\label{smallareaest1}
\hat{Y}_{G,g3reg}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \hat{Y}_1(x)+
\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}(\hat{Y}(x)-\hat{Y}_1(x))+
\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}(Y(x)-\hat{Y}(x))
\end{equation}
 and treat the internal model as an external one to obtain from [\ref{varexterny0greg}] the variance estimate
 \begin{eqnarray}\label{extsmallareaestvariance1}
 \hat{\VAR}_{ext1}(\hat{Y}_{G,g3reg})&=& \frac{1}{n_{0,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(Y(x)-\hat{\bar{Y}}_G)^2 \nonumber \\
 &+& (1-\frac{n_{1,G}}{n_{0,G}}){\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
 (\hat{R}_1(x)-\hat{\bar{R}}_{1,G})^2 \nonumber \\
 &+& (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\hat{\bar{R}}_G)^2
 \end{eqnarray}
where $\hat{\bar{Y}}_G=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}}Y(x)$, $\hat{\bar{R}}_{1,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}_1(x)$ and $\hat{\bar{R}}_G=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)$. This variance estimate neglects the uncertainty of the regression coefficients but there is empirical evidence that this is acceptable in large samples (see \cite{mandallaz3} for examples with $\hat{Y}_{greg}$).\\
If the empirical covariance between predictions and residuals can be assumed to be zero over $G$ (it is over $F$) then one could use [\ref{varexterny0gregbis}] instead to obtain
\begin{eqnarray}\label{extsmallareaestvariance2}
 \hat{\VAR}_{ext2}(\hat{Y}_{G,g3reg})&=& \frac{1}{n_{0,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2 \nonumber \\
 &+& {\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
 (\hat{R}_1(x)-\hat{\bar{R}}_{1,G})^2 \nonumber \\
 &+& (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\hat{\bar{R}}_G)^2
 \end{eqnarray}
In the first term $\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2 $ is the empirical variance of the predictions in $s_{2,G}$ and can be replaced by the variance of the prediction in $s_{0,G}$, i.e. by $\frac{1}{n_{0,G}-1}\sum_{x\in{s_{0,G}}}(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2$ (with
$\hat{\bar{Y}}_{1,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \hat{Y}_1(x)$ ) to obtain the following variance estimate under the external model assumption.
 \begin{eqnarray}\label{extsmallareaestvariance3}
 \hat{\VAR}_{ext}(\hat{Y}_{G,g3reg})&=& \frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2 \nonumber \\
 &+& {\frac{1}{n_{1,G}(n_{2,G}-1)}\sum_{x\in{s}_{2,G}}}
 (\hat{R}_1(x)-\hat{\bar{R}}_{1,G})^2 \nonumber \\
 &+& (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}(\hat{R}(x)-\hat{\bar{R}}_G)^2
 \end{eqnarray}
 \noindent with
 \begin{eqnarray*}
 \hat{\bar{Y}}_{1,G}&=&\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}}\hat{Y}_1(x) \\
 \hat{\bar{R}}_{1,G}&=&\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}_1(x) \\
 \hat{\bar{R}}_{G}&=&\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)
 \end{eqnarray*}
 \noindent Note that the first term in [\ref{extsmallareaestvariance3}] can be expected to be more stable than the first terms in [\ref{extsmallareaestvariance1}] and [\ref{extsmallareaestvariance2}]. Also, as we shall see,  $\hat{\VAR}_{ext3}(\hat{Y}_{G,0,greg})$ is asymptotically equivalent to a g-weight version of the variance, and should therefore be preferred to [\ref{extsmallareaestvariance1}] and [\ref{extsmallareaestvariance2}].

The main difficulty to obtain better variance estimate is that the residual third term in [\ref{smallareaest1}] is no longer zero in general. The most elegant way to bypass this difficulty is by extending the model with the indicator variable $I_G(x)$ of the small area $G$. Since the procedure has been discussed in details in \cite{mandallazreport1, mandallazreport2, mandallaz3, mandallaz4} we can here briefly sketch the procedure.\\
We have the extended models with auxiliary vectors:
$\pmb{\mathcal{Z}}^t(x)=(\pmb{\mathcal{Z}}^{(1)t}(x),\pmb{\mathcal{Z}}^{(2)t}(x))$,
where $\pmb{\mathcal{Z}}^{(1)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$ and $\pmb{\mathcal{Z}}^{(2)t}(x)=\pmb{Z}^{(2)t}(x)$.
\begin{enumerate}
 \item
 The large extended model $M$
 \begin{equation*}
 Y(x)=\pmb{\mathcal{Z}}(x)^t\pmb{\theta}+ \mathcal{R(}x)=\pmb{\mathcal{Z}}^{(1)t}(x)\pmb{\theta}^{(1)}
 +\pmb{\mathcal{Z}}^{(2)t}(x)\pmb{\theta}^{(2)}+ \mathcal{R}(x)
 \end{equation*}
 \noindent with $\pmb{\theta}^t=({\pmb{\theta}^{(1)t}}, {\pmb{\theta}^{(2)t}})$.
 The intercept term is contained in $\pmb{\mathcal{Z}}^{(1)}(x)$ or it is a linear combination of its components.
 \item
 The reduced extended model $M_1$
 \begin{equation*}
 Y(x)=\pmb{\mathcal{Z}}^{(1)t}(x)\pmb{\gamma} + \mathcal{R}_1(x)
 \end{equation*}
\end{enumerate}
We can obviously apply mutatis mutandis all the previous results. The estimated regression coefficients are
\begin{eqnarray}\label{extcoeff1}
\hat{\pmb{\gamma}}_2 &=& \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathcal{Z}}^{(1)}(x)\pmb{\mathcal{Z}}^{(1)t}(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}^{(1)}(x)\nonumber\\
&:=&(\pmb{\mathcal{A}}^{(1)}_2)^{-1}
\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}^{(1)}(x)
\end{eqnarray}
and
\begin{equation}\label{extcoeff2}
\hat{\pmb{\theta}}_{2}=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{\mathcal{Z}}(x)\pmb{\mathcal{Z}}^t(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}(x):
=\pmb{\mathcal{A}}^{-1}_2\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{\mathcal{Z}}(x)
\end{equation}
The estimated covariance matrices are
\begin{eqnarray}\label{robustvar3}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_2}&=&\pmb{\mathcal{A}}_2^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathcal{R}}^2(x)\pmb{\mathcal{Z}}(x)
\pmb{\mathcal{Z}}^t(x)\Big)\pmb{\mathcal{A}}_2^{-1} \nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_2}&=&(\pmb{\mathcal{A}}^{(1)}_1)^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{\mathcal{R}}_1^2(x)\pmb{\mathcal{Z}}^{(1)}(x)
\pmb{\mathcal{Z}}^{(1)t}(x)\Big)(\pmb{\mathcal{A}}^{(1)}_1)^{-1} \nonumber \\
\end{eqnarray}
\noindent
where $\hat{\mathcal{R}}(x)=Y(x)-\pmb{\mathcal{Z}}^t(x)\hat{\pmb{\theta}}_2$ and
$\hat{\mathcal{R}}_1(x)=Y(x)-\pmb{\mathcal{Z}}^{(1)t}(x)\hat{\pmb{\gamma}}_2$ are the residuals. \\

\noindent Because the sum of the residuals over $s_{2,G}$ is now zero we can write the new small-area estimator $\hat{\tilde{Y}}_{G,g3reg}$ as in [\ref{ygreg2}]

\begin{equation}\label{smallarea0greg}
\hat{\tilde{Y}}_{G,g3reg}
= (\hat{\bar{\pmb{\mathcal{Z}}}}_{0,G}^{(1)}-\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^{(1)})^t\hat{\pmb{\gamma}}_2 +
\hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^t\hat{\pmb{\theta}}_{2}
\end{equation}
where we have set
\begin{eqnarray*}
 \hat{\bar{\pmb{\mathcal{Z}}}}_{0,G}^{(1)}&=&\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}} \pmb{\mathcal{Z}}^{(1)}(x) \\ \hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}^{(1)}&=&
 \frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\pmb{\mathcal{Z}}^{(1)}(x)\\
 \hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}&=&
 \frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}\pmb{\mathcal{Z}}(x)
 \end{eqnarray*}
 To get an estimate of the design-based variance we use mutatis mutandis [\ref{estvarianceY0greg}]
and we obtain
\begin{eqnarray}\label{estvarianceY_G0greg}
\hat{\VAR}_{0,1,2}(\hat{Y}_{G,g3reg})&=&\hat{\pmb{\gamma}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}}
\hat{\pmb{\gamma}}_2 +\frac{n_2}{n_1}\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)t}_{0,G}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{2}}\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G} \nonumber \\ &+& (1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_2}
\hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,G}
\end{eqnarray}
where
\begin{equation}\label{estcovarextz-overG}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}}=
\frac{1}{n_{0,G}}\frac{1}{n_{0,G}-1}  \sum_{x\in{s_{0,G}}}(\pmb{\mathcal{Z}}^{(1)}(x)-
\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G})(\pmb{\mathcal{Z}}(x)-\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G})^t
\end{equation}
\noindent is the empirical covariance matrix of $\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}$ over $G$.\\
The first term in [\ref{estvarianceY_G0greg}] can be rewritten as the empirical variance of the predictions
$\hat{\tilde{Y}}_1(x)=\pmb{\mathcal{Z}}^{(1)t}\hat{\pmb{\gamma}}_2$, that is
\begin{equation*}
\hat{\pmb{\gamma}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}}\hat{\pmb{\gamma}}_2=
\frac{1}{n_{0,G}-1}\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}}(\hat{\tilde{Y}}_1(x)-\hat{\bar{\tilde{Y}}}_{1,G})^2
\end{equation*}
\noindent with $\hat{\bar{\tilde{Y}}}_{1,G}=\frac{1}{n_{0,G}}\sum_{x\in{s_{0,G}}}\hat{\tilde{Y}}_1(x)$.\\
It is shown in section \ref{proofofequivalence} that the external variance estimate [\ref{extsmallareaestvariance3}]
and the g-weight variance estimate [\ref{estvarianceY_G0greg}] are asymptotically equivalent, which is not obvious at first glance.\\
For a very small area $H$ the number of points $n_{2,H}$ may be to small or even zero. In such a case one could imbed $H$ in a somewhat larger area $G\supset H$ with $n_{2,G}$ sufficiently large and consider the synthetic estimator in the extended model with respect to $G$: i.e. one replaces $\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}$ by $\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,H}$, $\hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,G}$ by $\hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,H}$  in equations [\ref{smallarea0greg}], [\ref{estvarianceY_G0greg}] and
$\hat{\pmb{\gamma}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,G}}\hat{\pmb{\gamma}}_2$ by
$\hat{\pmb{\gamma}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{0,H}}\hat{\pmb{\gamma}}_2
=\frac{1}{n_{0,H}-1}\frac{1}{n_{0,H}}\sum_{x\in{s_{0,H}}}(\hat{\tilde{Y}}_1(x)-\hat{\bar{\tilde{Y}}}_{1,H})^2$ with
$\hat{\bar{\tilde{Y}}}_{1,H}=\frac{1}{n_{0,H}}\sum_{x\in{s_{0,H}}}\hat{\tilde{Y}}_1(x)$.\\
The variance $\hat{\VAR}_{0,1,2}(\hat{Y}_{H,g3reg})$ will be small but at the cost of a potential bias, which can be expected to be smaller than the bias of the synthetic estimator with respect to $G=F$.\\
As shown in \cite{mandallazreport1} it is straightforward to consider simultaneously several
small areas by extending the model with as many small area indicator variables.\\

\section{Generalization to cluster sampling}
We follow the description of cluster sampling as defined in \cite{mandallaz} (especially section 5.5) and \cite{mandallazreport1}. A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$. The geometry of the cluster is given by $M$ vectors $e_1,\ldots e_M$ defining the random cluster $x_l=x+e_l$. $M(x)=\sum_{l=1}^MI_{F}(x_l)$ is the random number of points of the cluster falling into the forest area $F$. We define the local density at the cluster level by $Y_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)Y(x_l)}{M(x)}$, likewise we set $\pmb{Z}_c(x)=\frac{\sum_{l=1}^MI_{F}(x_l)\pmb{Z}(x_l)}{M(x)}$. The set $\tilde{F}$ above can be mathematically defined as the smallest set $\{x\in{\mathcal{R}}^2 \mid M(x) \ne 0 \}$. In the null phase we have $n_0$ clusters identified by $x\in{s_0}$, in the first phase $n_1$ clusters identified by $x\in{s_1}$, and in the second phase $n_2$ clusters with $x\in{s_2}$, obtained by simple random sampling in $s_0$, or $s_1$ respectively.
As the estimators $\hat{Y}_{reg},\hat{Y}_{F,greg}$ and $\hat{\tilde{Y}}_{G,greg}$ have been generalized to cluster sampling in \cite{mandallazreport2,mandallaz3,mandallaz4} we will only sketch the main features. \\
The design-unbiased estimate $\hat{\pmb{\beta}}_{c,2}$  for $\pmb{\beta}_c$ (the equivalent of $\pmb{\beta}$ in simple random sampling) is
 \begin{eqnarray}\label{estbetacluster1}
 \hat{\pmb{\beta}}_{c,2}&=&
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}_c(x)\pmb{Z}_c^t(x)\Big)^{-1}
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \nonumber\\
 &:=&\pmb{A}^{-1}_{c,s_2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \end{eqnarray}
For the reduced model with $\pmb{Z}^{(1)}(x)$ alone we obtain similarly
\begin{eqnarray}\label{estalphacluster1}
 \hat{\pmb{\alpha}}_{c,2}&=&
 \Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{Z}^{(1)}_c(x)\pmb{Z}_c^{(1)t}(x)\Big)^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}_c(x)\Big)
 \nonumber\\
 &:=&(\pmb{A}^{(1)}_{c,2})^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)Y_c(x)\pmb{Z}^{(1)}_c(x)\Big)
 \end{eqnarray}
 The residuals for the reduced model are $\hat{R}_{1,c}(x)=Y_c(x)-\hat{Y}_{1,c}$ with the predictions $\hat{Y}_{1,c}=\pmb{Z}^{(1)t}_c(x)\hat{\pmb{\alpha}}_{c,2}$.\\
 \noindent Using mutatis mutandis exactly the same arguments as in simple random sampling we get the asymptotic robust design-based estimated covariance matrices
\begin{eqnarray}\label{estcovmatrixcluster1}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,2}}&=&\pmb{A}^{-1}_{c,2}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}^2_c(x)
\pmb{Z}_c(x)\pmb{Z}^t_c(x)\Big)\pmb{A}^{-1}_{c,2}\nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{c,2}}&=&(\pmb{A}^{(1)}_{c,1})^{-1}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}M^2(x)\hat{R}_{1,c}^2(x)
\pmb{Z}_c(x)\pmb{Z}^t_c(x)\Big)(\pmb{A}^{(1)}_{c,1})^{-1}
\end{eqnarray}
where $\pmb{A}^{(1)}_{c,1}=\frac{1}{n_1}\sum_{x\in{s_1}}M(x)\pmb{Z}^{(1)}_c(x)\pmb{Z}_c^{(1)t}(x)$.\\

We define the \textbf{generalized regression estimator in cluster sampling} directly by adapting [\ref{ygreg2}]
\begin{equation}\label{y0gregcluster}
\hat{Y}_{c,g3reg}= (\hat{\bar{\pmb{Z}}}_{c,0}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_{c,1})^t\hat{\pmb{\alpha}}_{c,2} +
\hat{\bar{\pmb{Z}}}^t_{c,1}\hat{\pmb{\beta}}_{c,2}
\end{equation}
where
\begin{eqnarray*}
\hat{\bar{\pmb{Z}}}^{(1)}_{c,0}&=&\frac{\sum_{x\in{s_0}}M(x)\pmb{Z}^{(1)}_c(x)}{\sum_{x\in{s_0}}M(x)} \\
\hat{\bar{\pmb{Z}}}^{(1)}_{c,1}&=&\frac{\sum_{x\in{s_1}}M(x)\pmb{Z}^{(1)}_c(x)}{\sum_{x\in{s_1}}M(x)} \\
\hat{\bar{\pmb{Z}}}_{c,1}&=& \frac{\sum_{x\in{s_1}}M(x)\pmb{Z}_c(x)}{\sum_{x\in{s_1}}M(x)}
\end{eqnarray*}

With the same technique as in simple random sampling we obtain the estimated design-based variance
\begin{eqnarray}\label{estvarclustery0greg}
\hat{\VAR}_{0,1,2}(\hat{Y}_{c,g3reg})&=& \hat{\pmb{\alpha}}^t_{c,2}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_{c,0}}\hat{\pmb{\alpha}}_{c,2} \nonumber \\
&+& \frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}_{c,0}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{c,2}}
\hat{\bar{\pmb{Z}}}_{c,0}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{c,1}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,2}}
\hat{\bar{\pmb{Z}}}_{c,1}
\end{eqnarray}
where
\begin{equation*}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_{c,0}}=
\frac{1}{n_0(n_0-1)}\sum_{x\in{s_0}}\big(\frac{M(x)}{\bar{M}_0}\big)^2(\pmb{Z}^{(1)}_c(x)-\hat{\bar{\pmb{Z}}}^{(1)}_{c,0})
(\pmb{Z}^{(1)}_c(x)-\hat{\bar{\pmb{Z}}}^{(1)}_{c,0})^t
\end{equation*}
\noindent is the empirical covariance matrix of the $\pmb{Z}^{(1)}_c(x)$ in $s_0$.\\
The first term in [\ref{estvarclustery0greg}] can be rewritten as the empirical variance over $s_0$ of the predictions of the reduced model, i.e.
\begin{equation*}
\frac{1}{n_0(n_0-1)}\sum_{x\in{s_0}}\big(\frac{M(x)}{\bar{M}_0}\big)^2(\hat{Y}_{1,c}(x)-
\hat{\bar{Y}}_{1,0})^2
\end{equation*}
with $\hat{\bar{Y}}_{1,0}=\frac{\sum_{x\in{s_0}}M(x)\hat{Y}_{1,c}(x)}{\sum_{x\in{s_0}}M(x)}$ and
$\bar{M}_0=\frac{1}{n_0}\sum_{x\in{s_0}}M(x)$.\\
The external version of the variance estimate can be easily adapted from [\ref{extexternvary0greg}] after introducing the weights $\big(\frac{M(x)}{\bar{M}_k}\Big)^2$.\\
For small-area estimation we shall work with the extended model
\begin{equation*}
\pmb{\mathcal{Z}}^t(x)=(\pmb{\mathcal{Z}}^{(1)t}(x), \pmb{\mathcal{Z}}^{(2)t}(x))
\end{equation*}
 with
$\pmb{\mathcal{Z}}^{(1)t}(x)=(\pmb{Z}^{(1)t}(x), I_G^t(x))$, $\pmb{\mathcal{Z}}^{(2)t}(x)=\pmb{Z}^{(2)t}(x)$ and $I_G(x)$ is the indicator of the small area $G$. At the cluster level we have $\pmb{\mathcal{Z}}_c^{(1)t}(x)=(\pmb{Z}_c^{(1)t}(x),I^t_{c,G}(x))$ where $I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}$. In extensive inventories we can reasonably assume that all the points of a cluster lying in the forest area $F$ will belong to the same small area $G$ so that in fact $I_{c,G}(x) \equiv 1$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$. This ensures again that we will have zero mean residual over $F$ and $G$.  For the regression estimates in the extended model we have

\begin{equation}\label{regbettaclusterext}
\hat{\pmb{\gamma}}_{c,2}=(\pmb{\mathcal{A}}_{c,2}^{(1)})^{-1}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)Y(x)\Big)
\end{equation}
with $\pmb{\mathcal{A}}_{c,2}^{(1)}=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)\pmb{\mathcal{Z}}^{(1)t}_c(x)$. Likewise we get
\begin{equation}\label{regbettaclusterext}
\hat{\pmb{\theta}}_{c,2}=\pmb{\mathcal{A}}^{-1}_{c,2}\Big(\frac{1}{n_2}\sum_{x\in{s_2}}M(x)
\pmb{\mathcal{Z}}_c(x)Y(x)\Big)
\end{equation}
with $\pmb{\mathcal{A}}_{c,2}=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)$.\\
We define as in [\ref{y0gregcluster}] the generalized small-area estimator by
\begin{equation}\label{extclustersmallareaY_0greg}
\hat{\tilde{Y}}_{c,G,g3reg}
= (\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^{(1)})^t\hat{\pmb{\gamma}}_{c,2} +
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,2}
\end{equation}
where we have set
\begin{eqnarray*}
 \hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{c,0,G}&=&\frac{\sum_{x\in{s_{0,G}}}M(x)\mathcal{Z}_c^{(1)}(x)}{\sum_{x\in{s_{0,G}}}M(x)} \\ \hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^{(1)}&=&
 \frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c^{(1)}(x)}{\sum_{x\in{s_{1,G}}}M(x)} \\
 \hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}&=&
 \frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c(x)}{\sum_{x\in{s_{1,G}}}M(x)}
\end{eqnarray*}
The estimated design-based covariance matrix are now
\begin{eqnarray}\label{gammathetaclusterextestvar}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{c,2}}&=&
(\pmb{\mathcal{A}}^{(1)}_{c,1})^{-1}\Big(\frac{1}{n^2_2}\sum_{\in{s_2}}M^2(x)\hat{\mathcal{R}}_{1,c}^2(x)
\pmb{\mathcal{Z}}^{(1)}_c(x)\pmb{\mathcal{Z}}^{(1)t}_c(x)\Big)(\pmb{\mathcal{A}}^{(1)}_{c,1})^{-1}\nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,2}}&=&
\pmb{\mathcal{A}}^{-1}_{c,2}\Big(\frac{1}{n^2_2}\sum_{\in{s_2}}M^2(x)\hat{\mathcal{R}}_c^2(x)
\pmb{\mathcal{Z}}_c(x)\pmb{\mathcal{Z}}^t_c(x)\Big)\pmb{\mathcal{A}}^{-1}_{c,2}
\end{eqnarray}
\noindent where
\begin{eqnarray*}
\hat{\mathcal{R}}_{1,c}(x)&=&Y_c(x)-\hat{\tilde{Y}}_{1,c}(x):=Y_c(x)-\pmb{\mathcal{Z}}^{(1)t}_c(x)\hat{\pmb{\gamma}}_2\\
\hat{\mathcal{R}}_{c}(x)&=&Y_c(x)-\hat{\tilde{Y}}_c(x):=Y_c(x)-\pmb{\mathcal{Z}}^{t}_c(x)\hat{\pmb{\theta}}_2
\end{eqnarray*}
We obtain as in [\ref{estvarclustery0greg}] the estimated design-based variance

\begin{eqnarray}\label{extendedclustersmallestvariance}
\hat{\VAR}_{0,1,2}(\hat{Y}_{c,G,g3reg}) &=&\hat{\pmb{\gamma}}^t_{c,2}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}}\hat{\pmb{\gamma}}^t_{c,2} \nonumber \\
&+& \frac{n_2}{n_1}\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\gamma}}_{c,2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,2}}
\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}
\end{eqnarray}
\noindent where
\begin{equation*}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}}=
\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}\Big(\frac{M(x)}{\bar{M}_0}\Big)^2
(\pmb{\mathcal{Z}}^{(1)}_c(x)-\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{c,0,G})(\pmb{\mathcal{Z}}^{(1)}_c(x)-\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)}_{c,0,G})^t
\end{equation*}
is the empirical covariance of $\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}$ over $G$.\\
The first term in [\ref{extendedclustersmallestvariance}] can be rewritten as the variance of the predictions, that is
\begin{equation*}
\hat{\pmb{\gamma}}^t_{c,2}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,0,G}^{(1)}}\hat{\pmb{\gamma}}^t_{c,2}
=\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}\Big(\frac{M(x)}{\bar{M}_0}\Big)^2
(\hat{\tilde{Y}}_{1,c}(x)-\bar{\tilde{Y}}_{1,c})^2
\end{equation*}
\noindent with $\bar{\tilde{Y}}_{1,c}=\frac{\sum_{x\in{s_{0,G}}}M(x)\hat{\tilde{Y}}_{1,c}(x)}{\sum_{x\in{s_{0,G}}}M(x)}$. \\
The external version of the variance estimate can be easily adapted from [\ref{extsmallareaestvariance3}]: replace all occurring expressions $U(x)$ by the cluster means $U_c(x)$, calculate the mean values according to
$\bar{U}=\frac{\sum_x M(x)Uc_(x)}{\sum_x M(x)}$ and introduce the weights $\Big(\frac{M(x)}{\bar{M}_k}\Big)^2$.\\

\section{Generalization to two-stage sampling at the plot level}
In many applications costs to measure the response variable $Y_i$ are high. For instance,
a good determination of the volume may require that one records
$DBH$, as well as the diameter at $7m$ above ground and total height
in order to utilize a three-way volume function. However, one could rely
on a coarser, but cheaper, approximation of the volume based only on
$DBH$. Nonetheless, it may be most sensible to assess those three
parameters only on a sub-sample of trees. We now briefly formalize this
simple idea, which is used in the Swiss National Forest Inventory. The reader is referred to (\cite{mandallaz}, section 4.4, 4.5, 5.4 and 9.5) for details. For each point $x\in{s_2}$ trees are drawn with
probabilities $\pi_i$. The set of selected trees is denoted by
$s_{2}(x)$. From each of the selected trees $i\in{s_{2}(x)}$ one
gets an approximation $Y_i^*$ of the exact value $Y_i$. From the
finite set $s_{2}(x)$ one draws a sub-sample
$s_{3}(x)\subset{s_{2}(x)}$ of trees by Poisson sampling. For each tree $i\in{s_{3}(x)}$
one then measures the exact variable $Y_i$. Let us now define the
second stage indicator variable
\begin{equation}
 J_i(x)=\begin{cases}&1 \text{ if $i\in s_{3}(x)$}\\
                      &0 \text{ if $i\not\in s_{3}(x)$}
         \end{cases}
\end{equation}

 \par To construct a good point estimate, we must have
 the residual $R_i=Y_i-Y_i^*$ which is known only for trees
 $i\in{s_{3}(x)}$. The generalized local density $Y^*(x)$ is defined
 according to    \index{Generalized local density}
 \begin{eqnarray}\label{gdens}
 Y^*(x)&=&\frac{1}{\LF}\left( \sum_{i=1}^N \frac{I_{i}(x)Y_i^*}{\pi_i} +
 \sum_{i=1}^N \frac{I_{i}(x)J_{i}(x)R_i}{\pi_{i}p_i}\right)\nonumber \\
 &=&\frac{1}{\LF}\left( \sum_{i\in{s_{2}(x)}} \frac{Y_i^*}{\pi_i}
 +\sum_{i\in{s_{3}(x)}} \frac{R_i}{\pi_{i}p_i}\right)
 \end{eqnarray}
 where the $p_i$ are the conditional inclusion probabilities for the the second stage sampling, i.e. $p_i=\PR(J_i(x)=1 \mid I_i(x)=1)$.\\
  One can use all the previous results by making the following changes:\\
  \begin{itemize}
  \item
   $Y(x)\rightarrow Y^{*}(x)$
    \item $\hat{\pmb{\beta}}\rightarrow \hat{\pmb{\beta}}_2^{*}=\pmb{A}_2^{-1}\Big(\frac{1}{n_2}\sum_{x\in_{s_2}}Y^{*}(x)\pmb{Z}(x)\Big)$,
  $\hat{\pmb{\alpha}}_2^{*}=(\pmb{A}^{(1)}_2)^{-1}\Big(\frac{1}{n_2}\sum_{x\in_{s_2}}Y^{*}(x)\pmb{Z}^{(1)}(x)\Big)$
\item
  $\hat{R}_1(x)\rightarrow\hat{R}_1^{*}(x)=Y^{*}(x)-\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2^{*}$,
   $\hat{R}(x)\rightarrow \hat{R}^{*}(x)=Y^{*}(x)-\pmb{Z}^t(x)\hat{\pmb{\beta}}_2^{*}$
    \item
   Similarly in cluster sampling with $Y(x_l)\rightarrow Y^{*}(x_l)$, $Y_c(x)\rightarrow Y_c^{*}(x)$ etc.
   \end{itemize}
 One obtains for the resulting three-phase two-stage estimator $\hat{Y}^{*}_{F,g3reg}$ the theoretical variance
 \begin{equation}\label{3phasetwostegevariance}
 \VAR_{0,1,2,3}(\hat{Y}^{*}_{F,g3reg})=\VAR_{0,1,2,3}(\hat{Y}_{F,g3reg})+\frac{1}{n_2}\EX_{x\in{F}}V(x)
 \end{equation}
 \noindent where $\EX_{x\in{F}}V(x)=\frac{1}{\lambda^2(F)}\sum_{i=1}^N\frac{R_i^2(1-p_i)}{\pi_ip_i}$ is the second stage variance. The subscript $3$ in $\VAR_{0,1,2,3}$ refers to the second-stage random selection at points $x\in{s_2}$.\\
 The proof of [\ref{3phasetwostegevariance}] is straightforward under the external model assumption (including the small-area version), it suffices to use the techniques presented in (\cite{mandallaz}, section 4.4 pp. 71-73, section 4.5 pp. 82-83), which also show that one can use the variance estimates [\ref{extsmallareaestvariance2}] and [\ref{extsmallareaestvariance3}] (valid also for $G=F$), after making the afore mentioned changes $Y(x)\rightarrow Y^{*}(x)$ etc. The second-stage variance is automatically taken into account. The same applies to cluster sampling (by using \cite{mandallaz}, section 5.3 pp. 86-83). Because of the algebraic asymptotic equivalence established in section \ref{proofofequivalence} of the Appendix this will also hold true for all the g-weights variance estimates.

 \bibliography{biblio1}
\begin{appendix}
\clearpage
\newpage
\section{Variance of $\hat{Y}_{F,g3reg}$}\label{proofofvariance}
According to \cite{mandallaz} (pp 124-125) the asymptotic covariance matrices of the regression coefficients are given
by
\begin{equation}\label{lemma1}
\pmb{\Sigma}_{\hat{\pmb{\beta}}_k}=\pmb{A}^{-1}
\Big(\frac{1}{n_k}\EX R^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}^{-1}
\end{equation}
\begin{equation}\label{lemma2}
\pmb{\Sigma}_{\hat{\pmb{\alpha}}_k}=(\pmb{A}^{(1)})^{-1}
\Big(\frac{1}{n_k}\EX R_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)})^{-1}
\end{equation}
\noindent where $\pmb{A}=\EX_x \pmb{Z}(x)\pmb{Z}^t(x)$  and $\pmb{A}^{(1)}=\EX_x \pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)$, which can be consistently estimated by $\pmb{A}_k$ ($k=1,2$) and  $\pmb{A}_k^{(1)}$ ($k=0,1,2$). To ensure the calibration properties we will use $\pmb{A}_2$ and $\pmb{A}_1^{(1)}$.\\
We start with the calculation of
$T_3=\VAR_0(\EX_{1 \mid 0}\EX_{1,2 \mid 0}(\hat{Y}_{F,g3reg}))\approx \VAR_0(\hat{\bar{\pmb{Z}}}_0^{(1t)}\hat{\pmb{\alpha}}_0)$. Since
$\EX_0(\hat{\bar{\pmb{Z}}}_0^{(1t)}\hat{\pmb{\alpha}}_0) \approx \bar{\pmb{Z}}^{(1t)}\pmb{\alpha}=\bar{Y}$
we consider
\begin{equation*}
\Delta=\hat{\bar{\pmb{Z}}}^{(1)t}\hat{\pmb{\alpha}}_{0}-\bar{\pmb{Z}}^{(1)t}\pmb{\alpha}=
\hat{\bar{\pmb{Z}}}_0^{(1)t}(\hat{\pmb{\alpha}}_{0}-\pmb{\alpha})+ (\hat{\bar{\pmb{Z}}}_0^{(1)t}-\bar{\pmb{Z}}^{(1)t})\pmb{\alpha}
\end{equation*}

Asymptotically we get
\begin{equation*}
\EX \Delta^2=\bar{\pmb{Z}} ^{(1)t}\pmb{\Sigma}_{\hat{\pmb{\alpha}}_{0}}\bar{\pmb{Z}}^{(1)}+
\pmb{\alpha}^t\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}\pmb{\alpha}+
2\EX\Big((\hat{\pmb{\alpha}}_{0}-\pmb{\alpha})^t\hat{\bar{\pmb{Z}}}^{(1)}_{0}(\hat{\bar{\pmb{Z}}}^{(1)}_{0}
-\bar{\pmb{Z}}^{(1)})^t\pmb{\alpha}\Big)
\end{equation*}
\noindent where $\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}$ is the covariance matrix of
$\hat{\bar{\pmb{Z}}}^{(1)}_{0}$. The first two terms are of order $n_0^{-1}$. For the third term we note that
$\EX (\hat{\bar{\pmb{Z}}}^{(1)}_{0}(\hat{\bar{\pmb{Z}}}^{(1)}_{0}-\bar{\pmb{Z}}^{(1)})^t)$ is equal to the covariance matrix $\pmb{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}$ and therefore of order $n_0^{-1}$ and that
$\hat{\pmb{\alpha}}_{0}-\pmb{\alpha}$ is of order $n_0^{-\frac{1}{2}}$. The last term is therefore of smaller order than the first two. Since $\pmb{\Sigma}_{\hat{\alpha}_0}=\frac{n_2}{n_0}\pmb{\Sigma}_{\hat{\alpha}_2}$ we get the following asymptotically consistent estimate
\begin{equation}\label{appendixhatt3}
\hat{T}_3 =
\frac{n_2}{n_0}\hat{\bar{\pmb{Z}}}^{(1)t}_{0}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(1)}_{0}
+ \hat{\pmb{\alpha}}_{2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}\hat{\pmb{\alpha}}_{2}
\end{equation}

We have to calculate
\begin{equation*}
T_2=\EX_0\VAR_{1 \mid 0}\EX_{2\mid 0,1}(\hat{Y}_{F,g3reg})\approx \EX_0\VAR_{1 \mid 0}(\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_1)
\end{equation*}
because, as we have seen, $\EX_{2\mid 0,1}(\hat{Y}_{F,g3reg})\approx \hat{\bar{\pmb{Z}}}^{(1)t}_0\hat{\pmb{\alpha}}_1$.
 First, we note that
\begin{eqnarray*}
 \VAR_{1 \mid 0}(\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_1)&=&
\EX_{ 1\mid 0}(\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_1-\EX_{1 \mid 0}\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_1)^2 \nonumber\\
& \approx & \EX_{1 \mid 0}(\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_1-\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\alpha}}_0)^2 \nonumber \\
&=& \hat{\bar{\pmb{Z}}}_0^{(1)t}\EX_{1 \mid 0}(\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0)(\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0)^t\hat{\bar{\pmb{Z}}}_0^{(1)t}
\end{eqnarray*}
\noindent Using the Taylor expansion given in \cite{mandallaz} (p. 124, but at point $(\pmb{A}^{(1)}_0,\;\hat{\pmb{\alpha}}_0)$ instead of $(\pmb{A}^{(1)},\;\pmb{\alpha}))$ we get (with the notation defined in [\ref{coeff1}])
\begin{eqnarray*}
\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0 &\approx &
-(\pmb{A}^{(1)}_0)^{-1}\big(\pmb{A}^{(1)}_1-\pmb{A}^{(1)}_0\big)(\pmb{A}^{(1)}_0)^{-1}\pmb{U}_0+
(\pmb{A}^{(1)}_0)^{-1}(\pmb{U}_1-\pmb{U}_0) \nonumber \\
&=& (\pmb{A}^{(1)}_0)^{-1}(\pmb{U}_1-\pmb{A}^{(1)}_1\hat{\pmb{\alpha}}_0)
\end{eqnarray*}
Since $\hat{\pmb{\alpha}}_0=\pmb{\alpha} + O(n_0^{-\frac{1}{2}})$ in probability this is asymptotically equivalent
to
\begin{eqnarray}\label{tayloralpha0}
\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0 & \approx &
(\pmb{A}^{(1)}_0)^{-1}\Big(\frac{1}{n_1}\sum_{x\in{s_1}}(Y(x)-\pmb{Z}^{(1)t}(x)\pmb{\alpha})\pmb{Z}^{(1)}(x)\Big)\nonumber \\
&=&(\pmb{A}^{(1)}_0)^{-1}\Big(\frac{1}{n_1}\sum_{x\in{s_1}}R_1(x)\pmb{Z}^{(1)}(x)\Big)
\end{eqnarray}
With the indicator variables $J(x)=1$ if $x\in{s_0}$ and $J(x)=0$ if $x\not\in{s_0}$ we obtain
\begin{eqnarray*}
\EX_{1 \mid 0}(\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0 )(\hat{\pmb{\alpha}}_1-\hat{\pmb{\alpha}}_0)^t & \approx &
(\pmb{A}^{(1)}_0)^{-1}\frac{1}{n^2_1}\EX_{1 \mid 0}\sum_{x\in{s_0}}J(x)R_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\nonumber \\
&+& (\pmb{A}^{(1)}_0)^{-1}\frac{1}{n^2_1}\EX_{1 \mid 0}\sum_{x\ne y\in{s_0}}J(x)J(y)R_1(x)\pmb{Z}^{(1)}(x)R_1(y)
\pmb{Z}^{(1)t}(y)
\end{eqnarray*}
Using simple random sampling within the set $s_0$ we have
\begin{eqnarray*}
\EX_{1 \mid 0}J(x)&=&\frac{n_1}{n_0}\nonumber \\
\EX_{1 \mid 0}J(x)J(y)&=& \frac{n_1(n_1-1)}{n_0(n_0-1)} \\
\sum_{x\ne y\in{s_0}}J(x)J(y)R_1(x)\pmb{Z}^{(1)}(x)R_1(y)
\pmb{Z}^{(1)t}(y)&=&(\sum_{x\in{s_0}}J(x)R_1(x)\pmb{Z}^{(1)}(x))^2 \\
&-&\sum_{x\in{s_0}}R^2_1(x)
\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x) \\
\EX_0\EX_{1 \mid 0}\sum_{x\in{s_0}}J(x)R_1(x)\pmb{Z}^{(1)}(x)&=&0\quad  \text{(orthogonality relationship)}
\end{eqnarray*}
With the approximation $\frac{n_1-1}{n_0-1}\approx \frac{n_1}{n_0}$ we obtain after tedious but simple algebra
\begin{eqnarray}\label{appendixt2}
T_2=(1-\frac{n_1}{n_0})\bar{\pmb{Z}}^{(1)t}(\pmb{A}^{(1)})^{-1}\Big(\frac{1}{n_1}\EX_x( R^2_1 (x)\pmb{Z}^{(1)}(x)
\pmb{Z}^{(1)t}(x))\Big)(\pmb{A}^{(1)})^{-1}\bar{\pmb{Z}}^{(1)}
\end{eqnarray}
\noindent and the 'plug-in' asymptotically consistent estimate

\begin{eqnarray}\label{appendixhatt2}
\hat{T}_2 & \approx &
(1-\frac{n_1}{n_0})\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}_0^{(1)t}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}\hat{\bar{\pmb{Z}}}_0^{(1)} \nonumber \\
&\approx & (1-\frac{n_1}{n_0})\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}_0^{(1)t}(\pmb{A}^{(1)}_1)^{-1}\Big(\frac{1}{n_2^2}
\sum_{x\in{s_2}}\hat{R}^2_1(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)}_1)^{-1}\hat{\bar{\pmb{Z}}}_0^{(1)}
\end{eqnarray}
We now proceed to calculate the first term $T_1$. First we note that
\begin{equation}\label{equivalence2}
\hat{Y}_{F,g3reg}-\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})\approx (\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t
(\hat{\pmb{\alpha}}_{2}-\hat{\pmb{\alpha}}_{1})+\hat{\bar{\pmb{Z}}}^{t}_1(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1)
\approx \hat{\bar{\pmb{Z}}}^{t}_1(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1)
\end{equation}
The last asymptotic equivalence in [\ref{equivalence2}] holds because  $(\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t(\hat{\pmb{\alpha}}_{2}-\hat{\pmb{\alpha}}_{1})$ is the product of two terms of order $O(n_2^{-\frac{1}{2}})$ in probability whereas the second is of order $O(n_2^{-\frac{1}{2}})$. One gets therefore
\begin{eqnarray*}
\VAR_{2 \mid 0,1}(\hat{Y}_{F,g3reg}-\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})) &\approx &
\hat{\bar{\pmb{Z}}}^{t}_1\EX_{2 \mid 0,1}(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1)(\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1)^t\hat{\bar{\pmb{Z}}}_1 \nonumber \end{eqnarray*}
Using mutatis mutandis [\ref{tayloralpha0}] (i.e. Taylor expansion at $\hat{\pmb{\beta}}_1$) we have
\begin{equation}\label{taylorbeta1}
\hat{\pmb{\beta}}_2-\hat{\pmb{\beta}}_1  \approx \pmb{A}^{-1}_1\Big(\frac{1}{n_2}\sum_{x\in{s_2}}R(x)\pmb{Z}(x)\Big)
\end{equation}
Using the same techniques as for $T_2$ (simple random sampling within the set $s_1$ instead of $s_0$)
we have
\begin{eqnarray}\label{appendixt1}
T_1=(1-\frac{n_2}{n_1})\bar{\pmb{Z}}^{t}\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX_x( R^2 (x)\pmb{Z}(x)
\pmb{Z}^{t}(x))\Big)\pmb{A}^{-1}\bar{\pmb{Z}}
\end{eqnarray}
and
\begin{equation}\label{appendixt1}
T_1 \approx (1-\frac{n_2}{n_1})\bar{\pmb{Z}}^{t}\pmb{\Sigma}_{\pmb{\hat{\beta}}_2}\bar{\pmb{Z}}
\end{equation}
which leads to the 'plug-in' asymptotically consistent estimate
\begin{eqnarray}\label{appendixhatt1}
\hat{T}_1 &\approx& (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^{t}\hat{\pmb{\Sigma}}_{\pmb{\hat{\beta}}_2}\hat{\bar{\pmb{Z}}}_1\nonumber \\
&=&(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^{t}\pmb{A}_2^{-1}\big(\frac{1}{n^2_2}\sum_{x\in{s_2}}
\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}_2^{-1}\hat{\bar{\pmb{Z}}}_1
\end{eqnarray}
Hence, one obtains the asymptotically consistent estimate of the variance
\begin{equation*}\label{estvary0greg1}
\hat{\VAR}(\hat{Y}_{F,g3reg}) \approx  \hat{T}_3 + \hat{T}_2 + \hat{T}_1
\end{equation*}
Because two terms cancel out we finally obtain
\begin{equation}\label{estvary0greg2}
\hat{\VAR}(\hat{Y}_{F,g3reg}) \approx  \hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_0}\hat{\pmb{\alpha}}_2
+ \frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}_0^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}\hat{\bar{\pmb{Z}}}_0^{(1)}
+ (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_2}\hat{\bar{\pmb{Z}}}_1
\end{equation}
Since the variance-covariance matrix of the $\pmb{Z}^{(1)}(x)$ is given by
\begin{equation*}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}=
\frac{1}{n_{0}}\frac{\sum_{x\in{s_{0}}}(\pmb{Z}^{(1)}(x)-
\hat{\bar{\pmb{Z}}}^{(1)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(1)}_{0})^t}{n_{0}-1}
\end{equation*}
it is easily checked that the first term in [\ref{estvary0greg2}] is equal to the variance of the predictions under the reduced model, i.e.
\begin{equation*}
\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1}
\end{equation*}
\noindent with $\hat{Y}_1(x)=\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_2$ and
$\hat{\bar{Y}}_1=\frac{1}{n_0}\sum_{x\in{s_0}}\hat{Y}_1(x)$.

\section{Asymptotic equivalence of variance estimates for small areas} \label{proofofequivalence}
We define the g-weights
\begin{eqnarray}\label{smallgweights}
\tilde{g}(x)&=& \hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,G}\pmb{\mathcal{A}}_2^{-1}\pmb{\mathcal{Z}}(x) \nonumber \\
\tilde{g}^{(1)}(x)&=&\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)t}_{0,G}(\pmb{\mathcal{A}}^{(1)}_1)^{-1}\pmb{\mathcal{Z}}^{(1)}(x)
\end{eqnarray}
First, we note that [\ref{estvarianceY_G0greg}] can be rewritten as
\begin{eqnarray*}
\hat{\VAR}_{0,1,2}(\hat{Y}_{G,g3reg})&=&\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2 \\
&+& \frac{1}{n_1n_2}\sum_{x\in{s_2}} (\tilde{g}^{(1)}(x))^2\hat{\mathcal{R}}_1^2(x)
+(1-\frac{n_2}{n_1})\frac{1}{n_2^2}\sum_{x\in{s_2}}\tilde{g}^2(x)\hat{\mathcal{R}}^2(x)
\end{eqnarray*}
Second, we consider
\begin{equation*}
\hat{\bar{\pmb{\mathcal{Z}}}}^{(1)t}_{0,G}\hat{\pmb{\gamma}}_1=\frac{1}{n_1}\sum_{x\in{s_1}}\tilde{g}^{(1)}(x)Y(x)\approx
\frac{1}{n_{1,G}}\sum_{x\in{s_{1,G}}}Y(x)
\end{equation*}
 \noindent because all terms converge towards $\bar{Y}_G$ (the fact that they are not observable is irrelevant in this respect). Hence, we see that $\tilde{g}^{(1)}(x) \to 0$ for $x\not\in{G}$ and $\tilde{g}^{(1)}(x)\to \frac{n_1}{n_{1,G}}$ for $x\in{G}$. Likewise we have
\begin{equation*}
\hat{\bar{\pmb{\mathcal{Z}}}}^{t}_{1,G}\hat{\pmb{\theta}}_2=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}(x)Y(x)\approx
\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}Y(x)
\end{equation*}
\noindent so that $g(x) \to 0$ for $x\not\in{G}$ and $g(x)\to \frac{n_2}{n_{2,G}}$ for $x\in{G}$.  Note that the behavior of the g-weights for $G$ and for $F$ is completely different. \\
Since obviously $\frac{n_{2,G}}{n_{1,G}}\approx \frac{n_2}{n_1}$ we get
\begin{eqnarray}
\hat{\VAR}_{0,1,2}(\hat{Y}_{G,g3reg})&\approx&\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
(\hat{Y}_1(x)-\hat{\bar{Y}}_{1,G})^2 \\
&+& \frac{1}{n_{1,G}n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{\mathcal{R}}_1^2(x)
+(1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}^2}\sum_{x\in{s_{2,G}}}\hat{\mathcal{R}}^2(x)
\end{eqnarray}
\noindent Finally, using the relation between $\hat{\pmb{\beta}}_2$ and $\hat{\pmb{\theta}}_2$ given in equation [42]
of \cite{mandallazreport1}, we see that $\hat{\mathcal{R}}(x)=\hat{R}(x)-\phi(x)\hat{\bar{R}}_G$ with $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\phi(x)=1$, and likewise $\hat{\mathcal{R}}_1(x)=\hat{R}_1(x)-\phi_1(x)\hat{\bar{R}}_{1,G}$ with $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\phi_1(x)=1$. Asymptotically we can replace $\hat{\mathcal{R}}(x)$ by
$\hat{R}(x)-\hat{\bar{R}}_G$ and $\hat{\mathcal{R}}_1(x)$ by
$\hat{R}_1(x)-\hat{\bar{R}}_{1,G}$, which leads to
$\hat{\VAR}_{0,1,2}(\hat{Y}_{G,g3reg})\approx \hat{\VAR}_{ext3}(\hat{Y}_{G,g3reg})$ and the asymptotic equivalence
of external and g-weights variance estimates.

\end{appendix}

\end{document}

### 2018-08-30 texfiles Daniel/techreporttwophase.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

--- 2017-07-28 texfiles Daniel - orig/reporttwophaserevised.tex
@@
 \doublespacing
-\nolinenumbers
-\pagestyle{empty}
-
-\title{Design-based properties of some small-area estimators in forest inventory with two-phase sampling \\ \small{revised version}}
-\author{Daniel Mandallaz, Andreas Hill, Alexander Massey \\Deparment of Environmental Systems Science \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
-\date{January 2016}
+\pagestyle{plain}
+\pagenumbering{arabic}
+
+\title{Design-based properties of some small-area estimators in forest inventory with two-phase sampling}
+\author{Daniel Mandallaz \thanks{Tel. ++41(0)44 6323186 e-mail
+    daniel.mandallaz@env.ethz.ch} \\Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland}
+\date{}
 
@@
 \newpage
-$ $
-\newpage
+\nolinenumbers
 \begin{abstract}
@@
 Nous consid\`{e}rons le probl\`{e}me de l'estimation pour petits domaines dans le contexte d'inventaires forestiers en deux phases.
-Nous proposons une am\'{e}lioration simple de l'estimateur synth\'{e}tique quand la moyenne des variables auxiliaires dans le petit domaine doit \^{e}tre estim\'{e}e en premier lieu, de m\^{e}me pour l'estimateur pour petit domain bas\'{e} sur les r\'{e}sidus.
+Nous proposons une am\'{e}lioration simple de l'estimateur synth\'{e}tique quand la moyenne des variables auxiliaires dans le petit domaine doit \^{e}tre estim\'{e}e en premier lieu     , de m\^{e}me pour l'estimateur pour petit domain bas\'{e} sur les r\'{e}sidus.
 Nous calculons la variance sous le plan de sondage de ces nouveaux estimateurs  en tenant compte de la variance des coefficients de r\'{e}gression.
@@
 \clearpage
-\section*{Foreword}
-The first version of this technical report has been published in the e-collection in September 2012. Since then many technical reports on further developments have been published in the e-collection
- (\cite{mandallazreport2,mandallazreport3,mandallazreport4,mandallazreport5,masseyphd}) as well as papers in the Canadian Journal of Forest Research, CJFR, (\cite{mandallaz3,mandallaz4,mandallaz5,massey1,massey2,massey3}.
-Recently, Dr.
-Alexander Massey and Andreas Hill, MSc, started writing an R-package implementing the various procedures discussed in the aforementioned publications.
-Writing this software revealed some minor notational ambiguities used in cluster sampling as well as small numerical discrepancies between the new case study results produced in R and the original ones obtained with the IML procedure of the SAS package.
-Detective work revealed that some SAS programs contained a minor error in a small-area variance formulae that caused the mean number of points per cluster in the small area to be taken from the second-phase instead of from the first-phase sample.
-We emphasize that the underlying concept, i.e.
-"borrowing strength" by estimating regression coefficients over an unrestricted large area to improve the calculation of point and variance estimates in a restricted small area, remains precisely the same and that all the original conclusions of this technical report, including all relevant implications on subsequent works, remain valid.
-This revised version improves the notation used for cluster sampling, presents updated tables with the corrected results for the case study and provides an enhanced discussion of the results.
-Some comments have been added to take into account results obtained since 2012.
-The mathematical derivations have also been carefully checked and, so far, confirmed.
-
-\newpage
-\pagestyle{plain}
-\pagenumbering{arabic}
 \section{Introduction}\label{introduction}
@@
 However, the resulting point estimates are not always intuitively convincing and not optimal in the model-dependent framework.
-Besides, they are not available from the usual statistical software packages.
+Beside, they are not available from the usual statistical software packages.
 For these reasons we shall only work with $\pmb{A}_{s_2}$.
@@
 \end{equation}
-with the predictions $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta_0}$ and the residuals $R_0(x)=Y(x)-\hat{Y}_0(x)$, where $\pmb{\beta_0}$ is the given external regression coefficient, ideally obtained from another similar inventory whose sample points are independent of the current inventory's sample points.
+with the predictions $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta_0}$ and the residuals $R_0(x)=Y(x)-\hat{Y}_0(x)$, where $\pmb{\beta_0}$ is the given external regression coefficient, ideally obtained from another similar inventory.
 Note that in this case the mean residual will not necessarily be zero.
@@
 where $\bar{\pmb{Z}}_G=\frac{1}{\lambda(G)}\int_G\pmb{Z}(x)dx$ is the true mean of the auxiliary vector over the small area $G$, which is available only if the first phase is exhaustive.
-$\hat{Y}_{G,synth}$ is unbiased under the model, but not optimal as it does not the take the model-dependent spatial correlation of the $Y(x)$ into account.
+$\hat{Y}_{G,synt}$ is unbiased under the model, but not optimal as it does not the take the model-dependent spatial correlation of the $Y(x)$ into account.
 Let us emphasize the fact that the model, i.e.
@@
 \\
-\textbf{In this paper we shall investigate the properties of $\hat{Y}_{G,synth}$ in the design-based inference framework}. \\
-First, let us note that $\hat{Y}_{G,synth}$ is a design-based consistent sample copy of
+\textbf{In this paper we shall investigate the properties of $\hat{Y}_{G,synt}$ in the design-based inference framework}. \\
+First, let us note that $\hat{Y}_{G,synt}$ is a design-based consistent sample copy of
  $$\frac{1}{\lambda(G)}\int_G \hat{Y}(x)dx=\frac{1}{\lambda(G)}\int_G(Y(x)-R(x))dx=\bar{Y}_G-\frac{1}{\lambda(G)}\int_G R(x)dx$$
- Consequently, the synthetic estimator $\hat{Y}_{G,synth}$ has a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_GR(x)$, which is not zero unless $G=F$ (we have zero mean residual over the entire domain, see [\ref{linearmodel2}]) or, which is unlikely, zero mean residual over the small area of interest.
+ Consequently, the synthetic estimator $\hat{Y}_{G,synt}$ has a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_GR(x)$, which is not zero unless $G=F$ (we have zero mean residual over the entire domain, see [\ref{linearmodel2}]) or, which is unlikely, zero mean residual over the small area of interest.
 Using [\ref{synthetic1}] and [\ref{estvarmatrix}] \textbf{the estimated design-based variance of the synthetic estimator} is
@@
 \begin{equation}\label{smallareaest}
-\hat{Y}_{G,small}=\hat{Y}_{G,synth}+
+\hat{Y}_{G,small}=\hat{Y}_{G,synt}+
 \frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}\hat{R}(x)
@@
 \end{equation}
-which is clearly asymptotically equivalent to $\hat{Y}_{G,synth}$ as $n_1 \to \infty$ and its design-based expected value tends to  $\bar{\pmb{Z}}^t_G\pmb{\beta}$.
+which is clearly asymptotically equivalent to $\hat{Y}_{G,synt}$ as $n_1 \to \infty$ and its design-based expected value tends to  $\bar{\pmb{Z}}^t_G\pmb{\beta}$.
 To calculate the asymptotic variance we use the decomposition (actually the first order Taylor expansion)
@@
 \end{eqnarray}
-For very large $n_{1,G}$ it is clear that the external version will underestimate the variance
+For very large $n_{1,G}$ is is clear that the external version will underestimate the variance
 as it neglects the first term in [\ref{pseudosmallvar2}], which albeit is also small for large $n_2$. \\
@@
 113 and the properties of the g-weights discussed below) so that both versions [\ref{varyregsynt1}] and [\ref{varexternalsimpleG3}] are asymptotically equivalent.
-However, version [\ref{varyregsynt1}] estimates the variance of the predictions in the large sample, which is better, and it rests upon the g-weights for the residual part, which is known to have better conditional properties (see \cite{mandallaz} p.
+However, version [\ref{varyregsynt1}] estimates the variance of the predictions in the large sample, which is better, and it rests upon the g-weights for the residual part, which it is known to have better conditional properties (see \cite{mandallaz} p.
 84 and section \ref{poststrat} for the important special case of stratification).\\
@@
 {\hat{p}_{2,G}-\hat{p}^2_{2,G}\hat{\bar{\pmb{Z}}}_{2,G}^t
-\pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$.\\
+pmb{A}^{-1}_{s_2}\hat{\bar{\pmb{Z}}}_{2,G}}$.\\
 By [\ref{estvarpseudosynth1}] we get immediately the following consistent estimate of the design-based variance
@@
  \item
- By considering formally the trivial constant local density $Y(x)\equiv 1$ and solving the normal equations one sees that for any $G$ (i.e. also for $G=F$) one has
+ By considering formally the trivial constant local density $Y(x)\equiv 1$ and solving the normal equations one sees that
  $\frac{1}{n_2}\sum_{x\in{s_2}}g_{G,1}(x)=
@@
  \item
-  When $G=F$ the estimator $\hat{Y}_{reg}$ and the sample mean $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ both converge towards the true value for an arbitrary $Y(x)$ and therefore one gets $\lim_{n_2\to\infty}g_{F,1}(x)=1>0$.
+  When $G=F$ the estimator $\hat{Y}_{reg}$ is asymptotically equivalent to the sample mean over $F$, i.e.
+to $\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$.
+This must hold for an arbitrary density $Y(x)$ and therefore one gets $\lim_{n_2\to\infty}g_{F,1}(x)=1>0$.
 This is not true for a proper subset $G\subset F$.
-In this case, $\hat{\tilde{Y}}_{G,synth}$ and the sample mean $\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}Y(x)$ converge towards the true value for an arbitrary $Y(x)$, which implies that $\tilde{g}_{G,1}(x)$ will tend to $0$ for $x\not\in{G}$ (negative values are possible) and to $\frac{n_2}{n_{2,G}} $ for $x\in{G}$.
+In this case, $\hat{\tilde{Y}}_{G,synt}$ is asymptotically equivalent to the sample mean over the small domain, i.e.
+to $\hat{\bar{Y}}_G$.
+Hence, $\tilde{g}_{G,1}(x)$ will tend to $0$  for $x\not\in{G}$ (negative values are possible) and to $\frac{n_2}{n_{2,G}} $ for $x\in{G}$.
  \end{enumerate}
-In simple random sampling one can derive non-parametric versions of the above procedures, i.e.
-kernel-based regression estimators or $k$ nearest neighbors estimators (\textbf{knn}).
-However, they are more difficult to implement than the regression estimators described above, primarily because on must use re-sampling (bootstrap) techniques to obtain reliable variance estimators because analytical results are beyond reach and the classical external variance formula severely underestimate the true variance as shown by simulations.
-Also, the performances of the non-parametric estimators are not better than the regression estimators (for global and small-area estimation) if the coefficient of determination $R^2$ is sufficient large, say above $0.7$, which is the case for timber volume assessment using modern remote sensing data as auxiliary information (e.g.
-canopy height parameters obtain via e.g.
-LiDAR).
-Details can be found in \cite{mandallazreport4}, \cite{massey2} and the PhD thesis \cite{masseyphd}.
-In the next section we generalize the previous results on regression estimators to cluster sampling, widely used in national inventories.
-The main ideas remain the same but the formulae are slightly more cumbersome due to the random cluster size.
-The non-parametric estimators are not yet available for cluster sampling.
-\newpage
+In the next section we generalize the previous results to cluster sampling. The main ideas remain the same but the formulae are slightly more cumbersome due to the random cluster size.
 \section{Generalization to cluster sampling}
-We follow the description of cluster sampling as defined in \cite{mandallaz} (especially section 5.5).
+We follow the description of cluster sampling as define in \cite{mandallaz} (especially section 5.5).
 A cluster is identified by its origin $x$, uniformly distributed in $\tilde{F}\supset F$.
@@
 $$\EX_{x\in{\tilde{F}}}M(x)(Y_c(x)-\pmb{\beta}^{t}\pmb{Z}_c(x))^2$$
-In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (though close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details).
+In the pure design-based approach the weights will be $M^2(x)$ but this leads to non-zero mean residual (thought close zero in practice), and the definitions of the regression estimator and of the normal equation are slightly different (see \cite{mandallaz}, section 5.5 for details).
 The choice of $M(x)$ rather than $M^2(x)$ as weights is suggested by the model-dependent approach.
@@
 \end{equation}
-The small-area estimation set up is slightly more complicated because the points of a cluster with $M(x) \ge 0, \;x\in{\tilde{F}}$ can be spread over more than one small area.
-The notation used in the 2012 version of this technical report was not sufficiently explicit in this respect.
-This does not affect the regression coefficients, always based on the entire sample in $F$ (the borrowing strength philosophy), but the formulation "after restricting the samples to the small area of interest" is now more explicit in the notation.
-We define the following quantities:
-\begin{eqnarray} \label{extranotation}
-M_G(x) &=& \sum_{l=1}^L  I_G(x+e_l) \nonumber \\
-Y_{c,G}(x)&=&\frac{\sum_{l=1}^L I_G(x+e_l)Y(x+e_l)}{M_G(x)} \nonumber \\
-\hat{\bar{Y}}_{c,2,G} &=& \frac{\sum_{x\in{s_{2,G}}}M_G(x)Y_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}\nonumber \\
-\pmb{Z}_{c,G}(x)&=& \frac{\sum_{l=1}^L I_G(x+e_l)\pmb{Z}(x+e_l)}{M_G(x)} \nonumber \\
-\hat{Y}_{c,G}(x)&=&\pmb{Z}^t_{c,G}(x)\hat{\pmb{\beta}}_{c,s_2} \nonumber \\
-\hat{R}_{c,G}&=& Y_{c,G}(x)-\hat{Y}_{c,G}(x) \nonumber \\
-\hat{\bar{\pmb{Z}}}_{c,1,G}&=&\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{Z}_{c,G}(x)}{\sum_{x\in{s_{1,G}}}M_G(x)}
-\end{eqnarray}
-
-The estimated variance of the one-phase estimator is given by (\cite{mandallaz}, section 4.3).
-\begin{equation}\label{onephaseclustervariance}
-\hat{\VAR}(\hat{\bar{Y}}_{c,2,G})=\frac{1}{n_{2,G}(n_{2,G}-1))}
-\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2(Y_{c,G}(x)-\hat{\bar{Y}}_{c,2,G})^2
-\end{equation}
-\noindent where $\bar{M}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)}{n_{2,G}}$.\\
-The estimate of the covariance matrix of the auxiliary variables is given by
+In two-phase sampling we estimate the mean of the auxiliary information over the small area $G$ by
+$$\hat{\bar{\pmb{Z}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{Z}_c(x)}{\sum_{x\in{S_{1,G}}}M(x)}$$
+with estimated covariance matrix
 \begin{equation}\label{estcovauxiliarycluster}
 \hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
-\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})
-(\pmb{Z}_{c,G}(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})^t
-\end{equation}
-\noindent where $\bar{M}_{1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)}{n_{1,G}}$.\\
+\sum_{x\in{s_{1,G}}}\big(\frac{M(x)}{\bar{M}_{1,G}}\big)^2(\pmb{Z}_c(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})
+(\pmb{Z}_c(x)-\hat{\bar{\pmb{Z}}}_{c,1,G})^t
+\end{equation}
+according to (\cite{mandallaz}, section 4.3).
 The \textbf{pseudo-synthetic estimate} is then
@@
 \end{equation}
-The pseudo-synthetic estimate is generally design-biased. Adjusting for the residuals we get the  \textbf{pseudo small-area estimator}
+The pseudo-synthetic estimate is generally design-biased. Adjusting for the residuals we get the  small-area estimator
 \begin{equation}\label{psmallclusterest}
-\hat{Y}_{c,G,psmall}=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\beta}}_{c,s_2}
-+ \frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}
-\end{equation}
-It is asymptotically design-unbiased and, by analogy with simple random sample (see [\ref{pseudosmallvar}]) its variance can be expected to be well approximated by
+\hat{Y}_{c,G,psmall}=\hat{\bar{\pmb{Z}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{c,s_2}}
++ \frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_{2,G}}}M(x)}
+\end{equation}
+It is asymptotically design-unbiased and intuitively its variance can be expected to be approximated by
 
@@
 + \hat{\pmb{\beta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{c,1,G}}\hat{\pmb{\beta}}_{c,s_2}
-\nonumber \\ &+& \frac{1}{n_{2,G}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\big(\frac{M_G(x)}{\bar{M}_{2,G}}\big)^2
-(\hat{R}_{c,G}(x)-\bar{\hat{R}}_{2,G})^2
-\end{eqnarray}
-where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M_G(x)$ and
-$\bar{\hat{R}}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}$.
+\nonumber \\ &+& \frac{1}{n_{2,g}(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\big(\frac{M(x)}{\bar{M}_{2,G}}\big)^2
+(\hat{R}_c(x)-\bar{\hat{R}}_{2,G})^2
+\end{eqnarray}
+where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M(x)$ and
+$\bar{\hat{R}}_{2,G}=\frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_c(x)}{\sum_{x\in{s_{2,G}}}M(x)}$.
 One can rewrite [\ref{estvarpsmallcluster}] with g-weights and predictions as in [\ref{pseudosmallvar2}].
 \par
-
 As in simple two-phase sampling we can transform the above estimator into a synthetic estimator by considering the extended model $\pmb{\mathcal{Z}}_c^t(x)=(\pmb{Z}_c^t(x),I_{c,G}(x))\in{\mathcal{R}}^{(p+1)}$ with
-$I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}=\frac{M_G(x)}{M(x)}$.
-We shall assume that $M_G(x)\equiv M(x)$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$.
-If we have a partition of the forest in small areas $G_k, k=1,2,\ldots L$, then $\tilde{F} \subset \cup_{k=1}^L \tilde{G}_k$ but it is unlikely in general that the $\tilde{G}_k$ are disjoined.
-We can reasonably hope that in extensive forest inventory the surface area of points where $M_G(x)< M(x)$ is negligible.
+$I_{c,G}(x)=\frac{\sum_{l=1}^M I_G(x_l)}{M(x)}$.
+In extensive inventories we can reasonably assume that all the points of a cluster lying in the forest area $F$ will belong to the same small area $G$ so that in fact $I_{c,G}(x) \equiv 1$ for all $x\in{\tilde{G}}=\{x \mid \sum_{l=1}^M I_G(x_l)>0\}$.
 The theoretical normal equation
@@
 \begin{equation}\label{estimatorclusterextended}
-\hat{\tilde{Y}}_{c,G,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,s_2}
+\hat{\tilde{Y}}_{c,psynth}=\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\theta}}_{c,s_2}
 \end{equation}
 where
-$\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)\pmb{\mathcal{Z}}_{c,G}(x)}
-{\sum_{x\in{s_{1,G}}}M_G(x)}$
+$\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}=\frac{\sum_{x\in{s_{1,G}}}M(x)\pmb{\mathcal{Z}}_c(x)}
+{\sum_{x\in{s_{1,G}}}M(x)}$
 is the mean of the extended auxiliary vector over the small area.\\
+Obviously a decomposition similar to [\ref{extsynthethic3}] will hold so that $\hat{\tilde{Y}}_{c,psynth}$ and $\hat{Y}_{c,G,small}$ in [\ref{psmallclusterest},\ref{estimatorclusterextended}]
+are asymptotically equivalent.\\
 As in [\ref{estvarpseudosynth1cluster}] the estimated variance is given by
@@
 \hat{\VAR}(\hat{\tilde{Y}}_{c,G,psynth})=
-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}
-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}
+\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{c,s_2}}\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}
 + \hat{\pmb{\theta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}\hat{\pmb{\theta}}_{c,s_2}
@@
 \hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}=\frac{1}{n_{1,G}(n_{1,G}-1)}
-\sum_{x\in{s_{1,G}}}\big(\frac{M_G(x)}{\bar{M}_{1,G}}\big)^2(\pmb{\mathcal{Z}}_{c,G}(x)-
+\sum_{x\in{s_{1,G}}}\big(\frac{M(x)}{\bar{M}_{1,G}}\big)^2(\pmb{\mathcal{Z}}_c(x)-
 \hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})
-(\pmb{\mathcal{Z}}_{c,G}(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})^t
-\end{equation}
-\textbf{Remark}:\\
-Even if the condition $M_G(x)\equiv M(x)$ is fulfilled in $s_{2,G}$ we cannot expect $\hat{\tilde{Y}}_{c,G,psynth}$ to be asymptotically unbiased (although the sample copy of the normal equation converges towards the theoretical normal equation this does not guarantee zero mean residuals over $G$ as we have seen).
-However the bias can be expected to be negligible.\\[0.5cm]
-Obviously a decomposition similar to [\ref{extsynthethic3}] will hold so that $\hat{\tilde{Y}}_{c,psynth}$ and $\hat{Y}_{c,G,psmall}$ in [\ref{psmallclusterest},\ref{estimatorclusterextended}]
-will be close to each other in large samples.\\
+(\pmb{\mathcal{Z}}_c(x)-\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G})^t
+\end{equation}
 Defining the g-weights at the cluster level as
 \begin{equation}
-\tilde{g}_{c,1,G}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}^t_{c,1,G}\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
+\tilde{g}_{G,c,1}(x)=\hat{\bar{\pmb{\mathcal{Z}}}}^t_{c,1,G}\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
 \pmb{\mathcal{Z}}_c(x)
@@
 \begin{equation}\label{gweightclusterpointestextended}
-\hat{\tilde{Y}}_{c,G,psynth}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)Y_c(x)
+\hat{\tilde{Y}}_{c,G,psynth}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,c,1}(x)Y_c(x)
 \end{equation}
@@
 \hat{\VAR}(\hat{\tilde{Y}}_{c,G,psynth})=\frac{1}{n_2^2}\sum_{x\in{s_2}}
-\tilde{g}^2_{c,1,G}(x)\hat{\mathcal{R}}^2(x)+
+\tilde{g}^2_{G,c,1}(x)\hat{\mathcal{R}}^2(x)+
 \hat{\pmb{\theta}}_{c,s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{\mathcal{Z}}}}_{c,1,G}}\hat{\pmb{\theta}}_{c,s_2}
 \end{equation}
+
 The synthetic estimator in the extended model corresponds formally to $n_1=\infty$, i.e.
@@
 \begin{eqnarray}
-\tilde{g}_{c,G}(x)&=&\bar{\pmb{\mathcal{Z}}}_G^t\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
+\tilde{g}_{G,c}(x)&=&\bar{\pmb{\mathcal{Z}}}_G^t\pmb{\mathcal{A}}^{-1}_{c,s_2}M(x)
 \pmb{\mathcal{Z}}_c(x)\nonumber \\
-\hat{\tilde{Y}}_{c,G,synth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,G}(x)Y_c(x)\nonumber \\
+\hat{\tilde{Y}}_{c,G,synth}&=&\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{G,c}(x)Y_c(x)\nonumber \\
 \hat{\VAR}(\hat{\tilde{Y}}_{c,G,synth})&=&\frac{1}{n_2^2}\sum_{x\in{s_2}}
-\tilde{g}^2_{c,G}(x)\hat{\mathcal{R}}^2(x)
+\tilde{g}^2_{G,c}(x)\hat{\mathcal{R}}^2(x)
 \end{eqnarray}
@@
  We have
- $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)\pmb{Z}_c(x)=
- \hat{\bar{\pmb{Z}}}_{c,1,G}$  and  $\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)\pmb{\mathcal{Z}}_c(x)=
+ $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,G,1}(x)\pmb{Z}_c(x)=
+ \hat{\bar{\pmb{Z}}}_{c,1,G}$  and  $\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,G,1}(x)\pmb{\mathcal{Z}}_c(x)=
  \hat{\bar{\pmb{\mathcal{Z}}}}_{1,G}$.
@@
  By considering $Y_c(x)\equiv 1$ one gets
- $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,1,G}(x)=
-  \frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,1,G}(x)=1$, i.e. the g-weights have means equal to 1.
- \item
-  When $G=F$ the estimator $\hat{Y}_{c,F,psynth}$ and the sample mean $\frac{\sum_{x\in{s_2}}M(x)Y(x)}{\sum_{x\in{s_2}}M(x)}$ both converge towards the true value over $F$.
-Hence, for large $n_2$, $g_{c,1,F}(x)\approx \frac{M(x)}{\bar{M}_2}$ with $\bar{M}_2=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)$.
-Likewise for $\hat{\tilde{Y}}_{G,psynth}$ and the sample mean over $G$, $\hat{\bar{Y}}_{c,2,G}=\frac{\sum_{x\in{s_{2,G}}}M_G(x) Y_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}$.
-Hence, for large $n_2$ we get $\tilde{g}_{c,1,G}(x)\approx 0$ for $x\not\in{\tilde{G}}$ (negative values are possible) and to $\tilde{g}_{c,1,G}(x)\approx \frac{M_G(x)}{\bar{M}_{2,G}}$ for $x\in{\tilde{G}}$, where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M_G(x)$.
+ $\frac{1}{n_2}\sum_{x\in{s_2}}g_{c,G,1}(x)=
+  \frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{c,G,1}(x)=1$, i.e. the g-weights have means equal to 1.
+ \item
+  When $G=F$ the estimator $\hat{Y}_{c,reg}$ is asymptotically equivalent to the sample mean over $F$, i.e.
+to $\frac{\sum_{x\in{s_2}}M(x)_cY(x)}{\sum_{x\in{s_2}}M(x)}$.
+Thus, for large $n_2$, $g_{c,F,1}(x)\approx \frac{M(x)}{\bar{M}_2}$ with $\bar{M}_2=\frac{1}{n_2}\sum_{x\in{s_2}}M(x)$.
+Likewise, $\hat{\tilde{Y}}_{G,synt}$ is asymptotically equivalent to the sample mean over the small domain, i.e.
+to
+  $\bar{Y}_{c,G,2}=\frac{\sum_{x\in{s_{2,G}}}M(x)_cY(x)}{\sum_{x\in{s_{2,G}}}M(x)}$.
+Hence, for large $n_2$ we get $\tilde{g}_{c,G,1}(x)\approx 0$ for $x\not\in{\tilde{G}}$ (negative values are possible) and to $\tilde{g}_{c,G,1}(x)\approx \frac{M(x)}{\bar{M}_{2,G}}$ for $x\in{\tilde{G}}$, where $\bar{M}_{2,G}=\frac{1}{n_{2,G}}\sum_{x\in{s_{2,G}}}M(x)$.
  \end{enumerate}
-\textbf{Remarks:}\\
-With simple random sampling the construction of design-unbiased small-area estimators as synthetic or pseudo-synthetic estimators in the extended model containing the indicator variable of the small area of interest is mathematically more convenient than the classical approach.
-The mathematical approximation of the variances is also more satisfactory than simply treating the internal model as an external one and it can be formulated within the g-weight technique, which is known to offer several theoretical advantages.
-The same essentially holds in cluster sampling, with a slight advantage maybe for $\hat{Y}_{c,G,psmall}$,with the external variance, in terms of simplicity and validity, since one cannot guarantee in general zero mean residuals over $G$ for the pseudo-synthetic estimator $\hat{\tilde{Y}}_{c,G,psynth}$ in the extended model.
-It is certainly good practice to calculate the different small-area estimators and their estimated variances, particularly for cluster sampling.
-Large differences could reveal unsuspected patterns in the data (or eventually even programming errors).\\
+
+The construction of design-unbiased small-area estimators as synthetic or pseudo-synthetic estimators in the extended model containing the indicator variable of the small area of interest is mathematically more convenient than the classical approach.
+The mathematical approximation of the variances is also more satisfactory than simply treating the internal model as an external one and it can be formulated within the g-weight technique, which is known to offer several theoretical advantages.\\
 It is mathematically clear that one can generalize all the previous results to the simultaneous estimation of $q \ge 2$ small areas by extending the model with $q$ indicator variables (combined extended model).
 One can conjecture that the combined model will be less efficient, for any given small area, than the individual estimation and, on the other hand, that it will smooth out the residual pattern.\\
-To calculate confidence intervals we recommend to use the Student's $T$ distribution on $ n_2$ d.f.
-for $F$ and $n_{2,G}-1$ d.f.
-for any small area $G$.
-If $n_{2,G}$ is to small one can use the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, which will yield short confidence intervals, at the cost of a potential bias.
-In this case it might instructive to embed $G$ in a larger domain $G^*$ with $n_{2,G^*}$ large enough and to calculate the pseudo-synthetic estimate for $G$ in the model extended with the indicator variable of $G^*$.
-\newpage
+
 \section{Generalization to two-stage sampling}
@@
 The second-stage variance is automatically taken into account.
-More mathematical details are given in \cite{mandallazreport5}.
- \newpage
 \section{Examples}
@@
 $\hat{\tilde{Y}}_{G,psynth}$ in the extended model.
-\newpage
 \subsection{Case study}
-We reanalyze the case study described with in Chapter 9 of \cite{mandallaz}.
+We reanalyze the case study described with full details in Chapter 9 of \cite{mandallaz}.
 The inventoried area covered $218ha$.
-The auxiliary information is based on 16 stands defined by the following qualitative variables obtained from the manual interpretation of aerial photographs:
+The auxiliary information is based on 16 stands defined by the following qualitative variables:
 \begin{enumerate}
@@
 \end{itemize}
-Hence, we have an additive ANOVA model with 7 parameters, as compared with 16 parameters for the full stratification model.
-The coefficient of determination $R^2$ (in the classical model-dependent approach, i.e.
-without taking the cluster structure into account) was satisfactory for the stem density ($0.53$) but not for the basal area $(0.19)$.
-This case study was performed in 1989 and is the only data available for cluster sampling in Switzerland (the reason for using cluster sampling was to provide good estimates of the variograms for the geostatistical Kriging procedures, see \cite{mandallaz} and \cite{dmhabil} for details).
-It should be emphasized that with modern remote sensing techniques (e.g.
-LiDAR or digital aerial photographs) one can obtain much better $R^2$'s, e.g.
-around $0.70$ or more for timber volume (mean canopy height being the most important explanatory variable).\\
+Hence, we have an additive ANOVA model with 7 parameters, as compared with 16 parameters for the full stratification model.\\
 We shall consider 5 small areas:
@@
 \item
+Small area $G_{21}\subset G_2$ ($\approx 7ha$) is a small subset in the central part of $G_2$ chosen to have a small number ($3$) of complete clusters ($I_{c,G}(x)\equiv 1$) spread over many different stands.
+\item
 Small area $G_3$ ($\approx 46ha$) is the most southern part of the forest.
@@
 \end{itemize}
-We have $F=G_2\cup G_3\cup G_4\cup G_5$ and $I_{c,G_k}(x) \equiv 1$ for $k=2,3,4,5$.
+We have $F=G_2\cup G_3\cup G_4\cup G_5$ and $I_{c,G_k} \equiv 1$ for $k=2,3,4,5$.
 Fig. \ref{zbergterrestrial} displays the terrestrial plots according to the domains $G_2-G_5$. Stand map of $F$ and detailed maps of $G_1$ are given in \cite{mandallaz} Chapter 8.
@@
 
-Tables \ref{result1} and \ref{result3} display the results for the basal area and the stem density for the small areas $G_1$, $G_2$, $G_3$, $G_4$ and $G_5$.\\
-The external variances for $\hat{Y}_{c,G,psmall}$ are given by the equations
+Tables \ref{newresult1} and \ref{newresult2} displays the result for the basal area and the stem density for the small areas $G_1$, $G_2$, $G_{21}$ and $G_3$.\\
+The standard error for $\hat{Y}_{c,G,small}$ are given within $(-)$ when considering the internal model as an external one, i.e. by using the formulae (see \cite{mandallaz} p. 87):
 \begin{equation*}
-\hat{Y}_{c,G,psmall}=\frac{\sum_{x\in{s_{1,G}}}M_G(x)\hat{Y}_{c,G}(x)}{\sum_{x\in{s_{1,G}}}M_G(x)}
-+\frac{\sum_{x\in{s_{2,G}}}M_G(x)\hat{R}_{c,G}(x)}{\sum_{x\in{s_{2,G}}}M_G(x)}
+\hat{Y}_{c,G,psmall}=\frac{\sum_{x\in{s_{1,G}}}M(x)\hat{Y}_{c}(x)}{\sum_{x\in{s_{1,G}}}M(x)}
++\frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_{c}(x)}{\sum_{x\in{s_{2,G}}}M(x)}
 \end{equation*}
-and
 \begin{eqnarray*}
-\widehat{\VAR}_{ext}(\widehat{Y}_{c,G,psmall})&=&\left(1-\frac{n_{2,G}}{n_{1,G}}\right)
+\widehat{\VAR}(\widehat{Y}_{c,G,psmall})&=&\left(1-\frac{n_{2,G}}{n_{1,G}}\right)
 \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}
-\sum_{x\in{s_{2,G}}}\left(\frac{M_G(x)}{\bar{M}_{2,G}}\right)^{2}(\hat{R}_{c,G}(x)-\hat{\bar{R}}_{c,2,G})^2\\
+\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}\right)^{2}(\hat{R}_c(x)-\hat{\bar{R}}_{2,G})^2\\
 &+&
-\frac{1}{n_{1,G}}\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\left(\frac{M_G(x)}{\bar{M}_{2,G}}
-\right)^{2}(Y_{c,G}(x)-\hat{Y}_{c,2,G})^2
+\frac{1}{n_{1,G}}\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}
+\right)^{2}(Y_c(x)-\hat{Y}_{2,G})^2
 \end{eqnarray*}
-The resulting standard errors are given in $(-)$.\\
-Tables \ref{result1} (basal area) and \ref{result3} (density of stems) summarize the results for the sample mean $\hat{\bar{Y}}_G$, the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, the pseudo small-area estimator $\hat{Y}_{c,G,psmall}$ and the pseudo-synthetic estimator in the extended model  $\hat{\tilde{Y}}_{c,G,psynth}$ for $F$ and the small areas $G_1$, $G_2$, $G_3$, $G_4$, $G_5$.
-\\
-The extended model for the $\hat{\tilde{Y}}_{c,G_k,psynth}$ contains only the indicator variable of the corresponding small area $G_k$.
-For this reason the corresponding estimates for the entire domain $F$ depend on the selected $G_k$ and they are given separately in Tables \ref{result2} and \ref{result4}.\\
+The standard errors given in $[-]$ refer to the various Taylor expansions given in [\ref{estvarpseudosynth1cluster}], [\ref{estvarpsmallcluster}], [\ref{estvarpsynth1clusterextended}] or their equivalent g-weights versions.\\
+
+The extended model for the
+$\hat{\tilde{Y}}_{c,G_k,psynt}$ contain only the indicator variable of the corresponding small area $G_k$.
+For this reason the corresponding estimates for the entire domain $F$ are given for $G_1$ $G_2$ and $G_3$ separately (in this order).\\
  We also consider the joint estimation of $F$ and the small areas $G_2$, $G_3$, $G_4$ and $G_5$, which form a partition of $F$.
-The corresponding model $\pmb{\mathcal{Z}}(x)$ contains the $4$ indicator variables $I_{G_k}(x)$ ($k=2,3,4,5$), the previous components $Z_l(x), l=2,3,4,5,6$ but no longer the intercept term $Z_1(x) \equiv 1$ (otherwise $\pmb{\mathcal{A}}_{s_2}$ would be singular because $Z_1(x)$ is a linear combination of the $I_{G_k}(x)$).
-The results for this estimator, denoted by $ \hat{\tilde{Y}}_{c,G,cpsynth}$, are displayed in Table  \ref{result5}.\\
- All the calculations were performed with the linear algebra procedure \textbf{proc IML} of the statistical software package SAS.
-
+The corresponding model $\pmb{\mathcal{Z}}(x)$ contains the $4$ indicator variable $I_{G_k}(x)$ ($k=2,3,4,5$), the previous components $Z_l(x), l=2,3,4,5,6$ but no longer the intercept term $Z_1(x) \equiv 1$ (otherwise $\pmb{\mathcal{A}}_{s_2}$ would be singular because $Z_1(x)$ is a linear combination of the $I_{G_k}(x)$).
+The results for this estimator, denoted by $ \hat{\tilde{Y}}_{c,G,cpsynt}$, are displayed in Table  \ref{newresult3}.\\
+ All the calculations were performed with the linear algebra procedure \textbf{proc iml} of the statistical software package SAS.
 \section{Discussion and conclusions}
- The results for the basal area are disappointing, except maybe for global estimation, in the sense that the regression estimators, with the exception of the pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$, frequently have a larger error than the sample mean.
-This is of course due to the very low coefficient of determination $R^2=0.19$ and we shall therefore not pursue the discussion for basal area, simply noting that all point estimates were close to each other, with the exception of the small area $G_1$ with the full census, for which the pseudo-synthetic estimator was closer to the true values (also for the density of stems).
-As confirmed by simulations this was due to the fact that the plots within this small area were in the lower tail of the distribution for both basal area and the density of stem.
-\\
-As expected for the stem density (with an acceptable $R^2=0.53$) the two-phase estimators perform much better than the sample mean.
-For the classical small-area estimator $\hat{Y}_{c,G,psmall}$ the standard errors based on the external model assumption were smaller than their counterparts based on the g-weights (i.e.
-on the design-based covariance matrix of the regression coefficients) for $F,G_3,G_4,G_5$ and slightly larger for $G_1,G_2$, but the differences were small, a reassuring result.
-The point estimates of  $\hat{Y}_{c,G,psmall}$ were very close to the point estimates of the pseudo-synthetic estimator $\hat{\tilde{Y}}_{c,G,psynth}$ (in the extended model with the single indicator variable of the small area of interest), which has slightly smaller g-weights standard errors, except for $F$ and $G_1$.
-From a practical point of view $\hat{Y}_{c,G,psmall}$ and $\hat{\tilde{Y}}_{c,G,psynth}$ are essentially equivalent.
-The estimators $\hat{\tilde{Y}}_{c,G,psynth}$, $\hat{Y}_{c,G,psmall}$ and the combined estimator $\hat{\tilde{Y}}_{c,G,cpsynth}$ (model without intercept but extended with the 4 indicator variables of $G_2,G_3,G_4,G_5$) were also practically equivalent for $F,G_2,G_3,G_4,G_5$, the combined estimator having slightly smaller g-weights standard errors.\\
-The pseudo-synthetic estimator $\hat{Y}_{c,G,psynth}$ yields point estimates close to those of
-$\hat{Y}_{c,G,psmall}$, $\hat{\tilde{Y}}_{c,G,psynth}$, $\hat{\tilde{Y}}_{c,G,cpsynth}$ (recall that
-$\hat{Y}_{c,F,psmall}=\hat{Y}_{c,F,psynth}$) for $F, G_2,G_3,G_4,G_5$ but, of course, with much smaller g-weights based standard errors, at a cost of a potential bias.
-For $G_1$ the point estimate $\hat{Y}_{c,G_1,psynth}$ was very close to the true value (for the density of stem and the basal area), precisely because the clusters hitting $G_1$ were, as afore mentioned, not "representative" for $G_1$.
-This can happen!\\
-From a mathematical point of view the g-weights technique in the models extended by the indicator variables of the small areas is without any doubts the most elegant approach in simple random sampling: it bypasses the residuals terms and allows for a straightforward calculation of the asymptotic variance that takes into account the errors of the regression coefficients.
-The same is likely to hold in large sample also with cluster sampling.
-The results from the case study show that the differences between the asymptotically unbiased estimators are very small from a practical point of the view, and that a potentially biased estimator can by chance be closer to the truth.
+In the case study all point estimates were close to each other and do not differ significantly from each other.
+In the small area with full census the synthetic estimator was closer to the true values.
+As confirmed by simulations this was due to the fact that the plots within this small area were in the lower tail of the distribution for basal area and stem density.
+Of course, the synthetic estimators always had the smallest standard errors but at the potential cost of a local bias.\\
+For the classical small-area estimator the standard errors based on the external model assumption were usually, but not always, smaller than their counterparts based on the g-weights (equivalent to the Taylor asymptotic expansions), but the differences were small, a reassuring result.
+The g-weights based standard errors of the synthetic estimators in the extended model for one single small area were usually smaller than their g-weights counterparts of the classical small-area estimator but generally still larger than under the external model assumption.
+The g-weights based standard errors in the extended model with several small areas were comparable to those derived specifically for one single small area.
+In this case study  the various methods can be regarded as practically almost equivalent, which should be confirmed or eventually invalidated by further examples.\\
+From a mathematical point of view the Taylor based g-weights technique in the models extended by the indicator variables of the small areas is without any doubts the most elegant approach: it bypasses the residuals terms and allows for a straightforward calculation of the asymptotic variances that takes into account the errors of the regression coefficients.
 
@@
 \newpage
-\begin{table}[h]
-\begin{center}
-\caption{ \label{result1}\textbf{Results for basal area}}
-\end{center}
-\begin{center}
- \begin{tabular}{l|c|c|c|c|c|}\hline
-  \textbf{Domain}     & $n_1:n_2$  & $\hat{\bar{Y}}_G$ & $\hat{Y}_{c,G,psynth}$  & $\hat{Y}_{c,G,psmall}$ & $ \hat{\tilde{Y}}_{c,G,psynth}$ \\[1.0ex]
-  & $n_1\bar{M}_1:n_2\bar{M}_2$   & &   & & \\[1.0ex]
-  $ $ & $ $ & $ $ &  $ $ & $ $ &  \\ \hline\hline
- $F$     & $ 298:73$     & $31.90$   &  $31.34$     &  $31.34$       & see Table \ref{result2}    \\ [1.0ex]
-         & $1203:298$    & $(1.08)$  &  $[0.94]$    &  $[0.94]$      & $ $     \\ [1.0ex]
-         &               &           &  $(0.89)$    &  $(0.89)$      & $ $     \\ \hline\hline
- $G_1$   & $29:8$        &  $24.54$  & $30.28$      &  $23.99$       & $25.55$            \\ [1.0ex]
- true=29.60  & $92:19$   &  $(3.59)$ & $[1.37]$      &  $[3.96]$      & $[3.76]$          \\ [1.0ex]
-         & $ $           &           &              &  $(3.68)$      &  $ $             \\ \hline \hline
- $G_2$ & $49:9$         & $30.69$     &  $28.27$    &  $29.32 $      & $29.31$     \\ [1.0ex]
-       & $185:41$       & $(2.05)$    & $[1.47]$    &  $[2.55]$      & $[2.33]$     \\ [1.0ex]
-       & $ $            &   $ $   &   $ $           &  $(2.08)$      &  $ $         \\ \hline \hline
- $G_3$ & $73:18$   &  $32.32$         & $31.62$     &  $31.46$       & $31.46$      \\ [1.0ex]
-       &$250:66$   &  $(2.05)$        & $[1.54]$    &  $[2.37]$      & $[2.18] $    \\[1.0ex]
-       &           &   $ $   &                      &  $(1.87)$      & $ $               \\ \hline \hline
- $G_4$ & $81:17$   &  $27.63$         & $29.72$     &  $27.78$       & $27.77$          \\ [1.0ex]
-       & $306:69$  &  $(2.15)$        & $[1.15] $   &  $[2.27]$      & $[1.98]$        \\ [1.0ex]
-       & $ $       &  $ $    &  $ $                 &  $(2.00)$      & $ $            \\ \hline \hline
- $G_5$ &  $125:29$ &  $34.49$         &  $33.50$    & $34.33$        & $34.33$           \\ [1.0ex]
-       & $462:122$ &  $(1.78)$        &$[1.00]$     & $[1.55]$       & $[1.30]$           \\ [1.0ex]
-       &           &  $ $             & $$          & $(1.35)$       & $ $    \\    \hline \hline
-
-\end{tabular}
-\end{center}
-\textbf{Design-based standard errors:} $[-]$ (g-weights based) and $(-)$ (based on external or sample variance)\\
-For $G_1$ the mean residual of $ \hat{\tilde{Y}}_{c,G,psynth}$ was $-1.59$ and $0$, as expected, for all other domains.
-\end{table}
-
-\begin{table}[h]
-\begin{center}
-\caption{ \label{result2}\textbf{Results basal area for $F$ with extended models}}
-\end{center}
-\begin{center}
- \begin{tabular}{|c|c|c|c|c|}\hline
-  $\hat{\tilde{Y}}^{(1)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(2)}_{c,F,psynth}$& $\hat{\tilde{Y}}^{(3)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(4)}_{c,F,psynth}$ &  $\hat{\tilde{Y}}^{(5)}_{c,F,psynth}$ \\[1.0ex] \hline \hline
-  $31.30$    & $31.36$   & $31.35$   & $31.30$  & $31.33$ \\[1.0ex]
-  $[0.92] $  & $[0.93]$  & $[0.94]$  & $[0.92]$ & $[0.94]$ \\[1.0ex] \hline \hline
-\end{tabular}
-\end{center}
-$\hat{\tilde{Y}}^{(k)}_{c,F,psynth}$ is the pseudo-synthetic estimate for $F$ in the model extended with the indicator variable of domain $G_k$, $k=1,2,3,4,5$.
-The g-weights based standard errors are given in $[-]$.
-\end{table}
-
-\newpage
-
-\begin{table}[h]
-\begin{center}
-\caption{ \label{result3}\textbf{Results for stem density}}
-\end{center}
-\begin{center}
- \begin{tabular}{l|c|c|c|c|c|}\hline
-  \textbf{Domain}     & $n_1:n_2$  & $\hat{\bar{Y}}_G$ & $\hat{Y}_{c,G,psynth}$  & $\hat{Y}_{c,G,psmall}$ & $ \hat{\tilde{Y}}_{c,G,psynth}$ \\[1.0ex]
-  & $n_1\bar{M}_1:n_2\bar{M}_2$   & &   & & \\[1.0ex]
-  $ $ & $ $  & $ $ & $ $   & $ $  &  \\ \hline\hline
- $F$  &    $ 298:73$ & $321.03$  &  $325.79$   &  $325.79 $       & see Table \ref{result4}     \\ [1.0ex]
-      & $1203:298$   & $(18.44)$ & $[12.80 ]$  &  $[12.80 ]$      & $ $     \\ [1.0ex]
-      &              &           & $(12.09 )$  &  $(12.09)$      & $ $     \\ \hline\hline
- $G_1$ & $29:8$      & $245.61$  & $279.54 $   &  $257.34$       & $258.20$            \\ [1.0ex]
- true=$280.23$ & $92:19$ & $(65.51)$ & $[27.26]$   &  $[48.26]$       & $[49.56]$          \\ [1.0ex]
-       & $ $         &           &             &  $(48.29 )$      &  $ $         \\ \hline \hline
- $G_2$ & $49:9$      & $393.50$  & $400.49 $   &  $406.47 $       & $406.41$     \\ [1.0ex]
-       & $185:41$    & $(79.69)$ &$ [30.95 ]$  &  $[42.96 ]$      & $[39.36 ]$     \\ [1.0ex]
-       & $ $       &   $ $       &   $ $       &  $(43.49 )$      &   $ $         \\ \hline \hline
- $G_3$ & $73:18$   &  $275.25$   & $279.75 $   &  $282.46 $       & $282.40$      \\ [1.0ex]
-       &$250:66$   &  $(20.98)$  & $[17.16 ]$  &  $[22.68 ]$      & $[20.77]$    \\[1.0ex]
-       &           &   $ $       &             &  $(16.56)$       & $ $       \\ \hline \hline
- $G_4$ & $81:17$   &  $284.06$   &  $307.85$   &  $274.79$        & $274.53 $          \\ [1.0ex]
-       & $306:69$  &  $(39.51)$  &  $[19.69]$  &  $[26.64]$       & $[23.55]$        \\ [1.0ex]
-       & $ $       &  $ $        &  $ $        &  $(24.12)$       & $ $            \\ \hline \hline
- $G_5$ &  $125:29$ &  $342.35$   &  $332.66$   &  $347.89$        & $347.91 $           \\ [1.0ex]
-       & $462:122$ &  $(24.50)$  &  $[15.33]$  &  $[21.26]$       & $[18.65]$           \\ [1.0ex]
-       &           &  $ $     & $$             &  $(17.49)$       & $ $       \\    \hline \hline
-
-\end{tabular}
-\end{center}
-\textbf{Design-based standard errors:} $[-]$ (g-weights based) and $(-)$ (based on external or sample variance)\\
-For $G_1$ the mean residual of $ \hat{\tilde{Y}}_{c,G,psynth}$ was $-0.99$ and $0$, as expected, for all other domains.
-\end{table}
-
-\begin{table}[h]
-\begin{center}
-\caption{ \label{result4}\textbf{Results for stem density for $F$ with extended models}}
-\end{center}
-\begin{center}
- \begin{tabular}{|c|c|c|c|c|}\hline
-  $\hat{\tilde{Y}}^{(1)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(2)}_{c,F,psynth}$& $\hat{\tilde{Y}}^{(3)}_{c,F,psynth}$  & $\hat{\tilde{Y}}^{(4)}_{c,F,psynth}$ &  $\hat{\tilde{Y}}^{(5)}_{c,F,psynth}$ \\[1.0ex] \hline \hline
-  $325.62$    & $325.88$  & $325.72$   & $325.15$   & $325.56$ \\[1.0ex]
-  $[12.81]$   & $[12.84]$ & $[12.85]$  & $[12.67]$  & $[12.62]$ \\[1.0ex] \hline \hline
-\end{tabular}
-\end{center}
-$\hat{\tilde{Y}}^{(k)}_{c,F,psynth}$ is the pseudo-synthetic estimate for $F$ in the model extended with the indicator variable of domain $G_k$, $k=1,2,3,4,5$.
-The g-weights based standard errors are given in $[-]$.
-\end{table}
-
-\begin{table}[h]
-\begin{center}
-\caption{ \label{result5}\textbf{Two-phase combined estimates}}
-\end{center}
-\begin{center}
- \begin{tabular}{|l|c|c|c|}
-   \textbf{Domain} & \textbf{sample sizes} & \textbf{basal area}  & \textbf{stem density}\\[1.0ex]
- $ $   & $n_1:n_2$    &  & \ \\[1.0ex]
- $ $   & $n_1\bar{M}_1:n_2\bar{M}_2$  & $ \hat{\tilde{Y}}_{c,G,cpsynth}$ &$\hat{\tilde{Y}}_{c,G,cpsynth}$ \\ \hline
- $F$            & $ 298:73$   &  $31.32 $    &  $325.17$       \\ [1.0ex]
-                & $1203:298$  &  $[0.93]$    &  $[12.62]$       \\ \hline \hline
- $G_2$          &  $49:9$     &  $29.39 $    &  $407.84$        \\ [1.0ex]
-                &  $185:41$   &  $[2.23]$    &  $[39.04]$       \\ \hline \hline
- $G_3$          &  $73:18$     &  $31.57 $    &  $284.17$         \\ [1.0ex]
-                & $250:66$    &  $[2.09]$    &  $[18.09]$       \\ \hline \hline
- $G_4$          &  $81:17$     &  $27.77 $    &  $274.59$        \\ [1.0ex]
-                & $306:69$    &  $[1.99]$   &  $[23.49]$        \\ \hline \hline
- $G_5$          &  $125:29$    &  $34.31$    &  $347.76$          \\ [1.0ex]
-                & $462:122$   &  $[1.24]$   &  $[16.61]$         \\ \hline \hline
-\end{tabular}
-\end{center}
-The g-weights based standard errors are given in $[-]$.
-As expected on mathematical grounds all extended models yielded empirical means of the residuals over the entire domain and over one or many small areas which were equal to zero ($<10^{-12})$.
-\end{table}
-
-\end{document}
-
 \begin{table}[h]
@@
 
-The standard error for $\hat{Y}_{c,G,small}$ are given within $(-)$ when considering the internal model as an external one, i.e. by using the formulae (see \cite{mandallaz} p. 87):
-\begin{equation*}
-\hat{Y}_{c,G,psmall}=\frac{\sum_{x\in{s_{1,G}}}M(x)\hat{Y}_{c}(x)}{\sum_{x\in{s_{1,G}}}M(x)}
-+\frac{\sum_{x\in{s_{2,G}}}M(x)\hat{R}_{c}(x)}{\sum_{x\in{s_{2,G}}}M(x)}
-\end{equation*}
-\begin{eqnarray*}
-\widehat{\VAR}(\widehat{Y}_{c,G,psmall})&=&\left(1-\frac{n_{2,G}}{n_{1,G}}\right)
-\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}
-\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}\right)^{2}(\hat{R}_c(x)-\hat{\bar{R}}_{2,G})^2\\
-&+&
-\frac{1}{n_{1,G}}\frac{1}{(n_{2,G}-1)}\sum_{x\in{s_{2,G}}}\left(\frac{M(x)}{\bar{M}_{2,G}}
-\right)^{2}(Y_c(x)-\hat{Y}_{2,G})^2
-\end{eqnarray*}
-The standard errors given in $[-]$ refer to the various Taylor expansions given in [\ref{estvarpseudosynth1cluster}], [\ref{estvarpsmallcluster}], [\ref{estvarpsynth1clusterextended}] or their equivalent g-weights versions.\\
-
-The extended model for the
-$\hat{\tilde{Y}}_{c,G_k,psynt}$ contain only the indicator variable of the corresponding small area $G_k$.
-For this reason the corresponding estimates for the entire domain $F$ are given for $G_1$ $G_2$ and $G_3$ separately (in this order).\\
-
 \end{document}

### 2018-08-30 texfiles Daniel/two_threephase_twostage.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

\documentclass[a4paper,12pt,leqno, titlepage]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\bibliographystyle{mystyle2}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\VAR}{\mathbb{V}}
\newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{R}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}
\setlength{\topmargin}{0cm}
\setlength{\headheight}{3mm}
\setlength{\headsep}{1cm}
\setlength{\topskip}{0cm}
\setlength{\textwidth}{16cm}
\setlength{\textheight}{22cm}
\setlength{\oddsidemargin}{0cm}
\makeatletter
\def\tagform@#1{\maketag@@@{[\ignorespaces#1\unskip\@@italiccorr]}}
\makeatother

\begin{document}
\doublespacing
\pagestyle{plain}
\pagenumbering{arabic}

\title{Mathematical details of two-phase/two-stage and three-phase/two-stage regression estimators in forest inventories: design-based Monte Carlo approach}
\author{Daniel Mandallaz \thanks{Tel. ++41(0)44 6323186 e-mail
    daniel.mandallaz@env.ethz.ch} \\Department of Environmental Systems Science\\
    Chair of Land Use Engineering\\ETH Zurich\\CH 8092 Zurich, Switzerland \\}
\date{Mai 2015}

\maketitle
\newpage
\nolinenumbers
\begin{center}
\section*{Foreword}\label{foreword}
\end{center}
This technical report discusses some mathematical details of regression estimators under two or three-phase sampling with two-stage sampling of trees at the plot level. Discussions with colleagues and graduate students convinced me that the mathematical derivations presented (and often simply sketched) in previous work deserved clarification. This is particularly true, because of space restriction, for the papers \cite{mandallaz3}, \cite{mandallaz4}, \cite{mandallaz5} and to a lesser degree for the the book \cite{mandallaz} and the technical reports (\cite{mandallazreport1}, \cite{mandallazreport2}, \cite{mandallazreport3}). The great physicist and Nobel prize laureate Richard Feynman once said, after a life-time frequentation of mathematicians, that he finally understood what the word obvious really means, namely proved. Of course, the level he was referring to is stratospheric in comparison with the down-to-earth properties of regression estimators in forest inventory. Nevertheless, I thought it would not be harmful, a few months before my retirement, to clarify some theoretical issues in survey sampling, in particular the differences between design-based model-assisted versus model-dependent inference, which are not always well understood by practitioners, as I can now judge after 25 years experience as a reviewer and author. It turns out that all the previously published results are indeed correct, at least asymptotically. The main point is that one can use mutatis mutandis the two-phase/three-phase one-stage formulae also with two-stage sampling at the plot level. The second-stage variance resulting from Poisson sampling at the plot level is so to speak automatically taken into account. This holds in particular for the easy to use so-called external variances. In contrast to previous work the notation is more pedantic, which is the price to pay to be as clear and accurate as possible.
\newpage
\section{Introduction}\label{intro}
In the forested area $F$ we consider a well-defined population $ \mathcal{P}$ of $N$ trees with response variable $Y_i,\;i=1,2 \ldots$, e.g. the timber volume.  \textbf{The objective is to estimate the overall spatial mean}  $\bar{Y}=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i$, where $\lambda(\cdot)$ denotes the surface area (usually in ha) and \textbf{the spatial mean over F} is defined as
 \begin{equation}\label{small1}
 \bar{Y}_F=\frac{1}{\lambda(F)}\sum_{i=1}^N Y_i=:\frac{1}{\lambda(F)}\sum_{i\in{F}}Y_i
 \end{equation}
Let us first recall the definitions of local density and generalized local density.
For a random point $x$  uniformly distributed in $F$ trees are selected at $x$ from the population $\mathcal{P}$ for instance with concentric circles or angle count techniques. Each tree has its associated inclusion circle $K_i$ centered on the tree. The
set of trees selected at point $x$ is denoted by $s_{2}(x)$. The first-stage inclusion probabilities are $\pi_i=\frac{\lambda(K_i\cap F}{\lambda(F)}$. For each of the
selected trees $i\in{s_{2}(x)}$ one determines $Y_i$. The indicator variable $I_i$ is defined as
\begin{equation}\label{1stage}
 I_i(x)=\begin{cases}&1 \text{ if $i\in s_{2}(x)$}\\
                      &0 \text{ if $i\not\in s_{2}(x)$}
         \end{cases}
\end{equation}
At each point  the
terrestrial inventory provides the \textbf{local density} $Y(x)$
\begin{equation}\label{truelocaldensity}
 Y(x) =\frac{1}{\lambda(F)}\sum_{i=1}^N \frac{I_i(x)Y_i}{\pi_i}=\frac{1}{\lambda(F)}\sum_{i\in{s}_2(x)} \frac{Y_i}{\pi_i}
 \end{equation}
 The term $\frac{1}{\lambda(F)\pi_i}$ is the tree extrapolation factor $f_i$ with dimension $ha^{-1}$. One must include possible boundary adjustments at the forest edge: $\lambda(F)\pi_i=\lambda(F \cap K_i)$. In the infinite population approach (now better known as Monte Carlo approach) one samples the function $Y(x)$ (\cite{mandallaz}) for which the following important relation holds:
 \begin{equation}\label{montecarlo}
 \EX_{x} (Y(x))=\frac{1}{\lambda(F)}\int_{F} Y(x)dx=\frac{1}{\lambda(F)}\sum_{i=1}^NY_i=\bar{Y}_F
 \end{equation}
 Where $\EX_x$ denotes the expectation with respect to a random point $x$ uniformly distributed in $F$. This establishes the link between the infinite population (continuum) $\{x\in{F} \mid Y(x)\}$ and the finite population of trees $\{i=1,2 \ldots N \mid Y_i\}$. The infinite populations and super-populations models were introduced in my PhD thesis (\cite{dmphd}) to solve the obvious contradiction that a domain of the plane cannot be viewed (i.e. partitioned) as a finite population of single or concentric circles (even infinitely many with the angle count). Furthermore, the local density $Y(x)$ as defined above is a regionalized variable in the sense of Matheron (\cite{matheron1}), who developed the purely model-dependent geostatistical Kriging techniques with primarily the estimation of mining or oil resources in mind (where one has to estimate 3-dimensional integrals). The application of geostatistical methods in forest inventory was therefore quite natural in the Monte Carlo approach and led in particular to the model-dependent Double Kriging procedure discussed in my habilitation thesis (\cite{dmhabil}).\\
 Readers totally unfamiliar with concepts such as Monte Carlo approach, design-based versus model-dependent inference (unfortunately sometimes also called model-based inference) should maybe consult for a first perusal the paper of \cite{gregoire3} on general survey sampling and the text book of \cite{mandallaz} (chapter 4, 6 and 7) in the context of Monte Carlo approach to forest inventory. \\[0.5cm]
 In many applications costs to measure the response variable $Y_i$ are high. For instance,
a good determination of the volume may require that one records
$DBH$, as well as the diameter $D_7$  at $7m$ above ground and total height $H$
in order to utilize a three-way volume function. However, one could rely
on a coarser, but cheaper, approximation of the volume based only on
$DBH$ and species alone. Nonetheless, it may be most sensible to assess those three
parameters only on a sub-sample of trees. We now briefly formalize this
simple idea, which is used in the Swiss National Forest Inventory. The reader is referred to (\cite{mandallaz}, section 4.4, 4.5, 5.4 and 9.5) for details. \\
From each of the selected trees $i\in{s_{2}(x)}$ one
determines coarse approximation $\hat{Y}_i$ (e.g. based on species and $DBH$) of the pseudo exact value $Y_i$ (pseudo exact in the sense that is is a better approximation of the unknown true bole volume based on e.g. species, $DBH$, $D_7$ and $H$). From the
finite set $s_{2}(x)$ one draws a sub-sample
$s_{3}(x)\subset{s_{2}(x)}$ of trees by Poisson sampling. For each tree $i\in{s_{3}(x)}$
one then determines the (pseudo) exact variable $Y_i$. The second stage indicator variable is
\begin{equation}
 J_i(x)=\begin{cases}&1 \text{ if $i\in s_{3}(x)$}\\
                      &0 \text{ if $i\not\in s_{3}(x)$}
         \end{cases}
\end{equation}

 The residual s at the tree level is denoted by $R_i=Y_i-\hat{Y}_i$ which is known only for trees
 $i\in{s_{3}(x)}$. The \textbf{generalized local density} $Y^*(x)$ is defined
 according to
 \begin{eqnarray}\label{gdens}
 Y^*(x)&=&\frac{1}{\LF}\left( \sum_{i=1}^N \frac{I_{i}(x)\hat{Y}_i}{\pi_i} +
 \sum_{i=1}^N \frac{I_{i}(x)J_{i}(x)R_i}{\pi_{i}p_i}\right)\nonumber \\
 &=&\frac{1}{\LF}\left( \sum_{i\in{s_{2}(x)}} \frac{\hat{Y}_i}{\pi_i}
 +\sum_{i\in{s_{3}(x)}} \frac{R_i}{\pi_{i}p_i}\right)
 \end{eqnarray}
 where the $p_i$ are the conditional inclusion probabilities for the the second stage sampling, i.e. $p_i=\PR(J_i(x)=1 \mid I_i(x)=1)$. The second term is an Horwitz-Thomson estimator and the generalized local density can be viewed as a difference estimator (\cite{sarndal}, section 6.3). The expected value and the variance under Poisson sampling at point $x$ are easily found to be
 \begin{eqnarray}\label{poisson_atx}
  \EX_{Poisson}(Y^*(x)) &=& Y(x) \nonumber \\
  \VAR_{Poisson}(Y^*(x))&=&\frac{1}{\LFC}\big(\sum_{i\in{s_{2}(x)}} \frac{R_i^2(1-p_i)}{\pi_i^{2}p_i} \big)=:V(x)
 \end{eqnarray}
If necessary $V(x)$ can be estimated unbiasedly by
\begin{equation}\label{estvarvofx}
\hat{V}(x)=\frac{1}{\LFC}\big(\sum_{i\in{s_{3}(x)}} \frac{R_i^2(1-p_i)}{\pi^2_ip^2_i}\big)
\end{equation}
\clearpage \newpage

\section{Two-phase sampling}\label{twophase}
 The \textbf{first phase} draws a large sample $s_1$ of $n_1$ points that are independently and uniformly distributed within the forest area $F$. At each point $x\in{s}_1$ auxiliary
information is collected and denoted by $\mathbf{Z}(x)\in{\RE ^{p}}$. The \textbf{second phase} draws a small sample $s_2\subset{s_1}$ of
$n_2$ points from $s_1$ according to \textbf{equal probability
sampling without replacement, or simple random sampling SRS}.
 The first phase randomization procedure will be indicated by the subscript $1$, the second phase by the subscripts $2$ or $x$ and the second-stage procedure at a given point $x$ by the subscript $3$. Note that the points $x\in{s_2}$ inherit from $s_1$ the property of being i.i.d. uniformly distributed in $F$. One has directly from the definition of Poisson sampling and the decomposition rule
 $$\VAR_{2,3}Y^*(x)=\VAR_2 \EX_{3 \mid 2}Y^*(x) + \EX_2 \VAR_{3 \mid 2}(Y^*(x))$$
 the following properties
 \begin{eqnarray}\label{twostage}
 \EX_{3\mid 2} Y^*(x)&=& Y(x)\nonumber \nonumber \\
 \EX_{2,3}Y^*(x)&=&\EX_x(Y(x))=\bar{Y}_F \nonumber \\
 \VAR_{2,3}(Y^*(x))&=&\VAR_x(Y(x)) + \EX_x(V(x)) \nonumber \\
 &=&\VAR_x(Y(x))+\frac{1}{\LFC}\Big(\sum_{i=1}^N\frac{(R_i^2(1-p_i)}{\pi_ip_i}\Big)
 \end{eqnarray}

 \subsection{The model}
 In the \textbf{design-based approach} we set
 \begin{equation}\label{linearmodel1}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R_{\pmb{\beta}}(x)=:\hat{Y}_{\pmb{\beta}}(x)+R_{\pmb{\beta}}(x)
 \end{equation}
 the quantities $Y(x), \pmb{Z}(x), R_{\pmb{\beta}}(x)$ are random because the point $x$ is random and uniformly distributed in $F$. Equation [\ref{linearmodel1}] can be viewed, for any fixed given $\pmb{\beta}$, as simply defining the residual term $R_{\pmb{\beta}}(x)$. \textbf{The true regression coefficient} $\pmb{\beta}_0$ \textbf{is by definition the theoretical least squares estimate}
 \begin{equation}\label{lstheoretical}
 \pmb{\beta}_0=\arg \min_{\pmb{\beta}}\int_F (Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2dx
 \end{equation}
 It satisfies the normal equation
 \begin{equation}\label{normaleq1}
 \Big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\Big)\pmb{\beta}_0=\int_F Y(x)\pmb{Z}(x)dx
 \end{equation}
 and the orthogonality relationship
 \begin{equation}\label{linearmodel2}
 \int_F R_0(x)\pmb{Z}(x)dx=\pmb{0}
 \end{equation}
 where $R_0(x)=R_{\pmb{\beta}_0}=Y(x)-\pmb{Z}^t(x)\pmb{\beta}_0=:Y(x)-\hat{Y}_0(x)$.
  We shall assume that $\pmb{Z}(x)$ contains the intercept term $1$, or, more generally, that the intercept can be expressed as a linear combination of the component of $\pmb{Z}(x)$, which then insures the following properties
  \begin{equation}\label{truerespred}
  \frac{1}{\lambda(F)}\int_F R_0(x)dx=0\; \text{and}\;\frac{1}{\lambda(F)}\int_{F}\hat{Y}_0(x)=\frac{1}{\lambda(F)}\int_{F}Y(x)dx=\bar{Y}
  \end{equation}
In the \textbf{model-dependent approach} one postulates a model
 \begin{equation}\label{linearmodel1mod_dep}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R(x)
 \end{equation}
 \noindent
 where \textbf{for a fixed point $x$} the quantities $Y(x)$ and $R(x)$ are viewed random variables, $\pmb{R}(x)$ with zero mean and a given covariance structure. The vector of auxiliary $\pmb{Z}(x)$ can have for a given $x$ non-random or random components. We emphasize the fact that in the design-based model-assisted approach the model [\ref{linearmodel1}] is not viewed as the true complex stochastic process generating the $Y(x)$, but, more pragmatically, \textbf{simply as a tool to get better estimators of $\bar{Y}$ than the ordinary sample mean (which corresponds to the model $\pmb{Z}(x)$ containing only the intercept and therefore no auxiliary information at all}. Of course, ideally, the model should capture qualitatively the main features of the underlying natural phenomenon. Hence, one could say that in the design-based approach the forest under investigation is a \textbf{fixed entity observed at random points}, whereas in the model-dependent set-up the actual forest is viewed as \textbf{the realization of a complex stochastic process observed at given fixed points}. That these two totally different approaches yield usually (but not always, particularly for small area estimation) similar results in practice is in some sense a miracle.\\
 To simplify the notation we set
 set $\pmb{A}=\EX_x\pmb{Z}(x)\pmb{Z}^t(x)$, $\pmb{U}(x)=Y(x)\pmb{Z}(x)$.
 The normal equation then reads
 $$\pmb{A}\pmb{\beta}_0=\EX_x\pmb{U}(x):=\pmb{U}$$
 Of course, only a sample-based normal equation is available, i.e.
 $$\pmb{A}_{s_2}\hat{\pmb{\beta}}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{U}(x)=\pmb{U}_{s_2}$$
 where we have set
 $$\pmb{A}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x)$$ and
 $$\pmb{U}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)$$
 The theoretical and empirical regression vector parameters are
 \begin{eqnarray}\label{linearmodel3}
 \pmb{\beta}_0&=&\pmb{A}^{-1}\pmb{U} \nonumber \\
 \hat{\pmb{\beta}}_{s_2}&=&\pmb{A}_{s_2}^{-1}\pmb{U}_{s_2}
 \end{eqnarray}
$\hat{\pmb{\beta}}_{s_2}$ is asymptotically design-unbiased for
$\pmb{\beta}_0$.
It can be shown that the asymptotic design-base variance-covariance matric of $\hat{\pmb{\beta}}_{s_2}$ is given by
\begin{equation}\label{varmatrix}
\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}=
\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX_xR_0^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}^{-1}
\end{equation}
which can be estimated by replacing the theoretical residuals $R_0(x)=Y(x)-\pmb{Z}^t(x)\pmb{\beta}_0$
with their empirical counterparts
$\hat{R}_{s_2}(x)=Y(x)-\hat{Y}_{s_2}(x)$, with $\hat{Y}_{s_2}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$, and $\pmb{A}$ with $\pmb{A}_{s_2}$. We then get the \textbf{estimated design-based
variance-covariance matrix} as
\begin{equation}\label{estvarmatrix}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\pmb{A}_{s_2}^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{s_2}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}_{s_2}^{-1}
\end{equation}
Interestingly this is precisely the \textbf{robust estimate of the model-dependent covariance matrix} proposed by \cite{huber1} and also discussed by \cite{gregoire2} in a forestry context (see also \cite{mandallaz}, pp. 123-125 for a proof).\\
With two-stage sampling one replaces the true local density $Y(x)$ by its estimate $Y^*(x)$ and considers
\begin{equation}\label{beta2star}
\hat{\pmb{\beta}}^*_{s_2}=\pmb{A}_{s_2}^{-1}\pmb{U}^*_{s_2}
\end{equation}
where we have set $\pmb{U}^*_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}Y^*(x)\pmb{Z}(x)$. One has $\EX_{3\mid 2}(\hat{\pmb{\beta}}^*_{s_2})=\hat{\pmb{\beta}}_{s_2}$.\\
 \noindent
We can now derive the design-based variance covariance matrix of $\hat{\pmb{\beta}}^*_{s_2}$. One has
\begin{eqnarray*}
\pmb{\Sigma}_{\hat{\pmb{\beta}}^*_{s_2}} &=& \EX_{2,3}\big((\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})^t\big)
+ \EX_{2,3}(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}_0)(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}_0)^t\\
&+& \EX_{2,3}\big(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}_0)
+ \EX_{2,3}\big(\hat{\pmb{\beta}}_{s_2}-\pmb{\beta}_0)(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2}) \\
&=&\EX_2\VAR_{3\mid 2}(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})+ \pmb{\Sigma}_{\pmb{\beta}_{s_2}} + 0 +0 \\
&=& \pmb{\Sigma}_{\pmb{\beta}_{s_2}} + \EX_2 \pmb{A}_2^{-1}(\frac{1}{n_2^2}\sum_{x\in{s_2}}V(x)\pmb{Z}(x)\pmb{Z}^t(x))\pmb{A}_2^{-1}\\
&\approx & \pmb{\Sigma}_{\pmb{\beta}_{s_2}} + \pmb{A}^{-1}(\frac{1}{n_2}\EX_xV(x)\pmb{Z}(x)\pmb{Z}^t(x))\pmb{A}^{-1}\\
&=&\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX_x\big((V(x)+R^2(x))\big)\pmb{Z}(x)\pmb{Z}(x)\big)^t\Big)\pmb{A}^{-1}
\end{eqnarray*}
where we have use the independence of Poisson sampling at different points, equations [\ref{varmatrix}] and the identity
$$\hat{\pmb{\beta}}^*_{s_2}-\pmb{\beta}_0=\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2}
+\hat{\pmb{\beta}}_{s_2}- \pmb{\beta}_0$$
Hence, we have
\begin{equation}\label{twostagedbcovmatrix}
\pmb{\Sigma}_{\hat{\pmb{\beta}}^*_{s_2}}=
\pmb{A}^{-1}\Big(\frac{1}{n_2}\EX_x\big((V(x)+R^2(x))\big)\pmb{Z}(x)\pmb{Z}(x)\big)^t\Big)\pmb{A}^{-1}
\end{equation}
The main point is that the variance of $\hat{\pmb{\beta}}^*_{s_2}$ decrease as $n_2^{-1}$, like the variance of $\hat{\pmb{\beta}}_{s_2}$, so that $\hat{\pmb{\beta}}^*_{s_2}$  is also an asymptotically unbiased and consistent estimator of $\pmb{\beta}_0$. The true prediction $\hat{Y}_0=\pmb{Z}(x)^t\pmb{\beta}_0$ is of course not available and we have the following possibilities
\begin{eqnarray}\label{predictions}
\hat{Y}_{\pmb{\beta}}(x)&=&\pmb{Z}^t(x)\pmb{\beta} \\ \nonumber
\hat{Y}_{s_2}(x)&=&\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}=\hat{Y}_0(x) +
O_p\big(\frac{1}{\sqrt{n_2}}\big) \\ \nonumber
\hat{Y}_{s_2}^*(x)&=&\pmb{Z}^t(x)\hat{\pmb{\beta}}^*_{s_2}=\hat{Y}_0(x) +
O_p\big(\frac{1}{\sqrt{n_2}}\big)
\end{eqnarray}
\noindent
where we have set $\hat{Y}_0(x)=\pmb{Z}^t(x)\pmb{\beta}_0$.\\
Therefore, we can treat all the above predictions asymptotically as if they were depending on $\pmb{Z}(x)$ only and not on the sample, which is precisely the definition of the external variance assumption.\\
Setting now $\hat{R}_{s_2}^*(x)=Y^*(x)-\hat{Y}_{s_2}^*(x)\approx Y^*(x)-\hat{Y}_0(x)$ we get first $\EX_{3 \mid x}\hat{R}_{s_2}^*(x)=\hat{R}_{s_2}(x)$, $\VAR_{3 \mid 2}\hat{R}_{s_2}^*(x)=\VAR_{3 \mid 2}Y^*(x)=V(x)$ and finally $\EX_{3 \mid x}(R_{s_2}^*(x))^2=\hat{R}_{s_2}^2(x)+V(x)$. From this we have the following consistent estimator of the design-based variance covariance matrix of $\hat{\pmb{\beta}}^*_{s_2}$
\begin{equation}\label{dbvariancetwostage}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}^*_{s_2}} =\pmb{A}_{s_2}^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}(\hat{R}_{s_2}^*(x))^2\pmb{Z}(x)\pmb{Z}(x)^t\Big)\pmb{A}_{s_2}^{-1}
\end{equation}

\subsection{The two-phase one-stage regression estimators}\label{twophaseonestagereg}

If the prediction model is \textbf{external}, i.e. not fitted with the inventory data at hand, the regression estimate is defined as
\begin{equation}\label{external1}
\hat{Y}_{reg}=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_{\pmb{\beta}}(x)+ \frac{1}{n_2}\sum_{x\in{s_2}}R_{\pmb{\beta}}(x)
\end{equation}
with the predictions $\hat{Y}_{\pmb{\beta}}(x)=\pmb{Z}^t(x)\pmb{\beta}$ and the residuals $R_{\pmb{\beta}}(x)=Y(x)-\hat{Y}_{\pmb{\beta}}(x)$, where $\pmb{\beta}$ is the given external regression coefficient, ideally obtained from another similar inventory. Note that in this case the mean residual will not necessarily be zero.
To calculate the variance one uses the decomposition
\begin{equation}\label{external2}
\VAR_{1,2}(\hat{Y}_{reg})=\VAR_1\EX_{2\mid 1}(\hat{Y}_{reg})+\EX_1\VAR_{2 \mid 1}(\hat{Y}_{reg})
\end{equation} to obtain
\begin{equation}\label{external3}
\VAR(\hat{Y}_{reg})=\frac{1}{n_1}\VAR(Y(x))+(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R_{\pmb{\beta}}(x))
 \end{equation}
 which can be unbiasedly estimated with
\begin{equation}\label{varexternalsimple}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R_{\pmb{\beta}}(x)-\bar{R}_{\pmb{\beta},2})^2
\end{equation}
where $\bar{Y}_2=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ and  $\bar{R}_{\pmb{\beta},2}=\frac{1}{n_2}\sum_{x\in{s_2}}R_{\pmb{\beta}}(x)$.\\
If the model is \textbf{internal} the regression coefficients are estimated with the inventory data at hand to obtain first $\hat{\pmb{\beta}}_{s_2}$, second the predictions $\hat{Y}_{s_2}(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$
and residuals $\hat{R}_{s_2}(x)=Y(x)-\hat{Y}_{s_2}(x)$ and finally the two-phase one-stage regression estimator

\begin{equation}\label{internal1}
\hat{Y}_{reg}=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_{s_2}(x)+ \frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{s_2}(x)
\end{equation}
It can be shown (\cite{mandallaz3} and \cite{mandallazreport1} for details) that asymptotically ($n_2 \to \infty$) that one can treat an internal model as an external one, so that we have the consistent variance estimate
\begin{equation}\label{estvarinternal1}
\hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_{s_2})^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}_{s_2}(x)-\bar{\hat{R}}_{s_2})^2
\end{equation}
where we have set $\bar{Y}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)$ and $\bar{\hat{R}}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{s_2}(x)$.
Because predictions and residuals are orthogonal by construction we also have
\begin{equation}\label{estvarinternal3}
 \hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{Y}_{s_2}(x)-\bar{\hat{Y}}_{s_2})^2+
\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}_{s_2}(x)-\bar{\hat{R}}_{s_2})^2
\end{equation}
where =$\bar{\hat{Y}}_{s_2}=\frac{1}{n_2}\sum_{x\in{s_2}}\hat{Y}_{s_2}(x)$. The first term contains the empirical variance of the predictions in the small sample $s_2$, which could be replace by the empirical variance in the large sample $s_1$. Furthermore, by construction the mean of the residuals in the second term is zero by construction so that one can also use
\begin{equation}\label{estvarinternal4}
 \hat{\VAR}(\hat{Y}_{reg})=\frac{1}{n_1}\frac{1}{n_1-1}\sum_{x\in{s_1}}(\hat{Y}_{s_2}(x)-\bar{\hat{Y}}_{s_1})^2+
\frac{1}{n^2_2}\sum_{x\in{s_2}}\hat{R}^2_{s_2}(x)
\end{equation}
where $\bar{\hat{Y}}_{s_1}=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_{s_2}(x)$. This version is more in line with the variance obtained via the design-based covariance matrix derived in (\cite{mandallazreport1} eqn. [28]) and which reads
\begin{equation}\label{varyreggweight}
\VAR(\hat{Y}_{reg})=
\bar{\pmb{Z}}^t\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}
+ \pmb{\beta}^t\Sigma_{\hat{\bar{\pmb{Z}}}_1}\pmb{\beta}
\end{equation}
which yields the asymptotically consistent estimate of the variance
\begin{equation}\label{estvaryreggweight}
\hat{\VAR}(\hat{Y}_{reg}) =
\hat{\bar{\pmb{Z}}}_{1}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_{1}
+ \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}\hat{\pmb{\beta}}_{s_2}
\end{equation}
with the estimated variance-covariance matrix of the $\pmb{Z}(x)$ given by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
\end{equation}
with $\hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s_1}}\pmb{Z}(x)$ .\\
The \textbf{g-weights} are defined as
\begin{equation}\label{gweight1}
g_{2}(x)=\hat{\bar{\pmb{Z}}}^t_{1}\pmb{A}^{-1}_{s_2}\pmb{Z}(x)
\end{equation}
They satisfy the \textbf{calibration properties}
\begin{equation}\label{gweight2}
\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)\pmb{Z}(x)=\hat{\bar{\pmb{Z}}}_{1}
\end{equation}
\noindent In particular $\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)=1$. The regression estimator can be rewritten, because of the zero mean residuals (i.e. $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{s_2}(x)=0$) as the weighted mean
\begin{equation}\label{gweight3}
\hat{Y}_{reg}=\frac{1}{n_2}\sum_{x\in{s_2}}g_{2}(x)Y(x)
\end{equation}
and also as
\begin{equation}\label{gweight3bis}
\hat{Y}_{reg}=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{2}(x)Y(x)
\end{equation}
with the g-weights
\begin{equation}\label{gweight3bis}
\tilde{g}_{2}(x)=1+(\hat{\bar{\pmb{Z}}}_{1}-\hat{\bar{\pmb{Z}}}_{2})^t\pmb{A}^{-1}_{s_2}\pmb{Z}(x)
\end{equation}
Hence, for all $Y(x)$, we have
$$\frac{1}{n_2}\sum_{x\in{s_2}}g_{2}(x)Y(x)=\frac{1}{n_2}\sum_{x\in{s_2}}\tilde{g}_{2}(x)Y(x)$$
Since the g-weights depend only on $\pmb{Z}(x)$ and not $Y(x)$ we must have
\begin{equation}\label{equalityofgweights}
g_2(x)=\tilde{g}_{2}(x)
\end{equation}
Because $\hat{\bar{\pmb{Z}}}_{1}-\hat{\bar{\pmb{Z}}}_{2}$ tends to zero almost surely and in quadratic mean, we see that $\tilde{g}_{2}(x)=g_2(x)=1+O_p(\frac{1}{\sqrt{n_2}})$. We have thus proved the \textbf{fundamental property that the g-weights are asymptotically equal to $1$}.\\
Furthermore the estimated variance can be expressed as
\begin{eqnarray}\label{gweight4}
\hat{\VAR}(\hat{Y}_{reg})&=&\frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{2}(x)\hat{R}_{s_2}^2(x) + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}\hat{\pmb{\beta}}_{s_2}  \nonumber \\
&=& \frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{2}(x)\hat{R}_{s_2}^2(x)+ \frac{1}{n_{1}(n_{1}-1)}
\sum_{x\in{s_{1}}}(\hat{Y}_{s_2}(x)-\bar{\hat{Y}}_{1})^2
\end{eqnarray}
with $\bar{\hat{Y}}_{1}=\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_{s_2}(x)$
 The second term in the last equation is the variance of the predictions over $F$. Since the g-weights tend to $1$ asymptotically  [\ref{estvarinternal4}] and [\ref{gweight4}] are asymptotically equivalent. \\
 \noindent \textbf{Remarks:}
 \begin{itemize}
 \item
 The notation $g_2(x)$ for g-weights can be misleading in the sense that they do not depent solely on $x$, and consequently $\pmb{Z}(x)$ but also on the sample $s_1$ (because of $\pmb{A}_2$ and $s_1$ (because of $\hat{\bar{\pmb{Z}}}_{1}$. Hence, a better notation would be $g_{s_1,s_2}(x)$, which is cumbersome to use in complicated formulae.
 \item
 If the first phase is exhaustive, i.e. $n_1=\infty$, we have $g_{2}(x)=\bar{\pmb{Z}}^t\pmb{A}^{-1}_{s_2}\pmb{Z}(x)$. With $Y(x)=\pmb{Z}^t(x)\pmb{\beta}_0+R_0(x)$ and $\bar{Y}_F=\frac{1}{\lambda(F)}\int_F \hat{Y}_0(x)dx=\bar{\pmb{Z}}^t\pmb{\beta}_0$ we get, using
 [\ref{gweight2}]
 \begin{eqnarray*}
 \hat{Y}_{reg}-\bar{Y}_F &=&\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)Y(x)-\bar{\pmb{Z}}^t\pmb{\beta}_0  \\
 &=&\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)(\pmb{Z}^t(x)\pmb{\beta}_0+R_0(x))-\bar{\pmb{Z}}^t(x)\pmb{\beta}_0\\
 &=& \bar{\pmb{Z}}^t\pmb{\beta}_0+\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)R_0(x)-\bar{\pmb{Z}}^t\pmb{\beta}_0\\
 &=& \sum_{x\in{s_2}}g_2(x)R_0(x)
 \end{eqnarray*}
 We also have $\EX_x g_2(x)R_0(x)\approx \bar{\pmb{Z}}^t\pmb{A}^{-1}\EX_x \pmb{Z}(x)R_0(x)=0$
 because of the orthogonality relationship. Comparing with [\ref{gweight4}] we see that for calculating the variance we can consider the $g_2(x)R_0(x)$ and then the $g_{2}(x)\hat{R}(x)$ as i.i.d with zero mean.
 \end{itemize}
 \subsection{The two-phase two-stage regression estimator}\label{twpphasetwostagereg}
 We will now show that treating an internal model as if it were external is also asymptotically correct in the two-phase two-stage procedures. This was up to certain point taken for granted in \cite{mandallaz},section 5.2, and subsequent work, where the predictions were considered to depend only on $\pmb{Z}(x)$, which is only asymptotically correct.\\
 The predictions are now calculated according to $\hat{Y}_{s_2}^*(x)=\pmb{Z}^t(x)\hat{\pmb{\beta}}^*_{s_2}$
 and the two-phase two-stage regression estimator reads now
 \begin{eqnarray}\label{twophasetwostagereg}
 \hat{Y}^*_{reg}&=&\frac{1}{n_1}\sum_{x\in{s_1}}\hat{Y}_{s_2}^*(x)+\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{s_2}^*(x) \nonumber \\
 &=&\hat{\bar{\pmb{Z}}}_1\hat{\pmb{\beta}}^*_{s_2}
 \end{eqnarray}
 with $\hat{R}_{s_2}^*(x)=Y_{s_2}^*(x)-\hat{Y}_{s_2}^*(x)$. Obviously one has
 $\EX_{3 \mid 2,1}\hat{Y}^*_{reg}=\hat{Y}_{reg}$. Using the variance decomposition
 \begin{equation}\label{vardecomposition1}
 \VAR_{1,2,3}(\hat{Y}^*_{reg})=\VAR_{1,2}\EX_{3 \mid 2,1}(\hat{Y}^*_{reg})+\EX_{1,2}\VAR_{3 \mid 2,1}(\hat{Y}^*_{reg})
 \end{equation}
 we get
 \begin{equation}\label{vardecomposition2}
 \VAR_{1,2,3}(\hat{Y}^*_{reg})=\VAR_{1,2}(\hat{Y}_{reg})+\EX_{1,2}\VAR_{3 \mid 2,1}(\hat{\bar{\pmb{Z}}}_1\hat{\pmb{\beta}}^*_{s_2})
 \end{equation}
 The first term is given by [\ref{varyreggweight}] and for the third we need the conditional variance-covariance matrix $\Sigma_{\hat{\pmb{\beta}}^*_{s_2}\mid{2,1}}$, which is defined as
 \begin{equation*}
 \Sigma_{\hat{\pmb{\beta}}^*_{s_2}\mid{2,1}}=\EX_{3 \mid 2,1}(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})^t
 \end{equation*}
 Because Poisson sampling is performed independently at points $x\in{s_2}$ we have with [\ref{beta2star}] and [\ref{poisson_atx}] $\VAR_{3\mid 2,1}\sum_{x\in{s_2}}Y^*(x)=\sum_{x\in{s_2}}V(x)$
 \begin{equation*}
 \VAR_{3\mid 2,1}\Big(\sum_{x\in{s_2}}Y^*(x)\Big) =\sum_{x\in{s_2}}V(x)
 \end{equation*}
 This gives after some algebra
 \begin{equation}\label{conditioncovariancematrix1}
 \Sigma_{\hat{\pmb{\beta}}^*_{s_2}\mid 2,1}=\pmb{A}^{-1}_{s_2}\Big(
 \frac{1}{n^2_2}\sum_{x\in{s_2}}V(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}^{-1}_{s_2}
 \end{equation}
 Using [\ref{vardecomposition2}], [\ref{varyreggweight}], [\ref{varmatrix}] and collecting the pieces we obtain obtain the asymptotic variance

 \begin{eqnarray}\label{vartwophasetwostageyreg1}
 \VAR_{1,2,3}(\hat{Y}^*_{reg})&=&\frac{1}{n_1}\VAR(\hat{Y}_0(x))+
  \bar{\pmb{Z}}^t\pmb{A}^{-1}\Big(
 \frac{1}{n_2}\EX_x( R_0^2(x)+V(x))\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}^{-1}\bar{\pmb{Z}} \nonumber \\
 &=&\frac{1}{n_1}\VAR(\hat{Y}_0(x))+\bar{\pmb{Z}}^t\pmb{\Sigma}_{\pmb{\beta}^*_{s_2}}\bar{\pmb{Z}}
  \end{eqnarray}
 and consequently the following consistent estimate of the variance
 \begin{eqnarray}\label{esttwophasetwostageasymptvar}
 \hat{\VAR}(\hat{Y}^*_{reg})&=&\hat{\pmb{\beta}^*}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}\hat{\pmb{\beta}^*}_{s_2}
 +\hat{\bar{Z}}_1^t\pmb{A}_{s_2}^{-1}\big(\frac{1}{n^2_2}\sum_{x\in{s_2}}(\hat{R}^*(x))^2\pmb{Z}(x)\pmb{Z}^t(x)\big) \pmb{A}_{s_2}^{-1}\hat{\bar{\pmb{Z}}}_1 \nonumber \\
 &=&\frac{1}{n_1}\frac{1}{n_1-1}\sum_{x\in{s_2}}(\hat{Y}^*(x)-\bar{\hat{Y}}^*_1)^2 +
 \frac{1}{n^2_2}\sum_{x\in{s_2}}g^2_{2}(x)(\hat{R}^*(x))^2
 \end{eqnarray}
 which is algebraically the same as [\ref{gweight2}] except for the $*$. Since the $g^2_{2}(x)$ tend to $1$ we have proved that the external variance estimate is also equivalent to the consistent g-weight variance in two-phase two-stage sampling.\\
 In the model-dependent geostatistical set-up the situation is more complicated, even in simple one-phase inventory, and one has to use Kriging with measurement error instead (see section 7.3 in \cite{mandallaz}). \\
 For timber volume assessment in the the Swiss National Inventory the second stage variance is remarkably small (around $3\%$ in comparison to the overall variance ( i.e.
 $\frac{\EX V(x)}{\VAR(Y^*(x))}\approx 0.03$. However,the slight increase in variance is more than counterbalanced by drastic savings in measurement costs (\cite{mandallazmassey1} and \cite{massey} for details).\\
 \noindent
 \textbf{Remarks:}\label{remarkstwophase}
 \begin{enumerate}
 \item
 One-phase two-stage sampling is the special case with $\pmb{Z}(x)$ consisting only of the intercept, which leads to the results given in (\cite{mandallaz}, pp.70-71)
 \begin{eqnarray*}
 \hat{Y}^*&=&\frac{1}{n_2}\sum_{x\in{s_2}}Y^*(x) \\
 \VAR(\hat{Y}^*)&=&\frac{1}{n_2}\VAR_x(Y(x))+\frac{1}{n_2}\EX_x V(x) \\
 \hat{\VAR}(\hat{Y}^*)&=&\frac{1}{n_2(n_2-1)}\sum_{x\in{s_2}}(Y^*(x)-\hat{Y}^*)^2 \\
 \EX_{2,3}(\hat{\VAR}(\hat{Y}^*)&=& \VAR(\hat{Y}^*)
 \end{eqnarray*}
 These results are exact and not only asymptotically correct in contrast to the general case.
 \item
 From a theoretical point of view the g-weights variance estimates have better statistical properties. The simple post-stratification example is particularly instructive in this respect: the external variance estimate does not use the strata weights estimated from the large sample and the variance is not inversely proportional to the number of $s_2$-points in the strata, in contrast to the g-weight variance estimate (see \cite{mandallaz}, pp. 84-86). Also, it is closer, under regularity conditions, to the model-dependent variance if the model is correct (see \cite{mandallaz}, pp. 114-116).
 \item
 In practice one uses primarily systematic grids, for which no design-based variance estimate exists. In this case the only mathematically absolutely correct approach is to use the geostatistical Kriging procedures: treating systematic grids as random samples will usually overestimate the variance, particularly in one-phase sampling and for small-area estimation. The situation is better in two-phase sampling because the variance of the residuals is generally the dominating term,  with a much shorter range of the spatial correlation than the range of the observations $Y(x)$. For global estimation the two-phase regression estimator and its geostatistical counterpart, the Double-Kriging estimator, seem to give very similar results (both for point and variance estimates). With two-stage sampling one has to use Kriging with measurement errors (see \cite{mandallaz}, chapter 7, for an introduction to geostatistical methods in forest inventory and \cite{dmhabil} for a detailed mathematical treatment). It is fair to say that geostatistical techniques are far more difficult to implement than the design-based procedures and, in my opinion, they are not suitable for routine work in forest inventory. They are undoubtedly very successful in oil and mining resource assessment, where the financial stakes are completely different. However, the increasing demand for small-area estimation and the availability of free geostatistical software packages within the R-project (20 years ago the software packages, essentially used in the mining and oil industry, were extremely costly) with good interfaces to GIS software will certainly promote geostatistical techniques in forest inventories.
 \end{enumerate}

 \clearpage \newpage
\section{Three-phase sampling}\label{threephase}
\subsection{Background}\label{background}
The \textbf{null phase} draws a very large sample $s_0$ of $n_0$ points
$x_{i}\in{s_0}$ ($i=1,2\ldots n_0$) that are independently and uniformly distributed
within the forest area $F$. At each of those points auxiliary
information is collected, very often coding information of qualitative nature
(e.g. following the  interpretation of aerial photographs) or quantitative (e.g. timber volume estimates  based on LiDAR measurements). We shall assume that the auxiliary information at point $x$
is described by the column vector $\pmb{Z}^{(1)}(x)\in{\Re^p}$. The case $n_0=\infty$, i.e. $\pmb{Z}^{(1)}(x)$ is  \textbf{exhaustive}, has been investigated in \cite{mandallaz4}. The \textbf{first phase} draws a large sample $s_1 \subset s_0$ of $n_1 <<n_0$ points by simple random sampling in $s_0$. Note that the points $x\in{s_1}$ are also uniformly independently distributed in $F$. For each point in the first phase a further component $\pmb{Z}^{(2)}(x)\in{\Re^q}$ of the auxiliary information is available and hence also the vector $\pmb{Z}^t(x)=(\pmb{Z}^{(1)t}(x),\pmb{Z}^{(2)t}(x))\in{\Re^{p+q}}$ (the upper index $t$ denotes the transposition operator). The \textbf{second phase} draws a small sample $s_2\subset{s_1}$ of
$n_2$ points from $s_1$ by simple random sampling and consists of the terrestrial inventory. Note that we have used the terms null, first and second phases instead of first, second and third phases simply to ensure compatibility with the terminology used in previous work, in which the null-phase was exhaustive.\\
To set the stage the component $\pmb{Z}^{(1)}(x)\in{\Re^p}$ can be based on the interpretation of aerial photographs or on simple characteristics of the canopy height obtained from LiDAR data (such as mean canopy height and eventually quantiles thereof), whereas $\pmb{Z}^{(2)}(x)\in{\Re^p}$ is based on other computationally intensive characteristics of the canopy requiring individual tree detection (e.g. tree species ortree volume prediction based on tree height).
The reason for introducing the null-phase sample $s_0$ is that the component $\pmb{Z}^{(1)}(x)$ can be computationally  prohibitive to calculate exhaustively in extensive forest inventories (see \cite{mandallaz3} for a case study with LiDAR data).
\subsection{The models}\label{models3phase}
 We shall work with the following linear models
 \begin{enumerate}
 \item
 The large model $M$
 \begin{equation}\label{largemodel}
 Y(x)=\pmb{Z}^t(x)\pmb{\beta}+ R_{\pmb{\beta}}(x)=\pmb{Z}^{(1)t}(x)\pmb{\beta}^{(1)}
 +\pmb{Z}^{(2)t}(x)\pmb{\beta}^{(2)}+ R_{\pmb{\beta}}(x)=:\hat{Y}_{\pmb{\beta}}(x)+R_{\pmb{\beta}}(x)
 \end{equation}
 \noindent with $\pmb{\beta}^t=({\pmb{\beta}^{(1)t}}, {\pmb{\beta}^{(2)t}})$ and the predictions
 $\hat{Y}_{\pmb{\beta}}(x)=\pmb{Z}^t(x)\pmb{\beta}$ and residuals $R_{\pmb{\beta}}=Y(x)-\hat{Y}_{\pmb{\beta}}(x)$.\\
 The intercept term is contained in $\pmb{Z}^{(1)}(x)$ or a linear combination of the components of $\pmb{Z}^{(1)}(x)$ is constant equal to $1$.\\
 The theoretical regression parameter $\pmb{\beta}_0$ minimizes
 $\int_F (Y(x)-\pmb{Z}^t(x)\pmb{\beta})^2dx$. It satisfies the normal equation
 $\big(\int_F\pmb{Z}(x)\pmb{Z}^t(x)dx\big)\pmb{\beta}_0=\int_F Y(x)\pmb{Z}(x)dx$ and the orthogonality relationship
 $\int_F R_{\pmb{\beta}_0}(x)\pmb{Z}(x)dx=\pmb{0}$, in particular the zero mean residual property
 $\frac{1}{\lambda(F)}\int_F R_{\pmb{\beta}_0}(x)dx=0$.
 \item
 The reduced model $M_1$
 \begin{equation}\label{reducedmodel}
 Y(x)=\pmb{Z}^{(1)t}(x)\pmb{\alpha} + R_{1,\pmb{\alpha}}(x)=:\hat{Y}_{1,\pmb{\alpha}}(x)+ R_{1,\pmb{\alpha}}(x)
 \end{equation}
 The theoretical regression parameter $\pmb{\alpha}_0$ minimizes
 $\int_F (Y(x)-\pmb{Z}^{(1)t}(x)\pmb{\alpha})^2dx$. It satisfies the normal equation
 $\big(\int_F \pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)dx\big)\pmb{\alpha}_0=\int_FY(x)\pmb{Z}^{(1)}(x)dx$ and the orthogonality relationship
 $\int_F R_{1,\pmb{\alpha}_0}(x)\pmb{Z}^{(1)}(x)dx=\pmb{0}$, in particular the zero mean residual property $ \frac{1}{\lambda(F)}\int_F R_{1,\pmb{\alpha}_0}(x)dx=0$.
\end{enumerate}
\subsection{The three-phase one-stage generalized regression estimator}\label{threephaseregest}
We consider the following design-based least squares estimators of the regression coefficients of the reduced model, which are solutions of sample copies of the normal equations
\begin{eqnarray}\label{coeff1}
\hat{\pmb{\alpha}}_{s_k} &=& \Big(\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)
\Big)^{-1}\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)\nonumber\\
&:=&(\pmb{A}^{(1)}_k)^{-1}
\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}^{(1)}(x)=:(\pmb{A}^{(1)}_k)^{-1}\pmb{U}^{(1)}_k,\quad  k=0,1,2
\end{eqnarray}
Likewise for the large large model we set
\begin{eqnarray}\label{coeff2}
\hat{\pmb{\beta}}_{s_k}&=&\Big(\frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}(x)\pmb{Z}^t(x)
\Big)^{-1}\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}(x)\nonumber \\
&=&\pmb{A}^{-1}_k\frac{1}{n_k}\sum_{x\in{s}_k}Y(x)\pmb{Z}(x)=:\pmb{A}^{-1}_k\pmb{U}_k,\quad k=0,1,2
\end{eqnarray}
 Note that only $\hat{\pmb{\alpha}}_{s_2}$ and $\hat{\pmb{\beta}}_{s_2}$ are observable, because $Y(x)$ is only available at $x\in{s_2}$, and that in general the vector consisting of the first $p$ components of $\hat{\pmb{\beta}}_{s_2}$ is not equal to $\hat{\pmb{\alpha}}_{s_2}$.\\
The empirical predictions and residuals are defined, in a slightly simplified notation, as
\begin{eqnarray*}
\hat{Y}(x)&=&\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2} \\ \hat{Y}_1(x)&=&\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}_{s_2}\\
\hat{R}(x)&=&Y(x)-\hat{Y}(x)\\
\hat{R}_1(x)&=&Y(x)-\hat{Y}_1(x)
\end{eqnarray*}
Consistent estimates of the design-based covariance matrices are given by
\begin{eqnarray}\label{robustvar}
\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}&=&\pmb{A}_2^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)\pmb{A}_2^{-1}\nonumber \\
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=&(\pmb{A}^{(1)}_1)^{-1}
\Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_1^2(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)t}(x)\Big)(\pmb{A}^{(1)}_1)^{-1}
\end{eqnarray}
Note that the  matrix $\pmb{A}^{(1)}_1$ is used for $\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}$ instead of $\pmb{A}^{(1)}_2$, both being asymptotically equivalent. This is necessary in order to have compatibility with the g-weight version of the estimated variance of the regression estimators discussed below. We need the following notation for the various means of the auxiliary vectors
\begin{eqnarray}\label{meanvalues}
\bar{\pmb{Z}}^{(1)} &=& \frac{1}{\lambda(F)}\int_F \pmb{Z}^{(1)}(x)dx \nonumber \\
\hat{\bar{\pmb{Z}}}^{(1)}_0 &=& \frac{1}{n_0}\sum_{x\in{s_0}}\pmb{Z}^{(1)}(x) \nonumber \\
\hat{\bar{\pmb{Z}}}^{(1)}_1 &=& \frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(1)}(x)  \nonumber \\
\hat{\bar{\pmb{Z}}}_k &=& \frac{1}{n_k}\sum_{x\in{s}_k}\pmb{Z}(x),\;k=1,2
\end{eqnarray}
The g-weights are defined as
\begin{eqnarray}\label{gweight5}
g^{(1)}_1(x)&=&\hat{\bar{\pmb{Z}}}_0^{t(1)}(\pmb{A}^{(1)}_1)^{-1}\pmb{Z}^{(1)}(x)\nonumber \\
g_2(x) &=& \hat{\bar{\pmb{Z}}}^t_{1}\pmb{A}^{-1}_{2}\pmb{Z}(x)
\end{eqnarray}
Note that $\pmb{A}^{(1)}_1$ instead of $\pmb{A}^{(1)}_2$ is used in the definition of $g^{(1)}_1(x)$.
 The g-weights satisfy the calibration properties
\begin{eqnarray}\label{gweight6}
\frac{1}{n_1}\sum_{x\in{s_1}}g^{(1)}_1(x)\pmb{Z}^{(1)}(x)&=&\hat{\bar{\pmb{Z}}}^{(1)}_0 \nonumber \\
\frac{1}{n_2}\sum_{x\in{s_2}}g_2(x)\pmb{Z}(x)&=&\hat{\bar{\pmb{Z}}}_{1}
\end{eqnarray}
The mean of the g-weights over $s_1$ and $s_2$ are equal to $1$. We have already proved that the $g_2(x)$ tend to $1$. With an obvious modification of the proof it can be shown that this also holds for $g^{(1)}_1(x)$.\\
The \textbf{generalized regression estimate} discussed by \cite{mandallaz4} is defined as
\begin{eqnarray}\label{ygreg1}
\hat{Y}_{greg} &=& \frac{1}{\lambda(F)}\int_F \hat{Y}_1(x)dx
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}(x)-\hat{Y}_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x)) \nonumber \\
&=& (\bar{\pmb{Z}}^{(1)}-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2}
\end{eqnarray}
Here, $\pmb{Z}^{(1)}(x)$ is exhaustive, i.e. known at all $x\in{F}$, though in practice it suffices to know the true mean $\bar{\pmb{Z}}^{(1)}$. \\
Using the results given in \cite{mandallaz4} (see \cite{mandallazreport3} for the proofs) one has under the \textbf{external model assumption} the variance estimate
\begin{equation}\label{estexternalvarYgreg}
\hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{1}^2(x))
+ (1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s_2}}\hat{R}^2(x)
\end{equation}
Using the estimated design-based variance-covariance matrices one obtains alternatively the consistent estimate
\begin{equation}\label{estvarygreggweight1}
\hat{\VAR}(\hat{Y}_{greg})=\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(1)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_1
\end{equation}
which can be rewritten in the g-weight form
\begin{equation}\label{estvarygreggweight2}
\hat{\VAR}(\hat{Y}_{greg})=\frac{1}{n_1n_2}\sum_{x\in{s_2}}(g_1^{(1)}(x))^2\hat{R}_{1}^2(x))
+ (1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s_2}}g_2^2(x)\hat{R}^2(x)
\end{equation}
Because the g-weights tend to $1$ the variance estimates [\ref{estexternalvarYgreg}] and [\ref{estvarygreggweight2}] are asymptotically equivalent.\\
If $\pmb{Z}^{(1)}(x)$ is no longer exhaustive we replace the first term in [\ref{ygreg1}] by its sample mean in the null phase sample and we define the new three-phase estimator as
\begin{eqnarray}\label{yg3reg}
\hat{Y}_{g3reg} &=& \frac{1}{n_0} \sum_{x\in{s_0}} \hat{Y}_1(x)
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}(x)-\hat{Y}_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y(x)-\hat{Y}(x)) \nonumber \\
&=& (\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2}
\end{eqnarray}
where $\hat{\bar{\pmb{Z}}}^{(1)}_0=\frac{1}{n_0}\sum_{x\in{s_0}}\pmb{Z}^{(1)}(x)$.\\
Note that the third terms in [\ref{ygreg1}] and [\ref{yg3reg}] vanish for internal models because by assumption the intercept term is contained in $\pmb{Z}^{(1)}$ (or it can be expressed as a linear combination of its components) so that the residuals sum up to zero over $s_2$.\\
By using the properties of conditional expectations and variances  we see that
$\hat{Y}_{F,g3reg}$ is asymptotically design unbiased and that under the external model assumption
one has
\begin{eqnarray}\label{extvaryg3reg1}
\VAR_{0,1,2}(\hat{Y}_{g3reg})&=&\EX_0 \EX_{1 \mid 0}\VAR_{2 \mid 0,1}(\hat{Y}_{F,g3reg})+
\EX_0 \VAR_{1 \mid 0}\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})
+\VAR_0 \EX_{1 \mid 0}\EX_{2 \mid 0,1}(\hat{Y}_{F,g3reg})\nonumber \\
&=& \frac{1}{n_0}\VAR_{x}(Y(x))+(1-\frac{n_1}{n_0})\frac{1}{n_1}\VAR_x(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR_x(R(x))
\end{eqnarray}
The above equation holds for any fixed given $\pmb{\alpha}$ and $\pmb{\beta}$. At the true  value $\pmb{\alpha}_0$ the predictions $\hat{Y}_1(x)$ and the residuals $R_1(x)$ are orthogonal in the design-based sense, which implies $\VAR_x(Y(x))=\VAR_x(\hat{Y}_1(x))+\VAR_x(R_1(x))$ and the following external variance
\begin{equation}\label{theoreticalexternalvaryg3reg}
\VAR_{0,1,2}(\hat{Y}_{g3reg})=\frac{1}{n_0}\VAR_{x}(\hat{Y}_1(x))+\frac{1}{n_1}\VAR_x(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x))
\end{equation}
As the empirical residuals based on $\hat{\pmb{\alpha}}_{s_2}$ and $\hat{\pmb{\beta}}_{s_2}$ have zero mean we get the external variance estimate
\begin{equation}\label{estimatedexternalvaryg3reg}
\hat{\VAR}_{ext}(\hat{Y}_{g3reg})=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1}
+ \frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}\hat{R}^2_1(x)+
(1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s}_2}\hat{R}^2(x)
\end{equation}
One can also derive the asymptotic theoretical variance via the design-based covariance matrices, which gives
\begin{equation}\label{theoreticalgweightvaryg3reg}
\VAR_{0,1,2}(\hat{Y}_{g3reg})= \frac{1}{n_0}\VAR(\hat{Y}_{1,\pmb{\alpha}}(x))
+\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}
\pmb{\Sigma}_{\hat{\pmb{\alpha}}_{s_2}}\bar{\pmb{Z}}^{(1)} + (1-\frac{n_2}{n_1})\bar{\pmb{Z}}^{t}\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}^{t}
\end{equation}
which can be consistently estimated by
\begin{equation}\label{estgweightvaryg3reg}
\hat{\VAR}_{0,1,2}(\hat{Y}_{g3reg})=\hat{\pmb{\alpha}}_{s_2}^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}_{s_2}
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(1)t}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}\hat{\bar{\pmb{Z}}}^{(1)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}^{t}_1
\end{equation}
with the design-based covariance matrix of the first component
\begin{equation}\label{estcovarz}
\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_{0}}=
\frac{1}{n_{0}}\frac{\sum_{x\in{s_{0}}}(\pmb{Z}^{(1)}(x)-
\hat{\bar{\pmb{Z}}}^{(1)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(1)}_{0})^t}{n_0-1}
\end{equation}
The proof of [\ref{theoreticalgweightvaryg3reg}] is rather intricate and given in \cite{mandallazreport3}.
One can rewrite [\ref{estgweightvaryg3reg}] as
\begin{eqnarray}\label{estgweightvaryg3regfinal}
\hat{\VAR}(\hat{Y}_{g3reg})&=&\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1}
\nonumber\\
&+&\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(g_1^{(1)}(x))^2\hat{R}^2_1(x)+
(1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s}_2}g_2^2(x)\hat{R}^2(x)
\end{eqnarray}
\noindent where $\hat{\bar{Y}}_1=\frac{1}{n_0}\sum_{x\in{s_0}}\hat{Y}_1(x)$. \\
Because the g-weights tend to $1$ [\ref{estgweightvaryg3regfinal}] is asymptotically equivalent to the external variance estimate given in [\ref{estimatedexternalvaryg3reg}].
\subsection{The three-phase two-stage generalized regression estimator}\label{threephasetwostageregest}
For the point estimate one simply replaces $Y(x)$ by $Y^*(x)$ in [\ref{coeff2}] to obtain $\hat{\pmb{\alpha}}_{s_2}^*$, $\hat{\pmb{\beta}}_{s_2}^*$. The predictions and residuals are defined as
\begin{eqnarray*}
\hat{Y}^*(x)&=&\pmb{Z}^t(x)\hat{\pmb{\beta}}^*_{s_2} \\ \hat{Y}^*_1(x)&=&\pmb{Z}^{(1)t}(x)\hat{\pmb{\alpha}}^*_{s_2}\\
\hat{R}^*(x)&=&Y^*(x)-\hat{Y}^*(x) \\
\hat{R}_1^*(x)&=&Y^*(x)-\hat{Y}_1^*(x)
\end{eqnarray*}
The tree-phase two-stage regression estimator reads
\begin{eqnarray}\label{yg3regtwostage}
\hat{Y}^*_{g3reg} &=& \frac{1}{n_0} \sum_{x\in{s_0}} \hat{Y}^*_1(x)
+\frac{1}{n_1}\sum_{x\in{s}_1}(\hat{Y}^*(x)-\hat{Y}^*_1(x))
+ \frac{1}{n_2}\sum_{x\in{s}_2}(Y^*(x)-\hat{Y}^*(x)) \nonumber \\
&=& (\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t\hat{\pmb{\alpha}}^*_{s_2} +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}^*_{s_2}
\end{eqnarray}
Since $\hat{\pmb{\alpha}}^*_{s_2}=\pmb{\alpha}_0 + O_p(\frac{1}{n_2})$ and $\hat{\pmb{\alpha}}^*_{s_2}=\pmb{\beta }_0 + O_p(\frac{1}{n_2})$ the external variance assumption is asymptotically valid. To calculate the overall variance we use the standard decomposition and [\ref{extvaryg3reg1}]
\begin{eqnarray}\label{extvary3regtwostage}
\VAR_{0,1,2,3}(\hat{Y}^*_{g3reg})&=&\VAR_{0,1,2}\EX_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})+\EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg}) \nonumber \\
&=&\VAR_{0,1,2}(\hat{Y}_{g3reg})+\frac{1}{n_2}\EX_x (V(x))\nonumber \\
&=&\frac{1}{n_0}\VAR_{x}(\hat{Y}_1(x))+\frac{1}{n_1}\VAR_x(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x))+ \frac{1}{n_2}\EX_x (V(x))
\end{eqnarray}
To obtain an estimate of the external variance we adapt [\ref{estimatedexternalvaryg3reg}]
\begin{eqnarray}\label{estexternvaryg3reg}
\hat{\VAR}_{ext}(\hat{Y}^*_{g3reg})&=&\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}^*_1(x)-\hat{\bar{Y}}^*_1)^2}{n_0-1}
\nonumber\\
&+&\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(\hat{R}^*_1(x))^2+
(1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s}_2}(\hat{R}^*(x))^2
\end{eqnarray}
As in two-phase sampling we use the asymptotically valid relations
\begin{eqnarray}\label{residualsquaredtwostage}
\EX_{3 \mid 0,1,2}(\hat{R}^*(x))^2 &=& R^2(x)+V(x) \nonumber \\
\EX_{3 \mid 0,1,2}(\hat{R}_1^*(x))^2&=&R^2_1(x)+V(x)
 \end{eqnarray}
Hence, we get for the expected value of the external variance estimate
\begin{eqnarray}\label{estexternvaryg3regproof}
\EX_{0,1,2,3}\hat{\VAR}_{ext}(\hat{Y}^*_{g3reg})&=&
\frac{1}{n_0}\VAR(\hat{Y}_1(x)) + \frac{1}{n_1}\VAR(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x)) \nonumber \\
&+ &\frac{1}{n_1}\EX_x V(x) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\EX_x V(x)\nonumber \\
&=& \frac{1}{n_0}\VAR(\hat{Y}_1(x)) + \frac{1}{n_1}\VAR(R_1(x))+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\VAR(R(x)) +\frac{1}{n_2}\EX_x V(x)\nonumber \\
&=& \VAR_{ext}(\hat{Y}_{g3reg})+\frac{1}{n_2}\EX_x V(x)
\end{eqnarray}
which is precisely [\ref{extvary3regtwostage}]. Hence, the external variance estimate [\label{estexternvaryg3reg}] is asymptotically unbiased.\\
We can also derive a g-weight version of the variance of $\hat{Y}^*_{g3reg}$ by using
\begin{eqnarray*}\label{theoreticgweightvarthreephasetwostage}
\VAR(\hat{Y}^*_{g3reg})&=&\VAR_{0,1,2}\EX_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})+ \EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg}) \nonumber \\
&=& \VAR_{0,1,2}(\hat{Y}_{g3reg})+\EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg}) \nonumber \\
&=& \frac{1}{n_0}\VAR(\hat{Y}_{1,\pmb{\alpha}_2 }(x))
+\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}
\pmb{\Sigma}_{\hat{\pmb{\alpha}}_{s_2}}\bar{\pmb{Z}}^{(1)} + (1-\frac{n_2}{n_1})\bar{\pmb{Z}}^{t}\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}^{t} + \EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})
\end{eqnarray*}
The increase in variance due to two-stage sampling is therefore $\EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})$. Recalling that the Poisson sampling schemes at different points are independent we get for the extra term
\begin{eqnarray*}\label{extraterm}
\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})&=&\EX_{3 \mid 0,1,2}\Big(
(\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t(\hat{\pmb{\alpha}}^*_{s_2}-\hat{\pmb{\alpha}}_{s_2})
+\hat{\bar{\pmb{Z}}}^{(1)}_1(\hat{\pmb{\beta}}^*_{s_2}-\hat{\pmb{\beta}}_{s_2})\Big)^2 \\
&=& (\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t(\pmb{A}_2^{(1)})^{-1}\big(\frac{1}{n^2_2}
\sum_{x\in{s_2}}V(x)\pmb{Z}^{(1)}(x)\pmb{Z}^{(1)}(x)\big)(\pmb{A}_2^{(1)})^{-1}
(\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1) \\
&+& 2(\hat{\bar{\pmb{Z}}}^{(1)}_0-\hat{\bar{\pmb{Z}}}^{(1)}_1)^t(\pmb{A}_2^{(1)})^{-1}\big(\frac{1}{n_2^2}
\sum_{x\in{s_2}}V(x)\pmb{Z}^{(1)}(x)\pmb{Z}(x)\big)\pmb{A}_2^{-1}\hat{\bar{\pmb{Z}}}_1 \\
&+ & \hat{\bar{\pmb{Z}}}_1\pmb{A}_2^{-1}\big(\frac{1}{n^2_2}\sum_{x\in{s_2}}V(x)\pmb{Z}(x)\pmb{Z}^t(x)\big)
\pmb{A}_2^{-1}\hat{\bar{\pmb{Z}}}_1
\end{eqnarray*}
Taking the expectation $\EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})$ we see that the first and second term are of smaller order than the third, so that we get asymptotically for the increase in variance
\begin{equation}\label{extratermasymptotic}
\EX_{0,1,2}\VAR_{3 \mid 0,1,2}(\hat{Y}^*_{g3reg})=
 \bar{\pmb{Z}}^t\pmb{A}^{-1}\big(\frac{1}{n_2}\EX_x (V(x)\pmb{Z}(x)\pmb{Z}^t(x))\big)\pmb{A}^{-1}\bar{\pmb{Z}}
 \end{equation}
 Writing its estimate with the g-weight $g(x)$ tending to 1 we see that the extra term is asymptotically equal to $\frac{1}{n_2}\EX_x V(x)$, which is the same as the value obtained with the external variance in [\ref{estexternvaryg3regproof}].\\
 Intuitively it is  appealing to use [\ref{theoreticalgweightvaryg3reg}] with the matrices $\pmb{\Sigma}_{\pmb{\alpha}^*_2}$ and $\pmb{\Sigma}_{\pmb{\beta}^*_2}$
 to get the variance estimate
 \begin{eqnarray}\label{estimatedthreephasetwostagevariance}
\hat{\VAR}_{0,1,2,3}(\hat{Y}^*_{g3reg})&=&\hat{\pmb{\alpha}}^{*t}_2\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}\hat{\pmb{\alpha}}^*_{s_2}
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(1)t}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}^*_{s_2}}\hat{\bar{\pmb{Z}}}^{(1)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}^*_{s_2}}\hat{\bar{\pmb{Z}}}^{t}_1 \nonumber \\
&=&\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}^*_1(x)-\hat{\bar{Y}}^*_1)^2}{n_0-1}
\nonumber\\
&+&\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s}_2}(g_1^{(1)}(x))^2(\hat{R}^*_1(x))^2+
(1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s}_2}g_2^2(x)(\hat{R}^*(x))^2
\end{eqnarray}

Using [\ref{residualsquaredtwostage}] we get
\begin{eqnarray}\label{threetwogweightvar1}
\EX_{1,2,3}\big(\hat{\VAR}(\hat{Y}^*_{g3reg})\big)&=& \frac{1}{n_0}\VAR(\hat{Y}_1(x)) \nonumber \\
&+&\EX_{1,2}\Big(\frac{1}{n_1n_2}\sum_{x\in{s}_2}(g_1^{(1)}(x))^2(R_1^2(x)+V(x))\Big) \nonumber \\
&+&\EX_{1,2}\Big(\frac{1}{n^2_2}\sum_{x\in{s_2}}g_2^2(x)(R^2(x)+V(x))\Big) \nonumber \\
&-&\EX_{1,2}\Big(\frac{1}{n_1n_2}\sum_{x\in{s_2}}g_2^2(x)(R^2(x)+V(x))\Big)
\end{eqnarray}
Grouping the terms and since the g-weights tend to $1$ we get
\begin{eqnarray}\label{threetwogweightvar}
\EX_{1,2,3}\big(\hat{\VAR}(\hat{Y}^*_{g3reg})\big)&=& \frac{1}{n_0}\VAR(\hat{Y}_1(x))  \nonumber \\
&+& \EX_{1,2}\frac{1}{n_1n_2}\sum_{x\in{s}_2}((g_1^{(1)}(x))^2-g_2^2(x))V(x) \nonumber \\
&+&\frac{n_2}{n_1}\bar{\pmb{Z}}^{(1)t}
\pmb{\Sigma}_{\hat{\pmb{\alpha}}_{s_2}}\bar{\pmb{Z}}^{(1)} + (1-\frac{n_2}{n_1})\bar{\pmb{Z}}^{t}\pmb{\Sigma}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}^{t} \nonumber \\
&+& \frac{1}{n_2}\EX_x(V(x)) \nonumber \\
&=& \VAR(\hat{Y}_{g3reg})+ \frac{1}{n_2}\EX_x(V(x))
\end{eqnarray}
We have proved that in three-phase two-stage sampling the external variance estimate [\ref{estexternvaryg3reg}] and the g-weight variance estimate [\ref{estimatedthreephasetwostagevariance}] are asymptotically consistent and equivalent. \\
\textbf{One can obtain the results for two-phase two-stage and three-phase two-stage sampling by replacing in all formulae the local density $Y(x)$ by the general local density $Y^*(x)$ in all the corresponding formulae valid for two-phase one-stage and three-phase one-stage sampling schemes.}\\
The remarks on two-phase systematic sampling remain a fortiori valid for three-phase systematic sampling: the impact of the long range spatial correlation of the $Y(x)$ or $\hat{Y}(x)$ being smaller than under two-phase sampling ($\frac{1}{n_0}<<\frac{1}{n_1}$), and because the residuals $R(x)$, $R_1(x)$ have a spatial correlation usually very small in comparison to $F$, likewise for the better statistical properties of the g-weight variance estimates (see \cite{mandallazreport3} pp. 24-29 for a stratification example with exhaustive first-phase).

\section{Concluding remarks}
It is clear that the same results will apply to the estimation of a small area $G\subset F$: for the external variance this is trivially achieved by restricting the samples to the small area, whereas in the g-weight approach one extends the reduced model by the indicator variable of $G$ (thus ensuring zero mean residuals over $G$) and replaces the empirical means of the auxiliary variables over $F$ by their means over $G\subset F$. The resulting g-weights $g^{(1)}(x)$, respectively $g_2(x)$, tend to $0$ for $x\not\in{G}$ and to $\frac{n_1}{n_{1,G}}$, respectively $\frac{n_2}{n_{2,G}}$, for $x\in{G}$.\\
The results extend trivially to cluster sampling if the number of plots per cluster is constant. The ideas are the same but the algebra is slightly more cumbersome if this not the case (see \cite{mandallaz} section 4.5 to get the flavor of the required modifications). Formal proof for two/three phase one-stage sampling schemes are given in the technical reports \cite{mandallazreport1}, \cite{mandallazreport2} and \cite{mandallazreport3} and it should be clear from the arguments given in this technical report that the same results will also hold in two/three phase two-stage cluster sampling: details are left as an exercise for the reader. Interestingly, Feynman did not comment this lame excuse, also over-used in mathematical books or papers!\\
Current work is devoted to the development of geostatistical Kriging versions of the three-phase design-based estimators.

\newpage
\bibliography{biblio1}
\end{document}

### 2018-08-31 revision/reviews.tex
Further comments about reviewer 1's coverage rate objections.  I included code for a simulation under the inf pop approach of different coverage rates by sampling fraction that demonstrate his point.

@@
 In fact, there is some automatic recoding which in part does just that (page 27).
-Some related discussion is found on page 28.\\
+Some related discussion is found on page 28.
+\\
+
+\answerfin{The reviewer makes a valid point that it would be easier from a "plug-and-play" perspective to simply interpret what the phase memberships were via the locations of NAs.
+However, NAs in covariates may have nothing to do with the sampling design e.g.
+non-response due to inaccessibility or lidar error.
+We feel that the current implementation of phase\_id forces the user to be more attentive when these situations occur that may interfere with the uniform and independent sampling assumption for the sampling design.
+To assist the user in understanding how function calls are made we have included several examples for the usage of all estimators in both the documentation and the vignette using real-life examples.
+It should also be noted that defining the phase\_id for all observations with the same (terrestrial) id essentially does the same as the reviewer has described with a warning message given.
+However we do not recommend this usage for the aforementioned reasons.}
+\\
 
@@
 Below I include three numerical examples where the nominal 95\% confidence intervals never exceed a coverage of 85\%.\\
+
+\answerfin{The reviewer is concerned that more attention needs to be given concerning issues in boundary cases with the coverage rates.
+Although the given explanation seems to have switched $\alpha$ and $1-\alpha$ in a few places the point is helpful and clear especially with the generously provided examples included for the finite population case.
+We have reproduced similar results using simulation in the infinite population 2-phase context and have concluded that this indeed occurs with the sampling fraction between the first and second phases is less than $\alpha$.
+For this reason we have expanded the confidence interval section to inform the reader of this situation.}
+\\
 

### 2018-08-31 coverage_rates.R

@@ -1,0 +1,119 @@
+################
+
+# Artificial data:
+
+set.seed(0)
+d <- data.frame(
+  elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
+  vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observe in s2 only)
+  vol2 = rnorm(e, mean=300, sd=100)*(e<310),
+  resp3 = runif(e) < 0.03
+)
+
+# Coverage of true means of response variables by 95% confidence
+# intervals obtained from 10000 samples s2 after two-phase fitting:
+
+library(forestinventory)
+
+# PROBLEM 1: Small sample size (n2=5)
+true_value <- mean(d$vol1)
+d$phase <- 1
+cover <- replicate(1000, {
+  d$phase[sample(length(d$phase), 5)] <- 2
+  ci <- confint( twophase( vol1 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+# PROBLEM 2: Effect of explanatory variable
+true_value <- mean(d$vol2)
+cover <- replicate(10000, {
+  d$phase <- 1
+  d$phase[sample(length(d$phase), 80)] <- 2
+  ci <- confint( twophase( vol2 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+# PROBLEM 3: Distribution of response
+true_value <- mean(d$resp3)
+cover <- replicate(10000, {
+  d$phase <- 1
+  d$phase[sample(length(d$phase), 80)] <- 2
+  ci <- confint( twophase( resp3 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+
+
+##################################
+## Infinite Population Approach ##
+##################################
+
+target.suface <- function(x0, y0){
+  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
[…]

### 2018-09-04 coverage_rates.R
I am trying to make a clean simulation example to demonstrate the effect of sample size and sampling fraction on coverage rate using the simulation example.  The simulation takes a couple hours on my machine so I will have to push the finished graphic at a later time.

@@ -7,7 +7,7 @@ d <- data.frame(
   elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
   vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observe in s2 only)
   vol2 = rnorm(e, mean=300, sd=100)*(e<310),
-  resp3 = runif(e) < 0.03
+  resp3 = as.numeric(runif(e) < 0.03)
 )
 
 # Coverage of true means of response variables by 95% confidence
@@ -60,7 +60,7 @@ target.suface <- function(x0, y0){
 # The true mean value here is 39.17
 
 sample.generator <- function(n1, n2, target.surface=target.surface){
-  realization <- matrix(NA,n1,6)
+  realization <- matrix(NA,n1,11)
   for(i in 1:n1){
     x0=2*runif(1)
     y0=3*runif(1)
@@ -70,9 +70,14 @@ sample.generator <- function(n1, n2, target.surface=target.surface){
     realization[i,4]<-x0*x0
     realization[i,5]<-x0*y0
     realization[i,6]<-y0*y0
+    realization[i,7]<-x0 + rnorm(1, 1, 4)
+    realization[i,8]<-y0 + rnorm(1, -3, 7)
+    realization[i,9]<-x0*x0 + rnorm(1, 2, 3)
+    realization[i,10]<-x0*y0 + rnorm(1, 0.5, 1)
+    realization[i,11]<-y0*y0 + rnorm(1, -0.5, 2)
   }
   realization <- as.data.frame(realization) # warning: still contains true response variable for all 1st phase
-  names(realization) <- c("response","x","y","xx","xy","yy")
+  names(realization) <- c("response","x","y","xx","xy","yy","x_perturbed","y_perturbed","xx_perturbed","xy_perturbed","yy_perturbed")
   realization$phase <- 1
   realization$phase[sample(nrow(realization), n2)] <- 2
   realization
@@ -82,7 +87,7 @@ sample.generator <- function(n1, n2, target.surface=target.surface){
 # Set sample sizes
 num_replications <- 10000
 sampling_fractions <- 1:15/100
-n1_seq <- 5:15*100
+n1_seq <- seq(250, 3000, by = 250)
 true_value <- 39.17
 coverage_rates <- data.frame(n1=integer(), n2=integer(), sampling_fractions=numeric(), coverage_rates=numeric())
 
[…]

### 2018-09-05 revision/reviews.tex
coverage rate script and more comments to the reviewers

@@
 \answerfin{The reviewer is concerned that more attention needs to be given concerning issues in boundary cases with the coverage rates.
-Although the given explanation seems to have switched $\alpha$ and $1-\alpha$ in a few places the point is helpful and clear especially with the generously provided examples included for the finite population case.
-We have reproduced similar results using simulation in the infinite population 2-phase context and have concluded that this indeed occurs with the sampling fraction between the first and second phases is less than $\alpha$.
-For this reason we have expanded the confidence interval section to inform the reader of this situation.}
+The given explanation seems to have switched $\alpha$ and $1-\alpha$ in a few places.
+Regardless the concept of having a sampling fraction too low for the confidence level isn't really applicable in the infinite population approach where the total population of points is infinity.
+We have performed simulations under the infinite population approach and could not reproduce this point.
+However, the reviewer does have a valid point that discussing coverage rates would improve the paper.
+For this reason we have expanded the confidence interval section to include some simulated examples of coverage rates to develop the reader's intuition about when the sample size is too low to produce reliable results.}
 \\
@@
 \\
-\answerfin{Reviewer 1 made a very valid criticism about designed-based confidence intervals when the sampling fraction is too small that we would like to include in the paper and Section 5 is the most logical place to put it.} \\
+\answerfin{Reader 1 suggested that this section in fact be extended to include discussion about coverage rates.  We prefer to go in this direction.} \\
 
@@
 
-\answerfin{}
+\answerfin{$n_2G < 6$ is just a rule of thumb.
+In Section 2.4, the goal was merely to introduce the small area problem in a general sense for readers that were not familiar.
+Rules of thumb should not be blindly obeyed though so we thought it best to refer to it when examples are present to give context allow the reader to develop intuition.
+Further discussion of this rule of thumb including simulation examples are included in the confidence interval section in the context of coverage rates.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{This is usually the case, but the estimators are both random and the opposite may sometimes occur.
+They are asymptotically equivalent though.
+The point is that one should prefer the g-weight because it attempts to account for sampling variation in the coefficients whereas the external variance completely ignores it.
+This is explained on page 11 after the example for the global two-phase estimator.}
 % --------------------------- %
@@
 
-\answerfin{In this context "a biased estimate" is an estimate that is produced from a biased estimator.
-It would be a bit tedious to explain every time that an given estimate was produced from a biased estimator.}
+\answerfin{Unbiased point estimates are point estimates produced by unbiased point estimators.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The extended synthetic estimator does not necessarily need to implement an ancova model.
+That depends on the types of variables that the user has specified.
+The extended synthetic estimator merely adds an indicator variable for the small area in addition to the other variables that the user specified.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The reviewers comments refer to the version of the package that is already available on CRAN and has over 5000 downloads.
+We prefer not to alter the function parameters in order to preserve backwards compatibility.}
 % --------------------------- %
@@
 
-\answerfin{It is elegant in the sense that the SAE estimator is unbiased merely by modifying the design matrix.
-It is not necessary to refit every small area with an indicator, but it is  how the package implemented the estimator.}
+\answerfin{It is elegant in the sense that the estimator is unbiased merely by modifying the design matrix.
+It is not necessary to refit every small area with an indicator, but this is it was implemented in the package.}
 % --------------------------- %
@@
 There may not even be observations available from within small area G.
-Please describe and define the term $\theta_s2$ in more detail.}.
+Please describe and define the term $\theta_{s_2}$ in more detail.}.
 
-\answerfin{It says that "the formulas look similar to the synthetic estimators", because they are exactly same.  They merely have a modified design matrix.}
+\answerfin{It says that "the formulas look similar to the synthetic estimators", because the formulas are exactly same.
+They merely have a modified design matrix that ensures that the residual correction term is zero by construction and doesn't need to be included.
+$\theta_{s_2}$ is now clearly defined in Eq.
+13 and the preceeding paragraph.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The reviewer's comment refers to the help documentation of the R-package.
+The twophase() function can be used to estimate in both the global and small area contexts.
+$Z_bar_1G$ refers to the small area context whereas $Z_bar_1$ is the terminology used for global estimation (i.e.
+when G = F).
+In the future release of the package we will try to include a short reference to this distinction in the help documentation but given that the package already has over 5000 downloads we prefer not to change the names of the function outputs.}
 % --------------------------- %

### 2018-09-05 coverage_rates.R

@@ -60,7 +60,7 @@ target.suface <- function(x0, y0){
 # The true mean value here is 39.17
 
 sample.generator <- function(n1, n2, target.surface=target.surface){
-  realization <- matrix(NA,n1,11)
+  realization <- matrix(NA,n1,6)
   for(i in 1:n1){
     x0=2*runif(1)
     y0=3*runif(1)
@@ -70,14 +70,9 @@ sample.generator <- function(n1, n2, target.surface=target.surface){
     realization[i,4]<-x0*x0
     realization[i,5]<-x0*y0
     realization[i,6]<-y0*y0
-    realization[i,7]<-x0 + rnorm(1, 1, 4)
-    realization[i,8]<-y0 + rnorm(1, -3, 7)
-    realization[i,9]<-x0*x0 + rnorm(1, 2, 3)
-    realization[i,10]<-x0*y0 + rnorm(1, 0.5, 1)
-    realization[i,11]<-y0*y0 + rnorm(1, -0.5, 2)
   }
   realization <- as.data.frame(realization) # warning: still contains true response variable for all 1st phase
-  names(realization) <- c("response","x","y","xx","xy","yy","x_perturbed","y_perturbed","xx_perturbed","xy_perturbed","yy_perturbed")
+  names(realization) <- c("response","x","y","xx","xy","yy")
   realization$phase <- 1
   realization$phase[sample(nrow(realization), n2)] <- 2
   realization
@@ -87,45 +82,25 @@ sample.generator <- function(n1, n2, target.surface=target.surface){
 # Set sample sizes
 num_replications <- 10000
 sampling_fractions <- 1:15/100
-n1_seq <- seq(250, 3000, by = 250)
+n1_seq <- seq(300, 2700, by = 300)
 true_value <- 39.17
 coverage_rates <- data.frame(n1=integer(), n2=integer(), sampling_fractions=numeric(), coverage_rates=numeric())
+pb <- txtProgressBar(min = 0, max = 3000, style = 3)
 
 for(n1 in n1_seq){
   for(n2 in n1*sampling_fractions){
     cover <- replicate(num_replications, {
       realization <- sample.generator(n1, n2)
-      ci <- confint( twophase( response ~ x + y + xx, data = realization,
+      ci <- confint( twophase( response ~ x + xy, data = realization,
                                phase_id = list(phase.col="phase", terrgrid.id = 2) ), level=0.95 )
[…]

### 2018-09-11 revision/reviews.tex
update 11_09_2018(andi)

--- 2018-08-30 revision/reviews.tex
@@
 
-	\answerfin{Has been changed accordingly.}
+	\answerfin{Has been added.}
 	% --------------------------- %
@@
 
-	\answerfin{Has been changed accordingly.}
+	\answerfin{Has been changed.}
 	% --------------------------- %
@@
 
-	\answerfin{Has been changed accordingly.}
+	\answerfin{}
 	% --------------------------- %
@@
 
-	\answerfin{Whereas we agree that they might be slightly more readable without the n2 fractions, the decision to include them was based on the fact that this is how the formulas were originally presented in Mandallaz's papers.
-We feel that the marginal loss in readability is offset by perfect continuity with the original notation.}
+	\answerfin{}
 	% --------------------------- %
@@
 
-	\answerfin{It has been added that the models require intercept terms.}
+	\answerfin{}
 	% --------------------------- %
@@
 Describing just once that in all cases both types are possible (as done for cluster sampling), would make the text a lot easier to read.
-\\
-\\
-\answerfin{Different versions of this article were written that attempted to shorten it in ways just like the review mentioned.
-However, we ultimately decided that the purpose of this article is to bridge the gap between likely users of the software (e.g.
-forest engineers and technical practitioners) and the mathematical details that might be beyond their skillset.
-For this reason, we decided to ensure that every function call was connected explicitly to the underlying formula that it applies and a citation is given for those interested in further technical details.
-We believe this also enables the reader to write their own routines using the functions which can be used to error check our routines.
-Considering that the current length is not excessively long compared to other JSS papers, we strongly prefer to not to shorten sections merely for the sake of making the article shorter.} \\
-
-Also section 5 on CIs could be considerably shortened. Needs to be a section on its own right? Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\
-\\
-\answerfin{Reviewer 1 made a very valid criticism about designed-based confidence intervals when the sampling fraction is too small that we would like to include in the paper and Section 5 is the most logical place to put it.} \\
+Also section 5 on CIs could be considerably shortened.
+Needs to be a section on its own right?
+Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\
 
@@
 
-\answerfin{When the package was orginally beta tested at the Swiss Federal Institute of Forest, Snow and Landscape Management on likely operational users as well as forest engineers, the most contentious area for confusion was the calculation of explanatory variables in the exhaustive case.
-The reviewer appears to have confirmed that confusion in the previous paragraph, which is explained in the section on calculating explanatory variables.
-It is explictly stated that the exact exhaustive mean vector can only be calculated for explanatory variables based on means (note that polygon layers are means of binary variables).
-It has now been explicitly added as an example that order statistics are not included in this category with the reviewers example given.
-It is already explicitly stated that the size of the support should ideally be chosen to achieve a best possible explanatory power of the regression model on page 9 and a citation has been added for the interested reader.} \\
-
 Related to the previous comment is whether the difference between the infinite and finite population approaches is relevant from a practical point of view.
@@
 
-\answerfin{The population doesn't need to be tessellated and can be allowed to overlap or not use all the wall-to-wall map in the infinite population approach.
-This is already mentioned on page 9.
-Section 2.1 discusses the Infinite Population Approach.
-Some brief discussion comparing to the finite is included, but as this package is about software written for the infinite not finite approach, we prefer not to delve too deeply into a topic that makes little practical difference.} \\
-
 When providing two estimators for the same purpose such as the synthetic or extended synthetic, some (extreme) users might just chose what results in the smallest standard error.
 Some more guidance or discussion on when to use which estimator may be needed.\\
-
-\answerfin{This is true, but it is already stated many times that the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias making them seem deceptively optimistic (see pages 5, 13 (twice), 14, 16 (this is explained here using a concrete example), 21, 26 and 30)} \\
 
@@
 
-\answerfin{Removing outliers is more of a common practice in model-dependent estimation, which is beyond the scope of this article.
-You should never remove outliers in design-based estimation unless you can defend the belief that the resulting (and self-inflicted) non-response is missing at random.
-The observed value of the the response/explanatory variable is fixed at each $x$ so deleting an observation effectively means the plot at x has been removed from the forest.
-For temporal mismatches this is pretty much indefensible because the outliers likely represent harvesting after the aerial measurement.
-We initially included a discussion comparing design-based with model-dependent estimation but we ultimately decided that the paper has more focus if we concentrate only on explaining how to use this software under the design-based infinite population approach.
-In any case, many citations are currently included that already discuss these issues.} \\
-
 Dealing with sample plots close to borders (split plots) is a common issue in forest inventories. It would be helpful to mention how to handle them in the estimators.\\
-
-\answerfin{A mention of sample plot boundary adjustments is now included in Section 2.1 with a citation.} \\
 
@@
 
-\answerfin{Multiphase is now used consistently to match package's spelling}
+\note{\textit{Andi}: There is a slight inconsistency in our article here: Our package says 'multiphase', but here we used 'multi-phase'. However, should we then also change 'two-phase' to 'twophase'?}
+
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{Good point.  Has been changed accordingly.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{That suggestion only can be done if the plots form a perfect tesselation.
-Circular units, the most common in NFIs, can't do this.
-The rest of the sentence actually says "with respect to the forest area".
-This is true.
-Points can fill any forested area even if it is a fractal whereas a counter example forest area can always be given for any plot shape you put forward.
-Extending your tesselation beyond the forest edge is possible, but this increases the weights for boundary cases which likely increases the variance.
-Many of the cited articles in the infinite population approach already discuss this but the focus for this article is not a comparison of the finite and infinite approaches, but a presentation of a software package written under the infinite approach.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{The g-weight is the name of a variance estimator for the point estimate in eq.2 as opposed to another possible variance estimator here named the external estimator.
-The origin of the g-weight name is more obvious in the citation provided in the same paragraph.
-The g-weight is analogous to a hat matrix and can be extracted from the more compact form that was included in the article.
-It the spirit of not being too verbose we chose the compact form.
-}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{In this context "a biased estimate" is an estimate that is produced from a biased estimator.
-It would be a bit tedious to explain every time that an given estimate was produced from a biased estimator.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{The "survey" package is in the finite population approach (aka list sampling).  It is mentioned on page 2.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{It is elegant in the sense that the SAE estimator is unbiased merely by modifying the design matrix.
-It is not necessary to refit every small area with an indicator, but it is  how the package implemented the estimator.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{It says that "the formulas look similar to the synthetic estimators", because they are exactly same.  They merely have a modified design matrix.}
+\answerfin{}
 % --------------------------- %
@@
 
-\answerfin{Change has been implemented.}
+\answerfin{}
 % --------------------------- %

### 2018-09-11 rnw/Acknowledgements.Rnw
update 11_09_2018(andi)

@@
 We also want to thank Daniel Mandallaz for his support in completing the range of the already published estimators in the frame of the three-phase small area estimators, as well as many helpful discussions and advice throughout the implementation of our package.
-Our thanks also go to Meinrad Abegg for proofreading the manuscript, and to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of grisons for providing the example data.
+Our thanks also go to Meinrad Abegg for proofreading the manuscript, and to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of Grisons for providing the example data.

### 2018-09-11 rnw/Main structure and modules of the Package.Rnw
update 11_09_2018(andi)

--- 2017-10-27 rnw/Main structure and modules of the Package.Rnw
@@
 \resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
-\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}.}
+\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}. All estimators are also available for cluster sampling.}
 \label{fig:struct_package}

### 2018-09-11 revision/simulation/simulation_alex.R

@@ -1,0 +1,262 @@
+################
+
+# Artificial data:
+
+set.seed(0)
+d <- data.frame(
+  elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
+  vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observe in s2 only)
+  vol2 = rnorm(e, mean=300, sd=100)*(e<310),
+  resp3 = runif(e) < 0.03
+)
+
+# Coverage of true means of response variables by 95% confidence
+# intervals obtained from 10000 samples s2 after two-phase fitting:
+
+library(forestinventory)
+
+# PROBLEM 1: Small sample size (n2=5)
+true_value <- mean(d$vol1)
+d$phase <- 1
+cover <- replicate(1000, {
+  d$phase[sample(length(d$phase), 5)] <- 2
+  ci <- confint( twophase( vol1 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+# PROBLEM 2: Effect of explanatory variable
+true_value <- mean(d$vol2)
+cover <- replicate(10000, {
+  d$phase <- 1
+  d$phase[sample(length(d$phase), 80)] <- 2
+  ci <- confint( twophase( vol2 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+# PROBLEM 3: Distribution of response
+true_value <- mean(d$resp3)
+cover <- replicate(10000, {
+  d$phase <- 1
+  d$phase[sample(length(d$phase), 80)] <- 2
+  ci <- confint( twophase( resp3 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+} )
+mean(cover)
+
+
+
+##################################
+## Infinite Population Approach ##
+##################################
+
+target.suface <- function(x0, y0){
+  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
[…]

### 2018-09-12 revision/reviews.tex
This is all my final comments and changes covering all issues except for the development of the coverage rate examples in the CI section.  I tried to include some of Daniel's comments too.

--- 2018-09-05 revision/reviews.tex
@@
 \answerfin{The reviewer is concerned that more attention needs to be given concerning issues in boundary cases with the coverage rates.
-The given explanation seems to have switched $\alpha$ and $1-\alpha$ in a few places.
-Regardless the concept of having a sampling fraction too low for the confidence level isn't really applicable in the infinite population approach where the total population of points is infinity.
-We have performed simulations under the infinite population approach and could not reproduce this point.
-However, the reviewer does have a valid point that discussing coverage rates would improve the paper.
-For this reason we have expanded the confidence interval section to include some simulated examples of coverage rates to develop the reader's intuition about when the sample size is too low to produce reliable results.}
+The given explanation seems to have switched $\alpha$ and $1-\alpha$ in a few places but made clear that there was a link between too low sampling fractions and coverage rates under the finite population approach.
+However, the concept of having a sampling fraction too low for the confidence level isn't really applicable in the infinite population approach where the total population is a continuum of points and is thus infinity.
+We have performed simulations under the infinite population approach and could not reproduce the reviewers examples, but did show that extremely low sample sizes could have an influence.
+The cited literature includes many simulations and artificial examples demonstrating the distribution of the parameter of interest (the integral of the local density function) is beyond analytical treatment under the sampling schemes but can be well approximated by t-distribution (and consequently the normal in large samples).
+The reviewer does have a very valid point that discussing coverage rates would improve the paper.
+For this reason we have expanded the confidence interval section to include some new simulated examples of coverage rates to develop the reader's intuition about when the sample size is too low to produce reliable results.}
 \\
@@
 
-\answerfin{When the package was orginally beta tested at the Swiss Federal Institute of Forest, Snow and Landscape Management on likely operational users as well as forest engineers, the most contentious area for confusion was the calculation of explanatory variables in the exhaustive case.
-The reviewer appears to have confirmed that confusion in the previous paragraph, which is explained in the section on calculating explanatory variables.
+\answerfin{This is explained in the section on calculating explanatory variables.
 It is explictly stated that the exact exhaustive mean vector can only be calculated for explanatory variables based on means (note that polygon layers are means of binary variables).
@@
 \answerfin{The estimators presented in our article particularly rely on the assumption that the forest area $\lambda$ is known.
-The reason for this is that the local density $Y(x)$ is actually the Horwitz-Thompson}
-
-\note{\textit{Andi}: Whereas the forest area $\lambda(F)$ appears in the Horwitz-Thompson estimator, we actually only have to know the forest boundaries to compute the intersection of the circle and the forest boundary.
 The exact forest area is only needed to compute the total by multiplying, e.g., the sample mean with $\lambda(F)$.
 The author is, however, right by saying that in most NFIs the forest area is rather estimated than known exactly.
-A solution, as applied in the Swiss NFI, is to estimate the forest area by multiplying the \textit{known} country area by the percentage of (randomly and uniformly distributed) sample points falling within the forest (decision made by field crews) -> see Mandallaz 2014: Note on the estimation of totals.
-So actually, its no big deal "to do it in the infinite population approach".
-However, in the case of multiphase sampling, also $s_1$ or $s_0$ points can no longer be restricted to the forest, which can lead to loss of efficiency (see also Mandallaz 2014).}
+A solution, as applied in the Swiss NFI, is to estimate the forest area by multiplying the \textit{known} country area by the percentage of (independent and uniformly distributed) sample points falling within the forest (decision made by field crews) -> see Mandallaz 2014: Note on the estimation of totals.
+So actually, its no big deal "to do it in the infinite population approach" but how the user wants to estimate forest area is up to them.
+A citation noting where the reader can find more information on this topic has been included but further discussion was not included as it was outside the scope of this paper.}
 
@@
 
-\answerfin{}
+\answerfin{The reviewer is correct and this change has been implemented.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{The term is consistent with much of the already published literature included in the citations.}
 % --------------------------- %
@@
 \answerfin{The reviewers comments refer to the version of the package that is already available on CRAN and has over 5000 downloads.
-We prefer not to alter the function parameters in order to preserve backwards compatibility.}
+We prefer not to alter the function parameters in order to preserve backwards compatibility.
+In the future if the usefulness of more general package names becomes apparent a wrapper package could be created.}
 % --------------------------- %
@@
 \answerfin{It is elegant in the sense that the estimator is unbiased merely by modifying the design matrix.
-It is not necessary to refit every small area with an indicator, but this is it was implemented in the package.}
+It is not necessary to refit every small area with an indicator, but this is how it was implemented in the package.}
 % --------------------------- %
@@
 
-\answerfin{}
+\answerfin{Parts of the equation are directly referenced in the sentence to make it easier to follow.}
 % --------------------------- %

### 2018-09-12 rnw/Main structure and modules of the Package.Rnw
This is all my final comments and changes covering all issues except for the development of the coverage rate examples in the CI section.  I tried to include some of Daniel's comments too.

--- 2018-08-30 rnw/Main structure and modules of the Package.Rnw
@@
 Our strategy is thus to take a sample of $n_2$ points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.
-The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.
+The total timber volume can then be obtained by multiplying $\hat{Y}$ by the known surface area of the forest, $\lambda(F)$.
+It should be noted $\lambda(F)$ is often estimated in practice (see \cite{mandallaz2014}) for further explanation) but that is beyond the scope of this article.
 All estimators included in \pkg{forestinventory} rest upon this theoretical perspective.
@@
 They are therefore also referred to as model-assisted.
-It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to minimize travelling costs in the terrestrial survey.
+It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to achieve a more spatially balanced sample in the terrestrial survey.
 However, there is reasonable evidence that softening this assumption is acceptable for point and variance estimation as long as the grid does not interact with periodic features in the forest structure \citep{mandallaz2008}.

### 2018-09-12 rnw/Threephase_Estimators and their Application.Rnw
This is all my final comments and changes covering all issues except for the development of the coverage rate examples in the CI section.  I tried to include some of Daniel's comments too.

@@
 
-Intuitively, the three phase estimator is simply the mean of the predictions using the reduced model, corrected by the mean difference between the reduced model predictions and the more accurate full model predictions, corrected by the mean difference between the ground truth and the full model predictions.
+Intuitively, the three phase estimator is simply the mean of the predictions using the reduced model (i.e.
+$\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}$), corrected by the mean difference between the reduced model predictions and the more accurate full model predictions (i.e.
+$\frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2})$), corrected by the mean difference between the ground truth and the full model predictions (i.e.
+$\frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})$).
 For the compact version of the formula in the non-exhaustive case, the estimated means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the estimated mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq.

### 2018-09-14 revision/reviews.tex
update Andi 14_09_2018(1)
update Andi 14_09_2018(1)

@@
 The g-weight is analogous to a hat matrix and can be extracted from the more compact form that was included in the article.
-It the spirit of not being too verbose we chose the compact form.
+In the spirit of not being too verbose we chose the compact form.
 }
@@
 That depends on the types of variables that the user has specified.
-The extended synthetic estimator merely adds an indicator variable for the small area in addition to the other variables that the user specified.}
+The extended synthetic estimator merely adds an indicator variable for the small area in addition to the other variables that the user specified.
+This can in fact extend the model to an ancova model if the explanatory variables contain at least one continuous predictor.
+However, the reason of adding the indicator variable exclusively is to ensure the zero-mean-residual property in both the global domain F and small area G, and not to improve the model fit.
+We think that linking the indicator variable to the anova or ancova modeling technique might probably cause misunderstanding of the indicator purpose on the side of the readers.}
 % --------------------------- %
@@
 
-\answerfin{The term is consistent with much of the already published literature included in the citations.}
+\answerfin{The term is consistent with much of the already published literature included in the citations (e.g., von L{\"u}pke et al. 2012).}
 % --------------------------- %
@@
 
-\answerfin{The "survey" package is in the finite population approach (aka list sampling).  It is mentioned on page 2.}
+\answerfin{The \textit{survey} package is in the finite population approach (aka list sampling).  It is mentioned on page 2.}
 % --------------------------- %

### 2018-09-14 rnw/Introduction.Rnw
update Andi 14_09_2018(1)
update Andi 14_09_2018(1)

@@
 The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017.
-The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}.
+The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2018b}.
 The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}.

### 2018-09-16 JStatSoft_forestinventory_master-concordance.tex
update andi 16_09_2018(1)
update andi 16_09_2018(1)

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
-1 2 1 1 4 38 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
+1 2 1 1 4 38 1 1 2 11 0 1 2 80 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1309:%
 141 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1237:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1314:%
 1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
@@
 0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1472:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1549:%
 147 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1477:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1554:%
 1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
@@
 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1666:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
 153 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1671:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1683:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
 159 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1688:%
-1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1693:%
-165 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
+1 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1772:%
+165 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1777:%
+1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
+3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 22 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1956:%
+171 11 1}

### 2018-09-16 rnw/Appendix.Rnw
update andi 16_09_2018(1)
update andi 16_09_2018(1)

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:
<<echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
@

\section*{Appendix}
\label{sec:appendix}

\subsection*{Simulation for global two-phase estimation}

\begin{small}
<<eval=FALSE>>=
# function to create density surface:
target.suface <- function(x0, y0){
  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
@
\end{small}

\begin{small}
<<eval=FALSE>>=
# --- True spatial mean for Global and Small Area Surface Area: --- #
library(rmutil)
int2(target.suface, a = c(0,0), b=c(3,2))/6 # 39.17
int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) # 37.16243
@
\end{small}

\begin{small}
<<eval=FALSE>>=
# --- sample generator for simple sampling: --- #
sample.generator<- function(n1, n2, target.surface=target.surface){
  realization <- matrix(NA,n1,7)
  for(i in 1:n1){
    x0=2*runif(1)
    y0=3*runif(1)
    realization[i,1] <- target.suface(x0, y0)
    realization[i,2]<-x0
    realization[i,3]<-y0
    realization[i,4]<-x0*x0
    realization[i,5]<-x0*y0
    realization[i,6]<-y0*y0
    realization[i,7]<-0
  }
  realization <- as.data.frame(realization)
  names(realization) <- c("response","x","y","xx","xy","yy", "area")
  realization$phase <- 1
  realization$phase[sample(nrow(realization), n2)] <- 2
  realization$response[realization$phase == 1]<- NA
  realization$area[realization$x >= 0.3 & realization$x <= 1.3 & 
                   realization$y >= 0.5 & realization$y <= 2]<- 1
  return(realization)
}
@
\end{small}

\begin{small}
<<eval=FALSE>>=

# --- function for global simulations --- #
run.global.sim <- function(num_replications, n1.vec){

  sfrac <- seq(0.02, 0.26, by=0.02) # sampling fractions
  true_value<- 39.17 # true population mean

  # data.frame to store results:
  coverage_rates <- data.frame(n1=integer(), n2=integer(), 
                               sampling_fractions=numeric(), 
                               avg.error.ext=numeric(),
                               avg.error.g=numeric(),
                               covrate.ext=numeric(),
                               covrate.g=numeric(),
                               no.valid.sims = integer())

  # -------------- #
  # simulation loop:

  for(i in 1:length(sfrac)){

    # variables to store temporary results:
    n2 <- integer()
    error.ext <- numeric()
    error.g <- numeric()
    ci.ext.logical <- logical()
    ci.g.logical <- logical()
    no.valid.sims. <- integer()

    for(j in 1:num_replications){

      realization <- sample.generator(n1 = n1.vec, n2 = round(n1.vec*sfrac[i]))

      est <- twophase(formula = response ~ y + x + xx, data=realization,
                      phase_id = list(phase.col = "phase", terrgrid.id = 2))

      ci <- confint(est)
      ci.ext.logical <- append(ci.ext.logical, 
                               ci$ci$ci_lower_ext <= true_value & 
                               true_value <= ci$ci$ci_upper_ext)
      ci.g.logical <- append(ci.g.logical, 
                             ci$ci$ci_lower_g <= true_value & 
                             true_value <= ci$ci$ci_upper_g)
      error.ext <- append(error.ext, 
                          sqrt(est$estimation$ext_variance)/est$estimation$estimate)
      error.g <- append(error.g, 
                        sqrt(est$estimation$g_variance)/est$estimation$estimate)

      n2 <- append(n2, est$estimation$n2)
      no.valid.sims. <- sum(n2 > 1)

    }

    # calculate cove.rates, avg. n2G etc:
    coverage_rates <- rbind(coverage_rates, 
                            data.frame(n1=n1.vec, n2=round(mean(n2)), 
                                       sampling_fractions=sfrac[i], 
                                       avg.error.ext=mean(error.ext, na.rm = TRUE),
                                       avg.error.g=mean(error.g, na.rm = TRUE),
                                       covrate.ext=mean(ci.ext.logical, na.rm = TRUE), 
                                       covrate.g=mean(ci.g.logical, na.rm = TRUE),
                                       no.valid.sims=no.valid.sims.))

  } # end of sim

return(coverage_rates)

} # end of function
@
\end{small}

\newpage

 \begin{small}
<<eval=FALSE>>= 
# --- run global simulations --- #
library(forstinventory)
global.sims<- data.frame()
n1_sims<- c(250, 500, 750, 1000, 1500, 2000) # n1 sample sizes
  for (i in 1:length(n1_sims)){
    res<- run.global.sim(num_replications = 10000, n1.vec = n1_sims[i])
    global.sims <- rbind(global.sims, res)
  }
@
\end{small}

\newpage

### 2018-09-16 rnw/Calculation of Confidence Intervals.Rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.
update andi 16_09_2018(1)
update andi 16_09_2018(1)

@@
 
-\vspace{8mm}
+\vspace{4mm}
 
@@
 
-\vspace{7mm}
+\vspace{4mm}
 
@@
 In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).
+Note that in case of synthetic estimations (SYNTH, PSYNTH), the degrees of freedom are $n_{2}-p$ as is the case for global estimation.
 In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function.
@@
 
+% ---------------------------------------------- %
+% Coverage rates
+
+% notes:
+%
+% - address link between too low sampling fractions and coverage rates under the finite population approach
+% - we did show that extremely low sample sizes could have an influence
+% - aim:  develop the reader’s intuition about when the sample size is too low to produce reliable results
+%
+
+A commonly raised question of practitioners regarding the application of design-unbiased estimators is which minimum terrestrial sample size should be provided.
+Answering this question is highly related to ensuring the validity of the confidence coverage rates by a sufficiently large sample size.
+It has to be noted that giving a generally valid sample size for the asymptotic validity range is, unfortunately, infeasible because the convergence to the \textit{t}- or normal distribution also depends on the hetereogeneity of the underlying population.
+However, simulation studies using an artificial local density surface were conducted in \citet{mandallaz2013a} and \citet{mandallaz2013b} to empirically check the asymptotic properties of the estimators.
+In order to develop the reader's intuition about when the sample size is too low to produce reliable results, we re-evaluated these simulations for a broader range of sample sizes.
+For each fixed sample size $n_1$ and sampling fraction, 10000 samples $s_2$ were conducted by simple random sampling and the 95\% confidence interval based on the g-weight variance was calculated for each sample after applying the global two-phase estimator (Section \ref{sec:twophase_globest}).
+The $R^2$ of the regression model was 0.81.
+For further details on the simulation, the reader is referred to the original articles mentioned above.
+The results can also be reproduced or expanded using the \proglang{R}-code given in the Appendix.\par
+
+From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for sample sizes larger equals $n_2$=50 regardless of the chosen sample size $n_1$.
+The realized coverage rate only drops significantly below 90\% if considerably small sampling fractions (i.e., the ratio ${n_2}/{n_1}$) < 0.1 are used that lead to small terrestrial sample sizes of 20 and less.
+The simulation examples also confirmed that in such cases, one should whenever possible consider one of the design-unbiased small area estimators (P)SMALL or EXT(P)SYNTH for which the nominal coverage rates in the simulations were already met for sample sizes $n_{2,G}$ > 5.
+Finally, it should also be noted that applying the design-based estimators under systematic sampling will lead to an inflation of the estimated variance \citep{mandallaz1993}.
+This will also effect the actual coverage rates to be higher than those derived under simple random sampling.
+
+\begin{figure}[h]
+\centering
+\resizebox{0.80\hsize}{!}{\includegraphics{fig/globalsim.png}}
+\caption{Coverage rates realized undervarying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}.
+The nominal confidence level was 95\% and for each setup, 10000 simulations were performed using simple random sampling.}
+\label{fig:covrate_sim}
+\end{figure}
+
+% 
+% 
+% A commonly raised critic on the proposed design-unbiased estimators are their asymptotic 638
+% properties, i.e. the validity of the conﬁdence coverage rates is only ensured if the sample size in a small 639
+% area is sufﬁciently large. Giving a generally valid sample size for the asymptotic validity range is, 640
+% unfortunately, infeasible due to the dependency on the heterogeneity of the underlying population 641
+% which is per se unknown. However, simulation studies for simple random sampling presented in 642
+% Mandallaz et al. [25] suggested that a minimum sample size of 6 within a small area is sufﬁcient 643
+% to ensure the nominal coverage rates of the conﬁdence intervals for PSMALL and EXTPSYNTH. 644
+% Re-evaluating the same simulation example recently conﬁrmed the same results for cluster sampling.
+%
+
 \newpage

### 2018-09-16 simuations covrates/graphic.R

@@ -1,0 +1,66 @@
+global.sims<- readRDS("D:\\Holzvorratschaetzung_RLP\\Artikel forestinventory\\vignette_forestinventory\\simuations covrates\\simres_global.rds")
+
+library(ggplot2)
+
+globalsim.fig<- 
+ggplot(data=global.sims) + 
+  geom_line(aes(x = n2, y=covrate.g, group=as.factor(n1), color=as.factor(n1)), size=rel(1)) +
+  geom_point(aes(x = n2, y=covrate.g, group=as.factor(n1), color=as.factor(n1)),  size=rel(3)) +
+  geom_hline(yintercept=0.95) +
+  guides(color=guide_legend(title="Sample size n1"))+ 
+  theme_bw() +
+  theme(axis.title.x =element_text(size=rel(3)), # size of axis-labels text
+        axis.title.y =element_text(size=rel(3)),
+        text = element_text(size=rel(7)), # text-size in panels
+        axis.text.x = element_text(size=rel(3)),
+        axis.text.y = element_text(size=rel(3)),
+        legend.title=element_text(size=rel(2.5)),
+        legend.text=element_text(size=rel(2)),
+        legend.key.size=unit(3, "line")) +
+  xlab("Terrestrial sample size n2") +
+  ylab("Coverage rate")
+
+
+setwd("D:\\Holzvorratschaetzung_RLP\\Artikel forestinventory\\vignette_forestinventory\\fig\\")
+ggsave("globalsim.png", 
+       globalsim.fig, width = 10, height = 7, dpi = 600)
+
+
+
+
+
+
+
+library(ggplot2)
+ggplot(data=global.sims) + 
+  geom_line(aes(x = sampling_fractions, y=covrate.g, group=as.factor(n1), color=as.factor(n1))) +
+  geom_point(aes(x = sampling_fractions, y=covrate.g, group=as.factor(n1), color=as.factor(n1)))
+
+# --------------------- #
+
+simres_sae<- readRDS("D:\\Holzvorratschaetzung_RLP\\Artikel forestinventory\\vignette_forestinventory\\simuations covrates\\simres_sae.rds")
+simres_sae_2<- readRDS("D:\\Holzvorratschaetzung_RLP\\Artikel forestinventory\\vignette_forestinventory\\simuations covrates\\simres_sae_2.rds")
+
+
+simres_sae<- rbind(simres_sae, simres_sae_2)
+
+
+
+library(ggplot2)
+ggplot(data=simres_sae) + 
[…]

### 2018-09-16 simuations covrates/simulation_coverage rates_2.R

@@ -1,0 +1,275 @@
+
+# -------------------------------------------------------------------------- #
+# -------------------------------------------------------------------------- #
+
+# function to create density surface:
+
+target.suface <- function(x0, y0){
+  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
+# The true mean value is derived from an integral can be calculated in analytically or in Matlab.
+# The true mean value here is 39.17
+
+# --- True spatial mean for Small area surface area: --- #
+library(rmutil)
+int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) # 37.16243
+
+
+# -------------------------------------------------------------------------- #
+# -------------------------------------------------------------------------- #
+
+
+# --- sample generator for simple sampling: --- #
+sample.generator<- function(n1, n2, target.surface=target.surface){
+  realization <- matrix(NA,n1,7)
+  for(i in 1:n1){
+    x0=2*runif(1)
+    y0=3*runif(1)
+    realization[i,1] <- target.suface(x0, y0)
+    realization[i,2]<-x0
+    realization[i,3]<-y0
+    realization[i,4]<-x0*x0
+    realization[i,5]<-x0*y0
+    realization[i,6]<-y0*y0
+    realization[i,7]<-0
+  }
+  realization <- as.data.frame(realization) # warning: still contains true response variable for all 1st phase
+  names(realization) <- c("response","x","y","xx","xy","yy", "area")
+  realization$phase <- 1
+  realization$phase[sample(nrow(realization), n2)] <- 2
+  realization$response[realization$phase == 1]<- NA
+  realization$area[realization$x >= 0.3 & realization$x <= 1.3 & realization$y >= 0.5 & realization$y <= 2]<- 1
+  return(realization)
+}
+
+# -------------------------------------------------------------------------- #
+# -------------------------------------------------------------------------- #
+
+# simulation for small area edge cases:
+
+library(forestinventory)
+num_replications <- 10000
+n1.vec <- c(100, 150, 200, 250, 500, 750, 1000) 
+sfrac <- 0.25
+true_value.sae<- 37.16
+
+# data.frame to store results:
+coverage_rates <- data.frame(n1=integer(), n2=integer(), 
+                             sampling_fractions=numeric(), 
[…]

### 2018-11-11 revision/reviews.tex
update 11_11_2018
update 11_11_2018

@@
 
-\title{Reviews}
+\title{Answer to Reviewers}
 
@@
 However, NAs in covariates may have nothing to do with the sampling design e.g.
-non-response due to inaccessibility or lidar error.
+non-response due to inaccessibility or sensor failures (see e.g.
+Hill, A., Mandallaz, D., \& Langshausen, J.
+(2018).
+A Double-Sampling Extension of the German National Forest Inventory for Design-Based Small Area Estimation on Forest District Levels.
+\textit{Remote Sensing}, 10(7), 1052.).
 We feel that the current implementation of phase\_id forces the user to be more attentive when these situations occur that may interfere with the uniform and independent sampling assumption for the sampling design.
@@
 My main concern is that the limitations of the methodology are nowhere clearly specified in the manuscript.
-On page 5, it is suggested that a consequence of the design-based approach "is that the estimation properties of design-based regression estimators (e.
+On page 5, it is suggested that a consequence of the design-based approach is that 'the estimation properties of design-based regression estimators (e.
 g.
-unbiasedness) typically hold regardless of the model that is chosen." While this is basically true for unbiasedness, unbiasedness by itself means little, and, more important, the statement does not extend to relevant properties such as coverage rates of confidence intervals.
+unbiasedness) typically hold regardless of the model that is chosen.' While this is basically true for unbiasedness, unbiasedness by itself means little, and, more important, the statement does not extend to relevant properties such as coverage rates of confidence intervals.
 Indeed, if s2 contains less than 1-$\alpha$ of the entire population, any true 1-$\alpha$ confidence interval for the population mean necessarily involves distributional assumptions about the population because otherwise any fixed Y(x0), which has probability > $\alpha$ of not being included in the sample, can be anything, so any conclusion about the mean would be wrong with probability > $\alpha$.
@@
 The vague invocation of the central limit theorem (page 24) won't help here.
-Below I include three numerical examples where the nominal 95\% confidence intervals never exceed a coverage of 85\%.\\
+Below I include three numerical examples where the nominal 95\% confidence intervals never exceed a coverage of 85\%.
+As I personally regard solid knowledge of the limitations of a statistical method as a prerequisite of its use in science and in addition believe it would help many potential users of the package, I ask the authors to consider adding some clarification regarding typical situations where the methods can be or should not be applied.
+For example, showing a few artificial test cases highlighting the most important requirements would probably not be difficult.\\
 
@@
 However, the concept of having a sampling fraction too low for the confidence level isn't really applicable in the infinite population approach where the total population is a continuum of points and is thus infinity.
-We have performed simulations under the infinite population approach and could not reproduce the reviewers examples, but did show that extremely low sample sizes could have an influence.
-The cited literature includes many simulations and artificial examples demonstrating the distribution of the parameter of interest (the integral of the local density function) is beyond analytical treatment under the sampling schemes but can be well approximated by t-distribution (and consequently the normal in large samples).
+We have performed simulations under the infinite population approach and could not reproduce the reviewer's examples, but did show that extremely low sample sizes could have an influence.
+The cited literature includes many simulations and artificial examples demonstrating that the distribution of the parameter of interest (the integral of the local density function) is beyond analytical treatment under the sampling schemes but can be well approximated by the \textit{t}-distribution (and consequently the normal distribution in large samples).
 The reviewer does have a very valid point that discussing coverage rates would improve the paper.
@@
 \\
-
-As I personally regard solid knowledge of the limitations of a statistical method as a prerequisite of its use in science and in addition believe it would help many potential users of the package, I ask the authors to consider adding some clarification regarding typical situations where the methods can be or should not be applied.
-For example, showing a few artificial test cases highlighting the most important requirements would probably not be difficult.\\
 
@@
 \\
-\answerfin{Reader 1 suggested that this section in fact be extended to include discussion about coverage rates.  We prefer to go in this direction.} \\
+\answerfin{Reader 1 suggested that this section should in fact be extended to include discussion about coverage rates.  We prefer to go in this direction.} \\
 
@@
 
-\answerfin{The population doesn't need to be tessellated and can be allowed to overlap or not use all the wall-to-wall map in the infinite population approach.
+\answerfin{The population doesn't need to be tessellated and can be allowed to overlap or not use all the wall-to-wall map information in the infinite population approach.
 This is already mentioned on page 9.
@@
 The author is, however, right by saying that in most NFIs the forest area is rather estimated than known exactly.
-A solution, as applied in the Swiss NFI, is to estimate the forest area by multiplying the \textit{known} country area by the percentage of (independent and uniformly distributed) sample points falling within the forest (decision made by field crews) -> see Mandallaz 2014: Note on the estimation of totals.
-So actually, its no big deal "to do it in the infinite population approach" but how the user wants to estimate forest area is up to them.
-A citation noting where the reader can find more information on this topic has been included but further discussion was not included as it was outside the scope of this paper.}
-
+A solution, as applied in the Swiss NFI, is to estimate the forest area by multiplying the \textit{known} country area by the percentage of (independent and uniformly distributed) sample points falling within the forest (decision made by field crews).
+A citation where the reader can find more information on this topic has been included.}
 % --------------------------- %
@@
 This can in fact extend the model to an ancova model if the explanatory variables contain at least one continuous predictor.
-However, the reason of adding the indicator variable exclusively is to ensure the zero-mean-residual property in both the global domain F and small area G, and not to improve the model fit.
-We think that linking the indicator variable to the anova or ancova modeling technique might probably cause misunderstanding of the indicator purpose on the side of the readers.}
+However, the reason of adding the indicator variable is purely to ensure the zero-mean-residual property in both the global domain F and small area G, and not to improve the model fit.
+We think that linking the indicator variable to the anova or ancova modeling technique might probably cause a misunderstanding of the indicator purpose on the side of the readers.}
 % --------------------------- %

### 2018-11-11 rnw/Appendix.Rnw
update 11_11_2018
update 11_11_2018

@@
 # --- run global simulations --- #
-library(forstinventory)
+library(forestinventory)
 global.sims<- data.frame()

### 2018-11-11 rnw/Calculation of Confidence Intervals.Rnw
update 11_11_2018
update 11_11_2018

@@
 In order to develop the reader's intuition about when the sample size is too low to produce reliable results, we re-evaluated these simulations for a broader range of sample sizes.
-For each fixed sample size $n_1$ and sampling fraction, 10000 samples $s_2$ were conducted by simple random sampling and the 95\% confidence interval based on the g-weight variance was calculated for each sample after applying the global two-phase estimator (Section \ref{sec:twophase_globest}).
+For each fixed sample size $n_1$ and sampling fraction, 10000 samples were conducted by simple random sampling and the 95\% confidence interval based on the g-weight variance was calculated for each sample after applying the global two-phase estimator (Section \ref{sec:twophase_globest}).
 The $R^2$ of the regression model was 0.81.
@@
 
-From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for sample sizes larger equals $n_2$=50 regardless of the chosen sample size $n_1$.
+From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for large sample sizes (e.g., $n_2$=50) regardless of the chosen sample size $n_1$.
 The realized coverage rate only drops significantly below 90\% if considerably small sampling fractions (i.e., the ratio ${n_2}/{n_1}$) < 0.1 are used that lead to small terrestrial sample sizes of 20 and less.
@@
 \resizebox{0.80\hsize}{!}{\includegraphics{fig/globalsim.png}}
-\caption{Coverage rates realized undervarying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}.
+\caption{Coverage rates realized under varying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}.
 The nominal confidence level was 95\% and for each setup, 10000 simulations were performed using simple random sampling.}

### 2018-11-11 2nd revision/alex_response.R

@@ -1,0 +1,157 @@
+# Editor: The editor does not understand why the authors were not able to reproduce the code examples given by the reviewer. Is there a bug, or just a misunderstanding between authors on one side and reviewer + editor on the other? Please check the code and explain.
+# 
+# Reviewer A:
+#   In the revised version of the manuscript (ID 3280), the authors include a new discussion on coverage rates of confidence intervals (p. 25), which, however, appears to only partially embrace the methodological concerns I raised in the earlier review. In particular, the statement that "the nominal coverage rate of 95% can [be] expected to be met for large sample sizes (e.g., n2=50)" is in conflict with examples 2 and 3 included from my review (n2=80, coverage rate ≤ 85%). I am not convinced by the authors' claim that they could not reproduce the examples.
+# 
+# On the other hand, the new discussion and in particular Figure 5 are very helpful in illustrating the issue of coverage rates, and I believe this will for most readers be sufficient to make them aware the related problems.
+# 
+# Based on my rather short reading of the manuscript and given that a formal review is not required at this stage (after previous recommendation "Revisions Required"), I would therefore agree that the manuscript can now be accepted.
+# 
+# #####
+# 
+# 
+# Response:
+# The reviewers code runs without bugs but the simulation was incorrectly formulated.  First of all, the three examples were posed in the finite population context rather than the infinite population context.  The package was written using estimators derived specifically in the infinite population approach.  Let's look at the first simulation example to see the fundamental problem with doing this:
+  
+  ###
+  # Reviewer's artificial data:
+  ###
+  set.seed(0)
+d <- data.frame(
+  elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
+  vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
+  vol2 = rnorm(e, mean=300, sd=100)*(e<310),
+  resp3 = runif(e) < 0.03
+)
+
+
[…]

### 2018-11-11 JSS_3280_resubmission/replication_rcode.R

@@ -1,0 +1,761 @@
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples                                                     #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
+
+library("forestinventory")
+
+# **************************************************************************** #
+# ------ SECTION 3: TWO-PHASE ESTIMATORS -------------------------------------
+# **************************************************************************** #
+
+
+# ----- 3.1 Global Estimations --------
+
+
+## load grisons data set:
+data("grisons", package = "forestinventory")
+
+## see structure of grisons:
+str(grisons)
+
+## see first 12 rows of grisons:
+head(grisons, n = 12)
+
+
+# **********************************************************
+## non-exhaustive (NEX) estimator with boundary weight adjustment
+reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                      data = grisons,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights")
+
+## get summary of estimation results:
+summary(reg2p_nex)
+
+
+# *****************************************************
+## exhaustive (EX) estimator with boundary weight adjustment
+## first, retrieve order of columnames that the lm()-function processes a formula object:
+#  --> actually the order they appear in the design-matrix ...
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+
+## store true means of explanatory variables:
+true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
+
+## apply estimator:
+reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                     data = grisons,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                     exhaustive = true.means.Z)
+
+## get summary of estimation results:
+summary(reg2p_ex)
+
[…]

### 2018-11-12 2nd revision/answer_to_editor.Rnw
update 2nd revision 12.11.2018
update 2nd revision 12.11.2018

\documentclass{article}
\usepackage{authblk}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[ngerman]{babel}
\usepackage[normalem]{ulem}

%============================================ Define Titlepage & packages =============================================%

\title{Answer to Editor}

\author{The R Package forestinventory: design-based global and small area estimations for multi-phase forest inventories\\
Andreas Hill, Alexander Massey}

\usepackage{fancyhdr}     
\usepackage{amsmath} %Paket für erweiterte math. Formeln
\usepackage[labelfont=bf]{caption}
\usepackage[font=footnotesize]{caption}
\usepackage[font=footnotesize]{subcaption}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[final]{pdfpages}
\usepackage{color}

\usepackage{geometry}
\geometry{
	a4paper,
	left=25mm,
	right=25mm,
	top=30mm,
	bottom=30mm
}

\setlength{\parindent}{0em} % Einzug bei neuen Absätzen

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Main Document------------------------------------------ %

\begin{document}

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Tex Settings ------------------------------------------ %

\maketitle
\thispagestyle{empty}
\newpage

\pagenumbering{arabic}
\setcounter{page}{1}

\pagestyle{fancy} %Kopfzeile und Fusszeile
\fancyfoot[C]{\thepage}
\setlength{\headsep}{15mm}

\definecolor{mybrown}{rgb}{0.6, 0.15, 0.1}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\definecolor{mygreen}{rgb}{0.1, 0.4, 0.4}
\newcommand{\answer}[1]{\small \color{mybrown}{#1} \color{black}}
\newcommand{\note}[1]{\textit{\small \color{amaranth} \textbf{Note:} #1} \color{black}}
\newcommand{\todo}[1]{\color{red}{#1} \color{black}}
\newcommand{\answerfin}[1]{\small \color{mygreen}{#1} \color{black}}
\newcommand{\code}[1]{\texttt{#1}}

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Review -----------------------------------------------------%

\section*{Feedback from Editor and Reviewer A}

Editor:\\
The editor does not understand why the authors were not able to reproduce the code examples given by the reviewer. Is there a bug, or just a misunderstanding between authors on one side and reviewer + editor on the other? Please check the code and explain.\\

Reviewer A:\\
In the revised version of the manuscript (ID 3280), the authors include a new discussion on coverage rates of confidence intervals (p. 25), which, however, appears to only partially embrace the methodological concerns I raised in the earlier review. In particular, the statement that 'the nominal coverage rate of 95\% can [be] expected to be met for large sample sizes (e.g., n2=50)' is in conflict with examples 2 and 3 included from my review (n2=80, coverage rate ≤ 85\%). I am not convinced by the authors' claim that they could not reproduce the examples. On the other hand, the new discussion and in particular Figure 5 are very helpful in illustrating the issue of coverage rates, and I believe this will for most readers be sufficient to make them aware the related problems. Based on my rather short reading of the manuscript and given that a formal review is not required at this stage (after previous recommendation 'Revisions Required'), I would therefore agree that the manuscript can now be accepted.

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Answer -----------------------------------------------------%

\section{Answer}

The reviewers code runs without bugs but the simulation was incorrectly formulated. One fundamental problem with the reviewers simulation is that he seems to pose all three examples in the \textit{finite} population context rather than the \textit{infinite} population context. We can see this right at the beginning:\\

Here, the reviewer sets up a dataframe that he then uses in his simulation examples:

\begin{small}
<<>>=
# Reviewer's artificial data:
set.seed(0)
d <- data.frame(
          elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
          vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
          vol2 = rnorm(e, mean=300, sd=100)*(e<310),
          resp3 = runif(e) < 0.03)
@
\end{small}

Let's look at the first simulation example to explain where the fundamental problem is:

\begin{small}
<<eval=FALSE>>=
library(forestinventory)
# PROBLEM 1: Small sample size (n2=5)
true_value <- mean(d$vol1)
cover <- replicate(10000, {
d$phase <- 1
d$phase[sample(length(d$phase), 5)] <- 2
ci <- confint( twophase( vol1 ~ elv, data = d,
phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
} )
mean(cover)
@
\end{small}

As obvious, the true population mean (\code{true\_value}) that has to be estimated is here calculated by taking the mean of a \textit{finite} number of response variable values ($N=500$, i.e. \code{nrow(d)}). This shows that the reviewer actually defined a finite population of the response variable (\code{vol1}), and he does the same for the auxiliary variable (\code{elv}). In his simulation, he then samples the set of observed response variable values $s_2$ (and also the auxiliary variable) from a finite population. This is a fundamental misunderstanding of the infinite population theory, which was exclusively used to derive all estimators contained in \code{forestinventory}. First of all, the true value should be an integral (i.e. spatial mean over infinite points, as described in section 2.1 in the article), but by using the finite mean, the reviewer has effectively acknowledged that actually $N=500$ instead of infinity. Second, we then see that every population unit is assigned to the first phase (\code{d\$phase<- 1}) indicating the intention to apply an \textit{exhaustive} two-phase estimator. Notice that in the \code{twophase()}-function call, the reviewer, however, did not to set the exhaustive parameter making this a non-exhaustive function call. This means that by simulation design $s_1$ does not get resampled with every replicate so the variability coming from resampling the first phase is ignored.\\

All 3 examples given by the reviewer made these same errors so the results clearly can't be considered conclusive on their own merits.  This is why we attempted to reformulate the simulation examples correctly in the infinite population context and mimic the examples intended effects. The following is some sample code (similar code was added to the Appendix in the last submission) to demonstrate simulation in the infinite population context.  This technique has already applied in Mandallaz (2013a), Mandallaz et al. (2013) and Massey et al. (2015):

\begin{small}
<<eval=FALSE>>=
library(forestinventory)
library(rmutil)  # used to calculate the integral in the spatial mean
#
#
set.seed(1)
#
# --- function to create density surface: --- #
target.surface <- function(x0, y0){
  local.density <- 30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)
  local.density
}
#
# --- True spatial mean for Global Surface Area: --- #
true_value <- int2(target.surface, a = c(0,0), b=c(3,2))/6 # 39.16667  #true mean
#
# --- sample generator for simple two-phase sampling: --- #
sample.generator<- function(n1, n2){

  realization <- matrix(NA,n1,6)
  for(i in 1:n1){
    x0=2*runif(1) # NOTE: The randomness comes from the uniform independent selection of points
    y0=3*runif(1)
    realization[i,1] <- target.surface(x0, y0)
    realization[i,2] <- x0
    realization[i,3] <- y0
    realization[i,4] <- x0*x0
    realization[i,5] <- x0*y0
    realization[i,6] <- y0*y0
  }
  realization <- as.data.frame(realization)
  names(realization) <- c("response","x","y","xx","xy","yy")
  realization$phase <- 1
  realization$phase[sample(nrow(realization), n2)] <- 2
  realization$response[realization$phase == 1] <- NA
  realization
}
@
\end{small}

As obvious, we take a finite random sample $s_2$ and $s_1$ from an infinite number of points. The randomness in the sampling procedure comes from the uniform independent selection of points (\code{x0, y0}) that are randomly distributed over the density surface defining the distribution of the target variable at every possible point in the plane.

\subsection{Reviewer Problem \#1}

\begin{small}
<<eval=FALSE>>=
# --- Run 10000 simulations with n1 = 500 and n2 = 5 (small sample size n2) --- #
n1 <- 500
n2 <- 5
num_sims <- 10000
ci.g.logical <- rep(NA, num_sims)   # two-phase coverage indicators for g-weight formula
ci.ext.logical <- rep(NA, num_sims) # two-phase coverage indicators for external formula
ci.op.logical <- rep(NA, num_sims)  # one-phase coverage indicators
#
for(j in 1: num_sims){
  realization <- sample.generator(n1 = n1, n2 = n2)
  est <- twophase(formula = response ~ y + x + xx, data=realization, 
                  phase_id = list(phase.col = "phase", terrgrid.id = 2))
  est_onephase <- onephase(formula = response ~ 1, data=realization, 
                           phase_id = list(phase.col = "phase", terrgrid.id = 2))
  ci <- confint(est)
  ci_op <- confint(est_onephase)
  ci.g.logical[j] <- ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
  ci.op.logical[j] <- ci_op$ci$ci_lower_op <= true_value & true_value <= ci_op$ci$ci_upper_op
}
#
# results for coverage rates:
mean(ci.g.logical)   # 0.6738, g-weight coverage rate of 2-phase estimator
mean(ci.op.logical)  # 0.9457, coverage rate of one-phase estimator
@
\end{small}

The one-phase estimator has the correct coverage rate even with sample size 5 (it would also have it for $n_2$ = 2) but it will probably have a very large margin of variance (i.e. estimation error). The g-weight variance has lower than nominal coverage rates because the contribution of the variance by calculating the beta vector is based on approximations (i.e. a first order taylor approximation for the g-weight) which are not adequate for this sample size.  The article has already included a graphic with sample code that addresses the sample size issue and the reviewer has already accepted it so we can leave this issue alone for the time being.

\subsection{Reviewer Problem \#2 and \#3}

The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values and non-zero where the auxiliary information yields a low value.  Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain is at elevations where no trees can grow.  This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers. However, a quick look at the reported R-squared, which would be very close to zero, should trigger the  the question why the model is included in the first place.\\

Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables. This means that there is per-definition no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information at all. Nevertheless, such a situation may occur if say the inventorist wanted to count how many trees are in the desert where there are just a handful of uniformly distributed oases.  Even in this case, the nonzero plots will again be highly influential leverage points in the presence of so many zeros. This issue is easily seen by low R-squared values as well as by a look at the residuals of the linear model.\\

These two examples are also pathological for the one-phase estimator because a significant proportion of the samples drawn will have zeros for ALL of s2 making the estimate = 0 and the confidence interval just a point with no chance of covering the true value.  If this occurred in practice the sample would immediately be considered suspect.  It is hard to imagine a situation where this would accidentally happen.  Furthermore, trying to measure forest volume in areas where 98\% of the land has no trees is very unlikely and would be difficult to accidentally do.

\end{document}

### 2018-11-12 2nd revision/answer_to_editor.tex
update 2nd revision 12.11.2018
update 2nd revision 12.11.2018

--- 2018-11-12 2nd revision/answer_to_editor.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
@@
 25), which, however, appears to only partially embrace the methodological concerns I raised in the earlier review.
-In particular, the statement that 'the nominal coverage rate of 95\% can [be] expected to be met for large sample sizes (e.g., n2=50)' is in conflict with examples 2 and 3 included from my review (n2=80, coverage rate ≤ 85\%).
+In particular, the statement that 'the nominal coverage rate of 95\% can [be] expected to be met for large sample sizes (e.g., n2=50)' is in conflict with examples 2 and 3 included from my review (n2=80, coverage rate = 85\%).
 I am not convinced by the authors' claim that they could not reproduce the examples.
@@
 \begin{small}
-<<>>=
-# Reviewer's artificial data:
-set.seed(0)
-d <- data.frame(
-          elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
-          vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
-          vol2 = rnorm(e, mean=300, sd=100)*(e<310),
-          resp3 = runif(e) < 0.03)
-@
+\begin{Schunk}
+\begin{Sinput}
+> # Reviewer's artificial data:
+> set.seed(0)
+> d <- data.frame(
++           elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
++           vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
++           vol2 = rnorm(e, mean=300, sd=100)*(e<310),
++           resp3 = runif(e) < 0.03)
+\end{Sinput}
+\end{Schunk}
 \end{small}
@@
 \begin{small}
-<<eval=FALSE>>=
-library(forestinventory)
-# PROBLEM 1: Small sample size (n2=5)
-true_value <- mean(d$vol1)
-cover <- replicate(10000, {
-d$phase <- 1
-d$phase[sample(length(d$phase), 5)] <- 2
-ci <- confint( twophase( vol1 ~ elv, data = d,
-phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
-ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
-} )
-mean(cover)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library(forestinventory)
+> # PROBLEM 1: Small sample size (n2=5)
+> true_value <- mean(d$vol1)
+> cover <- replicate(10000, {
++ d$phase <- 1
++ d$phase[sample(length(d$phase), 5)] <- 2
++ ci <- confint( twophase( vol1 ~ elv, data = d,
++ phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
++ ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
++ } )
+> mean(cover)
+\end{Sinput}
+\end{Schunk}
 \end{small}
@@
 \begin{small}
-<<eval=FALSE>>=
-library(forestinventory)
-library(rmutil)  # used to calculate the integral in the spatial mean
-#
-#
-set.seed(1)
-#
-# --- function to create density surface: --- #
-target.surface <- function(x0, y0){
-  local.density <- 30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)
-  local.density
-}
-#
-# --- True spatial mean for Global Surface Area: --- #
-true_value <- int2(target.surface, a = c(0,0), b=c(3,2))/6 # 39.16667  #true mean
-#
-# --- sample generator for simple two-phase sampling: --- #
-sample.generator<- function(n1, n2){
-
-  realization <- matrix(NA,n1,6)
-  for(i in 1:n1){
-    x0=2*runif(1) # NOTE: The randomness comes from the uniform independent selection of points
-    y0=3*runif(1)
-    realization[i,1] <- target.surface(x0, y0)
-    realization[i,2] <- x0
-    realization[i,3] <- y0
-    realization[i,4] <- x0*x0
-    realization[i,5] <- x0*y0
-    realization[i,6] <- y0*y0
-  }
-  realization <- as.data.frame(realization)
-  names(realization) <- c("response","x","y","xx","xy","yy")
-  realization$phase <- 1
-  realization$phase[sample(nrow(realization), n2)] <- 2
-  realization$response[realization$phase == 1] <- NA
-  realization
-}
-@
+\begin{Schunk}
+\begin{Sinput}
+> library(forestinventory)
+> library(rmutil)  # used to calculate the integral in the spatial mean
+> #
+> #
+> set.seed(1)
+> #
+> # --- function to create density surface: --- #
+> target.surface <- function(x0, y0){
++   local.density <- 30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)
++   local.density
++ }
+> #
+> # --- True spatial mean for Global Surface Area: --- #
+> true_value <- int2(target.surface, a = c(0,0), b=c(3,2))/6 # 39.16667  #true mean
+> #
+> # --- sample generator for simple two-phase sampling: --- #
+> sample.generator<- function(n1, n2){
++   
++   realization <- matrix(NA,n1,6)
++   for(i in 1:n1){
++     x0=2*runif(1) # NOTE: The randomness comes from the uniform independent selection of points
++     y0=3*runif(1)
++     realization[i,1] <- target.surface(x0, y0)
++     realization[i,2] <- x0
++     realization[i,3] <- y0
++     realization[i,4] <- x0*x0
++     realization[i,5] <- x0*y0
++     realization[i,6] <- y0*y0
++   }
++   realization <- as.data.frame(realization)
++   names(realization) <- c("response","x","y","xx","xy","yy")
++   realization$phase <- 1
++   realization$phase[sample(nrow(realization), n2)] <- 2
++   realization$response[realization$phase == 1] <- NA
++   realization
++ }
+\end{Sinput}
+\end{Schunk}
 \end{small}
@@
 \begin{small}
-<<eval=FALSE>>=
-# --- Run 10000 simulations with n1 = 500 and n2 = 5 (small sample size n2) --- #
-n1 <- 500
-n2 <- 5
-num_sims <- 10000
-ci.g.logical <- rep(NA, num_sims)   # two-phase coverage indicators for g-weight formula
-ci.ext.logical <- rep(NA, num_sims) # two-phase coverage indicators for external formula
-ci.op.logical <- rep(NA, num_sims)  # one-phase coverage indicators
-#
-for(j in 1: num_sims){
-  realization <- sample.generator(n1 = n1, n2 = n2)
-  est <- twophase(formula = response ~ y + x + xx, data=realization, 
-                  phase_id = list(phase.col = "phase", terrgrid.id = 2))
-  est_onephase <- onephase(formula = response ~ 1, data=realization, 
-                           phase_id = list(phase.col = "phase", terrgrid.id = 2))
-  ci <- confint(est)
-  ci_op <- confint(est_onephase)
-  ci.g.logical[j] <- ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
-  ci.op.logical[j] <- ci_op$ci$ci_lower_op <= true_value & true_value <= ci_op$ci$ci_upper_op
-}
-#
-# results for coverage rates:
-mean(ci.g.logical)   # 0.6738, g-weight coverage rate of 2-phase estimator
-mean(ci.op.logical)  # 0.9457, coverage rate of one-phase estimator
-@
+\begin{Schunk}
+\begin{Sinput}
+> # --- Run 10000 simulations with n1 = 500 and n2 = 5 (small sample size n2) --- #
+> n1 <- 500
+> n2 <- 5
+> num_sims <- 10000
+> ci.g.logical <- rep(NA, num_sims)   # two-phase coverage indicators for g-weight formula
+> ci.ext.logical <- rep(NA, num_sims) # two-phase coverage indicators for external formula
+> ci.op.logical <- rep(NA, num_sims)  # one-phase coverage indicators
+> #
+> for(j in 1: num_sims){
++   realization <- sample.generator(n1 = n1, n2 = n2)
++   est <- twophase(formula = response ~ y + x + xx, data=realization, 
++                   phase_id = list(phase.col = "phase", terrgrid.id = 2))
++   est_onephase <- onephase(formula = response ~ 1, data=realization, 
++                            phase_id = list(phase.col = "phase", terrgrid.id = 2))
++   ci <- confint(est)
++   ci_op <- confint(est_onephase)
++   ci.g.logical[j] <- ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
++   ci.op.logical[j] <- ci_op$ci$ci_lower_op <= true_value & true_value <= ci_op$ci$ci_upper_op
++ }
+> #
+> # results for coverage rates:
+> mean(ci.g.logical)   # 0.6738, g-weight coverage rate of 2-phase estimator
+> mean(ci.op.logical)  # 0.9457, coverage rate of one-phase estimator
+\end{Sinput}
+\end{Schunk}
 \end{small}

### 2018-11-12 rnw/Calculation of Confidence Intervals.Rnw
update 2nd revision 12.11.2018
update 2nd revision 12.11.2018

@@
 \centering
-\resizebox{0.80\hsize}{!}{\includegraphics{fig/globalsim.png}}
-\caption{Coverage rates realized under varying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}.
-The nominal confidence level was 95\% and for each setup, 10000 simulations were performed using simple random sampling.}
+\resizebox{0.78\hsize}{!}{\includegraphics{fig/globalsim.png}}
+\caption{Coverage rates of the global two-phase estimator realized under varying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}.
+The confidence intervals were based on the g-weight variance with a nominal confidence level of 95\%.
+For each setup, 10000 simulations were performed using simple random sampling.}
 \label{fig:covrate_sim}

### 2018-11-15 2nd revision/answer_to_editor.Rnw
update 15_11_2018 2nd revision
update 15_11_2018 2nd revision

@@
 
-\section{Answer}
+\section*{Answer}
 
@@
 
-Here, the reviewer sets up a dataframe that he then uses in his simulation examples:
+Here, the reviewer sets up a dataframe defining a \textit{finite} population that he then uses in his simulation examples:
 
@@
 \code{nrow(d)}).
-This shows that the reviewer actually defined a finite population of the response variable (\code{vol1}), and he does the same for the auxiliary variable (\code{elv}).
-In his simulation, he then samples the set of observed response variable values $s_2$ (and also the auxiliary variable) from a finite population.
+This shows that the reviewer actually defined a finite population for the response variable (\code{vol1}), and likewise for the auxiliary variable (\code{elv}).
+In every iteration of his simulation, he then resamples from a finite set of observed response variable values $s_2$.
 This is a fundamental misunderstanding of the infinite population theory, which was exclusively used to derive all estimators contained in \code{forestinventory}.
-First of all, the true value should be an integral (i.e.
-spatial mean over infinite points, as described in section 2.1 in the article), but by using the finite mean, the reviewer has effectively acknowledged that actually $N=500$ instead of infinity.
-Second, we then see that every population unit is assigned to the first phase (\code{d\$phase<- 1}) indicating the intention to apply an \textit{exhaustive} two-phase estimator.
-Notice that in the \code{twophase()}-function call, the reviewer, however, did not to set the exhaustive parameter making this a non-exhaustive function call.
+First of all, the true population mean (\code{true\_value}) should be an integral (i.e.
+spatial mean) over infinite points, as described in section 2.1 in the article, but by using the finite mean, the reviewer has effectively acknowledged that actually $N=500$ instead of infinity.\par
+Second, we then see that every population unit of his finite population is assigned to the first phase (\code{d\$phase<- 1}).
+Basically, this means that the auxiliary information is not resampled in every simulation iteration, but the entire population ($N=500$) of auxiliary information \code{elv} is used.
+This indicates that the reviewer intended to apply an \textit{exhaustive} two-phase estimator where the true population mean of the auxiliary information is used.
+Notice that in the \code{twophase()}-function call, the reviewer, however, did \textit{not} set the exhaustive parameter as demonstrated in section 3.1 pp.
+11-12 in the article, making this a non-exhaustive function call.
 This means that by simulation design $s_1$ does not get resampled with every replicate so the variability coming from resampling the first phase is ignored.\\
@@
 
-As obvious, we take a finite random sample $s_2$ and $s_1$ from an infinite number of points.
-The randomness in the sampling procedure comes from the uniform independent selection of points (\code{x0, y0}) that are randomly distributed over the density surface defining the distribution of the target variable at every possible point in the plane.
+As obvious, we take a finite random sample $s_2$ and $s_1$ from an infinite number of points (note that the number of possible points, i.e., coordinate pairs \code{x0, y0}) as input for the density function \code{target.surface} is infinite).
+The randomness in the sampling procedure comes from the uniform independent selection of points that are randomly distributed over the density surface defining the distribution of the target variable at every possible point in the plane.
 
-\subsection{Reviewer Problem \#1}
+\subsection*{Reviewer Problem \#1}
 
@@
 
-\subsection{Reviewer Problem \#2 and \#3}
+\subsection*{Reviewer Problem \#2 and \#3}
 
@@
 This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers.
-However, a quick look at the reported R-squared, which would be very close to zero, should trigger the  the question why the model is included in the first place.\\
+However, a quick look at the reported R-squared, which would be very close to zero, should immediately trigger the  question why the model is included in the first place.\\
 

### 2018-11-15 JStatSoft_forestinventory_master-concordance.tex
update 15_11_2018 2nd revision
update 15_11_2018 2nd revision

@@
 1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
-3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 22 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1956:%
+3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1945:%
 171 11 1}

### 2018-12-30 2nd revision/answer_to_editor.Rnw
update 30_12_2018
update 30_12_2018 (version berfore 2nd revsion)

@@
 
-The reviewers code runs without bugs but the simulation was incorrectly formulated.
-One fundamental problem with the reviewers simulation is that he seems to pose all three examples in the \textit{finite} population context rather than the \textit{infinite} population context.
+\textbf{The reviewers code runs without bugs but the simulation was incorrectly formulated.} One fundamental problem with the reviewers simulation is that he seems to pose all three examples in the \textit{finite} population context rather than the \textit{infinite} population context.
 We can see this right at the beginning:\\
@@
 In every iteration of his simulation, he then resamples from a finite set of observed response variable values $s_2$.
-This is a fundamental misunderstanding of the infinite population theory, which was exclusively used to derive all estimators contained in \code{forestinventory}.
+\textbf{This is a fundamental misunderstanding of the infinite population theory, which was exclusively used to derive all estimators contained in \code{forestinventory}}.
 First of all, the true population mean (\code{true\_value}) should be an integral (i.e.
@@
 
-All 3 examples given by the reviewer made these same errors so the results clearly can't be considered conclusive on their own merits.
-This is why we attempted to reformulate the simulation examples correctly in the infinite population context and mimic the examples intended effects.
-The following is some sample code (similar code was added to the Appendix in the last submission) to demonstrate simulation in the infinite population context.
+\textbf{All 3 examples given by the reviewer made these same errors so the results clearly can't be considered conclusive on their own merits.
+This is why we attempted to reformulate the simulation examples correctly in the infinite population context and mimic the examples intended effects.} The following is some sample code (similar code was added to the Appendix in the last submission) to demonstrate simulation in the infinite population context.
 This technique has already applied in Mandallaz (2013a), Mandallaz et al.
@@
 estimation error).
-The g-weight variance has lower than nominal coverage rates because the contribution of the variance by calculating the beta vector is based on approximations (i.e.
+However, the g-weight variance has lower than nominal coverage rates because the contribution of the variance by calculating the beta vector is based on approximations (i.e.
 a first order taylor approximation for the g-weight) which are not adequate for this sample size.
-The article has already included a graphic with sample code that addresses the sample size issue and the reviewer has already accepted it so we can leave this issue alone for the time being.
+\textbf{This particular issue has been addressed in the added paragraph about coverage rates in section 5 "Calculation of confidence intervals", including a graphic, the simulation code in the Appendix as well as a hint for the reader that in such cases of very small sample sizes, one should prefer to apply the small area estimators which ensure the nominal coverage rates even under such limited sample sizes.
+The reviewer has allready aggreed with this solution.}
 
@@
 
-The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values and non-zero where the auxiliary information yields a low value.
-Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain is at elevations where no trees can grow.
-This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers.
+The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values (i.e., for his population) and non-zero where the auxiliary information yields a low value.
+Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain (here 98\%) is at elevations where no trees can actually grow.
+He calls this 'Effect of explanatory variable'.
+As described in Mandallaz (2014), it is somewhat common in practice to sample information over an entire country, i.e.
+also at places outside forest even including areas where no trees can grow.
+This is caused by the fact that in order to calculate totals (by multiplication of the estimated density, e.g.
+volume per hectare, with the area value of the inventory domain), the area of a country is well known, whereas the area of the forest is often not known precisely.
+In the mentioned article it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros (i.e.
+places where the target-value is sampled outside forest).
+It has been mathematically shown that by this procedure, the R$^2$ of the regression model is overoptimistic as the zeros are easy to predict.
+This is e.g.
+the case if - just as in the reviewer's example - the occurence of zeros depend on a known explanatory variable like the elevation.
+Furthermore, this procedure is inefficient as it inflates the estimated variance, and a suggested solution in Mandallaz (2014) is to filter out most of the non-forested areas where one can certainly assume that no forest exists (this was consequently done in Hill et al.
+2018).\par
+
+However, the reviewer's example goes beyond that topic by contructing an artifical example where \textit{only 2\%} of the entire population (i.e.
+also appr.
+2\% of the inventory area) consist of forest/trees.
+It is extremely hard to imagine that such an inventory would actually be conducted by anyone - unless there is no information at all about the environment where the forest inventory is going to be conducted (i.e., wether it is a desert with only few oasis, an alpine mountain side above the tree level, ...).
+This is hard to imagine, especially in times where aerial photographs and further geodata are availabe and used in the planing process of every forest inventory.
+Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
+This means that now there is per-definition even no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information in the first place.
+Nevertheless, such a situation may occur if say the inventorist wanted to count how many trees are in the desert where there are just a handful of uniformly distributed oases, and the auxiliary information does not provide any information about their location.
+These two examples are also pathological for the one-phase estimator because a significant proportion of the samples drawn will have zeros for ALL of $s_2$ making the estimate = 0 and the confidence interval just a point with no chance of covering the true value.
+If this occurred in practice the sample would immediately be considered suspect.
+It is hard to imagine a situation where this would accidentally happen.
+Furthermore, trying to measure forest volume in areas where 98\% of the land has no trees is very difficult to accidentally do.
+Concluding, \textbf{the fact that the estimators did not produce the nominal coverage rates under this artificial setup is clearly not dedicated to a limitation of the estimators, but to conducting a totally pathological inventory that would never come into anyone's mind in practice.}\\
+
+This being said, it is impossible and has clearly not been our intention to address any imaginable misuse of the implemented estimators in our article.
+However, we want to emphasize that we already addressed edge-cases that can - from our experience -  expected to actually occur in practice in section 6 'Special cases and scenarios'.
+\textbf{We now have also added a short paragraph \textit{'Restricting the inventory domain to forested area'} to section 3.1 in order to address the issue of 'sampling outside forest', as mentioned above.}
+
+\subsection*{Final Remark}
+
+Since it's first release in July 2016, the package has been downloaded over 6000 times (1000 downloads within the last 4 month).
+Considering that the field of forest inventory addresses a rather small user community compared to other R-packages, we think that these numbers support the importance and the added-value of the package to the field.
+It also underlines the unique characteristic of our package among the existing set of R-Packages.
+The development of the package as well as the writing of the article was accompanied by many feedbacks and discussions with external experts in the field, and many of their questions and criticisms have been taken up in the article.
+We are convinced that the revision process helped to further improve the article.
+After carefully checking the reviewers' questions and concerns, we addressed as much of their requests as possible without altering the general objective of our article.\\
+
+We hope that the editor can agree with our suggestions, and would be happy if the article can now be published in the \textit{Journal of Statistical Software}.\\
+
+Yours sincerely,\\
+
+Andreas Hill \& Alexander Massey
+
+\begin{small}
+<<eval=FALSE, echo=FALSE>>=
+true_value <- mean(d$vol2)
+cover <- replicate(10000, {
+d$phase <- 1
+truemean<- c(1,mean(d$elv)) 
+  d$phase[sample(length(d$phase), 80)] <- 2
+  ci <- confint( twophase( vol2 ~ elv, data = d,
+                           phase_id = list(phase.col="phase", terrgrid.id=2),
+                           exhaustive =truemean), 
+                           level=0.95)
+  ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
+
+ r2<-summary(lm( vol2 ~ elv, data = d[d$phase==2,]))$r.squared
+
+ d$phase <- 1
+ d$phase[sample(length(d$phase), 80)] <- 2
+ plot(vol2 ~ elv, data = d[d$phase==2,])
+ summary(mod<- lm( vol2 ~ elv, data = d[d$phase==2,]))
+ abline(mod, col="red")
+
+} )
+mean(cover)
+mean(r2)
+@
+\end{small}
+
+% This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers.
 However, a quick look at the reported R-squared, which would be very close to zero, should immediately trigger the  question why the model is included in the first place.\\
 
-Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
-This means that there is per-definition no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information at all.
+% Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
+This means that there is per-definition no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information in the first place.
 Nevertheless, such a situation may occur if say the inventorist wanted to count how many trees are in the desert where there are just a handful of uniformly distributed oases.
@@
 
-These two examples are also pathological for the one-phase estimator because a significant proportion of the samples drawn will have zeros for ALL of s2 making the estimate = 0 and the confidence interval just a point with no chance of covering the true value.
+% These two examples are also pathological for the one-phase estimator because a significant proportion of the samples drawn will have zeros for ALL of s2 making the estimate = 0 and the confidence interval just a point with no chance of covering the true value.
 If this occurred in practice the sample would immediately be considered suspect.
@@
 
+%Considering this being said, it has clearly not been our intention to address any imaginable pathological inventory case in our article.
+However, we specifically addressed edge-cases that can - from our experience -  expected to actually occur in section 6 "Special cases and scenarios".
+
 \end{document}

### 2018-12-30 2nd revision/answer_to_editor.tex
update 30_12_2018
update 30_12_2018 (version berfore 2nd revsion)

@@
 \begin{document}
+\input{answer_to_editor-concordance}
 
@@
 
-The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values and non-zero where the auxiliary information yields a low value.
-Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain is at elevations where no trees can grow.
-This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers.
+The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values (i.e., for his population) and non-zero where the auxiliary information yields a low value.
+Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain (here 98\%) is at elevations where no trees can actually grow.
+He calls this 'Effect of explanatory variable'.
+As described in Mandallaz (2014), it is somewhat common in practice to sample information over an entire country, i.e.
+also at places outside forest even including areas where no trees can grow.
+This is caused by the fact that in order to calculate totals (by multiplication of the estimated density, e.g.
+volume per hectare, with the area value of the inventory domain), the area of a country is well known, whereas the area of the forest is often not known precisely.
+In the mentioned article it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros (i.e.
+places where the target-value is sampled outside forest).
+It has been mathematically shown that by this procedure, the R$^2$ of the regression model is overoptimistic as the zeros are easy to predict.
+This is e.g.
+the case if - just as in the reviewer's example - the occurence of zeros depend on a known explanatory variable like the elevation.
+Furthermore, this procedure is inefficient as it inflates the estimated variance, and a suggested solution in Mandallaz (2014) is to filter out most of the non-forested areas where one can certainly assume that no forest exists (this was consequently done in Hill et al.
+2018).\par
+
+However, the reviewer's example goes beyond that topic by contructing an artifical example where \textit{only 2\%} of the entire population (i.e.
+also appr.
+2\% of the inventory area) consist of forest/trees.
+It is extremely hard to imagine that such an inventory would actually be conducted by anyone - unless there is no information at all about the environment where the forest inventory is going to be conducted (i.e., wether it is a desert with only few oasis, an alpine mountain side above the tree level, ...).
+This is hard to imagine, especially in times where aerial photographs and further geodata are availabe and used in the planing process of every forest inventory.
+Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
+This means that now there is per-definition even no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information in the first place.
+Nevertheless, such a situation may occur if say the inventorist wanted to count how many trees are in the desert where there are just a handful of uniformly distributed oases, and the auxiliary information does not provide any information about their location.
+These two examples are also pathological for the one-phase estimator because a significant proportion of the samples drawn will have zeros for ALL of $s_2$ making the estimate = 0 and the confidence interval just a point with no chance of covering the true value.
+If this occurred in practice the sample would immediately be considered suspect.
+It is hard to imagine a situation where this would accidentally happen.
+Furthermore, trying to measure forest volume in areas where 98\% of the land has no trees is very difficult to accidentally do.
+Concluding, \textbf{the fact that the estimators did not produce the nominal coverage rates under this artificial setup is clearly not dedicated to a limitation of the estimators, but to conducting a totally pathological inventory that would never come into anyone's mind in practice.}\\
+
+This being said, it is impossible and has clearly not been our intention to address any imaginable misuse of the implemented estimators in our article.
+However, we want to emphasize that we already addressed edge-cases that can - from our experience -  expected to actually occur in practice in section 6 'Special cases and scenarios'.
+\textbf{We now have also added a short paragraph \textit{'Restricting the inventory domain to forested area'} to section 3.1 in order to address the issue of 'sampling outside forest', as mentioned above.}
+
+\subsection*{Final Remark}
+
+Since it's first release in July 2016, the package has been downloaded over 6000 times (1000 downloads within the last 4 month).
+Considering that the field of forest inventory addresses a rather small user community compared to other R-packages, we think that these numbers support the importance and the added-value of the package to the field.
+It also underlines the unique characteristic of our package among the existing set of R-Packages.
+The development of the package as well as the writing of the article was accompanied by many feedbacks and discussions with external experts in the field, and many of their questions and criticisms have been taken up in the article.
+We are convinced that the revision process helped to further improve the article.
+After carefully checking the reviewers' questions and concerns, we addressed as much of their requests as possible without altering the general objective of our article.\\
+
+We hope that the editor can agree with our suggestions, and would be happy if the article can now be published in the \textit{Journal of Statistical Software}.\\
+
+Yours sincerely,\\
+
+Andreas Hill \& Alexander Massey
+
+\begin{small}
+\end{small}
+
+% This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers.
 However, a quick look at the reported R-squared, which would be very close to zero, should immediately trigger the  question why the model is included in the first place.\\
 
-Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
-This means that there is per-definition no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information at all.
+% Problem 3 is actually the exact same example except that the response variable takes exclusively the value \code{1} at non-zero plots and these non-zero plots (and thus plots with response variable value=1) are uniformly distributed among the auxiliary variables.
+This means that there is per-definition no causal correlation between the auxiliary variable values and the response value, which arises the question why one should use the auxiliary information in the first place.
 Nevertheless, such a situation may occur if say the inventorist wanted to count how many trees are in the desert where there are just a handful of uniformly distributed oases.

### 2018-12-30 JStatSoft_forestinventory_master-concordance.tex
update 30_12_2018
update 30_12_2018 (version berfore 2nd revsion)

--- 2018-08-30 JStatSoft_forestinventory_master-concordance.tex
@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
-1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
+1 2 1 1 4 114 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
 0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
@@
 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 824:%
 133}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 825:%
 1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
@@
 0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1176:%
 135 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
-1 2 1 1 4 38 1 1 2 11 0 1 2 3 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1181:%
+1 2 1 1 4 38 1 1 2 11 0 1 2 78 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1314:%
 141 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1237:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1319:%
 1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
@@
 0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1472:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1554:%
 147 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1477:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1559:%
 1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
@@
 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1666:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
 153 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1671:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
 1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1683:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
 159 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1688:%
-1 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1693:%
-165 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
+1 6 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1777:%
+165 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1782:%
+1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
+3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1950:%
+171 11 1}

### 2018-12-30 rnw/Calculation of Confidence Intervals.Rnw
update 30_12_2018
update 30_12_2018 (version berfore 2nd revsion)

@@
 For further details on the simulation, the reader is referred to the original articles mentioned above.
-The results can also be reproduced or expanded using the \proglang{R}-code given in the Appendix.\par
-
-From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for large sample sizes (e.g., $n_2$=50) regardless of the chosen sample size $n_1$.
-The realized coverage rate only drops significantly below 90\% if considerably small sampling fractions (i.e., the ratio ${n_2}/{n_1}$) < 0.1 are used that lead to small terrestrial sample sizes of 20 and less.
-The simulation examples also confirmed that in such cases, one should whenever possible consider one of the design-unbiased small area estimators (P)SMALL or EXT(P)SYNTH for which the nominal coverage rates in the simulations were already met for sample sizes $n_{2,G}$ > 5.
+The results can also be reproduced or expanded using the \proglang{R}-code given in the Appendix.
+From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for sample sizes $n_2$ of 50 and higher regardless of the chosen sample size $n_1$.
+The realized coverage rate only drops significantly below 90\% if considerably small sampling fractions (i.e., the ratio ${n_2}/{n_1}$) < 0.1 are used that lead to small terrestrial sample sizes $n_2$ of 20 and less.
+Re-evaluating the same simulations for small area estimation confirmed that in such cases, one should whenever possible consider one of the design-unbiased small area estimators (P)SMALL or EXT(P)SYNTH for which the nominal coverage rates in the simulations were already met for sample sizes $n_{2,G}$ > 5.
 Finally, it should also be noted that applying the design-based estimators under systematic sampling will lead to an inflation of the estimated variance \citep{mandallaz1993}.

### 2018-12-30 rnw/Twophase_Estimators and their Application.Rnw
update 30_12_2018
update 30_12_2018 (version berfore 2nd revsion)

@@
 \end{equation}
+
+% --------------------------- %
+\subsubsection{Restricting the inventory domain to forested area}
+
+In some national forest inventories it has been common to sample information over the entire country, i.e., also at places outside forest even including areas such as high mountain sides where no trees can actually grow.
+This is caused by the fact that in order to calculate totals by multiplying the estimated density (e.g., timber volume per hectare) with the area value of the inventory domain, the area of a country is well known, whereas the precise area of the forest is often much harder to determine.
+In \citet{mandallaz2014} it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros for the target variable (i.e., at sample locations outside forest).
+It has been mathematically shown that by this procedure, the coefficient of determination (R$^2$) of the regression model can be overoptimistic as the zeros are easy to predict.
+This is particularly the case if the occurence of zeros correlate with an explanatory variable, such as the elevation above sea level.
+Furthermore, \citet{mandallaz2014} showed that this procedure is inefficient as it inflates the estimated variance.
+In order to mitigate this effect, a solution is to filter out most of the non-forested areas where one can certainly assume that no forest exists.
 

### 2018-12-30 2nd revision/problem no 3 analysed.R

@@ -1,0 +1,59 @@
+
+# Reviewer's problem No. 3:
+
+crit.height<- 320
+
+set.seed(0)
+d <- data.frame(
+  elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
+  vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
+  #vol2 = rnorm(e, mean=300, sd=100)*(e<310),
+  vol2 = rnorm(e, mean=300, sd=100)*(e<crit.height),
+  resp3 = runif(e) < 0.03)
+
+
+# repeat manually to see what happens to data and model R2:
+d$phase <- 1
+d$phase[sample(length(d$phase), 80)] <- 2
+plot(vol2 ~ elv, data = d[d$phase==2,])
+print("mod1 (red)", quote=F);summary(mod1<- lm( vol2 ~ elv, data = d[d$phase==2, ]))$r.squared # fit without masking zeros (thesis: higher R2 than mod2)
+print("mod2 (green)", quote=F);summary(mod2<- lm( vol2 ~ elv, data = d[d$phase==2 & d$elv< crit.height, ]))$r.squared  # fit after masking zeros
+abline(mod1, col="red")
+abline(mod2, col="green")
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2019-04-27 Final Edit/jss3280_manuscript/jss3280.tex
formatting changes for jss

\documentclass[article]{jss}
%\documentclass[nojss]{jss} % for creating the vignette

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Andreas Hill \And Alexander Massey}
\title{The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Andreas Hill, Alexander Massey} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
\Shorttitle{\pkg{forestinventory}: Design-Based Global and Small Area Estimations in \proglang{R}} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by field crews. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of multiphase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}

\Keywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}
\Plainkeywords{forest inventory, design-based, infinite population approach, two- and three-phase sampling, regression estimators, small area estimation}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Andreas Hill\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{andreas.hill@forestinventory.de}\\

  \noindent Alexander Massey\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{afmass@gmail.com}\\}

%% for those who use Sweave please include the following line (with % symbols):
%\usepackage{Sweave}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Tex Settings ------------------------------------------ %

% used packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathptmx} 
\usepackage{latexsym}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{flafter}
\usepackage{bm}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{geometry}
 \geometry{
 a4paper,
 left=28mm,
 bottom=28mm
 }

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\var}{\mathbb{V}}
\newcommand{\EX}{\mathbb{E}}

\setlength{\parindent}{0em} % Einzug bei neuen Absätzen

%------------------------------------------------------------------------------------------------%
% -------------------------------------- Main Document------------------------------------------ %

\begin{document}\sloppy
% \input{JStatSoft_forestinventory_master-concordance}

%------------------------------------------------------------------------------------------------%
% -------------------------------------- R Settings -------------------------------------------- %

% set global options for R-output:

% execute everything that goes without echo:

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Introduction ---------------------------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Introduction}
\label{sec:intro}

In many countries, forest inventories have become an indispensable tool for evaluating the current state of forests as well as for tracking their development over time. They provide accurate quantitative information that can be used to define management actions, to adapt forest management strategies according to guidelines on national and international levels and to support in monitoring carbon change in the context of climate reporting and mitigation. As conducting a full census of all trees within any sizable forest area is clearly infeasible due to time and cost restrictions, forest inventories usually gather their information by means of statistical sampling methods. Typically this means that discrete sample locations (sample plots) are randomly chosen in the forest, making up the framework of a terrestrial inventory.  This terrestrial sample data is then used to make estimates for the entire forested area and provide a measure of precision for those estimates in the form of confidence intervals. There is a broad range of literature describing the concepts and methods regarding the choice of different estimators under various sample designs \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}.

Terrestrial inventories have the benefit of being very flexible in the sense that they can be used to produce high quality estimates for a wide-variety of different forest attributes. However, they have the downside of being very expensive.  Improving the precision of the estimates by increasing the number of sample plots essentially means that travel costs will rise as trained inventorists are sent to more and more plot locations. This is why the number of terrestrial samples is often limited. Although national inventories usually possess a sufficiently large terrestrial sample size to provide high estimation accuracies for larger areas, this is often not the case for smaller areas, such as forest management units.  As a result, there has been an increasing need for alternative inventory methods that can maintain the same estimation precision at lower costs, or achieve higher estimation precision at identical costs \citep{vonluepke2013}. A method which has become particularly attractive is called multiphase sampling. The core concept is to enlarge the sample size in order to gain higher estimation precision without enlarging the terrestrial sample size. This is done by using predictions of the terrestrial target variable at additional sample locations where the terrestrial information has not been gathered. These predictions are produced by regression models that use explanatory variables derived from auxiliary
data, commonly in the form of spatially exhaustive remote sensing data in the inventory area. Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLuebke2012}. Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par

% section: briefly give a review of existing R-packages for forest inventories (not much around)
% While commercial enterprises have already been offering multiphase inventory analyses, freely available software for application of multiphase estimators has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multiphase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par
% 
% The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management. The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

Multistage and multiphase estimation has already been implemented in commercial as well as open-source software, such as the survey sampling procedures in \proglang{SAS} \citep{sas} and the \pkg{survey} package in \proglang{R} \citep{survey}. However, both are targeted towards list-sampling as it is applied in official statistics. Available software providing multiphase sampling methods better suited for forest inventories has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the \pkg{maSAE} package by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2018b}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

 \newpage

%------------------------------------------------------------------------------------------------%
% --------------------------- Methods and structure of the Package ----------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Methods and structure of the package}
\label{sec:str_and_mod}

\subsection{Forest inventory in the infinite population approach}

Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by coordinates. The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g., timber volumes) to local plot densities (e.g., the timber density in m$^3$/ha) including potential boundary adjustments (see \cite{mandallaz2008} for details). The estimators implemented in \pkg{forestinventory} use the so-called infinite population approach in order to bridge this inventory process to the mathematics behind the estimators. This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig. \ref{fig:inf_pop_apr}. The population total of the target variable (e.g., the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig. \ref{fig:inf_pop_apr} as the volume under the density surface. From this perspective, the practical challenge is that the form of this function is unknown. Theoretically, we could get the total timber volume by observing the function value, i.e., the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum. However, this is impossible because there is an infinite number of points in the forest area. Our strategy is thus to take a sample of $n_2$ points, $s_2$, from an infinite population of possible points and use their associated local densities $Y(x)$ to estimate the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.  The total timber volume can then be obtained by multiplying $\hat{Y}$ by the known surface area of the forest, $\lambda(F)$. It should be noted $\lambda(F)$ is often estimated in practice (see \cite{mandallaz2014}) for further explanation) but that is beyond the scope of this article. All estimators included in \pkg{forestinventory} rest upon this theoretical perspective. The key point in the infinite population approach is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area. Sampling the actual plot areas follows the so-called finite population approach, where the sample units are usually assumed to be either circular or rectangular plots.  Although choosing between infinite and finite population schemas does not likely make much practical difference, the former does has some theoretical advantages.  This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape. Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area.  Also, cluster sampling is much more easily generalized in the infinite population approach. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

\begin{figure}[htb]
\centering
\resizebox{0.42\hsize}{!}{\includegraphics{fig/localdensity_final-01.eps}}
\caption{Artificial representation of a local density surface. The spatial distribution of a hypothetical density function for every point in a forested area is represented as a wavy piecewise constant green surface.  Sample plots (white dots) inform the inventorist of the value of the density function at that point. Note that the plateaus of constant $Y(x)$ values here have the shape of squares whereas in reality they are likely to be formed by the intersection of circles around trees.}
\label{fig:inf_pop_apr}
\end{figure}

\subsection{Two-phase sampling}

The two-phase or double-sampling estimators use inventory information from two nested samples which are commonly referred to as phases (Fig. \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e., the local density $Y(x)$ (e.g., the timber volume per hectare). For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some model, which in \pkg{forestinventory} is always a linear model fit in $s_2$ using ordinary least squares (OLS). The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed. In the design-based context, the two-phase estimator is typically unbiased regardless of the model used to produce the predictions. This property comes from the assumption that each phase's sample is selected via simple random sampling (see Section \ref{sec:db_vs_md}).

\subsection{Three-phase sampling}

Three-phase estimators extend the principle of two-phase sampling and use inventory information from three nested samples (phases) (Fig. \ref{fig:concmphase_and_sae_a}). The basic setup is that the explanatory variables calculated from the auxiliary information are available in two different frequencies. The phase $s_0$ provides a large number $n_0$ of auxiliary data, whereas the phase $s_1$ provides additional auxiliary data that are only available at $n_1$ subsamples of $s_0$. The terrestrial information is then collected at a further subsample $s_2$ of $s_1$. The motivation for three-phase sampling is that the additional set of explanatory variables available at $s_1$, now denoted $\pmb{Z}^{(1)}(x)$, adds considerable explanatory power to the set of variables available at all sample locations $x \in s_0$, denoted $\pmb{Z}^{(0)}(x)$. From that it follows that we can define two nested regression models. The full set of predictor variables $\pmb{Z}^{\top}(x)=(\pmb{Z}^{(0){\top}}(x),\pmb{Z}^{(1){\top}}(x))$ can be used to calculate the predictions $\hat{Y}(x)$ of $Y(x)$ at all sample locations $x \in s_1$. The regression model applicable to the $s_1$ phase is thus referred to as the full model. Less accurate predictions, $\hat{Y}^{(0)}(x)$, can be produced at all the sample locations $x \in s_0$ using only the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$. If there is a significant gain in model precision between the reduced and the full model and the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{(a) Concept of multiphase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate available information. (b) Illustration of the small area estimation problem.}
\label{fig:concmphase_and_sae}
\end{figure}

\subsection{Small area estimation}

Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample. To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig. \ref{fig:concmphase_and_sae_b}). As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multiphase estimation can be efficient. However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals. The idea is then to borrow strength from the entire terrestrial sample $s_2$ of $F$ to fit the model, and to apply this model to the small area. The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area. If there are no terrestrial plots in $G$ (i.e., $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator. These are called synthetic estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.

\subsection{Design-based vs. model-dependent approach}
\label{sec:db_vs_md}

The subject of model selection gets a lot of attention in the field of forest inventory. This is why it is important to understand that the mathematical interpretation of how a model is used to produce estimates is fundamentally different between the design-based and model-dependent approach. In the model-dependent (also known as model-based) framework, the sample locations $x$ are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the forest is assumed to be the realization of a stochastic process. Although the model does not need to be fit from a probability sample, i.e., the sample locations could arbitrarily be chosen, the model should adequately describe the underlying stochastic process in order to efficiently ensure unbiased results. In practice this means that special attention must be made to ensure that the variable selection is appropriate to avoid overfitting, important variables are not omitted and all model assumptions are reasonably met through empirical verification. If a model is misspecified then estimation based on inference from that model may not be reliable. In the model-dependent framework one thus has to trust the model. In contrast, the design-based approach, on which all \pkg{forestinventory} estimators are based, rests upon the randomization of the sample locations $x$. While the sample locations $x$ are independently and uniformly distributed in the forest, the forest itself and thus the values of the local density surface at any location $x \in F$ are fixed and not the result of a stochastic process. A selected observation $Y(x)$ still remains a random variable, but solely due to the random sample mechanism. A consequence of this approach is that the estimation properties of design-based regression estimators (e.g., unbiasedness) typically hold regardless of the model that is chosen. The philosophy of the design-based approach is thus to use prediction models to improve the efficiency of the estimators without having to rely on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as model-assisted. It should be noted that the randomization of sample locations upon which design-based inference depends, is in practice often replaced by systematic grids to achieve a more spatially balanced sample in the terrestrial survey. However, there is reasonable evidence that softening this assumption is acceptable for point and variance estimation as long as the grid does not interact with periodic features in the forest structure \citep{mandallaz2008}. The variance will in most cases be slightly overestimated and lead to wider, more conservative confidence intervals \citep{mandallaz2013a}.

%\newpage

\subsection{Package structure}
\label{sec:packstruc}

In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig. \ref{fig:struct_package}). Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications. The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (global estimators) or only for subregions of the entire inventory area (small area estimators). In the second case, the package offers three small area estimators that will in detail be described in the following sections. The estimators are available under exhaustive and non-exhaustive use of the auxiliary data. Additionally, the package can also calculate one-phase estimates solely based on terrestrial samples. All estimators are also available for cluster sampling, in which case a sample unit consists of multiple, spatially agglomerated samples. The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict discussion to simple sampling, while the formulas for cluster sampling are available in the technical reports \citep{mandallaz2016, mandallaz2013techa, mandallaz2013techb}. A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.

\begin{figure}[htb]
\centering
\resizebox{0.9\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
\caption{Structure of the multiphase estimators in the \proglang{R} package \pkg{forestinventory}. All estimators are also available for cluster sampling.}
\label{fig:struct_package}
\end{figure}

\newpage

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Estimators and their Application -------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:

\section{Two-phase estimators and their application}
\label{sec:twophase_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Global estimators}
\label{sec:twophase_globest}

\subsubsection{Mathematical background}

The vector of regression coefficients of the OLS regression model is found by using the following solution to the sample-based normal equation:

\begin{equation}\label{normequ_simple}
  \hat{\pmb{\beta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
\end{equation}

The individual predictions can then be calculated as $\hat{Y}(x)=\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Unless stated otherwise, \pkg{forestinventory} only uses internal models to calculate estimates. This means that the model fit, i.e., $\hat{\pmb{\beta}}_{s_2}$, is derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} functions.  While virtually all inventorists fit their models using the current inventory data, sometimes there is reason to use formulas derived from external models where the sample used to train the model is assumed to be taken from an independent source \citep{massey2015a}. However, this usually occurs when using a model other than the OLS regression model and is beyond the scope of the package at this time.\par

The package provides the calculation of point estimates under exhaustive (EX) and non-exhaustive (NEX) use of the auxiliary information, which means to respectively apply $\hat{\pmb{\beta}}_{s_2}$ to $\bar{\pmb{Z}}$, i.e., the exact spatial mean of $\pmb{Z}(x)$, or to $\hat{\bar{\pmb{Z}}}$, i.e., an estimate of the spatial mean of $\pmb{Z}(x)$:

\begin{subequations}\label{eq:pointest_2p_reg}
\begin{align}
  \hat{Y}_{reg2p,EX} & =\bar{\pmb{Z}}^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
  \hat{Y}_{reg2p,NEX} & =\hat{\bar{\pmb{Z}}}^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
\end{align}
\end{subequations}

Note that for internal linear models containing intercept terms, the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (zero mean residual property) which is why it does not appear in the point estimate. More explanation about how to obain the auxiliary means is given in the next subsection.

The \pkg{forestinventory} package implements two kinds of variances for each of these point estimates: the g-weight formulation that accounts for the fact that our model is in fact internal, and the external variance formulation that assumes a true external regression model and thus neglects the uncertainty in the regression coefficients \citep{mandallaz2016}.

The g-weight formulation is

\begin{subequations}\label{eq:gw_var_2p_reg}
\begin{align}
  \hat{\var}(\hat{Y}_{reg2p,EX}) & :=\bar{\pmb{Z}}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}} \label{eq:gw_var_2p_reg_ex}\\
  \hat{\var}(\hat{Y}_{reg2p,NEX}) & :=
  \hat{\bar{\pmb{Z}}}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
\end{align}
\end{subequations}

where the g-weight variance-covariance matrix of $\hat{\pmb{\beta}}_{s_2}$ is calculated as

\begin{equation}\label{eq:estvarmatrix}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^{\top}(x)\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
\end{equation}

and the uncertainty caused by using the $s_1$ sample to estimate $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is accounted for by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}=
\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}})^{\top}
\end{equation}

The external variance formulation for linear regression models is

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{reg2p,EX}) & = \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}_{ext}(\hat{Y}_{reg2p,NEX}) & = \frac{1}{n_1}\hat{\var}_{s_1}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \nonumber  \label{eq:varexternal_2p_reg_nex}
\end{align}
\end{subequations}
where $\hat{\var}_{s_2}$ and $\hat{\var}_{s_1}$ indicate taking the sample variance over $s_2$ and $s_1$ respectively.

Note that when applied to internal linear regression models, the external variance is asymptotically unbiased and usually slightly smaller than the g-weight variance, where the uncertainty of the regression coefficients is accounted for by the variance-covariance matrix (Eq. \ref{eq:estvarmatrix}).  The external variances are provided in the package \pkg{forestinventory} in case the user wants to compare linear models to another model type where no g-weight formulation is possible, as is the case with non-parametric models like kNN.

\subsubsection{Calculation of explanatory variables}

We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive cases. Fig. \ref{fig:exh_nexh_and_boundweights_b} depicts how the non-exhaustive case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles) and densified to a larger sample $s_1$ (the points). For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ is calculated using a defined spatial extent of auxiliary information around that point called the support (the dark green square tiles). We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the sample point whereas the support is the spatial extent of the auxiliary information used to calculate those values. So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points. The \pkg{forestinventory} package calculates the empirical mean of $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par

The exhaustive case requires a closer look. In the infinite population approach, $\pmb{Z}(x)$ refers to the sample point $x$ and not the area around it. Deriving the exact spatial mean, $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}(x)dx= (\frac{1}{\lambda(F)}\int_{F} z_1(x)dx, ..., \frac{1}{\lambda(F)}\int_{F} z_p(x)dx)^{\top}$, implies that we need to calculate the spatial mean of each component of $\pmb{Z}(x)$ using all possible points in $F$. This is much like the situation we had with calculating the mean of the local density surface for $Y(x)$ in that we need to find the mean of $\pmb{Z}(x)$ over an infinite number of sample points (i.e., $n_1=\infty$). Although it is practically infeasible to assess $\pmb{Z}(x)$ for every $x$, there are few cases where the exact mean can in fact be precisely calculated. The first case is when the explanatory variables are provided by polygon layers (e.g., map of development stages). In this case, one can calculate the exact mean as the area-weighted average of each categorical variable. The second case is when the exact mean can be calculated in one step, e.g., taking the mean of all height pixels of a raster canopy height model will exactly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}. However, for most types of explanatory variables, e.g. order statistics such as maximum or median canopy height in the spatial extent, this approach will fail.  For example, the mean of the maximum pixel heights across all supports is clearly not equal to the global maximum pixel height across the entire area.  In such cases, we will try to get an approximation of $\bar{\pmb{Z}}$ that is only negligibly different. \par

One implementation to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig. \ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited. Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the finite population approach, i.e., deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units. While in the infinite population approach this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large. However, the perfect tessellation implementation can also impose drawbacks. One is that a perfect tessellation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size. Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model \citep{hill2018} a perfect tessellation might often not be feasible. In the infinite population frame, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$. With this respect, the infinite population approach provides more flexibility than the finite approach.

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of (a) exhaustive and (b) non-exhaustive calculation of explanatory variables including boundary adjustment at the support level. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon).}
\label{fig:exh_nexh_bw}
\end{figure}

\subsubsection{Boundary adjustment}

An extension to the so-far published estimators by Mandallaz is the consideration of a boundary adjustment. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g., by a polygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (see Fig. \ref{fig:exh_nexh_bw}). This method was suggested in \citet{mandallaz2013b} and led to an improvement in estimation precision. In order to ensure an unbiased calculation of either $\hat{\bar{\pmb{Z}}}$ or $\bar{\pmb{Z}}$, the respective means have then to be calculated as the weighted mean (Eq. \ref{eq:wmeanZ}) where the weight $w(x)$ is equal to the percentage of forested area within the support of sample location $x$.

\begin{equation}\label{eq:wmeanZ}
  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_1}}w(x)}
\end{equation}

% --------------------------- %
\subsubsection{Restricting the inventory domain to forested area}

In some national forest inventories it has been common to sample information over the entire country, i.e., also at places outside forest even including areas such as high mountain sides where no trees can actually grow. This is caused by the fact that in order to calculate totals by multiplying the estimated density (e.g., timber volume per hectare) with the area value of the inventory domain, the area of a country is well known, whereas the precise area of the forest is often much harder to determine. In \citet{mandallaz2014} it is demonstrated that by doing so, one usually inflates the sample by a large number of zeros for the target variable (i.e., at sample locations outside forest). It has been mathematically shown that by this procedure, the coefficient of determination (R$^2$) of the regression model can be overoptimistic as the zeros are easy to predict. This is particularly the case if the occurence of zeros correlates with an explanatory variable, such as the elevation above sea level. Furthermore, \citet{mandallaz2014} showed that this procedure is inefficient as it inflates the estimated variance. In order to mitigate this effect, a solution is to filter out most of the non-forested areas where one can certainly assume that no forest exists.

% --------------------------- %
\subsubsection{Application}

To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data set that comes with installing the package from the CRAN repository. The data set contains data from a simple (i.e., non-cluster) two-phase forest inventory conducted in 2007 that was used in \citet{mandallaz2013b} as a case study. The $s_1$ sample is comprised of 306 sample locations arranged on a systematic grid containing auxiliary information in the form of airborne laserscanning (LiDAR) canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}). For a systematic subsample of 67 ($s_2$ sample), terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey. We can load \pkg{forestinventory} and examine the \code{grisons} data set in the \proglang{R} environment as follows:

\begin{Schunk}
\begin{Sinput}
R> library("forestinventory")
R> data("grisons", package = "forestinventory")
R> head(grisons)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Soutput}
   phase_id_2p boundary_weights  mean stddev   max   q75 smallarea   tvol
1            2             1.00  9.30  11.84 40.87 21.14         C 107.80
2            1             1.00 12.16  11.35 39.80 21.54         A     NA
3            2             1.00  5.25   5.74 23.82  9.53         D  63.77
4            1             1.00  7.53   9.33 34.10 13.02         A     NA
5            1             0.67  6.11   5.87 23.33 10.55         B     NA
6            1             1.00 12.15  10.16 33.76 20.97         C     NA
7            2             1.00  6.38   4.72 17.96 10.14         D 154.10
8            1             1.00  1.25   3.79 22.72  0.00         B     NA
9            1             1.00 21.56   7.49 32.66 27.81         A     NA
10           2             1.00 13.55   7.20 36.14 18.59         A 256.15
\end{Soutput}
\end{Schunk}

Estimates can be made using the \code{onephase()}, \code{twophase()} or \code{threephase()} functions. The data frame inputted to these functions must have the structure where each row corresponds to a unique sample location and the columns specify the attributes associated to that respective sample location. Attributes that are missing, e.g., because they are associated with sample locations that were not selected in the subsample for the subsequent phase, should be designated as \code{NA} and the phase membership is encoded as numeric.

For global two-phase estimation, we have to specify

\begin{itemize}
  \itemsep0em
  \item the regression model (\code{formula}) as specified in the \code{lm}()-function \citep{R}.
  \item the inputted \code{data.frame} containing the inventory information (\code{data}).
  \item the \code{list}-object \code{phase_id} containing: the \code{phase.col} argument identifying the name of the column specifying membership to $s_1$ or $s_2$, and the \code{terrgrid.id} argument specifying which numeric value indicates $s_2$ membership in that column. Note that \pkg{forestinventory} implicitly assumes that all rows not indcated as $s_2$ belong to the $s_1$ phase.
  \item the name of the column containing the weights $w(x)$ of the boundary adjustments (optional).
\end{itemize}

The non-exhaustive estimator with boundary weight adjustment can thus be applied as follows:

\begin{Schunk}
\begin{Sinput}
R> reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}. A readable summary of the estimation results can be obtained by passing this object to the \code{summary()} function, which automatically interprets what type of estimator was used and returns pertinent information such as the regression model formula, the point estimate (\code{estimate}), the g-weight and external variance (\code{g_variance} and \code{ext_variance}) as well as the sample sizes and the model R$^2$:

\begin{Schunk}
\begin{Sinput}
R> summary(reg2p_nex)
\end{Sinput}
\begin{Soutput}
Two-Phase global estimation

Call: 
twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2), 
    boundary_weights = "boundary_weights")

Method used:
Non-exhaustive global estimator

Regression Model:
tvol ~ mean + stddev + max + q75

Estimation results:
 estimate ext_variance g_variance  n1 n2 r.squared
 383.5354      279.954   271.5057 306 67 0.6428771

'boundary_weight'- option was used to calculate weighted means of auxiliary variables
\end{Soutput}
\end{Schunk}

For practical use, one should normally always prefer the g-weight variance over the external variance. This is because when we use internal models, the regression coefficients actually depend on the terrestrial sample realized by the sampling design. In contrast to the external variance, the g-weight variance accounts for this sampling variability which results in more reliable point and variance estimates and also enjoys better statistical calibration properties (g-weights). The external and g-weight variances are asymptotically equivalent but the external variance is really only included here in case the user wants to compare to another estimator where no g-weight variance exists.

The exhaustive estimator can be applied by additionally passing a vector containing the exact means of the explanatory variables, i.e., $\bar{\pmb{Z}}$, to the optional argument \code{exhaustive}. This vector must be calculated beforehand in such a way that any desired boundary adjustment has already been applied. Note that the vector input to \code{exhaustive} must be in the same order that the \code{lm()}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1. Particular caution must be taken if categorical variables are present because the \code{lm()}-function, which is internally used to set up the design-matrix, automatically creates dummy variables with one of the categories used as a reference. Using our \code{grisons} example, the correct order can always be extracted by the following \proglang{R}-code:

\begin{Schunk}
\begin{Sinput}
R> colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+    x = TRUE)$x)
\end{Sinput}
\end{Schunk}

The exhaustive estimator can be applied after defining the vector of exact means $\bar{\pmb{Z}}$ taken from \citet{mandallaz2013b}, denoted as \code{true.means.Z}:

\begin{Schunk}
\begin{Sinput}
R> true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
R> reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+    data = grisons,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    exhaustive = true.means.Z)
\end{Sinput}
\end{Schunk}

An alternative way to look at the estimation results without using the \code{summary()} is to query \code{reg2p_ex} directly:

\begin{Schunk}
\begin{Sinput}
R> reg2p_ex$estimation
\end{Sinput}
\begin{Soutput}
  estimate ext_variance g_variance  n1 n2 r.squared
1 376.7426     202.5602   187.2787 Inf 67 0.6428771
\end{Soutput}
\end{Schunk}

Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is essentially because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

%--------------------------------------------------------------------------------------------------%
% ################################################################################################ %
%--------------------------------------------------------------------------------------------------%

\subsection{Small area estimators}
\label{sec:twophase_sae}

\subsubsection{Mathematical background}

The \pkg{forestinventory} package provides three types of small area estimators each of which has an exhaustive and non-exhaustive form. We will use a different nomenclature for the non-exhaustive case in small area estimation since much of the existing literature shows preference for the label pseudo to indicate that the mean of the explanatory variables within the small area was based on a finite sample. The main idea for all these small area estimators is to calculate the regression coefficient vector $\hat{\pmb{\beta}}_{s_2}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ on the entire $s_2$ sample according to Eq. \ref{normequ_simple} and \ref{eq:estvarmatrix}, and subsequently use that to make predictions for sample locations restricted to small area $G$.\par

%----------------------------------------------- %
% \textbf{Small and Pseudo Small Area Estimator}\par

We first introduce the small area estimator (SMALL), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the pseudo small area estimator (PSMALL).

\begin{subequations}\label{eq:pest_2p_small_psmall}
\begin{align}
  \hat{Y}_{G,SMALL,2p} & =\bar{\pmb{Z}}_G^{\top}\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_small} \\
  \hat{Y}_{G,PSMALL,2p} & =\hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pepsmall}
\end{align}
\end{subequations}

In the equations for the point estimates (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$. A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data also outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$. This is called the bias or residual correction term.

The package provides the g-weight variance for SMALL and PSMALL respectively (Eq. \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) as well as the  external variance (Eq. \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}). Again note that all components are restricted to those available at the sample locations in the small area ($s_{1,G}$ and $s_{2,G}$), with exception of the regression coefficient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.

\begin{subequations}\label{eq:var_2p_small_psmall}
\begin{align}
  \hat{\var}(\hat{Y}_{G,SMALL,2p}) & := \bar{\pmb{Z}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G
    + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_small} \\
  \hat{\var}(\hat{Y}_{G,PSMALL,2p}) & := \hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}
  + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:var_2p_reg_psmall}
\end{align}
\end{subequations}

\begin{subequations}\label{eq:varext_2p_small_psmall}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,SMALL,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
  \hat{\var}_{ext}(\hat{Y}_{G,PSMALL,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
\end{align}
\end{subequations}
where $\hat{\var}_{s_{2,G}}$ indicates taking the sample variance over $s_{2,G}$. If boundary adjustment is applied, the simple mean of the explanatory variable vector over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.

%------------------------------------------------------------- %
% \textbf{Synthetic and Pseudo Synthetic Estimator}\par

The synthetic estimator (SYNTH) and pseudo synthetic estimator (PSYNTH) are commonly applied when no terrestrial sample is available within the small area $G$ (i.e., $n_{2,G}=0$). In this case, the point estimates (Eq. \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based only on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively. However, the bias correction using the observed residuals $\hat{R}(x)$ is not applied as was the case in the small and pseudo small area estimator (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}). Thus, the (pseudo) synthetic estimator has a potentially unobservable design-based bias. Also note that the residual variation can no longer be considered in the g-weight variance (Eq. \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}). Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias. Due to the absence of available residuals in $G$, there is also no external variance form for the synthetic and pseudo synthetic estimator.

% Left out for simplicity reasons in article, maybe include in Diss:
% However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$.

\begin{subequations}\label{eq:pest_2p_synth_psynth}
\begin{align}
  \hat{Y}_{G,SYNTH,2p} & =\bar{\pmb{Z}}_G^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
  \hat{Y}_{G,PSYNTH,2p} & =\hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth} \\
  \hat{\var}(\hat{Y}_{G,SYNTH,2p}) & = \bar{\pmb{Z}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
  \hat{\var}(\hat{Y}_{G,PSYNTH,2p}) & = \hat{\bar{\pmb{Z}}}_G^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
\end{align}
\end{subequations}

where the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
\begin{equation}\label{estvarcovaux_G}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
\frac{1}{n_{1,G}(n_{1,G}-1)}\sum_{x\in{s_{1,G}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^{\top}
\end{equation}

%------------------------------------------------------------- %
% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par

The synthetic estimators, SYNTH and PSYNTH, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively optimistic. The SMALL and PSMALL estimators overcome this issue by using a bias correction term, i.e., $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$. The motivation behind the extended synthetic and extended pseudo synthetic estimator (EXTSYNTH and EXTPSYNTH) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while ensuring that the mean of the empirical prediction model residuals in the entire area $F$ and the small area $G$ are by construction both zero at the same time. This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0. Recalling that linear models fitted using OLS have zero mean residual property by construction also if categorical variables are used, this leads to unbiased point estimates. The new extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{Z}^{\top}(x),I^{\top}_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^{\top}(x)$ whereever it is used in Eq. \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}. Note that the package functions internally extend the data set by the indicator variable if the EXTSYNTH or EXTPSYNTH estimator is called.

Not every equation needs to be re-written here, but to give an example of the notational change, the regression coefficient under extended model approach becomes

\begin{equation}\label{ext_normequ_simple}
  \hat{\pmb{\theta}}_{s_2}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{\mathbb{Z}}(x)\pmb{\mathbb{Z}}^{\top}(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{\mathbb{Z}}(x)\Big)
\end{equation}

The point estimates and their g-weight variances can then be re-written as

\begin{subequations}\label{eq:pest_2p_extsynth_extpsynth}
\begin{align}
\hat{Y}_{G,EXTSYNTH,2p} & = \bar{\pmb{\mathbb{Z}}}^{\top}_{G}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
\hat{Y}_{G,EXTPSYNTH,2p} & =\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^{\top}\hat{\pmb{\theta}}_{s_2} \label{eq:pointest_2p_extsynth} \\
\hat{\var}(\hat{Y}_{G,EXTSYNTH,2p}) & = \bar{\pmb{\mathbb{Z}}}^{\top}_{G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}\bar{\pmb{\mathbb{Z}}}_{G} \label{eq:var_2p_extsynth} \\
\hat{\var}(\hat{Y}_{G,EXTPSYNTH,2p})& =
\hat{\bar{\pmb{\mathbb{Z}}}}_{G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\theta}}_{s_2}}
\hat{\bar{\pmb{\mathbb{Z}}}}_{G}
+ \hat{\pmb{\theta}}^{\top}_{s_2}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{\mathbb{Z}}}}_{G}}\hat{\pmb{\theta}}_{s_2} \label{eq:var_2p_extpsynth}
\end{align}
\end{subequations}

While the formulas look similar to the synthetic estimators, note that a decomposition of $\hat{\pmb{\theta}}_{s_2}$ reveals that the residual correction term is now included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016} and thus the estimates are asymptotically design-unbiased.

The package also provides the external variance for both the extended synthetic and extended pseudo synthetic estimator. Note that neither the extended model approach nor external variance estimates are possible in the absence of terrestrial samples and thus model residuals in $G$, which is precisely when one must rely on the (pseudo) synthetic estimates. The external variance forms of EXTSYNTH and EXTPSYNTH are

\begin{subequations}\label{eq:ext_varexternal_2p_extsynth}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,EXTSYNTH,2p}) & = \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extsynth} \\
  \hat{\var}_{ext}(\hat{Y}_{G,EXTPSYNTH,2p}) & = \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extpsynth}
\end{align}
\end{subequations}
where $\hat{\mathbb{R}}(x)$ are the empirical residuals under the extended auxiliary vector.

To summarize, the synthetic estimators SYNTH and PSYNTH can be applied whether terrestrial inventory sample is found in the small area or not, but has a deceptively small g-weight variance due to its potential bias.  When terrestrial sample is observed in the small area, we can produce (asymptotically) design-unbiased estimates and variances using either SMALL or PSMALL which remove this bias explicitly with a mean residual term, or more elegantly with EXTSYNTH or EXTPSYNTH which simply use the same synthetic formulas while including an indicator variable for the small area in the model formula to remove the bias by construction in OLS.

%------------------------------------------------------------- %
\subsubsection{Application}

Small area estimates in the \pkg{forestinventory} package can be applied by specifying the optional argument \code{small_area}. The input data set has to include an additional column of class \code{factor} that describes the small area membership of the sample location represented by that row. The argument \code{small_area} requires a \code{list}-object that comprises

\begin{itemize}
  \itemsep0em
  \item the name of the column specifiying the small area of each observation (\code{sa.col}).
  \item a vector specifying the small area(s) for which estimations are desired (\code{areas}).
  \item the argument \code{unbiased} that controls which of the three available estimators is applied.
\end{itemize}

In order to apply the pseudo small area estimator (PSMALL) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:

\begin{Schunk}
\begin{Sinput}
R> psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+    unbiased = TRUE), psmall = TRUE, boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> summary(psmall_2p)
\end{Sinput}
\begin{Soutput}
Two-phase small area estimation

Call: 
twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2), 
    small_area = list(sa.col = "smallarea", areas = c("A", "B"), 
        unbiased = TRUE), boundary_weights = "boundary_weights", 
    psmall = TRUE)

Method used:
Pseudo small area estimator

Regression Model:
tvol ~ mean + stddev + max + q75

Estimation results:
 area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
    A 393.9713     1009.034   1308.117 306 67  94  19 0.6428771
    B 419.6416     1214.035   1259.472 306 67  81  17 0.6428771

'boundary_weight'- option was used to calculate weighted means of auxiliary variables
\end{Soutput}
\end{Schunk}

The small area functions all return an \code{S3} object of class \code{"twophase"} with subclass \code{"smallarea"}. In addition to global estimation, the \code{estimation} object now comprises the estimates and variances for all small areas (column \code{area}). We can view the sample sizes by looking into the object itself

\begin{Schunk}
\begin{Sinput}
R> psmall_2p$samplesizes
\end{Sinput}
\begin{Soutput}
$A
      n1G n2G  n1 n2
plots  94  19 306 67

$B
      n1G n2G  n1 n2
plots  81  17 306 67
\end{Soutput}
\end{Schunk}

The extended pseudo synthetic estimator (EXTPSYNTH) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:

\begin{Schunk}
\begin{Sinput}
R> extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+    unbiased = TRUE), boundary_weights = "boundary_weights")
R> extpsynth_2p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
1    A 391.9356     995.5602   1017.633 306 67  94  19 0.6526503
2    B 419.7231    1214.6053   1019.191 306 67  81  17 0.6428854
\end{Soutput}
\end{Schunk}

The \pkg{forestinventory} package automatically includes the indicator variable for the small area behind the scenes so there is no need for the user to implement it. Notice that the $R^2$s (\code{r.squared}) under the EXTPSYNTH estimator vary between the small areas, while they are identical under the PSMALL estimator. This is because under the EXTPSYNTH estimator, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrix. In case of the PSMALL estimator, the regression model stays the same for each small area estimation. Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling (see Section \ref{sec:speccas_and_scen}).\par

Setting the argument \code{unbiased=FALSE} applies the pseudo synthetic estimator to the selected small areas. Note that in the \code{grisons} data set, all small areas possess much more than the suggested minimum number of terrestrial observations (a rule of thumb is that $n_{2,G} \geq 6$) required to produce reliable design-unbiased estimates. Hence, choosing to use PSYNTH is probably not desireable and is just applied here for demonstration purposes. In this case the residual correction will not be applied.

\begin{Schunk}
\begin{Sinput}
R> psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"),
+    unbiased = FALSE), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> psynth_2p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
1    A 421.8863           NA   546.8651 306 67  94  19 0.6428771
2    B 418.7399           NA   566.3361 306 67  81  17 0.6428771
\end{Soutput}
\end{Schunk}

We see here that the PSYNTH variances are almost only half the variances of the PSMALL and EXTPSYNTH estimator. However, PSMALL and EXTPSYNTH are design unbiased and their variances reflect the fact that they account for potential bias of the regression model predictions. The g-weight variance of PSYNTH completely neglects a potential bias and as a result risks severely overstating the estimation precision.\par

The exhaustive versions of the small area estimators (Eq. \ref{eq:pointest_2p_small}, \ref{eq:var_2p_reg_small}, \ref{eq:varext_2p_reg_small}, \ref{eq:pointest_2p_reg_synth}, \ref{eq:var_2p_reg_synth}) are specified via the optional argument \code{exhaustive}. Its application requires that we know the exact means of all explanatory variables within the small area(s) of interest. In contrast to the global estimators, the exact means have now to be delivered in the form of a \code{data.frame}, where each row corresponds to a small area, and each column specifies the exact mean of the respective explanatory variable. Note that likewise the case of global estimation, the order of the explanatory variables in the data frame has to match the order in which they appear in the design matrix defined by the \code{lm()}-function in \proglang{R}. In order to tell \proglang{R} which row describes which small area, the row names have to match the respective names of the small areas specified in the \code{areas} argument.

For the \code{grisons} data set, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by

\begin{Schunk}
\begin{Sinput}
R> colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, 
+    x = TRUE)$x)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> true.means.Z.G <- data.frame(Intercept = rep(1, 4),
+    mean = c(12.85, 12.21, 9.33, 10.45),
+    stddev = c(9.31, 9.47, 7.90, 8.36),
+    max = c(34.92, 35.36, 28.81, 30.22),
+    q75 = c(19.77, 19.16, 15.40, 16.91))
R> rownames(true.means.Z.G) <- c("A", "B", "C", "D")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> true.means.Z.G
\end{Sinput}
\begin{Soutput}
  Intercept  mean stddev   max   q75
A         1 12.85   9.31 34.92 19.77
B         1 12.21   9.47 35.36 19.16
C         1  9.33   7.90 28.81 15.40
D         1 10.45   8.36 30.22 16.91
\end{Soutput}
\end{Schunk}

The extended synthetic estimator (EXTSYNTH) can then be applied by

\begin{Schunk}
\begin{Sinput}
R> extsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col ="smallarea", areas = c("A", "B"),
+    unbiased = TRUE), exhaustive = true.means.Z.G)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extsynth_2p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
1    A 372.6930     744.3658   696.5739 Inf 67 Inf  19 0.6526503
2    B 387.5116     693.8576   708.1105 Inf 67 Inf  17 0.6428854
\end{Soutput}
\end{Schunk}

Just as in the global case, we see that the variance has again been significantly decreased by substituting the exact auxiliary means and both first phase sample sizes are now infinity. Note that the function extracts the required exact means for small area \code{"A"} and \code{"B"} from the complete set of exact means defined in \code{true.means.Z.G}.

\newpage

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:

\section{Three-phase estimators and their application}
\label{sec:threephase_and_appl}

% ---------------------------------------------------------------------------- %
% ---------------------------------------------------------------------------- %
\subsection{Global estimators}
\label{sec:glob_est_3p}

% ------------------------------------- %
\subsubsection{Mathematical background}

Solving the sample-based normal equations, the vector of regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ for the reduced model, i.e., using the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$ available at $x \in s_0$, and likewise the vector of regression coefficients $\hat{\pmb{\beta}}_{s_2}$ for the full model, i.e., using the full set of explanatory variables $\pmb{Z}^{\top}(x)=(\pmb{Z}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$ available only at a subset $x \in s_1 \subset s_0$, are derived as

\begin{subequations}\label{eq:normequ_3p}
\begin{align}
\hat{\pmb{\alpha}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)  \label{eq:normequ_redmod} \\
\hat{\pmb{\beta}}_{s_2}&=\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^{\top}(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x) \label{eq:normequ_fullmod}
\end{align}
\end{subequations}

The package allows for the calculation of point estimates under exhaustive and non-exhaustive use of the auxiliary information in the $s_0$ phase. Fitting the model using $s_2$ (i.e., internally) ensures the zero mean residual property over $s_2$.

\begin{subequations}\label{eq:reg3p}
\begin{align}
\hat{Y}_{reg3p,EX}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
\hat{Y}_{reg3p,NEX}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2}  +
\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
\end{align}
\end{subequations}

Intuitively, the three phase estimator is simply the mean of the predictions using the reduced model (i.e. $\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}$), corrected by the mean difference between the reduced model predictions and the more accurate full model predictions (i.e. $\frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2})$), corrected by the mean difference between the ground truth and the full model predictions (i.e. $\frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})$). For the compact version of the formula in the non-exhaustive case, the estimated means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the estimated mean of $\pmb{Z}(x)$ over the $s_1$ phase are calculated according to Eq. \ref{meanvalues3pnex}. If the exact mean over $s_0$ is known, the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ can simply be replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$. Note that in case of applied boundary adjustment (Section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to Eq. \ref{eq:wmeanZ}.

\begin{equation}\label{meanvalues3pnex}
\hat{\bar{\pmb{Z}}}^{(0)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(0)}(x) ,
\quad \hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}(x)
\end{equation}

The package again provides the g-weight and external variances. The g-weight variance formulation is

\begin{subequations}\label{eq:dbvar_reg3p}
\begin{align}
\hat{\var}(\hat{Y}_{reg3p,EX})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_ex} \\
\hat{\var}(\hat{Y}_{reg3p,NEX})& =\hat{\pmb{\alpha}}_{s_2} ^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_{s_2}
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
\end{align}
\end{subequations}

with the variance-covariance matrix of $\hat{\bar{\pmb{Z}}}^{(0)}_0$ and the variance-covariance matrices of the regression coefficients $\hat{\pmb{\alpha}}_{s_2}$ and $\hat{\pmb{\beta}}_{s_2}$:

\begin{subequations}\label{eq:covar3p}
\begin{align}
  \hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}&=
  \frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
  (\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^{\top} \\
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^{(0)2}(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x)\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)\top}(x) \Big)^{-1} \label{eq:covar_alpha} \\
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}&=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^{\top}(x)\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^{\top}(x) \Big)^{-1} \label{eq:covar_beta}
\end{align}
\end{subequations}

Note that $\hat{R}(x)=Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}^{(0)}(x)=Y(x)-\pmb{Z}^{(0)\top}\hat{\pmb{\alpha}}_{s_2}$ denotes the empirical residuals of the reduced model. The external variance form under linear regression models is defined as

\begin{subequations}\label{eq:extvar_reg3p}
\begin{align}
\hat{\var}_{ext}(\hat{Y}_{reg3p,EX})&=\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x))\label{eq:extvar_reg3p_ex} \\
\hat{\var}_{ext}(\hat{Y}_{reg3p,NEX})&=\frac{1}{n_0}\hat{\var}_{s_0}(\hat{Y}^{(0)}(x))
+\frac{1}{n_1}\hat{\var}_{s_2}(\hat{R}^{(0)}(x)) + (1-\frac{n_2}{n_1})\frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:extvar_reg3p_nex}
\end{align}
\end{subequations}
where $\hat{\var}_{s_0}$ indicates taking the sample variance over $s_0$.

% ------------------------------------- %
\subsubsection{Application}

In order to demonstrate the three-phase estimators in the package, we created an artificial three-phase scenario by recoding the phase indicators in the \code{grisons} data set (column \code{phase_id_3p}) according to the terminology used in this article (\code{0} for $s_0$, \code{1} for $s_1$, \code{2} for $s_2$). We now assume that the mean canopy height (\code{mean}) is available at all 306 sample locations $x \in s_0$, whereas we have the explanatory variables \code{stddev}, \code{max} and \code{q75} only at 128 subsamples $s_1$ of $s_0$. At 40 further subsamples $s_2$ we have the observations $Y(x)$ from the field inventory. Based on this setup, we can now define the reduced and full regression model formulas to be used in the \code{threephase()} function (note that the models are nested):

\begin{Schunk}
\begin{Sinput}
R> formula.rm <- tvol ~ mean
R> formula.fm <- tvol ~ mean + stddev + max + q75
\end{Sinput}
\end{Schunk}

Compared to the \code{twophase()}-function, we now have to specify two regression models, i.e., the nested reduced (\code{formula.s0}) and full (\code{formula.s1}) regression model. In addition, we also have to specify the indication of the $s_1$ phase (\code{s1.id}) in the argument \code{phase_id} (note that \pkg{forestinventory} implicitly assumes that all other rows in the input data set belong to $s_0$). The global three-phase estimation can thus be applied by

\begin{Schunk}
\begin{Sinput}
R> reg3p_nex <- threephase(formula.s0 = formula.rm, formula.s1 = formula.fm, 
+    data = grisons, phase_id = list(phase.col = "phase_id_3p", s1.id = 1,  
+    terrgrid.id = 2), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> summary(reg3p_nex)
\end{Sinput}
\begin{Soutput}
Three-phase global estimation

Call: 
threephase(formula.s0 = formula.rm, formula.s1 = formula.fm, 
    data = grisons, phase_id = list(phase.col = "phase_id_3p", 
        s1.id = 1, terrgrid.id = 2), boundary_weights = "boundary_weights")

Method used:
Non-exhaustive global estimator

Full Regression Model:
tvol ~ mean + stddev + max + q75

Reduced Regression Model:
tvol ~ mean

Estimation results:
 estimate ext_variance g_variance  n0  n1 n2 r.squared_reduced
 372.0896     454.4064   451.3626 306 128 40          0.527363
 r.squared_full
      0.7166608

'boundary_weight'- option was used to calculate weighted means of auxiliary variables
\end{Soutput}
\end{Schunk}

The \code{summary()} of a \code{threephase()}-function now recalls both regression model formulas and also gives the $R^2$ for both the reduced (\code{r.squared_reduced}) and the full (\code{r.squared_full}) models. We are told that including \code{stddev}, \code{max} and \code{q75} yields a 20 \% improvement in R$^2$. When comparing to using only \code{mean} under a two-phase approach, we would see a considerable reduction in variance by the three-phase extension.

%--------------------------------------------------------------------------------------------------%
% ################################################################################################ %
%--------------------------------------------------------------------------------------------------%

% ---------------------------------------------------------------------------- %
\subsection{Small area estimators}
\label{sec:threephase_sae}

% ------------------------------------- %
\subsubsection{Mathematical background}

The three two-phase small area estimators described in Section \ref{sec:twophase_sae} can also be extended to the three-phase scenario. The general principle thereby stays the same, i.e., the regession coefficients of the reduced and full model and their variance-covariance matrices are calculated on the entire $s_2$ sample according to Eq. \ref{eq:normequ_redmod}, \ref{eq:normequ_fullmod}, \ref{eq:covar_alpha} and \ref{eq:covar_beta}, and are subsequently used to make predictions for sample locations restricted to small area $G$.

%----------------------------------------------- %
% \textbf{Small and Pseudo Small Area Estimator}\par

The unbiased point estimates of the SMALL and PSMALL estimator are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then corrected for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e., $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^{\top}(x)\hat{\pmb{\beta}}_{s_2}$, to the point estimate. The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a correction term likewise in the global estimation (Eq. \ref{eq:reg3p}).

\begin{subequations}\label{eq:pest_3p_small_psmall}
\begin{align}
\hat{Y}_{G,SMALL,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_small} \\
\hat{Y}_{G,PSMALL,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}+\frac{1}{n_{2,G}}\hat{R}_{G}(x) \label{eq:pe_3p_psmall}
\end{align}
\end{subequations}

The g-weight variance is then calculated as

\begin{subequations}\label{eq:var_3p_small_psmall}
\begin{align}
\hat{\var}(\hat{Y}_{G,SMALL,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} +
\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
\hat{\var}(\hat{Y}_{G,PSMALL,3p})& = \hat{\pmb{\alpha}}_{s_2}^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_{s_2} +
\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} + \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
\end{align}
\end{subequations}

with the variance-covariance matrix

\begin{equation}\label{estvarcovaux3pG}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}=
\frac{1}{n_{0,G}(n_{0,G}-1)}\sum_{x\in{s_{0,G}}}
(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0,G})^{\top}
\end{equation}

The external variance is defined as:

\begin{subequations}\label{eq:pest_3p_small_psmall}
\begin{flalign}
\hat{\var}_{ext}(\hat{Y}_{G,SMALL,3p})&= \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) +
 (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_small}&\\
\hat{\var}_{ext}(\hat{Y}_{G,PSMALL,3p})&= \frac{1}{n_{0,G}}\hat{\var}_{s_{2,G}}(Y(x)) + (1-\frac{n_{1,G}}{n_{0,G}}) \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(\hat{R}^{(0)}(x)) \nonumber &\\
 &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x)) \label{eq:var_3p_reg_psmall}
\end{flalign}
\end{subequations}
where $\hat{R}^{(0)}(x)=Y(x)-\hat{Y}^{(0)}(x)$ with $\hat{Y}^{(0)}(x)=\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}$.

%------------------------------------------------------------- %
% \textbf{Synthetic and Pseudo Synthetic Estimator}\par

The synthetic (SYNTH) and pseudo synthetic (PSYNTH) estimator can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$. Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq. \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq. \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas. The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be ensured within small area $G$. Also the variance will be small but to the cost of ignoring the model uncertainties. Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.

\begin{subequations}\label{eq:3p_synth_psynth}
\begin{align}
\hat{Y}_{G,SYNTH,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
\hat{Y}_{G,PSYNTH,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
\hat{\var}(\hat{Y}_{G,SYNTH,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
\hat{\var}(\hat{Y}_{G,PSYNTH,3p})& = \hat{\pmb{\alpha}}_2^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
\end{align}
\end{subequations}

%------------------------------------------------------------- %
% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par

The extended synthetic (EXTSYNTH) and extended pseudo synthetic (EXTPSYNTH) estimator ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$. This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0. The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{\mathbb{Z}}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$, where $\pmb{\mathbb{Z}}^{(0)\top}(x)=(\pmb{Z}^{(0)\top}(x), I_G^{\top}(x))$. In other words, when the extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for all observations in the input data frame (i.e., $s_0$). The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq. \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}. Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficients resulting from the random sample $s_2$.

%------------------------------------------------------------- %
\subsubsection{Application}

We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the EXTSYNTH and SYNTH estimator to the \code{grisons} data set. The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}). However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}. We thus first define the true means for each small area just as we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):

\begin{Schunk}
\begin{Sinput}
R> truemeans.G <- data.frame(Intercept = rep(1, 4),
+    mean = c(12.85, 12.21, 9.33, 10.45))
R> rownames(truemeans.G) <- c("A", "B", "C", "D")
\end{Sinput}
\end{Schunk}

Three-phase small area estimation in the package can in general be applied by additionally specifying the \code{small_area} list argument. The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument. The EXTSYNTH estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):

\begin{Schunk}
\begin{Sinput}
R> extsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE),
+    exhaustive = truemeans.G, boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extsynth_3p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n0  n1 n2 n0G n1G n2G
1    A 382.6405     1642.055   1518.741 Inf 128 40 Inf  38  12
2    B 368.9013     1501.211   1530.576 Inf 128 40 Inf  34  11
  r.squared_reduced r.squared_full
1         0.5454824      0.7242913
2         0.5354637      0.7171512
\end{Soutput}
\end{Schunk}

The SYNTH estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.

\begin{Schunk}
\begin{Sinput}
R> synth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = FALSE),
+    exhaustive = truemeans.G, boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> synth_3p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n0  n1 n2 n0G n1G n2G
1    A 409.3390           NA   410.7529 Inf 128 40 Inf  38  12
2    B 375.4608           NA   461.8250 Inf 128 40 Inf  34  11
  r.squared_reduced r.squared_full
1          0.527363      0.7166608
2          0.527363      0.7166608
\end{Soutput}
\end{Schunk}

We see that the \code{threephase()}-function returns the sample sizes in the entire inventory area as well as within each small area. The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively. If we compare the two results, we see that the SYNTH estimation again yields a much smaller variance than the EXTSYNTH estimation, but at the cost of a potential bias.

We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$. To do this, we additionally compute the EXTPSYNTH estimates. As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.

\begin{Schunk}
\begin{Sinput}
R> extpsynth_3p <- threephase(formula.rm, formula.fm, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B"), unbiased = TRUE), 
+    boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extpsynth_3p$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n0  n1 n2 n0G n1G n2G
1    A 395.1882     1901.211   1858.204 306 128 40  94  38  12
2    B 389.8329     1846.995   1816.655 306 128 40  81  34  11
  r.squared_reduced r.squared_full
1         0.5454824      0.7242913
2         0.5354637      0.7171512
\end{Soutput}
\end{Schunk}

\newpage

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Calculation of Confidence Intervals ----------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:

\section{Calculation of confidence intervals}
\label{sec:confint_calc}

Converting the estimated variance into a 95\% confidence interval (CI) allows for a more practical interpretation of a point estimate's precision. The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value. In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sample design. Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples and calculate the 95\%  CI from each sample, then 95\% of the CIs are expected to contain the true population parameter. The confidence level $1-\alpha$ (e.g., 95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate. The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries. Based on the central limit theorem it can be assumed that under hypothetical repeated sampling the point estimates will asymptotically follow a normal distribution. However, on the recommendation of \citet{mandallaz2013a}, better confidence intervals can obtained using the Student's \textit{t} distribution defined as

\vspace{4mm}

One-phase estimation:

\begin{equation}\label{ci_1phase}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

% ------------------------------------------ %
Two-phase and three-phase global estimation:

\begin{equation}\label{ci_2phase_global}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

% ------------------------------------------ %
Two-phase and three-phase small area estimation:

\begin{equation}\label{ci_2phase_sae}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

\vspace{4mm}

where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). Note that in case of synthetic estimations (SYNTH, PSYNTH), the degrees of freedom are $n_{2}-p$ as is the case for global estimation. In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function. For example, the 95\% confidence interval for the small area estimates by the EXTPSYNTH estimator (Section \ref{sec:twophase_sae}) are calculated by:

\begin{Schunk}
\begin{Sinput}
R> confint(extpsynth_2p)
\end{Sinput}
\begin{Soutput}
95% Confidence Intervals for twophase small area estimation

  area estimate ci_lower_ext ci_upper_ext ci_lower_g ci_upper_g
1    A 391.9356     325.6463     458.2250   324.9155   458.9558
2    B 419.7231     345.8418     493.6043   352.0456   487.4006
\end{Soutput}
\end{Schunk}

% ---------------------------------------------- %
% Coverage rates

% notes:
%
% - address link between too low sampling fractions and coverage rates under the finite population approach
% - we did show that extremely low sample sizes could have an influence
% - aim:  develop the reader’s intuition about when the sample size is too low to produce reliable results
%

A commonly raised question of practitioners regarding the application of design-unbiased estimators is which minimum terrestrial sample size should be provided. Answering this question is highly related to ensuring the validity of the confidence coverage rates by a sufficiently large sample size. It has to be noted that giving a generally valid sample size for the asymptotic validity range is, unfortunately, infeasible because the convergence to the \textit{t}- or normal distribution also depends on the hetereogeneity of the underlying population. However, simulation studies using an artificial local density surface were conducted in \citet{mandallaz2013a} and \citet{mandallaz2013b} to empirically check the asymptotic properties of the estimators. In order to develop the reader's intuition about when the sample size is too low to produce reliable results, we re-evaluated these simulations for a broader range of sample sizes. For each fixed sample size $n_1$ and sampling fraction, 10000 samples were conducted by simple random sampling and the 95\% confidence interval based on the g-weight variance was calculated for each sample after applying the global two-phase estimator (Section \ref{sec:twophase_globest}). The $R^2$ of the regression model was 0.81. For further details on the simulation, the reader is referred to the original articles mentioned above. The results can also be reproduced or expanded using the \proglang{R}-code given in the Appendix. From the simulation example for global estimation (Figure \ref{fig:covrate_sim}) it is obvious that the nominal coverage rate of 95\% can expected to be met for sample sizes $n_2$ of 50 and higher regardless of the chosen sample size $n_1$. The realized coverage rate only drops significantly below 90\% if considerably small sampling fractions (i.e., the ratio ${n_2}/{n_1}$) < 0.1 are used that lead to small terrestrial sample sizes $n_2$ of 20 and less. Re-evaluating the same simulations for small area estimation confirmed that in such cases, one should whenever possible consider one of the design-unbiased small area estimators (P)SMALL or EXT(P)SYNTH for which the nominal coverage rates in the simulations were already met for sample sizes $n_{2,G}$ > 5. Finally, it should also be noted that applying the design-based estimators under systematic sampling will lead to an inflation of the estimated variance \citep{mandallaz1993}. This will also effect the actual coverage rates to be higher than those derived under simple random sampling.

\begin{figure}[h]
\centering
\resizebox{0.78\hsize}{!}{\includegraphics{fig/globalsim.png}}
\caption{Coverage rates of the global two-phase estimator realized under varying sample sizes $n_1$ and sampling fractions for an artificial simulation example given in \citet{mandallaz2013a}. The confidence intervals were based on the g-weight variance with a nominal confidence level of 95\%. For each setup, 10000 simulations were performed using simple random sampling.}
\label{fig:covrate_sim}
\end{figure}

\newpage

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Special Cases and Scenarios ------------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Special cases and scenarios}
\label{sec:speccas_and_scen}

% --------------------------------------------------------- %
\subsection{Post-stratification}

A special case of multiphase regression estimation is post-stratification, which can further be divided into the cases of multiphase sampling for stratification and multiphase sampling for regression within strata. Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.

To demonstrate post-stratification, we first create an artificial categorical variable development stage (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:

\begin{Schunk}
\begin{Sinput}
R> grisons$stage <- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
\end{Sinput}
\end{Schunk}

Two-phase sampling for stratification is applied if the model only contains categorical variables, in this case the factor variable \code{stage}. Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multiphase regression estimators, is equivalent to post-stratification. For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).

\begin{Schunk}
\begin{Sinput}
R> twophase(formula = tvol ~ stage, data = grisons, 
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

Two-phase sampling for regression within strata implies the combination of continuous and categorical variables within the model (i.e., we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of OLS regression techniques in the multiphase estimators, including higher order terms and transformations of the explanatory variables, which makes them very flexible.

\begin{Schunk}
\begin{Sinput}
R> twophase(formula=tvol ~ mean + stddev + max + q75 + stage, data = grisons,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

The variance of all design-based estimators included in \pkg{forestinventory} can be decreased by reducing the sum of squared residuals of the regression model. In case of post-stratification, this particularly implies minimizing the within strata residual square sum. Also, for post-stratification, the g-weight variance should be trusted over the external variance because it has the advantage that the strata weights are estimated from the large sample rather than the terrestrial sample $s_2$.

% --------------------------------------------------------- %
\subsection{Small area estimation under cluster sampling}

As mentioned in Section \ref{sec:packstruc}, cluster sampling is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points. One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$. The idea of cluster sampling is to increase the amount of information without increasing the travel costs of the terrestrial campaign. However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e., the cluster origin $x$. Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights. Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e., defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random. The \pkg{forstinventory} package identifies clusters via a unique cluster ID that is assigned to a column in the input data set. Its column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e., only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the extended and pseudo extended synthetic estimator (see EXTSYNTH and EXTPSYNTH in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, it is adviseable to use the (pseudo) small area estimator (SMALL or PSMALL) where the zero mean residual property is still ensured.

In order to keep track of such cases, \pkg{forestinventory} tells the user to do so by returning a warning message:

\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust <- twophase(formula = basal ~ stade + couver + melange, 
+    data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster",  small_area = list(sa.col = "ismallold", areas = c("1"),
+    unbiased = TRUE))
Warning message:
At least one terrestrial cluster not entirely included within small area 1.
Zero mean residual assumption for small area maybe violated.
Check mean_Rc_x_hat_G and consider alternative estimator 'psmall' 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> psmall.clust <- twophase(formula = basal ~ stade + couver + melange, 
+    data = zberg, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster", small_area = list(sa.col = "ismallold", areas = c("1"),
+    unbiased = TRUE), psmall = TRUE)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extpsynth.clust$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
1    1 25.54748     14.03806   14.16853 298 73  29   8  0.205741
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> psmall.clust$estimation
\end{Sinput}
\begin{Soutput}
  area estimate ext_variance g_variance  n1 n2 n1G n2G r.squared
1    1 23.98581     16.30509   15.69473 298 73  29   8 0.1873795
\end{Soutput}
\end{Schunk}

Comparing the EXTPSYNTH and PSMALL estimates, we see that in this particular case the point estimates are close and more important, the external as well as the g-weight variances only differ marginally. This can be taken as evidence that the violation of the zero mean residual property can here be expected to have negligible consequences.

% --------------------------------------------------------------------- %
\subsection{Violation of nesting in sample design}

As explained in Section \ref{sec:str_and_mod}, a basic prerequisite for the application of multiphase estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are nested in each other. The correct nesting thereby concerns the spatial arrangement of the sample phases (Fig. \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location. For the latter, \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first-aid adjustments to the inventory data set to prevent the calculations from failing. We will demonstrate possible nesting violations by applying the global three-phase estimator to the \code{grisons} and \code{zberg} data sets.

\subsubsection*{Violation 1}

Based on the nesting rule, $s_2 \subset s_1 \subset s_0$, each $s_2$ and $s_1$ sample location must have all explanatory variables available that are used in the full (and thus reduced) regression model. If e.g., an $s_2$ and/or $s_1$ point misses a variable which is used in the full and reduced model (in this case \code{mean}), the function will delete this sample point from the dataset and produce the following messages:

\begin{Schunk}
\begin{Sinput}
R> grisons[which(grisons$phase_id_3p == 2)[1], "mean"] <- NA
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the zero phase (s0) is missing
3: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (1 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}

\subsubsection*{Violation 2}

However, if an $s_2$ and/or $s_1$ point is missing a variable which is only used in the full regression model (in this example \code{q75}), the function will recode the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model. If this concerns an $s_2$ sample location, the associated value of the response variable can no longer be used.

\begin{Schunk}
\begin{Sinput}
R> grisons[which(grisons$phase_id_3p == 2)[1], "q75"] <- NA
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col="phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning messages:
1: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Sample design not nested: for 1 terrestrial plots at least one auxiliary 
  parameter of the first phase (s1) is missing
2: In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  Changed the phase_id for 1 rows to the zero phase (s0) due to missingness in 
  the set of auxiliary parameters for the first phase (s1) (1 terrestrial 
  information no longer usable by this change)
\end{Sinput}
\end{Schunk}

\subsubsection*{Violation 3}

If an $s_0$ point misses at least one of the explanatory variables used in the reduced model, the sample locations are deleted from the data set.

\begin{Schunk}
\begin{Sinput}
R> grisons[which(grisons$phase_id_3p == 0)[1], "mean"] <- NA
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> threephase(formula.s0 = tvol ~ mean, 
+    formula.s1 = tvol ~  mean + stddev + max + q75,  data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    boundary_weights = "boundary_weights")
Warning message:
In threephase(formula.s0 = tvol ~ mean, formula.s1 = tvol ~ mean +  :
  1 rows deleted due to missingness in the set of auxiliary parameters for the 
  zero phase (s0) (0 terrestrial plots affected by deletion)
\end{Sinput}
\end{Schunk}

Note that all the automatic data adjustments (deletion, recoding) have to be accepted with caution. Recapitulating, the unbiasedness of estimators in the design-based framework is based on the uniform and independent randomization of the sample locations. This means that every possible location within the forest area $F$, as well as pairs of locations, have inclusion and joint inclusion probabilities greater than zero. Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures. If any information should be missing at the sample locations, one should clarify the reason for this and make sure that the information can reasonably be assumed to be completely missing at random.

\subsubsection*{Violation 4}

If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample. Therefore, an internal regression model cannot be calculated and the function stops with the following error message:

\begin{Schunk}
\begin{Sinput}
R> zberg <- zberg[-which(zberg.n$phase_id_2p == 2 & zberg.n$stade == "300"), ]
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> twophase(formula = basal ~ stade + couver + melange, data = zberg,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    cluster = "cluster")
Error in check.mandatoryInputs(formula, data, phase_id) : 
 Level '300' of factor variable 'stade' existing in s1(s0)-but not in s2 sample.
 Calculation of coefficient not feasible.
\end{Sinput}
\end{Schunk}

\newpage

%------------------------------------------------------------------------------------------------%
% ----------------------------- Visualization and Analysis of Estimation Results --------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:

\section{Analysis and visualization}
\label{sec:anal_and_vis}

\subsection{Analysis}

We often want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, which can be easily accomplished in \pkg{forestinventory} using the \code{estTable()} function. This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase EXTPSYNTH and PSYNTH estimator for the \code{grisons} data set:

\begin{Schunk}
\begin{Sinput}
R> op <- onephase(formula = tvol~1, data = grisons,
+    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
+    unbiased = TRUE), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, 
+    data = grisons, phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+    unbiased = FALSE), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> extpsynth_3p <- threephase(formula.s0 = tvol ~ mean,
+    formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+    phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+    small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+    unbiased = TRUE), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> psynth_3p <- threephase(formula.s0 = tvol ~ mean,
+   formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
+   phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
+   small_area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
+   unbiased = FALSE), boundary_weights = "boundary_weights")
\end{Sinput}
\end{Schunk}

We can then aggregate all estimation objects in a \code{list} and pass it to the \code{estTable()}-function:

\begin{Schunk}
\begin{Sinput}
R> grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
+    extpsynth_3p, psynth_3p), sae = TRUE, vartypes = c("variance", "g_variance", 
+    "ext_variance"))
\end{Sinput}
\end{Schunk}

The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}. The \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances. If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}. Note however that \pkg{forestinventory} provides several \code{S3} generic methods specifically for the class \code{"esttable"}.

The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package. However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance that was applied (\code{g_} for \textit{g-weight} and \code{ext_} for external variance). By default, the confidence intervals are also added.

\begin{Schunk}
\begin{Sinput}
R> str(grisons.sae.table)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
List of 20
 $ area             : chr [1:28] "A" "A" "A" "A" ...
 $ domain           : Factor w/ 2 levels "global","smallarea": 1 2 2 2 2 2 2 1 2  ...
 $ method           : Factor w/ 3 levels "onephase","twophase",..: 1 3 3 3 2 2 2  ...
 $ estimator        : Factor w/ 3 levels "onephase","psynth extended",..: 1 2 2 3 ...
 $ vartype          : Factor w/ 3 levels "ext_variance",..: 3 1 2 2 1 2 2 3 1 2 ...
 $ estimate         : num [1:28] 410 395 395 422 392 ...
 $ variance         : num [1:28] 1987 1901 1858 726 996 ...
 $ std              : num [1:28] 44.6 43.6 43.1 26.9 31.6 ...
 $ error            : num [1:28] 10.86 11.03 10.91 6.39 8.05 ...
 $ n2               : num [1:28] 19 40 40 40 67 67 67 17 40 40 ...
 $ n2G              : num [1:28] NA 12 12 12 19 19 19 NA 11 11 ...
 $ n1               : num [1:28] NA 128 128 128 306 306 306 NA 128 128 ...
 $ n1G              : num [1:28] NA 38 38 38 94 94 94 NA 34 34 ...
 $ n0               : int [1:28] NA 306 306 306 NA NA NA NA 306 306 ...
 $ n0G              : int [1:28] NA 94 94 94 NA NA NA NA 81 81 ...
 $ r.squared        : num [1:28] NA NA NA NA 0.653 ...
 $ r.squared_reduced: num [1:28] NA 0.545 0.545 0.527 NA ...
 $ r.squared_full   : num [1:28] NA 0.724 0.724 0.717 NA ...
 $ ci_lower         : num [1:28] 317 299 300 367 326 ...
 $ ci_upper         : num [1:28] 504 491 490 476 458 ...
 - attr(*, "row.names")= int [1:28] 1 2 3 4 5 6 7 8 9 10 ...
 - attr(*, "class")= chr [1:3] "list" "esttable" "smallarea"
\end{Sinput}
\end{Schunk}

Note that \code{estTable()} also returns the estimation error (\code{error}) that is defined as the standard error devided by the point estimate:

\begin{equation}\label{eq:esterror}
  error_{[\%]}=\frac{\sqrt{\hat{\var}(\hat{Y})}}{\hat{Y}}\cdot100
\end{equation}

As multiphase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multiphase global or small area estimate compared to its respective one-phase estimate. The function takes an \code{esttable} object as input and returns a summary of which multiphase method and estimator performed best using the precision from the one-phase estimator as a baseline. If the \code{esttable} object contains more than one multiphase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation. The argument \code{pref.vartype} can be used to define what type of variance (g-weight or external) should be used for the comparison. Synthetic estimates (SYNTH and PSYNTH estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.

\begin{Schunk}
\begin{Sinput}
R> mphase.gain(grisons.sae.table, pref.vartype = "g_variance")
\end{Sinput}
\begin{Soutput}
  area var_onephase var_multiphase     method       estimator gain  rel.eff
1    A     1987.117      1017.6327   twophase psynth extended 48.8 1.952686
2    B     3175.068      1019.1913   twophase psynth extended 67.9 3.115281
3    C     1180.853       763.0731 threephase psynth extended 35.4 1.547496
4    D     2290.652      1110.2454   twophase psynth extended 51.5 2.063194
\end{Soutput}
\end{Schunk}

% Interpretation: (gain and rel.eff)
The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multiphase estimator (\code{var_multiphase}). The multiphase estimation procedure is again specified in the \code{method} and \code{estimator} column. The last two columns quantify the potential benefit of the multiphase estimation. The \code{gain} is the reduction (if its value is positive) in variance when applying the multiphase as alternative to the one-phase estimation. For example, it is indicated that the two-phase extended PSYNTH estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure. The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multiphase variance:

\begin{equation}\label{eq:releff}
  rel.eff=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}
  \end{equation}

The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multiphase estimator. For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the two-phase EXTPSYNTH estimator. If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multiphase estimation for forest inventories.

\subsection{Visualization}

The \pkg{forestinventory} package also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results in two ways: 1) the point estimates with overlayed confidence intervals, and 2) the estimation errors. Both plots can be obtained by passing the \code{esstable} object to the \code{plot()} function.

\begin{Schunk}
\begin{Sinput}
R> plot(grisons.sae.table, ncol = 2)
\end{Sinput}
\end{Schunk}

\begin{figure}[h]
\centering
\resizebox{0.85\hsize}{!}{\includegraphics{fig/plot_error.png}}
\end{figure}

\newpage

Whereas the estimation errors are plotted by default, the point estimates and confidence intervals are returned when setting the argument \code{yvar = "estimate"}. Note that the graphics can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.

\begin{Schunk}
\begin{Sinput}
R> plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
+    ylab("Timber Volume [m3/ha]")
\end{Sinput}
\end{Schunk}

\begin{figure}[h]
\centering
\resizebox{0.85\hsize}{!}{\includegraphics{fig/plot_est_ci.png}}
\end{figure}

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Future Plans---- ------------------------------------ %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\newpage

\section{Future plans}
\label{sec:future}

The \pkg{forestinventory} package currently provides a fairly well-rounded toolkit for forestry inventorists to integrate auxiliary information into their estimates using the model-assisted methods under the design-based approach.  Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future. As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub \citep{github_forestinventory}. Currently planned extensions include:

\begin{itemize}
\item Implement parallel procedures for efficiently calculating many small areas.
\item Allow functions to accept objects of class \code{data.table} from the \pkg{data.table} package \citep{dt2017} to improve memory efficiency.
\item Enable the user to choose other types of models than linear regressions fitted with OLS.
\end{itemize}

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Acknowledgements ------------------------------------ %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section*{Acknowledgements}

We want to express our gratitude to Prof. H. Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study and providing the possibility of working on the package. We also want to thank Daniel Mandallaz for his support in completing the range of the already published estimators in the frame of the three-phase small area estimators, as well as many helpful discussions and advice throughout the implementation of our package. Our thanks also go to Meinrad Abegg for proofreading the manuscript, and to the Amt f{\"u}r Wald und Naturgefahren of the Swiss canton of Grisons for providing the example data. We also want to thank two anonymous reviewers for their valuable support to improve the manuscript.

\newpage

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Appendix -------------------------------------------- %

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:

\section*{Appendix}
\label{sec:appendix}

\subsection*{Simulation for global two-phase estimation}

\begin{Schunk}
\begin{Sinput}
R> target.suface <- function(x0, y0){
+    30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> library("rmutil")
R> int2(target.suface, a = c(0,0), b=c(3,2))/6
R> int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> sample.generator<- function(n1, n2, target.surface=target.surface){
+    realization <- matrix(NA,n1,7)
+    for(i in 1:n1){
+      x0=2*runif(1)
+      y0=3*runif(1)
+      realization[i,1] <- target.suface(x0, y0)
+      realization[i,2]<-x0
+      realization[i,3]<-y0
+      realization[i,4]<-x0*x0
+      realization[i,5]<-x0*y0
+      realization[i,6]<-y0*y0
+      realization[i,7]<-0
+    }
+    realization <- as.data.frame(realization)
+    names(realization) <- c("response","x","y","xx","xy","yy", "area")
+    realization$phase <- 1
+    realization$phase[sample(nrow(realization), n2)] <- 2
+    realization$response[realization$phase == 1]<- NA
+    realization$area[realization$x >= 0.3 & realization$x <= 1.3 & 
+                     realization$y >= 0.5 & realization$y <= 2]<- 1
+    return(realization)
+  }
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> run.global.sim <- function(num_replications, n1.vec){
+  
+    sfrac <- seq(0.02, 0.26, by=0.02) 
+    true_value<- 39.17 
+    
+    coverage_rates <- data.frame(n1=integer(), n2=integer(), 
+                                 sampling_fractions=numeric(), 
+                                 avg.error.ext=numeric(),
+                                 avg.error.g=numeric(),
+                                 covrate.ext=numeric(),
+                                 covrate.g=numeric(),
+                                 no.valid.sims = integer())
+    
+    for(i in 1:length(sfrac)){
+        
+      n2 <- integer()
+      error.ext <- numeric()
+      error.g <- numeric()
+      ci.ext.logical <- logical()
+      ci.g.logical <- logical()
+      no.valid.sims. <- integer()
+        
+        
+      for(j in 1:num_replications){
+          
+        realization <- sample.generator(n1 = n1.vec, n2 = round(n1.vec*sfrac[i]))
+          
+        est <- twophase(formula = response ~ y + x + xx, data=realization,
+                        phase_id = list(phase.col = "phase", terrgrid.id = 2))
+          
+        ci <- confint(est)
+        ci.ext.logical <- append(ci.ext.logical, 
+                                 ci$ci$ci_lower_ext <= true_value & 
+                                 true_value <= ci$ci$ci_upper_ext)
+        ci.g.logical <- append(ci.g.logical, 
+                               ci$ci$ci_lower_g <= true_value & 
+                               true_value <= ci$ci$ci_upper_g)
+        error.ext <- append(error.ext, 
+                            sqrt(est$estimation$ext_variance)/est$estimation$estimate)
+        error.g <- append(error.g, 
+                          sqrt(est$estimation$g_variance)/est$estimation$estimate)
+          
+        n2 <- append(n2, est$estimation$n2)
+        no.valid.sims. <- sum(n2 > 1)
+          
+      }
+      
+      coverage_rates <- rbind(coverage_rates, 
+                              data.frame(n1=n1.vec, n2=round(mean(n2)), 
+                                         sampling_fractions=sfrac[i], 
+                                         avg.error.ext=mean(error.ext, na.rm = TRUE),
+                                         avg.error.g=mean(error.g, na.rm = TRUE),
+                                         covrate.ext=mean(ci.ext.logical, na.rm = TRUE), 
+                                         covrate.g=mean(ci.g.logical, na.rm = TRUE),
+                                         no.valid.sims=no.valid.sims.))
+        
+    } 
+    
+  return(coverage_rates)
+  
+  } 
\end{Sinput}
\end{Schunk}

\newpage

\begin{Schunk}
\begin{Sinput}
R> library(forestinventory)
R> global.sims<- data.frame()
R> n1_sims<- c(250, 500, 750, 1000, 1500, 2000) 
R>   for (i in 1:length(n1_sims)){
+      res<- run.global.sim(num_replications = 10000, n1.vec = n1_sims[i])
+      global.sims <- rbind(global.sims, res)
+    }
\end{Sinput}
\end{Schunk}

\newpage

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Literature ------------------------------------------ %

\bibliography{bib/jss3280}

%------------------------------------------------------------------------------------------------%
\end{document}

### 2019-04-27 JStatSoft_forestinventory_master-concordance.tex
formatting changes for jss

@@
 \Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
-1 105 1 1 4 1 1 1 4 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
+1 104 1 1 4 1 1 1 4 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 112:%
 1 64 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
-121 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 177:%
+120 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 182:%
 1 67 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
-127 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 250:%
+126 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 255:%
 1 2 1 1 4 114 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
@@
 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 824:%
-133}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 825:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 823:%
+132}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 824:%
 1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
@@
 0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1176:%
-135 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1181:%
-1 2 1 1 4 38 1 1 2 11 0 1 2 78 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1314:%
-141 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1319:%
-1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1175:%
+134 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1180:%
+1 2 1 1 4 36 1 1 2 11 0 1 2 78 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1311:%
+140 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1316:%
+1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 15 1 1 5 %
 7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
 0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1554:%
-147 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1559:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1550:%
+146 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1555:%
 1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
@@
 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
-153 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
-1 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
-159 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1744:%
+152 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1749:%
+1 13 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1763:%
+158 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1768:%
 1 6 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1777:%
-165 4 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1782:%
-1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
-3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 11 1}
-\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1950:%
-171 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1775:%
+164 4 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1780:%
+1 2 1 1 4 8 1 1 3 5 0 1 2 2 1 1 2 1 0 2 1 3 0 1 2 2 1 1 23 25 0 1 2 %
+2 1 1 63 64 0 1 2 6 1 1 3 1 0 2 1 1 4 6 0 1 2 11 1}
+\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1935:%
+170 11 1}

### 2019-04-27 JStatSoft_forestinventory_master.rnw
update andi 16_09_2018(1)
update andi 16_09_2018(1)
formatting changes for jss

@@
   8092 Z\"urich, Switzerland\\
-  E-mail: \email{andreas.hill@usys.ethz.ch}\\
-  Telephone: +41/44/632 32 36\\
+  E-mail: \email{andreas.hill@forestinventory.de}\\
 
@@
 %------------------------------------------------------------------------------------------------%
+% ---------------------------------------- Appendix -------------------------------------------- %
+
+\SweaveInput{rnw/Appendix.Rnw}
+
+%------------------------------------------------------------------------------------------------%
 % ---------------------------------------- Literature ------------------------------------------ %

### 2019-04-27 rnw/Analysis and Visualization.Rnw
formatting changes for jss

@@
 
-\begin{small}
 <<>>=
@@
 @
-\end{small}
 
-\begin{small}
 <<>>=
@@
 @
-\end{small}
 
@@
 
-\begin{small}
 <<eval=FALSE>>=
@@
 @
-\end{small}
 
-\begin{small}
 \begin{Schunk}
@@
 \end{Schunk}
-\end{small}
 

### 2019-04-27 rnw/Appendix.Rnw
update 15_11_2018 2nd revision
update 15_11_2018 2nd revision
formatting changes for jss

--- 2018-09-16 rnw/Appendix.Rnw
@@
 
-\begin{small}
 <<eval=FALSE>>=
-# function to create density surface:
 target.suface <- function(x0, y0){
@@
 @
-\end{small}
 
-\begin{small}
 <<eval=FALSE>>=
-# --- True spatial mean for Global and Small Area Surface Area: --- #
-library(rmutil)
-int2(target.suface, a = c(0,0), b=c(3,2))/6 # 39.17
-int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) # 37.16243
+library("rmutil")
+int2(target.suface, a = c(0,0), b=c(3,2))/6
+int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) 
 @
-\end{small}
 
-\begin{small}
 <<eval=FALSE>>=
-# --- sample generator for simple sampling: --- #
 sample.generator<- function(n1, n2, target.surface=target.surface){
@@
 @
-\end{small}
 
-\begin{small}
 <<eval=FALSE>>=
 
-# --- function for global simulations --- #
 run.global.sim <- function(num_replications, n1.vec){
 
-  sfrac <- seq(0.02, 0.26, by=0.02) # sampling fractions
-  true_value<- 39.17 # true population mean
+  sfrac <- seq(0.02, 0.26, by=0.02) 
+  true_value<- 39.17
 
-  # data.frame to store results:
   coverage_rates <- data.frame(n1=integer(), n2=integer(), 
@@
 
-  # -------------- #
-  # simulation loop:
-
   for(i in 1:length(sfrac)){
 
-    # variables to store temporary results:
     n2 <- integer()
@@
 
-    # calculate cove.rates, avg. n2G etc:
     coverage_rates <- rbind(coverage_rates, 
@@
 
-  } # end of sim
+  }
 
@@
 
-} # end of function
+} 
 @
-\end{small}
 
@@
 
- \begin{small}
 <<eval=FALSE>>=
-# --- run global simulations --- #
-library(forstinventory)
+
+library(forestinventory)
 global.sims<- data.frame()
-n1_sims<- c(250, 500, 750, 1000, 1500, 2000) # n1 sample sizes
+n1_sims<- c(250, 500, 750, 1000, 1500, 2000) 
   for (i in 1:length(n1_sims)){

### 2019-04-27 rnw/Introduction.Rnw
formatting changes for jss

@@
 data, commonly in the form of spatially exhaustive remote sensing data in the inventory area.
-Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLüpke2012}.
+Regression estimators using this concept can consider either one additional sample of plot locations (two-phase or double-sampling) or two additional samples available in different sample sizes (three-phase or triple-sampling) \citep{gregoire2007, saborowski2010, mandallaz2013a, mandallaz2013c, vonLuebke2012}.
 Their application to existing forest inventory systems have already showed their efficiency in terms of cost reduction and gain in estimation precision \citep{breidenbach2012, vonLuebke2014, mandallaz2013b, magnussen2014, massey2014a}.\par

### 2019-04-27 rnw/Special Cases and Scenarios.Rnw
formatting changes for jss

--- 2018-08-30 rnw/Special Cases and Scenarios.Rnw
@@
 
-\begin{small}
 \begin{Schunk}
@@
 \end{Schunk}
-\end{small}
 
-\begin{small}
 <<>>=

### 2019-04-27 rnw/Twophase_Estimators and their Application.Rnw
formatting changes for jss

@@
 It has been mathematically shown that by this procedure, the coefficient of determination (R$^2$) of the regression model can be overoptimistic as the zeros are easy to predict.
-This is particularly the case if the occurence of zeros correlate with an explanatory variable, such as the elevation above sea level.
+This is particularly the case if the occurence of zeros correlates with an explanatory variable, such as the elevation above sea level.
 Furthermore, \citet{mandallaz2014} showed that this procedure is inefficient as it inflates the estimated variance.
@@
 @
-\end{small}
 
-\begin{small}
 <<eval=TRUE, echo=FALSE>>=
@@
 @
-\end{small}
 
 An alternative way to look at the estimation results without using the \code{summary()} is to query \code{reg2p_ex} directly:
-\begin{small}
+
 <<>>=
@@
 @
-\end{small}
+
 Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation.
@@
 In order to apply the pseudo small area estimator (PSMALL) with boundary adjustment, we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
-\begin{small}
+
 <<>>=
@@
 We can view the sample sizes by looking into the object itself
-\begin{small}
+
 <<>>=
@@
 For the \code{grisons} data set, the exact means of the explanatory variables for the small areas used in \citet{mandallaz2013b} are thus defined by
-\begin{small}
+
 <<eval=FALSE>>=
@@
 @
-\end{small}
-\begin{small}
+
 <<>>=
@@
 @
-\end{small}
 
 The extended synthetic estimator (EXTSYNTH) can then be applied by
-\begin{small}
+
 <<>>=

### 2019-04-27 Final Edit/jss3280_replicationcode/replication_rcode.R

@@ -1,0 +1,796 @@
+#----------------------------------------------------------------------------------- #
+#  Journal of Statistical Software                                                   #
+#  Design-Based Global and Small Area Estimations for Multi-phase Forest Inventories #
+#  Andreas Hill & Alexander Massey                                                   #
+#                                                                                    #
+#  Code for running all examples                                                     #
+#                                                                                    #
+#----------------------------------------------------------------------------------- #
+
+library("forestinventory")
+
+# **************************************************************************** #
+# ------ SECTION 3: TWO-PHASE ESTIMATORS -------------------------------------
+# **************************************************************************** #
+
+
+# ----- 3.1 Global Estimations --------
+
+
+## load grisons data set:
+data("grisons", package = "forestinventory")
+
+## see structure of grisons:
+str(grisons)
+
+## see first 12 rows of grisons:
+head(grisons, n = 12)
+
+
+# **********************************************************
+## non-exhaustive (NEX) estimator with boundary weight adjustment
+reg2p_nex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                      data = grisons,
+                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                      boundary_weights = "boundary_weights")
+
+## get summary of estimation results:
+summary(reg2p_nex)
+
+
+# *****************************************************
+## exhaustive (EX) estimator with boundary weight adjustment
+## first, retrieve order of columnames that the lm()-function processes a formula object:
+#  --> actually the order they appear in the design-matrix ...
+colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
+
+## store true means of explanatory variables:
+true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)
+
+## apply estimator:
+reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
+                     data = grisons,
+                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
+                     exhaustive = true.means.Z)
+
+## get summary of estimation results:
+summary(reg2p_ex)
+
+## get data frame with estimation results
+reg2p_ex$estimation
+
[…]

### 2019-04-27 r/replication_rcode.R

@@ -589,3 +589,208 @@ plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
 
 
 
+
+# -------------------------------------------------------------------------- #
+# ----- Simulation Code --------
+# -------------------------------------------------------------------------- #
+
+# function to create density surface:
+
+target.suface <- function(x0, y0){
+  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
+# The true mean value is derived from an integral can be calculated in analytically or in Matlab.
+# The true mean value here is 39.17
+
+# --- True spatial mean for Small area surface area: --- #
+library(rmutil)
+int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) # 37.16243
+
+
+# -------------------------------------------------------------------------- #
+# -------------------------------------------------------------------------- #
+
+
+# --- sample generator for simple sampling: --- #
+sample.generator<- function(n1, n2, target.surface=target.surface){
+  realization <- matrix(NA,n1,7)
+  for(i in 1:n1){
+    x0=2*runif(1)
+    y0=3*runif(1)
+    realization[i,1] <- target.suface(x0, y0)
+    realization[i,2]<-x0
+    realization[i,3]<-y0
+    realization[i,4]<-x0*x0
+    realization[i,5]<-x0*y0
+    realization[i,6]<-y0*y0
+    realization[i,7]<-0
+  }
+  realization <- as.data.frame(realization) # warning: still contains true response variable for all 1st phase
+  names(realization) <- c("response","x","y","xx","xy","yy", "area")
+  realization$phase <- 1
+  realization$phase[sample(nrow(realization), n2)] <- 2
+  realization$response[realization$phase == 1]<- NA
+  realization$area[realization$x >= 0.3 & realization$x <= 1.3 & realization$y >= 0.5 & realization$y <= 2]<- 1
+  return(realization)
+}
+
+# -------------------------------------------------------------------------- #
+# -------------------------------------------------------------------------- #
+
+# simulation for small area edge cases:
+
+library(forestinventory)
+num_replications <- 10000
+n1.vec <- c(100, 150, 200, 250, 500, 750, 1000) 
+sfrac <- 0.25
+true_value.sae<- 37.16
+
+# data.frame to store results:
+coverage_rates <- data.frame(n1=integer(), n2=integer(), 
+                             sampling_fractions=numeric(), 
[…]
